### 2017-03-18 Manuscript/manuscript.tex
Add figures and summary of current results in Manuscript folder.

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

%\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

%\usepackage{graphicx}

%\draft % marks overfull lines with a black rule on the right
%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{csquotes}

\begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

\title{NMRlipids VI: New PC lipid model with correct cation binding affinity} %Title of paper

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s, 
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command. 
% The \affiliation command should follow the other information.

\author{Josef Melcr}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}

\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
% insert abstract here
  Cellular membranes are complex systems that are hard to study both experimentally and theoretically. Classical molecular dynamics simulations (MD)
give detailed information about membrane structure, dynamics and its interactions with other biomolecules. However,  there is still a room for
improvements in the current force fields, especially in the hydrophilic region of membrane.  According to the results reported in NMRlipids, Open Collaboration project
running at nmrlipids.blogspot.fi, biologically relevant Na+ and Ca2+ generally overbind to zwitterionic phophatidylcholine (PC) membranes in current MD models [Ollila 2016].
Here we suggest that the PC lipid membrane-ion interactions can be corrected by implicitly inducing electronic polarizability in the lipid models by
using a simple electrostatic continuum correction (ECC) for partial charges [Leontyev 2010], which is successfully applied already to monovalent and divalent ions in bulk water [Kohagen, 2015].
Our results show that applying ECC to existing force fields significantly improved PC lipid--cation interactions when compared against NMR data
by using the electrometer concept [Ollila 2016,Seelig87]. However, also other force field parameters have to be made consistent
for ECC correction to preserve the area per molecule and other relevant bilayer properties. This work is carried as Open Collaboration
in NMRlipids VI project (http://nmrlipids.blogspot.fi).
\end{abstract}

%\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands

%\label{}
\section{Introduction}

Cations were shown to generally overbind in PC lipid bilayers in
NMRlipids II project. Here we propose that the cation overbinding can
be corrected by implicitly inducing electronic polarizability in
lipid headgroups by scaling the partial charges.

%\section{Methods}

%\subsection{Molecular dynamics simulations}

\section{Effect of scaling of headgroup and glycerol backbone partial charges on lipid bilayer properties and cation binding}

Headgroup and glycerol backbone partial charges of Slipids and Lipid14 models were scaled
with various scaling factors and cation binding was monitored by using the electrometer
concept. Headgroup order parameter changes as a function of ion concentration are
shown in Fig. \ref{OrderParameterIONSchangesSCALED}. The order parameter decrease
seemst to be proportional to the headgroup charges: the smaller the charges
(larger scaling factor), the less change is observed in order parameters.
According to the electrometer concept, this means that the headgroup scaling
decreases cation binding affinity. However, from this data only we cannot
exclude the possibility that order parameters are changing less because
headgroup becomes less sensitive for cation binding.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/OrderParameterIONSchangesSCALED.eps}
  \caption{\label{OrderParameterIONSchangesSCALED}
    Headgroup order parameter changes as a function of cations from models
    with modified headgroup partial charges.}
\end{figure}

The best models from Fig.  \ref{OrderParameterIONSchangesSCALED} are shown in
Fig. \ref{OrderParameterCHANGESnewMODELS}. Based on these results we choose
two models for more careful studies: Slipids and Lipid 14 models with headroup
and glycerol backbone partial charges scaled with 0.8 and 0.85, respectively.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
  \caption{\label{OrderParameterCHANGESnewMODELS}
    Headgroup order parameter changes as a function of cations from models
    with modified headgroup partial charges.}
\end{figure}

While scaling seems to improve ion binding, it changes the lipid bilayer properties
without ions. The glycerol backbone and headgroup order parameters are plotted in
Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
The area per lipid from different models are shown in Table \ref{apls}
The conclusion is that the charge scaling has only a marginal effect on
headgroup order parameters, while area per lipid is significantly decreased.
Area per lipid decrease is more pronounced in Lipid14 than in Slipids.

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodels.eps}
  \caption{\label{HGopsNEWmodels}
    Headgroup and glycerol backbone order parameters from standard and EEC models.}
\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodelsCHANGE.eps}
  \caption{\label{HGopsNEWmodelsCHANGE}
    Changes in headgroup and glycerol backbone order parameters due to EEC.}
\end{figure}

\begin{table}
  \caption{Area per lipid from different models for POPC without ions\label{apls} }
  \begin{tabular}{c c}
    model          & A (Å$^2$)    \\
    Lipid14 (literature)  & 65.6$\pm$ 0.5   \\
    Lipid14eec0.85 & 55.5      \\
    Slipids (literature T=303K)       & 64.6 $\pm$ 0.4   \\
    SlipidsEEC0.8  & 57.8  \\
  \end{tabular}
\end{table}

The Calcium densities from new models and standard models are compared in Fig \ref{CAdensities}.
It seems to me that Ca binding in Lipid14 model is not reduced by scaling the
partial charges even thought the order parameter response is better.
On the other hand, in Slipids the binding seems to be reduced.
Possible interpretation is that Lipid14 result gets better because headgroup
gets less sensitive for bound charge, while in Slipid the binding is reduced.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/CAdensities.eps}
  \caption{\label{CAdensities}
    Calcium densities from different simulations with new and standard models.
  }
\end{figure}

\section{Conclusions}

% Tables may be be put in the text as floats.
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
%
% \begin{table}
% \caption{\label{} }
% \begin{tabular}{}
% \end{tabular}
% \end{table}

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
% Put your acknowledgments here.
\end{acknowledgments}
\newpage
\appendix
\begin{center}
{\bf SUPPLEMENTARY INFORMATION}
\end{center}

% Create the reference section using BibTe
\bibliography{refs.bib}

%\newpage
%\section{APPENDIX: The NMR results reported by Tiago Ferreira}

\listoftodos

\end{document}
%
% ****** End of file aiptemplate.tex ******

### 2017-04-27 scripts/area-per-lipid_calculate_GMX_deffnm.sh
Ca2+ electrometer results from a -close-to-final- model of Lipid14-based POPC (q_sc=0.8, sigma_sc=0.875, all non-tail atoms), extra scripts for GMX deffnm simulations, Manuscript: title of the future paper.

@@ -1,0 +1,58 @@
+#!/bin/bash
+
+# wrapping BASH script for calc_apl.py
+#  uses GMX tools to calculate Area per Lipid (APL)
+# meant for use with NMRlipids projects
+
+scriptdir=`dirname $0`
+
+
+ die() {
+       echo $@
+       exit 1
+ }
+
+
+[ -z $1 ] && die "Default file name not specified.   Usage: this_script.sh DFNM"
+dfnm=$1
+
+traj_file_name=$dfnm".xtc" #"../traj.trr" 
+traj_pbc_nonwat_file_name=$dfnm"nonwat_pbc.xtc" #"../traj.trr" 
+top_file_name=$dfnm"_last_frame_nonwat.gro"
+tpr_file_name=$dfnm".tpr"
+#op_def_file="../../Headgroup_Glycerol_OPs.def"
+op_def_file="../../order_parameter_definitions_POPC_all.def"
+op_out_file="OrdPars.dat"
+top="topol.top"
+f_conc=55430  # in mM/L
+
+
+edr_file_name=$dfnm".edr"
+xvg_file_name=$dfnm"_energies.xvg"
+boxdim_file_name=$dfnm"_box_dim.txt"
+top="lipid.top"
+
+if ! [ -s $edr_file_name ] 
+then
+    echo "We really need " $edr_file_name " , but we can't find it!"
+    exit 1
+fi
+
+# should give box-x,y,z
+echo 17 18 19 | gmx energy -f $edr_file_name -o $xvg_file_name > $boxdim_file_name
+
+#getting no. POPC lipids from topol.top file (if exists)
+if [ -f $top ]
+then
+    nlip=`grep -e "molecules" -A 10 $top | grep -e "^POPC" | cut -d " "  -f1 --complement `
+    nlip_per_leaflet=`echo $nlip/2.00 | bc`
+else
+    echo "Topology probably not present, can't get no. POPC lipids, assuming 64."
+    nlip_per_leaflet=64
+fi
+
+#CALCULATE APL (per POPC only!)
+python $scriptdir/calc_apl.py -f $boxdim_file_name -n $nlip_per_leaflet > apl_${nlip_per_leaflet}lip-leaflet.dat
+
+echo "Lipids per leaflet: " $nlip_per_leaflet
+cat apl_${nlip_per_leaflet}lip-leaflet.dat

### 2017-04-27 scripts/calcOPchanges.sh

@@ -18,6 +18,7 @@ awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.75
 awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_OxySigma_0.90_OPC3.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_OxySigma_0.90_OPC3.dat 
 awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_OxySigma_0.90_OPC3_NpAT.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_OxySigma_0.90_OPC3_NpAT.dat 
 awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_OxySigma_FBopt_OPC3.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_OxySigma_FBopt_OPC3.dat 
+awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_HGglyc-Sigmas_0.875_OPC3.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_HGglyc-Sigmas_0.875_OPC3.dat 
 
 awk -f calcOPchanges.awk ../Data/POPC/NaCl/Slipids-VIS/OPvsNACLconsALLSCALED.dat  > ../Data/POPC/NaCl/Slipids-VIS/SLIPIDnaclCONSchangeALLSCALED.dat 
 awk -f calcOPchanges.awk ../Data/POPC/NaCl/Slipids-VIS/OPvsNACLconsHGSCALED.dat  > ../Data/POPC/NaCl/Slipids-VIS/SLIPIDnaclCONSchangeHGSCALED.dat 

### 2017-05-10 Manuscript/proto-manuscript.tex
Initial commit of the new branch.
Wrote the outline of a proto manuscript.

--- 2017-03-18 Manuscript/manuscript.tex
@@
 
-\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}
+%\documentclass[aip,jcp,twocolumn]{revtex4}
+\documentclass[aip,jcp]{revtex4}
+%\documentclass{article}
 
@@
 \usepackage{csquotes}
+\usepackage{mhchem}
+
+%\usepackage{markdown}
 
@@
 
-\title{NMRlipids VI: New PC lipid model with correct cation binding affinity} %Title of paper
+\title{Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization} %Title of paper
 
@@
 \author{Josef Melcr}
-%\homepage[]{Your web page}
+\author{Hector Martinez-Seara Monne}
+\author{Pavel Jungwirth}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 % insert abstract here
-  Cellular membranes are complex systems that are hard to study both experimentally and theoretically. Classical molecular dynamics simulations (MD)
-give detailed information about membrane structure, dynamics and its interactions with other biomolecules. However,  there is still a room for
-improvements in the current force fields, especially in the hydrophilic region of membrane.  According to the results reported in NMRlipids, Open Collaboration project
-running at nmrlipids.blogspot.fi, biologically relevant Na+ and Ca2+ generally overbind to zwitterionic phophatidylcholine (PC) membranes in current MD models [Ollila 2016].
-Here we suggest that the PC lipid membrane-ion interactions can be corrected by implicitly inducing electronic polarizability in the lipid models by
-using a simple electrostatic continuum correction (ECC) for partial charges [Leontyev 2010], which is successfully applied already to monovalent and divalent ions in bulk water [Kohagen, 2015].
-Our results show that applying ECC to existing force fields significantly improved PC lipid--cation interactions when compared against NMR data
-by using the electrometer concept [Ollila 2016,Seelig87]. However, also other force field parameters have to be made consistent
-for ECC correction to preserve the area per molecule and other relevant bilayer properties. This work is carried as Open Collaboration
-in NMRlipids VI project (http://nmrlipids.blogspot.fi).
+Classical molecular dynamics simulations give detailed information about membrane structure and dynamics.
+However, there is still a room for improvements in current force fields – it is known from the literature, that the binding of ions, especially cations, to phopholipid membranes is overestimated in all classical models [1].
+We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent ions yielding models that feature correct ion pairing [3].
+Using the electrometer concept [3, 4], our pilot simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models.
+Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer.
+Poster presents a new lipid model, "ECClipids", that shows accurate affinity to sodium and calcium cations.
+This work will continue as an open collaboration project NMRlipids IV (http://nmrlipids.blogspot.fi).
+
+[1] Catte, A., Girych, M., Javanainen, M., Melcr J., Miettinen, M. S., Oganesyan, V. S. and Ollila H. S.,  PCCP 18(47) 32560-32569 (2016)
+[2] Leontyev, I. V., and Stuchebrukhov, A. A., JCTC 6(5) 1498–1508 (2010)
+[3] Kohagen, M., Mason, P. E., and Jungwirth, P., J. Phys. Chem. B 120(8) 1454–60 (2015)
+[4] Seelig, J., MacDonald, P. M., and Scherer, P. G., Biochemistry 26(24) 7535–7541 (1987)
 \end{abstract}
@@
 % References should be done using the \cite, \ref, and \label commands
+
+% Here I write pseudo-article statements that will make the main argument.
+% Beautiful polished sentences will be formed only afte we agree on these basic things
+\textbf{Introduction}
+
+ motivation, significance of membranes, phospholipids and simulation.
+
+ assumptions (so that we have structure of the paper like in a mathematical proof) -- 
+ MD simulation is a good tool for studying molecules.
+ classical MD models can describe lipids accurately,
+
+ MD is ... and it serves ... it is useful for ... (describe throuhg references)
+
+ Lipid membranes, especially phospholipid membranes; their significance for life, sciences, society, pharma ...
+
+\textbf{MDEC and current FFs}
+
+ Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties
+
+ lipid force fields fail in description of membrane-cation interaction -- could be answered by ECC?
+
+MDEC -- or -- ECC
+{MD in electronic continuum as in \cite{Leontyev2015} or electronic continuum correction as in \cite{Jungwirth2015}
+
+Describe ECC, good physical concept for treating part of the polarizability (electronic) in a simple mean-field way.
+
+Works for cations \cite{Jungwirth2015,Kohagen..} -- motivation for its application to zwitterionic lipids like POPC.
+
+Hypothesis: ECC helps in describing even zwitterionic molecules like POPC, will be demonstrated through headgroup oreder parametere response to cationic molecules.
+
+\textbf{Main body starts:}
+
+ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structur and energetics, but will fail in good interactions -- already proved for cations)
+
+How I apply ECC on Lipid14 POPC and why such choice (i.e. $f_\sigma$, $f_q$ and definition of the scaled region).
+
+why Lipid 14 -- good ratio of $\alpha / \beta$ response.
+
+\textbf{Observations:}
+
+Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant.
+
+Preserves other relevant properties: APL, Order Parameters (OPs), diffusion?
+\todoi{diffusion can be even improved}, elasticity?
+\todoi{check elastic properties -- do they change, improve, worsen??} -- however, it is still garbage in-garbage out
+
+Gives correct scattering form factor -- good groud for interpretting experimental data.
+
+Provides correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
+
+\textbf{Discussion:}
+
+other lipids: charged lipid? -- ongoing research in our lab (starting to do something with Aniket on POPE for curved membranes)
+
+Role of water model: we use OPC3 (current best), it would be worth giving an estimate how results change when we use say SPCE or even TIP3p at least in the SI (so that the reader knows what errors to expect comming from these sub-optimal models).
+In addition, there is protein force field Amber15-FB, which uses water close to OPC3, TIP3pFB \todoii{REFs}{add references here}.
+
+questionable charge distribution from RESP -- a leeway for further tweaks of the FF
+
+acheivable accuracy of the MD engines themselves (mainly Hector's worry).
+
+application of the correction to other lipid models: The rule looks general, however, it depends on how accurate the original model was.
+
+\textbf{Conclusion:}
+
+Say what we did...
+
+This will be a foundation stone of a new open-collaboration project NPRlipid 6 in nmrlipids.blogspot.fi....
 

### 2017-05-11 Manuscript/proto-manuscript.tex

@@
 MDEC -- or -- ECC 
-{MD in electronic continuum as in \cite{Leontyev2015} or electronic continuum correction as in \cite{Jungwirth2015}
+\todoii{choose one}{We have to decide whether we will refer to the method as a correction (ECC) or as a MD simulation paradigm (MDEC) -- and choose on of these labels.}
+MD in electronic continuum as in \cite{Leontyev2015} or electronic continuum correction as in \cite{Jungwirth2015}
 
-Describe ECC, good physical concept for treating part of the polarizability (electronic) in a simple mean-field way.
+Describe ECC: good physical concept for treating part of the polarizability (electronic) in a simple mean-field way.
 
-Works for cations \cite{Jungwirth2015,Kohagen..} -- motivation for its application to zwitterionic lipids like POPC.
+Succesful application: Works for cations \cite{Jungwirth2015,Kohagen..} -- motivation for its application to zwitterionic lipids like POPC.
 
@@
 ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structur and energetics, but will fail in good interactions -- already proved for cations)
+In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC framework -- part of $\Delta G$ is already included in the polarization of electrons, and only the remining part of $\Delta G$ belongs to the polarization of the nuclei.
+Interestingly, the part of $\Delta G$ due to electrons is approximately $1/2$ of it, which is what I observe in my simulations -- if I assume that half of this energy is already accounted for by the water model (which is rather a MDEC model already), than I should recover approximately $3/4$ of the original interaction energy.
+When I write the formula for interaction energy for two particles at a distance $r=\sigma$ (approximate equilibrium position), I get
+\begin{equation}
+W(f_\sigma r,f_q q) = \kappa \frac{q_i q_j}{r} \frac{f_q^2}{f_\sigma}
+\end{equation}
+
+When I apply $3/4$-assumption for the intereaction energy, I require: 
+\begin{equation}
+\frac{3}{4} W(r,q) = W(f_\sigma r, f_q q)
+\end{equation}
+
+So for $f_q=0.8$ I obtain $f_\sigma ~= \frac{4}{3} f_q^2 \,\,\, = 0.853$ -- which is very close to the value I use, $f_\sigma=0.875$.
+This is a crude derivation neglecting the true position of energy minimum and the LJ mixing rules.
 

### 2017-05-11 Manuscript/proto-manuscript.tex
Small addition to the MS: simple formula for interaction energy with scaling.
Small edits in proto-MS

@@
 We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent ions yielding models that feature correct ion pairing [3].
-Using the electrometer concept [3, 4], our pilot simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models.
+Using the electrometer concept [3, 4], our simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models.
 Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer.
-Poster presents a new lipid model, "ECClipids", that shows accurate affinity to sodium and calcium cations.
+The new lipid model, "ECClipids", shows accurate affinity to sodium and calcium cations. 
 This work will continue as an open collaboration project NMRlipids IV (http://nmrlipids.blogspot.fi).
@@
 % Beautiful polished sentences will be formed only afte we agree on these basic things
-\textbf{Introduction}
+\textbf{General Introduction}
 
@@
 
-\textbf{MDEC and current FFs}
+\textbf{Specific introduction: MDEC and current FFs}
 
@@
 
-Succesful application: Works for cations \cite{Jungwirth2015,Kohagen..} -- motivation for its application to zwitterionic lipids like POPC.
+Succesful application: Works for cations \cite{Jungwirth2015,Kohagen..}\todoii{REF}{Missing references} -- motivation for its application to zwitterionic lipids like POPC.
 
-Hypothesis: ECC helps in describing even zwitterionic molecules like POPC, will be demonstrated through headgroup oreder parametere response to cationic molecules.
+Hypothesis: ECC helps in describing even zwitterionic molecules like POPC, will be demonstrated through headgroup order parameter response to cationic molecules.
 
@@
 
-ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structur and energetics, but will fail in good interactions -- already proved for cations)
+ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
 In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC framework -- part of $\Delta G$ is already included in the polarization of electrons, and only the remining part of $\Delta G$ belongs to the polarization of the nuclei.
@@
 This is a crude derivation neglecting the true position of energy minimum and the LJ mixing rules.
+However, the formula for $f_\sigma$ does not depend on the absolute position and I believe that the assumption that the minimum is relatively shifted by scaling with $f_\sigma$ holds pretty well.
 

### 2017-05-12 Manuscript/proto-manuscript.tex
Restructured text in proto-MS, now it follows only the main message with a few excerpts from Samuli's writing.

@@
 
-\textbf{Specific introduction: MDEC and current FFs}
+\textbf{Specific introduction: current FFs and MDEC/ECC}
 
- Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties
+ Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties.
 
  lipid force fields fail in description of membrane-cation interaction -- could be answered by ECC? 
+Cations were shown to generally overbind in PC lipid bilayers in
+NMRlipids II project. Here we propose that the cation overbinding can
+be corrected by implicitly inducing electronic polarizability in
+lipid headgroups by scaling the partial charges.
 
@@
 
-How I apply ECC on Lipid14 POPC and why such choice (i.e. $f_\sigma$, $f_q$ and definition of the scaled region).
+How do I apply ECC on Lipid14 POPC and why such choice (i.e. $f_\sigma$, $f_q$ and definition of the scaled region).
 
-why Lipid 14 -- good ratio of $\alpha / \beta$ response.
+why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response.
 
-\textbf{Observations:}
+\textbf{Observations, how does the system behave, what are its macroscopic properties:}
 
 Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant. 
+The response of the headgroup order parameters is improved both through diminished affinity towards the cations and through lower headgroup response to the bound charge (from simulations with \ce{DHMDMAB} surfactant).
+The best models from Fig.  \ref{OrderParameterIONSchangesSCALED} are shown in
+Fig. \ref{OrderParameterCHANGESnewMODELS}. Based on these results we choose
 
@@
 \todoi{check elastic properties -- do they change, improve, worsen??} -- however, it is still garbage in-garbage out
+The glycerol backbone and headgroup order parameters are plotted in
+Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
+i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
 
-Gives correct scattering form factor -- good groud for interpretting experimental data.
+Gives correct scattering form factor -- good ground for interpretting experimental data.
 
-Provides correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
+Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
 
@@
 
-questionable charge distribution from RESP -- a leeway for further tweaks of the FF
+questionable charge distribution from RESP -- a leeway for further tweaks of the FF.
+It is not obvious that RESP charges provide the best description, especially due to its non-uniqe solution. 
+Shall we solve RESP fitting with the constraint of full charges and then scale down, or shall we rather solve the fitting with a scaled total charge target?
 
-acheivable accuracy of the MD engines themselves (mainly Hector's worry).
+acheivable accuracy of the MD engines themselves (mainly Hector's worry) is another limiting factor in fine tuning parameters -- solid physical ground helps.
 
 application of the correction to other lipid models: The rule looks general, however, it depends on how accurate the original model was.
+From preliminary simulations with POPE, it looks that the rule works for Lipid14 FF, at least for zwitterionic headgroups.
 
@@
 
-Say what we did...
+Reiterate what we did...
 
@@
 
-%\label{}
-\section{Introduction}
-
-Cations were shown to generally overbind in PC lipid bilayers in
-NMRlipids II project. Here we propose that the cation overbinding can
-be corrected by implicitly inducing electronic polarizability in
-lipid headgroups by scaling the partial charges.
-
-%\section{Methods}
-
-%\subsection{Molecular dynamics simulations}
-
-\section{Effect of scaling of headgroup and glycerol backbone partial charges on lipid bilayer properties and cation binding}
-
-Headgroup and glycerol backbone partial charges of Slipids and Lipid14 models were scaled
-with various scaling factors and cation binding was monitored by using the electrometer
-concept. Headgroup order parameter changes as a function of ion concentration are
-shown in Fig. \ref{OrderParameterIONSchangesSCALED}. The order parameter decrease
-seemst to be proportional to the headgroup charges: the smaller the charges
-(larger scaling factor), the less change is observed in order parameters.
-According to the electrometer concept, this means that the headgroup scaling
-decreases cation binding affinity. However, from this data only we cannot
-exclude the possibility that order parameters are changing less because
-headgroup becomes less sensitive for cation binding.
 \begin{figure}[]
@@
 
-The best models from Fig.  \ref{OrderParameterIONSchangesSCALED} are shown in
-Fig. \ref{OrderParameterCHANGESnewMODELS}. Based on these results we choose
-two models for more careful studies: Slipids and Lipid 14 models with headroup
-and glycerol backbone partial charges scaled with 0.8 and 0.85, respectively.
 \begin{figure}[]
@@
 \end{figure}
-
-While scaling seems to improve ion binding, it changes the lipid bilayer properties
-without ions. The glycerol backbone and headgroup order parameters are plotted in
-Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
-i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
-The area per lipid from different models are shown in Table \ref{apls}
-The conclusion is that the charge scaling has only a marginal effect on
-headgroup order parameters, while area per lipid is significantly decreased.
-Area per lipid decrease is more pronounced in Lipid14 than in Slipids.
 
@@
     Lipid14 (literature)  & 65.6$\pm$ 0.5   \\
-    Lipid14eec0.85 & 55.5      \\
+    Lipid14ecc0.80+sigma0.875 & 64.3      \\
     Slipids (literature T=303K)       & 64.6 $\pm$ 0.4   \\
@@
 \end{table}
-
-The Calcium densities from new models and standard models are compared in Fig \ref{CAdensities}.
-It seems to me that Ca binding in Lipid14 model is not reduced by scaling the
-partial charges even thought the order parameter response is better.
-On the other hand, in Slipids the binding seems to be reduced.
-Possible interpretation is that Lipid14 result gets better because headgroup
-gets less sensitive for bound charge, while in Slipid the binding is reduced.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/CAdensities.eps}
-  \caption{\label{CAdensities}
-    Calcium densities from different simulations with new and standard models.
-  }
-\end{figure}
-
-\section{Conclusions}
 

### 2017-05-12 Manuscript/proto-manuscript.tex
Change in the reasoning for f_sigma value.

@@
 ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
-In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC framework -- part of $\Delta G$ is already included in the polarization of electrons, and only the remining part of $\Delta G$ belongs to the polarization of the nuclei.
-Interestingly, the part of $\Delta G$ due to electrons is approximately $1/2$ of it, which is what I observe in my simulations -- if I assume that half of this energy is already accounted for by the water model (which is rather a MDEC model already), than I should recover approximately $3/4$ of the original interaction energy.
-When I write the formula for interaction energy for two particles at a distance $r=\sigma$ (approximate equilibrium position), I get
-\begin{equation}
-W(f_\sigma r,f_q q) = \kappa \frac{q_i q_j}{r} \frac{f_q^2}{f_\sigma}
-\end{equation}
-
-When I apply $3/4$-assumption for the intereaction energy, I require: 
-\begin{equation}
-\frac{3}{4} W(r,q) = W(f_\sigma r, f_q q)
-\end{equation}
-
-So for $f_q=0.8$ I obtain $f_\sigma ~= \frac{4}{3} f_q^2 \,\,\, = 0.853$ -- which is very close to the value I use, $f_\sigma=0.875$.
-This is a crude derivation neglecting the true position of energy minimum and the LJ mixing rules.
-However, the formula for $f_\sigma$ does not depend on the absolute position and I believe that the assumption that the minimum is relatively shifted by scaling with $f_\sigma$ holds pretty well.
+In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC paradigm -- part of $\Delta G$ is already included in the polarization of electrons, and only the remining part of $\Delta G$ belongs to the polarization of the nuclei.
+Hence it is non trivial, how large should be the scaling factor $f_\sigma$ -- it should lay between $1$ and $f_q$, the limits for the original interaction energy and its limit when we neglect the $\Delta G_{el}$ term in $\Delta G_{hyd}$.
 

### 2017-05-12 Manuscript/proto-manuscript.tex
Included some of Samuli's comments

@@
 Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
+Samuli: the main selling argument would be that we solve the details (stoichiometry, coordination numbers, etc.) of cation binding in PC bilayer.
+
+\textbf{Required figures}
+
+Figure comparing order parameters and form factor between original Lipid14, new model and experiments,
+
+- Figure showing order parameters as a function of \ce{NaCl} and \ce{CaCl_2} concentrations from Lipid14, new model and experiments (like current Fig. 1, but only original and final models)
+
+- Figure(s) analysing binding details (density profiles, lipids bound per ion, something else?).
 
@@
 other lipids: charged lipid? -- ongoing research in our lab (starting to do something with Aniket on POPE for curved membranes)
+Samuli: From the point of view of this paper, the most relevant other lipid to study would be DPPC to see if we can reproduce the difference between POPC and DPPC in experiments.
 

### 2017-05-12 Manuscript/proto-manuscript.tex
Proto-MS edits after a very brief discussion with Pavel

--- 2017-05-12 Manuscript/proto-manuscript.tex
@@
 We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent ions yielding models that feature correct ion pairing [3].
-Using the electrometer concept [3, 4], our simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models.
+Using the electrometer concept [3, 4] and x-ray scattering form factors, our simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models.
 Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. 
-The new lipid model, "ECClipids", shows accurate affinity to sodium and calcium cations. 
+The new lipid model, "ECClipids", shows accurate binding affinity to sodium and calcium cations and headgroup order parameter response to bound charge. 
+We also provide for the first time a realistic stochiometry of bound calcium cations to a POPC membrane, and their binding sites. 
 This work will continue as an open collaboration project NMRlipids IV (http://nmrlipids.blogspot.fi).
@@
 
- Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties.
+ Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties.  -- write basic ideas from the lipid-FF whitepaper I did recently.
 
  lipid force fields fail in description of membrane-cation interaction -- could be answered by ECC? 
-Cations were shown to generally overbind in PC lipid bilayers in
-NMRlipids II project. Here we propose that the cation overbinding can
-be corrected by implicitly inducing electronic polarizability in
-lipid headgroups by scaling the partial charges.
+Cations were shown to generally overbind in PC lipid bilayers in NMRlipids II project. 
+Here we propose that the cation overbinding can be corrected by implicitly inducing electronic polarizability in lipid headgroups by scaling the partial charges -- i.e.
+MDEC/ECC~\cite{Leontyev2015}\todo{Bulid BibTex references database.} paradigm.
 
@@
 
-ECC and solvation $\Delta G$.(from \cite{Leontyev2015}: hydration $\Delta G$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
-In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC paradigm -- part of $\Delta G$ is already included in the polarization of electrons, and only the remining part of $\Delta G$ belongs to the polarization of the nuclei.
+ECC and solvation $\Delta G$.(from \cite{Leontyev2015} \todoi{Pavel: do not replicate Stuchebrukhov paper, write it very briefly}: hydration $\Delta  G_{hyd}$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
+In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC paradigm -- part of $\Delta G = \Delta G_{nuc} + \Delta G_{el}$ is already included in the polarization of electrons, and only the remining part of $\Delta G$,  $\Delta G_{nuc}$, belongs to the polarization of the nuclei.
 Hence it is non trivial, how large should be the scaling factor $f_\sigma$ -- it should lay between $1$ and $f_q$, the limits for the original interaction energy and its limit when we neglect the $\Delta G_{el}$ term in $\Delta G_{hyd}$.
@@
 How do I apply ECC on Lipid14 POPC and why such choice (i.e. $f_\sigma$, $f_q$ and definition of the scaled region).
+Region definition: headgroup requires optimization, tails appear already accurate enough. 
+Choice of $f_q$: value 0.8 reflects the fact that the charges in Lipid14 were derived in vacuum, whereas they should rather be the average of vacuum and solvated charges (so called IPolQ charges \cite{Cerutti2013}, both charge-sets can be taken from \cite{Maciejewski2014a}).
 
-why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response.
+why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response~\cite{NMRlipid2paper}.
 
@@
 
+Accurate binding of \ce{Na^+} and \ce{Ca^{2+}} cations.
+
 Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant.
-The response of the headgroup order parameters is improved both through diminished affinity towards the cations and through lower headgroup response to the bound charge (from simulations with \ce{DHMDMAB} surfactant).
-The best models from Fig.  \ref{OrderParameterIONSchangesSCALED} are shown in
-Fig. \ref{OrderParameterCHANGESnewMODELS}. Based on these results we choose
 
-Preserves other relevant properties: APL, Order Parameters (OPs), diffusion?
-\todoi{diffusion can be even improved}, elasticity?
-\todoi{check elastic properties -- do they change, improve, worsen??} -- however, it is still garbage in-garbage out
+What affects the headgroup OP response?: The response of the headgroup order parameters is improved both through diminished affinity towards the cations and through lower headgroup response to the bound charge (from simulations with \ce{DHMDMAB} surfactant).
+The responses of headgroup OPs are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+
+Gives correct scattering form factor -- reliable structural properties (especially in combination with OPs) and good ground for interpretting experimental data.
+
+Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
+Samuli: the main selling argument would be that we solve the details (stoichiometry, coordination numbers, etc.) of cation binding in PC bilayer.
+
+Preserves other relevant properties: APL, Order Parameters (OPs of headgroup and tails), 
+diffusion? \todoi{diffusion can be even improved -- and doesn't have to. We may as well omit it.}, 
+elasticity? \todoi{check elastic properties -- do they change, improve, worsen?? -- likely not crucial for the main message of the paper.} 
+-- however, it is still garbage in-garbage out
+
 The glycerol backbone and headgroup order parameters are plotted in
@@
 
-Gives correct scattering form factor -- good ground for interpretting experimental data.
+\begin{table}
+  \caption{Area per lipid from different models for POPC without ions\label{apls} }
+  \begin{tabular}{c c c}
+    model          & A (Å$^2$)   & Temperature [K] \\
+    Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
+    Lipid14ecc0.80+sigma0.875 & 63.7 -- 64.9 &  313    \\
+    experiment \cite{Jambeck2012}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
+    experiment  & 64.3  &  303    \\
+    experiment  & 67.3  &  323    \\
+    experiment  & 68.1  &  333    \\
+    experiment POPE  & 56.6 &  303    \\
+  \end{tabular}
+\end{table}
 
-Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
+\textbf{Required figures}
+
+- Figure showing the system: POPC, membrane, atoms involved in headgroup OPs.\todoi{This is probably sufficient as the last thing we would do -- and will probably provide a good basis for TOC graphic.}
+
+- Figure comparing order parameters (Fig.~\ref{HGopsNEWmodels}) and form factor \todoii{Add plot}{Add the form factor plot} between original Lipid14, new model and experiments,
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodels.eps}
+  \caption{\label{HGopsNEWmodels}
+    Headgroup and glycerol backbone order parameters from standard and EEC models.}
+\end{figure}
+
+- Figure showing order parameters as a function of \ce{NaCl} and \ce{CaCl_2} concentrations~
+\todoii{ongoing}{Actual concentration of cations in simulation has yet to be estimated. 
+If it varies too much from the nominal concentration, I may need to tweak the scaling factors, $f_q$ or only $f_\sigma$, to accomodate it. 
+However, it is very unlikely, response to the surfactant \ce{DHMDMAB} is OK. 
+Big patches with loads of solvent are running at the moment to guide me on the possible finite-size errors and this conc-error.} 
+from Lipid14, new model and experiments (like current Fig. 1, but only original and final models)
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=9.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
+  \caption{\label{OrderParameterCHANGESnewMODELS}
+    Headgroup order parameter changes as a function of cations from models
+    with modified headgroup partial charges.}
+\end{figure}
+
+- Figure(s) analysing binding details (density profiles, lipids bound per ion, something else?).
 
@@
 
-other lipids: charged lipid? -- ongoing research in our lab (starting to do something with Aniket on POPE for curved membranes)
+other lipids: charged lipids? -- ongoing research in our lab.
+I'm currently working on POPE with Aniket for curved (and flat) membranes.
+
+Samuli: From the point of view of this paper, the most relevant other lipid to study would be DPPC to see if we can reproduce the difference between POPC and DPPC in experiments.
+
+Joe: In addition, there are more experimental OP data on DPPC. However, it is not necessary and we can leave it to the community project.
 
@@
 This will be a foundation stone of a new open-collaboration project NPRlipid 6 in nmrlipids.blogspot.fi....
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/OrderParameterIONSchangesSCALED.eps}
-  \caption{\label{OrderParameterIONSchangesSCALED}
-    Headgroup order parameter changes as a function of cations from models
-    with modified headgroup partial charges.}
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
-  \caption{\label{OrderParameterCHANGESnewMODELS}
-    Headgroup order parameter changes as a function of cations from models
-    with modified headgroup partial charges.}
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodels.eps}
-  \caption{\label{HGopsNEWmodels}
-    Headgroup and glycerol backbone order parameters from standard and EEC models.}
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodelsCHANGE.eps}
-  \caption{\label{HGopsNEWmodelsCHANGE}
-    Changes in headgroup and glycerol backbone order parameters due to EEC.}
-\end{figure}
-
-\begin{table}
-  \caption{Area per lipid from different models for POPC without ions\label{apls} }
-  \begin{tabular}{c c}
-    model          & A (Å$^2$)    \\
-    Lipid14 (literature)  & 65.6$\pm$ 0.5   \\
-    Lipid14ecc0.80+sigma0.875 & 64.3      \\
-    Slipids (literature T=303K)       & 64.6 $\pm$ 0.4   \\
-    SlipidsEEC0.8  & 57.8  \\
-  \end{tabular}
-\end{table}
 

### 2017-05-17 Manuscript/proto-manuscript.tex
proto-MS update

@@
 MDEC -- or -- ECC 
-\todoii{choose one}{We have to decide whether we will refer to the method as a correction (ECC) or as a MD simulation paradigm (MDEC) -- and choose on of these labels.}
+\todoii{choose one}{We have to decide whether we will refer to the method as a correction (ECC) or as a MD simulation paradigm (MDEC) -- and choose on of these labels.
+-- Joe: use ECC as we apply MDEC rather as a correction to current state not as a new paradigm for FF development.}
 MD in electronic continuum as in \cite{Leontyev2015} or electronic continuum correction as in \cite{Jungwirth2015}
@@
     model          & A (Å$^2$)   & Temperature [K] \\
+    \hline
     Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
-    Lipid14ecc0.80+sigma0.875 & 63.7 -- 64.9 &  313    \\
+    Lipid14ecc0.80+sigma0.875 &        &  313    \\
+    GMX small patch           & 64.9   &         \\
+    GMX 4xbig patch           & 65.5   &         \\
+    oMM small patch           & 63.65  &         \\
+    oMM 4xbig patch           & 63.7   &         \\
+    \hline
     experiment \cite{Jambeck2012}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
@@
     experiment POPE  & 56.6 &  303    \\
+    \hline
   \end{tabular}
@@
   \centering
-  \includegraphics[width=9.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
+  \includegraphics[width=14.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
   \caption{\label{OrderParameterCHANGESnewMODELS}

### 2017-05-17 scripts/calcOPchanges.sh
New: topologies for L14 q_sc 0.8 sig_sc 0.88&0.89, topology for L14 q_sc 0.8 sig_sc 0.875, new data: order parameters for simulations with these models and larger patches

@@ -35,3 +35,7 @@ awk -f calcOPchanges.awk ../Data/POPC/NaCl/lipid14/OPvsNACLconsHGglycSCALED_0.85
 awk -f calcOPchanges.awk ../Data/POPC/NaCl/lipid14/OPvsNACLconsHGglycSCALED_0.85_OPC3.dat  > ../Data/POPC/NaCl/lipid14/LIPID14naclCONSchangeHGglycSCALED_0.85_OPC3.dat 
 awk -f calcOPchanges.awk ../Data/POPC/NaCl/lipid14/OPvsNACLconsHGglycSCALED_0.85_OPC3_NpAT.dat  > ../Data/POPC/NaCl/lipid14/LIPID14naclCONSchangeHGglycSCALED_0.85_OPC3_NpAT.dat 
 awk -f calcOPchanges.awk ../Data/POPC/NaCl/lipid14/OPvsNACLconsHGglycSCALED_0.80_HGglyc-Sigmas_0.875_OPC3.dat  > ../Data/POPC/NaCl/lipid14/LIPID14naclCONSchangeHGglycSCALED_0.80_HGglyc-Sigmas_0.875_OPC3.dat 
+
+cd ../Fig/scripts
+gnuplot plotOPchangesNEWmodels.gpl
+

### 2017-05-18 Manuscript/proto-manuscript.tex
MS: Figures updated

@@
   \centering
-  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodels.eps}
+  \includegraphics[width=9.0cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+  \caption{\label{fig:form-f}
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.}
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
   \caption{\label{HGopsNEWmodels}
-    Headgroup and glycerol backbone order parameters from standard and EEC models.}
+    Headgroup and glycerol backbone order parameters from standard Lipid14 and EECLipid17 models.}
 \end{figure}
@@
   \centering
-  \includegraphics[width=14.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
+  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2_surf.eps}
   \caption{\label{OrderParameterCHANGESnewMODELS}
-    Headgroup order parameter changes as a function of cations from models
-    with modified headgroup partial charges.}
+    Headgroup order parameter changes as a function of concentration of cations or cationic surfactant DHMDMAB. }
 \end{figure}

### 2017-05-18 Manuscript/proto-manuscript.tex

@@
 
-- Figure comparing order parameters (Fig.~\ref{HGopsNEWmodels}) and form factor \todoii{Add plot}{Add the form factor plot} between original Lipid14, new model and experiments,
+- Figure comparing order parameters (Fig.~\ref{HGopsNEWmodels}) and form factor \todoi{The original L14 form-f are worse than those in the L14 paper -- I'd better double check it and make sure I use the same settings and version for all simulations consistently for a good comparison.}between original Lipid14, new model and experiments,
 
@@
   \caption{\label{fig:form-f}
-    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.}
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
 \end{figure}
@@
 - Figure(s) analysing binding details (density profiles, lipids bound per ion, something else?).
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities.eps}
+  \caption{\label{fig:cacl-dens}
+    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Aquist parameters and with ECC ions and ECCLipids17 with ECC ions. }
+\end{figure}
 

### 2017-05-19 Manuscript/proto-manuscript.tex
Some updates

@@
 
-%\documentclass[aip,jcp,twocolumn]{revtex4}
-\documentclass[aip,jcp]{revtex4}
+\documentclass[aip,jcp,twocolumn]{revtex4}
+%\documentclass[aip,jcp]{revtex4}
 %\documentclass{article}
@@
 % Beautiful polished sentences will be formed only afte we agree on these basic things
-\textbf{General Introduction}
+\section{Introduction}
 
@@
 
-\textbf{Specific introduction: current FFs and MDEC/ECC}
+%\textbf{Specific introduction: current FFs and MDEC/ECC}
 
@@
 MDEC -- or -- ECC 
-\todoii{choose one}{We have to decide whether we will refer to the method as a correction (ECC) or as a MD simulation paradigm (MDEC) -- and choose on of these labels.
+\todo{choose one}{We have to decide whether we will refer to the method as a correction (ECC) or as a MD simulation paradigm (MDEC) -- and choose on of these labels.
 -- Joe: use ECC as we apply MDEC rather as a correction to current state not as a new paradigm for FF development.}
@@
 
-\textbf{Main body starts:}
+\section{Methods}
 
@@
 
-\textbf{Observations, how does the system behave, what are its macroscopic properties:}
+\section{Results and Discussion}
 
-Accurate binding of \ce{Na^+} and \ce{Ca^{2+}} cations.
-
-Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant.
-
-What affects the headgroup OP response?: The response of the headgroup order parameters is improved both through diminished affinity towards the cations and through lower headgroup response to the bound charge (from simulations with \ce{DHMDMAB} surfactant).
-The responses of headgroup OPs are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-
-Gives correct scattering form factor -- reliable structural properties (especially in combination with OPs) and good ground for interpretting experimental data.
-
-Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
-Samuli: the main selling argument would be that we solve the details (stoichiometry, coordination numbers, etc.) of cation binding in PC bilayer.
-
-Preserves other relevant properties: APL, Order Parameters (OPs of headgroup and tails), 
-diffusion? \todoi{diffusion can be even improved -- and doesn't have to. We may as well omit it.}, 
-elasticity? \todoi{check elastic properties -- do they change, improve, worsen?? -- likely not crucial for the main message of the paper.} 
--- however, it is still garbage in-garbage out
-
-The glycerol backbone and headgroup order parameters are plotted in
-Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
-i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
-
+Detailed and robust structural information of lipid bilayers can
+be reached with C-H bond order parameter from NMR and form factors from scattering
+experiments. The formed can be used to validate structures sampled by individual
+molecules and latter to validate overall bilayer properties, lateral density
+(area per molecule) and thickness \cite{ollila16}. The structural quality of
+new ECC lipid model against scattering form factors and NMR order parameters
+is evaluated in Fig. \ref{simVSexpNOions}
+\todo{Add acyl chain order parameters, POPC chemical structure and fix experimental alpha order parameter value}
+\todo{Why original Lipid14 models seems to deviate from experiments, in contrast to the original publication?}
+\begin{figure}[]
+  \centering
+  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+%  \caption{\label{fig:form-f}
+%    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
+%\end{figure}
+%
+%\begin{figure}[]
+%  \centering
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
+  \caption{\label{simVSexpNOions}
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models. 
+    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and EECLipid17 models
+    compared with experimental order parameters from \cite{ferreira13}.}
+\end{figure}
+Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
+are shown in Table \ref{apls}
+\todo{finalize figure and the discussion}
 \begin{table}
@@
 
+Preserves other relevant properties: APL, Order Parameters (OPs of headgroup and tails), 
+diffusion? \todoi{diffusion can be even improved -- and doesn't have to. We may as well omit it.}, 
+elasticity? \todoi{check elastic properties -- do they change, improve, worsen?? -- likely not crucial for the main message of the paper.} 
+-- however, it is still garbage in-garbage out
+The glycerol backbone and headgroup order parameters are plotted in
+Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
+i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
+
 \textbf{Required figures}
@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-  \caption{\label{fig:form-f}
-    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
-\end{figure}
+Accurate binding of \ce{Na^+} and \ce{Ca^{2+}} cations.
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
-  \caption{\label{HGopsNEWmodels}
-    Headgroup and glycerol backbone order parameters from standard Lipid14 and EECLipid17 models.}
-\end{figure}
+Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant.
+
+What affects the headgroup OP response?: The response of the headgroup order parameters is improved
+both through diminished affinity towards the cations and through lower headgroup response to the bound charge
+(from simulations with \ce{DHMDMAB} surfactant). 
+The responses of headgroup OPs are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+
+Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
+Samuli: the main selling argument would be that we solve the details (stoichiometry, coordination numbers, etc.) of cation binding in PC bilayer.
 
@@
 
-\begin{figure}[]
+\begin{figure*}[]
   \centering
@@
     Headgroup order parameter changes as a function of concentration of cations or cationic surfactant DHMDMAB. }
-\end{figure}
+\end{figure*}
 

### 2017-05-22 scripts/calcOPchanges.sh
Minor update on Data ECClipids17

@@ -19,6 +19,7 @@ awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80
 awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_OxySigma_0.90_OPC3_NpAT.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_OxySigma_0.90_OPC3_NpAT.dat 
 awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_OxySigma_FBopt_OPC3.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_OxySigma_FBopt_OPC3.dat 
 awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_HGglyc-Sigmas_0.875_OPC3.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_HGglyc-Sigmas_0.875_OPC3.dat 
+awk -f calcOPchanges.awk ../Data/POPC/CaCl/lipid14/OPvsCACLconsHGglycSCALED_0.80_HGglyc-Sigmas_0.89_OPC3.dat  > ../Data/POPC/CaCl/lipid14/LIPID14caclCONSchangeHGglycSCALED_0.80_HGglyc-Sigmas_0.89_OPC3.dat 
 
 awk -f calcOPchanges.awk ../Data/POPC/NaCl/Slipids-VIS/OPvsNACLconsALLSCALED.dat  > ../Data/POPC/NaCl/Slipids-VIS/SLIPIDnaclCONSchangeALLSCALED.dat 
 awk -f calcOPchanges.awk ../Data/POPC/NaCl/Slipids-VIS/OPvsNACLconsHGSCALED.dat  > ../Data/POPC/NaCl/Slipids-VIS/SLIPIDnaclCONSchangeHGSCALED.dat 
@@ -35,6 +36,7 @@ awk -f calcOPchanges.awk ../Data/POPC/NaCl/lipid14/OPvsNACLconsHGglycSCALED_0.85
[…]

### 2017-05-23 Manuscript/proto-manuscript.tex

@@
 Here we propose that the cation overbinding can be corrected by implicitly inducing electronic polarizability in lipid headgroups by scaling the partial charges -- i.e.
-MDEC/ECC~\cite{Leontyev2015}\todo{Bulid BibTex references database.} paradigm.
+MDEC/ECC~\cite{leontyev11}\todo{Bulid BibTex references database.} paradigm.
 
@@
 -- Joe: use ECC as we apply MDEC rather as a correction to current state not as a new paradigm for FF development.}
-MD in electronic continuum as in \cite{Leontyev2015} or electronic continuum correction as in \cite{Jungwirth2015}
+MD in electronic continuum as in \cite{leontyev11} or electronic continuum correction as in \cite{Jungwirth2015}
 
@@
 
-Succesful application: Works for cations \cite{Jungwirth2015,Kohagen..}\todoii{REF}{Missing references} -- motivation for its application to zwitterionic lipids like POPC.
+Succesful application: Works for cations \cite{Jungwirth2015,kohagen14,kohagen16}\todoii{REF}{Missing references} -- motivation for its application to zwitterionic lipids like POPC.
 
@@
 
-ECC and solvation $\Delta G$.(from \cite{Leontyev2015} \todoi{Pavel: do not replicate Stuchebrukhov paper, write it very briefly}: hydration $\Delta  G_{hyd}$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
+ECC and solvation $\Delta G$.(from \cite{leontyev11} \todoi{Pavel: do not replicate Stuchebrukhov paper, write it very briefly}: hydration $\Delta  G_{hyd}$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
 In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC paradigm -- part of $\Delta G = \Delta G_{nuc} + \Delta G_{el}$ is already included in the polarization of electrons, and only the remining part of $\Delta G$,  $\Delta G_{nuc}$, belongs to the polarization of the nuclei.
@@
 Region definition: headgroup requires optimization, tails appear already accurate enough. 
-Choice of $f_q$: value 0.8 reflects the fact that the charges in Lipid14 were derived in vacuum, whereas they should rather be the average of vacuum and solvated charges (so called IPolQ charges \cite{Cerutti2013}, both charge-sets can be taken from \cite{Maciejewski2014a}).
+Choice of $f_q$: value 0.8 reflects the fact that the charges in Lipid14 were derived in vacuum, whereas they should rather be the average of vacuum and solvated charges (so called IPolQ charges \cite{Cerutti2013}, both charge-sets can be taken from \cite{maciejewski14}).
 
-why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response~\cite{NMRlipid2paper}.
+why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response~\cite{catte16}.
 
@@
 be reached with C-H bond order parameter from NMR and form factors from scattering
-experiments. The formed can be used to validate structures sampled by individual
+experiments. The former can be used to validate structures sampled by individual
 molecules and latter to validate overall bilayer properties, lateral density
@@
 
-Preserves other relevant properties: APL, Order Parameters (OPs of headgroup and tails), 
-diffusion? \todoi{diffusion can be even improved -- and doesn't have to. We may as well omit it.}, 
-elasticity? \todoi{check elastic properties -- do they change, improve, worsen?? -- likely not crucial for the main message of the paper.} 
--- however, it is still garbage in-garbage out
-The glycerol backbone and headgroup order parameters are plotted in
-Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
-i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
+%Preserves other relevant properties: APL, Order Parameters (OPs of headgroup and tails), 
+%diffusion? \todoi{diffusion can be even improved -- and doesn't have to. We may as well omit it.}, 
+%elasticity? \todoi{check elastic properties -- do they change, improve, worsen?? -- likely not crucial for the main message of the paper.} 
+%-- however, it is still garbage in-garbage out
+%The glycerol backbone and headgroup order parameters are plotted in
+%Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
+%i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
 
-\textbf{Required figures}
+%\textbf{Required figures}
 
-- Figure showing the system: POPC, membrane, atoms involved in headgroup OPs.\todoi{This is probably sufficient as the last thing we would do -- and will probably provide a good basis for TOC graphic.}
+%- Figure showing the system: POPC, membrane, atoms involved in headgroup OPs.\todoi{This is probably sufficient as the last thing we would do -- and will probably provide a good basis for TOC graphic.}
 
-- Figure comparing order parameters (Fig.~\ref{HGopsNEWmodels}) and form factor \todoi{The original L14 form-f are worse than those in the L14 paper -- I'd better double check it and make sure I use the same settings and version for all simulations consistently for a good comparison.}between original Lipid14, new model and experiments,
+%- Figure comparing order parameters (Fig.~\ref{HGopsNEWmodels}) and form factor \todoi{The original L14 form-f are worse than those in the L14 paper -- I'd better double check it and make sure I use the same settings and version for all simulations consistently for a good comparison.}between original Lipid14, new model and experiments,
 
-Accurate binding of \ce{Na^+} and \ce{Ca^{2+}} cations.
+Ion binding in lipid bilayers can be accurately measured and compared
+between experiments and simulations by using lipid headgroup order
+parameters and ''electrometer concept'' introduced by Seelig et al. \cite{seelig87,catte16}.
+This is based on the experimental observations that order parameters
+for $\alpha$ and $\beta$ carbons are proportional to the amount of bound charge
+in lipid bilayer, which was rationalized as a carge induced tilt
+of headgroup dipole \cite{seelig87}. Later analysis including signs of the order
+parameters showed that the order parameters are actually decreasing with bound
+positive charge and {\it vice versa} for negative charge \cite{ollila16,catte16}.
 
-Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant.
+Order parameter changes as a function of bound charge are shown in
+Fig. \ref{OrderParameterCHANGESnewMODELS}. Approximately linear
+decrease of headgroup order parameters is observed in simulations and
+experiments with cationic surfactant. Quantitative comparison reveals
+that in original Lipid14 model the proportionality constant is
+overestimated, thus the headgroup is too sensitive to the bound cations.
+However, the new ECC lipid model shows better agrement, althoght
+$\beta$ responce is slightly underestimated.
 
-What affects the headgroup OP response?: The response of the headgroup order parameters is improved
+{\it What affects the headgroup OP response?: The response of the headgroup order parameters is improved
 both through diminished affinity towards the cations and through lower headgroup response to the bound charge
 (from simulations with \ce{DHMDMAB} surfactant). 
-The responses of headgroup OPs are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+The responses of headgroup OPs are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}. }
 
-Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.
-Samuli: the main selling argument would be that we solve the details (stoichiometry, coordination numbers, etc.) of cation binding in PC bilayer.
+%Accurate binding of \ce{Na^+} and \ce{Ca^{2+}} cations. 
+%Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant. 
+%- Figure showing order parameters as a function of \ce{NaCl} and \ce{CaCl_2} concentrations~
 
-- Figure showing order parameters as a function of \ce{NaCl} and \ce{CaCl_2} concentrations~
-\todoii{ongoing}{Actual concentration of cations in simulation has yet to be estimated. 
+\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
 If it varies too much from the nominal concentration, I may need to tweak the scaling factors, $f_q$ or only $f_\sigma$, to accomodate it. 
@@
 Big patches with loads of solvent are running at the moment to guide me on the possible finite-size errors and this conc-error.}
-from Lipid14, new model and experiments (like current Fig. 1, but only original and final models)
 
@@
   \caption{\label{OrderParameterCHANGESnewMODELS}
-    Headgroup order parameter changes as a function of concentration of cations or cationic surfactant DHMDMAB. }
+    Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration and
+    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$)
+    from simulations and experiments (DPPC \cite{akutsu81}, POPC \cite{altenbach84}, surfactant \cite{scherer89}).
+    Simulations with Lipid14 and Åqvist ion model from \cite{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC1000mMCaClfiles}.
+  }
+  \todo{Lipid14/Åqvist data to be added from https://github.com/NMRLipids/lipid\_ionINTERACTION/blob/master/Data/POPC/CaCl/LIPID14/LIPID14caclCONSchange.dat}
 \end{figure*}
 
-- Figure(s) analysing binding details (density profiles, lipids bound per ion, something else?).
+Ion density profiles between different simulation models are compared in Fig. \ref{fig:cacl-dens}.
+Density profiles from simulations with original Lipid14 and Åqvist ions and new ECC lipid model with
+scaled ions are not much different, thus the overestimated order parameter decrease is explained
+by the overestimated headgroup response on bound charge in original lipid model (see Fig. \ref{OrderParameterCHANGESnewMODELS}). 
+Introducing scaling in only on ion models increase the ion binding which is also seen in order parameter
+changes in Fig. \ref{OrderParameterCHANGESnewMODELS} \todo{Final check to be done when the data from Åqvist model is there}.
+The increased binding can be probably explained by the changes in the size of the ion due to scaling \todo{more detailed explanation?}.
+Using ECC approach also in lipid headgroups decreases the ion binding and also brings headgroup response
+on bound charge close to experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}. Alltogether,
+the response of headgroup order parameters as a function of CaCl$_2$ concentration is in agreement with
+experiments in new ECC lipid model simulated with scaled CaCl$_2$ model. This is a significant improvement
+on previous modela available for lipid--ion interactions studies \cite{catte16} and indicates that
+the new model can be used for detailed studies of lipid-ion interactions.
+
+\todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, etc, and
+finalize the discussion after these results.}
 

### 2017-05-24 Fig/convert_all_eps_to_jpg_pdf.sh
Improved figures, fixed alpha value, added densities

@@ -1,0 +1,5 @@
+for f in *.eps
+do 
+  convert -density 200 -alpha background $f ${f%.eps}.jpg
+  epstopdf $f ${f%.eps}-eps-converted-to.pdf
+done

### 2017-06-05 Manuscript/proto-manuscript.tex

@@
 
-why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response~\cite{catte16}.
-
-\section{Results and Discussion}
+why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response~\cite{catte16} but not perfect.
 
 Detailed and robust structural information of lipid bilayers can
-be reached with C-H bond order parameter from NMR and form factors from scattering
-experiments. The former can be used to validate structures sampled by individual
-molecules and latter to validate overall bilayer properties, lateral density
-(area per molecule) and thickness \cite{ollila16}. The structural quality of
-new ECC lipid model against scattering form factors and NMR order parameters
-is evaluated in Fig. \ref{simVSexpNOions}
-\todo{Add acyl chain order parameters, POPC chemical structure and fix experimental alpha order parameter value}
-\todo{Why original Lipid14 models seems to deviate from experiments, in contrast to the original publication?}
+be reached with C-H bond order parameter from NMR experiments,
+which tell on the lipid structure sampled by each individual molecule.
+On the other hand, form factors from scattering experiments validate overall bilayer properties, 
+lateral density (area per molecule) and thickness \cite{ollila16}. 
+The structural quality of
+new ECC lipid model is evaluated against scattering form factors and NMR order parameters
+in Fig. \ref{simVSexpNOions}
+\todo{Add acyl chain order parameters, POPC chemical structure}
+\todo{Why original Lipid14 model seems to deviate from experiments, in contrast to the original publication? Different simulation setup used?}
 \begin{figure}[]
@@
 are shown in Table \ref{apls}
-\todo{finalize figure and the discussion}
+\todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
 \begin{table}
@@
 
-%Preserves other relevant properties: APL, Order Parameters (OPs of headgroup and tails), 
-%diffusion? \todoi{diffusion can be even improved -- and doesn't have to. We may as well omit it.}, 
-%elasticity? \todoi{check elastic properties -- do they change, improve, worsen?? -- likely not crucial for the main message of the paper.} 
-%-- however, it is still garbage in-garbage out
-%The glycerol backbone and headgroup order parameters are plotted in
-%Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
-%i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
-
-%\textbf{Required figures}
-
-%- Figure showing the system: POPC, membrane, atoms involved in headgroup OPs.\todoi{This is probably sufficient as the last thing we would do -- and will probably provide a good basis for TOC graphic.}
-
-%- Figure comparing order parameters (Fig.~\ref{HGopsNEWmodels}) and form factor \todoi{The original L14 form-f are worse than those in the L14 paper -- I'd better double check it and make sure I use the same settings and version for all simulations consistently for a good comparison.}between original Lipid14, new model and experiments,
-
 Ion binding in lipid bilayers can be accurately measured and compared
@@
 for $\alpha$ and $\beta$ carbons are proportional to the amount of bound charge
-in lipid bilayer, which was rationalized as a carge induced tilt
-of headgroup dipole \cite{seelig87}. Later analysis including signs of the order
+in lipid bilayer, which was rationalized as a charge induced tilt
+of the headgroup dipole \cite{seelig87}. Later analysis including signs of the order
 parameters showed that the order parameters are actually decreasing with bound
@@
 
+\section{Results and Discussion}
+
 Order parameter changes as a function of bound charge are shown in
-Fig. \ref{OrderParameterCHANGESnewMODELS}. Approximately linear
-decrease of headgroup order parameters is observed in simulations and
-experiments with cationic surfactant. Quantitative comparison reveals
-that in original Lipid14 model the proportionality constant is
-overestimated, thus the headgroup is too sensitive to the bound cations.
-However, the new ECC lipid model shows better agrement, althoght
-$\beta$ responce is slightly underestimated.
-
-{\it What affects the headgroup OP response?: The response of the headgroup order parameters is improved
-both through diminished affinity towards the cations and through lower headgroup response to the bound charge
-(from simulations with \ce{DHMDMAB} surfactant). 
-The responses of headgroup OPs are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}. }
-
-%Accurate binding of \ce{Na^+} and \ce{Ca^{2+}} cations. 
-%Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant. 
-%- Figure showing order parameters as a function of \ce{NaCl} and \ce{CaCl_2} concentrations~
+Fig. \ref{OrderParameterCHANGESnewMODELS}. 
+Approximately linear decrease of headgroup order parameters is observed both in simulations and experiments with the cationic surfactant. 
+The use of a cationic surfactant has the benefit that it not a subject of partitioning between water and lipid phase. 
+This means that we know the exact amount of bound charge in the membrane both in experiment and simulaiton. 
+Quantitative comparison reveals that in the original Lipid14 model the response is overestimated for both order parameters, $\alpha$ and $\beta$. 
+This suggests that the observed overestimated response in \cite{catte16} 
+is at least in part due to the high sensitivity of the headgroup order parameter response, 
+not only due to overestimated binding of cations. 
+Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in experiments (VALexp vs VALsim\todoi{Add values of $S^\alpha/S^\beta$ response for sim and experiment}).
+This property is kept also by the newly derived model, ECClipid, and, hence, 
+it shows a perfect agreement in the response of the order parameter $S^\alpha$ 
+but underestimates slightly the response of $/S^\beta$.
 
@@
   }
-  \todo{Lipid14/Åqvist data to be added from https://github.com/NMRLipids/lipid\_ionINTERACTION/blob/master/Data/POPC/CaCl/LIPID14/LIPID14caclCONSchange.dat}
+  \todoii{Add Lipid14-Aquist data.}{Lipid14/Åqvist data to be added from https://github.com/NMRLipids/lipid\_ionINTERACTION/blob/master/Data/POPC/CaCl/LIPID14/LIPID14caclCONSchange.dat .
+Joe: I'm not sure what is meant -- I think Aquist data are actually plotted there.
+I recently changed it to L14+Dang ions (both data and plot).
+I think just one ion model is sufficient (and ECC-ions are based on Dang model).}
 \end{figure*}
@@
 Ion density profiles between different simulation models are compared in Fig. \ref{fig:cacl-dens}.
-Density profiles from simulations with original Lipid14 and Åqvist ions and new ECC lipid model with
-scaled ions are not much different, thus the overestimated order parameter decrease is explained
-by the overestimated headgroup response on bound charge in original lipid model (see Fig. \ref{OrderParameterCHANGESnewMODELS}). 
-Introducing scaling in only on ion models increase the ion binding which is also seen in order parameter
-changes in Fig. \ref{OrderParameterCHANGESnewMODELS} \todo{Final check to be done when the data from Åqvist model is there}.
-The increased binding can be probably explained by the changes in the size of the ion due to scaling \todo{more detailed explanation?}.
-Using ECC approach also in lipid headgroups decreases the ion binding and also brings headgroup response
-on bound charge close to experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}. Alltogether,
+Density profiles from simulations with original Lipid14 and Dang ions \cite{smith94,chang1999,dang2006} show a pronounced peak in the position of the phosphate moieties of POPC. 
+The use of a ECC-ion model~\cite{kohagen14,kohagen16} along with original Lipid14 does not significantly change it. 
+The new ECClipid model with scaled ions exhibits on the other hand smaller density in this region suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
+This demonstrates how importnat it is to account for electronic polarizability also in molecules with zero total charge. 
+Alltogether,
 the response of headgroup order parameters as a function of CaCl$_2$ concentration is in agreement with
-experiments in new ECC lipid model simulated with scaled CaCl$_2$ model. This is a significant improvement
-on previous modela available for lipid--ion interactions studies \cite{catte16} and indicates that
+experiments in the new ECClipid model simulated with ECC model of CaCl$_2$. This is a significant improvement
+over previous models available for lipid--ion interactions studies \cite{catte16} and indicates that
 the new model can be used for detailed studies of lipid-ion interactions.
-
-\todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, etc, and
-finalize the discussion after these results.}
 
@@
     Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Aquist parameters and with ECC ions and ECCLipids17 with ECC ions. }
+\end{figure}
+
+For example, there is a debate in the literature on the stoichiometry of \ce{Ca^{2+}} binding to POPC~\cite{Altenbach84} \todoii{REFs}{Add references on Ca2+:POPC stoichiometry}. 
+In line with the early experimental finding~\cite{Altenbach84}, we find that our data fit well a ternary complex model, which assumes 2 POPC molecules per 1 \ce{Ca^{2+}}. 
+In contrast to this experiment, however, we possess also a complete atomistic detail of this phenomenon.
+Hence we can support this empirical finding with a direct observation of transient complexes of \ce{Ca^{2+}} and POPC.
+The simulation suggests that the stoichiometry 2~POPC:1~\ce{Ca^{2+}} is the most common complex. 
+In addition, there also is a non-negligable probability of one \ce{Ca^{2+}}  cross-bridging three POPC molecules. 
+This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to the binding isotherm that assume 2~POPC per 1~\ce{Ca^{2+}}. 
+\todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, etc, and finalize the discussion after these results.}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=9.0cm,angle=0]{../Fig/CA_binding.eps}
+  \caption{\label{fig:cacl-bind}
+    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{Altenbach84} fits the simulation data nicely.}
 \end{figure}

### 2017-06-09 Manuscript/proto-manuscript.tex
Minor MS changes
proto-MS: added methods, ECC model development.

@@
 
-ECC and solvation $\Delta G$.(from \cite{leontyev11} \todoi{Pavel: do not replicate Stuchebrukhov paper, write it very briefly}: hydration $\Delta  G_{hyd}$ can provide good structure and energetics, but will fail in good interactions -- already proved for cations)
-In addition, $\Delta G_{hyd}$ -- the usual target for parametrization of small molecules in classical force fields -- is not the right target in the MDEC/ECC paradigm -- part of $\Delta G = \Delta G_{nuc} + \Delta G_{el}$ is already included in the polarization of electrons, and only the remining part of $\Delta G$,  $\Delta G_{nuc}$, belongs to the polarization of the nuclei.
-Hence it is non trivial, how large should be the scaling factor $f_\sigma$ -- it should lay between $1$ and $f_q$, the limits for the original interaction energy and its limit when we neglect the $\Delta G_{el}$ term in $\Delta G_{hyd}$.
+Simulations with molecular mechanics models represent nuclei as classical particles 
+and describe their interactions through a pariwise additive empirical potential \cite{SOME_review, FF_papers}. 
+According to the electronic continuum correction (ECC), electrons that surround 
+the nuclei shall be taken into account in such simulations as an electronic continuum. \cite{leontyev11}
+Such a continuum responds to the electrostatic interactions of the explicitly represented nuclei
+as a dielectric polarizing itself. 
+The dielectric constant of shuch a dielectric medium, $\epsilon _{el}$, will 
+then resduce the electrostatic interactions in between the nuclei. 
+This can be formally represented in standard simulations by scaling charges of particles with a factor $f_q = \epsilon _{el} ^{-1/2}$. 
+Note that the high frequency dielectric constant, which is experimentally measurable \cite{some_original_work, leontyev11}, represents
+the part of dielectric response only due to electrons, and it is approximately $\epsilon _{el} ~ 2$ for almost any biomaterial that is
+usually simulated/exists in a living cells \todo{choose between the two options}.
+The high frequency dielectric constant of water, 
+the most abundant molecule in any simulaiton in biophysics, 
+is $\epsilon _{el} = 1.78$, hence the scaling factor adopts the value of $f_q = 0.75$. \cite{some_orig_source, leontyev11}
 
-How do I apply ECC on Lipid14 POPC and why such choice (i.e. $f_\sigma$, $f_q$ and definition of the scaled region).
-Region definition: headgroup requires optimization, tails appear already accurate enough. 
-Choice of $f_q$: value 0.8 reflects the fact that the charges in Lipid14 were derived in vacuum, whereas they should rather be the average of vacuum and solvated charges (so called IPolQ charges \cite{Cerutti2013}, both charge-sets can be taken from \cite{maciejewski14}).
+The newly presented model is based on the POPC model from Lipid14~\cite{dickson14} force field. 
+There, the partial atomic charges were derived by 
+fitting the electrostatic potential of a model quantum chemistry representation in vacuum. 
+If such charges are obtained in an implicit solvent, they vary the most for the polar moieties \cite{maiejewski14}. 
+In order to represent this difference in polarity in a model with fixed charges, 
+we shall use the average charges from both environments (IpolQ charges) \cite{Cerutti2013}. 
+By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
+we represent the IpolQ charges in the electronic continuum correction of the Lipid14 model
+by increasing the scaling factor $f_q = 0.8$.
 
-why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response~\cite{catte16} but not perfect.
+Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}
+are found to interact only with the phospholipid head group,
+and the alkyl sidechains are found to be described accuratley enough,
+we exclude alkyl tails from our modifications.
+
+By scaling the particle charges alone, we reduce the 
+electrostatic interactions and, hence, their contribution to hydration.
+A simple approach to correct for this effect is to scale also 
+the sigma factors in the Van der Waals interaction that is 
+modelled with a Lennard-Jones potential. \cite{lennard-jones_work}
+In line with the electronic continnum correction,
+the scaling factor for sigma parameters, $f_\sigma$,
+has to lay between $f_q$ and 1.
+
+The optimal value for $f_q$ was found by matching the overall membrane structure, 
+scattering form factors to the experimental data.
+This ensures that the presented model describes the membrane strucutre correctly,
+adopts correct liquid disordered phase, 
+and provides area per lipid agreeing with experiments (see Fig~\ref{simVSexpNOions}).
 
@@
 positive charge and {\it vice versa} for negative charge \cite{ollila16,catte16}.
+To date, there is no lipid model that can describe such a behaviour, 
+the electrometer response and cation binding, quantitatively well~\cite{catte16}
 
@@
   \centering
-  \includegraphics[width=9.0cm,angle=0]{../Fig/CA_binding.eps}
+  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
   \caption{\label{fig:cacl-bind}
@@
 \textbf{Discussion:}
+
+\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
 

### 2017-06-10 Manuscript/proto-manuscript.tex
proto-MS: Added short conclusion and discussion, small polishes

@@
 Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. 
-The new lipid model, "ECClipids", shows accurate binding affinity to sodium and calcium cations and headgroup order parameter response to bound charge. 
+The new lipid model, "ECC-lipids", shows accurate binding affinity to sodium and calcium cations and headgroup order parameter response to bound charge. 
 We also provide for the first time a realistic stochiometry of bound calcium cations to a POPC membrane, and their binding sites. 
@@
 Such a continuum responds to the electrostatic interactions of the explicitly represented nuclei
-as a dielectric polarizing itself. 
+as a polarizable dielectric. 
 The dielectric constant of shuch a dielectric medium, $\epsilon _{el}$, will 
 then resduce the electrostatic interactions in between the nuclei. 
-This can be formally represented in standard simulations by scaling charges of particles with a factor $f_q = \epsilon _{el} ^{-1/2}$. 
+This can be formally represented in classical simulations through scaling of charges of particles with a factor $f_q = \epsilon _{el} ^{-1/2}$. 
 Note that the high frequency dielectric constant, which is experimentally measurable \cite{some_original_work, leontyev11}, represents
@@
 
-The newly presented model is based on the POPC model from Lipid14~\cite{dickson14} force field. 
+The newly presented ECC-POPC model is based on the POPC model from Lipid14~\cite{dickson14} force field. 
 There, the partial atomic charges were derived by 
-fitting the electrostatic potential of a model quantum chemistry representation in vacuum. 
+fitting the electrostatic potential of its model quantum chemistry representation in vacuum. 
 If such charges are obtained in an implicit solvent, they vary the most for the polar moieties \cite{maiejewski14}. 
@@
 we represent the IpolQ charges in the electronic continuum correction of the Lipid14 model
-by increasing the scaling factor $f_q = 0.8$.
+by increasing the scaling factor for charges to $f_q = 0.8$.
 
@@
 are found to interact only with the phospholipid head group,
-and the alkyl sidechains are found to be described accuratley enough,
+and the accuracy of alkyl sidechains in Lipid14 model are satisfactory, 
 we exclude alkyl tails from our modifications.
@@
 adopts correct liquid disordered phase, 
-and provides area per lipid agreeing with experiments (see Fig~\ref{simVSexpNOions}).
+and provides area per lipid agreeing with experiments (see Fig~\ref{simVSexpNOions} and Table~\ref{tab:apls}).
 
@@
 Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
-are shown in Table \ref{apls}
+are shown in Table \ref{tab:apls}.
 \todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
 \begin{table}
-  \caption{Area per lipid from different models for POPC without ions\label{apls} }
+  \caption{Area per lipid from different models for POPC without ions\label{tab:apls} }
   \begin{tabular}{c c c}
@@
 
-Order parameter changes as a function of bound charge are shown in
-Fig. \ref{OrderParameterCHANGESnewMODELS}. 
-Approximately linear decrease of headgroup order parameters is observed both in simulations and experiments with the cationic surfactant. 
-The use of a cationic surfactant has the benefit that it not a subject of partitioning between water and lipid phase. 
-This means that we know the exact amount of bound charge in the membrane both in experiment and simulaiton. 
+Order parameter changes from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
+as a function of bound charge are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}. 
+Approximately linear decrease of headgroup order parameters is observed both 
+in simulations and experiments with the cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$). 
+The use of a cationic surfactant has the benefit that it is not a subject of partitioning between water and lipid phase. 
+This means that we know the exact amount of bound charge in the membrane both in experiment and simulation. 
 Quantitative comparison reveals that in the original Lipid14 model the response is overestimated for both order parameters, $\alpha$ and $\beta$. 
@@
 Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in experiments (VALexp vs VALsim\todoi{Add values of $S^\alpha/S^\beta$ response for sim and experiment}).
-This property is kept also by the newly derived model, ECClipid, and, hence, 
-it shows a perfect agreement in the response of the order parameter $S^\alpha$ 
-but underestimates slightly the response of $/S^\beta$.
+This property is kept also by the newly derived model, ECC-lipid, and, hence, 
+it shows a perfect agreement in the response of the headgroup order parameter $S^\alpha$ 
+but underestimates slightly the response of $S^\beta$.
 
@@
 Ion density profiles between different simulation models are compared in Fig. \ref{fig:cacl-dens}.
-Density profiles from simulations with original Lipid14 and Dang ions \cite{smith94,chang1999,dang2006} show a pronounced peak in the position of the phosphate moieties of POPC. 
+Density profiles from simulations with original Lipid14 and Dang ions \cite{smith94,chang1999,dang2006} 
+show a pronounced peak in the position of the phosphate moieties of POPC. 
 The use of a ECC-ion model~\cite{kohagen14,kohagen16} along with original Lipid14 does not significantly change it. 
-The new ECClipid model with scaled ions exhibits on the other hand smaller density in this region suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
+The new ECC-lipid model with scaled ions exhibits on the other hand smaller density in this region 
+suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
 This demonstrates how importnat it is to account for electronic polarizability also in molecules with zero total charge. 
@@
 the response of headgroup order parameters as a function of CaCl$_2$ concentration is in agreement with
-experiments in the new ECClipid model simulated with ECC model of CaCl$_2$. This is a significant improvement
+experiments in the new ECC-lipid model simulated with ECC model of CaCl$_2$. 
+This is a significant improvement
 over previous models available for lipid--ion interactions studies \cite{catte16} and indicates that
@@
 
-For example, there is a debate in the literature on the stoichiometry of \ce{Ca^{2+}} binding to POPC~\cite{Altenbach84} \todoii{REFs}{Add references on Ca2+:POPC stoichiometry}. 
-In line with the early experimental finding~\cite{Altenbach84}, we find that our data fit well a ternary complex model, which assumes 2 POPC molecules per 1 \ce{Ca^{2+}}. 
+One possible example is to adress the discourse on the stoichiometry of \ce{Ca^{2+}} binding to POPC~\cite{Altenbach84} \todoii{REFs}{Add references on Ca2+:POPC stoichiometry}. 
+In line with the early experimental finding~\cite{Altenbach84}, we find that our data 
+fit well a ternary complex model, which assumes 2 POPC molecules per 1 \ce{Ca^{2+}}. 
 In contrast to this experiment, however, we possess also a complete atomistic detail of this phenomenon.
@@
 In addition, there also is a non-negligable probability of one \ce{Ca^{2+}} cross-bridging three POPC molecules. 
-This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to the binding isotherm that assume 2~POPC per 1~\ce{Ca^{2+}}. 
+This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to the binding isotherm that assumes 2~POPC per 1~\ce{Ca^{2+}}. 
 \todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, etc, and finalize the discussion after these results.}
+
+It is also suggested that the addition of \ce{NaCl} to the solution of \ce{CaCl_2} enhances the hedgroup order parameter response compared to the solution with only \ce{CaCl_2}. \cite{Altenbach84}
+\todo{Simulate this effect and discuss it further}
+
+\todo{The difference between DPPC and POPC -- simulate and compare with experiment. }
 
@@
 In addition, there is protein force field Amber15-FB, which uses water close to OPC3, TIP3pFB \todoii{REFs}{add references here}.
+On the other hand, it might be the best if we used a model that doesn't have the dielectric constant from nuclei 78, but rather 44 -- TIP4p2005 is the closest. \cite{OPC_paper, ForceBalance_paper}
 
@@
 
-\textbf{Conclusion:}
+\section{Conclusion}
 
-Reiterate what we did...
+%Reiterate what we did...
+We present models of POPC and DPPC lipids that for the first time exhibit accurate headgroup order parameter response to cation binding. 
+The models are derived from Lipid14 model \cite{dickson14} by applying the electronic continuum correction. 
+The models were optimized to represent correct membrane structure, 
+and they were validated with the use of electrometer concept \cite{seelig87,Altenbach84}. 
+Compared to the phenomenological observations of calcium:POPC stoichiometry in the experimental work \cite{Altenbach87}, 
+our simulations reveal the same stoichiometry of 2~POPC molecules per 1~calcium but through direct observation. 
+This tells us on the subtle effects of calcium in phospholipid membranes 
+possibly leading to a better understanding of their physical properties (elasticity etc.). 
+\todo{Improve this concluding-discussion with actual insights.}
 

### 2017-06-10 Manuscript/proto-manuscript.tex
fix: added bracket

@@
 
-Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}
+Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
 are found to interact only with the phospholipid head group,

### 2017-06-11 Manuscript/proto-manuscript.tex

@@
 Note that the high frequency dielectric constant, which is experimentally measurable \cite{some_original_work, leontyev11}, represents
-the part of dielectric response only due to electrons, and it is approximately $\epsilon _{el} ~ 2$ for almost any biomaterial that is
-usually simulated/exists in a living cells \todo{choose between the two options}.
+the part of dielectric response only due to electrons, and it is approximately $\epsilon _{el} \approx 2$ for almost any biomaterial that exists in living cells. 
 The high frequency dielectric constant of water, 
-the most abundant molecule in any simulaiton in biophysics, 
+the most abundant molecule in any cell, 
 is $\epsilon _{el} = 1.78$, hence the scaling factor adopts the value of $f_q = 0.75$. \cite{some_orig_source, leontyev11}
@@
 
-The optimal value for $f_q$ was found by matching the overall membrane structure, 
-scattering form factors to the experimental data.
-This ensures that the presented model describes the membrane strucutre correctly,
-adopts correct liquid disordered phase, 
+The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
+by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
+This ensures that the presented model describes correctly the membrane strucutre,
+adopts a liquid disordered phase, 
 and provides area per lipid agreeing with experiments (see Fig~\ref{simVSexpNOions} and Table~\ref{tab:apls}).
@@
 The models were optimized to represent correct membrane structure, 
-and they were validated with the use of electrometer concept \cite{seelig87,Altenbach84}. 
-Compared to the phenomenological observations of calcium:POPC stoichiometry in the experimental work \cite{Altenbach87}, 
+and they were validated with the use of electrometer concept \cite{seelig87,Altenbach85,Altenbach84}. 
+Compared to the phenomenological observations of calcium:POPC stoichiometry in the experimental work \cite{Altenbach84}, 
 our simulations reveal the same stoichiometry of 2~POPC molecules per 1~calcium but through direct observation. 

### 2017-06-11 Manuscript/proto-manuscript.tex
Updates

@@
 
+To date, there is no lipid model that can describe such a behaviour, 
+the electrometer response and cation binding, quantitatively well~\cite{catte16}
+
+Almost all lipid bilayer simulations reported in the literature 
+are done with force fields that represent interactions between atoms
+with pairwise additive empirical potential and exclude the electronic 
+polarizability \cite{SOME_review, FF_papers}. While the issue has been 
+considered higly relevant since the early days of lipid bilayer simulations,
+practical solutions have turned out to be complicated \cite{??}.
+Here we propose that electronic polarizability can be effectively included
+in lipid bilayer simulations by using the electronic contimuun correction \cite{leontyev11}.
+
 \section{Methods}
 
-Simulations with molecular mechanics models represent nuclei as classical particles 
-and describe their interactions through a pariwise additive empirical potential \cite{SOME_review, FF_papers}. 
-According to the electronic continuum correction (ECC), electrons that surround 
-the nuclei shall be taken into account in such simulations as an electronic continuum. \cite{leontyev11}
-Such a continuum responds to the electrostatic interactions of the explicitly represented nuclei
-as a polarizable dielectric. 
-The dielectric constant of shuch a dielectric medium, $\epsilon _{el}$, will 
-then resduce the electrostatic interactions in between the nuclei. 
-This can be formally represented in classical simulations through scaling of charges of particles with a factor $f_q = \epsilon _{el} ^{-1/2}$. 
-Note that the high frequency dielectric constant, which is experimentally measurable \cite{some_original_work, leontyev11}, represents
-the part of dielectric response only due to electrons, and it is approximately $\epsilon _{el} \approx 2$ for almost any biomaterial that exists in living cells. 
-The high frequency dielectric constant of water, 
-the most abundant molecule in any cell, 
-is $\epsilon _{el} = 1.78$, hence the scaling factor adopts the value of $f_q = 0.75$. \cite{some_orig_source, leontyev11}
+\subsection{Electronic continuum correction for lipid bilayers}
+According to the electronic continuum correction (ECC) the electronic
+polarizability can be effectively included in MD simulations by scaling
+the partial charges with a constant factor $f_q = \epsilon _{el} ^{-1/2}$, 
+where $\epsilon _{el}$ is the electronic part of the dielectric constant of 
+media \cite{leontyev11}. Measurements of high frequency 
+dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
+for almost any biomaterial \cite{some_original_work, leontyev11}. % that exists in living cells. 
+The value measured for water, $\epsilon _{el} = 1.78$, gives 
+a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}, which has been
+successfully used to improve the performance of ion force fields \cite{kohagen14,kohagen16,??}.
 
-The newly presented ECC-POPC model is based on the POPC model from Lipid14~\cite{dickson14} force field. 
-There, the partial atomic charges were derived by 
-fitting the electrostatic potential of its model quantum chemistry representation in vacuum. 
+Here we apply the same approach to lipid bilayers to accurately 
+describe the lipid headgroup response to Na+ and Ca2+ concentrations \cite{catte16}. 
+We use the Lipid14~\cite{dickson14} force field parameters as a starting point,
+because they give the most realistic headgroup response with added cations 
+and relatively realistic glycerol backbone and headgroup structures
+when compared with other state of the art lipid models \cite{botan15,catte16}.
+The partial atomic charges in Lipid14 were derived by fitting the electrostatic 
+potential to its model quantum chemistry representation in vacuum.
+\todo{I do not fully understand the end of the paragraph} 
 If such charges are obtained in an implicit solvent, they vary the most for the polar moieties \cite{maiejewski14}. 
@@
 by increasing the scaling factor for charges to $f_q = 0.8$. 
+The newly presented ECC-POPC model is based on the POPC model from
 
-Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
-are found to interact only with the phospholipid head group,
-and the accuracy of alkyl sidechains in Lipid14 model are satisfactory, 
-we exclude alkyl tails from our modifications.
+Hydrocarbon chains in Lipid14 and other lipids models are highly optimized and
+give generally a good description for hydrophobic part of lipid bilayers
+in various conditions \cite{ollila16},
+in contrast to glycerol backbone and headgroup regions which require
+some improvement in all available lipid models \cite{botan15}. 
+To minimize the intercalation with the highly optimized hydrocarbon
+parameters, we decided to apply ECC correction only
+to the headgroup, glycerol backbone and carbonyl in acyl chains.
+These regions are also most polar parts in lipids and are expected
+to most affect on cation binding.   
+%Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
+%are found to interact only with the phospholipid head group,
+%and the accuracy of alkyl sidechains in Lipid14 model are satisfactory, 
+%we exclude alkyl tails from our modifications.
 
-By scaling the particle charges alone, we reduce the 
-electrostatic interactions and, hence, their contribution to hydration.
-A simple approach to correct for this effect is to scale also 
-the sigma factors in the Van der Waals interaction that is 
-modelled with a Lennard-Jones potential. \cite{lennard-jones_work}
+The scaling of these partial charges by a factor of 0.8 reduce the 
+area per molecule to ??, which is significantly smaller than the
+experimental and original Lipid 14 values.  
+The decrease of area was speculated 
+\todo{Did you analyze the effect of this to hydration or did we only speculate?}
+to arise from reduced hydration
+of headgroup due to the lower polarity of molecules with scaled charges.
+The hydration can be increased by decreasing the radius of atoms
+by reducing $\sigma$ term in Lennart-Jones potential.
+This allows water molecules approach closer to lipid atoms,
+where electrostatic interactions are stronger and, thus, 
+increases hydration level.
+\todo{This effect may have an official name. In that case we should mention it.}
+\todo{Is there some justification for the below sentence?} 
 In line with the electronic continnum correction,
 the scaling factor for sigma parameters, $f_\sigma$,
-has to lay between $f_q$ and 1.
+has to lay between $f_q$ and 1. Scaling factor
+value of ?? for $\sigma$ was found to reproduce
+area per molecule values close to experiments and
+original Lipid14 model when charges were scaled with
+0.8.
 
+\subsection{Comparison to experimental data}
+Structures sampled by individual lipid molecules were compared
+to experimental data by using C-H bond order parameters \cite{ollila16}
+\begin{equation}
+S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
+\end{equation}
+where $\theta$ is the angle between C-H bond and membrane
+normal and average is taken over all sampled configurations.
+Bilayer dimensions were compared to experiments by using the
+scattering form factor \cite{ollila16}
+\todo{Add equation}.
+
+Ion binding was compared between experiments and simulations by 
+using lipid headgroup order parameters and ''electrometer concept'' 
+introduced by Seelig et al. \cite{seelig87,catte16}.
+The concept is based on the experimental observation that the 
+order parameters of $\alpha$ and $\beta$ carbons in lipid headgroup
+(see Fig. \ref{??}) are proportional to the amount of bound charge
+in lipid bilayer \cite{seelig87}. More recent analysis included also
+the order parameter signs and concluded that the order parameters  
+decrease with bound positive charge and increase with bound negative 
+charge \cite{ollila16,catte16}. The observations are rationalized 
+as a change of lipid headgroup dipole tilt to more vertical orientation
+with bound positive charge and {\it vice versa} for negative \cite{seelig87}.
+
+\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
+If it varies too much from the nominal concentration, I may need to tweak the scaling 
+factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
+to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
+moment to guide me on the possible finite-size errors and this conc-error.
+SAMULI: I thought about this a little bit. My current conclusion is that
+we can take the number concentration from simulations with large box having 
+a clear bulk concentration region and calculate the bulk concentration in mol/l
+unit like this}
+\begin{equation}
+  C_{eq}[mol/l]=\frac{C_{plateau}[mM]}{0.602}
+\end{equation}
+
+\subsection{Simulation details}
+
+\todo{To be written}
+
+\section{Results and Discussion}
+
+\subsection{Lipid bilayers without ions}
+The structural quality of
+new ECC lipid model is evaluated against scattering form factors and NMR order parameters
+in Fig. \ref{simVSexpNOions}
 The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 
-Detailed and robust structural information of lipid bilayers can
-be reached with C-H bond order parameter from NMR experiments,
-which tell on the lipid structure sampled by each individual molecule.
-On the other hand, form factors from scattering experiments validate overall bilayer properties, 
-lateral density (area per molecule) and thickness \cite{ollila16}. 
-The structural quality of
-new ECC lipid model is evaluated against scattering form factors and NMR order parameters
-in Fig. \ref{simVSexpNOions}
+Order parameter changes from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
+as a function of bound charge are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}. 
+Approximately linear decrease of headgroup order parameters is observed both 
+in simulations and experiments with the cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$). 
+The use of a cationic surfactant has the benefit that it is not a subject of partitioning between water and lipid phase. 
+This means that we know the exact amount of bound charge in the membrane both in experiment and simulation. 
+Quantitative comparison reveals that in the original Lipid14 model the response is overestimated for both order parameters, $\alpha$ and $\beta$. 
+This suggests that the observed overestimated response in \cite{catte16} 
+is at least in part due to the high sensitivity of the headgroup order parameter response, 
+not only due to overestimated binding of cations. 
+Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in 
+experiments (VALexp vs VALsim\todoi{Add values of $S^\alpha/S^\beta$ response for sim and experiment}).
+This property is kept also by the newly derived model, ECC-lipid, and, hence, 
+it shows a perfect agreement in the response of the headgroup order parameter $S^\alpha$ 
+but underestimates slightly the response of $S^\beta$. 
 \todo{Add acyl chain order parameters, POPC chemical structure}
@@
 
-Ion binding in lipid bilayers can be accurately measured and compared
-between experiments and simulations by using lipid headgroup order
-parameters and ''electrometer concept'' introduced by Seelig et al. \cite{seelig87,catte16}.
-This is based on the experimental observations that order parameters
-for $\alpha$ and $\beta$ carbons are proportional to the amount of bound charge
-in lipid bilayer, which was rationalized as a charge induced tilt
-of the headgroup dipole \cite{seelig87}. Later analysis including signs of the order
-parameters showed that the order parameters are actually decreasing with bound
-positive charge and {\it vice versa} for negative charge \cite{ollila16,catte16}.
-To date, there is no lipid model that can describe such a behaviour, 
-the electrometer response and cation binding, quantitatively well~\cite{catte16}
-
-\section{Results and Discussion}
-
-Order parameter changes from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
-as a function of bound charge are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}. 
-Approximately linear decrease of headgroup order parameters is observed both 
-in simulations and experiments with the cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$). 
-The use of a cationic surfactant has the benefit that it is not a subject of partitioning between water and lipid phase. 
-This means that we know the exact amount of bound charge in the membrane both in experiment and simulation. 
-Quantitative comparison reveals that in the original Lipid14 model the response is overestimated for both order parameters, $\alpha$ and $\beta$. 
-This suggests that the observed overestimated response in \cite{catte16} 
-is at least in part due to the high sensitivity of the headgroup order parameter response, 
-not only due to overestimated binding of cations. 
-Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in experiments (VALexp vs VALsim\todoi{Add values of $S^\alpha/S^\beta$ response for sim and experiment}).
-This property is kept also by the newly derived model, ECC-lipid, and, hence, 
-it shows a perfect agreement in the response of the headgroup order parameter $S^\alpha$ 
-but underestimates slightly the response of $S^\beta$.
-
-\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
-If it varies too much from the nominal concentration, I may need to tweak the scaling factors, $f_q$ or only $f_\sigma$, to accomodate it. 
-However, it is very unlikely, response to the surfactant \ce{DHMDMAB} is OK. 
-Big patches with loads of solvent are running at the moment to guide me on the possible finite-size errors and this conc-error.}
+\subsection{Cation binding in POPC bilayer}
 

### 2017-06-12 Manuscript/proto-manuscript.tex
Updates

@@
 The hydration can be increased by decreasing the radius of atoms
-by reducing $\sigma$ term in Lennart-Jones potential.
-This allows water molecules approach closer to lipid atoms,
-where electrostatic interactions are stronger and, thus, 
-increases hydration level.
+by reducing $\sigma$ term in Lennart-Jones potential, which
+allows water molecules approach closer to lipid atoms and
+have stonger electrostatic interactions with them.
 \todo{This effect may have an official name. In that case we should mention it.}
-\todo{Is there some justification for the below sentence?} 
-In line with the electronic continnum correction,
-the scaling factor for sigma parameters, $f_\sigma$,
-has to lay between $f_q$ and 1. Scaling factor
-value of ?? for $\sigma$ was found to reproduce
-area per molecule values close to experiments and
-original Lipid14 model when charges were scaled with
-0.8.
+Indeed, the scaling of $\sigma$ increased the area per molecule
+and the results close to experimental and original Lipid14 values were achieved
+with scaling factor of ?? for $\sigma$ and 0.8 for partial charges.
+%\todo{Is there some justification for the below sentence?} 
+%In line with the electronic continnum correction,
+%the scaling factor for sigma parameters, $f_\sigma$,
+%has to lay between $f_q$ and 1.
 
@@
 \subsection{Lipid bilayers without ions}
-The structural quality of
-new ECC lipid model is evaluated against scattering form factors and NMR order parameters
-in Fig. \ref{simVSexpNOions}
-The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
-by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
-This ensures that the presented model describes correctly the membrane strucutre,
-adopts a liquid disordered phase, 
-and provides area per lipid agreeing with experiments (see Fig~\ref{simVSexpNOions} and Table~\ref{tab:apls}).
+The scattering form factors, NMR order parameters and area per lipids calculated from 
+the ECC corrected lipid model for POPC are compared to experiments and original
+Lipid14 results in Fig. \ref{simVSexpNOions} and in Table~\ref{tab:apls}. 
+%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
+%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
+The structural quality is comparable to the state of art lipid models available in literature \cite{ollila16},
+thus we conclude that the ECC corrected lipid model reproduces the lipid bilayer structure
+in liquid disordered phase with similar accuracy than other available models
+\todo{Discussion to be finished when we have all the results in the figure}.
 
-Order parameter changes from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
-as a function of bound charge are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}. 
-Approximately linear decrease of headgroup order parameters is observed both 
-in simulations and experiments with the cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$). 
-The use of a cationic surfactant has the benefit that it is not a subject of partitioning between water and lipid phase. 
-This means that we know the exact amount of bound charge in the membrane both in experiment and simulation. 
-Quantitative comparison reveals that in the original Lipid14 model the response is overestimated for both order parameters, $\alpha$ and $\beta$. 
-This suggests that the observed overestimated response in \cite{catte16} 
-is at least in part due to the high sensitivity of the headgroup order parameter response, 
-not only due to overestimated binding of cations. 
+Headgroup order parameter response to bound charge was evaluated against
+experimental data measured with cationic surfactant
+(dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) \cite{scherer89}.
+The exact amount of bound charge in the membrane is known these systems, because
+practically all cationic surfactant molecules are embedded in lipid bilayer due to
+their amphiphilic nature. Thus, the headgroup order parameter changes as a function of
+mole fraction of cationic surfactants in  Fig. \ref{OrderParameterCHANGESnewMODELS} gives
+also the order parameter changes as a function of bound cations.
+\todo{I think that we should make a separate figure for this one and ion concentrations.
+  In the current figure it would be difficult to show this with a reasonable scale.
+  Currently Lipid14 results are cut out, thus the scale should be increased.
+  This would, however, make the changes with CaCl too small.
+  The change of headgroup tilt could be incorporated in the same figure with order parameter
+  response to cationic surfactant.
+}
+The headgroup order parameter response to bound cation concentration
+is approximately linear up to $\sim$0.3 mole fraction in experiments \cite{scherer89}.
+The linearity is also observed in simulations with original Lipid14 and with ECC correction.
+%from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
+%as a function of bound charge are shown in
+Quantitative comparison, however, reveals that the response is overestimated in 
+original Lipid14 for both segments, while the ECC corrected model gives a
+good agreement for the change of $\alpha$ segment order parameter, but slightly
+underestimates the $\beta$ segment order parameter change.
+The overestimation of order parameter changes with original Lipid14 model 
+suggests that the overestimated response with CaCl$_2$ concentration in \cite{catte16} 
+can be partly explained by the high sensitivity of the headgroup order parameter response
+to bound charge.
+
 Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in 
-experiments (VALexp vs VALsim\todoi{Add values of $S^\alpha/S^\beta$ response for sim and experiment}).
+experiments (VALexp vs VALsim
+\todo{Add values of $S^\alpha/S^\beta$ response for sim and experiment
+  SAMULI: Maybe these could be put in the same figure with the order parameter
+  response and P-N vector angle change with surfactants.
+}
+).
 This property is kept also by the newly derived model, ECC-lipid, and, hence, 
@@
 but underestimates slightly the response of $S^\beta$. 
-\todo{Add acyl chain order parameters, POPC chemical structure}
-\todo{Why original Lipid14 model seems to deviate from experiments, in contrast to the original publication? Different simulation setup used?}
 \begin{figure}[]
@@
     compared with experimental order parameters from \cite{ferreira13}.}
+  \todo{Add acyl chain order parameters, POPC chemical structure}
+  \todo{Should we add the results from original lipid14?}
 \end{figure}
@@
 \todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, etc, and finalize the discussion after these results.}
+
+not only due to overestimated binding of cations.
 

### 2017-06-13 Manuscript/proto-manuscript.tex
New intro and other updates

@@
   % insert abstract here
+  \todo{Abstract directly from Joe's conference abstracts. To be rewritten.}
 Classical molecular dynamics simulations give detailed information about membrane structure and dynamics. 
-However, there is still a room for improvements in current force fields – it is known from the literature, that the binding of ions, especially cations, to phopholipid membranes is overestimated in all classical models [1].
-We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent ions yielding models that feature correct ion pairing [3].
-Using the electrometer concept [3, 4] and x-ray scattering form factors, our simulations point out that our hypothesis is correct and ECC is indeed a missing important contribution in current classical lipid models.
+However, there is still a room for improvements in current force fields – it is known from the literature,
+that the binding of ions, especially cations, to phopholipid membranes is overestimated in all classical models [1]. 
+We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the
+lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent
+ions yielding models that feature correct ion pairing [3]. 
+Using the electrometer concept [3, 4] and x-ray scattering form factors, our simulations point out that our hypothesis is
+correct and ECC is indeed a missing important contribution in current classical lipid models. 
 Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. 
@@
 This work will continue as an open collaboration project NMRlipids IV (http://nmrlipids.blogspot.fi).
-
-[1] Catte, A., Girych, M., Javanainen, M., Melcr J., Miettinen, M. S., Oganesyan, V. S. and Ollila H. S.,  PCCP 18(47) 32560-32569 (2016)
-[2] Leontyev, I. V., and Stuchebrukhov, A. A., JCTC 6(5) 1498–1508 (2010)
-[3] Kohagen, M., Mason, P. E., and Jungwirth, P., J. Phys. Chem. B 120(8) 1454–60 (2015)
-[4] Seelig, J., MacDonald, P. M., and Scherer, P. G., Biochemistry 26(24) 7535–7541 (1987)
 \end{abstract}
@@
 \section{Introduction}
+Cation interactions with cellular membranes play a key role in several biological processes,
+like in signal propagation in neurons and vesicle fusion.
+Cation interactions with lipid bilayers serving as a model for cellular
+membranes are thus widely studied with experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
+and theoretical methods \cite{Berkowitz12,??}. General conclusion from experimental
+studies has been that multivalent ions and lithium have weak specific binding in
+phopholipid bilayers, while other monovalent ions do not essentially
+bind \cite{cevc90,tocanne90,seelig90,binder02}. The presence of anoinic lipids, like PS or PG,
+increase the concentration close to the bilayer and thus the amount of bound
+ions, but do not affect the specific binding constant \cite{seelig90}.
+The binding details, like binding sites and stoichiometry are not yet fully
+resolved but interpretation of NMR and scattering experiments suggest that one
+Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
+phospholipid molecules \cite{altenbach84}.
 
- motivation, significance of membranes, phospholipids and simulation.
+Classical molecular dynamics simulations have potential to solve the ion binding details
+in lipid bilayers and reveal its relevance to various biological
+problems \cite{bockmann03,bockmann04,??}. However, the available
+molecular dynamics simulation models have a strong tendency to overestimate
+cation binding on zwitterionic bilayers and none of them can reproduce the
+experimental data with the accuracy required for the interpretation of
+experiments \cite{catte16}. The overestimated specific cation binding also
+makes zwitterionic bilayers effectively positively charged, which could
+potentially lead to significant artefacts in applications.
+Thus, there is a vast demand for classical molecular dynamics simulation
+model which correctly reproduces cation binding in lipid bilayers.
 
- assumptions (so that we have structure of the paper like in a mathematical proof) -- 
- MD simulation is a good tool for studying molecules.
- classical MD models can describe lipids accurately,
+The lack of electronic polarizability from the classical MD simulation
+models is a potential source of artefact, which could lead to overbinding
+of cations. 
+%Almost all lipid bilayer simulations reported in the literature 
+%are done with force fields that represent interactions between atoms
+%with pairwise additive empirical potential and exclude the electronic 
+%polarizability \cite{SOME_review, FF_papers}.
+The issue has been considered higly relevant since the early days of
+lipid bilayer simulations and pioneering simulation studies scaled
+the partial charges to effectively include electronic
+polarizability \cite{jonsson86,egberts94}. The explicit inclusion of
+electronic polarizability has been, however, turned to be 
+practically complicated and implementations for lipids are rare \cite{chowdhary13}.
 
- MD is ... and it serves ... it is useful for ... (describe throuhg references)
-
- Lipid membranes, especially phospholipid membranes; their significance for life, sciences, society, pharma ...
-
-%\textbf{Specific introduction: current FFs and MDEC/ECC}
-
- Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties.  -- write basic ideas from the lipid-FF whitepaper I did recently.
-
- lipid force fields fail in description of membrane-cation interaction -- could be answered by ECC? 
-Cations were shown to generally overbind in PC lipid bilayers in NMRlipids II project. 
-Here we propose that the cation overbinding can be corrected by implicitly inducing electronic polarizability in lipid headgroups by scaling the partial charges -- i.e.
-MDEC/ECC~\cite{leontyev11}\todo{Bulid BibTex references database.} paradigm.
-
-MDEC -- or -- ECC 
-\todo{choose one}{We have to decide whether we will refer to the method as a correction (ECC) or as a MD simulation paradigm (MDEC) -- and choose on of these labels.
--- Joe: use ECC as we apply MDEC rather as a correction to current state not as a new paradigm for FF development.}
-MD in electronic continuum as in \cite{leontyev11} or electronic continuum correction as in \cite{Jungwirth2015}
-
-Describe ECC: good physical concept for treating part of the polarizability (electronic) in a simple mean-field way.
-
-Succesful application: Works for cations \cite{Jungwirth2015,kohagen14,kohagen16}\todoii{REF}{Missing references} -- motivation for its application to zwitterionic lipids like POPC.
-
-Hypothesis: ECC helps in describing even zwitterionic molecules like POPC, will be demonstrated through headgroup order parameter response to cationic molecules.
-
-To date, there is no lipid model that can describe such a behaviour, 
-the electrometer response and cation binding, quantitatively well~\cite{catte16}
-
-Almost all lipid bilayer simulations reported in the literature 
-are done with force fields that represent interactions between atoms
-with pairwise additive empirical potential and exclude the electronic 
-polarizability \cite{SOME_review, FF_papers}. While the issue has been 
-considered higly relevant since the early days of lipid bilayer simulations,
-practical solutions have turned out to be complicated \cite{??}.
-Here we propose that electronic polarizability can be effectively included
-in lipid bilayer simulations by using the electronic contimuun correction \cite{leontyev11}.
+In this work we show that the cation overbinding in classical MD simulations
+can be corrected by including electronic polarizability by using effective continuum
+correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules.
+This is essentially physicalle well justified version of partial charge scaling
+implemented in early days of lipid and surfctant simulations \cite{jonsson86,egberts94}. 
+The approach has been previously shown in to improve bulk performace of
+ion models against neutron scattering data \cite{Jungwirth2015,kohagen14,kohagen16}.
+The better bulk behaviour was not, however, sufficient to correct binding in lipid
+bilayers \cite{catte16}. 
+%motivation for its application to zwitterionic lipids like POPC.
 
@@
 \todo{Discussion to be finished when we have all the results in the figure}.
+\begin{figure}[]
+  \centering
+  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+%  \caption{\label{fig:form-f}
+%    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
+%\end{figure}
+%
+%\begin{figure}[]
+%  \centering
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
+  \caption{\label{simVSexpNOions}
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models. 
+    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and EECLipid17 models
+    compared with experimental order parameters from \cite{ferreira13}.}
+  \todo{Add acyl chain order parameters, POPC chemical structure}
+  \todo{Should we add the results from original lipid14?}
+\end{figure}
+%Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
+%are shown in Table \ref{tab:apls}.
+\todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
+\begin{table}
+  \caption{Area per lipid from different models for POPC without ions\label{tab:apls} }
+  \begin{tabular}{c c c}
+    model          & A (Å$^2$)   & Temperature [K] \\
+    \hline
+    Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
+    Lipid14ecc0.80+sigma0.875 &        &  313    \\
+    GMX small patch           & 64.9   &         \\
+    GMX 4xbig patch           & 65.5   &         \\
+    oMM small patch           & 63.65  &         \\
+    oMM 4xbig patch           & 63.7   &         \\
+    \hline
+    experiment \cite{Jambeck2012}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
+    experiment  & 64.3  &  303    \\
+    experiment  & 67.3  &  323    \\
+    experiment  & 68.1  &  333    \\
+    experiment POPE  & 56.6 &  303    \\
+    \hline
+  \end{tabular}
+\end{table}
 
@@
 also the order parameter changes as a function of bound cations.
-\todo{I think that we should make a separate figure for this one and ion concentrations.
+\todo{I think that we should make a separate figures for this one and ion concentrations.
   In the current figure it would be difficult to show this with a reasonable scale.
@@
 but underestimates slightly the response of $S^\beta$.
-\begin{figure}[]
-  \centering
-  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-%  \caption{\label{fig:form-f}
-%    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
-%\end{figure}
-%
-%\begin{figure}[]
-%  \centering
-  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
-  \caption{\label{simVSexpNOions}
-    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models. 
-    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and EECLipid17 models
-    compared with experimental order parameters from \cite{ferreira13}.}
-  \todo{Add acyl chain order parameters, POPC chemical structure}
-  \todo{Should we add the results from original lipid14?}
-\end{figure}
-Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
-are shown in Table \ref{tab:apls}.
-\todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
-\begin{table}
-  \caption{Area per lipid from different models for POPC without ions\label{tab:apls} }
-  \begin{tabular}{c c c}
-    model          & A (Å$^2$)   & Temperature [K] \\
-    \hline
-    Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
-    Lipid14ecc0.80+sigma0.875 &        &  313    \\
-    GMX small patch           & 64.9   &         \\
-    GMX 4xbig patch           & 65.5   &         \\
-    oMM small patch           & 63.65  &         \\
-    oMM 4xbig patch           & 63.7   &         \\
-    \hline
-    experiment \cite{Jambeck2012}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
-    experiment  & 64.3  &  303    \\
-    experiment  & 67.3  &  323    \\
-    experiment  & 68.1  &  333    \\
-    experiment POPE  & 56.6 &  303    \\
-    \hline
-  \end{tabular}
-\end{table}
 
 \subsection{Cation binding in POPC bilayer}
+
+Headgroup order parameter response to increasing CaCl$_2$ concentration
+from experiments, original Lipid14 model and ECC corrected model are shown
+in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
+is significantly overestimated in original Lipid14 model, while results from
+ECC corrected model are in good agreement with experiments.
+This is a significant improvement over previously available models,
+which always overstimate the order parameter response to CaCl$_2$
+concentration \cite{catte16}. The good agreement with experiments indicate
+that the binding details of Ca2+ are realistic in the ECC corrected model
+and it can be thus used to study lipid-ion interaction details.
 
@@
   }
-  \todoii{Add Lipid14-Aquist data.}{Lipid14/Åqvist data to be added from https://github.com/NMRLipids/lipid\_ionINTERACTION/blob/master/Data/POPC/CaCl/LIPID14/LIPID14caclCONSchange.dat .
-Joe: I'm not sure what is meant -- I think Aquist data are actually plotted there.
-I recently changed it to L14+Dang ions (both data and plot).
-I think just one ion model is sufficient (and ECC-ions are based on Dang model).}
+  \todo{Add Lipid14-Aquist data.
+    Lipid14/Åqvist data to be added from https://github.com/NMRLipids/lipid\_ionINTERACTION/blob/master/Data/POPC/CaCl/LIPID14/LIPID14caclCONSchange.dat.
+    Joe: I'm not sure what is meant -- I think Aquist data are actually plotted there. I recently changed it to L14+Dang ions (both data and plot).
+    I think just one ion model is sufficient (and ECC-ions are based on Dang model).
+    Samuli: I think that we could add also the Aqvist data. This is in the NMRlipids II publication so it might
+    make easier to follow for people who have red it publication.
+  }
 \end{figure*}
@@
 show a pronounced peak in the position of the phosphate moieties of POPC. 
-The use of a ECC-ion model~\cite{kohagen14,kohagen16} along with original Lipid14 does not significantly change it. 
+The use of a ECC-ion model~\cite{kohagen14,kohagen16} along with original Lipid14 does not significantly change it
+\todo{If we show density profile for this, we should show also the order parameter changes}.
 The new ECC-lipid model with scaled ions exhibits on the other hand smaller density in this region 
 suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
-This demonstrates how importnat it is to account for electronic polarizability also in molecules with zero total charge. 
-Alltogether,
-the response of headgroup order parameters as a function of CaCl$_2$ concentration is in agreement with
-experiments in the new ECC-lipid model simulated with ECC model of CaCl$_2$. 
-This is a significant improvement
-over previous models available for lipid--ion interactions studies \cite{catte16} and indicates that
-the new model can be used for detailed studies of lipid-ion interactions.
+This demonstrates that cation binding in zwitterionic phospholipid bilayer can be accurately
+described with classical MD simulation model with effective included of electronic polarizability.
 
@@
   \caption{\label{fig:cacl-dens}
-    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Aquist parameters and with ECC ions and ECCLipids17 with ECC ions. }
+    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Aquist parameters
+    and with ECC ions and ECCLipids17 with ECC ions. }
+  \todo{We should add the location of bilayer here somehow.
+    In NMRlipids II the location of phosphate was shown with green vertical line.}
+  \todo{I would add Aqvist data from NMRlipids II in here as well.}
 \end{figure}
@@
 \todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, etc, and finalize the discussion after these results.}
-
-not only due to overestimated binding of cations.
 

### 2017-06-13 Manuscript/proto-manuscript.tex
Updates

@@
 potential to its model quantum chemistry representation in vacuum.
-\todo{I do not fully understand the end of the paragraph} 
-If such charges are obtained in an implicit solvent, they vary the most for the polar moieties \cite{maiejewski14}. 
+\todo{SAMULI: I do not fully understand the end of this paragraph}
+If such charges are obtained in an implicit solvent, 
+they vary the most for the polar moieties \cite{maiejewski14}. 
 In order to represent this difference in polarity in a model with fixed charges, 
@@
 by increasing the scaling factor for charges to $f_q = 0.8$. 
+\todo{SAMULI: I tried to repeat this calculation but I was not able to get 0.8.
+This has to be explained better.}
 The newly presented ECC-POPC model is based on the POPC model from
@@
 order parameters of $\alpha$ and $\beta$ carbons in lipid headgroup
-(see Fig. \ref{??}) are proportional to the amount of bound charge
+(see Fig. \ref{??} 
+\todo{Figure with chemical structure and labeling to be added}) 
+are proportional to the amount of bound charge
 in lipid bilayer \cite{seelig87}. More recent analysis included also
@@
 
-\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
-If it varies too much from the nominal concentration, I may need to tweak the scaling 
-factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
-to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
-moment to guide me on the possible finite-size errors and this conc-error.
-SAMULI: I thought about this a little bit. My current conclusion is that
-we can take the number concentration from simulations with large box having 
-a clear bulk concentration region and calculate the bulk concentration in mol/l
-unit like this}
+%\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
+%If it varies too much from the nominal concentration, I may need to tweak the scaling 
+%factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
+%to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
+%moment to guide me on the possible finite-size errors and this conc-error.
+The used experimental data reports order parameters as a function of
+equilibrium cation concentration in bulk solvent \cite{akutsu81,altenbach84}.
+The box size in simulations was adjusted such that ion conceration reaches 
+a clear bulk plateau, which then gives the concentration that corresponds to the
+one reported in experiments. The concentrations in units of mol/l were then 
+determined as
 \begin{equation}
-  C_{eq}[mol/l]=\frac{C_{plateau}[mM]}{0.602}
+  C_{eq}=\frac{C_{plateau}}{0.602},
 \end{equation}
+where plateau concentration is the number density in unist of 1/(nm)$^3$.
 
@@
 
-One possible example is to adress the discourse on the stoichiometry of \ce{Ca^{2+}} binding to POPC~\cite{Altenbach84} \todoii{REFs}{Add references on Ca2+:POPC stoichiometry}. 
-In line with the early experimental finding~\cite{Altenbach84}, we find that our data 
+One possible example is to adress the discourse on the stoichiometry of \ce{Ca^{2+}} binding 
+to POPC~\cite{altenbach84}.
+In line with the early experimental finding~\cite{altenbach84}, we find that our data 
 fit well a ternary complex model, which assumes 2 POPC molecules per 1 \ce{Ca^{2+}}. 
@@
 In addition, there also is a non-negligable probability of one \ce{Ca^{2+}} cross-bridging three POPC molecules. 
-This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to the binding isotherm that assumes 2~POPC per 1~\ce{Ca^{2+}}. 
-\todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, etc, and finalize the discussion after these results.}
+This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to 
+the binding isotherm that assumes 2~POPC per 1~\ce{Ca^{2+}}. 
+\todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, 
+etc, and finalize the discussion after these results.}
 
-It is also suggested that the addition of \ce{NaCl} to the solution of \ce{CaCl_2} enhances the hedgroup order parameter response compared to the solution with only \ce{CaCl_2}. \cite{Altenbach84}
+It is also suggested that the addition of \ce{NaCl} to the solution of \ce{CaCl_2} enhances the hedgroup 
+order parameter response compared to the solution with only \ce{CaCl_2}. \cite{altenbach84}
 \todo{Simulate this effect and discuss it further}
@@
 
-\textbf{Discussion:}
+%\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
 
-\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
+%Role of water model: we use OPC3 (current best), it would be worth giving an estimate how results 
+%change when we use say SPCE or even TIP3p at least in the SI (so that the reader knows what errors 
+%to expect comming from these sub-optimal models). 
+%In addition, there is protein force field Amber15-FB, which uses water close to OPC3, TIP3pFB \todoii{REFs}{add references here}.
+%On the other hand, it might be the best if we used a model that doesn't have the dielectric constant 
+%from nuclei 78, but rather 44 -- TIP4p2005 is the closest. \cite{OPC_paper, ForceBalance_paper}
 
-other lipids: charged lipids? -- ongoing research in our lab.
-I'm currently working on POPE with Aniket for curved (and flat) membranes.
+%questionable charge distribution from RESP -- a leeway for further tweaks of the FF.
+%It is not obvious that RESP charges provide the best description, especially due to its non-uniqe solution. 
+%Shall we solve RESP fitting with the constraint of full charges and then scale down, or shall we rather 
+%solve the fitting with a scaled total charge target?
 
-Samuli: From the point of view of this paper, the most relevant other lipid to study would be DPPC to see if we can reproduce the difference between POPC and DPPC in experiments.
-
-Joe: In addition, there are more experimental OP data on DPPC. However, it is not necessary and we can leave it to the community project.
-
-Role of water model: we use OPC3 (current best), it would be worth giving an estimate how results change when we use say SPCE or even TIP3p at least in the SI (so that the reader knows what errors to expect comming from these sub-optimal models).
-In addition, there is protein force field Amber15-FB, which uses water close to OPC3, TIP3pFB \todoii{REFs}{add references here}.
-On the other hand, it might be the best if we used a model that doesn't have the dielectric constant from nuclei 78, but rather 44 -- TIP4p2005 is the closest. \cite{OPC_paper, ForceBalance_paper}
-
-questionable charge distribution from RESP -- a leeway for further tweaks of the FF.
-It is not obvious that RESP charges provide the best description, especially due to its non-uniqe solution. 
-Shall we solve RESP fitting with the constraint of full charges and then scale down, or shall we rather solve the fitting with a scaled total charge target?
-
-acheivable accuracy of the MD engines themselves (mainly Hector's worry) is another limiting factor in fine tuning parameters -- solid physical ground helps.
-
-application of the correction to other lipid models: The rule looks general, however, it depends on how accurate the original model was.
-From preliminary simulations with POPE, it looks that the rule works for Lipid14 FF, at least for zwitterionic headgroups.
-
-\section{Conclusion}
-
-%Reiterate what we did...
+\section{Conclusions}
+We show that the Na+ and Ca2+ binding in phospholipid bilayers can
+be accurately described with MD simulation models, where electronic
+polarization is effectively included by using ECC correction.
+This is a significant advantage to the other available lipid models,
+which all overestimate specific cation binding affinities.  
+The proposed model reproduces the lipid bilayer structural details
+with similar accuracy as the other state of the art lipid models.
+overestimated cation binding in classical molecular
+dynamics simulations can be corrected 
 We present models of POPC and DPPC lipids that for the first time exhibit accurate headgroup order parameter response to cation binding. 

### 2017-06-14 Manuscript/proto-manuscript.tex
Updates

@@
 If such charges are obtained in an implicit solvent, 
-they vary the most for the polar moieties \cite{maiejewski14}. 
+they vary the most for the polar moieties \cite{maciejewski14}. 
 In order to represent this difference in polarity in a model with fixed charges, 
@@
 The decrease of area was speculated 
-\todo{Did you analyze the effect of this to hydration or did we only speculate?}
+\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?}
 to arise from reduced hydration
@@
 have stonger electrostatic interactions with them.
-\todo{This effect may have an official name. In that case we should mention it.}
+\todo{SAMULI: This effect may have an official name. In that case we should mention it.}
 Indeed, the scaling of $\sigma$ increased the area per molecule
@@
 where plateau concentration is the number density in unist of 1/(nm)$^3$.
+\todo{SAMULI: Once we have to final results, we can probably say that
+  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
+  }
 
@@
 \todo{To be written}
+
+\subsubsection{Simulations with cationic surfactants}
+Automated topoly builder \cite{malde11} was first used to create pdb structure of
+dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
+The AmberTools \cite{amber} was then used to generate the Amber force field
+parameters from the pdb file. These were converted to Gromacs format by using
+acpype tool \cite{acpype}. The partial charges were then manually modified
+to approximately correspond the corresponding segments in Lipid14 \cite{dickson14}.
+The surfactants were randomly mized with lipids to form bilayer structures with
+mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfctant in POPC bilayer.
+Simulation parameters and force field for POPC was already used and
+described previously \cite{catte16} and were downloaded from \cite{lipid14POPC0mMNaClfiles}.
+The systems were simulated 100-200~ns at 313K and the first 20~ns of the trajectories
+were excluded from the analysis as an equilibration period.
+The simulations were ran with Gromacs 5~\cite{abraham15}. 
+Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
 
@@
     \hline
-    experiment \cite{Jambeck2012}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
+    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
     experiment  & 64.3  &  303    \\
@@
 also the order parameter changes as a function of bound cations.
-\todo{I think that we should make a separate figures for this one and ion concentrations.
+\todo{SAMULI: I think that we should make a separate figures for this one and ion concentrations.
   In the current figure it would be difficult to show this with a reasonable scale.
@@
 The use of a ECC-ion model~\cite{kohagen14,kohagen16} along with original Lipid14 does not significantly change it
-\todo{If we show density profile for this, we should show also the order parameter changes}.
+\todo{SAMULI: If we show density profile for this, we should show also the order parameter changes}.
 The new ECC-lipid model with scaled ions exhibits on the other hand smaller density in this region 
@@
 described with classical MD simulation model with effective included of electronic polarizability.
-
 \begin{figure}[]
@@
     and with ECC ions and ECCLipids17 with ECC ions. }
-  \todo{We should add the location of bilayer here somehow.
+  \todo{SAMULI: We should add the location of bilayer here somehow.
     In NMRlipids II the location of phosphate was shown with green vertical line.}
-  \todo{I would add Aqvist data from NMRlipids II in here as well.}
+  \todo{SAMULI: I would add Aqvist data from NMRlipids II in here as well.}
 \end{figure}
 
-One possible example is to adress the discourse on the stoichiometry of \ce{Ca^{2+}} binding 
-to POPC~\cite{altenbach84}.
-In line with the early experimental finding~\cite{altenbach84}, we find that our data 
-fit well a ternary complex model, which assumes 2 POPC molecules per 1 \ce{Ca^{2+}}. 
-In contrast to this experiment, however, we possess also a complete atomistic detail of this phenomenon.
-Hence we can support this empirical finding with a direct observation of transient complexes of \ce{Ca^{2+}} and POPC.
-The simulation suggests that the stoichiometry 2~POPC:1~\ce{Ca^{2+}} is the most common complex. 
-In addition, there also is a non-negligable probability of one \ce{Ca^{2+}} cross-bridging three POPC molecules. 
-This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to 
-the binding isotherm that assumes 2~POPC per 1~\ce{Ca^{2+}}. 
-\todo{Analyze stoichiometry for \ce{Na^+}, \ce{Ca^{2+}}, their interaction energies with the lipid membrane, 
-etc, and finalize the discussion after these results.}
+Good agreement of ECC corrected model with experiments encourages us to analyse the
+binding details from MD simulations. Direct analysis of contacts between ions and
+lipids from simulations suggest that most common complex are ones with stoichiometry
+of 2~POPC:1~\ce{Ca^{2+}}.
+\todo{SAMULI: Details of this analysis have to be added.}
+As shown in Fig. \ref{fig:cacl-bind} this is in agreement
+with the ternary complex model suggested based on headgroup order parameter
+experiments~\cite{altenbach84}.
+\todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
+  be the only option. I will recheck and come back to this.}
+%This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to 
+%the binding isotherm that assumes 2~POPC per 1~\ce{Ca^{2+}}. 
+\todo{SAMULI: I would also analyze how much there is contact between ions and different
+parts of the lipid (phosphase, carbonyl, etc.).}
+In addition to the ternary complexes, there also is a non-negligable probability
+of one \ce{Ca^{2+}} cross-bridging three POPC molecules
+\todo{SAMULI: I think we should quantify this, i.e. how much there are these.
+  Maybe also the other possible complexes? Maybe also the correlation betweem complexes
+  and binding cites, if it is not too much work.}.
 
-It is also suggested that the addition of \ce{NaCl} to the solution of \ce{CaCl_2} enhances the hedgroup 
-order parameter response compared to the solution with only \ce{CaCl_2}. \cite{altenbach84}
-\todo{Simulate this effect and discuss it further}
+\todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},
+  their interaction energies with the lipid membrane, etc,
+  and finalize the discussion after these results.}
 
-\todo{The difference between DPPC and POPC -- simulate and compare with experiment. }
+%It is also suggested that the addition of \ce{NaCl} to the solution of \ce{CaCl_2} enhances the hedgroup 
+%order parameter response compared to the solution with only \ce{CaCl_2}. \cite{altenbach84}
+%\todo{Simulate this effect and discuss it further}
+
+%\todo{The difference between DPPC and POPC -- simulate and compare with experiment. }
 
@@
   \caption{\label{fig:cacl-bind}
-    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{Altenbach84} fits the simulation data nicely.}
+    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits the simulation data nicely.}
 \end{figure}
@@
 with similar accuracy as the other state of the art lipid models.
-overestimated cation binding in classical molecular
-dynamics simulations can be corrected 
-We present models of POPC and DPPC lipids that for the first time exhibit accurate headgroup order parameter response to cation binding. 
-The models are derived from Lipid14 model \cite{dickson14} by applying the electronic continuum correction. 
-The models were optimized to represent correct membrane structure, 
-and they were validated with the use of electrometer concept \cite{seelig87,Altenbach85,Altenbach84}. 
-Compared to the phenomenological observations of calcium:POPC stoichiometry in the experimental work \cite{Altenbach84}, 
-our simulations reveal the same stoichiometry of 2~POPC molecules per 1~calcium but through direct observation. 
-This tells us on the subtle effects of calcium in phospholipid membranes 
-possibly leading to a better understanding of their physical properties (elasticity etc.). 
-\todo{Improve this concluding-discussion with actual insights.}
+The correction is applied here on Lipid14 POPC model \cite{dickson14},
+but we expect that the correction can be generalized also for other lipids
+and force fields.
+%The models were optimized to represent correct membrane structure, 
+%and they were validated with the use of electrometer concept \cite{seelig87,Altenbach85,Altenbach84}.
+
+Direct analysis of calcium binding details from MD simulations is in agreement
+with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
+In this model 1~calcium binds to 2~POPC molecules, which together form a ternary
+complex.
+%This tells us on the subtle effects of calcium in phospholipid membranes 
+%possibly leading to a better understanding of their physical properties (elasticity etc.). 
+%\todo{Improve this concluding-discussion with actual insights.}
 

### 2017-07-26 Manuscript/proto-manuscript.tex
MS: modified section Methods

@@
 \subsection{Electronic continuum correction for lipid bilayers}
-According to the electronic continuum correction (ECC) the electronic
-polarizability can be effectively included in MD simulations by scaling
-the partial charges with a constant factor $f_q = \epsilon _{el} ^{-1/2}$, 
-where $\epsilon _{el}$ is the electronic part of the dielectric constant of 
-media \cite{leontyev11}. Measurements of high frequency 
+According to the electronic continuum correction (ECC)\cite{leontyev11}, electronic
+polarizability can be included in classical MD simulations by
+placing all particles into a homogeneous dielectric continuum 
+with a dielectric constant $\epsilon _{el}$. 
+which is the electronic part of the dielectric constant of 
+the media \cite{leontyev11}. Measurements of high frequency 
 dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
 for almost any biomaterial \cite{some_original_work, leontyev11}. % that exists in living cells. 
+Such a continuum can be easily included in standard MD simulation by
+a formal transformation of partial charges 
+\begin{equation}
+  Q^{ECC} = f_q \cdot Q
+\end{equation}
+with a constant scaling factor $f_q = \epsilon _{el} ^{-1/2}$ 
+effectively representing the newly introduced electronic continuum. 
 The value measured for water, $\epsilon _{el} = 1.78$, gives 
@@
 
-Here we apply the same approach to lipid bilayers to accurately 
+Here we apply the electronic continuum correction to lipid bilayers to accurately 
 describe the lipid headgroup response to Na+ and Ca2+ concentrations \cite{catte16}. 
@@
 The partial atomic charges in Lipid14 were derived by fitting the electrostatic 
-potential to its model quantum chemistry representation in vacuum.
-\todo{SAMULI: I do not fully understand the end of this paragraph}
-If such charges are obtained in an implicit solvent, 
+potential to its model quantum chemistry representation (RESP\cite{RESP_paper}) in vacuum.
+If the charges are obtained by using RESP in an implicit solvent, 
 they vary the most for the polar moieties \cite{maciejewski14}. 
-In order to represent this difference in polarity in a model with fixed charges, 
-we shall use the average charges from both environments (IpolQ charges) \cite{Cerutti2013}. 
+In order to implicitly capture solvent induced polarization 
+in models with fixed charges, 
+we shall use the average of partial charges 
+from both environments (i.e. vaccum and solvent),
+so called implicitly polarized charges (IpolQ) \cite{Cerutti2013}. 
 By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
 we represent the IpolQ charges in the electronic continuum correction of the Lipid14 model
-by increasing the scaling factor for charges to $f_q = 0.8$. 
-\todo{SAMULI: I tried to repeat this calculation but I was not able to get 0.8.
-This has to be explained better.}
-The newly presented ECC-POPC model is based on the POPC model from
+by increasing the scaling factor $f_q$ to $0.8$. 
+\footnote{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom types being around 0.78--0.80.
+Even though the methods are almost identical, authors of Lipid 14 find lower partial charges in vacuum than here -- so I prefer the higher value.
+As the choice of charges is arbitrary anyway, I use 0.80 as an approximate round value.
+The use of 0.78---.79 might be more appropriate, though.}
 
@@
 some improvement in all available lipid models \cite{botan15}. 
-To minimize the intercalation with the highly optimized hydrocarbon
-parameters, we decided to apply ECC correction only
+To minimize the detuning of the highly optimized hydrocarbon
+parameters, we apply ECC only
 to the headgroup, glycerol backbone and carbonyl in acyl chains.
-These regions are also most polar parts in lipids and are expected
-to most affect on cation binding.   
+These regions are also the most polar parts in lipids and are expected
+to most affect cation binding.   
 %Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
@@
 
-The scaling of these partial charges by a factor of 0.8 reduce the 
-area per molecule to ??, which is significantly smaller than the
-experimental and original Lipid 14 values.  
+Mere scaling of partial charges in the modified region by the factor $f_q$ reduces the 
+area per molecule to ??\todo{find the value}, which is significantly smaller than the
+experimental value (\cite{}) and the original Lipid14 values (\cite{}).  
 The decrease of area was speculated 
-\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?}
+\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
+JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
 to arise from reduced hydration
@@
 The hydration can be increased by decreasing the radius of atoms
-by reducing $\sigma$ term in Lennart-Jones potential, which
-allows water molecules approach closer to lipid atoms and
+by reducing the $\sigma$ parameters in Lennard-Jones potential for the selected atoms.
+This allows water molecules to approach closer to lipid atoms and
 have stonger electrostatic interactions with them.
 \todo{SAMULI: This effect may have an official name. In that case we should mention it.}
-Indeed, the scaling of $\sigma$ increased the area per molecule
-and the results close to experimental and original Lipid14 values were achieved
-with scaling factor of ?? for $\sigma$ and 0.8 for partial charges.
+Indeed, by reducing $\sigma$ with a scaling factor $f_\sigma = 0.89$ 
+we increased the area per molecule
+to a level close to experimental and original Lipid14 values. 
 %\todo{Is there some justification for the below sentence?} 

### 2017-07-27 Manuscript/proto-manuscript.tex
MS: updated section Comparison with experiments

@@
 from both environments (i.e. vaccum and solvent),
-so called implicitly polarized charges (IpolQ) \cite{Cerutti2013}. 
+so called implicitly polarized charges (IPolQ) \cite{ipolq2013}. 
 By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
-we represent the IpolQ charges in the electronic continuum correction of the Lipid14 model
+we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
 by increasing the scaling factor $f_q$ to $0.8$. 
@@
 area per molecule to ??\todo{find the value}, which is significantly smaller than the
-experimental value (\cite{}) and the original Lipid14 values (\cite{}).  
+experimental value (\cite{}) and the original Lipid14 values (\cite{})\todo{Add the values}.  
 The decrease of area was speculated 
@@
 \subsection{Comparison to experimental data}
-Structures sampled by individual lipid molecules were compared
+Structures sampled by individual lipid molecules in simulations were compared
 to experimental data by using C-H bond order parameters \cite{ollila16}
@@
 normal and average is taken over all sampled configurations.
+
 Bilayer dimensions were compared to experiments by using the
 scattering form factor \cite{ollila16}
-\todo{Add equation}.
+\begin{equation}
+  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
+\end{equation}
+where $f_\alpha(q_z)$ is the density of atomic scattering length, 
+$\rho_s$ is the density of solvent scattering lenght,
+$n_\alpha (z)$ is the number density of atom $\alpha$ and
+$z$ is the distance from the membrane centre along its normal 
+spanning the membrane with thickness $D$.
 
@@
 as a change of lipid headgroup dipole tilt to more vertical orientation
-with bound positive charge and {\it vice versa} for negative \cite{seelig87}.
+with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}.
 
@@
 %moment to guide me on the possible finite-size errors and this conc-error.
-The used experimental data reports order parameters as a function of
-equilibrium cation concentration in bulk solvent \cite{akutsu81,altenbach84}.
-The box size in simulations was adjusted such that ion conceration reaches 
-a clear bulk plateau, which then gives the concentration that corresponds to the
-one reported in experiments. The concentrations in units of mol/l were then 
+The used experimental data report order parameters as a function of
+equilibrium cation concentration in the bulk solvent \cite{akutsu81,altenbach84}.
+Such a condition is reached in simulations by adjusting
+the simulation box size to dimensions large enough
+that ion conceration reaches a clear plateau in the bulk solvent.
+The concentrations in the units of $\mathrm{mol/l}$ were then 
 determined as
@@
 \end{equation}
-where plateau concentration is the number density in unist of 1/(nm)$^3$.
+where plateau concentration is the number density in the units of $\mathrm{nm}^{-3}$.
 \todo{SAMULI: Once we have to final results, we can probably say that

### 2017-07-27 Manuscript/proto-manuscript.tex
There was a manuscript section accidentaly erased This cherry-pick merge restores it (Sim details) from commit MS: wrote section Simulation details

@@
 We also provide for the first time a realistic stochiometry of bound calcium cations to a POPC membrane, and their binding sites. 
-This work will continue as an open collaboration project NMRlipids IV (http://nmrlipids.blogspot.fi).
+This work will continue as an open collaboration project NMRlipids VI (http://nmrlipids.blogspot.fi).
 \end{abstract}
@@
 like in signal propagation in neurons and vesicle fusion.
-Cation interactions with lipid bilayers serving as a model for cellular
-membranes are thus widely studied with experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
-and theoretical methods \cite{Berkowitz12,??}. General conclusion from experimental
-studies has been that multivalent ions and lithium have weak specific binding in
-phopholipid bilayers, while other monovalent ions do not essentially
-bind \cite{cevc90,tocanne90,seelig90,binder02}. The presence of anoinic lipids, like PS or PG,
-increase the concentration close to the bilayer and thus the amount of bound
-ions, but do not affect the specific binding constant \cite{seelig90}.
-The binding details, like binding sites and stoichiometry are not yet fully
-resolved but interpretation of NMR and scattering experiments suggest that one
-Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
-phospholipid molecules \cite{altenbach84}.
+Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
+to understand molecular level details of specific cation interactions with
+cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
+and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} \todo{We need to select references for these}.
+While relative binding affinity of different ions is agreed to follow Hoffmeister
+serie, the molecular details of binding and binding energetics are
+not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
+%General conclusion from experimental
+%studies has been that multivalent ions and lithium have weak specific binding in
+%phopholipid bilayers, while other monovalent ions do not essentially
+%bind 
+%The presence of anoinic lipids, like PS or PG,
+%increase the concentration close to the bilayer and thus the amount of bound
+%ions, but do not affect the specific binding constant \cite{seelig90}.
+Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
+and infrared scpectroscopy, give accurate information about ion binding
+in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
+Molecular level interpretation of the results, however, requires assumptions
+about binding model and is often not fully conclusive \cite{??}.
+On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
+have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
 
-Classical molecular dynamics simulations have potential to solve the ion binding details
-in lipid bilayers and reveal its relevance to various biological
-problems \cite{bockmann03,bockmann04,??}. However, the available
-molecular dynamics simulation models have a strong tendency to overestimate
-cation binding on zwitterionic bilayers and none of them can reproduce the
-experimental data with the accuracy required for the interpretation of
-experiments \cite{catte16}. The overestimated specific cation binding also
-makes zwitterionic bilayers effectively positively charged, which could
-potentially lead to significant artefacts in applications.
-Thus, there is a vast demand for classical molecular dynamics simulation
-model which correctly reproduces cation binding in lipid bilayers.
+%The binding details, like binding sites and stoichiometry are not yet fully
+%resolved but interpretation of NMR and scattering experiments suggest that one
+%Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
+%phospholipid molecules \cite{altenbach84}.
 
-The lack of electronic polarizability from the classical MD simulation
-models is a potential source of artefact, which could lead to overbinding
-of cations. 
-%Almost all lipid bilayer simulations reported in the literature 
-%are done with force fields that represent interactions between atoms
-%with pairwise additive empirical potential and exclude the electronic 
-%polarizability \cite{SOME_review, FF_papers}.
-The issue has been considered higly relevant since the early days of
-lipid bilayer simulations and pioneering simulation studies scaled
-the partial charges to effectively include electronic
-polarizability \cite{jonsson86,egberts94}. The explicit inclusion of
-electronic polarizability has been, however, turned to be 
-practically complicated and implementations for lipids are rare \cite{chowdhary13}.
-
-In this work we show that the cation overbinding in classical MD simulations
-can be corrected by including electronic polarizability by using effective continuum
-correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules.
-This is essentially physicalle well justified version of partial charge scaling
-implemented in early days of lipid and surfctant simulations \cite{jonsson86,egberts94}. 
+In this work we show that the accuracy of Calcium binding
+on zwitterionic PC lipid bilayer can be significantly improved by including electronic polarizability
+by using effective continuum correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules
+in classical MD simulation model.
+This is essentially physically well justified version of partial charge scaling
+implemented in early days of lipid and surfactant simulations \cite{jonsson86,egberts94}. 
 The approach has been previously shown in to improve bulk performace of
@@
 
+%Classical molecular dynamics simulations have potential to solve the ion binding details
+%in lipid bilayers and reveal its relevance to various biological
+%problems \cite{bockmann03,bockmann04,??}. However, the available
+%molecular dynamics simulation models have a strong tendency to overestimate
+%cation binding on zwitterionic bilayers and none of them can reproduce the
+%experimental data with the accuracy required for the interpretation of
+%experiments \cite{catte16}.
+
+The proposed MD simulation model with ECC-corrected lipids and ions
+reproduce the experimental NMR data of Calcium binding in PC lipid bilayers
+and mainly supports the proposed ternary complex model (2 Ca2+ bind in
+on POPC) \cite{akutsu81,altenbach84}. However, non-negligible contribution
+of one Ca2+ cross bridging three POPC molecules is also observed.
+The overestimated specific cation binding observed in current lipid models \cite{catte16}
+may lead to articifially positively charged memranes and significant artefacts in
+MD simualtions. Thus, the proposed approach to improve the accuracy of cation
+binding in current MD models is highly useful for future MD simulations
+with physiological salt conditions.
+
+%re is a vast demand for classical molecular dynamics simulation
+%model which correctly reproduces cation binding in lipid bilayers.
+
 \section{Methods}
@@
 \subsection{Electronic continuum correction for lipid bilayers}
-According to the electronic continuum correction (ECC)\cite{leontyev11}, electronic
+The lack of electronic polarizability from the classical MD simulation
+models has been considered higly relevant issue since the early days of
+lipid bilayer simulations and pioneering simulation studies scaled
+the partial charges by half to effectively include electronic
+polarizability \cite{jonsson86,egberts94}. Also approaches that
+explicitly include electronic polarizability has been introduced, but
+this has turned out to be practically complicated \cite{chowdhary13}.
+In this work we effectively include electronic polarizability in lipid
+bilayer simulations by using the electronic continuum correction
+(ECC)~\cite{leontyev11}. The approach is more sophisticated version of
+early studies with scaling to half was used \cite{jonsson86,egberts94},
+but significantly more simple than the explicit inclusion of electronic
+polarizability \cite{chowdhary13}.
+
+%is a potential source of artefact, which could lead to overbinding
+%of cations. 
+%Almost all lipid bilayer simulations reported in the literature 
+%are done with force fields that represent interactions between atoms
+%with pairwise additive empirical potential and exclude the electronic 
+%polarizability \cite{SOME_review, FF_papers}.
+
+According to ECC, electronic
 polarizability can be included in classical MD simulations by
 placing all particles into a homogeneous dielectric continuum 
-with a dielectric constant $\epsilon _{el}$. 
+with a dielectric constant $\epsilon _{el}$, 
 which is the electronic part of the dielectric constant of 
@@
 dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
-for almost any biomaterial \cite{some_original_work, leontyev11}. % that exists in living cells. 
+for almost any biomaterial \cite{some_original_work, leontyev11}.
 Such a continuum can be easily included in standard MD simulation by
@@
 
-Here we apply the electronic continuum correction to lipid bilayers to accurately 
-describe the lipid headgroup response to Na+ and Ca2+ concentrations \cite{catte16}. 
-We use the Lipid14~\cite{dickson14} force field parameters as a starting point,
-because they give the most realistic headgroup response with added cations 
-and relatively realistic glycerol backbone and headgroup structures
-when compared with other state of the art lipid models \cite{botan15,catte16}.
-The partial atomic charges in Lipid14 were derived by fitting the electrostatic 
-potential to its model quantum chemistry representation (RESP\cite{RESP_paper}) in vacuum.
-If the charges are obtained by using RESP in an implicit solvent, 
-they vary the most for the polar moieties \cite{maciejewski14}. 
-In order to implicitly capture solvent induced polarization 
-in models with fixed charges, 
-we shall use the average of partial charges 
-from both environments (i.e. vaccum and solvent),
-so called implicitly polarized charges (IPolQ) \cite{ipolq2013}. 
-By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
-we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
-by increasing the scaling factor $f_q$ to $0.8$. 
-\footnote{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom types being around 0.78--0.80.
-Even though the methods are almost identical, authors of Lipid 14 find lower partial charges in vacuum than here -- so I prefer the higher value.
-As the choice of charges is arbitrary anyway, I use 0.80 as an approximate round value.
-The use of 0.78---.79 might be more appropriate, though.}
+The main goal of this work is to apply ECC correction to accurately 
+describe the lipid headgroup response to Na$^+$ and Ca$^{2+}$ concentrations
+in MD simulation models when compared with NMR data \cite{catte16}. 
+The Lipid14~\cite{dickson14} force field parameters were used as a starting point,
+because their response to bound ions was most realistic against NMR data
+(see Fig.~5 in Ref.~\citenum{catte16}).
+Also glycerol backbone and headgroup structures in Lipid14 model were
+relatively realistic when compared with other state of the art lipid models \cite{botan15}.
 
-Hydrocarbon chains in Lipid14 and other lipids models are highly optimized and
+While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
+to improve model performance against scattering data \cite{kohagen14,kohagen16,??},
+it is not clear if the same factor should be used for lipid headgroups \todo{, because ??}.
+Here we estimated the correct scaling factor for Lipid14 model by using
+so called implicitly polarized charges (IPolQ) \cite{ipolq2013},
+which are the average of partial charges given by RESP~\cite{RESP_paper} calculations in
+vacuum and in implicit solvent.
+%In order to implicitly capture solvent induced polarization 
+%in models with fixed charges, 
+%we shall use the average of partial charges 
+%from both environments (i.e. vaccum and solvent),
+%so called implicitly polarized charges (IPolQ) \cite{ipolq2013}. 
+The partial charges in original Lipid14 parametrization were derived by fitting the electrostatic 
+potential to its model quantum chemistry representation (RESP\cite{RESP_paper}) in vacuum,
+while charges from calculations in an implicit solvent vary most for the polar moieties \cite{maciejewski14}. 
+Here we estimate the appropriate scaling factor for lipid hedgroups and glycerol backbone
+by comparing IPolQ charges (average of RESP calculation in vacuum and in implicit solvent) to
+the RESP results in vacuum. By using the results reported in \cite{maciejewski14} we conclude
+that the IPolQ charges can be achived from vacuum calculations by scaling roughly with the
+factor of $f_q=0.8$, which is the used to derive ECC-corrected Lipid14 model.
+\todo{SAMULI: please check that this is correctly descrived.}
+%By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
+%we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
+%by increasing the scaling factor $f_q$ to $0.8$. 
+\footnote{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
+  types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
+  partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
+  I use 0.80 as an approximate round value. The use of 0.78---.79 might be more appropriate, though.}
+
+Here we apply ECC correction with the proposed scaling factor only to the headgroup,
+glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
+are expected to have the largest contribution to the cation binding.
+The hydrocarbon chain parameters are not modified, because they are
+highly optimized in Lipid14 and other lipids models and
 give generally a good description for hydrophobic part of lipid bilayers
@@
 some improvement in all available lipid models \cite{botan15}. 
-To minimize the detuning of the highly optimized hydrocarbon
-parameters, we apply ECC only
-to the headgroup, glycerol backbone and carbonyl in acyl chains.
-These regions are also the most polar parts in lipids and are expected
-to most affect cation binding.   
+%To minimize the detuning of the highly optimized hydrocarbon
+%parameters, 
 %Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
@@
 %we exclude alkyl tails from our modifications. 
-
-Mere scaling of partial charges in the modified region by the factor $f_q$ reduces the 
+The scaling of partial charges in the polar region was found to decrease
 area per molecule to ??\todo{find the value}, which is significantly smaller than the
 experimental value (\cite{}) and the original Lipid14 values (\cite{})\todo{Add the values}.  
-The decrease of area was speculated 
-\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
-JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
-to arise from reduced hydration
-of headgroup due to the lower polarity of molecules with scaled charges.
-The hydration can be increased by decreasing the radius of atoms
-by reducing the $\sigma$ parameters in Lennard-Jones potential for the selected atoms.
-This allows water molecules to approach closer to lipid atoms and
-have stonger electrostatic interactions with them.
-\todo{SAMULI: This effect may have an official name. In that case we should mention it.}
-Indeed, by reducing $\sigma$ with a scaling factor $f_\sigma = 0.89$ 
-we increased the area per molecule
+The decrease of area was found to arise from lower hydration of the lipid headgroup region,
+%\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
+%JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
+which can be explained by the increased solvation free energy due to the lower polarity
+of molecules with scaled charges. The solvation free energy can be decreased
+and hydration increased by reducing the radius of beads by changing the 
+the $\sigma$ parameters in Lennard-Jones potential for the selected atoms.
+%This allows water molecules to approach closer to lipid atoms and
+%have stonger electrostatic interactions with them.
+%\todo{SAMULI: This effect may have an official name. In that case we should mention it.}
+Here we reduce $\sigma$ with a scaling factor of $f_\sigma = 0.89$ for the same atoms for which
+charges was scaled. This increased the area per molecule back
 to a level close to experimental and original Lipid14 values. 
@@
 
-\subsection{Comparison to experimental data}
-Structures sampled by individual lipid molecules in simulations were compared
-to experimental data by using C-H bond order parameters \cite{ollila16}
-\begin{equation}
-S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
-\end{equation}
-where $\theta$ is the angle between C-H bond and membrane
-normal and average is taken over all sampled configurations.
-
-Bilayer dimensions were compared to experiments by using the
-scattering form factor \cite{ollila16}
-\begin{equation}
-  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
-\end{equation}
-where $f_\alpha(q_z)$ is the density of atomic scattering length, 
-$\rho_s$ is the density of solvent scattering lenght,
-$n_\alpha (z)$ is the number density of atom $\alpha$ and
-$z$ is the distance from the membrane centre along its normal 
-spanning the membrane with thickness $D$.
-
+\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept}
 Ion binding was compared between experiments and simulations by 
@@
 are proportional to the amount of bound charge
-in lipid bilayer \cite{seelig87}. More recent analysis included also
+in lipid bilayer \cite{seelig87}.
+More recent analysis included also
 the order parameter signs and concluded that the order parameters  
@@
 with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
+Order parameters for C-H bonds can be accurately measured for each
+C-H bond of lipids by using H$^2$ NMR or $^{13}$C NMR techniques and
+calculated from MD simulations directly from definition
+\begin{equation}\label{OP}
+S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
+\end{equation}
+where $\theta$ is the angle between the bond and membrane
+normal and average is taken over all sampled configurations \cite{ollila16}.
+
+The change of the headgroup order parameters as a function of
+the amount of bound charge per lipid $X^\pm$ 
+can be written as~\cite{ferreira16}
+\begin{equation}\label{OPchangeEQ}
+\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) =\frac{4m_i }{3\chi}X^\pm,
+\end{equation}
+where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
+$i$ refers to either $\alpha$ or $\beta$ carbon,
+%the empirically observed linear relation %for the change in order parameters %with respect to a bilayer without bound charges              
+%\begin{equation}\label{electrometer_eq}                                                                                                   
+%S_{\rm{CH}}^{i}(X^\pm)=S_{\rm{CH}}^{i}(0) + \frac{4m_i}{3\chi} X^\pm,                                                                     
+%\end{equation}                                                                                                                            
+$m_i$ is an empirical constant depending on the valency and position of bound charge,
+and the value of the quadrupole coupling constant is $\chi \approx$\,167\,kHz.
+Combination of atomic absorption spectra, $^2$H NMR experiments
+and information about order parameter signs gave
+$m_\alpha=-20.5$  and $m_\beta=-10.0$ for Ca$^{2+}$ binding to POPC
+bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
+
+Recent work published by the NMRlipids project showed that the concept works
+qualitatively also in simulations and can be used to compare ion binding
+affinitity between simulations and experiments~\cite{catte16}. However,
+it also turned out that also the sensitivity of the order parameters response
+to bound charge has to be quantified more accurately for quantitative comparison
+of binding affinity. In this work we first we first quantify the response of
+headgroup order parameters against experimental data from cationinc surfactants
+embedded in PC bilayer~\cite{scherer89}. Essentially all surfactants locate in
+lipid bilayers in these experiments, thus the amount of bound charge is known
+and can be plotted against order parameter change.
 
@@
 
+\subsection{Comparison of lipid bilayer structure to experimental data}
+Lipid bilayer structure without ions was validated against NMR experiments
+by order parameters for C-H bonds and x-ray scattering experiments by using
+form factors. The former validates the structures sampled by individual
+lipid molecules in simulations with segmental resolution, while latter
+validates the dimensions of the lipid bilayer (thickness and area per
+molecule)~\cite{ollila16}. Form factor was calculated from simulations
+as 
+\begin{equation}
+  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
+\end{equation}
+where $f_\alpha(q_z)$ is the density of atomic scattering length, 
+$\rho_s$ is the density of solvent scattering lenght,
+$n_\alpha (z)$ is the number density of atom $\alpha$ and
+$z$ is the distance from the membrane centre along its normal 
+spanning the membrane with thickness $D$.
+
 \subsection{Simulation details}
 
-\todo{To be written}
+\subsubsection{Simulations with aqueous ions}
+The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
+(POPC) bilayer and an aqueous salt solutions of varying concentrations.
+Water molecules were described by OPC3 model \cite{Izadi16}.
+In order to test transferability of the newly developed ECC-lipids model, 
+we also performed several additional simulations with water models 
+OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}
+presented in Supporting Information (SI). 
+We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+Simulations with Lipid14 use ion model by Dang \cite{smith94,chang1999,dang2006}. 
+Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
+The simulation settings used in this work are 
+summarized in Table \ref{tbl:mdpar}, 
+and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}. 
+Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+
+\begin{table}[btp]
+  \caption{Simulation parameters}
+  \label{tbl:mdpar}
+  \begin{tabular}{ll}
+    simulation property & parameter   \\
+    \hline
+    time-step           & 2~fs         \\
+    equilibration time  & 100~ns  \\
+    simulation time     & 200~ns  \\
+    temperature         & 313~K       \\
+    thermostat          & v-rescale  \cite{bussi07}   \\
+    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
+    long-range electrostatics & PME  \cite{darden93}  \\
+    cut-off scheme      & Verlet \cite{Pall13}      \\
+    Coulomb and VdW cut-off & 1.0~nm \\
+    constraints         & LINCS, only Hydrogen atoms \cite{hess97} \\
+    constraints for water & SETTLE  \cite{miyamoto92} \\
+    \hline
+  \end{tabular}
+\end{table}
 
 \subsubsection{Simulations with cationic surfactants}
-Automated topoly builder \cite{malde11} was first used to create pdb structure of
+Automated topology builder \cite{malde11} was first used to create the structure of
 dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
-The AmberTools \cite{amber} was then used to generate the Amber force field
-parameters from the pdb file. These were converted to Gromacs format by using
+The code AmberTools \cite{amber} was then used to generate the Amber-type force field
+parameters. The parameters were converted to the Gromacs format by using
 acpype tool \cite{acpype}. The partial charges were then manually modified
-to approximately correspond the corresponding segments in Lipid14 \cite{dickson14}.
-The surfactants were randomly mized with lipids to form bilayer structures with
-mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfctant in POPC bilayer.
-Simulation parameters and force field for POPC was already used and
-described previously \cite{catte16} and were downloaded from \cite{lipid14POPC0mMNaClfiles}.
-The systems were simulated 100-200~ns at 313K and the first 20~ns of the trajectories
-were excluded from the analysis as an equilibration period.
-The simulations were ran with Gromacs 5~\cite{abraham15}. 
-Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
+The surfactants were randomly placed among the lipids to form bilayer structures with
+mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
+These systems were ran with original Lipid14 POPC model and with ECC-corrected model.
+\todo{Describe the changes in surfactant model used with ECC-corrected lipids.}
 
@@
 \subsection{Lipid bilayers without ions}
-The scattering form factors, NMR order parameters and area per lipids calculated from 
-the ECC corrected lipid model for POPC are compared to experiments and original
+The scattering form factors, NMR order parameters and area per lipid calculated  
+using the ECC-lipids model for POPC are compared to experiments and original
 Lipid14 results in Fig. \ref{simVSexpNOions} and in Table~\ref{tab:apls}. 
@@
 The structural quality is comparable to the state of art lipid models available in literature \cite{ollila16},
-thus we conclude that the ECC corrected lipid model reproduces the lipid bilayer structure
-in liquid disordered phase with similar accuracy than other available models
-\todo{Discussion to be finished when we have all the results in the figure}.
+thus we conclude that the ECC-lipids model reproduces the lipid bilayer structure
+in liquid disordered phase with similar accuracy to other available models.
+
 \begin{figure}[]
@@
   \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-%  \caption{\label{fig:form-f}
-%    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
-%\end{figure}
-%
-%\begin{figure}[]
-%  \centering
-  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17_q80_sig89.png}
   \caption{\label{simVSexpNOions}
-    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models. 
-    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and EECLipid17 models
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
+    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
     compared with experimental order parameters from \cite{ferreira13}.}
-  \todo{Add acyl chain order parameters, POPC chemical structure}
-  \todo{Should we add the results from original lipid14?}
+  \todo{Add acyl chain order parameters, POPC chemical structure} \\
+  \todo{We should run a simulation for the system withtout ions with the
+    same (or close) temperature as the experimental data. The NMR data is at 300K
+    so that would be good. I think that this would be also close enough for the scattering data at 303K.
+  }
 \end{figure}
-%Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
-%are shown in Table \ref{tab:apls}.
-\todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
+
 \begin{table}
-  \caption{Area per lipid from different models for POPC without ions\label{tab:apls} }
-  \begin{tabular}{c c c}
-    model          & A (Å$^2$)   & Temperature [K] \\
+  \caption{Area per lipid (APL) from different models of POPC without ions\label{tab:apls} }
+  \begin{tabular}{l|c c}
+    model          & APL (\AA$^2$)   & Temperature [K] \\
     \hline
-    Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
-    Lipid14ecc0.80+sigma0.875 &        &  313    \\
-    GMX small patch           & 64.9   &         \\
-    GMX 4xbig patch           & 65.5   &         \\
-    oMM small patch           & 63.65  &         \\
-    oMM 4xbig patch           & 63.7   &         \\
+    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
     \hline
-    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
-    experiment  & 64.3  &  303    \\
+    ECC-lipids &        &  \\
+    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
+    ($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
+    \hline
+    (6.4 nm)$^2$, 128 lipids patch, OPC3           & 64.2   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, SPC/E          & 65.1   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, OPC            & 64.4   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, TIP4p/2005     & 66.8   &  313       \\
+    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
+    %oMM small patch           & 63.65  &         \\
+    %oMM 4xbig patch           & 63.7   &         \\
+    \hline
+    %experiment   & 62.7  &  293    \\
+    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
     experiment  & 67.3  &  323    \\
-    experiment  & 68.1  &  333    \\
-    experiment POPE  & 56.6 &  303    \\
+    %experiment  & 68.1  &  333    \\
+    %experiment POPE  & 56.6 &  303    \\
     \hline
@@
 
-Headgroup order parameter response to bound charge was evaluated against
-experimental data measured with cationic surfactant
+Headgroup order parameter response to the bound charge was evaluated against
+experimental data measured with the cationic surfactant
 (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) \cite{scherer89}.
-The exact amount of bound charge in the membrane is known these systems, because
-practically all cationic surfactant molecules are embedded in lipid bilayer due to
-their amphiphilic nature. Thus, the headgroup order parameter changes as a function of
-mole fraction of cationic surfactants in  Fig. \ref{OrderParameterCHANGESnewMODELS} gives
-also the order parameter changes as a function of bound cations.
-\todo{SAMULI: I think that we should make a separate figures for this one and ion concentrations.
-  In the current figure it would be difficult to show this with a reasonable scale.
-  Currently Lipid14 results are cut out, thus the scale should be increased.
-  This would, however, make the changes with CaCl too small.
-  The change of headgroup tilt could be incorporated in the same figure with order parameter
-  response to cationic surfactant.
-}
-The headgroup order parameter response to bound cation concentration
-is approximately linear up to $\sim$0.3 mole fraction in experiments \cite{scherer89}.
-The linearity is also observed in simulations with original Lipid14 and with ECC correction.
-%from experiments \cite{seelig87} and simulations with Lipid14 model and with ECC-lipids model 
-%as a function of bound charge are shown in
+The exact amount of bound charge in the membrane is known in these systems, because
+practically all cationic surfactant molecules are embedded in the lipid bilayer due to
+their amphiphilic nature. Thus, the changes of headgroup order parameter as a function of
+mole fraction of cationic surfactants (Fig. \ref{OrderParameterCHANGESnewMODELS}) gives
+also the order parameter changes as a function of the amount of bound cations.
+\todo{Split the OP-changes figure in two: cationic surfactant -and- binding of cations.
+      The separate figure with the cationic surfactant shall be additionally accompanied
+      with a plot of Head group tilt (PN-vector?) change/response. }
+
+The headgroup order parameter response to the bound cation concentration
+is in experiments approximately linear up to $\sim$0.3 mole fraction of the cation \cite{scherer89}.
+The linearity is also observed in simulations with original Lipid14 and with newly derived ECC-lipids models.
 Quantitative comparison, however, reveals that the response is overestimated in 
-original Lipid14 for both segments, while the ECC corrected model gives a
-good agreement for the change of $\alpha$ segment order parameter, but slightly
+Lipid14 for both segments, while the ECC-lipids model gives a
+good agreement for the change of $\alpha$ segment order parameter, and slightly
 underestimates the $\beta$ segment order parameter change.
-The overestimation of order parameter changes with original Lipid14 model 
-suggests that the overestimated response with CaCl$_2$ concentration in \cite{catte16} 
-can be partly explained by the high sensitivity of the headgroup order parameter response
-to bound charge.
-
-Secondly, the $S^\alpha/S^\beta$ ratio of the response is found to be slightly larger than in 
+Such a finding is not surprising, as the $\Delta S^\alpha/\Delta S^\beta$ ratio 
+of the response of Lipid14 model is found to be slightly larger than in 
 experiments (VALexp vs VALsim
@@
 }
-).
-This property is kept also by the newly derived model, ECC-lipid, and, hence, 
-it shows a perfect agreement in the response of the headgroup order parameter $S^\alpha$ 
-but underestimates slightly the response of $S^\beta$.
+), and this property is also kept by the newly derived Lipid14-based model, ECC-lipids. 
+The present comparison of headgroup order parameter changes 
+suggests that the observed overestimated response of all models to 
+increasing CaCl$_2$ concentration in \cite{catte16} 
+can be explained at least in part by a high sensitivity of the headgroup order parameter response
+to the bound charge.
 
-\subsection{Cation binding in POPC bilayer}
+\subsection{Cation binding to POPC membrane}
 
 Headgroup order parameter response to increasing CaCl$_2$ concentration
-from experiments, original Lipid14 model and ECC corrected model are shown
+from experiments, original Lipid14 model and ECC-lipids model are shown
 in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
 is significantly overestimated in original Lipid14 model, while results from
-ECC corrected model are in good agreement with experiments.
-This is a significant improvement over previously available models,
-which always overstimate the order parameter response to CaCl$_2$
+ECC-lipids model are in a good agreement with the experiments.
+This is a significant improvement over current models,
+which in general overstimate the head group order parameter response to CaCl$_2$
 concentration \cite{catte16}. The good agreement with experiments indicate
-that the binding details of Ca2+ are realistic in the ECC corrected model
-and it can be thus used to study lipid-ion interaction details.
+that the binding details of \ce{Ca^2+} are realistic in the ECC-lipids model
+and, hence, it can be used to study lipid-ion interaction in detail.
 
@@
     Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration and
-    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$)
+    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) mole fraction
     from simulations and experiments (DPPC \cite{akutsu81}, POPC \cite{altenbach84}, surfactant \cite{scherer89}).
-    Simulations with Lipid14 and Åqvist ion model from \cite{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC1000mMCaClfiles}.
-  }
-  \todo{Add Lipid14-Aquist data.
-    Lipid14/Åqvist data to be added from https://github.com/NMRLipids/lipid\_ionINTERACTION/blob/master/Data/POPC/CaCl/LIPID14/LIPID14caclCONSchange.dat.
-    Joe: I'm not sure what is meant -- I think Aquist data are actually plotted there. I recently changed it to L14+Dang ions (both data and plot).
-    I think just one ion model is sufficient (and ECC-ions are based on Dang model).
-    Samuli: I think that we could add also the Aqvist data. This is in the NMRlipids II publication so it might
-    make easier to follow for people who have red it publication.
   }
@@
 
-Ion density profiles between different simulation models are compared in Fig. \ref{fig:cacl-dens}.
-Density profiles from simulations with original Lipid14 and Dang ions \cite{smith94,chang1999,dang2006} 
+Ion density profiles from simulations with different models are compared in Fig.~\ref{fig:cacl-dens}.
+The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
 show a pronounced peak in the position of the phosphate moieties of POPC. 
-The use of a ECC-ion model~\cite{kohagen14,kohagen16} along with original Lipid14 does not significantly change it
-\todo{SAMULI: If we show density profile for this, we should show also the order parameter changes}.
-The new ECC-lipid model with scaled ions exhibits on the other hand smaller density in this region 
+The use of a ECC-ion model~\cite{kohagen14,kohagen16} 
+or the standard Amber ion model by \AA{quist} 
+along with Lipid14 does not significantly change it 
+(Headgroup order parameter responses for these models in SI).
+\todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
+The new ECC-lipids model with ECC-ions exhibits on the other hand smaller density in this region 
 suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
-This demonstrates that cation binding in zwitterionic phospholipid bilayer can be accurately
-described with classical MD simulation model with effective included of electronic polarizability.
+Together with the accurate head group order parameter response (Fig.~\ref{OrderParameterCHANGESnewMODELS}),
+the decrease in cation density of ECC-lipids demonstrates that 
+ion binding to zwitterionic phospholipid membranes can be accurately
+described with classical MD simulation models that account for electronic polarizability.
+
 \begin{figure}[]
@@
   \caption{\label{fig:cacl-dens}
-    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Aquist parameters
-    and with ECC ions and ECCLipids17 with ECC ions. }
-  \todo{SAMULI: We should add the location of bilayer here somehow.
-    In NMRlipids II the location of phosphate was shown with green vertical line.}
-  \todo{SAMULI: I would add Aqvist data from NMRlipids II in here as well.}
+    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
+    and with ECC-ions and ECC-lipids model with ECC-ions. }
+  \todo{Change the figure so that it contains a membrane background} \\
+  \todo{What are the concentrations (calculated as in NMRlipids II) in this figure?} \\
+  \todo{These should be also compared to the results with Åqvist ions, which are shown
+    in NMRlipids II. NMRlipids II have results with 350mM and 1000mM. If we have the distributions with
+    the same concentrations Åqvist results could be added also in this figure. If not, the
+  comparison has to be made some other way.}
 \end{figure}
 
-Good agreement of ECC corrected model with experiments encourages us to analyse the
+Good agreement of ECC-lipids model with experiments encourages us to analyse the
 binding details from MD simulations. Direct analysis of contacts between ions and
-lipids from simulations suggest that most common complex are ones with stoichiometry
-of 2~POPC:1~\ce{Ca^{2+}}.
-\todo{SAMULI: Details of this analysis have to be added.}
+lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
+have stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
 As shown in Fig. \ref{fig:cacl-bind} this is in agreement
@@
 experiments~\cite{altenbach84}.
+In addition to the ternary complexes, there also is a non-negligable probability
+of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
+Technical details of the analysis are in the SI.
+
+\todo{Put details of the cation-binding stoichiometry analysis to SI.} \\
+\todo{Update the binding isotherm figure with new simulations} \\
 \todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
-  be the only option. I will recheck and come back to this.}
-%This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to 
-%the binding isotherm that assumes 2~POPC per 1~\ce{Ca^{2+}}. 
+  be the only option. I will recheck and come back to this.
+  SAMULI: This is written in \cite{macdonald87}:
+  ''Ca2+binding to POPC bilayers over the whole concentration range can be best described in terms
+  of formation of a ternary complex involving complexation of two lipids to one calcium ion (Altenbach and Seelig, 1984).
+  The addition of a sodium competition term has not changed this conclusion. However, if Ca2+ concentrations
+  up to 100 mM are considered, the data can be equally well explained by a 1:1 binding mechanism (cf. Figure 7).
+  In contrast, the Ca2+ binding to POPC-POPG mixtures can be best described by assuming a 1:l stoichiometry regardless of the range of Ca2+concentrations.''
+  We might or might not want to discuss about this.} \\
 \todo{SAMULI: I would also analyze how much there is contact between ions and different
-parts of the lipid (phosphase, carbonyl, etc.).}
-In addition to the ternary complexes, there also is a non-negligable probability
-of one \ce{Ca^{2+}} cross-bridging three POPC molecules
+parts of the lipid (phosphase, carbonyl, etc.).} \\
 \todo{SAMULI: I think we should quantify this, i.e. how much there are these.
   Maybe also the other possible complexes? Maybe also the correlation betweem complexes
-  and binding cites, if it is not too much work.}.
-
+  and binding sites, if it is not too much work.
+  JOE: This looks like a careful work for the next paper to me. 
+  I'd only add a relatively simple analysis of binding sites and probably the propensity of 1-2-3 membered clusters.
+  SAMULI: I agree that we should not use too much time on this now.}. \\
 \todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},
@@
   \caption{\label{fig:cacl-bind}
-    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits the simulation data nicely.}
+    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .}
 \end{figure}
@@
 
-%Role of water model: we use OPC3 (current best), it would be worth giving an estimate how results 
-%change when we use say SPCE or even TIP3p at least in the SI (so that the reader knows what errors 
-%to expect comming from these sub-optimal models). 
-%In addition, there is protein force field Amber15-FB, which uses water close to OPC3, TIP3pFB \todoii{REFs}{add references here}.
-%On the other hand, it might be the best if we used a model that doesn't have the dielectric constant 
-%from nuclei 78, but rather 44 -- TIP4p2005 is the closest. \cite{OPC_paper, ForceBalance_paper}
-
-%questionable charge distribution from RESP -- a leeway for further tweaks of the FF.
-%It is not obvious that RESP charges provide the best description, especially due to its non-uniqe solution. 
-%Shall we solve RESP fitting with the constraint of full charges and then scale down, or shall we rather 
-%solve the fitting with a scaled total charge target?
-
 \section{Conclusions}
-We show that the Na+ and Ca2+ binding in phospholipid bilayers can
-be accurately described with MD simulation models, where electronic
-polarization is effectively included by using ECC correction.
-This is a significant advantage to the other available lipid models,
-which all overestimate specific cation binding affinities.  
-The proposed model reproduces the lipid bilayer structural details
-with similar accuracy as the other state of the art lipid models.
-The correction is applied here on Lipid14 POPC model \cite{dickson14},
-but we expect that the correction can be generalized also for other lipids
-and force fields.
-%The models were optimized to represent correct membrane structure, 
-%and they were validated with the use of electrometer concept \cite{seelig87,Altenbach85,Altenbach84}.
+We show that the \ce{Na^+} and \ce{Ca^2+} binding in phospholipid bilayers can
+be accurately described with classical MD simulation models, where electronic
+polarization is effectively included by using electronic continuum correction.
+This is a significant improvement over other available lipid models,
+which all overestimate specific cation binding affinities \cite{catte16}.  
+The newly proposed model, which we denote as ''ECC-lipids~17'', 
+reproduces the lipid bilayer structural details
+with similar accuracy as other state of the art lipid models.
+ECC-lipids~17 exhibit accurate headgroup order parameter response to
+bound cations, monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
+and divalent \ce{Ca^2+}.
+Several water models 
+(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
+were used to exemplify the transferability of 
+the parameters of the new ECC-lipids~17 force field.
 
@@
 complex.
-%This tells us on the subtle effects of calcium in phospholipid membranes 
-%possibly leading to a better understanding of their physical properties (elasticity etc.). 
-%\todo{Improve this concluding-discussion with actual insights.}
 
-This will be a foundation stone of a new open-collaboration project NPRlipid 6 in nmrlipids.blogspot.fi....
+The electronic continuum correction is applied here on Lipid14 POPC model \cite{dickson14},
+but we expect that the correction can be generalized also for other lipids
+and force fields.
 
-% Tables may be be put in the text as floats.
-% Here is an example of the general form of a table:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
-% \begin{tabular}{} command.
-%
-% \begin{table}
-% \caption{\label{} }
-% \begin{tabular}{}
-% \end{tabular}
-% \end{table}
+This work will serve as a foundation stone of a 
+new open-collaboration project NMRlipids~VI in nmrlipids.blogspot.fi.
 

### 2017-07-27 Manuscript/proto-manuscript.tex
MS: wrote section Simulation details

--- 2017-07-27 Manuscript/proto-manuscript.tex
@@
 
-\todo{To be written}
+The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer and an aqueous salt solutions of varying concentrations.
+Water molecules were described by OPC3 model \cite{Izadi16}.
+In order to test transferability of the newly developed ECC-Lipids model, 
+we also performed several additional simulations with water models 
+OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}
+presented in Supporting Information (SI). 
+We used ECC-ions model for ions. \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} % cite order: [Ca_2s, Na_s, Cl_2s]
+Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
+The simulation settings used in this work are 
+are summarized in Table \ref{tbl:mdpar}, 
+and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}. 
+Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+
+\begin{table}[btp]
+  \caption{Simulation parameters}
+  \label{tbl:mdpar}
+  \begin{tabular}{ll}
+    simulation property & parameter   \\
+    \hline
+    time-step           & 2~fs         \\
+    equilibration time  & 100~ns  \\
+    simulation time     & 200~ns  \\
+    temperature         & 313~K       \\
+    thermostat          & v-rescale  \cite{bussi07}   \\
+    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
+    long-range electrostatics & PME  \cite{darden93}  \\
+    cut-off scheme      & Verlet \cite{Pall13}      \\
+    Coulomb and VdW cut-off & 1.0~nm \\
+    constraints         & LINCS, only Hydrogen atoms \cite{hess97} \\
+    constraints for water & SETTLE  \cite{miyamoto92} \\
+    \hline
+  \end{tabular}
+\end{table}
 
 \subsubsection{Simulations with cationic surfactants}
-Automated topoly builder \cite{malde11} was first used to create pdb structure of
+Automated topology builder \cite{malde11} was first used to create the structure of
 dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
-The AmberTools \cite{amber} was then used to generate the Amber force field
-parameters from the pdb file. These were converted to Gromacs format by using
+The code AmberTools \cite{amber} was then used to generate the Amber-type force field
+parameters. The parameters were converted to the Gromacs format by using
 acpype tool \cite{acpype}. The partial charges were then manually modified
-to approximately correspond the corresponding segments in Lipid14 \cite{dickson14}.
-The surfactants were randomly mized with lipids to form bilayer structures with
-mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfctant in POPC bilayer.
-Simulation parameters and force field for POPC was already used and
-described previously \cite{catte16} and were downloaded from \cite{lipid14POPC0mMNaClfiles}.
-The systems were simulated 100-200~ns at 313K and the first 20~ns of the trajectories
-were excluded from the analysis as an equilibration period.
-The simulations were ran with Gromacs 5~\cite{abraham15}. 
-Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
+The surfactants were randomly placed among the lipids to form bilayer structures with
+mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
 

### 2017-07-27 Manuscript/proto-manuscript.tex
MS: updated sections Results and discussion and Conclusions

@@
 placing all particles into a homogeneous dielectric continuum 
-with a dielectric constant $\epsilon _{el}$. 
+with a dielectric constant $\epsilon _{el}$, 
 which is the electronic part of the dielectric constant of 
@@
 Water molecules were described by OPC3 model \cite{Izadi16}.
-In order to test transferability of the newly developed ECC-Lipids model, 
+In order to test transferability of the newly developed ECC-lipids model, 
 we also performed several additional simulations with water models 
@@
 presented in Supporting Information (SI). 
-We used ECC-ions model for ions. \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} % cite order: [Ca_2s, Na_s, Cl_2s]
-Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
-The simulation settings used in this work are 
-are summarized in Table \ref{tbl:mdpar}, 
-and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}. 
-Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
-
-\begin{table}[btp]
-  \caption{Simulation parameters}
-  \label{tbl:mdpar}
-  \begin{tabular}{ll}
-    simulation property & parameter   \\
-    \hline
-    time-step           & 2~fs         \\
-    equilibration time  & 100~ns  \\
-    simulation time     & 200~ns  \\
-    temperature         & 313~K       \\
-    thermostat          & v-rescale  \cite{bussi07}   \\
-    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
-    long-range electrostatics & PME  \cite{darden93}  \\
-    cut-off scheme      & Verlet \cite{Pall13}      \\
-    Coulomb and VdW cut-off & 1.0~nm \\
-    constraints         & LINCS, only Hydrogen atoms \cite{hess97} \\
-    constraints for water & SETTLE  \cite{miyamoto92} \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\subsubsection{Simulations with cationic surfactants}
-Automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
+We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+Simulations with Lipid14 use ion model by Dang \cite{smith94,chang1999,dang2006}. 
 The code AmberTools \cite{amber} was then used to generate the Amber-type force field
@@
   \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-%  \caption{\label{fig:form-f}
-%    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models.  }
-%\end{figure}
-%
-%\begin{figure}[]
-%  \centering
   \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
   \caption{\label{simVSexpNOions}
-    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 and ECCLipids17 models. 
-    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and EECLipid17 models
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
+    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
     compared with experimental order parameters from \cite{ferreira13}.}
   \todo{Add acyl chain order parameters, POPC chemical structure}
-  \todo{Should we add the results from original lipid14?}
 \end{figure}
-%Area per molecules extracted from MD simulations and SPD model fitted to scattering data \todo{check that this the case for the used values}
-%are shown in Table \ref{tab:apls}.
-\todo{finalize figure (NMR headgr. OPs + SAXS, continue in the discussion.}
+
 \begin{table}
-  \caption{Area per lipid from different models for POPC without ions\label{tab:apls} }
-  \begin{tabular}{c c c}
-    model          & A (Å$^2$)   & Temperature [K] \\
+  \caption{Area per lipid (APL) from different models of POPC without ions\label{tab:apls} }
+  \begin{tabular}{l|c c}
+    model          & APL (\AA$^2$)   & Temperature [K] \\
     \hline
-    Lipid14 (literature)  & 65.6$\pm$ 0.5  &  303 \\
-    Lipid14ecc0.80+sigma0.875 &        &  313    \\
-    GMX small patch           & 64.9   &         \\
-    GMX 4xbig patch           & 65.5   &         \\
-    oMM small patch           & 63.65  &         \\
-    oMM 4xbig patch           & 63.7   &         \\
+    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
     \hline
-    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.}  & 62.7  &  293    \\
-    experiment  & 64.3  &  303    \\
+    ECC-lipids &        &  \\
+    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
+    ($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
+    \hline
+    (6.4 nm)$^2$, 128 lipids patch, OPC3           & 64.2   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, SPC/E          & 65.1   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, OPC            & 64.4   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, TIP4p/2005     & 66.8   &  313       \\
+    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
+    %oMM small patch           & 63.65  &         \\
+    %oMM 4xbig patch           & 63.7   &         \\
+    \hline
+    %experiment   & 62.7  &  293    \\
+    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
     experiment  & 67.3  &  323    \\
-    experiment  & 68.1  &  333    \\
-    experiment POPE  & 56.6 &  303    \\
+    %experiment  & 68.1  &  333    \\
+    %experiment POPE  & 56.6 &  303    \\
     \hline
@@
   \caption{\label{fig:cacl-dens}
-    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Aquist parameters
-    and with ECC ions and ECCLipids17 with ECC ions. }
-  \todo{SAMULI: We should add the location of bilayer here somehow.
-    In NMRlipids II the location of phosphate was shown with green vertical line.}
-  \todo{SAMULI: I would add Aqvist data from NMRlipids II in here as well.}
+    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
+    and with ECC-ions and ECC-lipids model with ECC-ions. }
+  \todo{Change the figure so that it contains a membrane background}
 \end{figure}
 
-Good agreement of ECC corrected model with experiments encourages us to analyse the
+Good agreement of ECC-lipids model with experiments encourages us to analyse the
 binding details from MD simulations. Direct analysis of contacts between ions and
-lipids from simulations suggest that most common complex are ones with stoichiometry
-of 2~POPC:1~\ce{Ca^{2+}}.
-\todo{SAMULI: Details of this analysis have to be added.}
+lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
+have stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
 As shown in Fig. \ref{fig:cacl-bind} this is in agreement
@@
 experiments~\cite{altenbach84}.
+In addition to the ternary complexes, there also is a non-negligable probability
+of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
+Technical details of the analysis are in the SI.
+
+\todo{Put details of the cation-binding stoichiometry analysis to SI.}
+\todo{Update the binding isotherm figure with new simulations}
 \todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
   be the only option. I will recheck and come back to this.}
-%This cannot be observed in experiments without atomistic detail as it makes only a small perturabation to 
-%the binding isotherm that assumes 2~POPC per 1~\ce{Ca^{2+}}. 
 \todo{SAMULI: I would also analyze how much there is contact between ions and different
 parts of the lipid (phosphase, carbonyl, etc.).}
-In addition to the ternary complexes, there also is a non-negligable probability
-of one \ce{Ca^{2+}} cross-bridging three POPC molecules
 \todo{SAMULI: I think we should quantify this, i.e. how much there are these.
   Maybe also the other possible complexes? Maybe also the correlation betweem complexes
-  and binding cites, if it is not too much work.}.
-
+  and binding sites, if it is not too much work.
+      JOE: This looks like a careful work for the next paper to me. 
+           I'd only add a relatively simple analysis of binding sites and probably the propensity of 1-2-3 membered clusters.}. 
 \todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},

### 2017-07-28 Manuscript/proto-manuscript.tex
Introduction updated

@@
 
+%Classical molecular dynamics simulations have potential to solve the ion binding details
+%in lipid bilayers and reveal its relevance to various biological
+%problems \cite{bockmann03,bockmann04,??}. However, the available
+%molecular dynamics simulation models have a strong tendency to overestimate
+%cation binding on zwitterionic bilayers and none of them can reproduce the
+%experimental data with the accuracy required for the interpretation of
+%experiments \cite{catte16}.
+
+The proposed MD simulation model with ECC-corrected lipids and ions
+reproduce the experimental NMR data of Calcium binding in PC lipid bilayers
+and mainly supports the proposed ternary complex model (2 Ca2+ bind in
+on POPC) \cite{akutsu81,altenbach84}. However, non-negligible contribution
+of one Ca2+ cross bridging three POPC molecules is also observed.
+The overestimated specific cation binding observed in current lipid models \cite{catte16}
+may lead to articifially positively charged memranes and significant artefacts in
+MD simualtions. Thus, the proposed approach to improve the accuracy of cation
+binding in current MD models is highly useful for future MD simulations
+with physiological salt conditions.
+
+%re is a vast demand for classical molecular dynamics simulation
+%model which correctly reproduces cation binding in lipid bilayers.
+
+%The lack of electronic polarizability from the classical MD simulation
+%models is a potential source of artefact, which could lead to overbinding
+%of cations. 
+%Almost all lipid bilayer simulations reported in the literature 
+%are done with force fields that represent interactions between atoms
+%with pairwise additive empirical potential and exclude the electronic 
+%polarizability \cite{SOME_review, FF_papers}.
+%The issue has been considered higly relevant since the early days of
+%lipid bilayer simulations and pioneering simulation studies scaled
+%the partial charges to effectively include electronic
+%polarizability \cite{jonsson86,egberts94}.
+
 \section{Methods}
@@
 \subsection{Electronic continuum correction for lipid bilayers}
+The explicit inclusion of
+electronic polarizability has been, however, turned to be 
+practically complicated and implementations for lipids are rare \cite{chowdhary13}.
+
 According to the electronic continuum correction (ECC)\cite{leontyev11}, electronic

### 2017-07-28 Manuscript/proto-manuscript.tex
Methods partly updated

@@
 
-%The lack of electronic polarizability from the classical MD simulation
-%models is a potential source of artefact, which could lead to overbinding
+\section{Methods}
+
+\subsection{Electronic continuum correction for lipid bilayers}
+The lack of electronic polarizability from the classical MD simulation
+models has been considered higly relevant issue since the early days of
+lipid bilayer simulations and pioneering simulation studies scaled
+the partial charges by half to effectively include electronic
+polarizability \cite{jonsson86,egberts94}. Also approaches that
+explicitly include electronic polarizability has been introduced, but
+this has turned out to be practically complicated \cite{chowdhary13}.
+In this work we effectively include electronic polarizability in lipid
+bilayer simulations by using the electronic continuum correction
+(ECC)~\cite{leontyev11}. The approach is more sophisticated version of
+early studies with scaling to half was used \cite{jonsson86,egberts94},
+but significantly more simple than the explicit inclusion of electronic
+polarizability \cite{chowdhary13}.
+
+%is a potential source of artefact, which could lead to overbinding
 %of cations. 
@@
 %polarizability \cite{SOME_review, FF_papers}.
-%The issue has been considered higly relevant since the early days of
-%lipid bilayer simulations and pioneering simulation studies scaled
-%the partial charges to effectively include electronic
-%polarizability \cite{jonsson86,egberts94}.
 
-\section{Methods}
-
-\subsection{Electronic continuum correction for lipid bilayers}
-The explicit inclusion of
-electronic polarizability has been, however, turned to be 
-practically complicated and implementations for lipids are rare \cite{chowdhary13}.
-
-According to the electronic continuum correction (ECC)\cite{leontyev11}, electronic
+According to ECC, electronic
 polarizability can be included in classical MD simulations by
@@
 
-Here we apply the electronic continuum correction to lipid bilayers to accurately 
-describe the lipid headgroup response to Na+ and Ca2+ concentrations \cite{catte16}. 
-We use the Lipid14~\cite{dickson14} force field parameters as a starting point,
-because they give the most realistic headgroup response with added cations 
-and relatively realistic glycerol backbone and headgroup structures
-when compared with other state of the art lipid models \cite{botan15,catte16}.
-The partial atomic charges in Lipid14 were derived by fitting the electrostatic 
+The main goal of this work is to apply ECC correction to accurately 
+describe the lipid headgroup response to Na$^+$ and Ca$^{2+}$ concentrations
+in MD simulation models when compared with NMR data \cite{catte16}. 
+The Lipid14~\cite{dickson14} force field parameters were used as a starting point,
+because their response to bound ions was most realistic against NMR data
+(see Fig.~5 in Ref.~\citenum{catte16}).
+Also glycerol backbone and headgroup structures in Lipid14 model were
+relatively realistic when compared with other state of the art lipid models \cite{botan15}.
+
+The partial atomic charges in original Lipid14 parametrization were derived by fitting the electrostatic 
 potential to its model quantum chemistry representation (RESP\cite{RESP_paper}) in vacuum.
@@
 by increasing the scaling factor $f_q$ to $0.8$. 
-\footnote{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom types being around 0.78--0.80.
-Even though the methods are almost identical, authors of Lipid 14 find lower partial charges in vacuum than here -- so I prefer the higher value.
-As the choice of charges is arbitrary anyway, I use 0.80 as an approximate round value.
-The use of 0.78---.79 might be more appropriate, though.}
+\footnote{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
+  types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
+  partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
+  I use 0.80 as an approximate round value. The use of 0.78---.79 might be more appropriate, though.}
 

### 2017-08-01 scripts/calcOrderParameters.py
Updated calcOrderParameters.py to contain Students T-Test - prints the p-value backwards compatibly.

@@ -1,9 +1,9 @@
 #!/usr/bin/env python
 """
- calculation of order parameters 
+ calculation of order parameters
  from a MD trajectory
  useful for example for lipid bilayers
- 
+
  meant for use with NMRlipids projects
 
 ------------------------------------------------------------
@@ -20,8 +20,9 @@
 
 import MDAnalysis as mda
 import numpy as np
+import scipy.stats
 import math
-import os, sys
+import os  #, sys
 from optparse import OptionParser
 
 
@@ -40,12 +41,13 @@ class OrderParameter:
        - name of the OP
        - residue name
        - involved atoms (exactly 2)
-       + extra: mean and std.dev. of the OP (when reading-in an already calculated result)
+       + extra: mean, std.dev. & err. estimation (p-value from Student's T-test)
+                of the OP (when reading-in an already calculated result)
     """
     def __init__(self, name, resname, atom_A_name, atom_B_name, *args):
         """
         Initialization of an instance of this class.
-        
+
         it doesn't matter which atom comes first,
         atom A or B, for OP calculation.
         """
@@ -61,15 +63,21 @@ class OrderParameter:
                 if not field.strip():
                     raise RuntimeError, "provided name >> {} << is empty! \n \
                     Cannot use empty names for atoms and OP definitions.".format(field)
-        # extra optional arguments allow setting avg,std values -- suitable for reading-in results of this script
+        # extra optional arguments allow setting avg,std,pval values -- suitable for reading-in results of this script
         if len(args) == 0:
-            self.avg = None
-            self.std = None
+            self.avg = None     # average/mean value
+            self.std = None     # standard deviation of the mean
+            self.pval = None    # p-value as an error estimate of the mean
         elif len(args) == 2:
             self.avg = args[0]
             self.std = args[1]
+            self.pval = None
+        elif len(args) == 3:
+            self.avg = args[0]
+            self.std = args[1]
+            self.pval = args[2]
         else:
-            raise UserWarning, "Number of optional positional arguments is {len}, not 2 or 0. Args: {args}\nWrong file format?".format(len=len(args), args=args)
+            raise UserWarning, "Number of optional positional arguments is {len}, not 3, 2 or 0. Args: {args}\nWrong file format?".format(len=len(args), args=args)
         self.traj = []  # for storing OPs
 
 
@@ -153,7 +161,7 @@ def parse_op_input(fname):
     fname : string
[…]

### 2017-08-01 scripts/calcOrderParameters.py
Updated calcOrderParameter.py to use Bayesian statistics to give mean, standard deviation and confidence interval (represented with symmetric error estimate in the file output). T-test was deleted. Should be backwards-compatible with previous outputs and definitoins.

@@ -41,7 +41,7 @@ class OrderParameter:
        - name of the OP
        - residue name
        - involved atoms (exactly 2)
-       + extra: mean, std.dev. & err. estimation (p-value from Student's T-test)
+       + extra: mean, std.dev. & err. estimation (from Bayesian statistics)
                 of the OP (when reading-in an already calculated result)
     """
     def __init__(self, name, resname, atom_A_name, atom_B_name, *args):
@@ -55,6 +55,10 @@ class OrderParameter:
         self.resname = resname       # name of residue atoms are in
         self.atAname = atom_A_name
         self.atBname = atom_B_name
+        # variables for Bayesian statistics
+        self.mean = None
+        self.var = None
+        self.confidence_int = None
         for field in self.__dict__:
             if not isinstance(field, str):
                 raise UserWarning, "provided name >> {} << is not a string! \n \
@@ -63,19 +67,19 @@ class OrderParameter:
                 if not field.strip():
                     raise RuntimeError, "provided name >> {} << is empty! \n \
                     Cannot use empty names for atoms and OP definitions.".format(field)
-        # extra optional arguments allow setting avg,std,pval values -- suitable for reading-in results of this script
+        # extra optional arguments allow setting avg,std,errest values -- suitable for reading-in results of this script
         if len(args) == 0:
             self.avg = None     # average/mean value
             self.std = None     # standard deviation of the mean
-            self.pval = None    # p-value as an error estimate of the mean
+            self.errest = None  # error estimate of the mean
         elif len(args) == 2:
             self.avg = args[0]
             self.std = args[1]
-            self.pval = None
+            self.errest = None
         elif len(args) == 3:
             self.avg = args[0]
             self.std = args[1]
-            self.pval = args[2]
+            self.errest = args[2]
         else:
             raise UserWarning, "Number of optional positional arguments is {len}, not 3, 2 or 0. Args: {args}\nWrong file format?".format(len=len(args), args=args)
         self.traj = []  # for storing OPs
@@ -97,13 +101,12 @@ class OrderParameter:
         return S
 
 
-    @property
-    def get_avg_std_OP(self):
+    def get_avg_std_OP(self, alpha_confidence=0.95):
[…]

### 2017-08-02 scripts/rename_residue_lipid14_to_PALM-POPC-OLE.py
Order parameter calculation updated: new simple script for Lipid14 POPC residue renamnig to PALM-POPC-OLE, updated OP-calculating script and definitions so that it calculates also tail OPs.

@@ -1,0 +1,52 @@
+#
+# simple python script for conversion of 
+# Lipid14 POPC residue names with duplicate atom names 
+# to a split-residue convention: 
+#   POPC will be split into 3 virtual residues with names
+#   PALM - POPC - OLE
+#   so that duplicate atom names can be distinquished at least
+#   by a different residue name
+#
+# script assumes the particular ordering as above with 
+# the head group segment beginning with O12 and ending with O22
+# the rest (begin/end) being the tails PALM/OLE.
+
+import MDAnalysis as mda
+from optparse import OptionParser
+
+# help message is automatically provided
+# type=string, action=store is default
+parser = OptionParser()
+parser.add_option('-i', '--inp',  dest='inp_fname',  help='input MDanalysis topology file name', default="last_frame_nonwat.gro")
+parser.add_option('-o', '--out',  dest='out_fname',  help='output MDanalysis topology file name', default="last_frame_nonwat.gro")
+opts, args = parser.parse_args()
+
+# conf_init.gro should always be present in the sim dir
+u = mda.Universe(opts.inp_fname)
+
+# find some POPC residue
+some_popc_res = None
+try:
+    for r in u.residues:
+        if r.name == "POPC":
+            some_popc_res = r
+            break
+except:
+    print "Couldn't find a POPC residue -- already renamed to PALM-POPC-OLE?"
+
+if some_popc_res != None:
+    # find the indices of the first/last head group atoms
+    headgr_begin_index = some_popc_res.names.tolist().index("O12")
+    headgr_end_index   = some_popc_res.names.tolist().index("O22")
+
+    # resname the resnames of PALM and OLE segments
+    for r in u.residues:
+        if r.name == "POPC":
+            for a in r.atoms[:headgr_begin_index]:
+                a.resname = "PALM"
+            for a in r.atoms[headgr_end_index+1:]:
+                a.resname = "OLE"
+
+    # write the new topology out
+    u.SYSTEM.write(opts.out_fname)
+

### 2017-08-02 Manuscript/proto-manuscript.tex
Update methods section

@@
 
-The partial atomic charges in original Lipid14 parametrization were derived by fitting the electrostatic 
-potential to its model quantum chemistry representation (RESP\cite{RESP_paper}) in vacuum.
-If the charges are obtained by using RESP in an implicit solvent, 
-they vary the most for the polar moieties \cite{maciejewski14}. 
-In order to implicitly capture solvent induced polarization 
-in models with fixed charges, 
-we shall use the average of partial charges 
-from both environments (i.e. vaccum and solvent),
-so called implicitly polarized charges (IPolQ) \cite{ipolq2013}. 
-By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
-we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
-by increasing the scaling factor $f_q$ to $0.8$. 
+While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
+to improve model performance against scattering data \cite{kohagen14,kohagen16,??},
+it is not clear if the same factor should be used for lipid headgroups \todo{, because ??}.
+Here we estimated the correct scaling factor for Lipid14 model by using
+so called implicitly polarized charges (IPolQ) \cite{ipolq2013},
+which are the average of partial charges given by RESP~\cite{RESP_paper} calculations in
+vacuum and in implicit solvent.
+%In order to implicitly capture solvent induced polarization 
+%in models with fixed charges, 
+%we shall use the average of partial charges 
+%from both environments (i.e. vaccum and solvent),
+%so called implicitly polarized charges (IPolQ) \cite{ipolq2013}. 
+The partial charges in original Lipid14 parametrization were derived by fitting the electrostatic 
+potential to its model quantum chemistry representation (RESP\cite{RESP_paper}) in vacuum,
+while charges from calculations in an implicit solvent vary most for the polar moieties \cite{maciejewski14}. 
+Here we estimate the appropriate scaling factor for lipid hedgroups and glycerol backbone
+by comparing IPolQ charges (average of RESP calculation in vacuum and in implicit solvent) to
+the RESP results in vacuum. By using the results reported in \cite{maciejewski14} we conclude
+that the IPolQ charges can be achived from vacuum calculations by scaling roughly with the
+factor of $f_q=0.8$, which is the used to derive ECC-corrected Lipid14 model.
+\todo{SAMULI: please check that this is correctly descrived.}
+%By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
+%we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
+%by increasing the scaling factor $f_q$ to $0.8$. 
 \footnote{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
@@
 
-Hydrocarbon chains in Lipid14 and other lipids models are highly optimized and
+Here we apply ECC correction with the proposed scaling factor only to the headgroup,
+glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
+are expected to have the largest contribution to the cation binding.
+The hydrocarbon chain parameters are not modified, because they are
+highly optimized in Lipid14 and other lipids models and
 give generally a good description for hydrophobic part of lipid bilayers
@@
 
+\subsection{Comparison of lipid bilayer structure to experimental data}
+Lipid bilayer structure without ions was validated against NMR experiments
+by order parameters for C-H bonds and x-ray scattering experiments by using
+form factors. The former validates the structures sampled by individual
+lipid molecules in simulations with segmental resolution, while latter
+validates the dimensions of the lipid bilayer (thickness and area per
+molecule)~\cite{ollila16}. Form factor was calculated from simulations
+as 
+\begin{equation}
+  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
+\end{equation}
+where $f_\alpha(q_z)$ is the density of atomic scattering length, 
+$\rho_s$ is the density of solvent scattering lenght,
+$n_\alpha (z)$ is the number density of atom $\alpha$ and
+$z$ is the distance from the membrane centre along its normal 
+spanning the membrane with thickness $D$.
+
 \subsection{Simulation details}
 
-The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer and an aqueous salt solutions of varying concentrations.
+\subsubsection{Simulations with aqueous ions}
+The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
+(POPC) bilayer and an aqueous salt solutions of varying concentrations.
 Water molecules were described by OPC3 model \cite{Izadi16}.
@@
 Simulations with Lipid14 use ion model by Dang \cite{smith94,chang1999,dang2006}. 
+\todo{Table about simulations details and some related text is deleted in commit
+  https://github.com/ohsOllila/NMRlipids\_VI-NewIonModel/commit/6526079062ca322e90f8ea0654597394b023041e
+  Maybe this is done accidentally, or is there some other plan with this?
+}
+
+\subsubsection{Simulations with cationic surfactants}
+Automated topology builder \cite{malde11} was first used to create the structure of
+dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
 The code AmberTools \cite{amber} was then used to generate the Amber-type force field
@@
 mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
+These systems were ran with original Lipid14 POPC model and with ECC-corrected model.
+\todo{Describe the changes in surfactant model used with ECC-corrected lipids.}
 

### 2017-08-02 Manuscript/proto-manuscript.tex
New IPython-notebook script for plotting all order parameters. Added recent simulation results to Data + updated OP-change figure.

--- 2017-07-28 Manuscript/proto-manuscript.tex
@@
 like in signal propagation in neurons and vesicle fusion.
-Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
-to understand molecular level details of specific cation interactions with
-cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
-and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} \todo{We need to select references for these}.
-While relative binding affinity of different ions is agreed to follow Hoffmeister
-serie, the molecular details of binding and binding energetics are
-not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
-%General conclusion from experimental
-%studies has been that multivalent ions and lithium have weak specific binding in
-%phopholipid bilayers, while other monovalent ions do not essentially
-%bind 
-%The presence of anoinic lipids, like PS or PG,
-%increase the concentration close to the bilayer and thus the amount of bound
-%ions, but do not affect the specific binding constant \cite{seelig90}.
-Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
-and infrared scpectroscopy, give accurate information about ion binding
-in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
-Molecular level interpretation of the results, however, requires assumptions
-about binding model and is often not fully conclusive \cite{??}.
-On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
-have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
+Cation interactions with lipid bilayers serving as a model for cellular
+membranes are thus widely studied with experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
+and theoretical methods \cite{Berkowitz12,??}. General conclusion from experimental
+studies has been that multivalent ions and lithium have weak specific binding in
+phopholipid bilayers, while other monovalent ions do not essentially
+bind \cite{cevc90,tocanne90,seelig90,binder02}. The presence of anoinic lipids, like PS or PG,
+increase the concentration close to the bilayer and thus the amount of bound
+ions, but do not affect the specific binding constant \cite{seelig90}.
+The binding details, like binding sites and stoichiometry are not yet fully
+resolved but interpretation of NMR and scattering experiments suggest that one
+Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
+phospholipid molecules \cite{altenbach84}.
 
-%The binding details, like binding sites and stoichiometry are not yet fully
-%resolved but interpretation of NMR and scattering experiments suggest that one
-%Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
-%phospholipid molecules \cite{altenbach84}.
+Classical molecular dynamics simulations have potential to solve the ion binding details
+in lipid bilayers and reveal its relevance to various biological
+problems \cite{bockmann03,bockmann04,??}. However, the available
+molecular dynamics simulation models have a strong tendency to overestimate
+cation binding on zwitterionic bilayers and none of them can reproduce the
+experimental data with the accuracy required for the interpretation of
+experiments \cite{catte16}. The overestimated specific cation binding also
+makes zwitterionic bilayers effectively positively charged, which could
+potentially lead to significant artefacts in applications.
+Thus, there is a vast demand for classical molecular dynamics simulation
+model which correctly reproduces cation binding in lipid bilayers.
 
-In this work we show that the accuracy of Calcium binding
-on zwitterionic PC lipid bilayer can be significantly improved by including electronic polarizability
-by using effective continuum correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules
-in classical MD simulation model.
-This is essentially physically well justified version of partial charge scaling
-implemented in early days of lipid and surfactant simulations \cite{jonsson86,egberts94}. 
+The lack of electronic polarizability from the classical MD simulation
+models is a potential source of artefact, which could lead to overbinding
+of cations. 
+%Almost all lipid bilayer simulations reported in the literature 
+%are done with force fields that represent interactions between atoms
+%with pairwise additive empirical potential and exclude the electronic 
+%polarizability \cite{SOME_review, FF_papers}.
+The issue has been considered higly relevant since the early days of
+lipid bilayer simulations and pioneering simulation studies scaled
+the partial charges to effectively include electronic
+polarizability \cite{jonsson86,egberts94}. The explicit inclusion of
+electronic polarizability has been, however, turned to be 
+practically complicated and implementations for lipids are rare \cite{chowdhary13}.
+
+In this work we show that the cation overbinding in classical MD simulations
+can be corrected by including electronic polarizability by using effective continuum
+correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules.
+This is essentially physicalle well justified version of partial charge scaling
+implemented in early days of lipid and surfctant simulations \cite{jonsson86,egberts94}. 
 The approach has been previously shown in to improve bulk performace of
@@
 
-%Classical molecular dynamics simulations have potential to solve the ion binding details
-%in lipid bilayers and reveal its relevance to various biological
-%problems \cite{bockmann03,bockmann04,??}. However, the available
-%molecular dynamics simulation models have a strong tendency to overestimate
-%cation binding on zwitterionic bilayers and none of them can reproduce the
-%experimental data with the accuracy required for the interpretation of
-%experiments \cite{catte16}.
-
-The proposed MD simulation model with ECC-corrected lipids and ions
-reproduce the experimental NMR data of Calcium binding in PC lipid bilayers
-and mainly supports the proposed ternary complex model (2 Ca2+ bind in
-on POPC) \cite{akutsu81,altenbach84}. However, non-negligible contribution
-of one Ca2+ cross bridging three POPC molecules is also observed.
-The overestimated specific cation binding observed in current lipid models \cite{catte16}
-may lead to articifially positively charged memranes and significant artefacts in
-MD simualtions. Thus, the proposed approach to improve the accuracy of cation
-binding in current MD models is highly useful for future MD simulations
-with physiological salt conditions.
-
-%re is a vast demand for classical molecular dynamics simulation
-%model which correctly reproduces cation binding in lipid bilayers.
-
-%The lack of electronic polarizability from the classical MD simulation
-%models is a potential source of artefact, which could lead to overbinding
-%of cations. 
-%Almost all lipid bilayer simulations reported in the literature 
-%are done with force fields that represent interactions between atoms
-%with pairwise additive empirical potential and exclude the electronic 
-%polarizability \cite{SOME_review, FF_papers}.
-%The issue has been considered higly relevant since the early days of
-%lipid bilayer simulations and pioneering simulation studies scaled
-%the partial charges to effectively include electronic
-%polarizability \cite{jonsson86,egberts94}.
-
 \section{Methods}
@@
 \subsection{Electronic continuum correction for lipid bilayers}
-The explicit inclusion of
-electronic polarizability has been, however, turned to be 
-practically complicated and implementations for lipids are rare \cite{chowdhary13}.
-
 According to the electronic continuum correction (ECC)\cite{leontyev11}, electronic
@@
   \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17.png}
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17_q80_sig89.png}
   \caption{\label{simVSexpNOions}

### 2017-08-02 Manuscript/proto-manuscript.tex
Some comments added as todo's in Results and Discussion

--- 2017-08-02 Manuscript/proto-manuscript.tex
@@
     compared with experimental order parameters from \cite{ferreira13}.}
-  \todo{Add acyl chain order parameters, POPC chemical structure}
+  \todo{Add acyl chain order parameters, POPC chemical structure} \\
+  \todo{We should run a simulation for the system withtout ions with the
+    same (or close) temperature as the experimental data. The NMR data is at 300K
+    so that would be good. I think that this would be also close enough for the scattering data at 303K.
+  }
 \end{figure}
@@
     and with ECC-ions and ECC-lipids model with ECC-ions. }
-  \todo{Change the figure so that it contains a membrane background}
+  \todo{Change the figure so that it contains a membrane background} \\
+  \todo{What are the concentrations (calculated as in NMRlipids II) in this figure?} \\
+  \todo{These should be also compared to the results with Åqvist ions, which are shown
+    in NMRlipids II. NMRlipids II have results with 350mM and 1000mM. If we have the distributions with
+    the same concentrations Åqvist results could be added also in this figure. If not, the
+  comparison has to be made some other way.}
 \end{figure}

### 2017-08-03 Manuscript/proto-manuscript.tex
Some comments about stoichimetry issues added

@@
 
-\todo{Put details of the cation-binding stoichiometry analysis to SI.}
-\todo{Update the binding isotherm figure with new simulations}
+\todo{Put details of the cation-binding stoichiometry analysis to SI.} \\
+\todo{Update the binding isotherm figure with new simulations} \\
 \todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
-  be the only option. I will recheck and come back to this.}
+  be the only option. I will recheck and come back to this.
+  SAMULI: This is written in \cite{macdonald87}:
+  ''Ca2+binding to POPC bilayers over the whole concentration range can be best described in terms
+  of formation of a ternary complex involving complexation of two lipids to one calcium ion (Altenbach and Seelig, 1984).
+  The addition of a sodium competition term has not changed this conclusion. However, if Ca2+ concentrations
+  up to 100 mM are considered, the data can be equally well explained by a 1:1 binding mechanism (cf. Figure 7).
+  In contrast, the Ca2+ binding to POPC-POPG mixtures can be best described by assuming a 1:l stoichiometry regardless of the range of Ca2+concentrations.''
+  We might or might not want to discuss about this.} \\
 \todo{SAMULI: I would also analyze how much there is contact between ions and different
-parts of the lipid (phosphase, carbonyl, etc.).}
+parts of the lipid (phosphase, carbonyl, etc.).} \\
 \todo{SAMULI: I think we should quantify this, i.e. how much there are these.
@@
   JOE: This looks like a careful work for the next paper to me. 
-           I'd only add a relatively simple analysis of binding sites and probably the propensity of 1-2-3 membered clusters.}. 
+  I'd only add a relatively simple analysis of binding sites and probably the propensity of 1-2-3 membered clusters.
+  SAMULI: I agree that we should not use too much time on this now.}. \\
 \todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},

### 2017-08-04 Manuscript/proto-manuscript.tex
MS: updated and polished overall. Small updates to figures.

@@
 Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. 
-The new lipid model, "ECC-lipids", shows accurate binding affinity to sodium and calcium cations and headgroup order parameter response to bound charge. 
+The new lipid model, "ECC-lipids", shows accurate binding affinity to sodium and calcium cations and head group order parameter response to bound charge. 
 We also provide for the first time a realistic stochiometry of bound calcium cations to a POPC membrane, and their binding sites. 
@@
 like in signal propagation in neurons and vesicle fusion.
+\todo{JOE: list more introductory examples, where it has interesting effects.}
 Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
@@
 While relative binding affinity of different ions is agreed to follow Hoffmeister
-serie, the molecular details of binding and binding energetics are
+series, the molecular details of binding and binding energetics are
 not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
@@
 have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
+The overestimated specific cation binding observed in current lipid models \cite{catte16}
+may lead to articifially positively charged memranes and significant artefacts in
+MD simualtions (e.g.~\ce{Ca^2+} induced charge inversion of bacterial membranes~\cite{Luan2016}).
 
@@
 
-In this work we show that the accuracy of Calcium binding
-on zwitterionic PC lipid bilayer can be significantly improved by including electronic polarizability
-by using effective continuum correction (ECC)~\cite{leontyev11} for polar region of zwitterionic lipid molecules
-in classical MD simulation model.
-This is essentially physically well justified version of partial charge scaling
-implemented in early days of lipid and surfactant simulations \cite{jonsson86,egberts94}. 
-The approach has been previously shown in to improve bulk performace of
-ion models against neutron scattering data \cite{Jungwirth2015,kohagen14,kohagen16}.
-The better bulk behaviour was not, however, sufficient to correct binding in lipid
-bilayers \cite{catte16}. 
-%motivation for its application to zwitterionic lipids like POPC.
-
-%Classical molecular dynamics simulations have potential to solve the ion binding details
-%in lipid bilayers and reveal its relevance to various biological
-%problems \cite{bockmann03,bockmann04,??}. However, the available
-%molecular dynamics simulation models have a strong tendency to overestimate
-%cation binding on zwitterionic bilayers and none of them can reproduce the
-%experimental data with the accuracy required for the interpretation of
-%experiments \cite{catte16}.
-
-The proposed MD simulation model with ECC-corrected lipids and ions
-reproduce the experimental NMR data of Calcium binding in PC lipid bilayers
-and mainly supports the proposed ternary complex model (2 Ca2+ bind in
-on POPC) \cite{akutsu81,altenbach84}. However, non-negligible contribution
-of one Ca2+ cross bridging three POPC molecules is also observed.
-The overestimated specific cation binding observed in current lipid models \cite{catte16}
-may lead to articifially positively charged memranes and significant artefacts in
-MD simualtions. Thus, the proposed approach to improve the accuracy of cation
-binding in current MD models is highly useful for future MD simulations
-with physiological salt conditions.
-
-%re is a vast demand for classical molecular dynamics simulation
-%model which correctly reproduces cation binding in lipid bilayers.
+In this work we show that the accuracy of ion binding, especially Calcium,
+in zwitterionic PC lipid bilayer can be significantly improved 
+for classical MD simulation models of zwitterionic lipids
+by including electronic polarizability
+in the polar region of lipid molecules
+using the electronic continuum correction (ECC)~\cite{leontyev11}. 
+This approach has been previously shown to improve performace of
+ion models against neutron scattering data in concentrated solutions \cite{Jungwirth2015,kohagen14,kohagen16}
+and ab-initio simulations~\cite{Jungwirth2017}. 
+The realistic structure of ions in a bulk solvent was not, however, sufficient to correct binding in lipid
+bilayers~\cite{catte16}. 
+To fix it, we propose
+a classical fixed-charge MD simulation model of 
+1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) lipid 
+that accounts for electronic polarizability by using ECC. 
+We will validate such a model by comparing its lipid bilayer structue
+to x-ray scattering data and its head group structure to NMR experiments. 
+We will then quantify the accuracy of head group order parameter response to bound charge
+and compare it to the experimental NMR data of Sodium and Calcium binding in PC lipid bilayers. 
+In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
+This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
+The proposed POPC model, ECC-lipids, and the approach to obtain it
+is, hence, highly useful for future MD simulations with physiological salt conditions, 
+and for effects where membrane electrostatics play a major role.
 
@@
 \subsection{Electronic continuum correction for lipid bilayers}
-The lack of electronic polarizability from the classical MD simulation
+The lack of electronic polarizability in the classical MD simulation
 models has been considered higly relevant issue since the early days of
-lipid bilayer simulations and pioneering simulation studies scaled
-the partial charges by half to effectively include electronic
-polarizability \cite{jonsson86,egberts94}. Also approaches that
-explicitly include electronic polarizability has been introduced, but
+lipid bilayer simulations. Some pioneering simulation studies scaled
+the partial charges of atoms by one half to effectively include 
+polarizability \cite{jonsson86,egberts94}. Also, approaches that
+explicitly include polarizability has been introduced, but
 this has turned out to be practically complicated \cite{chowdhary13}.
@@
 bilayer simulations by using the electronic continuum correction
-(ECC)~\cite{leontyev11}. The approach is more sophisticated version of
-early studies with scaling to half was used \cite{jonsson86,egberts94},
+(ECC)~\cite{leontyev11}. 
+Technically, it is a similar approach to the phenomenological charge-scaling as applied in 
+the early studies where a scaling factor one half was used \cite{jonsson86,egberts94}. 
+In contrast to those studies, however, the concept of ECC is 
+physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11},
 but significantly more simple than the explicit inclusion of electronic
-polarizability \cite{chowdhary13}.
-
-%is a potential source of artefact, which could lead to overbinding
-%of cations. 
-%Almost all lipid bilayer simulations reported in the literature 
-%are done with force fields that represent interactions between atoms
-%with pairwise additive empirical potential and exclude the electronic 
-%polarizability \cite{SOME_review, FF_papers}.
+polarizability \cite{chowdhary13} both for derivation and application.
 
@@
 The main goal of this work is to apply ECC correction to accurately 
-describe the lipid headgroup response to Na$^+$ and Ca$^{2+}$ concentrations
+describe the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
 in MD simulation models when compared with NMR data \cite{catte16}. 
@@
 (see Fig.~5 in Ref.~\citenum{catte16}).
-Also glycerol backbone and headgroup structures in Lipid14 model were
+Also glycerol backbone and head group structures in Lipid14 model were
 relatively realistic when compared with other state of the art lipid models \cite{botan15}.
@@
 While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
-to improve model performance against scattering data \cite{kohagen14,kohagen16,??},
-it is not clear if the same factor should be used for lipid headgroups \todo{, because ??}.
-Here we estimated the correct scaling factor for Lipid14 model by using
-so called implicitly polarized charges (IPolQ) \cite{ipolq2013},
-which are the average of partial charges given by RESP~\cite{RESP_paper} calculations in
-vacuum and in implicit solvent.
-%In order to implicitly capture solvent induced polarization 
-%in models with fixed charges, 
-%we shall use the average of partial charges 
-%from both environments (i.e. vaccum and solvent),
-%so called implicitly polarized charges (IPolQ) \cite{ipolq2013}. 
-The partial charges in original Lipid14 parametrization were derived by fitting the electrostatic 
-potential to its model quantum chemistry representation (RESP\cite{RESP_paper}) in vacuum,
-while charges from calculations in an implicit solvent vary most for the polar moieties \cite{maciejewski14}. 
-Here we estimate the appropriate scaling factor for lipid hedgroups and glycerol backbone
-by comparing IPolQ charges (average of RESP calculation in vacuum and in implicit solvent) to
-the RESP results in vacuum. By using the results reported in \cite{maciejewski14} we conclude
-that the IPolQ charges can be achived from vacuum calculations by scaling roughly with the
-factor of $f_q=0.8$, which is the used to derive ECC-corrected Lipid14 model.
-\todo{SAMULI: please check that this is correctly descrived.}
+to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
+it is not clear if the same factor should be used for lipid head groups. 
+Unlike the total charge of cations and anions, atomic partial charges used in
+classical MD simulations are not experimental observables and 
+can in principle adopt arbitrary values. 
+Here we estimate the correct scaling factor for Lipid14 model by using
+so called ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013},
+which are the average of partial charges given by 
+restrained electrostatic potential fitting (RESP~\cite{RESP_paper}) 
+in vacuum and in solvent.
+Using the partial atomic charges obtained from RESP fitting in vacuum and implicit solvent reported in \cite{maciejewski14} and
+considering that the partial charges in original Lipid14 parametrization were derived using RESP~\cite{RESP_paper} in vacuum,
+we conclude that the charges in Lipid14 model shall be scaled roughly by the factor $f_q=0.8$. 
 %By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
@@
 %by increasing the scaling factor $f_q$ to $0.8$. 
-\footnote{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
-  types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
-  partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
-  I use 0.80 as an approximate round value. The use of 0.78---.79 might be more appropriate, though.}
+%-----\footnote----------
+  %{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
+  %types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
+  %partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
+  %I use 0.80 as an approximate round value, which accounts for the already lower charges in Lipid14. 
+  %The use of 0.78---.79 might be more appropriate, though.}
 
-Here we apply ECC correction with the proposed scaling factor only to the headgroup,
+Here we apply the ECC correction with the proposed scaling factor only to the head group,
 glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
@@
 The hydrocarbon chain parameters are not modified, because they are
-highly optimized in Lipid14 and other lipids models and
+already highly optimized in most lipids models including Lipid14 and
 give generally a good description for hydrophobic part of lipid bilayers
-in various conditions \cite{ollila16},
-in contrast to glycerol backbone and headgroup regions which require
-some improvement in all available lipid models \cite{botan15}. 
+in various conditions \cite{ollila16}.
+In contrast, all available lipid models require
+some improvements in glycerol backbone and head group regions \cite{botan15}. 
 %To minimize the detuning of the highly optimized hydrocarbon
@@
 The scaling of partial charges in the polar region was found to decrease
-area per molecule to ??\todo{find the value}, which is significantly smaller than the
-experimental value (\cite{}) and the original Lipid14 values (\cite{})\todo{Add the values}.  
-The decrease of area was found to arise from lower hydration of the lipid headgroup region,
+area per molecule to 60~\AA$^2$, which is significantly smaller than the
+experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment}) and the original Lipid14 values $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}..  
+The decrease of area was found to arise from a lower hydration of the lipid head group region,
 %\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
 %JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
-which can be explained by the increased solvation free energy due to the lower polarity
-of molecules with scaled charges. The solvation free energy can be decreased
-and hydration increased by reducing the radius of beads by changing the 
-the $\sigma$ parameters in Lennard-Jones potential for the selected atoms.
+which can be explained by the decreased solvation free energy due to the lower polarity
+of molecules with scaled charges. 
+The solvation free energy and hydration is sensitive to Van der Waals interaction parameters,
+hence, they can be increased by reducing the radius of the selected atoms by changing the 
+the $\sigma$ parameters in Lennard-Jones potential. 
 %This allows water molecules to approach closer to lipid atoms and
@@
 Here we reduce $\sigma$ with a scaling factor of $f_\sigma = 0.89$ for the same atoms for which
-charges was scaled. This increased the area per molecule back
-to a level close to experimental and original Lipid14 values. 
+charges were scaled. This has lead to an increase of the area per molecule back
+close to the experimental value (see Table~\ref{tab:apls}). 
 %\todo{Is there some justification for the below sentence?} 
@@
 %has to lay between $f_q$ and 1.
-
-\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept}
-Ion binding was compared between experiments and simulations by 
-using lipid headgroup order parameters and ''electrometer concept'' 
-introduced by Seelig et al. \cite{seelig87,catte16}.
-The concept is based on the experimental observation that the 
-order parameters of $\alpha$ and $\beta$ carbons in lipid headgroup
-(see Fig. \ref{??} 
-\todo{Figure with chemical structure and labeling to be added}) 
-are proportional to the amount of bound charge
-in lipid bilayer \cite{seelig87}.
-More recent analysis included also
-the order parameter signs and concluded that the order parameters  
-decrease with bound positive charge and increase with bound negative 
-charge \cite{ollila16,catte16}. The observations are rationalized 
-as a change of lipid headgroup dipole tilt to more vertical orientation
-with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
-Order parameters for C-H bonds can be accurately measured for each
-C-H bond of lipids by using H$^2$ NMR or $^{13}$C NMR techniques and
-calculated from MD simulations directly from definition
-\begin{equation}\label{OP}
-S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
-\end{equation}
-where $\theta$ is the angle between the bond and membrane
-normal and average is taken over all sampled configurations \cite{ollila16}.
-
-The change of the headgroup order parameters as a function of
-the amount of bound charge per lipid $X^\pm$ 
-can be written as~\cite{ferreira16}
-\begin{equation}\label{OPchangeEQ}
-\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) =\frac{4m_i }{3\chi}X^\pm,
-\end{equation}
-where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
-$i$ refers to either $\alpha$ or $\beta$ carbon,
-%the empirically observed linear relation %for the change in order parameters %with respect to a bilayer without bound charges              
-%\begin{equation}\label{electrometer_eq}                                                                                                   
-%S_{\rm{CH}}^{i}(X^\pm)=S_{\rm{CH}}^{i}(0) + \frac{4m_i}{3\chi} X^\pm,                                                                     
-%\end{equation}                                                                                                                            
-$m_i$ is an empirical constant depending on the valency and position of bound charge,
-and the value of the quadrupole coupling constant is $\chi \approx$\,167\,kHz.
-Combination of atomic absorption spectra, $^2$H NMR experiments
-and information about order parameter signs gave
-$m_\alpha=-20.5$  and $m_\beta=-10.0$ for Ca$^{2+}$ binding to POPC
-bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-
-Recent work published by the NMRlipids project showed that the concept works
-qualitatively also in simulations and can be used to compare ion binding
-affinitity between simulations and experiments~\cite{catte16}. However,
-it also turned out that also the sensitivity of the order parameters response
-to bound charge has to be quantified more accurately for quantitative comparison
-of binding affinity. In this work we first we first quantify the response of
-headgroup order parameters against experimental data from cationinc surfactants
-embedded in PC bilayer~\cite{scherer89}. Essentially all surfactants locate in
-lipid bilayers in these experiments, thus the amount of bound charge is known
-and can be plotted against order parameter change.
-
-%\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
-%If it varies too much from the nominal concentration, I may need to tweak the scaling 
-%factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
-%to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
-%moment to guide me on the possible finite-size errors and this conc-error.
-The used experimental data report order parameters as a function of
-equilibrium cation concentration in the bulk solvent \cite{akutsu81,altenbach84}.
-Such a condition is reached in simulations by adjusting
-the simulation box size to dimensions large enough
-that ion conceration reaches a clear plateau in the bulk solvent.
-The concentrations in the units of $\mathrm{mol/l}$ were then 
-determined as
-\begin{equation}
-  C_{eq}=\frac{C_{plateau}}{0.602},
-\end{equation}
-where plateau concentration is the number density in the units of $\mathrm{nm}^{-3}$.
-\todo{SAMULI: Once we have to final results, we can probably say that
-  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
-  }
-
-\subsection{Comparison of lipid bilayer structure to experimental data}
-Lipid bilayer structure without ions was validated against NMR experiments
-by order parameters for C-H bonds and x-ray scattering experiments by using
-form factors. The former validates the structures sampled by individual
-lipid molecules in simulations with segmental resolution, while latter
-validates the dimensions of the lipid bilayer (thickness and area per
-molecule)~\cite{ollila16}. Form factor was calculated from simulations
-as 
-\begin{equation}
-  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
-\end{equation}
-where $f_\alpha(q_z)$ is the density of atomic scattering length, 
-$\rho_s$ is the density of solvent scattering lenght,
-$n_\alpha (z)$ is the number density of atom $\alpha$ and
-$z$ is the distance from the membrane centre along its normal 
-spanning the membrane with thickness $D$.
-
-\subsection{Simulation details}
-
-\subsubsection{Simulations with aqueous ions}
-The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
-(POPC) bilayer and an aqueous salt solutions of varying concentrations.
-Water molecules were described by OPC3 model \cite{Izadi16}.
-In order to test transferability of the newly developed ECC-lipids model, 
-we also performed several additional simulations with water models 
-OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}
-presented in Supporting Information (SI). 
-We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Simulations with Lipid14 use ion model by Dang \cite{smith94,chang1999,dang2006}. 
-\todo{Table about simulations details and some related text is deleted in commit
-  https://github.com/ohsOllila/NMRlipids\_VI-NewIonModel/commit/6526079062ca322e90f8ea0654597394b023041e
-  Maybe this is done accidentally, or is there some other plan with this?
-}
-
-\subsubsection{Simulations with cationic surfactants}
-Automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
-The code AmberTools \cite{amber} was then used to generate the Amber-type force field
-parameters. The parameters were converted to the Gromacs format by using
-acpype tool \cite{acpype}. The partial charges were then manually modified
-to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
-The surfactants were randomly placed among the lipids to form bilayer structures with
-mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
-These systems were ran with original Lipid14 POPC model and with ECC-corrected model.
-\todo{Describe the changes in surfactant model used with ECC-corrected lipids.}
-
-\section{Results and Discussion}
-
-\subsection{Lipid bilayers without ions}
-The scattering form factors, NMR order parameters and area per lipid calculated  
-using the ECC-lipids model for POPC are compared to experiments and original
-Lipid14 results in Fig. \ref{simVSexpNOions} and in Table~\ref{tab:apls}. 
-%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
-%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
-The structural quality is comparable to the state of art lipid models available in literature \cite{ollila16},
-thus we conclude that the ECC-lipids model reproduces the lipid bilayer structure
-in liquid disordered phase with similar accuracy to other available models.
-
-\begin{figure}[]
-  \centering
-  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Headgr_OPs_exp-L14-ECCL17_q80_sig89.png}
-  \caption{\label{simVSexpNOions}
-    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
-    Headgroup and glycerol backbone order parameters from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
-    compared with experimental order parameters from \cite{ferreira13}.}
-  \todo{Add acyl chain order parameters, POPC chemical structure} \\
-  \todo{We should run a simulation for the system withtout ions with the
-    same (or close) temperature as the experimental data. The NMR data is at 300K
-    so that would be good. I think that this would be also close enough for the scattering data at 303K.
-  }
-\end{figure}
 
@@
 
+\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept}
+Ion binding was compared between experiments and simulations by 
+using lipid head group order parameters and ''electrometer concept'' 
+introduced by Seelig et al. \cite{seelig87,catte16}.
+The concept is based on the experimental observation that the 
+order parameters of $\alpha$ and $\beta$ carbons in lipid head group
+(see Fig. \ref{??} 
+\todo{Figure with chemical structure and labeling to be added}) 
+are proportional to the amount of bound charge
+in lipid bilayer \cite{seelig87}.
+More recent analysis included also
+the order parameter signs and concluded that the order parameters  
+decrease with bound positive charge and increase with bound negative 
+charge \cite{ollila16,catte16}. The observations are rationalized 
+as a change of lipid head group dipole tilt to more vertical orientation
+with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
+Order parameters for C-H bonds can be accurately measured for each
+C-H bond of lipids by using H$^2$ NMR or $^{13}$C NMR techniques and
+calculated from MD simulations directly from definition
+\begin{equation}\label{OP}
+S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
+\end{equation}
+where $\theta$ is the angle between the bond and membrane
+normal and average is taken over all sampled configurations \cite{ollila16}.
+
+The change of the head group order parameters as a function of
+the amount of bound charge per lipid $X^\pm$ 
+can be written as~\cite{ferreira16}
+\begin{equation}\label{OPchangeEQ}
+\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) =\frac{4m_i }{3\chi}X^\pm,
+\end{equation}
+where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
+$i$ refers to either $\alpha$ or $\beta$ carbon,
+%the empirically observed linear relation %for the change in order parameters %with respect to a bilayer without bound charges              
+%\begin{equation}\label{electrometer_eq}                                                                                                   
+%S_{\rm{CH}}^{i}(X^\pm)=S_{\rm{CH}}^{i}(0) + \frac{4m_i}{3\chi} X^\pm,                                                                     
+%\end{equation}                                                                                                                            
+$m_i$ is an empirical constant depending on the valency and position of bound charge,
+and the value of the quadrupole coupling constant is $\chi \approx$\,167\,kHz.
+Combination of atomic absorption spectra, $^2$H NMR experiments
+and information about order parameter signs gave
+$m_\alpha=-20.5$  and $m_\beta=-10.0$ for Ca$^{2+}$ binding to POPC
+bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
+
+Recent work published by the NMRlipids project showed that the concept works
+qualitatively also in simulations and can be used to compare ion binding
+affinitity between simulations and experiments~\cite{catte16}. However,
+it also turned out that the sensitivity of the order parameters response
+to the bound charge has to be quantified more accurately for quantitative comparison
+of binding affinity. Hence, we first quantify the response of
+head group order parameters against experimental data from cationinc surfactants
+embedded in PC bilayer~\cite{scherer89}. Essentially all surfactants locate in
+lipid bilayers in these experiments, thus the amount of the bound charge is known
+and can be plotted against order parameter change.
+Secondly we evaluate the response of the head group order parameters to varying salt concentrations in solution
+and estimate the amount of membrane bound charge.
+
+%\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
+%If it varies too much from the nominal concentration, I may need to tweak the scaling 
+%factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
+%to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
+%moment to guide me on the possible finite-size errors and this conc-error.
+The used experimental data report order parameters as a function of
+equilibrium cation concentration in the bulk solvent \cite{akutsu81,altenbach84}.
+Such a condition is reached in simulations by adjusting
+the simulation box size to dimensions large enough
+that ion conceration reaches a clear plateau in the bulk solvent.
+The concentrations in the units of $\mathrm{mol/l}$ were then 
+determined as
+\begin{equation}
+  C_{eq}=\frac{C_{plateau}}{0.602},
+\end{equation}
+where plateau concentration is the number density in the units of $\mathrm{nm}^{-3}$.
+\todo{SAMULI: Once we have to final results, we can probably say that
+  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
+  }
+
+\subsection{Comparison of lipid bilayer structure to experimental data}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.png}
+  \caption{\label{simVSexpNOions}
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
+    Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
+    compared with experimental order parameters from \cite{ferreira13}.}
+  \todo{Add acyl chain order parameters, POPC chemical structure} \\
+  \todo{We should run a simulation for the system withtout ions with the
+    same (or close) temperature as the experimental data. The NMR data is at 300K
+    so that would be good. I think that this would be also close enough for the scattering data at 303K.
+  }
+\end{figure}
+
+Lipid bilayer structure without ions was validated against NMR experiments
+by order parameters for C-H bonds and x-ray scattering experiments by using
+form factors. The former validates the structures sampled by individual
+lipid molecules in simulations with segmental resolution, while the latter
+validates the dimensions of the lipid bilayer (thickness and area per
+molecule)~\cite{ollila16}. Form factor was calculated from simulations
+as 
+\begin{equation}
+  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
+\end{equation}
+where $f_\alpha(q_z)$ is the density of atomic scattering length, 
+$\rho_s$ is the density of solvent scattering lenght,
+$n_\alpha (z)$ is the number density of atom $\alpha$ and
+$z$ is the distance from the membrane centre along its normal 
+spanning the membrane with thickness $D$.
+
+The scattering form factors, NMR order parameters and area per lipid calculated  
+using the ECC-lipids model for POPC are compared to experiments and original
+Lipid14 results in Fig. \ref{simVSexpNOions} and in Table~\ref{tab:apls}. 
+%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
+%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
+The structural quality is comparable to the state of art lipid models available in literature \cite{ollila16},
+thus we conclude that the ECC-lipids model reproduces the lipid bilayer structure
+in liquid disordered phase with similar accuracy to other available models.
+
+\subsection{Simulation details}
+
+\subsubsection{Simulations with aqueous ions}
+The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
+(POPC) bilayer and an aqueous salt solutions of varying concentrations.
+Water molecules were described by OPC3 model \cite{Izadi16}.
+In order to test transferability of the newly developed ECC-lipids model, 
+we also performed several additional simulations with water models 
+OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}
+presented in Supporting Information (SI). 
+We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+Simulations with Lipid14 use ion model by Dang \cite{smith94,chang1999,dang2006}. 
+Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
+The simulation settings used in this work are 
+summarized in Table \ref{tbl:mdpar}, 
+and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}. 
+Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+
+\begin{table}[btp]
+  \caption{Simulation parameters}
+  \label{tbl:mdpar}
+  \begin{tabular}{ll}
+    simulation property & parameter   \\
+    \hline
+    time-step           & 2~fs         \\
+    equilibration time  & 100~ns  \\
+    simulation time     & 200~ns  \\
+    temperature         & 313~K       \\
+    thermostat          & v-rescale  \cite{bussi07}   \\
+    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
+    long-range electrostatics & PME  \cite{darden93}  \\
+    cut-off scheme      & Verlet \cite{Pall13}      \\
+    Coulomb and VdW cut-off & 1.0~nm \\
+    constraints         & LINCS, only Hydrogen atoms \cite{hess97} \\
+    constraints for water & SETTLE  \cite{miyamoto92} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\subsubsection{Simulations with cationic surfactants}
+Automated topology builder \cite{malde11} was first used to create the structure of
+dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
+The code AmberTools \cite{amber} was then used to generate the Amber-type force field
+parameters. The parameters were converted to the Gromacs format by using
+acpype tool \cite{acpype}. The partial charges were then manually modified
+to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
+The surfactants were randomly placed among the lipids to form bilayer structures with
+mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
+These systems were ran with original Lipid14 POPC model and with ECC-corrected model.
+\todo{Describe the changes in surfactant model used with ECC-corrected lipids.}
+
+\section{Results and Discussion}
+
+\subsection{POPC membrane and electrometer concept with cationic surfactant}
+
+\begin{figure*}[tbp]
+  \centering
+  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2_surf.eps}
+  \caption{\label{OrderParameterCHANGESnewMODELS}
+    Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration and
+    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) mole fraction
+    from simulations and experiments (DPPC \cite{akutsu81}, POPC \cite{altenbach84}, surfactant \cite{scherer89}).
+  }
+\end{figure*}
+
 Headgroup order parameter response to the bound charge was evaluated against
@@
 practically all cationic surfactant molecules are embedded in the lipid bilayer due to
-their amphiphilic nature. Thus, the changes of headgroup order parameter as a function of
+their amphiphilic nature. Thus, the changes of head group order parameter as a function of
 mole fraction of cationic surfactants (Fig. \ref{OrderParameterCHANGESnewMODELS}) gives
@@
 
-The headgroup order parameter response to the bound cation concentration
+The head group order parameter response to the bound cation concentration
 is in experiments approximately linear up to $\sim$0.3 mole fraction of the cation \cite{scherer89}.
@@
 ), and this property is also kept by the newly derived Lipid14-based model, ECC-lipids. 
-The present comparison of headgroup order parameter changes 
+The present comparison of head group order parameter changes 
 suggests that the observed overestimated response of all models to 
 increasing CaCl$_2$ concentration in \cite{catte16} 
-can be explained at least in part by a high sensitivity of the headgroup order parameter response
+can be explained at least in part by a high sensitivity of the head group order parameter response
 to the bound charge.
 
-\subsection{Cation binding to POPC membrane}
+\subsection{POPC membrane and electrometer concept with cation binding}
 
-Headgroup order parameter response to increasing CaCl$_2$ concentration
-from experiments, original Lipid14 model and ECC-lipids model are shown
-in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
-is significantly overestimated in original Lipid14 model, while results from
-ECC-lipids model are in a good agreement with the experiments.
-This is a significant improvement over current models,
-which in general overstimate the head group order parameter response to CaCl$_2$
-concentration \cite{catte16}. The good agreement with experiments indicate
-that the binding details of \ce{Ca^2+} are realistic in the ECC-lipids model
-and, hence, it can be used to study lipid-ion interaction in detail.
-
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2_surf.eps}
-  \caption{\label{OrderParameterCHANGESnewMODELS}
-    Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration and
-    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) mole fraction
-    from simulations and experiments (DPPC \cite{akutsu81}, POPC \cite{altenbach84}, surfactant \cite{scherer89}).
-  }
-\end{figure*}
-
-Ion density profiles from simulations with different models are compared in Fig.~\ref{fig:cacl-dens}.
-The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
-show a pronounced peak in the position of the phosphate moieties of POPC. 
-The use of a ECC-ion model~\cite{kohagen14,kohagen16} 
-or the standard Amber ion model by \AA{quist} 
-along with Lipid14 does not significantly change it 
-(Headgroup order parameter responses for these models in SI).
-\todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
-The new ECC-lipids model with ECC-ions exhibits on the other hand smaller density in this region 
-suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
-Together with the accurate head group order parameter response (Fig.~\ref{OrderParameterCHANGESnewMODELS}),
-the decrease in cation density of ECC-lipids demonstrates that 
-ion binding to zwitterionic phospholipid membranes can be accurately
-described with classical MD simulation models that account for electronic polarizability.
-
-\begin{figure}[]
+\begin{figure}[tbp]
   \centering
@@
 
+After achieving accurate head group order parameter response with cationic surfactant, 
+we advanced to the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}. 
+Headgroup order parameter response to increasing CaCl$_2$ concentration
+from experiments, original Lipid14 model and ECC-lipids model are shown
+in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
+is significantly overestimated in original Lipid14 model, while results from
+ECC-lipids model are in a good agreement with the experiments.
+This is a significant improvement over current models,
+which in general overstimate the head group order parameter response to CaCl$_2$
+concentration \cite{catte16}. The good agreement with experiments indicate
+that the binding details of \ce{Ca^2+} are realistic in the ECC-lipids model
+and, hence, it can be used to study lipid-ion interaction in detail.
+
+Ion density profiles from simulations with different models are compared in Fig.~\ref{fig:cacl-dens}.
+The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
+show a pronounced peak in the position of the phosphate moieties of POPC. 
+The use of a ECC-ion model~\cite{kohagen14,kohagen16} 
+or the standard Amber ion model by \AA{quist} 
+along with Lipid14 does not significantly change it 
+(Headgroup order parameter responses for these models in SI).
+\todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
+The new ECC-lipids model with ECC-ions exhibits on the other hand smaller density in this region 
+suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
+Together with the accurate head group order parameter response (Fig.~\ref{OrderParameterCHANGESnewMODELS}),
+we can conclude that the present model, ECC-lipids, exhibits accurate ion binding affinity to zwitterionic phospholipid membranes
+suggesting that the models that overestimate head group order parameter response~\cite{catte16} also overestimate the amount of bound cations.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
+  \caption{\label{fig:cacl-bind}
+    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .}
+\end{figure}
+
 Good agreement of ECC-lipids model with experiments encourages us to analyse the
@@
 have stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
-As shown in Fig. \ref{fig:cacl-bind} this is in agreement
-with the ternary complex model suggested based on headgroup order parameter
+As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
+with the ternary complex model suggested based on head group order parameter
 experiments~\cite{altenbach84}.
@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
-  \caption{\label{fig:cacl-bind}
-    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .}
-\end{figure}
-
 %\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
@@
 be accurately described with classical MD simulation models, where electronic
-polarization is effectively included by using electronic continuum correction.
+polarization is effectively included by using electronic continuum correction (ECC)~\cite{leontyev11}.
 This is a significant improvement over other available lipid models,
-which all overestimate specific cation binding affinities \cite{catte16}.  
+which all overestimate specific cation binding affinities~\cite{catte16}.  
 The newly proposed model, which we denote as ''ECC-lipids~17'', 
-reproduces the lipid bilayer structural details
-with similar accuracy as other state of the art lipid models.
-ECC-lipids~17 exhibit accurate headgroup order parameter response to
+exhibits accurate head group order parameter response to
 bound cations, monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
-and divalent \ce{Ca^2+}.
+and divalent \ce{Ca^2+}
+also quantifying their binding affinities.
+Moreover, ECC-lipids~17 reproduce the lipid bilayer structural details
+with similar accuracy as other state of the art lipid models~\cite{catte16}.
 Several water models 
@@
 complex.
+\todo{Continue summary using previous section once it is finished.}
 
@@
 and force fields.
+The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}.
 

### 2017-08-07 Manuscript/manuscript.tex
Ca2+ electrometer results from a -close-to-final- model of Lipid14-based POPC (q_sc=0.8, sigma_sc=0.875, all non-tail atoms), extra scripts for GMX deffnm simulations, Manuscript: title of the future paper.
Updates to plots: OP-change plot has corrected concentrations on x-axis; added experimental values for tails to OrdPar plot with actual values;  updated density plots; Added simulation data for model f_q=0.8, f_siq=0.89; debugged scripts for order-parameter calculation added simple pythonic script for the calculation of corrected ionic concentration.

@@
 
-\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}
+\documentclass[aip,jcp,twocolumn]{revtex4}
+%\documentclass[aip,jcp]{revtex4}
+%\documentclass{article}
 
@@
   % insert abstract here
-  Cellular membranes are complex systems that are hard to study both experimentally and theoretically. Classical molecular dynamics simulations (MD)
-give detailed information about membrane structure, dynamics and its interactions with other biomolecules. However,  there is still a room for
-improvements in the current force fields, especially in the hydrophilic region of membrane.  According to the results reported in NMRlipids, Open Collaboration project
-running at nmrlipids.blogspot.fi, biologically relevant Na+ and Ca2+ generally overbind to zwitterionic phophatidylcholine (PC) membranes in current MD models [Ollila 2016].
-Here we suggest that the PC lipid membrane-ion interactions can be corrected by implicitly inducing electronic polarizability in the lipid models by
-using a simple electrostatic continuum correction (ECC) for partial charges [Leontyev 2010], which is successfully applied already to monovalent and divalent ions in bulk water [Kohagen, 2015].
-Our results show that applying ECC to existing force fields significantly improved PC lipid--cation interactions when compared against NMR data
-by using the electrometer concept [Ollila 2016,Seelig87]. However, also other force field parameters have to be made consistent
-for ECC correction to preserve the area per molecule and other relevant bilayer properties. This work is carried as Open Collaboration
-in NMRlipids VI project (http://nmrlipids.blogspot.fi).
+  \todo{Abstract directly from Joe's conference abstracts. To be rewritten.}
+Classical molecular dynamics simulations give detailed information about membrane structure and dynamics. 
+However, there is still a room for improvements in current force fields – it is known from the literature,
+that the binding of ions, especially cations, to phopholipid membranes is overestimated in all classical models [1]. 
+We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the
+lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent
+ions yielding models that feature correct ion pairing [3]. 
+Using the electrometer concept [3, 4] and x-ray scattering form factors, our simulations point out that our hypothesis is
+correct and ECC is indeed a missing important contribution in current classical lipid models. 
+Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. 
+The new lipid model, "ECC-lipids", shows accurate binding affinity to sodium and calcium cations and head group order parameter response to bound charge. 
+We also provide for the first time a realistic stochiometry of bound calcium cations to a POPC membrane, and their binding sites. 
+This work will continue as an open collaboration project NMRlipids VI (http://nmrlipids.blogspot.fi).
 \end{abstract}
@@
 
-%\label{}
+% Here I write pseudo-article statements that will make the main argument.
+% Beautiful polished sentences will be formed only afte we agree on these basic things
 \section{Introduction}
+Cation interactions with cellular membranes play a key role in several biological processes,
+like in signal propagation in neurons and vesicle fusion.
+\todo{JOE: list more introductory examples, where it has interesting effects.}
+Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
+to understand molecular level details of specific cation interactions with
+cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
+and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} \todo{We need to select references for these}.
+While relative binding affinity of different ions is agreed to follow Hoffmeister
+series, the molecular details of binding and binding energetics are
+not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
+%General conclusion from experimental
+%studies has been that multivalent ions and lithium have weak specific binding in
+%phopholipid bilayers, while other monovalent ions do not essentially
+%bind 
+%The presence of anoinic lipids, like PS or PG,
+%increase the concentration close to the bilayer and thus the amount of bound
+%ions, but do not affect the specific binding constant \cite{seelig90}.
+Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
+and infrared scpectroscopy, give accurate information about ion binding
+in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
+Molecular level interpretation of the results, however, requires assumptions
+about binding model and is often not fully conclusive \cite{??}.
+On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
+have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
+The overestimated specific cation binding observed in current lipid models \cite{catte16}
+may lead to articifially positively charged memranes and significant artefacts in
+MD simualtions (e.g.~\ce{Ca^2+} induced charge inversion of bacterial membranes~\cite{Luan2016}).
 
-Cations were shown to generally overbind in PC lipid bilayers in
-NMRlipids II project. Here we propose that the cation overbinding can
-be corrected by implicitly inducing electronic polarizability in
-lipid headgroups by scaling the partial charges.
+%The binding details, like binding sites and stoichiometry are not yet fully
+%resolved but interpretation of NMR and scattering experiments suggest that one
+%Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
+%phospholipid molecules \cite{altenbach84}.
 
-%\section{Methods}
+In this work we show that the accuracy of ion binding, especially calcium,
+in zwitterionic PC lipid bilayer can be significantly improved 
+for classical MD simulation models of zwitterionic lipids
+by including electronic polarizability
+in the polar region of lipid molecules
+using the electronic continuum correction (ECC)~\cite{leontyev11}. 
+This approach has been previously shown to improve performace of
+ion models against neutron scattering data in concentrated solutions \cite{Jungwirth2015,kohagen14,kohagen16}
+and ab-initio simulations~\cite{Jungwirth2017}. 
+The realistic structure of ions in a bulk solvent was not, however, sufficient to correct binding in lipid
+bilayers~\cite{catte16}. 
+To fix it, we propose
+a classical fixed-charge MD simulation model of 
+1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) lipid 
+that accounts for electronic polarizability by using ECC. 
+We will validate such a model by comparing its lipid bilayer structue
+to x-ray scattering data and its head group structure to NMR experiments. 
+We will then quantify the accuracy of head group order parameter response to bound charge
+and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
+In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
+This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
+The proposed POPC model, ECC-lipids, and the approach to obtain it
+is, hence, highly useful for future MD simulations with physiological salt conditions, 
+and for effects where membrane electrostatics play a major role.
 
-%\subsection{Molecular dynamics simulations}
+\section{Methods}
 
-\section{Effect of scaling of headgroup and glycerol backbone partial charges on lipid bilayer properties and cation binding}
+\subsection{Electronic continuum correction for lipid bilayers}
+The lack of electronic polarizability in the classical MD simulation
+models has been considered higly relevant issue since the early days of
+lipid bilayer simulations. Some pioneering simulation studies scaled
+the partial charges of atoms by one half to effectively include 
+polarizability \cite{jonsson86,egberts94}. Also, approaches that
+explicitly include polarizability has been introduced, but
+this has turned out to be practically complicated \cite{chowdhary13}.
+In this work we effectively include electronic polarizability in lipid
+bilayer simulations by using the electronic continuum correction
+(ECC)~\cite{leontyev11}. 
+Technically, it is a similar approach to the phenomenological charge-scaling as applied in 
+the early studies where a scaling factor one half was used \cite{jonsson86,egberts94}. 
+In contrast to those studies, however, the concept of ECC is 
+physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11},
+but significantly more simple than the explicit inclusion of electronic
+polarizability \cite{chowdhary13} both for derivation and application.
 
-Headgroup and glycerol backbone partial charges of Slipids and Lipid14 models were scaled
-with various scaling factors and cation binding was monitored by using the electrometer
-concept. Headgroup order parameter changes as a function of ion concentration are
-shown in Fig. \ref{OrderParameterIONSchangesSCALED}. The order parameter decrease
-seemst to be proportional to the headgroup charges: the smaller the charges
-(larger scaling factor), the less change is observed in order parameters.
-According to the electrometer concept, this means that the headgroup scaling
-decreases cation binding affinity. However, from this data only we cannot
-exclude the possibility that order parameters are changing less because
-headgroup becomes less sensitive for cation binding.
-\begin{figure}[]
+According to ECC, electronic
+polarizability can be included in classical MD simulations by
+placing all particles into a homogeneous dielectric continuum 
+with a dielectric constant $\epsilon _{el}$, 
+which is the electronic part of the dielectric constant of 
+the media \cite{leontyev11}. Measurements of high frequency 
+dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
+for almost any biomaterial \cite{some_original_work, leontyev11}.
+Such a continuum can be easily included in standard MD simulation by
+a formal transformation of partial charges 
+\begin{equation}
+  Q^{ECC} = f_q \cdot Q
+\end{equation}
+with a constant scaling factor $f_q = \epsilon _{el} ^{-1/2}$ 
+effectively representing the newly introduced electronic continuum. 
+The value measured for water, $\epsilon _{el} = 1.78$, gives 
+a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}, which has been
+successfully used to improve the performance of ion force fields \cite{kohagen14,kohagen16,??}.
+
+The main goal of this work is to apply ECC correction to accurately 
+describe the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
+in MD simulation models when compared with NMR data \cite{catte16}. 
+The Lipid14~\cite{dickson14} force field parameters were used as a starting point,
+because their response to bound ions was most realistic against NMR data
+(see Fig.~5 in Ref.~\citenum{catte16}).
+Also glycerol backbone and head group structures in Lipid14 model were
+relatively realistic when compared with other state of the art lipid models \cite{botan15}.
+
+While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
+to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
+it is not clear if the same factor should be used for lipid head groups. 
+Unlike the total charge of cations and anions, atomic partial charges used in
+classical MD simulations are not experimental observables and 
+can in principle adopt arbitrary values. 
+Here we estimate the correct scaling factor for Lipid14 model by using
+so called ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013},
+which are the average of partial charges given by 
+restrained electrostatic potential fitting (RESP~\cite{RESP_paper}) 
+in vacuum and in solvent.
+Using the partial atomic charges obtained from RESP fitting in vacuum and implicit solvent reported in \cite{maciejewski14} and
+considering that the partial charges in original Lipid14 parametrization were derived using RESP~\cite{RESP_paper} in vacuum,
+we conclude that the charges in Lipid14 model shall be scaled roughly by the factor $f_q=0.8$. 
+%By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
+%we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
+%by increasing the scaling factor $f_q$ to $0.8$. 
+%-----\footnote----------
+  %{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
+  %types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
+  %partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
+  %I use 0.80 as an approximate round value, which accounts for the already lower charges in Lipid14. 
+  %The use of 0.78---.79 might be more appropriate, though.}
+
+Here we apply the ECC correction with the proposed scaling factor only to the head group,
+glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
+are expected to have the largest contribution to the cation binding.
+The hydrocarbon chain parameters are not modified, because they are
+already highly optimized in most lipids models including Lipid14 and
+give generally a good description for hydrophobic part of lipid bilayers
+in various conditions \cite{ollila16}.
+In contrast, all available lipid models require
+some improvements in glycerol backbone and head group regions \cite{botan15}. 
+%To minimize the detuning of the highly optimized hydrocarbon
+%parameters, 
+%Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
+%are found to interact only with the phospholipid head group,
+%and the accuracy of alkyl sidechains in Lipid14 model are satisfactory, 
+%we exclude alkyl tails from our modifications. 
+The scaling of partial charges in the polar region was found to decrease
+area per molecule to 60~\AA$^2$, which is significantly smaller than the
+experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment}) and the original Lipid14 values $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}..  
+The decrease of area was found to arise from a lower hydration of the lipid head group region,
+%\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
+%JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
+which can be explained by the decreased solvation free energy due to the lower polarity
+of molecules with scaled charges. 
+The solvation free energy and hydration is sensitive to Van der Waals interaction parameters,
+hence, they can be increased by reducing the radius of the selected atoms by changing the 
+the $\sigma$ parameters in Lennard-Jones potential. 
+%This allows water molecules to approach closer to lipid atoms and
+%have stonger electrostatic interactions with them.
+%\todo{SAMULI: This effect may have an official name. In that case we should mention it.}
+Here we reduce $\sigma$ with a scaling factor of $f_\sigma = 0.89$ for the same atoms for which
+charges were scaled. This has lead to an increase of the area per molecule back
+close to the experimental value (see Table~\ref{tab:apls}). 
+%\todo{Is there some justification for the below sentence?} 
+%In line with the electronic continnum correction,
+%the scaling factor for sigma parameters, $f_\sigma$,
+%has to lay between $f_q$ and 1.
+
+\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept}
+Ion binding was compared between experiments and simulations by 
+using lipid head group order parameters and ''electrometer concept'' 
+introduced by Seelig et al. \cite{seelig87,catte16}.
+The concept is based on the experimental observation that the 
+order parameters of $\alpha$ and $\beta$ carbons in lipid head group
+(see Fig. \ref{??} 
+\todo{Figure with chemical structure and labeling to be added}) 
+are proportional to the amount of bound charge
+in lipid bilayer \cite{seelig87}.
+More recent analysis included also
+the order parameter signs and concluded that the order parameters  
+decrease with bound positive charge and increase with bound negative 
+charge \cite{ollila16,catte16}. The observations are rationalized 
+as a change of lipid head group dipole tilt to more vertical orientation
+with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
+Order parameters for C-H bonds can be accurately measured for each
+C-H bond of lipids by using H$^2$ NMR or $^{13}$C NMR techniques and
+calculated from MD simulations directly from definition
+\begin{equation}\label{OP}
+S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
+\end{equation}
+where $\theta$ is the angle between the bond and membrane
+normal and average is taken over all sampled configurations \cite{ollila16}.
+
+The change of the head group order parameters as a function of
+the amount of bound charge per lipid $X^\pm$ 
+can be written as~\cite{ferreira16}
+\begin{equation}\label{OPchangeEQ}
+\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) =\frac{4m_i }{3\chi}X^\pm,
+\end{equation}
+where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
+$i$ refers to either $\alpha$ or $\beta$ carbon,
+%the empirically observed linear relation %for the change in order parameters %with respect to a bilayer without bound charges              
+%\begin{equation}\label{electrometer_eq}                                                                                                   
+%S_{\rm{CH}}^{i}(X^\pm)=S_{\rm{CH}}^{i}(0) + \frac{4m_i}{3\chi} X^\pm,                                                                     
+%\end{equation}                                                                                                                            
+$m_i$ is an empirical constant depending on the valency and position of bound charge,
+and the value of the quadrupole coupling constant is $\chi \approx$\,167\,kHz.
+Combination of atomic absorption spectra, $^2$H NMR experiments
+and information about order parameter signs gave
+$m_\alpha=-20.5$  and $m_\beta=-10.0$ for Ca$^{2+}$ binding to POPC
+bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
+
+Recent work published by the NMRlipids project showed that the concept works
+qualitatively also in simulations and can be used to compare ion binding
+affinitity between simulations and experiments~\cite{catte16}. However,
+it also turned out that the sensitivity of the order parameters response
+to the bound charge has to be quantified more accurately for quantitative comparison
+of binding affinity. Hence, we first quantify the response of
+head group order parameters against experimental data from cationinc surfactants
+embedded in PC bilayer~\cite{scherer89}. Essentially all surfactants locate in
+lipid bilayers in these experiments, thus the amount of the bound charge is known
+and can be plotted against order parameter change.
+Secondly we evaluate the response of the head group order parameters to varying salt concentrations in solution
+and estimate the amount of membrane bound charge.
+
+%\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
+%If it varies too much from the nominal concentration, I may need to tweak the scaling 
+%factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
+%to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
+%moment to guide me on the possible finite-size errors and this conc-error.
+The used experimental data report order parameters as a function of
+equilibrium cation concentration in the bulk solvent \cite{akutsu81,altenbach84}.
+Such a condition is reached in simulations by adjusting
+the simulation box size to dimensions large enough
+that ion conceration reaches a clear plateau in the bulk solvent.
+The concentrations in the units of $\mathrm{mol/l}$ were then 
+determined as
+\begin{equation}
+  C_{eq}=\frac{C_{plateau}}{0.602},
+\end{equation}
+where plateau concentration is the number density in the units of $\mathrm{nm}^{-3}$.
+\todo{SAMULI: Once we have to final results, we can probably say that
+  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
+  }
+
+\subsection{Comparison of lipid bilayer structure to experimental data}
+
+Lipid bilayer structure without ions was validated against NMR experiments
+by order parameters for C-H bonds and x-ray scattering experiments by using
+form factors. The former validates the structures sampled by individual
+lipid molecules in simulations with segmental resolution, while the latter
+validates the dimensions of the lipid bilayer (thickness and area per
+molecule)~\cite{ollila16}. Form factor was calculated from simulations
+as 
+\begin{equation}
+  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
+\end{equation}
+where $f_\alpha(q_z)$ is the density of atomic scattering length, 
+$\rho_s$ is the density of solvent scattering lenght,
+$n_\alpha (z)$ is the number density of atom $\alpha$ and
+$z$ is the distance from the membrane centre along its normal 
+spanning the membrane with thickness $D$.
+
+\subsection{Simulation details}
+
+\subsubsection{Simulations with aqueous ions}
+The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
+(POPC) bilayer and an aqueous salt solutions of varying concentrations.
+Water molecules were described by OPC3 model \cite{Izadi16}.
+In order to test transferability of the newly developed ECC-lipids model, 
+we also performed several additional simulations with water models 
+OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}
+presented in Supporting Information (SI). 
+We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+Simulations with Lipid14 use ion model by Dang \cite{smith94,chang1999,dang2006}. 
+Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
+The simulation settings used in this work are 
+summarized in Table \ref{tbl:mdpar}, 
+and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}. 
+Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+
+\begin{table}[btp]
+  \caption{Simulation parameters}
+  \label{tbl:mdpar}
+  \begin{tabular}{ll}
+    simulation property & parameter   \\
+    \hline
+    time-step           & 2~fs         \\
+    equilibration time  & 100~ns  \\
+    simulation time     & 200~ns  \\
+    temperature         & 313~K       \\
+    thermostat          & v-rescale  \cite{bussi07}   \\
+    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
+    long-range electrostatics & PME  \cite{darden93}  \\
+    cut-off scheme      & Verlet \cite{Pall13}      \\
+    Coulomb and VdW cut-off & 1.0~nm \\
+    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
+    constraints for water & SETTLE  \cite{miyamoto92} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\subsubsection{Simulations with cationic surfactants}
+Automated topology builder \cite{malde11} was first used to create the structure of
+dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
+The code AmberTools \cite{amber} was then used to generate the Amber-type force field
+parameters. The parameters were converted to the Gromacs format by using
+acpype tool \cite{acpype}. The partial charges were then manually modified
+to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
+The surfactants were randomly placed among the lipids to form bilayer structures with
+mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
+These systems were ran with original Lipid14 POPC model and with ECC-corrected model.
+\todo{Describe the changes in surfactant model used with ECC-corrected lipids.}
+
+\section{Results and Discussion}
+
+\subsection{POPC membrane and its structure}
+
+\begin{figure}[tbp]
   \centering
-  \includegraphics[width=9.0cm]{../Fig/OrderParameterIONSchangesSCALED.eps}
-  \caption{\label{OrderParameterIONSchangesSCALED}
-    Headgroup order parameter changes as a function of cations from models
-    with modified headgroup partial charges.}
+  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.png}
+  \caption{\label{simVSexpNOions}
+    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
+    Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
+    compared with experimental order parameters from \cite{ferreira13}.}
+  \todo{Add acyl chain order parameters, POPC chemical structure} \\
+  \todo{We should run a simulation for the system withtout ions with the
+    same (or close) temperature as the experimental data. The NMR data is at 300K
+    so that would be good. I think that this would be also close enough for the scattering data at 303K.
+  }
 \end{figure}
 
-The best models from Fig.  \ref{OrderParameterIONSchangesSCALED} are shown in
-Fig. \ref{OrderParameterCHANGESnewMODELS}. Based on these results we choose
-two models for more careful studies: Slipids and Lipid 14 models with headroup
-and glycerol backbone partial charges scaled with 0.8 and 0.85, respectively.
-\begin{figure}[]
+\begin{table}
+  \caption{Area per lipid (APL) from different models of POPC without ions\label{tab:apls} }
+  \begin{tabular}{l|c c}
+    model          & APL (\AA$^2$)   & Temperature [K] \\
+    \hline
+    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
+    \hline
+    ECC-lipids &        &  \\
+    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
+    ($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
+    \hline
+    (6.4 nm)$^2$, 128 lipids patch, OPC3           & 64.2   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, SPC/E          & 65.1   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, OPC            & 64.4   &  313       \\
+    (6.4 nm)$^2$, 128 lipids patch, TIP4p/2005     & 66.8   &  313       \\
+    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
+    %oMM small patch           & 63.65  &         \\
+    %oMM 4xbig patch           & 63.7   &         \\
+    \hline
+    %experiment   & 62.7  &  293    \\
+    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
+    experiment  & 67.3  &  323    \\
+    %experiment  & 68.1  &  333    \\
+    %experiment POPE  & 56.6 &  303    \\
+    \hline
+  \end{tabular}
+\end{table}
+
+In order to validate the newly developed model, ECC-lipids, 
+we compared our simulation results without ions to
+NMR order parameters measurements and x-ray scattering form factors 
+(Fig.~\ref{simVSexpNOions} and Table~\ref{tab:apls}). 
+The tail order parameters being already highly optimized in the original Lipid14 model~\cite{dickson14}
+are found to match the experimental values nicely.
+The head group and glycerol backbone order parameters accuracy is 
+comparable to the state of art lipid models available in literature \cite{ollila16},
+The agreement between the x-ray scattering form factors and areas per molecule from simulations and experiments
+confirm that the membrane structural properties are well captured. 
+%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
+%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
+
+\subsection{POPC membrane and electrometer concept with cationic surfactant}
+
+\begin{figure*}[tbp]
   \centering
-  \includegraphics[width=9.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
+  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2_surf.eps}
   \caption{\label{OrderParameterCHANGESnewMODELS}
-    Headgroup order parameter changes as a function of cations from models
-    with modified headgroup partial charges.}
+    Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration and
+    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) mole fraction
+    from simulations and experiments (DPPC \cite{akutsu81}, POPC \cite{altenbach84}, surfactant \cite{scherer89}).
+  }
+\end{figure*}
+
+Headgroup order parameter response to the bound charge was evaluated against
+experimental data measured with the cationic surfactant
+(dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) \cite{scherer89}.
+The exact amount of bound charge in the membrane is known in these systems, because
+practically all cationic surfactant molecules are embedded in the lipid bilayer due to
+their amphiphilic nature. Thus, the changes of head group order parameter as a function of
+mole fraction of cationic surfactants (Fig. \ref{OrderParameterCHANGESnewMODELS}) gives
+also the order parameter changes as a function of the amount of bound cations.
+\todo{Split the OP-changes figure in two: cationic surfactant -and- binding of cations.
+      The separate figure with the cationic surfactant shall be additionally accompanied
+      with a plot of Head group tilt (PN-vector?) change/response. }
+
+The head group order parameter response to the bound cation concentration
+is in experiments approximately linear up to $\sim$0.3 mole fraction of the cation \cite{scherer89}.
+The linearity is also observed in simulations with original Lipid14 and with newly derived ECC-lipids models.
+Quantitative comparison, however, reveals that the response is overestimated in 
+Lipid14 for both segments, while the ECC-lipids model gives a
+good agreement for the change of $\alpha$ segment order parameter, and slightly
+underestimates the $\beta$ segment order parameter change.
+Such a finding is not surprising, as the $\Delta S^\alpha/\Delta S^\beta$ ratio 
+of the response of Lipid14 model is found to be slightly larger than in 
+experiments (VALexp vs VALsim
+\todo{Add values of $S^\alpha/S^\beta$ response for sim and experiment
+  SAMULI: Maybe these could be put in the same figure with the order parameter
+  response and P-N vector angle change with surfactants.
+}
+), and this property is also kept by the newly derived Lipid14-based model, ECC-lipids. 
+The present comparison of head group order parameter changes 
+suggests that the observed overestimated response of all models to 
+increasing CaCl$_2$ concentration in \cite{catte16} 
+can be explained at least in part by a high sensitivity of the head group order parameter response
+to the bound charge.
+
+\subsection{POPC membrane and electrometer concept with cation binding}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities.eps}
+  \caption{\label{fig:cacl-dens}
+    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
+    and with ECC-ions and ECC-lipids model with ECC-ions. }
+  \todo{Change the figure so that it contains a membrane background} \\
+  \todo{What are the concentrations (calculated as in NMRlipids II) in this figure?} \\
+  \todo{These should be also compared to the results with Åqvist ions, which are shown
+    in NMRlipids II. NMRlipids II have results with 350mM and 1000mM. If we have the distributions with
+    the same concentrations Åqvist results could be added also in this figure. If not, the
+  comparison has to be made some other way.}
 \end{figure}
 
-While scaling seems to improve ion binding, it changes the lipid bilayer properties
-without ions. The glycerol backbone and headgroup order parameters are plotted in
-Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
-i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
-The area per lipid from different models are shown in Table \ref{apls}
-The conclusion is that the charge scaling has only a marginal effect on
-headgroup order parameters, while area per lipid is significantly decreased.
-Area per lipid decrease is more pronounced in Lipid14 than in Slipids.
+After achieving accurate head group order parameter response with cationic surfactant, 
+we advanced to the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}. 
+Headgroup order parameter response to increasing CaCl$_2$ concentration
+from experiments, original Lipid14 model and ECC-lipids model are shown
+in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
+is significantly overestimated in original Lipid14 model, while results from
+ECC-lipids model are in a good agreement with the experiments.
+This is a significant improvement over current models,
+which in general overstimate the head group order parameter response to CaCl$_2$
+concentration \cite{catte16}. The good agreement with experiments indicate
+that the binding details of \ce{Ca^2+} are realistic in the ECC-lipids model
+and, hence, it can be used to study lipid-ion interaction in detail.
+
+Ion density profiles from simulations with different models are compared in Fig.~\ref{fig:cacl-dens}.
+The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
+show a pronounced peak in the position of the phosphate moieties of POPC. 
+The use of a ECC-ion model~\cite{kohagen14,kohagen16} 
+or the standard Amber ion model by \AA{quist} 
+along with Lipid14 does not significantly change it 
+(Headgroup order parameter responses for these models in SI).
+\todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
+The new ECC-lipids model with ECC-ions exhibits on the other hand smaller density in this region 
+suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
+Together with the accurate head group order parameter response (Fig.~\ref{OrderParameterCHANGESnewMODELS}),
+we can conclude that the present model, ECC-lipids, exhibits accurate ion binding affinity to zwitterionic phospholipid membranes
+suggesting that the models that overestimate head group order parameter response~\cite{catte16} also overestimate the amount of bound cations.
 
@@
   \centering
-  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodels.eps}
-  \caption{\label{HGopsNEWmodels}
-    Headgroup and glycerol backbone order parameters from standard and EEC models.}
+  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
+  \caption{\label{fig:cacl-bind}
+    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .}
 \end{figure}
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/HGopsNEWmodelsCHANGE.eps}
-  \caption{\label{HGopsNEWmodelsCHANGE}
-    Changes in headgroup and glycerol backbone order parameters due to EEC.}
-\end{figure}
+Good agreement of ECC-lipids model with experiments encourages us to analyse the
+binding details from MD simulations. Direct analysis of contacts between ions and
+lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
+have stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
+As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
+with the ternary complex model suggested based on head group order parameter
+experiments~\cite{altenbach84}.
+In addition to the ternary complexes, there also is a non-negligable probability
+of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
+Technical details of the analysis are in the SI.
 
-\begin{table}
-  \caption{Area per lipid from different models for POPC without ions\label{apls} }
-  \begin{tabular}{c c}
-    model          & A (Å$^2$)    \\
-    Lipid14 (literature)  & 65.6$\pm$ 0.5   \\
-    Lipid14eec0.85 & 55.5      \\
-    Slipids (literature T=303K)       & 64.6 $\pm$ 0.4   \\
-    SlipidsEEC0.8  & 57.8  \\
-  \end{tabular}
-\end{table}
+\todo{JOE: Put details of the cation-binding stoichiometry analysis to SI.} \\
+\todo{JOE: Update the binding isotherm figure with new simulations} \\
+\todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
+  be the only option. I will recheck and come back to this.
+  SAMULI: This is written in \cite{macdonald87}:
+  ''Ca2+binding to POPC bilayers over the whole concentration range can be best described in terms
+  of formation of a ternary complex involving complexation of two lipids to one calcium ion (Altenbach and Seelig, 1984).
+  The addition of a sodium competition term has not changed this conclusion. However, if Ca2+ concentrations
+  up to 100 mM are considered, the data can be equally well explained by a 1:1 binding mechanism (cf. Figure 7).
+  In contrast, the Ca2+ binding to POPC-POPG mixtures can be best described by assuming a 1:l stoichiometry regardless of the range of Ca2+concentrations.''
+  We might or might not want to discuss about this.} \\
+\todo{SAMULI: I would also analyze how much there is contact between ions and different
+parts of the lipid (phosphase, carbonyl, etc.).
+JOE: Pavel suggest using volumetric map (e.g. in VMD) to visualise the amounts of Ca, Na and Cl bound to different moieties.} \\
+\todo{SAMULI: I think we should quantify this, i.e. how much there are these.
+  Maybe also the other possible complexes? Maybe also the correlation betweem complexes
+  and binding sites, if it is not too much work.
+  JOE: This looks like a careful work for the next paper to me. 
+  I'd only add a relatively simple analysis of binding sites and probably the propensity of 1-2-3 membered clusters.
+  SAMULI: I agree that we should not use too much time on this now.}. \\
+\todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},
+  their interaction energies with the lipid membrane, etc,
+  and finalize the discussion after these results.}
 
-The Calcium densities from new models and standard models are compared in Fig \ref{CAdensities}.
-It seems to me that Ca binding in Lipid14 model is not reduced by scaling the
-partial charges even thought the order parameter response is better.
-On the other hand, in Slipids the binding seems to be reduced.
-Possible interpretation is that Lipid14 result gets better because headgroup
-gets less sensitive for bound charge, while in Slipid the binding is reduced.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/CAdensities.eps}
-  \caption{\label{CAdensities}
-    Calcium densities from different simulations with new and standard models.
-  }
-\end{figure}
+%It is also suggested that the addition of \ce{NaCl} to the solution of \ce{CaCl_2} enhances the hedgroup 
+%order parameter response compared to the solution with only \ce{CaCl_2}. \cite{altenbach84}
+%\todo{Simulate this effect and discuss it further}
+
+%\todo{The difference between DPPC and POPC -- simulate and compare with experiment. }
+
+%\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
 
 \section{Conclusions}
+We show that the \ce{Na^+} and \ce{Ca^2+} binding in phospholipid bilayers can
+be accurately described with classical MD simulation models, where electronic
+polarization is effectively included by using electronic continuum correction (ECC)~\cite{leontyev11}.
+This is a significant improvement over other available lipid models,
+which all overestimate specific cation binding affinities~\cite{catte16}.  
+The newly proposed model, which we denote as ''ECC-lipids~17'', 
+exhibits accurate head group order parameter response to
+bound cations, monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
+and divalent \ce{Ca^2+}
+also quantifying their binding affinities.
+Moreover, ECC-lipids~17 reproduce the lipid bilayer structural details
+with similar accuracy as other state of the art lipid models~\cite{catte16}.
+Several water models 
+(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
+were used to exemplify the transferability of 
+the parameters of the new ECC-lipids~17 force field.
 
-% Tables may be be put in the text as floats.
-% Here is an example of the general form of a table:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
-% \begin{tabular}{} command.
-%
-% \begin{table}
-% \caption{\label{} }
-% \begin{tabular}{}
-% \end{tabular}
-% \end{table}
+Direct analysis of calcium binding details from MD simulations is in agreement
+with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
+In this model 1~calcium binds to 2~POPC molecules, which together form a ternary
+complex.
+\todo{Continue summary using previous section once it is finished.}
+
+The electronic continuum correction is applied here on Lipid14 POPC model \cite{dickson14},
+but we expect that the correction can be generalized also for other lipids
+and force fields.
+The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}.
+
+This work will serve as a foundation stone of a 
+new open-collaboration project NMRlipids~VI in nmrlipids.blogspot.fi.
 

### 2017-08-07 Data/POPC/no-ions/Lipid14/sim22_T7H-42p_SPCE_non-tail-atoms_q-sc_0.80_sig-sc_0.89/submit.sh

@@ -1,0 +1,132 @@
+#!/bin/bash
+
+#PBS -N L14_headgr80_0mM
+#PBS -q gpu_long
+#PBS -l walltime=167h
+#PBS -l mem=20gb
+#PBS -l nodes=1:ppn=1:gpu=1:nfs4:cl_doom
+#PBS -l scratch=10gb
+#PBS -j oe
+
+
+
+export stob3='/storage/brno3-cerit/home/melcrj'
+export storage=$stob3
+export workfolder=$PBS_O_WORKDIR   # $storage/wilCleGFP_3lipids_POPS/md12_dAFED/val-met_dist_dAFED
+
+
+# suggestion from https://wiki.metacentrum.cz/wiki/Pr%C3%A1ce_s_daty_v_%C3%BAloze
+trap 'cp -r $SCRATCH/* $workfolder/ && clean_scratch' TERM
+
+
+
+
+echo 'load modules ... '
+#  also loads gcc-4.8.1 automatically
+module load cmake-3.6.1 swig-3.0.8 cuda-8.0  doxygen-1.8.6 fftw-3.3.4omp gcc-4.8.4 
+module load gcc-4.9.2 
+module load gcc-5.3.0 
+
+# get rid of annoying warning from openMPI
+export OMPI_MCA_mpi_warn_on_fork=0
+
+
+
+# Just a basic initial printout
+echo
+echo 'Basic job -- info'
+echo '-----------------'
+echo 'job name: ' $PBS_JOBNAME
+echo 'job id: ' $PBS_JOBID
+echo 'name: ' $name
+echo 'running on: ' `hostname`
+echo 'proc per node: ' $PBS_NUM_PPN
+echo 'num nodes: ' $PBS_NUM_NODES
+echo 'Req-d walltime: ' $TORQUE_RESC_TOTAL_WALLTIME
+# echo 'nodelist: (PBS_NODEFILE) '
+# cat $PBS_NODEFILE
+echo 
+
+cd $workfolder          || exit 1
+echo 'cwd: ' `pwd`
+echo
+
+
+echo `env` > $infofile.env
+
+
+
+# check hwether previous run ended/copied OK
+if [ -e ___COPY_FAILED___ ] 
+then
+   echo previous copy had failed, ___COPY_FAILED___ file exist.
+   echo Solve the problem and remove this file after that. 
+   exit 3
+fi
+
+
[…]

### 2017-08-07 Data/POPC/no-ions/Lipid14/sim22_T7H-ECC-42p_SPCE_non-tail-atoms_q-sc_0.80_sig-sc_0.89/omm-md_NpT_heavyH_CUDA_gaiwan.py

@@ -1,0 +1,67 @@
+from simtk.openmm.app import *
+from simtk.openmm import *
+from simtk.unit import *
+from sys import stdout
+import os
+from datetime import datetime
+#import openmmplumed
+import IPython
+#import omm_vfswitch
+#import omm_readparams
+
+seconds_per_day = 86400
+NSTEPS = 250000 # 1ns with 4fs time step
+
+
+dt = 0.004*picoseconds
+temp = 313.15
+
+print "Read coordinates and parameters"
+gro = GromacsGroFile('conf_init.gro')
+top = GromacsTopFile('topol.top', periodicBoxVectors=gro.getPeriodicBoxVectors())
+system = top.createSystem(nonbondedMethod=PME, nonbondedCutoff=1.0*nanometer, constraints=HBonds, hydrogenMass=4*atomic_mass_unit)
+
+print "Created system"
+print "Set-up barostat and integrator"
+
+# barostat adopted from charmm-gui scripts
+barostat = MonteCarloMembraneBarostat(1*bar, 0*bar*nanometer, temp*kelvin, MonteCarloMembraneBarostat.XYIsotropic, MonteCarloMembraneBarostat.ZFree, 100)
+system.addForce(barostat)
+
+integrator = LangevinIntegrator(temp*kelvin, 1/picosecond, dt)
+
+#%%
+########## SIMULATION ##########
+
+print "Set Simulation class"
+#simulation_run = Simulation(top.topology, system, integrator, platform=openmm.Platform_getPlatformByName('CUDA'))   #, platformProperties={"CudaCompiler":"nvcc -isystem /usr/local/gcc/5.4.0/include/c++/5.4.0/ -ccbin /usr/local/gcc/5.4.0/bin/gcc"}) ##'OpenCL')) ##'CPU')) #
+simulation_run = Simulation(top.topology, system, integrator, platform=openmm.Platform_getPlatformByName('CUDA'), platformProperties={"CudaCompiler":"nvcc -isystem /usr/local/gcc/5.4.0/include/c++/5.4.0/ -ccbin /usr/local/gcc/5.4.0/bin/gcc"}) ##'OpenCL')) ##'CPU')) #
+simulation_run.context.setPositions(gro.positions)
+simulation_run.loadState('state_checkpoint.openmm')
+
+print "Set Reporter"
[…]

### 2017-08-07 Data/POPC/no-ions/Lipid14/sim22a_0mM_NaCl_OPC4_non-tail-atoms_q-sc_0.80_sig-sc_0.89_noIons-ref/form_f_analysis/calc_FORM_FACTOR.sh

@@ -1,0 +1,38 @@
+#LIPIDname=$(grep M_POPC_M ../mappingFILE.txt | awk '{printf "%5s\n",$2}')
+#grep ' 1POPC' ../conf.gro | awk '{print $2}' > tmpEL.dat
+#cat tmpEL.dat | awk '{if($0~"C") print $1" =6";if($0~"H") print $1" =1"; if($0~"H") print $1" =1"; if($0~"O") print $1" =8"; if($0~"P") print $1" =15"; if($0~"N") print $1" =7";}END{print "OW =6";print "HW1 =1";print "HW2 =1";} 
+#' > electrons.dat
+#Enumber=$(cat electrons.dat | awk '{sum=sum+1}END{print sum}')
+#echo $Enumber >> tmp.dat
+#cat electrons.dat >> tmp.dat 
+#mv tmp.dat electrons.dat
+
+mappingFILE='../../../mappingFILE.txt'
+
+/mnt/ssd/MATCH-git/scripts/centerTheBilayer.sh $mappingFILE ../topol.tpr ../traj_comp.xtc centered.xtc
+cp ../../../electronsLIPID.dat ./electrons.dat
+SOLname=$(grep M_SOL_M $mappingFILE | awk '{printf "%5s\n",$2}')
+SOLname=SOL
+echo $SOLname | gmx density -f centered.xtc -s ../topol.tpr -ei electrons.dat -dens electron -o electronDENSITYsol.xvg -xvg none -sl 100
+LIPIDname=$(grep M_POPC_M $mappingFILE | awk '{printf "%5s\n",$2}')
+LIPIDname=POPC
+echo $LIPIDname | gmx density -f centered.xtc -s ../topol.tpr -ei electrons.dat -dens electron -o electronDENSITYlipid.xvg -xvg none -sl 100
+#cp ../electronsCHOL.dat ./electrons.dat
+#CHOLname=$(grep M_CHOL_M $mappingFILE | awk '{printf "%5s\n",$2}')
+#echo $CHOLname | gmx density -f centered.xtc -s ../topol.tpr -ei electrons.dat -dens electron -o electronDENSITYchol.xvg -xvg none -sl 100
+if [ -e "electronDENSITYchol.xvg" ];
+then
[…]

### 2017-08-07 Data/POPC/no-ions/Lipid14/sim22a_0mM_NaCl_OPC4_non-tail-atoms_q-sc_0.80_sig-sc_0.89_noIons-ref/form_f_analysis/makeREPORT.sh

@@ -1,0 +1,42 @@
+#Location of data
+#ELECTRONS_FILE=/wrk/ollilas1/HGmodel/NMRlipids/NmrLipidsCholXray/scratch/POPCberger/electronsBERGER.dat
+CALC_FORM_FACTOR=/mnt/ssd/scaledip/scripts/MATCH/scratch/report/calc_FORM_FACTOR.sh
+REPORT_TEMPLATE=/mnt/ssd/scaledip/scripts/MATCH/scratch/report/report.tex
+CALC_ORDER_PARAMETERS=/mnt/ssd/scaledip/scripts/MATCH/scratch/report/calc_ORDER_PARAMETERS.sh
+GRO_OP=/mnt/ssd/scaledip/scripts/MATCH/scratch/report/gro_OP.awk
+
+cp $REPORT_TEMPLATE ./
+
+echo '0 0' > Electron_Density_From_Simulation.dat
+echo '0 0' > Form_Factor_From_Simulation.dat
+echo 'label  Order_Parameter_1  Order_Parameter_2
+0 0 0' >  Headgroup_Glycerol_Order_Parameters_Simulation.dat
+echo 'label  Order_Parameter_1  Order_Parameter_2
+0 0 0' > sn-1_Order_Parameters_Simulation.dat
+echo 'label  Order_Parameter_1  Order_Parameter_2
+0 0 0' > sn-2_Order_Parameters_Simulation.dat
+
+#mkdir ANALYSISdirectory
+#cp $TRAJECTORY ANALYSISdirectory/trajectory.trr
+#cp $ELECTRONS_FILE ANALYSISdirectory/electrons.dat
+#cp $CALC_FORM_FACTOR ANALYSISdirectory/
+#cp $CALC_ORDER_PARAMETERS ANALYSISdirectory/
+#cp $GRO_OP ANALYSISdirectory/
+
+#Calculate Form Factor from simulations
+#cd ANALYSISdirectory 
+#sh calc_FORM_FACTOR.sh
+#cd ..
+
+#Update Report
+#pdflatex report.tex
+
+#Calculate order parameters
+#cp ffgmx2.hdb ANALYSISdirectory
+#cd ANALYSISdirectory
+#chmod a+x ./calc_ORDER_PARAMETERS.sh
+#./calc_ORDER_PARAMETERS.sh
+#cd ..
+
+#Update Report
+pdflatex report.tex

### 2017-08-07 Data/POPC/no-ions/Lipid14/sim22a_251mM_CaCl-2s_OPC4_non-tail-atoms_q-sc_0.80_sig-sc_0.89/submit.sh

@@ -1,0 +1,197 @@
+#!/bin/bash
+
+#PBS -N sim22a_sim
+#PBS -q gpu
+#PBS -l walltime=12:00:00
+#PBS -l select=1:ncpus=8:ngpus=1:cl_doom=True:scratch_local=10gb:mem=20gb
+#PBS -j oe
+
+
+
+export stob3='/storage/brno3-cerit/home/melcrj'
+export storage=$stob3
+export workfolder=$PBS_O_WORKDIR   # $storage/wilCleGFP_3lipids_POPS/md12_dAFED/val-met_dist_dAFED
+
+export SCRATCH=$SCRATCHDIR
+
+
+# suggestion from https://wiki.metacentrum.cz/wiki/Pr%C3%A1ce_s_daty_v_%C3%BAloze
+trap 'cp -r $SCRATCH/* $workfolder/ && clean_scratch' TERM
+
+
+
+
+echo 'load modules ... '
+#  also loads gcc-4.8.1 automatically
+module load intelcdk-15 openmpi-1.8.2-gcc cuda-7.5 cmake-3.2.3 boost-1.57-gcc 
+
+
+
+########################################
+# load my build of Gromacs 5.1.2 with Plumed2
+########################################
+
+mpirun_command=""
+
+if [ $PBS_NUM_NODES -eq 1 ]
+then
+	#
+	# don't use MPI (intel-15 compiler)
+	#
+	# if "avx" is not contained in /proc/cpuinfo, load sse4.1 version of Gromacs
+	if [ `grep -e 'avx' /proc/cpuinfo -c` -eq 0 ]
+	then
+	   # sse4.1
+	   source $stob3/prog/BIN/gromacs-5.1.2+plumed2_ownFFTW_noMPI_intel15_sse4/bin/GMXRC
+	   export gmx_mdrun="gmx mdrun"
+	else
+	   # avx_256 mdrun only
+	   export PATH=$stob3/prog/BIN/gromacs-5.1.2+plumed2_ownFFTW_noMPI_intel15_avx_256/bin:$PATH
+	   export gmx_mdrun="mdrun"
+	fi
+else
+	mpirun_command="mpirun -n" $PBS_NUM_NODES
+	#
+	# use MPI (gcc 4.8.1 compiler)
+	#
+	# if "avx" is not contained in /proc/cpuinfo, load sse4.1 version of Gromacs
+	if [ `grep -e 'avx' /proc/cpuinfo -c` -eq 0 ]
+	then
+	   # sse4.1
[…]

### 2017-08-07 Data/POPC/no-ions/Lipid14/sim22a_978mM_NaCl_OPC3_non-tail-atoms_q-sc_0.80_sig-sc_0.89/rename_residue_POPC_lipid14_palm_ole.py

@@ -1,0 +1,36 @@
+#
+# simple python script for conversion of 
+# Lipid14 POPC residue names with duplicate atom names 
+# to a split-residue convention: 
+#   POPC will be split into 3 virtual residues with names
+#   PALM - POPC - OLE
+#   so that duplicate atom names can be distinquished at least
+#   by a different residue name
+#
+# script assumes the particular ordering as above with 
+# the head group segment beginning with O12 and ending with O22
+# the rest (begin/end) being the tails PALM/OLE.
+
+import MDAnalysis as mda
+
+# conf_init.gro should always be present in the sim dir
+u = mda.Universe("conf_init.gro")
+# find some POPC residue
+for r in u.residues:
+    if r.name == "POPC":
+        some_popc_res = r
+        break
+# find the indices of the first/last head group atoms
+headgr_begin_index = some_popc_res.names.tolist().index("O12")
+headgr_end_index   = some_popc_res.names.tolist().index("O22")
+
+# resname the resnames of PALM and OLE segments
+for r in u.residues:
+    if r.name == "POPC":
+        for a in r.atoms[:headgr_begin_index]:
+            a.resname = "PALM"
+        for a in r.atoms[headgr_end_index+1:]:
+            a.resname = "OLE"
+
+# write the new topology out
+u.SYSTEM.write("conf_palm.gro")

### 2017-08-08 Manuscript/manuscript.tex
Manuscript update after group discussion. Concentration corrected to actual values, not nominal. Ord.Pars include exp. values for tails. Updated Ca-density, included water profile.

@@
 cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
-and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} \todo{We need to select references for these}.
+and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} 
+\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
 While relative binding affinity of different ions is agreed to follow Hoffmeister
@@
 Molecular level interpretation of the results, however, requires assumptions
-about binding model and is often not fully conclusive \cite{??}.
+about binding models and is often not fully conclusive \cite{??}.
 On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
@@
 may lead to articifially positively charged memranes and significant artefacts in
-MD simualtions (e.g.~\ce{Ca^2+} induced charge inversion of bacterial membranes~\cite{Luan2016}).
+MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016} \todo{JOE: this may be a too daring statement.}).
 
@@
 This approach has been previously shown to improve performace of
-ion models against neutron scattering data in concentrated solutions \cite{Jungwirth2015,kohagen14,kohagen16}
-and ab-initio simulations~\cite{Jungwirth2017}. 
+ion models against neutron scattering data in concentrated solutions 
+and ab-initio simulations~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
 The realistic structure of ions in a bulk solvent was not, however, sufficient to correct binding in lipid
@@
 to x-ray scattering data and its head group structure to NMR experiments. 
-We will then quantify the accuracy of head group order parameter response to bound charge
+We will then quantify the accuracy of the head group order parameter response to the bound charge
 and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
@@
 to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
-it is not clear if the same factor should be used for lipid head groups. 
+it is not clear if the same factor should be used for molecules, i.e. lipid head groups in our case. 
 Unlike the total charge of cations and anions, atomic partial charges used in
@@
 can in principle adopt arbitrary values. 
-Here we estimate the correct scaling factor for Lipid14 model by using
-so called ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013},
-which are the average of partial charges given by 
+In cases where the partial charges have some physical meaning 
+(e.g. approximate electrostatic potential as in RESP~\cite{RESP_paper}),
+we can merely estimate the scaling factor for molecules, $f_q$, 
+to adopt a value between $\epsilon _{el} ^{-1/2}$ and $1$.
+Searching in this parameter space, we empirically discovered that
+the scaling factor $f_q = 0.8$ is optimal for applying ECC on Lipid14 model. 
+Such a finding is in line with the estimate considering the
+so called ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013}.
+IPolQ charges are obtained as the average of partial charges given by 
 restrained electrostatic potential fitting (RESP~\cite{RESP_paper}) 
-in vacuum and in solvent.
+in vacuum and in a solvent.
 Using the partial atomic charges obtained from RESP fitting in vacuum and implicit solvent reported in \cite{maciejewski14} and
 considering that the partial charges in original Lipid14 parametrization were derived using RESP~\cite{RESP_paper} in vacuum,
-we conclude that the charges in Lipid14 model shall be scaled roughly by the factor $f_q=0.8$. 
+we find out that the charges in Lipid14 model shall be scaled roughly by the factor $f_q=0.8$ 
+-- agreeing with our empirical finding. 
 %By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
@@
 %we exclude alkyl tails from our modifications.
-The scaling of partial charges in the polar region was found to decrease
-area per molecule to 60~\AA$^2$, which is significantly smaller than the
-experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment}) and the original Lipid14 values $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}..  
+
+The scaling of partial charges in the polar region was found to 
+significantly decrease cation binding and
+the area per molecule to $\approx$60~\AA$^2$, which is significantly smaller than the
+experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment}) and the original Lipid14 values $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
 The decrease of area was found to arise from a lower hydration of the lipid head group region,
@@
 \begin{equation}
+\label{eq:conc}
   C_{eq}=\frac{C_{plateau}}{0.602},
@@
 
-\subsection{Comparison of lipid bilayer structure to experimental data}
+\subsection{Comparison of lipid bilayer structure to experiments}
 
@@
   \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+  \includegraphics[width=7.0cm]{../Fig/POPCstructure.eps}
   \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.png}
@@
     compared with experimental order parameters from \cite{ferreira13}.}
-  \todo{Add acyl chain order parameters, POPC chemical structure} \\
   \todo{We should run a simulation for the system withtout ions with the
@@
 In order to validate the newly developed model, ECC-lipids, 
-we compared our simulation results without ions to
+we compared our simulation results without any ions to
 NMR order parameters measurements and x-ray scattering form factors 
@@
 The head group and glycerol backbone order parameters accuracy is 
-comparable to the state of art lipid models available in literature \cite{ollila16},
-The agreement between the x-ray scattering form factors and areas per molecule from simulations and experiments
+comparable to the state of art lipid models available in literature \cite{ollila16}.
+The agreement between the x-ray scattering form factors and the areas per molecule from simulations and experiments
 confirm that the membrane structural properties are well captured. 
+A structural comparison of ECC-lipids with Lipid14 can be found 
+in SI along with results with other water models. 
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
   \caption{\label{fig:cacl-dens}
-    Density profiles of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
-    and with ECC-ions and ECC-lipids model with ECC-ions. }
-  \todo{Change the figure so that it contains a membrane background} \\
+    Profiles of normalized number density of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
+    and with ECC-ions and ECC-lipids model with ECC-ions. 
+    Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}), 2 for divalent ions (i.e.~\ce{Ca^2+}), and $1/100$ for water. 
+    }
+  \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
+  \todo{JOE: Change the figure so that it contains a membrane background} \\
   \todo{What are the concentrations (calculated as in NMRlipids II) in this figure?} \\
@@
 
-After achieving accurate head group order parameter response with cationic surfactant, 
-we advanced to the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}. 
+After achieving accurate head group order parameter response with a cationic surfactant, 
+we advanced to quantifying the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}. 
 Headgroup order parameter response to increasing CaCl$_2$ concentration
+(as defined in Equation~\ref{eq:conc})
 from experiments, original Lipid14 model and ECC-lipids model are shown
@@
 show a pronounced peak in the position of the phosphate moieties of POPC. 
-The use of a ECC-ion model~\cite{kohagen14,kohagen16} 
-or the standard Amber ion model by \AA{quist} 
+The use of a ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
+or the standard Amber ion model by \AA{qvist}~\cite{aqvist90}
 along with Lipid14 does not significantly change it 
@@
 \todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
-The new ECC-lipids model with ECC-ions exhibits on the other hand smaller density in this region 
+The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16} exhibits on the other hand smaller density in this region 
 suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
@@
   \caption{\label{fig:cacl-bind}
-    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .}
+    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .
+    }
 \end{figure}
 
-Good agreement of ECC-lipids model with experiments encourages us to analyse the
-binding details from MD simulations. Direct analysis of contacts between ions and
+The good agreement of ECC-lipids model with experiments encourages us to analyse the
+molecular binding details from MD simulations. Direct analysis of contacts between ions and
 lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
-have stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
+has stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
 As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
@@
 The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. 
+We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
+which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
 

### 2017-08-08 Manuscript/manuscript.tex

@@
 
-Ion density profiles from simulations with different models are compared in Fig.~\ref{fig:cacl-dens}.
+Based on the accurate head group order parameter response of the new model, ECC-lipids, (Fig.~\ref{OrderParameterCHANGESnewMODELS})
+we suggest that it also exhibits accurate ion binding affinity to zwitterionic phospholipid membranes.
+This is directly related to the amount of surface excess charge 
+as shown along with ion density profiles in Fig.~\ref{fig:cacl-dens}.
 The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
@@
 suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
-Together with the accurate head group order parameter response (Fig.~\ref{OrderParameterCHANGESnewMODELS}),
-we can conclude that the present model, ECC-lipids, exhibits accurate ion binding affinity to zwitterionic phospholipid membranes
-suggesting that the models that overestimate head group order parameter response~\cite{catte16} also overestimate the amount of bound cations.
+It is then likely that the models that overestimate head group order parameter response~\cite{catte16} also overestimate the amount of bound cations.
 

### 2017-08-08 scripts/order_parameters_calculate.sh
Added P-N vector calculation and density analysis into the main analysis script and pythonic programs.

@@ -14,8 +14,9 @@ traj_file_name="traj_comp.xtc" #"../traj.trr"
 traj_pbc_nonwat_file_name="traj_nonwat_pbc.xtc" #"../traj.trr" 
 top_file_name="last_frame_nonwat.gro"
 tpr_file_name="topol.tpr"
+dens_file_name="density_ca_cl_water.xvg"
 #op_def_file="../../Headgroup_Glycerol_OPs.def"
-op_def_file="../../order_parameter_definitions_POPC_all.def"
+op_def_file=${scriptdir}"/order_parameter_definitions_Lipid14_POPC_all.def"
 op_out_file="OrdPars.dat"
 top="topol.top"
 f_conc=55430  # in mM/L
@@ -27,7 +28,9 @@ then
 fi
 
 # remove PBC:
-! [ -s $traj_pbc_nonwat_file_name ] && echo non-water | gmx trjconv -f $traj_file_name -s topol.tpr -o $traj_pbc_nonwat_file_name -pbc mol #-n index
+# center the trajectory around POPC (should be no. 2)
+# and start at 100ns
+! [ -s $traj_pbc_nonwat_file_name ] && echo 2 non-water | gmx trjconv -f $traj_file_name -s topol.tpr -o $traj_pbc_nonwat_file_name -pbc mol -center -b 100000 #-n index
 
 # get a non-water gro-file (topology)
 if ! [ -s $top_file_name ] 
@@ -41,22 +44,34 @@ then
     fi
 fi
 
+# get densities of cation - anion - water centered around POPC
+# and start at 100ns
+seq 2 5 | gmx density -sl 900 -dens number -ng 3 -f $traj_file_name -center -symm -relative -o $dens_file_name -xvg none -b 100000 
+
 # rename resnames of palmitoyl and oleoyl segments 
 python $scriptdir/rename_residue_lipid14_to_PALM-POPC-OLE.py -i $top_file_name -o $top_file_name 
+
 #CALCULATE ORDER PARAMETERS
 python $scriptdir/calcOrderParameters.py -i $op_def_file -t $top_file_name -x $traj_pbc_nonwat_file_name -o $op_out_file && rm $traj_pbc_nonwat_file_name
 
+# getting actual bulk concentration from number density profile
+python $scriptdir/get_conc_ion_bulk.py 
+conc=`cut -d " " -f1 conc_ion_bulk_mmolL.dat`
+# write the actual bulk concentration into the OP-output file
[…]

### 2017-08-09 Manuscript/manuscript.tex

@@
   }
+  \todo{This will be replaced with new figure with surfactants and P-N vector analysis when it is done.}
 \end{figure*}
@@
 \subsection{POPC membrane and electrometer concept with cation binding}
+
+\begin{figure*}[tbp]
+  \centering
+  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2.eps}
+  \caption{\label{OrderParameterCHANGESnewMODELS}
+    Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration
+    from simulations with different force fields and experiments (DPPC \cite{akutsu81} and POPC \cite{altenbach84}).
+    Simulations with Lipid14 parameters and Aqvist ions are directly from Ref. \cite{catte16}, except that
+    the concentration are determined with Eq. \ref{eq:conc}.
+  }
+  \todo{Simulation results with original Slipids and Dang ions to be added when we have the results.}
+\end{figure*}
 

### 2017-08-09 scripts/calcOrderParameters.py
Order parameters python script has one dirty hack mended (P-N vector calculation).

@@ -101,19 +101,18 @@ class OrderParameter:
         return S
 
 
-    def calc_angle(self, atoms):
+    def calc_angle(self, atoms, z_dim=45.0):
         """
         calculates the angle between the vector and z-axis in degrees
         no PBC check!
-        assuming POPC centred around z=0 (discriminates leaflets)
+        assuming a sim-box-centred membrane --> it's centre ~ z_dim/2
         """
         vec = atoms[1].position - atoms[0].position
         d = math.sqrt(np.square(vec).sum())
         cos = vec[2]/d
         # values for the bottom leaflet are inverted so that 
         # they have the same nomenclature as the top leaflet
-        ## DIRTY HACK! HARD CODED NUMBER 45 -- approximate centre of the bilayer
-        cos *= math.copysign(1.0, atoms[0].position[2]-45.0)
+        cos *= math.copysign(1.0, atoms[0].position[2]-z_dim*0.5)
         try:
             angle = math.degrees(math.acos(cos))
         except ValueError:
@@ -174,7 +173,7 @@ def read_trajs_calc_OPs(ordPars, top, trajs):
         for op in ordPars.values():
             for residue in op.selection:
                 if "vec" in op.name:
-                    S = op.calc_angle(residue)
+                    S = op.calc_angle(residue, z_dim=frame.dimensions[2])
                 else:
                     S = op.calc_OP(residue)
                 op.traj.append(S)

### 2017-08-09 Manuscript/manuscript.tex
New version of density plot and updates on text (I am still working on it)

@@
 may lead to articifially positively charged memranes and significant artefacts in
-MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016} \todo{JOE: this may be a too daring statement.}).
+MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
+\todo{JOE: this may be a too daring statement. \\
+  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
+  but it might happen with too small concentrations. 
+  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).
 
@@
 
-The main goal of this work is to apply ECC correction to accurately 
-describe the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
-in MD simulation models when compared with NMR data \cite{catte16}. 
-The Lipid14~\cite{dickson14} force field parameters were used as a starting point,
-because their response to bound ions was most realistic against NMR data
-(see Fig.~5 in Ref.~\citenum{catte16}).
+While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
+to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
+it is not clear if the same factor should be used for partial charges in molecules,
+i.e. lipids in our case. Unlike the total charge of ions, atomic partial charges in
+classical MD simulations are not experimental observables.
+Instead they are based on theoretical calculations and the
+derived values depend, for example, on the conformational
+sampling of the molecules and description of solvent 
+in the calculations \cite{RESP_paper,??}. 
+%In cases where the partial charges have some physical meaning 
+%(e.g. approximate electrostatic potential as in RESP~\cite{RESP_paper}),
+Depending to the details, partial charge calculations may partially
+incude some of the solvent electronic polarizability effects. Thus, we expect that
+the correct ECC scaling factor, $f_q$, for molecular partial charges would lie
+between $\epsilon _{el} ^{-1/2}$ (no electronic polarizability in calculation)
+and $1$ (all electronic polarizability in calculation).
+
+In this work, we empirically sample this parameter space
+to find a atomistic MD simulation force field parameters for lipids that
+accurately describes the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
+when compared with NMR data \cite{catte16}. This data can used to accurately asses
+ion binding in PC bilayers, as discussed in section \ref{section:electrometer}.
+We chose the Lipid14~\cite{dickson14} force field parameters as a starting point,
+because their response to bound ions was apparently most realistic against NMR data
+in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
 Also glycerol backbone and head group structures in Lipid14 model were
 relatively realistic when compared with other state of the art lipid models \cite{botan15}.
+The ECC correction was applied to Lipid14 parameters by scaling partial charges of the headgroup,
+glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
+are expected to have the largest contribution to the cation binding.
+The hydrocarbon chain parameters are not modified, because they are
+already highly optimized and give generally a good description for
+hydrophobic part of lipid bilayers in various conditions
+in most lipids models, including Lipid14 \cite{ollila16}.
+In contrast, improvements in glycerol backbone and headgroup
+parameters are required in all available lipid models~\cite{botan15}.
 
-While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
-to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
-it is not clear if the same factor should be used for molecules, i.e. lipid head groups in our case. 
-Unlike the total charge of cations and anions, atomic partial charges used in
-classical MD simulations are not experimental observables and 
-can in principle adopt arbitrary values. 
-In cases where the partial charges have some physical meaning 
-(e.g. approximate electrostatic potential as in RESP~\cite{RESP_paper}),
-we can merely estimate the scaling factor for molecules, $f_q$, 
-to adopt a value between $\epsilon _{el} ^{-1/2}$ and $1$.
-Searching in this parameter space, we empirically discovered that
-the scaling factor $f_q = 0.8$ is optimal for applying ECC on Lipid14 model. 
-Such a finding is in line with the estimate considering the
-so called ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013}.
-IPolQ charges are obtained as the average of partial charges given by 
-restrained electrostatic potential fitting (RESP~\cite{RESP_paper}) 
-in vacuum and in a solvent.
-Using the partial atomic charges obtained from RESP fitting in vacuum and implicit solvent reported in \cite{maciejewski14} and
-considering that the partial charges in original Lipid14 parametrization were derived using RESP~\cite{RESP_paper} in vacuum,
-we find out that the charges in Lipid14 model shall be scaled roughly by the factor $f_q=0.8$ 
--- agreeing with our empirical finding. 
-%By taking the charges of oxygen atoms in vacuum and implicit water solvent for POPC from \cite{maciejewski14}, 
-%we represent the IPolQ charges in the electronic continuum correction of the Lipid14 model
-%by increasing the scaling factor $f_q$ to $0.8$. 
+Exploring different scaling factor values, applied to headgroup, glycerol backbone
+and carbonyl charges of Lipid14 model, we found out that ion binding and
+related headgroup order parameter response became weaker with the scaling of partial
+charges. The optimal behaviour of ion binding was observed with
+the scaling factor $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
+Interestingly, this scaling factor is in line with the estimate
+given by ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013}
+combined with RESP calculations in vacuum and implicit solvent reported
+in \cite{maciejewski14}. IPolQ charges are obtained as the average of
+partial charges given by RESP calculation~\cite{RESP_paper}
+in vacuum and in a solvent. Applying the scaling factor of 0.75 to
+IPolQ charges calculated from the data in Ref.~\cite{maciejewski14},
+gives similar partial charges to ones obtained by scaling Lipid14 charges
+with a factor 0.8.
+\todo{SAMULI: I am still not sure if this is correct and/or makes sense.}
 %-----\footnote----------
@@
 
-Here we apply the ECC correction with the proposed scaling factor only to the head group,
-glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
-are expected to have the largest contribution to the cation binding.
-The hydrocarbon chain parameters are not modified, because they are
-already highly optimized in most lipids models including Lipid14 and
-give generally a good description for hydrophobic part of lipid bilayers
-in various conditions \cite{ollila16}.
-In contrast, all available lipid models require
-some improvements in glycerol backbone and head group regions \cite{botan15}. 
-%To minimize the detuning of the highly optimized hydrocarbon
-%parameters, 
-%Because the biologically relevant cations like \ce{Na^+} and \ce{Ca^{2+}}
-%are found to interact only with the phospholipid head group,
-%and the accuracy of alkyl sidechains in Lipid14 model are satisfactory, 
-%we exclude alkyl tails from our modifications.
+While, the charge scaling improved the behaviour of lipid-ion interactions,
+it reduced the area per molecule of lipid bilayer without ions below experimental
+values. Simulations with Lipid14 parameters having partial charges of headgroup, glycerol
+backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
+which is significantly smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
+and the original Lipid14 value $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
+The decrease of area was found to arise from a lower hydration of the lipid head group region,
+which can be explained by the decreased solvation free energy due to the lower polarity
+of molecules with scaled charges. The hydration can be increased
+by reducing the effective radius of atoms by changing the 
+the $\sigma$ parameters in Lennard-Jones potential for the selected atoms.
+This increases solvation free energy by allowing water molecules to
+approach closer to lipid atoms and have stonger electrostatic interactions with them.
+After reducing the $\sigma$ parameters of the same atoms for which charges were scaled
+with a scaling factor of $f_\sigma = 0.89$, the area per molecule value was
+againg in agreement with experimental value (see Table~\ref{tab:apls}).
 
-The scaling of partial charges in the polar region was found to 
-significantly decrease cation binding and
-the area per molecule to $\approx$60~\AA$^2$, which is significantly smaller than the
-experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment}) and the original Lipid14 values $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
-The decrease of area was found to arise from a lower hydration of the lipid head group region,
-%\todo{SAMULI: Did you analyze the effect of this to hydration or did we only speculate?
-%JOE: I found a systematic decrease of water density in the heagroup region even if I kept APL constant.}
-which can be explained by the decreased solvation free energy due to the lower polarity
-of molecules with scaled charges. 
-The solvation free energy and hydration is sensitive to Van der Waals interaction parameters,
-hence, they can be increased by reducing the radius of the selected atoms by changing the 
-the $\sigma$ parameters in Lennard-Jones potential. 
-%This allows water molecules to approach closer to lipid atoms and
-%have stonger electrostatic interactions with them.
-%\todo{SAMULI: This effect may have an official name. In that case we should mention it.}
-Here we reduce $\sigma$ with a scaling factor of $f_\sigma = 0.89$ for the same atoms for which
-charges were scaled. This has lead to an increase of the area per molecule back
-close to the experimental value (see Table~\ref{tab:apls}). 
-%\todo{Is there some justification for the below sentence?} 
-%In line with the electronic continnum correction,
-%the scaling factor for sigma parameters, $f_\sigma$,
-%has to lay between $f_q$ and 1.
-
-\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept}
+\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept} \label{section:electrometer}
 Ion binding was compared between experiments and simulations by 
@@
 with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
-Order parameters for C-H bonds can be accurately measured for each
-C-H bond of lipids by using H$^2$ NMR or $^{13}$C NMR techniques and
-calculated from MD simulations directly from definition
+Order parameters for all C-H bonds in lipid molecules
+can be accurately measured by using H$^2$ NMR or $^{13}$C NMR techniques.
+The same order parameters can be calculated from MD simulations directly
+from the definition
 \begin{equation}\label{OP}
@@
 and the value of the quadrupole coupling constant is $\chi \approx$\,167\,kHz.
-Combination of atomic absorption spectra, $^2$H NMR experiments
-and information about order parameter signs gave
+Atomic absorption spectra and $^2$H NMR data, combined with the
+information about order parameter signs gave
 $m_\alpha=-20.5$  and $m_\beta=-10.0$ for Ca$^{2+}$ binding to POPC
@@
 qualitatively also in simulations and can be used to compare ion binding
-affinitity between simulations and experiments~\cite{catte16}. However,
-it also turned out that the sensitivity of the order parameters response
-to the bound charge has to be quantified more accurately for quantitative comparison
-of binding affinity. Hence, we first quantify the response of
-head group order parameters against experimental data from cationinc surfactants
-embedded in PC bilayer~\cite{scherer89}. Essentially all surfactants locate in
-lipid bilayers in these experiments, thus the amount of the bound charge is known
-and can be plotted against order parameter change.
-Secondly we evaluate the response of the head group order parameters to varying salt concentrations in solution
-and estimate the amount of membrane bound charge.
+affinitity between simulations and experiments~\cite{catte16}.
+However, only systems with lipid bilayers and aqueous cations were used in that
+study. The headgroup response to bound charge then depends on both, the
+binding affinity and the response of the headgroup on bound charge.
+Thus, the headgroup response to bound charge has to be correctly
+described in the model if the concept is used for quantitative studies of
+binding affinity. Hence, we first quantify the response of head group order
+parameters to the bound charge by using experimental data measured from mixtures of cationinc surfactants
+and PC lipids~\cite{scherer89}. The amount of the bound charge is known excatly in
+these systems, because essentially all surfactants with unit charge locate in
+lipid bilayers. Thus, the headgroup order parameter response to bound charge can be first
+quantified aginst experimental data before proceeding to binding affinity studies.
+%Secondly we evaluate the response of the head group order parameters
+%to varying salt concentrations in solution
+%and estimate the amount of membrane bound charge.
 
-%\todo{ongoing,Actual concentration of cations in simulation has yet to be estimated. 
-%If it varies too much from the nominal concentration, I may need to tweak the scaling 
-%factors, $f_q$ or only $f_\sigma$, to accomodate it. However, it is very unlikely, response 
-%to the surfactant \ce{DHMDMAB} is OK. Big patches with loads of solvent are running at the 
-%moment to guide me on the possible finite-size errors and this conc-error.
-The used experimental data report order parameters as a function of
-equilibrium cation concentration in the bulk solvent \cite{akutsu81,altenbach84}.
-Such a condition is reached in simulations by adjusting
-the simulation box size to dimensions large enough
-that ion conceration reaches a clear plateau in the bulk solvent.
-The concentrations in the units of $\mathrm{mol/l}$ were then 
-determined as
+The experimental data used to quantify cation binding reports
+order parameters as a function of equilibrium cation concentration
+in the bulk solvent \cite{akutsu81,altenbach84}. Recent work
+by NMRlipids project estimated this concentration by using
+the molar concentration in water
+[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M,\\
+\todo{SAMULI: I call here concentration we used to call ''nominal'' as ''molar concentration in water''.
+  I think that the ''nominal'' is not a good term for this. We need to decide how we will call this.} \\
+which was accurate enough for the conclusions in that study \cite{catte16}.
+Since we are now targeting a model with quantitatively accurate binding
+affinity, we used larger simulation box dimensions that allow the
+determination of bulk ion conceration. Equilibrium concentration
+corresponding the experimental data in the units of $\mathrm{mol/l}$
+can be then determined from the ion concentration plateau in the bulk s
+olvent at the boundaries of the simulation box as
 \begin{equation}
@@
 dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
-The code AmberTools \cite{amber} was then used to generate the Amber-type force field
+AmberTools program \cite{amber} was then used to generate the Amber-type force field
 parameters. The parameters were converted to the Gromacs format by using
@@
 
-Headgroup order parameter response to the bound charge was evaluated against
-experimental data measured with the cationic surfactant
-(dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) \cite{scherer89}.
-The exact amount of bound charge in the membrane is known in these systems, because
-practically all cationic surfactant molecules are embedded in the lipid bilayer due to
-their amphiphilic nature. Thus, the changes of head group order parameter as a function of
-mole fraction of cationic surfactants (Fig. \ref{OrderParameterCHANGESnewMODELS}) gives
-also the order parameter changes as a function of the amount of bound cations.
-\todo{Split the OP-changes figure in two: cationic surfactant -and- binding of cations.
-      The separate figure with the cationic surfactant shall be additionally accompanied
-      with a plot of Head group tilt (PN-vector?) change/response. }
+Headgroup order parameter response to the amount of bound charge can
+be quantified by using ionic surfactants with a known valency~\cite{scherer89}.
+All surfactants locate in lipid bilayers due to their amphiphilic nature,
+thus the exact amount of bound charge in the membrane is known in these systems.
+The headgroup order parameter changes as a function of bound charges can be then
+explicitly determined from such systems.
 
+Changes of headgroup order parameter as a function of added cationic surfactant
+dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) from
+experiments~\cite{scherer89} and simulations are shown in
+Fig. \ref{OrderParameterCHANGESnewMODELS}).
 The head group order parameter response to the bound cation concentration
-is in experiments approximately linear up to $\sim$0.3 mole fraction of the cation \cite{scherer89}.
-The linearity is also observed in simulations with original Lipid14 and with newly derived ECC-lipids models.
+is approximately linear at least up to $\sim$0.3 mole fraction in experiments and
+in both simulation models, original Lipid14 and ECClipids.
+
+of the cation \cite{scherer89}.
+The linearity is also observed in simulations with original Lipid14 and with
+newly derived ECC-lipids models \todo{Yet to be shown.}.
 Quantitative comparison, however, reveals that the response is overestimated in 
@@
 underestimates the $\beta$ segment order parameter change.
-Such a finding is not surprising, as the $\Delta S^\alpha/\Delta S^\beta$ ratio 
-of the response of Lipid14 model is found to be slightly larger than in 
-experiments (VALexp vs VALsim
-\todo{Add values of $S^\alpha/S^\beta$ response for sim and experiment
-  SAMULI: Maybe these could be put in the same figure with the order parameter
-  response and P-N vector angle change with surfactants.
-}
-), and this property is also kept by the newly derived Lipid14-based model, ECC-lipids. 
+%Such a finding is not surprising, as the $\Delta S^\alpha/\Delta S^\beta$ ratio 
+%of the response of Lipid14 model is found to be slightly larger than in 
+%experiments (VALexp vs VALsim
+%\todo{Add values of $S^\alpha/S^\beta$ response for sim and experiment
+%  SAMULI: Maybe these could be put in the same figure with the order parameter
+%  response and P-N vector angle change with surfactants.
+%}
+%), and this property is also kept by the newly derived Lipid14-based model, ECC-lipids. 
 The present comparison of head group order parameter changes 
@@
   \centering
-  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities.eps}
+  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities2.eps}
   \caption{\label{fig:cacl-dens}
-    Profiles of normalized number density of \ce{Ca^{2+}} and \ce{Cl^-} for Lipid14 model with Dang ions
-    and with ECC-ions and ECC-lipids model with ECC-ions. 
-    Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}), 2 for divalent ions (i.e.~\ce{Ca^2+}), and $1/100$ for water. 
+    Number density of \ce{Ca^{2+}} and \ce{Cl^-} as a function of membrane normal axis
+    for different force fields. Data for Lipid14 with Aqvist ions are taken directly from Ref. \citenum{catte16}
+    %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
+    Densities of~\ce{Cl^-}) and water are divided with 2 and 200, respectively, to visalize them
+    with the same scale as \ce{Ca^{2+}}. The molar concentration in water is 350~mM in all systems
+    presented here. 
     }
@@
   \todo{JOE: Change the figure so that it contains a membrane background} \\
-  \todo{What are the concentrations (calculated as in NMRlipids II) in this figure?} \\
-  \todo{These should be also compared to the results with Åqvist ions, which are shown
-    in NMRlipids II. NMRlipids II have results with 350mM and 1000mM. If we have the distributions with
-    the same concentrations Åqvist results could be added also in this figure. If not, the
-  comparison has to be made some other way.}
+  \todo{Resuls with Lipid14+Dang to be added once done.}
 \end{figure}

### 2017-08-09 Manuscript/manuscript.tex
Small updates

@@
 Depending to the details, partial charge calculations may partially
-incude some of the solvent electronic polarizability effects. Thus, we expect that
+include some of the solvent electronic polarizability effects. Thus, we expect that
 the correct ECC scaling factor, $f_q$, for molecular partial charges would lie
@@
 accurately describes the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
-when compared with NMR data \cite{catte16}. This data can used to accurately asses
-ion binding in PC bilayers, as discussed in section \ref{section:electrometer}.
+when compared with NMR data \cite{akutsu81,altenbach84}. This data can used to accurately asses
+ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
 We chose the Lipid14~\cite{dickson14} force field parameters as a starting point,
@@
 is approximately linear at least up to $\sim$0.3 mole fraction in experiments and
-in both simulation models, original Lipid14 and ECClipids.
+in both simulation models, original Lipid14 and ECClipids \todo{Yet to be shown.}.
 
-of the cation \cite{scherer89}.
-The linearity is also observed in simulations with original Lipid14 and with
-newly derived ECC-lipids models \todo{Yet to be shown.}.
 Quantitative comparison, however, reveals that the response is overestimated in 

### 2017-08-11 Manuscript/manuscript.tex

@@
 \subsection{Electronic continuum correction for lipid bilayers}
-The lack of electronic polarizability in the classical MD simulation
-models has been considered higly relevant issue since the early days of
-lipid bilayer simulations. Some pioneering simulation studies scaled
-the partial charges of atoms by one half to effectively include 
-polarizability \cite{jonsson86,egberts94}. Also, approaches that
-explicitly include polarizability has been introduced, but
-this has turned out to be practically complicated \cite{chowdhary13}.
-In this work we effectively include electronic polarizability in lipid
+The lack of electronic polarizability in the standard MD simulation
+force fields has been considered higly relevant issue since the early days of
+lipid bilayer simulations.
+%Some pioneering simulation studies scaled
+%the partial charges of atoms by one half to effectively include 
+%polarizability \cite{jonsson86,egberts94}. Also, approaches that
+%explicitly include polarizability has been introduced, but
+%this has turned out to be practically complicated \cite{chowdhary13}.
+In this work we circumvent the rather demanding explicit inclusion
+of electronic polariization effects \cite{chowdhary13} by effectively including
+electronic polarizability in lipid
 bilayer simulations by using the electronic continuum correction
@@
 the early studies where a scaling factor one half was used \cite{jonsson86,egberts94}. 
-In contrast to those studies, however, the concept of ECC is 
+However, the concept of ECC is 
 physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11},
@@
 for almost any biomaterial \cite{some_original_work, leontyev11}.
-Such a continuum can be easily included in standard MD simulation by
+Such a dielectric continuum can be easily included in standard MD simulation by
 a formal transformation of partial charges 
@@
 a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}, which has been
-successfully used to improve the performance of ion force fields \cite{kohagen14,kohagen16,??}.
+successfully used to improve the performance of force field for ions \cite{kohagen14,kohagen16,??}.
 
-While the scaling factor of $f_q = 0.75$ for ions in water is well justified and shown
-to improve model performance against scattering data \cite{kohagen14,kohagen16, Pluharova2014},
+While the scaling factor of $f_q = 0.75$ for ions in water is well justified and 
+improves model performance against experimental scattering data \cite{kohagen14,kohagen16, Pluharova2014},
 it is not clear if the same factor should be used for partial charges in molecules,
-i.e. lipids in our case. Unlike the total charge of ions, atomic partial charges in
+i.e, lipids in our case. Unlike the total charge of ions, atomic partial charges in
 classical MD simulations are not experimental observables.
@@
 the correct ECC scaling factor, $f_q$, for molecular partial charges would lie
-between $\epsilon _{el} ^{-1/2}$ (no electronic polarizability in calculation)
-and $1$ (all electronic polarizability in calculation).
+between $\epsilon _{el} ^{-1/2}$ (no electronic polarizability in charge calculation)
+and $1$ (full electronic polarizability in charge calculation).
 
@@
 to find a atomistic MD simulation force field parameters for lipids that
-accurately describes the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
-when compared with NMR data \cite{akutsu81,altenbach84}. This data can used to accurately asses
+accurately describe the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
+when compared with NMR data \cite{akutsu81,altenbach84}. This data can be used to accurately asses
 ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
@@
 by reducing the effective radius of atoms by changing the 
-the $\sigma$ parameters in Lennard-Jones potential for the selected atoms.
+the $\sigma$ parameters in Lennard-Jones potential for the selected atoms
+similarly as done for free ions in solution \cite{kohagen14,kohagen16,??}.
 This increases solvation free energy by allowing water molecules to
@@
 After reducing the $\sigma$ parameters of the same atoms for which charges were scaled
-with a scaling factor of $f_\sigma = 0.89$, the area per molecule value was
+with a factor of $f_\sigma = 0.89$, the area per molecule value was
 againg in agreement with experimental value (see Table~\ref{tab:apls}).
@@
   }
+
+The binding affinities between different models were compared by calculating
+the relative surface excess between water and ions $\Gamma_w^i$ \cite{chattorajBOOK}
+\begin{equation}
+  \Gamma_w^i=\Gamma_w^x-\Gamma_i^x\frac{c_w-c_w'}{c_i-c_i'},
+\end{equation}
+where $c_w$ and $c_i$ are concentrations in bulk water, 
+$c_w'$ and $c_i'$ are concentrations inside bilayer, and
+$\Gamma_w^x$ and $\Gamma_i^x$ are surface excess of water and
+ions, respectively. Here consider the whole simulation box
+as a interface region, thus the surface excesses
+are $\Gamma_w^x=n_w/A$ and $\Gamma_i^x=n_i/A$, where
+$A$ is the area of the simulation box, and $n_w$ and
+$n_i$ are total number of water and ions in simulation box,
+respectively. The concentrations of water and ions inside
+bilayer can be expected to vanish, thus $c_w'=c_i'=0$.
+Then we get the relative surface excess for lipid bilayer
+simulations
+\begin{equation}\label{surfexcess}
+  \Gamma_w^i=\frac{n_w}{A}-\frac{n_i}{A}\frac{c_w}{c_i},
+\end{equation}
+where concentrations in bulk, $c_w$ and $c_i$, can be determined
+from the simulation box boundary.
 
@@
 
-\subsection{POPC membrane and electrometer concept with cationic surfactant}
+\subsection{Response of headgroup on bound charge}
 
-\begin{figure*}[tbp]
+\begin{figure}[tbp]
   \centering
-  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2_surf.eps}
-  \caption{\label{OrderParameterCHANGESnewMODELS}
-    Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration and
-    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) mole fraction
-    from simulations and experiments (DPPC \cite{akutsu81}, POPC \cite{altenbach84}, surfactant \cite{scherer89}).
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.png}
+  \caption{\label{OrderParameterCHANGESsurf}
+    Headgroup order parameter changes and P-N vector orientation as a function of
+    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$)
+    in PC bilayer from simulations and experiments \cite{scherer89}.
   }
-  \todo{This will be replaced with new figure with surfactants and P-N vector analysis when it is done.}
-\end{figure*}
+  \todo{Labels are missing.} \\
+  \todo{Experimental data points are missing.}\\
+  \todo{Order should be the same as in other figures, i.e, $\beta$ segment on top.}
+\end{figure}
 
@@
 experiments~\cite{scherer89} and simulations are shown in
-Fig. \ref{OrderParameterCHANGESnewMODELS}).
+Fig. \ref{OrderParameterCHANGESsurf}).
 The head group order parameter response to the bound cation concentration
@@
 
-Quantitative comparison, however, reveals that the response is overestimated in 
-Lipid14 for both segments, while the ECC-lipids model gives a
-good agreement for the change of $\alpha$ segment order parameter, and slightly
-underestimates the $\beta$ segment order parameter change.
-%Such a finding is not surprising, as the $\Delta S^\alpha/\Delta S^\beta$ ratio 
-%of the response of Lipid14 model is found to be slightly larger than in 
-%experiments (VALexp vs VALsim
-%\todo{Add values of $S^\alpha/S^\beta$ response for sim and experiment
-%  SAMULI: Maybe these could be put in the same figure with the order parameter
-%  response and P-N vector angle change with surfactants.
-%}
-%), and this property is also kept by the newly derived Lipid14-based model, ECC-lipids. 
-The present comparison of head group order parameter changes 
-suggests that the observed overestimated response of all models to 
-increasing CaCl$_2$ concentration in \cite{catte16} 
-can be explained at least in part by a high sensitivity of the head group order parameter response
-to the bound charge.
+The slope is, however, too steep in Lipid14 model indicating that the
+headgroup order parameters are too sensitive for bound charge in this model.
+This has to be taken into account when interpreting the
+headgroup order parameter response to CaCl$_2$ concentration in Lipid14 model.
+%observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
+The ECC-lipids model gives a slope in very good agreement with experiments
+for the $\alpha$ segment, while the slope is slightly
+underestimated for the $\beta$ segment.
+%The present comparison of head group order parameter changes 
+%suggests that the observed overestimated response of all models to 
+%increasing CaCl$_2$ concentration in \cite{catte16} 
+%can be explained at least in part by a high sensitivity of the head group order parameter response
+%to the bound charge.
 
-\subsection{POPC membrane and electrometer concept with cation binding}
+The headgroup P-N vector angle is shown in Fig.~\ref{??} as
+function of bound cations.
+As expected~\cite{seelig87}, the headgroup order parameter decrease due 
+to bound cations is related to the orientration of headgroup more parallel
+to membrane normal. The headgroup orientational effect is stronger
+in Lipid14 parameters than in ECClipid parameters, which is in line
+with smaller dipole moment and order parameter changes in ECClipid model.
+
+\subsection{Cation binding affinity in POPC}
 
@@
 \end{figure*}
+
+%After achieving accurate head group order parameter response with a cationic surfactant, 
+%we advanced to quantifying the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}.
+The headgroup order parameter response to aqueous ionic concentration
+can be used to measure ion binding affinity in experiments and simulations,
+as discussed in section \ref{??} and in previous literature~\cite{seelig87,catte16}.
+The headgroup order parameter response of PC lipids to NaCl and CaCl$_2$ concentrations
+from experiments (DPPC \cite{akutsu81} and POPC \cite{altenbach84}) and different simulation
+models are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+The simulation results of Lipid14 with Aqvist ions \cite{??} taken directly from
+Ref. \citenum{catte16} are shown together with the results from Lipid14 simulation
+with Dang ions \cite{??} and from ECClipid simulations with ECC-ions.
+
+As already discussed in the recent work by NMRlipids project~\cite{catte16},
+the headgroup order parameters are almost unaffected by addition of NaCl
+in Lipid14 simulations with Aqvist ion model due a very weak binding of Na$^+$
+in PC bilayer. This is in agremeent with experimental data, in contrast
+to almost all the other lipid models, which significantly overestimate
+Na$^+$ binding in PC bilayers~\cite{catte16}. Despite of good performance
+with NaCl, the response of headgroup order parameters to CaCl$_2$ concentration are
+significantly overestimated in Lipid14 simulations with Aqvist ion model,
+see Fig.~\ref{OrderParameterCHANGESnewMODELS} and \citenum{catte16}.
+
+Wide range of data presented in previous work \cite{catte16} suggested
+that CaCl$_2$ binding behaviour cannot be corrected by only improving
+ion models. This was not, however, tested for Lipid14 model, so we
+ran also Lipid14 simulations with Dang \cite{smith94,chang1999,dang2006} and ECC corrected \cite{??}
+ions having more realistic bulk behaviour as Aqvist model \cite{??}.
+In line with previous work, the headgroup order parameters response
+to CaCl$_2$ concentartion was overestimated also with these ion models,
+as seen in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively
+\todo{Add OP-response of Lipid14+Dang in Fig. \ref{OrderParameterCHANGESnewMODELS}
+  and OP-response of Lipid14+ECC-ions plot in SI}.
+
+Applying ECC correction also for lipids, introduced in section \ref{??},
+brings the order parameter response to CaCl$_2$ concentration to a
+good agreement with experiments, as shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+This is a significant improvement over all models introduced previously
+which in general overstimate the head group order parameter response to CaCl$_2$
+concentration \cite{catte16}. The good agreement with experiments indicates
+that the ECC corrected lipid model has a sufficient accuracy for realistic
+studies of ion binding details on lipid bilayers.
+
+Ion density profiles along membrane normal coordinate
+from different simulations are shown in Fig.~\ref{fig:cacl-dens}.
+\todo{This paragraph is to be finalized when we have the surface excess analysis
+  done by using Eq.~\ref{surfexcess}.}
+This is directly related to the amount of surface excess charge 
+as shown along with ion density profiles in  Fig.~\ref{fig:cacl-dens}.
+The density profiles from simulations with original Lipid14 \cite{dickson14}
+and Dang ions \cite{smith94,chang1999,dang2006} 
+show a pronounced peak in the position of the phosphate moieties of POPC. 
+The use of a ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
+or the standard Amber ion model by \AA{qvist}~\cite{aqvist90}
+along with Lipid14 does not significantly change it 
+(Headgroup order parameter responses for these models in SI).
+The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
+exhibits on the other hand smaller density in this region 
+suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
+It is then likely that the models that overestimate head group order parameter response~\cite{catte16}
+also overestimate the amount of bound cations.
 
@@
 \end{figure}
-
-After achieving accurate head group order parameter response with a cationic surfactant, 
-we advanced to quantifying the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}. 
-Headgroup order parameter response to increasing CaCl$_2$ concentration
-(as defined in Equation~\ref{eq:conc})
-from experiments, original Lipid14 model and ECC-lipids model are shown
-in Fig. \ref{OrderParameterCHANGESnewMODELS}. The order parameter response
-is significantly overestimated in original Lipid14 model, while results from
-ECC-lipids model are in a good agreement with the experiments.
-This is a significant improvement over current models,
-which in general overstimate the head group order parameter response to CaCl$_2$
-concentration \cite{catte16}. The good agreement with experiments indicate
-that the binding details of \ce{Ca^2+} are realistic in the ECC-lipids model
-and, hence, it can be used to study lipid-ion interaction in detail.
-
-Based on the accurate head group order parameter response of the new model, ECC-lipids, (Fig.~\ref{OrderParameterCHANGESnewMODELS})
-we suggest that it also exhibits accurate ion binding affinity to zwitterionic phospholipid membranes.
-This is directly related to the amount of surface excess charge 
-as shown along with ion density profiles in Fig.~\ref{fig:cacl-dens}.
-The density profiles from simulations with original Lipid14 \cite{dickson14} and Dang ions \cite{smith94,chang1999,dang2006} 
-show a pronounced peak in the position of the phosphate moieties of POPC. 
-The use of a ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
-or the standard Amber ion model by \AA{qvist}~\cite{aqvist90}
-along with Lipid14 does not significantly change it 
-(Headgroup order parameter responses for these models in SI).
-\todo{Add OP-response of Lipid14+ECC-ions into the OP-changes plot in SI}
-The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16} exhibits on the other hand smaller density in this region 
-suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
-It is then likely that the models that overestimate head group order parameter response~\cite{catte16} also overestimate the amount of bound cations.
 

### 2017-08-11 Manuscript/manuscript.tex

@@
 
-\subsection{Electronic continuum correction for lipid bilayers}
+\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
 The lack of electronic polarizability in the standard MD simulation
@@
 with a factor 0.8.
-\todo{SAMULI: I am still not sure if this is correct and/or makes sense.}
 %-----\footnote----------
@@
 approach closer to lipid atoms and have stonger electrostatic interactions with them.
-After reducing the $\sigma$ parameters of the same atoms for which charges were scaled
-with a factor of $f_\sigma = 0.89$, the area per molecule value was
-againg in agreement with experimental value (see Table~\ref{tab:apls}).
+After reducing the $\sigma$ parameters with a factor of $f_\sigma = 0.89$
+for the same atoms for which charges were scaled, the area per molecule value was
+again in agreement with experimental value (see Table~\ref{tab:apls}).
 
@@
 with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
-Order parameters for all C-H bonds in lipid molecules
+Order parameters for $\alpha$ and $\beta$ segments in headgroup, as well as for
+all C-H bonds in lipid molecules,
 can be accurately measured by using H$^2$ NMR or $^{13}$C NMR techniques.
-The same order parameters can be calculated from MD simulations directly
-from the definition
+The experimental order parameters can be compared with the the ones
+calculated from MD simulations by using the definition
 \begin{equation}\label{OP}
@@
 
-The change of the head group order parameters as a function of
+According to the electrometer concept,
+the change of the head group order parameters as a function of
 the amount of bound charge per lipid $X^\pm$ 
@@
 bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-
 Recent work published by the NMRlipids project showed that the concept works
@@
 affinitity between simulations and experiments~\cite{catte16}.
-However, only systems with lipid bilayers and aqueous cations were used in that
-study. The headgroup response to bound charge then depends on both, the
-binding affinity and the response of the headgroup on bound charge.
-Thus, the headgroup response to bound charge has to be correctly
-described in the model if the concept is used for quantitative studies of
-binding affinity. Hence, we first quantify the response of head group order
-parameters to the bound charge by using experimental data measured from mixtures of cationinc surfactants
-and PC lipids~\cite{scherer89}. The amount of the bound charge is known excatly in
-these systems, because essentially all surfactants with unit charge locate in
-lipid bilayers. Thus, the headgroup order parameter response to bound charge can be first
-quantified aginst experimental data before proceeding to binding affinity studies.
-%Secondly we evaluate the response of the head group order parameters
-%to varying salt concentrations in solution
-%and estimate the amount of membrane bound charge.
 
@@
 by NMRlipids project estimated this concentration by using
-the molar concentration in water
-[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M,\\
-\todo{SAMULI: I call here concentration we used to call ''nominal'' as ''molar concentration in water''.
-  I think that the ''nominal'' is not a good term for this. We need to decide how we will call this.} \\
+the relative molar concentration in water
+[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M,
 which was accurate enough for the conclusions in that study \cite{catte16}.
@@
 
-The binding affinities between different models were compared by calculating
+The ion binding affinities between different models were compared by calculating
 the relative surface excess between water and ions $\Gamma_w^i$ \cite{chattorajBOOK}
@@
 $\Gamma_w^x$ and $\Gamma_i^x$ are surface excess of water and
-ions, respectively. Here consider the whole simulation box
-as a interface region, thus the surface excesses
-are $\Gamma_w^x=n_w/A$ and $\Gamma_i^x=n_i/A$, where
-$A$ is the area of the simulation box, and $n_w$ and
+ions, respectively. The whole simulation box is considered here
+as interfacial region, thus $\Gamma_w^x=n_w/A$ and $\Gamma_i^x=n_i/A$,
+where $A$ is the area of the simulation box, and $n_w$ and
 $n_i$ are total number of water and ions in simulation box,
@@
 validates the dimensions of the lipid bilayer (thickness and area per
-molecule)~\cite{ollila16}. Form factor was calculated from simulations
-as 
+molecule)~\cite{ollila16}.
+
+Order parameters were calculated from simulations for all C-H bonds
+by using Eq. \ref{OP} and form factors as
 \begin{equation}
@@
 (POPC) bilayer and an aqueous salt solutions of varying concentrations.
-Water molecules were described by OPC3 model \cite{Izadi16}.
+Water molecules were described by OPC3 model \cite{Izadi16}
+\todo{Explain why were OPC3 chosen}.
 In order to test transferability of the newly developed ECC-lipids model, 
@@
 and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}. 
-Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}. \\
+\todo{System sizes has to reported somewhere}
 
@@
   \end{tabular}
+  \todo{This could be moved to SI. Only simulation lengths needs to be mentioned in the main paper.}
 \end{table}
@@
 mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
-These systems were ran with original Lipid14 POPC model and with ECC-corrected model.
+All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
+6, 14, 21, 35 or 50 surfactants per leaflet.
+These systems were first simulated 200~ns by using Lipid14 model.
+First 20~ns was omitted from the analysis.
+
+The same systems where also simulated with ECC-corrected lipi model...
 \todo{Describe the changes in surfactant model used with ECC-corrected lipids.}
@@
 
-\subsection{Response of headgroup on bound charge}
+\subsection{Response of POPC headgroup on bound charge}
 
@@
   \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.png}
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.eps}
   \caption{\label{OrderParameterCHANGESsurf}
@@
   \todo{Labels are missing.} \\
-  \todo{Experimental data points are missing.}\\
-  \todo{Order should be the same as in other figures, i.e, $\beta$ segment on top.}
+  \todo{Order should be the same as in other figures, i.e, $\beta$ segment on top.} \\
+  \todo{I would put x-axis from 0 to 51 and maybe zoom y-axis little bit as well.}
 \end{figure}
 
-Headgroup order parameter response to the amount of bound charge can
-be quantified by using ionic surfactants with a known valency~\cite{scherer89}.
-All surfactants locate in lipid bilayers due to their amphiphilic nature,
-thus the exact amount of bound charge in the membrane is known in these systems.
-The headgroup order parameter changes as a function of bound charges can be then
-explicitly determined from such systems.
+The headgroup response to bound charge depends on both, the
+binding affinity and the response of the headgroup on bound charge.
+Thus, the headgroup response to bound charge has to be correctly
+described in the model if the concept is used for quantitative studies of
+binding affinity. Hence, we first quantify the response of head group order
+parameters to the bound charge by using experimental data measured from
+mixtures of cationinc surfactants and PC lipids~\cite{scherer89}.
+The amount of the bound charge is known excatly in these systems,
+because essentially all surfactants with unit charge locate in lipid bilayers.
+%Thus, the headgroup order parameter response to bound charge can be first
+%quantified aginst experimental data before proceeding to binding affinity studies.
+%Headgroup order parameter response to the amount of bound charge can
+%be quantified by using ionic surfactants with a known valency~\cite{scherer89}.
+%All surfactants locate in lipid bilayers due to their amphiphilic nature,
+%thus the exact amount of bound charge in the membrane is known in these systems.
+%The headgroup order parameter changes as a function of bound charges can be then
+%explicitly determined from such systems.
 
@@
 Fig. \ref{OrderParameterCHANGESsurf}).
-The head group order parameter response to the bound cation concentration
+As expected from Eq. \ref{OPchangeEQ}, the head group order parameter response to the bound cation concentration
 is approximately linear at least up to $\sim$0.3 mole fraction in experiments and
 in both simulation models, original Lipid14 and ECClipids \todo{Yet to be shown.}. 
+\todo{We might want to calculate the slopes from simulations.}
 
@@
 
-The headgroup P-N vector angle is shown in Fig.~\ref{??} as
-function of bound cations.
-As expected~\cite{seelig87}, the headgroup order parameter decrease due 
-to bound cations is related to the orientration of headgroup more parallel
-to membrane normal. The headgroup orientational effect is stronger
+Also the angle of the headgroup P-N vector with respect to membrane normal
+is shown in Fig.~\ref{OrderParameterCHANGESsurf} as a function of bound cations.
+As expected~\cite{seelig87}, the headgroup orients more parallel with membrane
+normal with increasing amount of bound cations. The effect is stronger
 in Lipid14 parameters than in ECClipid parameters, which is in line
-with smaller dipole moment and order parameter changes in ECClipid model.
+with the reduced dipole moment and the lower sensitivity to bound charges
+in the ECClipid model.
+\todo{It might be a good idea to establish a relation between order parameter change
+and P-N vector angle by using ECClipids model.}
 
@@
 can be used to measure ion binding affinity in experiments and simulations,
-as discussed in section \ref{??} and in previous literature~\cite{seelig87,catte16}.
+as discussed in section \ref{section:electrometer} and in previous literature~\cite{seelig87,catte16}.
 The headgroup order parameter response of PC lipids to NaCl and CaCl$_2$ concentrations
@@
 models are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-The simulation results of Lipid14 with Aqvist ions \cite{??} taken directly from
-Ref. \citenum{catte16} are shown together with the results from Lipid14 simulation
+The simulation results of Lipid14 with Aqvist ions \cite{??} (from Ref. \citenum{catte16})
+are shown together with the results from Lipid14 simulation
 with Dang ions \cite{??} and from ECClipid simulations with ECC-ions.
@@
 
-Applying ECC correction also for lipids, introduced in section \ref{??},
+Applying ECC correction also for lipids, introduced in section \ref{section:ecc},
 brings the order parameter response to CaCl$_2$ concentration to a
 good agreement with experiments, as shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-This is a significant improvement over all models introduced previously
-which in general overstimate the head group order parameter response to CaCl$_2$
+This is a significant improvement over previously introduced models,
+which all overestimate the head group order parameter response to CaCl$_2$
 concentration \cite{catte16}. The good agreement with experiments indicates

### 2017-08-14 Manuscript/manuscript.tex
MS partial update (Intro and methods in part), added few data, improved ipynb plots.

@@
 series, the molecular details of binding and binding energetics are
-not fully understood \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
+not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
 %General conclusion from experimental
@@
 In this work we circumvent the rather demanding explicit inclusion
-of electronic polariization effects \cite{chowdhary13} by effectively including
+of electronic polariization effects \cite{chowdhary13} by implicitly including
 electronic polarizability in lipid
@@
 While the scaling factor of $f_q = 0.75$ for ions in water is well justified and 
-improves model performance against experimental scattering data \cite{kohagen14,kohagen16, Pluharova2014},
+improves the model performance against experimental scattering data \cite{kohagen14,kohagen16, Pluharova2014},
 it is not clear if the same factor should be used for partial charges in molecules,
-i.e, lipids in our case. Unlike the total charge of ions, atomic partial charges in
-classical MD simulations are not experimental observables.
-Instead they are based on theoretical calculations and the
-derived values depend, for example, on the conformational
-sampling of the molecules and description of solvent 
-in the calculations \cite{RESP_paper,??}. 
+i.e, lipids in our case. 
+Unlike the total charge of molecules, atomic partial charges within each molecule are not experimental observables.
+Although there have been proposed several schemes for the assignment of partial charges of molecules, 
+methods targetting at fitting of electrostatic potentials are 
+most popular among classical fixed charge force fields for biomolecules. \cite{Hu2007}
+Currently the most commonly employed scheme is the RESP scheme~\cite{RESP_paper,Singh1984}, 
+which fits the electrostatic potential from quantum mechanical calculations. 
+The resulting set of fixed partial atomic charges is hence 
+the best fit to electrostatic potentials of 
+an ensemble of conformations of a molecule. 
+Such a physical meaning is especially well suited for the application of ECC, 
+which targets at improving electrostatic interactions. 
+Depending to the details, partial charge calculations may also partially
+include some of the solvent electronic polarizability effects. 
+Thus, we expect that
+the correct ECC scaling factor, $f_q$, for molecular partial charges
+does not have to be striclty $f_q = \epsilon _{el} ^{-1/2}$, 
+but rather lies between this value (no electronic polarizability in the charge calculation)
+and $1$ (full electronic polarizability already included in the charge calculation).
+%
+% This is not really important for the paper:
+%
+%However, the RESP method has several remedies;  %and these are specific to RESP
+%RESP charges show a dependence on the orientation and position
+%and it does not provide numerically stable partial charges of burried atoms;~\cite{Hu2007}
+%the RESP-derived values depend, for example, on the conformational
+%sampling of the molecules and description of solvent 
+%in the calculations~\cite{RESP_paper, Singh1984, Hu2007, maciejewski14}. 
+%
+%Hirshfeld-E partitioning: \cite{Verstraelen2013}
 %In cases where the partial charges have some physical meaning 
-%(e.g. approximate electrostatic potential as in RESP~\cite{RESP_paper}),
-Depending to the details, partial charge calculations may partially
-include some of the solvent electronic polarizability effects. Thus, we expect that
-the correct ECC scaling factor, $f_q$, for molecular partial charges would lie
-between $\epsilon _{el} ^{-1/2}$ (no electronic polarizability in charge calculation)
-and $1$ (full electronic polarizability in charge calculation).
+%(e.g. approximate electrostatic potential as in RESP~~\cite{RESP_paper,Singh1984}),
+%
+%
 
-In this work, we empirically sample this parameter space
-to find a atomistic MD simulation force field parameters for lipids that
+In this work, we empirically sample this parameter space $f_q \in [0.75, 1.0]$
+to find an atomistic MD simulation force field parameters for lipids that
 accurately describe the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
-when compared with NMR data \cite{akutsu81,altenbach84}. This data can be used to accurately asses
+when compared with NMR data \cite{akutsu81,altenbach84}. These data can be used to accurately asses
 ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
-We chose the Lipid14~\cite{dickson14} force field parameters as a starting point,
+We chose Lipid14~\cite{dickson14} force field parameters as a starting point,
 because their response to bound ions was apparently most realistic against NMR data
@@
 and carbonyl charges of Lipid14 model, we found out that ion binding and
-related headgroup order parameter response became weaker with the scaling of partial
-charges. The optimal behaviour of ion binding was observed with
+related headgroup order parameter responses become weaker. 
+The optimal behaviour of ion binding was observed with
 the scaling factor $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
@@
 the $\sigma$ parameters in Lennard-Jones potential for the selected atoms
-similarly as done for free ions in solution \cite{kohagen14,kohagen16,??}.
+similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
 This increases solvation free energy by allowing water molecules to
@@
 Since we are now targeting a model with quantitatively accurate binding
-affinity, we used larger simulation box dimensions that allow the
-determination of bulk ion conceration. Equilibrium concentration
-corresponding the experimental data in the units of $\mathrm{mol/l}$
-can be then determined from the ion concentration plateau in the bulk s
-olvent at the boundaries of the simulation box as
+affinity, we rather use a definition that corresponds more closely to the 
+concentrations reported in experiments. 
+By using larger simulation box dimensions that allow the
+determination of bulk ion conceration, the equilibrium concentration
+corresponding to the experimental data in the units of $\mathrm{mol/l}$
+can be then determined from the simulations as
 \begin{equation}
 \label{eq:conc}
-  C_{eq}=\frac{C_{plateau}}{0.602},
+  C_{eq}=\frac{C_{bulk}}{0.602},
 \end{equation}
-where plateau concentration is the number density in the units of $\mathrm{nm}^{-3}$.
+where bulk concentration, $C_{bulk}$, is the number density in the units of $\mathrm{nm}^{-3}$.
 \todo{SAMULI: Once we have to final results, we can probably say that
@@
   \includegraphics[width=7.0cm]{../Fig/POPCstructure.eps}
-  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.png}
+  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
   \caption{\label{simVSexpNOions}

### 2017-08-15 Manuscript/manuscript.tex
Corrected surface excess Equation and statements around it. Surface excess was not written correctly, factor 1/2 missing in the final form for simulations, concentrations use same symbols as in previous paragraph. Corrected reference for chattorajBOOK.

@@
 By using larger simulation box dimensions that allow the
-determination of bulk ion conceration, the equilibrium concentration
+determination of bulk ion concentration, the equilibrium concentration
 corresponding to the experimental data in the units of $\mathrm{mol/l}$
@@
 \end{equation}
-where bulk concentration, $C_{bulk}$, is the number density in the units of $\mathrm{nm}^{-3}$.
+where the bulk concentration, $C_{bulk}$, is the number density in the units of $\mathrm{nm}^{-3}$
+in the bulk solvent far from the membrane interface. 
 \todo{SAMULI: Once we have to final results, we can probably say that
@@
 the relative surface excess between water and ions $\Gamma_w^i$ \cite{chattorajBOOK}
-\begin{equation}
-  \Gamma_w^i=\Gamma_w^x-\Gamma_i^x\frac{c_w-c_w'}{c_i-c_i'},
+\begin{eqnarray}
+  \Gamma_w^i   &=& \Gamma_w-\Gamma_i \frac{C_w-C_w'}{C_i-C_i'} \\
+  \Gamma_{a} &=& \frac{n_{a} - C_{a} V - C_{a}' V'}{A} \, \, ; \, a=i,w \, ,
+\end{eqnarray}
+where index $i$ and $w$ denote ions and water respectively,
+prime accent denotes the phase inside bilayer, 
+whereas non-prime symbols account for bulk water phase,
+$n$ is the total amount of water or ions,
+$C$ denotes bulk concentrations, 
+$V$ is the volume of the respective phase separated by the Gibbs divinding plane, 
+$A$ is the area of the interface, 
+and $\Gamma_a$ is the surface excess of water or ions. 
+As $\Gamma_w^i$ does not depend on the definition of the interfacial region
+and the position of the Gibbs dividing plane~\cite{chattorajBOOK},
+we consider the whole simulation box as an interface. 
+Note that phospholipid membranes are bilayers, 
+and hence the interface occurs in the simulation box twice
+providing a surface area $2A$.
+The concentrations of water and ions inside
+the bilayer can be expected to vanish, thus $c_{i,w}'=0$.
+This provides us a simplified relation for 
+the relative surface excess of lipid bilayers in simulations
+\begin{equation}\label{surfexcess}
+  \Gamma_w^i=\frac{1}{2A} \left ( n_w - n_i \frac{C_w}{C_i} \right ) ,
 \end{equation}
-where $c_w$ and $c_i$ are concentrations in bulk water, 
-$c_w'$ and $c_i'$ are concentrations inside bilayer, and
-$\Gamma_w^x$ and $\Gamma_i^x$ are surface excess of water and
-ions, respectively. The whole simulation box is considered here
-as interfacial region, thus $\Gamma_w^x=n_w/A$ and $\Gamma_i^x=n_i/A$,
-where $A$ is the area of the simulation box, and $n_w$ and
-$n_i$ are total number of water and ions in simulation box,
-respectively. The concentrations of water and ions inside
-bilayer can be expected to vanish, thus $c_w'=c_i'=0$.
-Then we get the relative surface excess for lipid bilayer
-simulations
-\begin{equation}\label{surfexcess}
-  \Gamma_w^i=\frac{n_w}{A}-\frac{n_i}{A}\frac{c_w}{c_i},
-\end{equation}
-where concentrations in bulk, $c_w$ and $c_i$, can be determined
-from the simulation box boundary.
+where the bulk concentrations, $C_w$ and $C_i$, can be determined
+from the simulation as $C_{bulk}$ in Equation~\ref{eq:conc}.
 

### 2017-08-16 Manuscript/manuscript.tex
MS update: improved reasoning around PN-vector and S_CD connection, explained simulation system composition, creation of ECC-surfactant, added many references to ion models, corrected some names, checked scattering form factor equiation sanity.

@@
 Lipid bilayer structure without ions was validated against NMR experiments
-by order parameters for C-H bonds and x-ray scattering experiments by using
+by order parameters for C-H bonds and \mbox{x-ray} scattering experiments by using
 form factors. The former validates the structures sampled by individual
@@
 where $f_\alpha(q_z)$ is the density of atomic scattering length, 
-$\rho_s$ is the density of solvent scattering lenght,
+$\rho_s$ is the density of solvent scattering length,
 $n_\alpha (z)$ is the number density of atom $\alpha$ and
@@
 (POPC) bilayer and an aqueous salt solutions of varying concentrations.
-Water molecules were described by OPC3 model \cite{Izadi16}
-\todo{Explain why were OPC3 chosen}.
+In particular, the periodic orthorhombic simulation box conatined 
+128 POPC molecules and approximately 50~water molecules per each lipid.
+Water molecules were described by OPC3 model~\cite{Izadi16} 
+as it is currently the most accurate three site rigid water model. 
 In order to test transferability of the newly developed ECC-lipids model, 
 we also performed several additional simulations with water models 
-OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}
+OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}
 presented in Supporting Information (SI). 
 We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Simulations with Lipid14 use ion model by Dang \cite{smith94,chang1999,dang2006}. 
+Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
 Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
@@
 and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}. 
-Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}. \\
-\todo{System sizes has to reported somewhere}
+Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
 
@@
 6, 14, 21, 35 or 50 surfactants per leaflet.
-These systems were first simulated 200~ns by using Lipid14 model.
+These systems were first simulated for 200~ns using Lipid14 model.
 First 20~ns was omitted from the analysis. 
-
-The same systems where also simulated with ECC-corrected lipi model...
-\todo{Describe the changes in surfactant model used with ECC-corrected lipids.}
+Simulations with ECC-lipids used the same simulation setup.
+We also applied electronic continuum correction to the cationic surfactant 
+by scaling all its charges with the same factor as for ECC-lipids, $f_q=0.8$, 
+and by using the atom types with reduced $\sigma$ parameters from ECC-lipids.
 
@@
 
-\subsection{Response of POPC headgroup on bound charge}
+\subsection{Response of POPC headgroup to bound charge}
 
@@
 
+REWRITE:
 The headgroup response to bound charge depends on both, the
@@
 mixtures of cationinc surfactants and PC lipids~\cite{scherer89}.
+
 The amount of the bound charge is known excatly in these systems,
-because essentially all surfactants with unit charge locate in lipid bilayers.
+because essentially all surfactants with a unit charge locate in the lipid bilayers.
 %Thus, the headgroup order parameter response to bound charge can be first
@@
 
-Changes of headgroup order parameter as a function of added cationic surfactant
+Changes of headgroup order parameters as a function of added cationic surfactant
 dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) from
 experiments~\cite{scherer89} and simulations are shown in
-Fig. \ref{OrderParameterCHANGESsurf}).
+Fig. \ref{OrderParameterCHANGESsurf}.
 As expected from Eq. \ref{OPchangeEQ}, the head group order parameter response to the bound cation concentration
-is approximately linear at least up to $\sim$0.3 mole fraction in experiments and
-in both simulation models, original Lipid14 and ECClipids \todo{Yet to be shown.}. 
+is approximately linear at least up to $30\%$ mole fraction in experiments and
+in both simulation models, original Lipid14 and ECC-lipids.  
 \todo{We might want to calculate the slopes from simulations.}
 
-The slope is, however, too steep in Lipid14 model indicating that the
-headgroup order parameters are too sensitive for bound charge in this model.
+The slope of the head group order parameters response is, however, 
+too steep in Lipid14 model indicating that 
+its head group response is too sensitive to a bound charge. 
 This has to be taken into account when interpreting the
-headgroup order parameter response to CaCl$_2$ concentration in Lipid14 model.
+head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
 %observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
@@
 is shown in Fig.~\ref{OrderParameterCHANGESsurf} as a function of bound cations.
-As expected~\cite{seelig87}, the headgroup orients more parallel with membrane
-normal with increasing amount of bound cations. The effect is stronger
-in Lipid14 parameters than in ECClipid parameters, which is in line
-with the reduced dipole moment and the lower sensitivity to bound charges
-in the ECClipid model.
+As expected~\cite{seelig87}, the headgroup orients more parallel with the membrane
+normal with an increasing amount of bound cations. The effect is stronger
+in Lipid14 model than in ECC-lipids model, which is in line
+with the reduced charge--dipole interactions in the ECC framework  
+leading to the lower head group sensitivity to bound charges in ECC-lipids model.
+
 \todo{It might be a good idea to establish a relation between order parameter change
-and P-N vector angle by using ECClipids model.}
+and P-N vector angle by using ECC-lipids model.}
 
@@
     from simulations with different force fields and experiments (DPPC \cite{akutsu81} and POPC \cite{altenbach84}).
-    Simulations with Lipid14 parameters and Aqvist ions are directly from Ref. \cite{catte16}, except that
+    Simulations with Lipid14 parameters and \AA{}qvist ions are directly from Ref. \cite{catte16}, except that
     the concentration are determined with Eq. \ref{eq:conc}.
@@
 models are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-The simulation results of Lipid14 with Aqvist ions \cite{??} (from Ref. \citenum{catte16})
+The simulation results of Lipid14 with \AA{}qvist ions~\cite{aqvist90} (from Ref. \citenum{catte16})
 are shown together with the results from Lipid14 simulation
-with Dang ions \cite{??} and from ECClipid simulations with ECC-ions.
+with ions by Dang~\cite{smith94,chang1999,dang2006} and from ECC-lipids simulations with ECC-ions.
 
-As already discussed in the recent work by NMRlipids project~\cite{catte16},
+As already discussed in a recent work by NMRlipids project~\cite{catte16},
 the headgroup order parameters are almost unaffected by addition of NaCl
-in Lipid14 simulations with Aqvist ion model due a very weak binding of Na$^+$
+in Lipid14 simulations with \AA{}qvist ion model due to a very weak binding of Na$^+$
 in PC bilayer. This is in agremeent with experimental data, in contrast
 to almost all the other lipid models, which significantly overestimate
-Na$^+$ binding in PC bilayers~\cite{catte16}. Despite of good performance
+Na$^+$ binding in PC bilayers~\cite{catte16}. Despite the good performance
 with NaCl, the response of headgroup order parameters to CaCl$_2$ concentration are
-significantly overestimated in Lipid14 simulations with Aqvist ion model,
-see Fig.~\ref{OrderParameterCHANGESnewMODELS} and \citenum{catte16}.
+significantly overestimated in Lipid14 simulations with \AA{}qvist ion model,
+see Fig.~\ref{OrderParameterCHANGESnewMODELS} and Ref.~\citenum{catte16}.
 
@@
 that CaCl$_2$ binding behaviour cannot be corrected by only improving
-ion models. This was not, however, tested for Lipid14 model, so we
-ran also Lipid14 simulations with Dang \cite{smith94,chang1999,dang2006} and ECC corrected \cite{??}
-ions having more realistic bulk behaviour as Aqvist model \cite{??}.
+ion models. This was not, however, tested specifically for Lipid14 model, so we
+ran also Lipid14 simulations with ions by Dang~\cite{smith94,chang1999,dang2006} 
+and ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014},
+which have more realistic bulk behaviour than \AA{}qvist model~\cite{aqvist90}.
 In line with previous work, the headgroup order parameters response
-to CaCl$_2$ concentartion was overestimated also with these ion models,
+to CaCl$_2$ concentration was overestimated also with these ion models,
 as seen in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively
@@
 
-Applying ECC correction also for lipids, introduced in section \ref{section:ecc},
+Applying ECC correction also to lipids, introduced in section \ref{section:ecc},
 brings the order parameter response to CaCl$_2$ concentration to a
@@
 that the ECC corrected lipid model has a sufficient accuracy for realistic
-studies of ion binding details on lipid bilayers.
+studies of details of ions binding to lipid bilayers.
 
-Ion density profiles along membrane normal coordinate
+Ion density profiles along membrane normal 
 from different simulations are shown in Fig.~\ref{fig:cacl-dens}.
@@
 show a pronounced peak in the position of the phosphate moieties of POPC. 
-The use of a ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
-or the standard Amber ion model by \AA{qvist}~\cite{aqvist90}
+The use of ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
+or the ion model by \AA{qvist}~\cite{aqvist90}
 along with Lipid14 does not significantly change it 
@@
     Number density of \ce{Ca^{2+}} and \ce{Cl^-} as a function of membrane normal axis
-    for different force fields. Data for Lipid14 with Aqvist ions are taken directly from Ref. \citenum{catte16}
+    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
@@
 
-This work will serve as a foundation stone of a 
-new open-collaboration project NMRlipids~VI in nmrlipids.blogspot.fi.
+This work can be reached as a repository containing all data at \url{zenodo.org:\dots\dots\dots} and as project \mbox{NMRlipids~VI} in \url{nmrlipids.blogspot.fi}.
 

### 2017-08-16 Manuscript/manuscript.tex
Improved figures: x-ray scattering form-f and NMR OPs show simulation at 300K, there's a better POPC_structure picture, updated other figures, Simulation with Dang ions+L14 is too small, still have to update it. added topologies of extra water models.

@@
   \centering
-  \includegraphics[height=8.5cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
   \includegraphics[width=7.0cm]{../Fig/POPCstructure.eps}
-  \includegraphics[width=9.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
+  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
   \caption{\label{simVSexpNOions}

### 2017-08-16 scripts/surface_excess_calculate.sh
Surface excess script -- work in progress, glueing python and BASH scripts that already work together

@@ -1,0 +1,73 @@
+#!/bin/bash
+
+# Gromacs 5.x version of ...
+# bash wrapper for calculating Order Parameters of lipid bilayer 
+# python script/library calcOrderParameters.py
+# meant for use with NMRlipids projects
+#------------------------------------------------------------
+# Made by J.Melcr,  Last edit 2017/03/21
+#------------------------------------------------------------
+
+scriptdir=`dirname $0`
+
+traj_file_name="traj_comp.xtc" #"../traj.trr" 
+traj_pbc_nonwat_file_name="traj_nonwat_pbc.xtc" #"../traj.trr" 
+top_file_name="last_frame_nonwat.gro"
+tpr_file_name="topol.tpr"
+dens_file_name="density_ca_cl_water.xvg"
+#op_def_file="../../Headgroup_Glycerol_OPs.def"
+op_def_file=${scriptdir}"/order_parameter_definitions_Lipid14_POPC_all.def"
+op_out_file="OrdPars.dat"
+top="topol.top"
+f_conc=55430  # in mM/L
+
+if ! [ -s $tpr_file_name ] 
+then
+    echo "We really need " $tpr_file_name " , but we can't find it!"
+    exit 1
+fi
+
+# get densities of cation - anion - water centered around POPC
+# and start at 100ns
+if ! [ -s $dens_file_name ] 
+then
+    seq 2 5 | gmx density -sl 900 -dens number -ng 3 -f $traj_file_name -center -symm -relative -o $dens_file_name -xvg none -b 100000 
+fi
+
+
+# getting actual bulk concentration from number density profile
+python $scriptdir/get_conc_ion_bulk.py 
+conc=`cut -d " " -f1 conc_ion_bulk_mmolL.dat`
+conc_wat=`cut -d " " -f1 conc_wat_bulk_mmolL.dat`
+
+#getting nominal concentration from topol.top file (if exists)
+if [ -f $top ]
+then
+    nwat=`grep -e "molecules" -A 10 $top | grep -e "^SOL" -e "^TIP" -e "^SPCE" -e "^OPC3" | cut -d " " -f1 --complement `
+    nion=`grep -e "molecules" -A 10 $top | grep -e "^NA"  -e "^CA"  | cut -d " " -f1 --complement `
+    [ -z $nion ] && nion=0
+
+    concnom=`echo $f_conc "*" $nion / $nwat  | bc`
+    echo nominal conc: $concnom
[…]

### 2017-08-17 scripts/calc_apl.py
Update of the surface-excess script

@@ -32,6 +32,7 @@ def main():
        if line.startswith("Box-Y"):
            ydim=float(line.split()[1])
 
+   print "area= %6.3f nm2" % (xdim*ydim)
    print "apl= (%6.3f*%6.3f)/%3d = %6.3f nm2" % (xdim,ydim,nLipids, xdim*ydim/nLipids)
 
    sys.exit(0)

### 2017-08-17 scripts/surface_excess_calculate.sh

@@ -34,10 +34,13 @@ then
     seq 2 5 | gmx density -sl 900 -dens number -ng 3 -f $traj_file_name -center -symm -relative -o $dens_file_name -xvg none -b 100000 
 fi
 
+# calculate area and area-per-lipid
+area=`bash $scriptdir/area-per-lipid_calculate.sh | grep -e "area" | cut -d " " -f3`
 
 # getting actual bulk concentration from number density profile
 python $scriptdir/get_conc_ion_bulk.py 
 conc=`cut -d " " -f1 conc_ion_bulk_mmolL.dat`
+conc_anion=`cut -d " " -f1 conc_anion_bulk_mmolL.dat`
 conc_wat=`cut -d " " -f1 conc_wat_bulk_mmolL.dat`
 
 #getting nominal concentration from topol.top file (if exists)
@@ -47,6 +50,9 @@ then
     nion=`grep -e "molecules" -A 10 $top | grep -e "^NA"  -e "^CA"  | cut -d " " -f1 --complement `
     [ -z $nion ] && nion=0
 
+    nanion=`grep -e "molecules" -A 10 $top | grep -e "^CL"  | cut -d " " -f1 --complement `
+    [ -z $nanion ] && nanion=0
+
     concnom=`echo $f_conc "*" $nion / $nwat  | bc`
     echo nominal conc: $concnom
     #sed "s/conc/$conc/" -i ${op_out_file}.line
@@ -55,19 +61,10 @@ else
     echo "Topology probably not present, can't calculate concentration."
 fi
 
-surfexc=`echo $nwat "-" $nion '*' $conc_wat"/"$conc  | bc`
+surfexc=`echo scale=4 ; ($nwat "-" $nion '*' $conc_wat"/"$conc)/("2*"$area)  | bc`
+surfexcani=`echo scale=4 ; ($nwat "-" $nanion '*' $conc_wat"/"$conc_anion)/("2*"$area)  | bc`
 
-echo Relative surface excess of ions-water = $surfexc
+echo Relative surface excess of water--ions = $surfexc > surf_excess_wat_ions.dat
+echo Relative surface excess of water--anions = $surfexcani > surf_excess_wat_anions.dat
+echo Relative surface excess of water--ions = $surfexc
 
-exit 0
-
-# remove PBC:
-# center the trajectory around POPC (should be no. 2)
-# and start at 100ns
[…]

### 2017-08-17 scripts/get_conc_ion_bulk.py
Small update to python script for getting bulk concentrations - now provides values for cation, anion, water.

@@ -21,3 +21,16 @@ print "bulk conc. = ({: 4.0f} +/- {: 3.0f}) mmol/L".format(mean, stdev)
 with open("conc_ion_bulk_mmolL.dat", "w") as f:
     f.write(" ".join([str(mean), str(stdev)]))
 
+wat_dens = densities[:,3]
+mean_wat  = wat_dens[:10*i].mean() /denom_conc *1000.0
+stdev_wat = wat_dens[:10*i].std()  /denom_conc *1000.0
+
+with open("conc_wat_bulk_mmolL.dat", "w") as f:
+    f.write(" ".join([str(mean_wat), str(stdev_wat)]))
+
+anion_dens = densities[:,2]
+mean_anion  = anion_dens[:10*i].mean() /denom_conc *1000.0
+stdev_anion = anion_dens[:10*i].std()  /denom_conc *1000.0
+
+with open("conc_anion_bulk_mmolL.dat", "w") as f:
+    f.write(" ".join([str(mean_anion), str(stdev_anion)]))

### 2017-08-17 scripts/surface_excess_calculate.sh
Surface excess script ready -- works with 3-site water models, for 4-site models, it has to be hand-tuned.

@@ -35,7 +35,8 @@ then
 fi
 
 # calculate area and area-per-lipid
-area=`bash $scriptdir/area-per-lipid_calculate.sh | grep -e "area" | cut -d " " -f3`
+area=`bash $scriptdir/area-per-lipid_calculate.sh | grep -e "area" | cut -d " " -f2`
+echo area = $area nm2
 
 # getting actual bulk concentration from number density profile
 python $scriptdir/get_conc_ion_bulk.py 
@@ -43,10 +44,21 @@ conc=`cut -d " " -f1 conc_ion_bulk_mmolL.dat`
 conc_anion=`cut -d " " -f1 conc_anion_bulk_mmolL.dat`
 conc_wat=`cut -d " " -f1 conc_wat_bulk_mmolL.dat`
 
+
 #getting nominal concentration from topol.top file (if exists)
 if [ -f $top ]
 then
     nwat=`grep -e "molecules" -A 10 $top | grep -e "^SOL" -e "^TIP" -e "^SPCE" -e "^OPC3" | cut -d " " -f1 --complement `
+    # correct the number density of water with the number of its atoms
+    # unfortunately, this does not really work, all 4-site models I employ use name SOL, however, some 3-site use it too
+    nwat4=`grep -e "molecules" -A 10 $top | grep -e "^TIP4" -e "^OPC4" | cut -d " " -f1 --complement `
+    if ! [ -z $nwat4 ]
+    then
+       watdenom=4.0
+    else
+       watdenom=3.0
+    fi
+
     nion=`grep -e "molecules" -A 10 $top | grep -e "^NA"  -e "^CA"  | cut -d " " -f1 --complement `
     [ -z $nion ] && nion=0
 
@@ -61,10 +73,15 @@ else
     echo "Topology probably not present, can't calculate concentration."
 fi
 
-surfexc=`echo scale=4 ; ($nwat "-" $nion '*' $conc_wat"/"$conc)/("2*"$area)  | bc`
-surfexcani=`echo scale=4 ; ($nwat "-" $nanion '*' $conc_wat"/"$conc_anion)/("2*"$area)  | bc`
+# correct the number density of water with the number of its atoms
+conc_wat=`echo "scale=4 ; "$conc_wat "/" $watdenom | bc `
+echo water concentration = $conc_wat mmol/L
 
-echo Relative surface excess of water--ions = $surfexc > surf_excess_wat_ions.dat
[…]

### 2017-08-17 Manuscript/manuscript.tex
Changed the definition of Relative surf. excess from water-w.r.t.-ions to excess of ions-w.r.t.-water, which is more meaningful for us. Done this both in TeX and script.

@@
 The ion binding affinities between different models were compared by calculating
-the relative surface excess between water and ions $\Gamma_w^i$ \cite{chattorajBOOK}
+the relative surface excess of ions with respect to water, $\Gamma_i^w$, as \cite{chattorajBOOK}
 \begin{eqnarray}
-  \Gamma_w^i   &=& \Gamma_w-\Gamma_i \frac{C_w-C_w'}{C_i-C_i'} \\
+  \Gamma_i^w   &=& \Gamma_i-\Gamma_w \frac{C_i-C_i'}{C_w-C_w'} \\
   \Gamma_{a} &=& \frac{n_{a} - C_{a} V - C_{a}' V'}{A} \, \, ; \, a=i,w \, ,
@@
 and $\Gamma_a$ is the surface excess of water or ions. 
-As $\Gamma_w^i$ does not depend on the definition of the interfacial region
+As $\Gamma_i^w$ does not depend on the definition of the interfacial region
 and the position of the Gibbs dividing plane~\cite{chattorajBOOK},
@@
 \begin{equation}\label{surfexcess}
-  \Gamma_w^i=\frac{1}{2A} \left ( n_w - n_i \frac{C_w}{C_i} \right ) ,
+  \Gamma_i^w=\frac{1}{2A} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
 \end{equation}

### 2017-08-17 scripts/surface_excess_calculate.sh

@@ -77,11 +77,11 @@ fi
 conc_wat=`echo "scale=4 ; "$conc_wat "/" $watdenom | bc `
 echo water concentration = $conc_wat mmol/L
 
-surfexc=`echo "scale=4 ; ("$nwat "-" $nion '*' $conc_wat"/"$conc")/(2*"$area")"  | bc`
-surfexcani=`echo "scale=4 ; ("$nwat "-" $nanion '*' $conc_wat"/"$conc_anion")/(2*"$area")"  | bc`
+surfexc=`echo "scale=4 ; ("$nion "-" $nwat '*' $conc"/"$conc_wat")/(2*"$area")"  | bc`
+surfexcani=`echo "scale=4 ; ("$nanion "-" $nwat '*' $conc_anion"/"$conc_wat")/(2*"$area")"  | bc`
 
-echo Relative surface excess of water--ions = $surfexc nm-2 > surf_excess_wat_ions.dat
-echo Relative surface excess of water--anions = $surfexcani nm-2 > surf_excess_wat_anions.dat
-echo Relative surface excess of water--ions = $surfexc nm-2 
-echo Relative surface excess of water--anions = $surfexcani nm-2 
+echo Relative surface excess of ions  --water = $surfexc nm-2 > surf_excess_wat_ions.dat
+echo Relative surface excess of ions  --water = $surfexc nm-2 
+echo Relative surface excess of anions--water = $surfexcani nm-2 > surf_excess_wat_anions.dat
+echo Relative surface excess of anions--water = $surfexcani nm-2 
 

### 2017-08-18 Manuscript/manuscript.tex
MS update: added volmap figures with a small commentary paragraph.

@@
     Number density of \ce{Ca^{2+}} and \ce{Cl^-} as a function of membrane normal axis
-    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}
+    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
+    Relative surface excess for the respective simulations are: 
+    Lipid14 with \AA{}qvist ions $\Gamma_i^w = XX nm^{-3}$; 
+    Lipid14 with Dang ions $\Gamma_i^w = 0.3 nm^{-3}$; 
+    ECC-lipids with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$. 
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
-    Densities of~\ce{Cl^-}) and water are divided with 2 and 200, respectively, to visalize them
-    with the same scale as \ce{Ca^{2+}}. The molar concentration in water is 350~mM in all systems
+    Densities of~\ce{Cl^-} and water are divided with 2 and 200, respectively, to visalize them
+    with the same scale as \ce{Ca^{2+}}. The molar concentration of the ions in water is 350~mM in all systems
     presented here. 
     }
+  \todo{Calculate the surface excess for L14+Aqvist sim.}
   \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
   \todo{JOE: Change the figure so that it contains a membrane background} \\
-  \todo{Resuls with Lipid14+Dang to be added once done.}
+  \todo{Resuls with Lipid14+Dang to be added once done -- running with larger box.}
 \end{figure}
@@
 Technical details of the analysis are in the SI.
+
+In addition, we estimated the relative binding affinities of several moieties in POPC:
+phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
+Using the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
+we estimate that the largest contribution
+to the binding of \ce{Ca2+} to POPC membranes likely comes from the phosphate group. 
+Although the isodensity plots are relatively easy to interpret,
+the contours shown in Fig.~\ref{fig:volmaps} cannot conclusively tell on the details
+of \ce{Ca2+} binding to any of the two carbonyl moieties.
 
@@
 %\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} \\
+  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
+  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
+  \caption{\label{fig:volmaps}
+      Contours of probability isodensities of \ce{Ca2+} with respect to 
+      various moieties fixed in space (highlighted with transparent spheres): 
+      phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
+      Shown contours suggest that the dominant contribution 
+      to \ce{Ca2+} binding likely comes from the phosphate group, 
+      but remain inconclusive on the details of the interactions with any of the two carbonyl groups. 
+  }
+\end{figure}
 

### 2017-08-25 Manuscript/manuscript.tex

@@
 order parameters of $\alpha$ and $\beta$ carbons in lipid head group
-(see Fig. \ref{??} 
-\todo{Figure with chemical structure and labeling to be added}) 
+(see Fig.~\ref{simVSexpNOions}) 
 are proportional to the amount of bound charge
-in lipid bilayer \cite{seelig87}.
+in a lipid bilayer \cite{seelig87}.
 More recent analysis included also
@@
     compared with experimental order parameters from \cite{ferreira13}.}
-  \todo{We should run a simulation for the system withtout ions with the
-    same (or close) temperature as the experimental data. The NMR data is at 300K
-    so that would be good. I think that this would be also close enough for the scattering data at 303K.
-  }
 \end{figure}

### 2017-08-25 Manuscript/manuscript.tex
Updates in methods section and comments added.

@@
 Ion binding was compared between experiments and simulations by 
-using lipid head group order parameters and ''electrometer concept'' 
-introduced by Seelig et al. \cite{seelig87,catte16}.
-The concept is based on the experimental observation that the 
+using lipid head group order parameters and 
+''electrometer concept'' \cite{seelig87,catte16}.
+The concept is based on the experimental observation that the C-H bond
 order parameters of $\alpha$ and $\beta$ carbons in lipid head group
-(see Fig.~\ref{simVSexpNOions}) 
-are proportional to the amount of bound charge
-in a lipid bilayer \cite{seelig87}.
-More recent analysis included also
-the order parameter signs and concluded that the order parameters  
-decrease with bound positive charge and increase with bound negative 
-charge \cite{ollila16,catte16}. The observations are rationalized 
-as a change of lipid head group dipole tilt to more vertical orientation
+(see Fig.~\ref{simVSexpNOions}) are proportional to the amount 
+of bound charge in a lipid bilayer \cite{seelig87}.
+This is rationalized as a change of lipid head group dipole tilt to more vertical orientation
 with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
-Order parameters for $\alpha$ and $\beta$ segments in headgroup, as well as for
+More recent analysis in the NMRlipids project concluded 
+that the order parameters decrease with bound positive charge 
+and increase with bound negative charge when the signs are taken 
+in account \cite{ollila16,catte16}. The concept was also shown
+to be suitable for the comparison of ion binding affinity in lipid 
+bilayers between simulations and experiments~\cite{catte16}.
+
+The order parameters for $\alpha$ and $\beta$ segments in headgroup, as well as for
 all C-H bonds in lipid molecules,
-can be accurately measured by using H$^2$ NMR or $^{13}$C NMR techniques.
-The experimental order parameters can be compared with the the ones
-calculated from MD simulations by using the definition
+can be accurately measured by using $^2$H NMR or $^{13}$C NMR techniques.
+The order parameters can be calculated from MD simulations 
+by using the definition
 \begin{equation}\label{OP}
@@
 and the value of the quadrupole coupling constant is $\chi \approx$\,167\,kHz.
-Atomic absorption spectra and $^2$H NMR data, combined with the
-information about order parameter signs gave
+Atomic absorption spectra and $^2$H NMR data gave
 $m_\alpha=-20.5$  and $m_\beta=-10.0$ for Ca$^{2+}$ binding to POPC
 bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-Recent work published by the NMRlipids project showed that the concept works
-qualitatively also in simulations and can be used to compare ion binding
-affinitity between simulations and experiments~\cite{catte16}.
+Before using electrometer concept to analyze cation binding, 
+we calibrate the response of headgroup order parameters to the bound
+charge in the simulation models. This is done by comparing 
+the slope $m_i$ in Eq. \ref{OPchangeEQ} between simulations and
+experiments of cationic surfactants in POPC bilayer \cite{scherer89}. 
+Since essentially all surfactants locate in bilayer, the 
+molar ratio directly gives the amount of bound charge per 
+lipid $X^\pm$ is these systems.
 
-The experimental data used to quantify cation binding reports
-order parameters as a function of equilibrium cation concentration
-in the bulk solvent \cite{akutsu81,altenbach84}. Recent work
-by NMRlipids project estimated this concentration by using
-the relative molar concentration in water
-[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M,
-which was accurate enough for the conclusions in that study \cite{catte16}.
-Since we are now targeting a model with quantitatively accurate binding
-affinity, we rather use a definition that corresponds more closely to the 
-concentrations reported in experiments. 
-By using larger simulation box dimensions that allow the
-determination of bulk ion concentration, the equilibrium concentration
-corresponding to the experimental data in the units of $\mathrm{mol/l}$
-can be then determined from the simulations as
-\begin{equation}
-\label{eq:conc}
-  C_{eq}=\frac{C_{bulk}}{0.602},
-\end{equation}
-where the bulk concentration, $C_{bulk}$, is the number density in the units of $\mathrm{nm}^{-3}$
-in the bulk solvent far from the membrane interface. 
-\todo{SAMULI: Once we have to final results, we can probably say that
-  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
-  }
+\subsection{Ion concentrations and binding affinity}
+The experimental headgroup order parameter data used
+in this work is reported as a function of 
+cation concentration remaining in the bulk 
+solvent \cite{akutsu81,altenbach84}. 
+To determine the same concentration from simulations, we
+use large enough box to ensure that the ion concetration
+reaches a clear plateau and the electroneutrality
+is satisfied at the edge of simulation box. It should be noted that the recent work
+by NMRlipids project compared simulations to the same 
+experimental data by using the relative molar concentration in water
+[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M \cite{catte16}.
+This is not exactly the same as the concentration reported in the used experimental
+data, because the binding of ions in bilayer reduces the bulk concentration.
+The difference slightly moves points towards lower concentrations but do not 
+affect the conclusions in the previous study \cite{catte16}.
 
@@
 where the bulk concentrations, $C_w$ and $C_i$, can be determined
-from the simulation as $C_{bulk}$ in Equation~\ref{eq:conc}.
+from the concentration plateau at the edge of the simulation box.
 
@@
 
-Order parameters were calculated from simulations for all C-H bonds
-by using Eq. \ref{OP} and form factors as
+The order parameters were calculated from simulations for all C-H bonds
+in lipid molecule by using Eq. \ref{OP}. Form factors were calculated 
+from equation
 \begin{equation}
@@
 6, 14, 21, 35 or 50 surfactants per leaflet.
-These systems were first simulated for 200~ns using Lipid14 model.
+The systems were simulated for 200~ns using Lipid14 model for POPC.
 First 20~ns was omitted from the analysis.
-Simulations with ECC-lipids used the same simulation setup.
-We also applied electronic continuum correction to the cationic surfactant 
-by scaling all its charges with the same factor as for ECC-lipids, $f_q=0.8$, 
+
+The same systems were also simulated with ECC-lipid model for POPC \todo{How long?}.
+In these simulations the ECC correction was also applied to the cationic surfactant 
+by scaling all \todo{Were all scaled? Are there charges is tails? Were those scaled as well in this case?} 
+charges with the same factor as for ECC-lipids, $f_q=0.8$, 
 and by using the atom types with reduced $\sigma$ parameters from ECC-lipids.
@@
     compared with experimental order parameters from \cite{ferreira13}.}
+  \todo{There seems to a typo in experimental value for other order parameter for C2 carbon in sn-2 chain; it should be negative.
+  The scale of the y-axis should be then zoomed when this is fixed.} \\
+  \todo{x-axis scale in form factor figure from 0 to 0.6, until where experimental data is available.}
 \end{figure}
@@
   \end{tabular}
+  \todo{SAMULI: I would put here Lipid14 in 303K, ECC-lipid in 303K and experiment in 303K. Rest in SI.
+  The best experimental value would be the one analyzed from the form factor shown in previous figure, 
+  if available.}
 \end{table}
@@
   \caption{\label{OrderParameterCHANGESnewMODELS}
-    Headgroup order parameter changes as a function of NaCl, CaCl$_2$ concentration
-    from simulations with different force fields and experiments (DPPC \cite{akutsu81} and POPC \cite{altenbach84}).
-    Simulations with Lipid14 parameters and \AA{}qvist ions are directly from Ref. \cite{catte16}, except that
-    the concentration are determined with Eq. \ref{eq:conc}.
+    Changes of headgroup order parameters of POPC bilayer as a function of NaCl and CaCl$_2$ concentrations
+    are shown from simulations with different force fields together with experimental data 
+    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
+    Ion concentrations in bulk water are shown in x-axis. 
+    Values from simulations are calculated from the plateau of cation number density $C_{np}$
+    at the box edge as [ion]=$C_{np}/0.602$.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
   }
-  \todo{Simulation results with original Slipids and Dang ions to be added when we have the results.}
 \end{figure*}
@@
 as seen in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively
-\todo{Add OP-response of Lipid14+Dang in Fig. \ref{OrderParameterCHANGESnewMODELS}
-  and OP-response of Lipid14+ECC-ions plot in SI}.
+\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
 
@@
 studies of details of ions binding to lipid bilayers.
+
+\todo{SAMULI: Maybe we should discuss the repeat distances and
+  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
+  }
 

### 2017-08-28 Manuscript/manuscript.tex
MS: tiny change

--- 2017-08-18 Manuscript/manuscript.tex
@@
 to find an atomistic MD simulation force field parameters for lipids that
-accurately describe the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
-when compared with NMR data \cite{akutsu81,altenbach84}. These data can be used to accurately asses
+accurately describe the lipid head group response to varying concentrations of cations
+when compared to NMR data \cite{akutsu81,altenbach84,scherer89}. Such data can be used to accurately asses
 ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
@@
     Number density of \ce{Ca^{2+}} and \ce{Cl^-} as a function of membrane normal axis
-    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
-    Relative surface excess for the respective simulations are: 
-    Lipid14 with \AA{}qvist ions $\Gamma_i^w = XX nm^{-3}$; 
-    Lipid14 with Dang ions $\Gamma_i^w = 0.3 nm^{-3}$; 
-    ECC-lipids with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$. 
+    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
-    Densities of~\ce{Cl^-} and water are divided with 2 and 200, respectively, to visalize them
-    with the same scale as \ce{Ca^{2+}}. The molar concentration of the ions in water is 350~mM in all systems
+    Densities of~\ce{Cl^-}) and water are divided with 2 and 200, respectively, to visalize them
+    with the same scale as \ce{Ca^{2+}}. The molar concentration in water is 350~mM in all systems
     presented here. 
     }
-  \todo{Calculate the surface excess for L14+Aqvist sim.}
   \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
   \todo{JOE: Change the figure so that it contains a membrane background} \\
-  \todo{Resuls with Lipid14+Dang to be added once done -- running with larger box.}
+  \todo{Resuls with Lipid14+Dang to be added once done.}
 \end{figure}
@@
 Technical details of the analysis are in the SI.
-
-In addition, we estimated the relative binding affinities of several moieties in POPC:
-phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
-Using the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
-we estimate that the largest contribution
-to the binding of \ce{Ca2+} to POPC membranes likely comes from the phosphate group. 
-Although the isodensity plots are relatively easy to interpret,
-the contours shown in Fig.~\ref{fig:volmaps} cannot conclusively tell on the details
-of \ce{Ca2+} binding to any of the two carbonyl moieties.
 
@@
 %\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} \\
-  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
-  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
-  \caption{\label{fig:volmaps}
-      Contours of probability isodensities of \ce{Ca2+} with respect to 
-      various moieties fixed in space (highlighted with transparent spheres): 
-      phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
-      Shown contours suggest that the dominant contribution 
-      to \ce{Ca2+} binding likely comes from the phosphate group, 
-      but remain inconclusive on the details of the interactions with any of the two carbonyl groups. 
-  }
-\end{figure}
 

### 2017-08-30 Manuscript/manuscript.tex
Update 'Response of POPC headgroup to bound charge' section

@@
 
-REWRITE:
-The headgroup response to bound charge depends on both, the
-binding affinity and the response of the headgroup on bound charge.
-Thus, the headgroup response to bound charge has to be correctly
-described in the model if the concept is used for quantitative studies of
-binding affinity. Hence, we first quantify the response of head group order
-parameters to the bound charge by using experimental data measured from
-mixtures of cationinc surfactants and PC lipids~\cite{scherer89}.
+Before proceeding to ion binding affinity studies, we quantify
+the response of headgroup order parameters to the amount of 
+bound charge by using mixtures of monovalent cationic surfactants
+and POPC~\cite{scherer89}. The amount of bound charge per PC 
+in these systems is 
+given by the molar fraction of cationic surfactants, because essentially all surfactants 
+locate in the lipid bilayers. Comparison of headgroup order parameter
+changes between simulations and experiments 
+as a function of mole fraction of cationic surfactants 
+can be then used to validate the sensitivity of
+lipid headgroup to bound charge in simulation models.
 
-The amount of the bound charge is known excatly in these systems,
-because essentially all surfactants with a unit charge locate in the lipid bilayers.
+%depends on both, the
+%binding affinity and the response of the headgroup on bound charge.
+%Thus, the headgroup response to bound charge has to be correctly
+%described in the model if the concept is used for quantitative studies of
+%binding affinity. Hence, we first quantify the response of head group order
+%parameters to the bound charge by using experimental data measured from
+%mixtures of cationinc surfactants and PC lipids~\cite{scherer89}.
+%
 %Thus, the headgroup order parameter response to bound charge can be first
@@
 
-Changes of headgroup order parameters as a function of added cationic surfactant
-dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) from
-experiments~\cite{scherer89} and simulations are shown in
+The headgroup order parameter changes as a function of mole fraction
+of cationic surfactant dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) 
+is compared between experiments~\cite{scherer89} and simulations in
 Fig. \ref{OrderParameterCHANGESsurf}. 
-As expected from Eq. \ref{OPchangeEQ}, the head group order parameter response to the bound cation concentration
-is approximately linear at least up to $30\%$ mole fraction in experiments and
-in both simulation models, original Lipid14 and ECC-lipids.  
-\todo{We might want to calculate the slopes from simulations.}
-
-The slope of the head group order parameters response is, however, 
-too steep in Lipid14 model indicating that 
-its head group response is too sensitive to a bound charge. 
-This has to be taken into account when interpreting the
-head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
-%observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
+Approximately linear decrease of the order parameters is observed
+in simulations and experiments with increasing amount of cationic
+surfactants up to mole fractions of $\sim$50\%,
+as expected from Eq. \ref{OPchangeEQ}.
+The slope is, however, too steep in Lipid14 model indicating that 
+the headgroup order parameters are too sensitive to a bound charge.
 The ECC-lipids model gives a slope in very good agreement with experiments
@@
 underestimated for the $\beta$ segment.
+%This has to be taken into account when interpreting the
+%head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
+%observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
 %The present comparison of head group order parameter changes 
@@
 %to the bound charge.
+\todo{SAMULI: We could calculate the slopes from simulations, but I am not sure if
+we would actually learn anything useful from this.}
 
-Also the angle of the headgroup P-N vector with respect to membrane normal
-is shown in Fig.~\ref{OrderParameterCHANGESsurf} as a function of bound cations.
-As expected~\cite{seelig87}, the headgroup orients more parallel with the membrane
-normal with an increasing amount of bound cations. The effect is stronger
-in Lipid14 model than in ECC-lipids model, which is in line
-with the reduced charge--dipole interactions in the ECC framework  
-leading to the lower head group sensitivity to bound charges in ECC-lipids model.
-
-\todo{It might be a good idea to establish a relation between order parameter change
-and P-N vector angle by using ECC-lipids model.}
+Also the effect of bound cations on the
+headgroup P-N vector angle with respect to the membrane normal
+is shown in Fig.~\ref{OrderParameterCHANGESsurf}.
+The headgroup orients more parallel to the membrane
+normal with an increasing amount of bound cations, as expected~\cite{seelig87}. 
+The orientation effect is more pronounced in Lipid14 model than 
+in ECC-lipids model, which is in line with the order parameter 
+results and reduced charge--dipole interactions in the ECC-lipid model.
+Since the response of $\alpha$-order parameter is in very good agreement
+with experiments in ECC-lipid model, we can suggest a relation between
+order parameter and P-N vector angle changes. The data gives
+$\Delta \Theta_{\rm P-N}=?? \Delta S_{\rm CH}$, which can be used
+to estimate P-N vector angle changes from experimental results
+for $\alpha$ order parameters. 
+\todo{SAMULI: I think that we should establish the relation between 
+order parameter and P-N vector angle changes by using ECC-lipids model.
+This would be useful for people measuring aplha order parameter
+changes from experiments.}
 
@@
 
-%After achieving accurate head group order parameter response with a cationic surfactant, 
-%we advanced to quantifying the response to solutions with varying concentrations of salts, namely \ce{NaCl} and \ce{CaCl_2}.
-The headgroup order parameter response to aqueous ionic concentration
-can be used to measure ion binding affinity in experiments and simulations,
-as discussed in section \ref{section:electrometer} and in previous literature~\cite{seelig87,catte16}.
-The headgroup order parameter response of PC lipids to NaCl and CaCl$_2$ concentrations
-from experiments (DPPC \cite{akutsu81} and POPC \cite{altenbach84}) and different simulation
-models are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+The linear decrease of the headgroup order parameters, quantified in previous 
+section, can be used to measure the amount of bound aqueous cations in lipid 
+bilayers~\cite{seelig87,catte16}. The headgroup order parameter response 
+to NaCl and CaCl$_2$ concentrations from experiments (DPPC \cite{akutsu81} 
+and POPC \cite{altenbach84}) and different simulation models for POPC
+are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
 The simulation results of Lipid14 with \AA{}qvist ions~\cite{aqvist90} (from Ref. \citenum{catte16})
@@
     }
-  \todo{Calculate the surface excess for L14+Aqvist sim.}
+  \todo{Calculate the surface excess for L14+Aqvist sim.} \\
   \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
@@
 \end{figure}
+
+\subsection{Binding stoichiometry}
 

### 2017-09-03 Manuscript/manuscript.tex

@@
 are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-The simulation results of Lipid14 with \AA{}qvist ions~\cite{aqvist90} (from Ref. \citenum{catte16})
-are shown together with the results from Lipid14 simulation
-with ions by Dang~\cite{smith94,chang1999,dang2006} and from ECC-lipids simulations with ECC-ions.
+%The simulation results of Lipid14 with \AA{}qvist ions~\cite{aqvist90} (from Ref. \citenum{catte16})
+%are shown together with the results from Lipid14 simulation
+%with ions by Dang~\cite{smith94,chang1999,dang2006} and from ECC-lipids simulations with ECC-ions.
 
-As already discussed in a recent work by NMRlipids project~\cite{catte16},
-the headgroup order parameters are almost unaffected by addition of NaCl
-in Lipid14 simulations with \AA{}qvist ion model due to a very weak binding of Na$^+$
-in PC bilayer. This is in agremeent with experimental data, in contrast
-to almost all the other lipid models, which significantly overestimate
-Na$^+$ binding in PC bilayers~\cite{catte16}. Despite the good performance
-with NaCl, the response of headgroup order parameters to CaCl$_2$ concentration are
-significantly overestimated in Lipid14 simulations with \AA{}qvist ion model,
-see Fig.~\ref{OrderParameterCHANGESnewMODELS} and Ref.~\citenum{catte16}.
+As already discussed previously~\cite{catte16}, NaCl has a negligible effect 
+on the headgroup order parameters with submolar 
+concentrations due to the very weak binding in PC bilayers. 
+In contrast to almost all the other simulation models, this behaviour 
+is reproduced by Lipid14 model with \AA{}qvist ions \cite{catte16}. 
+The headgroup order parameter response to CaCl$_2$ concentration was,
+however, overestimated with the same combination of force field
+parameters~\cite{catte16}. The results were not improved when
+ion models by Dang et al.~\cite{smith94,chang1999,dang2006} or
+ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
+with more realistic bulk behaviour were used, as seen in 
+in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively.
+\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
+Together with extensive data in previous work \cite{catte16}, 
+the results suggest that CaCl$_2$ binding behaviour in PC bilayers
+cannot be corrected by only improving the ion models.
 
-Wide range of data presented in previous work \cite{catte16} suggested
-that CaCl$_2$ binding behaviour cannot be corrected by only improving
-ion models. This was not, however, tested specifically for Lipid14 model, so we
-ran also Lipid14 simulations with ions by Dang~\cite{smith94,chang1999,dang2006} 
-and ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014},
-which have more realistic bulk behaviour than \AA{}qvist model~\cite{aqvist90}.
-In line with previous work, the headgroup order parameters response
-to CaCl$_2$ concentration was overestimated also with these ion models,
-as seen in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively
-\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
-
-Applying ECC correction also to lipids, introduced in section \ref{section:ecc},
-brings the order parameter response to CaCl$_2$ concentration to a
-good agreement with experiments, as shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-This is a significant improvement over previously introduced models,
-which all overestimate the head group order parameter response to CaCl$_2$
-concentration \cite{catte16}. The good agreement with experiments indicates
-that the ECC corrected lipid model has a sufficient accuracy for realistic
-studies of details of ions binding to lipid bilayers.
-
-\todo{SAMULI: Maybe we should discuss the repeat distances and
-  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
-  }
-
-Ion density profiles along membrane normal 
-from different simulations are shown in Fig.~\ref{fig:cacl-dens}.
+Significant improvement is achieved when the developed ECC-lipid model 
+is used. The headgroup order parameter changes as a function of CaCl$_2$
+concentration model show a good agreement between ECC-lipid model and 
+experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+Ion binding affinity is quantified by using the relative surface
+excess between ions and water from Eq.~\ref{surfexcess} and  
+density profiles along membrane normal shown in Fig.~\ref{fig:cacl-dens}.
+%This is directly related to the amount of surface excess charge 
+%as shown along with ion density profiles in  Fig.~\ref{fig:cacl-dens}.
+The relative surface excess for ECC-lipid model with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$
+is significantly smaller than for Lipid14 with \AA{}qvist $\Gamma_i^w = XX nm^{-3}$
+and Dang ions $\Gamma_i^w = 0.3 nm^{-3}$.
+%The density profiles from simulations with original Lipid14 \cite{dickson14}
+%and Dang ions \cite{smith94,chang1999,dang2006} 
+%show a pronounced peak in the position of the phosphate moieties of POPC. 
+%The use of ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
+%or the ion model by \AA{qvist}~\cite{aqvist90}
+%along with Lipid14 does not significantly change it 
+%(Headgroup order parameter responses for these models in SI).
+%The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
+%exhibits on the other hand smaller density in this region 
+%suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
+%It is then likely that the models that overestimate head group order parameter response~\cite{catte16}
+%also overestimate the amount of bound cations.
 \todo{This paragraph is to be finalized when we have the surface excess analysis
   done by using Eq.~\ref{surfexcess}.}
-This is directly related to the amount of surface excess charge 
-as shown along with ion density profiles in  Fig.~\ref{fig:cacl-dens}.
-The density profiles from simulations with original Lipid14 \cite{dickson14}
-and Dang ions \cite{smith94,chang1999,dang2006} 
-show a pronounced peak in the position of the phosphate moieties of POPC. 
-The use of ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
-or the ion model by \AA{qvist}~\cite{aqvist90}
-along with Lipid14 does not significantly change it 
-(Headgroup order parameter responses for these models in SI).
-The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
-exhibits on the other hand smaller density in this region 
-suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
-It is then likely that the models that overestimate head group order parameter response~\cite{catte16}
-also overestimate the amount of bound cations.
 
@@
     for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
-    Relative surface excess for the respective simulations are: 
-    Lipid14 with \AA{}qvist ions $\Gamma_i^w = XX nm^{-3}$; 
-    Lipid14 with Dang ions $\Gamma_i^w = 0.3 nm^{-3}$; 
-    ECC-lipids with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$. 
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
@@
   \todo{JOE: Change the figure so that it contains a membrane background} \\
-  \todo{Resuls with Lipid14+Dang to be added once done -- running with larger box.}
+  \todo{Resuls with Lipid14+Dang to be added once done -- running with larger box.} \\
+  \todo{The current data for Dang simulation seems to contain more ions than others.}
 \end{figure}
+
+The good agreement with experiments indicates
+that the ECC corrected lipid model has a sufficient accuracy for realistic
+studies of details of ions binding to lipid bilayers.
+
+\todo{SAMULI: Maybe we should discuss the repeat distances and
+  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
+  }
 

### 2017-09-04 Manuscript/manuscript.tex
Manuscript small update to Samuli's writing

@@
 relatively realistic when compared with other state of the art lipid models \cite{botan15}.
-The ECC correction was applied to Lipid14 parameters by scaling partial charges of the headgroup,
+The ECC correction was applied to Lipid14 parameters by scaling partial charges of the head group,
 glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
@@
 in most lipids models, including Lipid14 \cite{ollila16}.
-In contrast, improvements in glycerol backbone and headgroup
+In contrast, improvements in glycerol backbone and head group
 parameters are required in all available lipid models~\cite{botan15}.
 
-Exploring different scaling factor values, applied to headgroup, glycerol backbone
+Exploring different scaling factor values, applied to head group, glycerol backbone
 and carbonyl charges of Lipid14 model, we found out that ion binding and
-related headgroup order parameter responses become weaker. 
+related head group order parameter responses become weaker. 
 The optimal behaviour of ion binding was observed with
@@
 it reduced the area per molecule of lipid bilayer without ions below experimental
-values. Simulations with Lipid14 parameters having partial charges of headgroup, glycerol
+values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
 backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
@@
 
-\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept} \label{section:electrometer}
+\subsection{Comparing head group order parameter response to bound charge in simulations and experiments by using the electrometer concept} \label{section:electrometer}
 Ion binding was compared between experiments and simulations by 
@@
 
-The order parameters for $\alpha$ and $\beta$ segments in headgroup, as well as for
+The order parameters for $\alpha$ and $\beta$ segments in head group, as well as for
 all C-H bonds in lipid molecules,
@@
 Before using electrometer concept to analyze cation binding, 
-we calibrate the response of headgroup order parameters to the bound
+we calibrate the response of head group order parameters to the bound
 charge in the simulation models. This is done by comparing 
@@
 \subsection{Ion concentrations and binding affinity}
-The experimental headgroup order parameter data used
+The experimental head group order parameter data used
 in this work is reported as a function of 
@@
 solvent \cite{akutsu81,altenbach84}. 
+\todo{JOE: Following is hard to understand 
+(even though I think I know the idea. 
+Should be rewritten so that it is clear.}
 To determine the same concentration from simulations, we
@@
 where the bulk concentrations, $C_w$ and $C_i$, can be determined
+\todo{JOE: I wouldn't use the term plateau. Term ``region where concentration is constant'' is preferred. Applies to the whole text -- replace everywhere. }
 from the concentration plateau at the edge of the simulation box.
@@
 The order parameters were calculated from simulations for all C-H bonds
-in lipid molecule by using Eq. \ref{OP}. Form factors were calculated 
+in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
 from equation
@@
 The systems were simulated for 200~ns using Lipid14 model for POPC.
-First 20~ns was omitted from the analysis.
+First 20~ns were omitted from the analysis.
 
-The same systems were also simulated with ECC-lipid model for POPC \todo{How long?}.
+The same systems were also simulated with ECC-lipid model for POPC using the same setup. 
 In these simulations the ECC correction was also applied to the cationic surfactant 
-by scaling all \todo{Were all scaled? Are there charges is tails? Were those scaled as well in this case?} 
-charges with the same factor as for ECC-lipids, $f_q=0.8$, 
+by scaling all charges with the same factor as for ECC-lipids, $f_q=0.8$, 
 and by using the atom types with reduced $\sigma$ parameters from ECC-lipids. 
+\todo{Were all scaled? Are there charges is tails? Were those scaled as well in this case? 
+      -- JOE: yes, all scaled for simplicity. }
 
@@
 
-\subsection{Response of POPC headgroup to bound charge}
+\subsection{Response of POPC head group to bound charge}
 
@@
 Before proceeding to ion binding affinity studies, we quantify
-the response of headgroup order parameters to the amount of 
+the response of head group order parameters to the amount of 
 bound charge by using mixtures of monovalent cationic surfactants
@@
 given by the molar fraction of cationic surfactants, because essentially all surfactants 
-locate in the lipid bilayers. Comparison of headgroup order parameter
+locate in the lipid bilayers. Comparison of head group order parameter
 changes between simulations and experiments 
@@
 can be then used to validate the sensitivity of
-lipid headgroup to bound charge in simulation models.
+lipid head group to bound charge in simulation models.
 
 %depends on both, the
-%binding affinity and the response of the headgroup on bound charge.
-%Thus, the headgroup response to bound charge has to be correctly
+%binding affinity and the response of the head group on bound charge.
+%Thus, the head group response to bound charge has to be correctly
 %described in the model if the concept is used for quantitative studies of
@@
 %
-%Thus, the headgroup order parameter response to bound charge can be first
+%Thus, the head group order parameter response to bound charge can be first
 %quantified aginst experimental data before proceeding to binding affinity studies.
@@
 %thus the exact amount of bound charge in the membrane is known in these systems.
-%The headgroup order parameter changes as a function of bound charges can be then
+%The head group order parameter changes as a function of bound charges can be then
 %explicitly determined from such systems.
 
-The headgroup order parameter changes as a function of mole fraction
+The head group order parameter changes as a function of mole fraction
 of cationic surfactant dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) 
@@
 The slope is, however, too steep in Lipid14 model indicating that 
-the headgroup order parameters are too sensitive to a bound charge.
+the head group order parameters are too sensitive to a bound charge.
 The ECC-lipids model gives a slope in very good agreement with experiments
@@
 Also the effect of bound cations on the
-headgroup P-N vector angle with respect to the membrane normal
+head group P-N vector angle with respect to the membrane normal
 is shown in Fig.~\ref{OrderParameterCHANGESsurf}.
-The headgroup orients more parallel to the membrane
+The head group orients more parallel to the membrane
 normal with an increasing amount of bound cations, as expected~\cite{seelig87}. 
 The orientation effect is more pronounced in Lipid14 model than 
-in ECC-lipids model, which is in line with the order parameter 
-results and reduced charge--dipole interactions in the ECC-lipid model.
+in ECC-lipids model, which is in line with the higher order parameter 
+response and reduced charge--dipole interactions in the ECC-lipid model.
 Since the response of $\alpha$-order parameter is in very good agreement
@@
   \caption{\label{OrderParameterCHANGESnewMODELS}
-    Changes of headgroup order parameters of POPC bilayer as a function of NaCl and CaCl$_2$ concentrations
+    Changes of head group order parameters of POPC bilayer as a function of NaCl and CaCl$_2$ concentrations
     are shown from simulations with different force fields together with experimental data 
@@
 
-The linear decrease of the headgroup order parameters, quantified in previous 
+The linear decrease of the head group order parameters, quantified in previous 
 section, can be used to measure the amount of bound aqueous cations in lipid 
-bilayers~\cite{seelig87,catte16}. The headgroup order parameter response 
+bilayers~\cite{seelig87,catte16}. The head group order parameter response 
 to NaCl and CaCl$_2$ concentrations from experiments (DPPC \cite{akutsu81} 
@@
 As already discussed previously~\cite{catte16}, NaCl has a negligible effect 
-on the headgroup order parameters with submolar 
+\todo{JOE: catte16 is not the original work that introduces NaCl effects -- on the
+           contrary, the presented simulations (which for the core of the original work there)
+           rather suggest that there is some effect. 
+           I think this has to be started from a different perspective.
+           For example, we can start with saying that we use Na and Ca as biologically relevant cations
+           and they also are representants of mono- and divalent cations. 
+           In addition, they also represent different behaviour accordin to experimental
+           measurements of head group order parameter. }
+on the head group order parameters with submolar 
 concentrations due to the very weak binding in PC bilayers. 
@@
 is reproduced by Lipid14 model with \AA{}qvist ions \cite{catte16}. 
-The headgroup order parameter response to CaCl$_2$ concentration was,
+The head group order parameter response to CaCl$_2$ concentration was,
 however, overestimated with the same combination of force field
@@
 
-Significant improvement is achieved when the developed ECC-lipid model 
-is used. The headgroup order parameter changes as a function of CaCl$_2$
-concentration model show a good agreement between ECC-lipid model and 
+Significant improvement is achieved with the newly developed ECC-lipid model. 
+The head group order parameter changes as a function of CaCl$_2$
+concentration show a good agreement between ECC-lipid model and 
 experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
@@
 and Dang ions $\Gamma_i^w = 0.3 nm^{-3}$.
+\todo{JOE: Add error estimates to Gamma values}
 %The density profiles from simulations with original Lipid14 \cite{dickson14}
@@
   \todo{JOE: Change the figure so that it contains a membrane background} \\
-  \todo{Resuls with Lipid14+Dang to be added once done -- running with larger box.} \\
   \todo{The current data for Dang simulation seems to contain more ions than others.}

### 2017-09-05 Manuscript/manuscript.tex
Added relative surface excess for L14+Aqvist into the MS with corresponding data into Data folder. Added a simple info-graphic (conceptual!) for the description of bindin sites in POPC. Only ECC-lipids model is presented (it's the simplest) Manuscript includes the new figure.

@@
 The relative surface excess for ECC-lipid model with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$
-is significantly smaller than for Lipid14 with \AA{}qvist $\Gamma_i^w = XX nm^{-3}$
-and Dang ions $\Gamma_i^w = 0.3 nm^{-3}$.
-\todo{JOE: Add error estimates to Gamma values}
+is significantly smaller than for Lipid14 with \AA{}qvist $\Gamma_i^w = 0.13 nm^{-3}$
+and Dang ions $\Gamma_i^w = 0.3 nm^{-3}$ (error estimate being below 10\%).
 %The density profiles from simulations with original Lipid14 \cite{dickson14}
@@
 \todo{This paragraph is to be finalized when we have the surface excess analysis
-  done by using Eq.~\ref{surfexcess}.}
+  done by using Eq.~\ref{surfexcess}.
+  JOE: Relative surface excess values are there, so let's finalize it.}
 
@@
 
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[height=6.0cm]{../Fig/POPC_binding_labels.eps} 
+  \includegraphics[height=6.0cm]{../Fig/POPC_binding_ECC-lipids.eps}
+  \caption{\label{fig:binding_conf}
+      Configurations of \ce{Ca2+} binding in POPC membranes in the ECC-lipids model.
+      The binding sites are: 
+      red circle for phosphate oxygens; 
+      blue square for sn-1 carbony oxygen; and green triangle for sn-2 carbonyl oxygen.
+      The configurations along with their probabilities are drawn in the following manner:
+      each row corresponds to an individual lipid the cation is bound to;
+      symbols denote the site the cations is bound to the respective lipid.
+      An empty configuration denotes an unbound cation in a solvent. 
+  }
+\end{figure}
+
 \section{Conclusions}

### 2017-09-05 Manuscript/manuscript.tex
Update binding affinity section, not ready yet

--- 2017-09-03 Manuscript/manuscript.tex
@@
 
-As already discussed previously~\cite{catte16}, NaCl has a negligible effect 
-on the headgroup order parameters with submolar 
-concentrations due to the very weak binding in PC bilayers. 
-In contrast to almost all the other simulation models, this behaviour 
-is reproduced by Lipid14 model with \AA{}qvist ions \cite{catte16}. 
-The headgroup order parameter response to CaCl$_2$ concentration was,
-however, overestimated with the same combination of force field
-parameters~\cite{catte16}. The results were not improved when
-ion models by Dang et al.~\cite{smith94,chang1999,dang2006} or
+Experimentally NaCl has a negligible effect on the headgroup order parameters
+with submolar concentrations due to the very weak binding of Na$^+$
+in PC bilayers \cite{akutsu81,catte16}. 
+While almost all the available simulation models overestimate cation
+binding in PC lipids, the weak binding of Sodium ions was reproduced by
+Lipid14 model when simulated with \AA{}qvist ions~\cite{catte16}.
+However, the headgroup order parameter response to CaCl$_2$ concentration was
+overestimated with the same combination of force field parameters~\cite{catte16}.
+Changing to ion models by Dang et al.~\cite{smith94,chang1999,dang2006} or
 ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
-with more realistic bulk behaviour were used, as seen in 
+with more realistic bulk behaviour were used did not improve the results
+for the CaCl$_2$ interactions with Lipid14 model, as seen in 
 in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively.
@@
 Significant improvement is achieved when the developed ECC-lipid model 
-is used. The headgroup order parameter changes as a function of CaCl$_2$
-concentration model show a good agreement between ECC-lipid model and 
-experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-Ion binding affinity is quantified by using the relative surface
-excess between ions and water from Eq.~\ref{surfexcess} and  
-density profiles along membrane normal shown in Fig.~\ref{fig:cacl-dens}.
-%This is directly related to the amount of surface excess charge 
-%as shown along with ion density profiles in  Fig.~\ref{fig:cacl-dens}.
-The relative surface excess for ECC-lipid model with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$
-is significantly smaller than for Lipid14 with \AA{}qvist $\Gamma_i^w = XX nm^{-3}$
-and Dang ions $\Gamma_i^w = 0.3 nm^{-3}$.
-%The density profiles from simulations with original Lipid14 \cite{dickson14}
-%and Dang ions \cite{smith94,chang1999,dang2006} 
-%show a pronounced peak in the position of the phosphate moieties of POPC. 
-%The use of ECC-ion model~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
-%or the ion model by \AA{qvist}~\cite{aqvist90}
-%along with Lipid14 does not significantly change it 
-%(Headgroup order parameter responses for these models in SI).
-%The new ECC-lipids model with ECC-ions~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}
-%exhibits on the other hand smaller density in this region 
-%suggesting overall weaker binding of cations (Fig.~\ref{OrderParameterCHANGESnewMODELS}). 
-%It is then likely that the models that overestimate head group order parameter response~\cite{catte16}
-%also overestimate the amount of bound cations.
-\todo{This paragraph is to be finalized when we have the surface excess analysis
-  done by using Eq.~\ref{surfexcess}.}
+is used together with ECC-ions. The headgroup order parameter changes
+as a function of CaCl$_2$ concentration show a good agreement between
+ECC-lipid model and experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+As discussed in previous section, also the response of headgroup order
+parameters to the bound charge is in good agreement with the experiments in this
+model. Thus, we conclude that the ECC-lipid model with ECC-ion parameters
+gives a realistic binding affinity for the Na$^+$ and Ca$^{2+}$ binding
+in POPC bilayer.
+
+Based on the ion density profiles along membrane normal
+in Fig.~\ref{fig:cacl-dens}, the Ca$^{2+}$ binding seems to
+be stronger in Lipid14 model with Dang and ECC-ions.
+This is in agreement 
+This is quantified by calculating the relative surface
+excess from Eq.~\ref{surfexcess}. This gives 
+$\Gamma_i^w = 0.07 nm^{-3}$ for ECC-lipid model simulated with ECC-ions,
+which is significantly smaller than $\Gamma_i^w = XX nm^{-3}$ for
+Lipid14 with \AA{}qvist and $\Gamma_i^w = 0.3 nm^{-3}$ with Dang ions.
+\todo{Surface excess for Lipid14 with Aqvist yet to be added.} \\
+\todo{Surface excess units are not correct. Maybe the numbers needs to be checked also?} \\
+Rough estimates for the free energy difference between bound and unbound cation
+can made from
+\begin{equation}
+  \Delta G={\rm k_b T}\log(\frac{p_o}{p_i}),
+\end{equation}
+where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
+in the maximum density in bilayer, respectively. The density profiles in
+Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
+between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
+in Lipid14 with \AA{}qvist.
+
+\todo{SAMULI: Maybe we should discuss the repeat distances and
+  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
+  }
 
@@
 
+Since ECC--lipid model gives a good agreement with experiments for the response
+of headgroup order parameters to the bound charge as well as to the CaCl$_2$
+concentration,
+
 The good agreement with experiments indicates
@@
 studies of details of ions binding to lipid bilayers.
-
-\todo{SAMULI: Maybe we should discuss the repeat distances and
-  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
-  }
 

### 2017-09-05 Manuscript/manuscript.tex
Update results and discussion, binding affinity section is not ready yet

@@
 
-Before proceeding to ion binding affinity studies, we quantify
+Before proceeding to the ion binding affinity studies, we quantify
 the response of headgroup order parameters to the amount of 
@@
 and POPC~\cite{scherer89}. The amount of bound charge per PC 
-in these systems is 
-given by the molar fraction of cationic surfactants, because essentially all surfactants 
-locate in the lipid bilayers. Comparison of headgroup order parameter
-changes between simulations and experiments 
-as a function of mole fraction of cationic surfactants 
-can be then used to validate the sensitivity of
-lipid headgroup to bound charge in simulation models.
+in these systems is given by the molar fraction of cationic 
+surfactants, because essentially all surfactants locate in the 
+lipid bilayers. Experimental data for these systems can be used to validate 
+the sensitivity of lipid headgroup order parameters
+to the amount of bound charge in simulations.
 
@@
 
-The headgroup order parameter changes as a function of mole fraction
-of cationic surfactant dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) 
+The headgroup order parameter changes with increasing amount of 
+cationic surfactant dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) 
 is compared between experiments~\cite{scherer89} and simulations in
 Fig. \ref{OrderParameterCHANGESsurf}. 
-Approximately linear decrease of the order parameters is observed
+Approximately linear decrease is observed
 in simulations and experiments with increasing amount of cationic
@@
 
-Also the effect of bound cations on the
-headgroup P-N vector angle with respect to the membrane normal
-is shown in Fig.~\ref{OrderParameterCHANGESsurf}.
+The headgroup P-N vector angle with respect to the membrane normal
+is also shown in Fig.~\ref{OrderParameterCHANGESsurf}
+as a function of cationic surfactant mole fraction.
 The headgroup orients more parallel to the membrane
-normal with an increasing amount of bound cations, as expected~\cite{seelig87}. 
-The orientation effect is more pronounced in Lipid14 model than 
+normal with increasing amount of bound cations, as expected~\cite{seelig87}. 
+The effect is more pronounced in Lipid14 than 
 in ECC-lipids model, which is in line with the order parameter 
-results and reduced charge--dipole interactions in the ECC-lipid model.
-Since the response of $\alpha$-order parameter is in very good agreement
-with experiments in ECC-lipid model, we can suggest a relation between
-order parameter and P-N vector angle changes. The data gives
-$\Delta \Theta_{\rm P-N}=?? \Delta S_{\rm CH}$, which can be used
-to estimate P-N vector angle changes from experimental results
-for $\alpha$ order parameters. 
+results and the reduced charge--dipole interactions in the ECC-lipid model.
+The response of $\alpha$-order parameter to bound positive charge
+in ECC-lipid model is in good agreement with experiments.
+Therefore we use the data in Fig.~\ref{OrderParameterCHANGESsurf}
+to suggest a linear relation between $\alpha$-carbon order parameter
+and average P-N vector angle changes $\Delta \Theta_{\rm P-N}=?? \Delta S_{\rm CH}^{\alpha}$.
+This can be used to estimate the change in P-N vector angle
+from experiments measuring changes in $\alpha$-carbon order parameter. 
 \todo{SAMULI: I think that we should establish the relation between 
@@
 changes from experiments.}
+For example, $\Delta S_{\rm CH}^{\alpha}=0.018$ measured due to the
+addition of 66.8 $\mu$M concentration of Mellittin in bulk water \cite{kuchinka89}
+gives $\Delta \Theta_{\rm P-N} = 2.5^{\circ}$.
+\todo{I looked this roughly from the figure, should be calculated from 
+the actual relation if we decide to put this.}
 
@@
 
-The linear decrease of the headgroup order parameters, quantified in previous 
-section, can be used to measure the amount of bound aqueous cations in lipid 
-bilayers~\cite{seelig87,catte16}. The headgroup order parameter response 
-to NaCl and CaCl$_2$ concentrations from experiments (DPPC \cite{akutsu81} 
+The binding affinity of aqueous cations in lipid 
+bilayers can be measured by using the headgroup 
+order parameters, because they decrease linearly 
+with the bound positive charge~\cite{seelig87,catte16}. 
+The headgroup order parameter responses 
+to aqueous NaCl and CaCl$_2$ concentrations 
+are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}
+from experiments (DPPC \cite{akutsu81} 
 and POPC \cite{altenbach84}) and different simulation models for POPC
-are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+
 %The simulation results of Lipid14 with \AA{}qvist ions~\cite{aqvist90} (from Ref. \citenum{catte16})
@@
 
-Experimentally NaCl has a negligible effect on the headgroup order parameters
-with submolar concentrations due to the very weak binding of Na$^+$
-in PC bilayers \cite{akutsu81,catte16}. 
-While almost all the available simulation models overestimate cation
-binding in PC lipids, the weak binding of Sodium ions was reproduced by
-Lipid14 model when simulated with \AA{}qvist ions~\cite{catte16}.
-However, the headgroup order parameter response to CaCl$_2$ concentration was
-overestimated with the same combination of force field parameters~\cite{catte16}.
-Changing to ion models by Dang et al.~\cite{smith94,chang1999,dang2006} or
+Negligible changes of the headgroup order parameters are measured
+with submolar concentrations of NaCl due to the very low affinity of Na$^+$
+in PC bilayers \cite{akutsu81}. While Na$^+$ binding and the related 
+headgroup order parameter changes were overestimated in
+almost all the available simulation models, the low affinity
+and negligible order parameter changes were reproduced by Lipid14 
+model when simulated with \AA{}qvist ions~\cite{catte16}.
+However, the same combination of force field parameters
+overestimated the headgroup order parameter response to CaCl$_2$ 
+concentration, which was the case also in all other models tested in
+Ref.~\citenum{catte16}.
+Using ion model by Dang et al.~\cite{smith94,chang1999,dang2006} or
 ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
-with more realistic bulk behaviour were used did not improve the results
+with more realistic bulk behaviour did not improve the results
 for the CaCl$_2$ interactions with Lipid14 model, as seen in 
@@
 \todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
-Together with extensive data in previous work \cite{catte16}, 
-the results suggest that CaCl$_2$ binding behaviour in PC bilayers
-cannot be corrected by only improving the ion models.
+The results support the conclusion of the previous work \cite{catte16}
+that improvements also in lipid models are needed to 
+correctly describe cation binding in PC bilayers.
 

### 2017-09-06 Manuscript/manuscript.tex
Update text, I am now done with the affinity section for now

@@
 According to ECC, electronic
-polarizability can be included in classical MD simulations by
+polarizability can be implicitly included in classical MD simulations by
 placing all particles into a homogeneous dielectric continuum 
@@
 Unlike the total charge of molecules, atomic partial charges within each molecule are not experimental observables.
+\todo{Text should be clarified from here to the end of the paragraph}
 Although there have been proposed several schemes for the assignment of partial charges of molecules, 
@@
 The decrease of area was found to arise from a lower hydration of the lipid head group region,
-which can be explained by the decreased solvation free energy due to the lower polarity
+which can be explained by the increased solvation free energy due to the lower polarity
 of molecules with scaled charges. The hydration can be increased
@@
 similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
-This increases solvation free energy by allowing water molecules to
+This decreases the solvation free energy by allowing water molecules to
 approach closer to lipid atoms and have stonger electrostatic interactions with them.
@@
 The headgroup P-N vector angle with respect to the membrane normal
-is also shown in Fig.~\ref{OrderParameterCHANGESsurf}
-as a function of cationic surfactant mole fraction.
+is also shown as a function of cationic surfactant mole fraction
+in Fig.~\ref{OrderParameterCHANGESsurf}.
 The headgroup orients more parallel to the membrane
@@
 
-Significant improvement is achieved when the developed ECC-lipid model 
-is used together with ECC-ions. The headgroup order parameter changes
-as a function of CaCl$_2$ concentration show a good agreement between
-ECC-lipid model and experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
-As discussed in previous section, also the response of headgroup order
-parameters to the bound charge is in good agreement with the experiments in this
-model. Thus, we conclude that the ECC-lipid model with ECC-ion parameters
-gives a realistic binding affinity for the Na$^+$ and Ca$^{2+}$ binding
-in POPC bilayer.
+Significant improvement can be achieved by using the ECC
+approach also for lipids. The headgroup order parameter changes
+as a function of CaCl$_2$ concentration from ECC-lipid model with ECC-ions
+show a good agreement with experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+As discussed in previous section, the model gives also
+a good agreement with experiments for
+the headgroup response to bound charge.
+Thus, the model can used for more detailed analysis of the
+binding affinity.
 
-Based on the ion density profiles along membrane normal
-in Fig.~\ref{fig:cacl-dens}, the Ca$^{2+}$ binding seems to
-be stronger in Lipid14 model with Dang and ECC-ions.
-This is in agreement 
-This is quantified by calculating the relative surface
-excess from Eq.~\ref{surfexcess}. This gives 
-$\Gamma_i^w = 0.07 nm^{-3}$ for ECC-lipid model simulated with ECC-ions,
+The binding affinities are quantified by using
+the ion density profiles along membrane normal shown
+in Fig.~\ref{fig:cacl-dens}. The density profiles show
+larger Ca$^{2+}$ density peak in lipid headgroup region for
+Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
+The relative surface excess calculated from Eq.~\ref{surfexcess} gives 
+$\Gamma_i^w = 0.07 nm^{-3}$ for the ECC-lipid model,
 which is significantly smaller than $\Gamma_i^w = XX nm^{-3}$ for
@@
 Rough estimates for the free energy difference between bound and unbound cation
-can made from
+are given by
 \begin{equation}
@@
 
-Since ECC--lipid model gives a good agreement with experiments for the response
-of headgroup order parameters to the bound charge as well as to the CaCl$_2$
-concentration,
+Since the lipid headgroup order parameter responses to the amount of bound charge
+and to the aqueous ion concentrations are both in good agreement with experiments
+in the ECC-lipid model with ECC-ion parameters, we consider the Na$^+$ and Ca$^{2+}$
+binding affinities to be realistic in this model. On the other hand, the Ca$^{2+}$
+binding affinity is overestimated by Lipid14 model when simulated with all the
+tested ion models. Similar conclusions were previously made based only on the headrogup
+order parameter data with aqueous cations~\cite{catte16}. However,
+the discrepancies with experiments in previous work could partly arise also from
+the inaccurate sensitivity of the headgroup to bound charge. Here we quantify this
+effect and conclude that the improvement due to ECC is partly, but not completely,
+caused by more realistic headgroup sensitivity to bound charge.
+This indicates that the issue should be carefully considered also when
+the electrometer concept is used to compare ion binding between experiments
+and simulations wiht other models.
 
-The good agreement with experiments indicates
-that the ECC corrected lipid model has a sufficient accuracy for realistic
-studies of details of ions binding to lipid bilayers.
+%The good agreement with experiments indicates
+%that the ECC corrected lipid model has a sufficient accuracy for realistic
+%studies of details of ions binding to lipid bilayers.
 

### 2017-09-06 Manuscript/manuscript.tex

@@
 
-\subsection{Comparison of ion binding affinity to experiments by using the electrometer concept} \label{section:electrometer}
+\subsection{Electrometer concept} \label{section:electrometer}
 Ion binding was compared between experiments and simulations by 
@@
 \subsection{Ion concentrations and binding affinity}
-The experimental headgroup order parameter data used
-in this work is reported as a function of 
-cation concentration remaining in the bulk 
-solvent \cite{akutsu81,altenbach84}. 
-To determine the same concentration from simulations, we
-use large enough box to ensure that the ion concetration
-reaches a clear plateau and the electroneutrality
-is satisfied at the edge of simulation box. It should be noted that the recent work
-by NMRlipids project compared simulations to the same 
-experimental data by using the relative molar concentration in water
-[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M \cite{catte16}.
-This is not exactly the same as the concentration reported in the used experimental
-data, because the binding of ions in bilayer reduces the bulk concentration.
-The difference slightly moves points towards lower concentrations but do not 
-affect the conclusions in the previous study \cite{catte16}.
+The salt concentrations are reported in two different ways
+in the used experimental data for the lipid head group order
+parameters. The data for DPPC is reported
+by using solution concentrations before solvating
+the lipids~\cite{akutsu81}, while the data
+for POPC~\cite{altenbach84} is reported by using
+supernatant concetration measured with atomic absorption
+spectroscopy after solvation of lipids.
+The former was used in comparison with experiments and
+calculated from simulations in previous work as 
+[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M~\cite{catte16}.
+Here we use the latter, because the focus is on POPC bilayers.
+This is determined from the ion concentrations, which reach
+a constant value at the edges of large enough simulation box.
+The two concentrations differ somewhat for CaCl$_2$ systems, because some
+amount of ions bind in lipids. However, the small difference do
+not affect the conclusions in this or in previous work~\cite{catte16}.
 
-The ion binding affinities between different models were compared by calculating
-the relative surface excess of ions with respect to water, $\Gamma_i^w$, as \cite{chattorajBOOK}
+The ion concentrations with constant value at the simulation box edges
+are also used to quantify the ion binding affinities is different simulation
+models by using the relative surface excess of ions with respect to water $\Gamma_i^w$.
+This is defined as \cite{chattorajBOOK}
 \begin{eqnarray}
@@
 we consider the whole simulation box as an interface. 
-Note that phospholipid membranes are bilayers, 
-and hence the interface occurs in the simulation box twice
-providing a surface area $2A$.
+Note that the area of the interface is given by the simulation box area
+multiplied with two $2A_b$, because bilayers have a interface
+in both leaflefts.
 The concentrations of water and ions inside
@@
 \begin{equation}\label{surfexcess}
-  \Gamma_i^w=\frac{1}{2A} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
+  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
 \end{equation}
-where the bulk concentrations, $C_w$ and $C_i$, can be determined
-from the concentration plateau at the edge of the simulation box.
+where the bulk concentrations, $C_w$ and $C_i$, are 
+concentrations with constant value at the simulation box edges.
 
-\subsection{Comparison of lipid bilayer structure to experiments}
+\subsection{Validation of lipid bilayer structure against experiments}
 
@@
 and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
+\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
+  this is done in Amber by default. If this is correct, the information in the
+  table do not match with this sentence. Based on semi-extensive testing I made few years
+ago this do not change anything. Anyway, this should be reported consistently.}
 Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
@@
 The systems were simulated for 200~ns using Lipid14 model for POPC.
-First 20~ns was omitted from the analysis.
+First 20~ns were omitted from the analysis.
 
-The same systems were also simulated with ECC-lipid model for POPC \todo{How long?}.
+The same systems were also simulated with ECC-lipid model for POPC using the same setup. 
 In these simulations the ECC correction was also applied to the cationic surfactant 
-by scaling all \todo{Were all scaled? Are there charges is tails? Were those scaled as well in this case?} 
-charges with the same factor as for ECC-lipids, $f_q=0.8$, 
+by scaling all charges with the same factor as for ECC-lipids, $f_q=0.8$, 
 and by using the atom types with reduced $\sigma$ parameters from ECC-lipids.
@@
     The scale of the y-axis should be then zoomed when this is fixed.} \\
-  \todo{x-axis scale in form factor figure from 0 to 0.6, until where experimental data is available.}
+  \todo{I would put the acyl chain order parameters also for original Lipid14}\\
+  \todo{x-axis scale in form factor figure would be better from 0 to 0.6. The experimental data is available for this range.} \\
+  \todo{$S_{CD} -> S_{CH}$}
 \end{figure}
@@
 %depends on both, the
-%binding affinity and the response of the headgroup on bound charge.
-%Thus, the headgroup response to bound charge has to be correctly
+%binding affinity and the response of the head group on bound charge.
+%Thus, the head group response to bound charge has to be correctly
 %described in the model if the concept is used for quantitative studies of
@@
 %thus the exact amount of bound charge in the membrane is known in these systems.
-%The headgroup order parameter changes as a function of bound charges can be then
+%The head group order parameter changes as a function of bound charges can be then
 %explicitly determined from such systems.
@@
     Ion concentrations in bulk water are shown in x-axis. 
-    Values from simulations are calculated from the plateau of cation number density $C_{np}$
-    at the box edge as [ion]=$C_{np}/0.602$.
+    Values from simulations are calculated from the of cation number density $C_{np}$
+    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
     Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
@@
 \todo{Surface excess units are not correct. Maybe the numbers needs to be checked also?} \\
+\todo{JOE: Add error estimates to Gamma values}
 Rough estimates for the free energy difference between bound and unbound cation

### 2017-09-07 Manuscript/manuscript.tex

@@
 we consider the whole simulation box as an interface. 
-Note that the area of the interface is given by the simulation box area
-multiplied with two $2A_b$, because bilayers have a interface
-in both leaflefts.
+Because bilayers have an interface
+in both leaflefts,
+the total area of the interface in the simulation
+ is given by the simulation box area multiplied by two, $2A_b$.
 The concentrations of water and ions inside
@@
     The scale of the y-axis should be then zoomed when this is fixed.} \\
-  \todo{I would put the acyl chain order parameters also for original Lipid14}\\
+  \todo{I would put the acyl chain order parameters also for original Lipid14; JOE: Already done, check whether your figures are up to date. I usually push only *.eps.}\\
   \todo{x-axis scale in form factor figure would be better from 0 to 0.6. The experimental data is available for this range.} \\
@@
 Fig. \ref{OrderParameterCHANGESsurf}. 
-Approximately linear decrease is observed
+Approximately linear decrease of the order parameters is observed
 in simulations and experiments with increasing amount of cationic
@@
 from experiments (DPPC \cite{akutsu81} 
-and POPC \cite{altenbach84}) and different simulation models for POPC
+and POPC \cite{altenbach84}) and different simulation models for POPC.
 
@@
 with more realistic bulk behaviour did not improve the results
-for the CaCl$_2$ interactions with Lipid14 model, as seen in 
+for the CaCl$_2$ interactions with Lipid14 model, as seen 
 in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively.
@@
 The relative surface excess calculated from Eq.~\ref{surfexcess} gives 
-$\Gamma_i^w = 0.07 nm^{-3}$ for the ECC-lipid model,
-which is significantly smaller than $\Gamma_i^w = XX nm^{-3}$ for
-Lipid14 with \AA{}qvist and $\Gamma_i^w = 0.3 nm^{-3}$ with Dang ions.
-\todo{Surface excess for Lipid14 with Aqvist yet to be added.} \\
-\todo{Surface excess units are not correct. Maybe the numbers needs to be checked also?} \\
-\todo{JOE: Add error estimates to Gamma values}
+$\Gamma_i^w = 0.07 \pm 0.01 nm^{-2}$ for the ECC-lipid model,
+which is significantly smaller than $\Gamma_i^w = 0.13 \pm 0.01 nm^{-2}$ for
+Lipid14 with \AA{}qvist and $\Gamma_i^w = 0.3 \pm 0.03 nm^{-2}$ with Dang ions.
 Rough estimates for the free energy difference between bound and unbound cation
@@
     }
-  \todo{Calculate the surface excess for L14+Aqvist sim.} \\
   \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
@@
 
+%\begin{figure}[tbp]
+%  \centering
+%  \includegraphics[height=6.0cm]{../Fig/POPC_binding_labels.eps} 
+%  \includegraphics[height=6.0cm]{../Fig/POPC_binding_ECC-lipids.eps}
+%  \caption{\label{fig:binding_conf}
+%      Configurations of \ce{Ca2+} binding in POPC membranes in the ECC-lipids model.
+%      The binding sites are: 
+%      red circle for phosphate oxygens; 
+%      blue square for sn-1 carbony oxygen; and green triangle for sn-2 carbonyl oxygen.
+%      The configurations along with their probabilities are drawn in the following manner:
+%      each row corresponds to an individual lipid the cation is bound to;
+%      symbols denote the site the cations is bound to the respective lipid.
+%      An empty configuration denotes an unbound cation in a solvent. 
+%  }
+%\end{figure}
+
 \section{Conclusions}

### 2017-09-07 Manuscript/manuscript.tex
Commit small changes before merge

--- 2017-09-06 Manuscript/manuscript.tex
@@
 \todo{JOE: list more introductory examples, where it has interesting effects.}
-Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
-to understand molecular level details of specific cation interactions with
-cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
+
+Zwitterionic phosphocholine (PC) lipid bilayers are common model systems for
+cellular membranes. Interactions of biological cations, especially Na$^+$ and
+Ca$^{2+}$ are with 
+have been widely used as
+model systems to study details of cation interactions with cellular membranes
+by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
 and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??}
-\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
+
 While relative binding affinity of different ions is agreed to follow Hoffmeister
@@
 not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
+\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
+
 %General conclusion from experimental

### 2017-09-07 Manuscript/manuscript.tex
MS: minor update

--- 2017-09-07 Manuscript/manuscript.tex
@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
-  \caption{\label{fig:cacl-bind}
-    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .
-    }
-\end{figure}
-
 The good agreement of ECC-lipids model with experiments encourages us to analyse the
@@
   }
+  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca2+}}
 \end{figure}

### 2017-09-07 Manuscript/manuscript.tex
Party updated intro, not ready yet

--- 2017-09-07 Manuscript/manuscript.tex
@@
 like in signal propagation in neurons and vesicle fusion.
-\todo{JOE: list more introductory examples, where it has interesting effects.}
+\todo{This could be maybe extended.}
+Since the direct measurements of ion-membrane interactions from biological systems
+are difficult, lipid bilayers are often used as models systems for cellular
+membranes. The detailed results from simple the model systems can be then
+used to understand the role of ions in complex biological systems.
 
-Zwitterionic phosphocholine (PC) lipid bilayers are common model systems for
+Zwitterionic phosphocholine (PC) lipid bilayers are commonly used model systems for
 cellular membranes. Interactions of biological cations, especially Na$^+$ and
-Ca$^{2+}$ are with 
-have been widely used as
-model systems to study details of cation interactions with cellular membranes
-by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
-and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??}
-
-While relative binding affinity of different ions is agreed to follow Hoffmeister
-series, the molecular details of binding and binding energetics are
-not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
-\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
-
+Ca$^{2+}$, with PC bilayers are widely studied with
+experiments~\cite{akutsu81,altenbach84,seelig90,cevc90,tocanne90,binder02,pabst07,uhrikova08}
+and classical MD simulations~\cite{bockmann03,bockmann04,Berkowitz12,melcrova16,javanainen17}.
+The details of ion binding are, however, not agreed in the literature.
+%not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
 %General conclusion from experimental
@@
 %ions, but do not affect the specific binding constant \cite{seelig90}.
-Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
+Non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
 and infrared scpectroscopy, give accurate information about ion binding
-in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
-Molecular level interpretation of the results, however, requires assumptions
-about binding models and is often not fully conclusive \cite{??}.
-On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
-have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
-The overestimated specific cation binding observed in current lipid models \cite{catte16}
-may lead to articifially positively charged memranes and significant artefacts in
-MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
-\todo{JOE: this may be a too daring statement. \\
-  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
-  but it might happen with too small concentrations. 
-  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).
+in PC lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
+Interpretation of these experiments suggests that Na$^+$ ions have
+negligible binding in PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
+binds to phosphate groups of two lipid molecules. Atomistic resolution
+molecular dynamics (MD) simulation models, however, predict significantly
+stronger binding for both ions and interactions with 3-4 lipids,
+including also interactions with carbonyl
+oxygens~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
+Some experiments have also been interpreted to support the predictions from
+MD simulations~\cite{bockmann03,vacha09a}.
+
+Recent work published by the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
+made an attempt to resolve the apparent controversies by introducing a direct
+method to compare ion binding affitity between simulations and
+experiments by using the electrometer concept~\cite{seelig87}.
+Using massive data collected by Open Collaboration method, it was concluded
+that the accuracy of the current state of the art lipid models for MD
+simulations is not sufficient for the detailed interpretation of the cation
+interactions with PC lipid bilayers~\cite{catte16}.
+
+%\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
+
+%While relative binding affinity of different ions in PC lipids is agreed to follow Hoffmeister
+%series, the molecular details of binding and binding energetics are
 
@@
 
-In this work we show that the accuracy of ion binding, especially calcium,
-in zwitterionic PC lipid bilayer can be significantly improved 
-for classical MD simulation models of zwitterionic lipids
-by including electronic polarizability
-in the polar region of lipid molecules
-using the electronic continuum correction (ECC)~\cite{leontyev11}. 
-This approach has been previously shown to improve performace of
-ion models against neutron scattering data in concentrated solutions 
-and ab-initio simulations~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
-The realistic structure of ions in a bulk solvent was not, however, sufficient to correct binding in lipid
-bilayers~\cite{catte16}. 
-To fix it, we propose
-a classical fixed-charge MD simulation model of 
-1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) lipid 
-that accounts for electronic polarizability by using ECC. 
+In this work we show that the cation binding behavior
+in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
+improved by implicitly including electronic polarizability
+in the polar region of lipid molecules. The electronic
+polarizability is included by using the electronic
+continuum correction (ECC)~\cite{leontyev11}, which
+has been previously shown to improve the behaviour of
+MD simulations of ions in bulk water~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}.
+
 We will validate such a model by comparing its lipid bilayer structue
@@
 and for effects where membrane electrostatics play a major role.
+
+The overestimated specific cation binding observed in current lipid models \cite{catte16}
+may lead to articifially positively charged memranes and significant artefacts in
+MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
+
+%\todo{JOE: this may be a too daring statement. \\
+%  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
+%  but it might happen with too small concentrations. 
+%  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).
 

### 2017-09-07 Manuscript/manuscript.tex
MS:small update, added figure with probabilities of ca-lipid complexes
New intro

@@
 are difficult, lipid bilayers are often used as models systems for cellular
-membranes. The detailed results from simple the model systems can be then
+membranes. The detailed results from simple model systems can be then
 used to understand the role of ions in complex biological systems.
@@
 molecular dynamics (MD) simulation models, however, predict significantly
-stronger binding for both ions and interactions with 3-4 lipids,
+stronger binding for the cations and interactions with 3-4 lipids,
 including also interactions with carbonyl
@@
 Recent work published by the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
-made an attempt to resolve the apparent controversies by introducing a direct
-method to compare ion binding affitity between simulations and
-experiments by using the electrometer concept~\cite{seelig87}.
+made an attempt to resolve the apparent controversies. The study presented 
+a direct comparison between ion binding affinity between simulations and
+experiments by using NMR data for lipid headgroup order 
+parameters and the electrometer concept~\cite{seelig87}.
 Using massive data collected by Open Collaboration method, it was concluded
@@
 MD simulations of ions in bulk water~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
+As a starting point we use the parameters from Lipid14 model \cite{dickson14},
+which gave the best cation binding behaviour in the previous study \cite{catte16}.
+The developed ECC-lipid parameters reproduce the experimentally
+measurable structural parameters of POPC lipid bilayer with the 
+accuracy comparable to the other state of the art lipid models.
 
-We will validate such a model by comparing its lipid bilayer structue
-to x-ray scattering data and its head group structure to NMR experiments. 
-We will then quantify the accuracy of the head group order parameter response to the bound charge
-and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
-In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
-This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
-The proposed POPC model, ECC-lipids, and the approach to obtain it
-is, hence, highly useful for future MD simulations with physiological salt conditions, 
-and for effects where membrane electrostatics play a major role.
+%We will validate such a model by comparing its lipid bilayer structure
+%to x-ray scattering data and its head group structure to NMR experiments. 
+%We will then quantify the accuracy of the head group order parameter response to the bound charge
+%and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
+%In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
+%This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
 
-The overestimated specific cation binding observed in current lipid models \cite{catte16}
-may lead to articifially positively charged memranes and significant artefacts in
-MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
+Since the cation binding affinity and the headgroup 
+order parameter changes are in good agreement with experiments
+in the proposed ECC-lipids, it can be used to interpret the
+related structural changes and ion binding stoichiometry.
+New lipid models with correct ion binding affinity in lipid
+bilayers are necessary in applications of 
+MD simulations with physiological salt conditions. 
+The overestimated cation binding in the current lipid models \cite{catte16}
+may lead to significant artefacts in MD simulations. For example,
+articifially positively charged membranes overestimate interactions
+with biomolecules having opposite sign.
+
+%MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
 

### 2017-09-08 Manuscript/manuscript.tex
Modified ternary complex binding plot for simulations -- now shows only simulations. Added a short discussion to it into SI in MS.
Started writing Binding stoichiometry section.

--- 2017-09-07 Manuscript/manuscript.tex
@@
 
+Binding stoichiometry of \ce{Ca2+} and POPC was 
+thoroughly studied in the experimental work~\citenum{altenbach84}.
+Several binding models were proposed and tested of which only one,
+ternary complex binding model, provided a good fit of the experimental observations. 
+In such a model, it is assumed that \ce{Ca2+} cations bind to a POPC membrane 
+with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
+Such a model also provides a good fit 
+to our simulations with ECC-lipids 
+(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
+Simulations, however, allow us to directly evaluate the stoichiometry 
+by calculating relative propensities of various \ce{Ca2+}:$n\times$POPC clusters
+In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
+the most probable binding mode of calcium at 285~mM concentration.
+
+Using our model, ECC-lipids, we fitted such a model to our simulations with a g
+In such a model, there is a linear relationship between quantities 
+$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt(C_b/C_I)$, 
+where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
+The concentration $C_b$ was obtained from an extrapolation of linear relation 
+between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
+Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
+remains constant throughout the extrapolation range. 
+The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
+However, Boltzmann theory yields inaccurate results
+for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
+An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
+Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
+our simulations likely lays in the fact that 
+the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
+gradually do not hold for higher concentrations of \ce{Ca2+}.
+
+Binding probability:
+\ce{Ca2+} complex with
+1 POPC - 13\%
+2 POPC - 18\%
+3 POPC - 12\%
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=6.0cm]{../Fig/number_bound_lipids_p_cation_Ecc-lipids.eps} \\
+  \caption{\label{fig:cacl_complexes}
+      Histogram of relative probabilities of existence of complexes
+      of \ce{Ca2+} and $n$ POPC lipids. 
+  }
+\end{figure}
+
 The good agreement of ECC-lipids model with experiments encourages us to analyse the
@@
 
+It was found in the original work~\cite{altenbach84} that 
+a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
+provides the best fit to experimental measurements of all considered models in that study. 
+In such a model, there is a linear relationship between quantities 
+$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt{C_b/C_I}$, 
+where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
+The concentration $C_b$ was obtained from an extrapolation of linear relation 
+between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
+Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
+remains constant throughout the extrapolation range. 
+The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
+However, Boltzmann theory yields inaccurate results
+for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
+An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
+Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
+our simulations likely lays in the fact that 
+the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
+gradually do not hold for higher concentrations of \ce{Ca2+}.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
+  \caption{\label{fig:cacl-bind}
+    Ternary complex binding model of \ce{Ca2+} to a POPC membrane 
+    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
+    provides a good fit to experimental measurements~\cite{altenbach84}
+    and it also provides a good fit to our simulation data. 
+    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
+    and, hence, the observed slope of the linear relationship is different.
+    }
+\end{figure}
+
 % Create the reference section using BibTe

### 2017-09-08 Manuscript/manuscript.tex
Small update to the section Stoich. binding in MS.

@@
 
+%Binding probability:
+%\ce{Ca2+} complex with \\
+%1 POPC - 13\% -- 30\% \\
+%2 POPC - 18\% -- 42\% \\
+%3 POPC - 12\% -- 28\% \\
+
 Binding stoichiometry of \ce{Ca2+} and POPC was 
@@
 the most probable binding mode of calcium at 285~mM concentration. 
-
-Using our model, ECC-lipids, we fitted such a model to our simulations with a g
-In such a model, there is a linear relationship between quantities 
-$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt(C_b/C_I)$, 
-where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
-The concentration $C_b$ was obtained from an extrapolation of linear relation 
-between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
-Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
-remains constant throughout the extrapolation range. 
-The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
-However, Boltzmann theory yields inaccurate results
-for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
-An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
-Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
-our simulations likely lays in the fact that 
-the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
-gradually do not hold for higher concentrations of \ce{Ca2+}.
-
-Binding probability:
-\ce{Ca2+} complex with
-1 POPC - 13\%
-2 POPC - 18\%
-3 POPC - 12\%
+Apart from this complex, we also find complexes with 1~and 3~lipids
+occuring with only a slightly lower but similar probability.
+The fractions of \ce{Ca2+}:$n\times$POPC complexes at 285~mM concetration 
+are then in order:
+$42\%$ for two lipids,
+$30\%$ for one lipid, and
+$28\%$ for three lipids.
+The symmetry of such a distribution --  
+i.e. almost equal probabilities of complexes with 1 or 3 lipids 
+that behave in the total average picture as complexes with 2 lipids -- 
+provides clues why ternary complex binding model 
+fits both simulation and experimental results relatively well, 
+although it is apparently incorrect.
 
@@
 \end{figure}
-
-The good agreement of ECC-lipids model with experiments encourages us to analyse the
-molecular binding details from MD simulations. Direct analysis of contacts between ions and
-lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
-has stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
-As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
-with the ternary complex model suggested based on head group order parameter
-experiments~\cite{altenbach84}.
-In addition to the ternary complexes, there also is a non-negligable probability
-of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
-Technical details of the analysis are in the SI.
 

### 2017-09-08 Manuscript/manuscript.tex
Finished writing observations and results to Stoichiometry section -- still has rugged style but the hard data shall be all there.

@@
 
-Binding stoichiometry of \ce{Ca2+} and POPC was 
-thoroughly studied in the experimental work~\citenum{altenbach84}.
+Binding stoichiometry of \ce{Ca^{2+}} and POPC was 
+thoroughly studied in the experimental work~\cite{altenbach84}.
 Several binding models were proposed and tested of which only one,
 ternary complex binding model, provided a good fit of the experimental observations. 
-In such a model, it is assumed that \ce{Ca2+} cations bind to a POPC membrane 
+In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
 with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
-Such a model also provides a good fit 
+In a later work~\cite{macdonald87}, 
+a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
+was found to provide as good fit as ternary complex model, 
+when only low concentrations of \ce{CaCl2} are considered.
+
+Ternary complex model also provides a good fit 
 to our simulations with ECC-lipids 
@@
 Simulations, however, allow us to directly evaluate the stoichiometry 
-by calculating relative propensities of various \ce{Ca2+}:$n\times$POPC clusters
+by calculating relative propensities of various \ce{Ca^{2+}}:$n\times$POPC clusters
 In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
@@
 occuring with only a slightly lower but similar probability.
-The fractions of \ce{Ca2+}:$n\times$POPC complexes at 285~mM concetration 
+The fractions of \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration 
 are then in order:
@@
       Histogram of relative probabilities of existence of complexes
-      of \ce{Ca2+} and $n$ POPC lipids. 
+      of \ce{Ca^{2+}} and $n$ POPC lipids. 
   }
@@
 
-In addition, we estimated the relative binding affinities of several moieties in POPC:
-phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
-Using the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
+In addition, we estimated relative binding affinities 
+of several moieties in POPC towards \ce{Ca^{2+}}. 
+Based on the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
 we estimate that the largest contribution
-to the binding of \ce{Ca2+} to POPC membranes likely comes from the phosphate group. 
+to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
 Although the isodensity plots are relatively easy to interpret,
 the contours shown in Fig.~\ref{fig:volmaps} cannot conclusively tell on the details
-of \ce{Ca2+} binding to any of the two carbonyl moieties.
+of \ce{Ca^{2+}} binding to any of the two carbonyl moieties
+but their apparently lower affinity compared to phosphate oxygens.
 
-\todo{JOE: Put details of the cation-binding stoichiometry analysis to SI.} \\
-\todo{JOE: Update the binding isotherm figure with new simulations} \\
+The residence time of \ce{Ca^{2+}} bound to a POPC membrane 
+is experimentally determined to be lower than $10\,\mu\mathrm{s}$~\cite{carrington67}. 
+Recent theoretical work with long enough simulations estimates this time
+in the order of $1\,\mu\mathrm{s}$~\cite{catte16}. 
+This is in contrast to our model, ECC-lipids, 
+which gives a mean residence time in the order of $10\,\mathrm{ns}$, 
+i.e. at least two orders of magnitude lower than previous estimates.
+Such a finding changes the point of view of calcium binding from
+very tight long-term stable binding with rare exchanges to 
+a relatively frequent exchange of cations 
+in equilibrium between membrane and solvent.
+
 \todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
@@
   SAMULI: This is written in \cite{macdonald87}:
-  ''Ca2+binding to POPC bilayers over the whole concentration range can be best described in terms
+  ''\ce{Ca^{2+}} binding to POPC bilayers over the whole concentration range can be best described in terms
   of formation of a ternary complex involving complexation of two lipids to one calcium ion (Altenbach and Seelig, 1984).
-  The addition of a sodium competition term has not changed this conclusion. However, if Ca2+ concentrations
+  The addition of a sodium competition term has not changed this conclusion. However, if \ce{Ca^{2+}} concentrations
   up to 100 mM are considered, the data can be equally well explained by a 1:1 binding mechanism (cf. Figure 7).
   In contrast, the Ca2+ binding to POPC-POPG mixtures can be best described by assuming a 1:l stoichiometry regardless of the range of Ca2+concentrations.''
-  We might or might not want to discuss about this.} \\
-\todo{SAMULI: I would also analyze how much there is contact between ions and different
-parts of the lipid (phosphase, carbonyl, etc.).
-JOE: Pavel suggest using volumetric map (e.g. in VMD) to visualise the amounts of Ca, Na and Cl bound to different moieties.} \\
-\todo{SAMULI: I think we should quantify this, i.e. how much there are these.
-  Maybe also the other possible complexes? Maybe also the correlation betweem complexes
-  and binding sites, if it is not too much work.
-  JOE: This looks like a careful work for the next paper to me. 
-  I'd only add a relatively simple analysis of binding sites and probably the propensity of 1-2-3 membered clusters.
-  SAMULI: I agree that we should not use too much time on this now.}. \\
+  We might or might not want to discuss about this.
+  JOE: I think this is adressed/acknowledged enough now.} \\
 \todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},
@@
   \caption{\label{fig:volmaps}
-      Contours of probability isodensities of \ce{Ca2+} with respect to 
+      Contours of probability isodensities of \ce{Ca^{2+}} with respect to 
       various moieties fixed in space (highlighted with transparent spheres): 
@@
       Shown contours suggest that the dominant contribution 
-      to \ce{Ca2+} binding likely comes from the phosphate group, 
-      but remain inconclusive on the details of the interactions with any of the two carbonyl groups. 
+      to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
+      whereas the interactions with any of the two carbonyl groups are considerably milder.
   }
-  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca2+}}
+  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca^{2+}}}
 \end{figure}
@@
 %  \caption{\label{fig:binding_conf}
-%      Configurations of \ce{Ca2+} binding in POPC membranes in the ECC-lipids model.
+%      Configurations of \ce{Ca^{2+}} binding in POPC membranes in the ECC-lipids model.
 %      The binding sites are: 
@@
 In such a model, there is a linear relationship between quantities 
-$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt{C_b/C_I}$, 
+$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
 where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
@@
 between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
-Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
+Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
 remains constant throughout the extrapolation range. 
@@
 However, Boltzmann theory yields inaccurate results
-for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
+for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
 An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
@@
 the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
-gradually do not hold for higher concentrations of \ce{Ca2+}.
+gradually do not hold for higher concentrations of \ce{Ca^{2+}}.
 
@@
   \caption{\label{fig:cacl-bind}
-    Ternary complex binding model of \ce{Ca2+} to a POPC membrane 
+    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
     that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 

### 2017-09-08 Manuscript/manuscript.tex
Modifcations after Hectors' comments.

--- 2017-09-08 Manuscript/manuscript.tex
@@
 like in signal propagation in neurons and vesicle fusion.
-\todo{JOE: list more introductory examples, where it has interesting effects.}
-Zwitterionic phosphocholine (PC) lipid bilayers have been widely used as model systems
-to understand molecular level details of specific cation interactions with
-cellular membranes by using experimental \cite{cevc90,tocanne90,binder02,pabst07,uhrikova08}
-and theoretical methods \cite{bockmann03,bockmann04,Berkowitz12,??} 
-\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
-While relative binding affinity of different ions is agreed to follow Hoffmeister
-series, the molecular details of binding and binding energetics are
-not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
+\todo{This could be maybe extended.}
+Since the direct measurements of ion-membrane interactions from biological systems
+are difficult, lipid bilayers are often used as models systems for cellular
+membranes. The detailed results from simple model systems can be then
+used to understand the role of ions in complex biological systems.
+
+Zwitterionic phosphocholine (PC) lipid bilayers are commonly used model systems for
+cellular membranes. Interactions of biological cations, especially Na$^+$ and
+Ca$^{2+}$, with PC bilayers are widely studied with
+experiments~\cite{akutsu81,altenbach84,seelig90,cevc90,tocanne90,binder02,pabst07,uhrikova08}
+and classical MD simulations~\cite{bockmann03,bockmann04,Berkowitz12,melcrova16,javanainen17}.
+The details of ion binding are, however, not agreed in the literature.
+%not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
 %General conclusion from experimental
@@
 %ions, but do not affect the specific binding constant \cite{seelig90}.
-Non-invansive spectroscopic methods, like nucelar magnetic resosnance (NMR), scattering
+Non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
 and infrared scpectroscopy, give accurate information about ion binding
-in lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02}. 
-Molecular level interpretation of the results, however, requires assumptions
-about binding models and is often not fully conclusive \cite{??}.
-On the other hand, the accuracy of the state of the art atomistic resolution lipid and ion models
-have turned out insufficient for the detailed interpretation of cation binding details \cite{catte16}.
-The overestimated specific cation binding observed in current lipid models \cite{catte16}
-may lead to articifially positively charged memranes and significant artefacts in
-MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
-\todo{JOE: this may be a too daring statement. \\
-  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
-  but it might happen with too small concentrations. 
-  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).
+in PC lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
+Interpretation of these experiments suggests that Na$^+$ ions have
+negligible binding in PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
+binds to phosphate groups of two lipid molecules. Atomistic resolution
+molecular dynamics (MD) simulation models, however, predict significantly
+stronger binding for the cations and interactions with 3-4 lipids,
+including also interactions with carbonyl
+oxygens~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
+Some experiments have also been interpreted to support the predictions from
+MD simulations~\cite{bockmann03,vacha09a}.
+
+Recent work published by the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
+made an attempt to resolve the apparent controversies. The study presented 
+a direct comparison between ion binding affinity between simulations and
+experiments by using NMR data for lipid headgroup order 
+parameters and the electrometer concept~\cite{seelig87}.
+Using massive data collected by Open Collaboration method, it was concluded
+that the accuracy of the current state of the art lipid models for MD
+simulations is not sufficient for the detailed interpretation of the cation
+interactions with PC lipid bilayers~\cite{catte16}.
+
+%\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
+
+%While relative binding affinity of different ions in PC lipids is agreed to follow Hoffmeister
+%series, the molecular details of binding and binding energetics are
 
@@
 
-In this work we show that the accuracy of ion binding, especially calcium,
-in zwitterionic PC lipid bilayer can be significantly improved 
-for classical MD simulation models of zwitterionic lipids
-by including electronic polarizability
-in the polar region of lipid molecules
-using the electronic continuum correction (ECC)~\cite{leontyev11}. 
-This approach has been previously shown to improve performace of
-ion models against neutron scattering data in concentrated solutions 
-and ab-initio simulations~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
-The realistic structure of ions in a bulk solvent was not, however, sufficient to correct binding in lipid
-bilayers~\cite{catte16}. 
-To fix it, we propose
-a classical fixed-charge MD simulation model of 
-1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) lipid 
-that accounts for electronic polarizability by using ECC. 
-We will validate such a model by comparing its lipid bilayer structue
-to x-ray scattering data and its head group structure to NMR experiments. 
-We will then quantify the accuracy of the head group order parameter response to the bound charge
-and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
-In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
-This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
-The proposed POPC model, ECC-lipids, and the approach to obtain it
-is, hence, highly useful for future MD simulations with physiological salt conditions, 
-and for effects where membrane electrostatics play a major role.
+In this work we show that the cation binding behavior
+in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
+improved by implicitly including electronic polarizability
+in the polar region of lipid molecules. The electronic
+polarizability is included by using the electronic
+continuum correction (ECC)~\cite{leontyev11}, which
+has been previously shown to improve the behaviour of
+MD simulations of ions in bulk water~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
+As a starting point we use the parameters from Lipid14 model \cite{dickson14},
+which gave the best cation binding behaviour in the previous study \cite{catte16}.
+The developed ECC-lipid parameters reproduce the experimentally
+measurable structural parameters of POPC lipid bilayer with the 
+accuracy comparable to the other state of the art lipid models.
+
+%We will validate such a model by comparing its lipid bilayer structure
+%to x-ray scattering data and its head group structure to NMR experiments. 
+%We will then quantify the accuracy of the head group order parameter response to the bound charge
+%and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
+%In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
+%This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
+
+Since the cation binding affinity and the headgroup 
+order parameter changes are in good agreement with experiments
+in the proposed ECC-lipids, it can be used to interpret the
+related structural changes and ion binding stoichiometry.
+New lipid models with correct ion binding affinity in lipid
+bilayers are necessary in applications of 
+MD simulations with physiological salt conditions. 
+The overestimated cation binding in the current lipid models \cite{catte16}
+may lead to significant artefacts in MD simulations. For example,
+articifially positively charged membranes overestimate interactions
+with biomolecules having opposite sign.
+
+%MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
+
+%\todo{JOE: this may be a too daring statement. \\
+%  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
+%  but it might happen with too small concentrations. 
+%  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).
 
@@
 Ion binding was compared between experiments and simulations by 
-using lipid head group order parameters and 
+using lipid head group order parameters and the
 ''electrometer concept'' \cite{seelig87,catte16}.
 The concept is based on the experimental observation that the C-H bond
-order parameters of $\alpha$ and $\beta$ carbons in lipid head group
+order parameters of $\alpha$ and $\beta$ carbons in PC lipid head group
 (see Fig.~\ref{simVSexpNOions}) are proportional to the amount 
-of bound charge in a lipid bilayer \cite{seelig87}.
-This is rationalized as a change of lipid head group dipole tilt to more vertical orientation
-with bound positive charge and {\it vice versa} for negative charge \cite{seelig87}. 
-More recent analysis in the NMRlipids project concluded 
+of unit charge bound per lipid,  $X^\pm$~\cite{seelig87}.
+Change in the order parameters measured with varying aqueous
+ion concentration can be then related to the amount of bound ions.
+
+The change of the head group order parameters is empirically quantified
+as~\cite{seelig87,ferreira16}
+\begin{equation}\label{OPchangeEQ}
+  \Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx \frac{4m_i }{3\chi}X^\pm,
+\end{equation}
+where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
+$i$ refers to either $\alpha$ or $\beta$ carbon,
+$m_i$ is an empirical constant depending on the valency and position of the bound charge,
+and the experimental value \cite{seelig77,davis83}, $\chi \approx$\,167\,kHz, is used for the quadrupole coupling constant.
+Atomic absorption spectra and $^2$H NMR data gave
+$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for Ca$^{2+}$ binding to POPC
+bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
+The slopes are negative, because recent analysis concluded 
 that the order parameters decrease with bound positive charge 
 and increase with bound negative charge when the signs are taken 
-in account \cite{ollila16,catte16}. The concept was also shown
-to be suitable for the comparison of ion binding affinity in lipid 
-bilayers between simulations and experiments~\cite{catte16}.
+in account \cite{ollila16,catte16}. This is rationalized as a change
+of lipid head group dipole tilt toward water phase with bound positive
+charge and {\it vice versa} with negative charge \cite{seelig87}.
 
-The order parameters for $\alpha$ and $\beta$ segments in head group, as well as for
-all C-H bonds in lipid molecules,
-can be accurately measured by using $^2$H NMR or $^{13}$C NMR techniques.
-The order parameters can be calculated from MD simulations 
-by using the definition
+The concept can be used to compare the ion binding affinity in lipid 
+bilayers between MD simulations and  NMR experiments, because the order
+parameters can be accurately determined from both techniques~\cite{ollila16}.
+The order parameters for all C-H bonds in lipid molecules, including
+$\alpha$ and $\beta$ segments in head group, can be accurately measured
+by using $^2$H NMR or $^{13}$C NMR techniques. From MD simulations the
+order parameters can be calculated by using the definition
 \begin{equation}\label{OP}
@@
 where $\theta$ is the angle between the bond and membrane
-normal and average is taken over all sampled configurations \cite{ollila16}.
+normal and the average is taken over all sampled configurations \cite{ollila16}.
 
-According to the electrometer concept,
-the change of the head group order parameters as a function of
-the amount of bound charge per lipid $X^\pm$ 
-can be written as~\cite{ferreira16}
-\begin{equation}\label{OPchangeEQ}
-\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) =\frac{4m_i }{3\chi}X^\pm,
-\end{equation}
-where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
-$i$ refers to either $\alpha$ or $\beta$ carbon,
-%the empirically observed linear relation %for the change in order parameters %with respect to a bilayer without bound charges              
-%\begin{equation}\label{electrometer_eq}                                                                                                   
-%S_{\rm{CH}}^{i}(X^\pm)=S_{\rm{CH}}^{i}(0) + \frac{4m_i}{3\chi} X^\pm,                                                                     
-%\end{equation}                                                                                                                            
-$m_i$ is an empirical constant depending on the valency and position of bound charge,
-and the value of the quadrupole coupling constant is $\chi \approx$\,167\,kHz.
-Atomic absorption spectra and $^2$H NMR data gave
-$m_\alpha=-20.5$  and $m_\beta=-10.0$ for Ca$^{2+}$ binding to POPC
-bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-Before using electrometer concept to analyze cation binding, 
-we calibrate the response of head group order parameters to the bound
-charge in the simulation models. This is done by comparing 
-the slope $m_i$ in Eq. \ref{OPchangeEQ} between simulations and
-experiments of cationic surfactants in POPC bilayer \cite{scherer89}. 
-Since essentially all surfactants locate in bilayer, the 
-molar ratio directly gives the amount of bound charge per 
-lipid $X^\pm$ is these systems.
+The measured order parameter change depends on the response of 
+headgroup on bound charge, i.e. $m_i$ in Eq.~\ref{OPchangeEQ},
+and the ion binding affinity. Thus, the former property has to be
+well quantified before using using the electrometer concept to
+analyze binding affinities. This is done experimentally for wide range
+of systems~\cite{seelig87,beschiasvili91}. To calibrate 
+the head group order parameter response also in simulations,
+we use experimental data for dihexadecyldimethylammonium (C$_{12}$C$_{16}$$^+$N2C$_1$Br$^-$)
+in POPC bilayer \cite{scherer89}. Dihexadecyldimethylammonium
+is a cation surfactant having two acyl chains and bearing a unit
+charge in the hydrophilic end. Thus, it is expected to locate
+in bilayer similarly to the phospholipids and the molar ratio then 
+gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems \cite{scherer87}.
 
@@
 are found to match the experimental values nicely.
-The head group and glycerol backbone order parameters accuracy is 
-comparable to the state of art lipid models available in literature \cite{ollila16}.
-The agreement between the x-ray scattering form factors and the areas per molecule from simulations and experiments
+\todo{The discussion about acyl chain and headgroup order parameters needs to be
+  finished when the value for C2, the scale are fixed in the figure and
+  size of the points are fixed in the figure.}
+The headgroup order parameters are slightly larger in the ECC-lipid model,
+which is in line with larger P-N vector angle in Fig.~\ref{OrderParameterCHANGESsurf}.
+With the current data we cannot, however, conclude which is the more realistic
+conformation since in both models other headgroup order parameter agrees
+with experiments within error bars and other does not.
+Slight changes are also observed in glycerol backbone order parameters,
+bu the accuracy of both models is comparable to the state of art lipid models
+available in literature \cite{ollila16}. The agreement between the x-ray scattering
+form factors and the areas per molecule from simulations and experiments
 confirm that the membrane structural properties are well captured. 
@@
 Fig. \ref{OrderParameterCHANGESsurf}. 
-Approximately linear decrease of the order parameters is observed
-in simulations and experiments with increasing amount of cationic
-surfactants up to mole fractions of $\sim$50\%,
-as expected from Eq. \ref{OPchangeEQ}.
+The observed order parameter decrease is simulations and experiments
+can be approximated to be linear at least with the mole fractions
+below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
 The slope is, however, too steep in Lipid14 model indicating that 
@@
 in ECC-lipid model is in good agreement with experiments.
-Therefore we use the data in Fig.~\ref{OrderParameterCHANGESsurf}
-to suggest a linear relation between $\alpha$-carbon order parameter
-and average P-N vector angle changes $\Delta \Theta_{\rm P-N}=?? \Delta S_{\rm CH}^{\alpha}$.
-This can be used to estimate the change in P-N vector angle
-from experiments measuring changes in $\alpha$-carbon order parameter. 
-\todo{SAMULI: I think that we should establish the relation between 
-order parameter and P-N vector angle changes by using ECC-lipids model.
-This would be useful for people measuring aplha order parameter
-changes from experiments.}
-For example, $\Delta S_{\rm CH}^{\alpha}=0.018$ measured due to the
-addition of 66.8 $\mu$M concentration of Mellittin in bulk water \cite{kuchinka89}
-gives $\Delta \Theta_{\rm P-N} = 2.5^{\circ}$.
-\todo{I looked this roughly from the figure, should be calculated from 
-the actual relation if we decide to put this.}
+The model can be thus used to study changes of lipid P-N vector
+in varying conditions.
+
+%
+%Therefore we use the data in Fig.~\ref{OrderParameterCHANGESsurf}
+%to suggest a linear relation between $\alpha$-carbon order parameter
+%and average P-N vector angle changes $\Delta \Theta_{\rm P-N}=?? \Delta S_{\rm CH}^{\alpha}$.
+%This can be used to estimate the change in P-N vector angle
+%from experiments measuring changes in $\alpha$-carbon order parameter. 
+%\todo{SAMULI: I think that we should establish the relation between 
+%order parameter and P-N vector angle changes by using ECC-lipids model.
+%This would be useful for people measuring aplha order parameter
+%changes from experiments.}
+%
+%For example, $\Delta S_{\rm CH}^{\alpha}=0.018$ measured due to the
+%addition of 66.8 $\mu$M concentration of Mellittin in bulk water \cite{kuchinka89}
+%gives $\Delta \Theta_{\rm P-N} = 2.5^{\circ}$.
+%\todo{I looked this roughly from the figure, should be calculated from 
+%the actual relation if we decide to put this.}
 
@@
 Lipid14 with \AA{}qvist and $\Gamma_i^w = 0.3 \pm 0.03 nm^{-2}$ with Dang ions.
+
+\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
+  do this analysis by using the density profiles, but it is not necessary.}
 Rough estimates for the free energy difference between bound and unbound cation
@@
 
-Binding stoichiometry of \ce{Ca2+} and POPC was 
-thoroughly studied in the experimental work~\citenum{altenbach84}.
-Several binding models were proposed and tested of which only one,
-ternary complex binding model, provided a good fit of the experimental observations. 
-In such a model, it is assumed that \ce{Ca2+} cations bind to a POPC membrane 
-with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
-Such a model also provides a good fit 
-to our simulations with ECC-lipids 
-(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
-Simulations, however, allow us to directly evaluate the stoichiometry 
-by calculating relative propensities of various \ce{Ca2+}:$n\times$POPC clusters
-In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
-the most probable binding mode of calcium at 285~mM concentration.
-
-Using our model, ECC-lipids, we fitted such a model to our simulations with a g
-In such a model, there is a linear relationship between quantities 
-$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt(C_b/C_I)$, 
-where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
-The concentration $C_b$ was obtained from an extrapolation of linear relation 
-between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
-Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
-remains constant throughout the extrapolation range. 
-The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
-However, Boltzmann theory yields inaccurate results
-for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
-An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
-Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
-our simulations likely lays in the fact that 
-the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
-gradually do not hold for higher concentrations of \ce{Ca2+}.
-
-Binding probability:
-\ce{Ca2+} complex with
-1 POPC - 13\%
-2 POPC - 18\%
-3 POPC - 12\%
-
-\begin{figure}[tbp]
+\begin{figure}[]
   \centering
-  \includegraphics[width=6.0cm]{../Fig/number_bound_lipids_p_cation_Ecc-lipids.eps} \\
-  \caption{\label{fig:cacl_complexes}
-      Histogram of relative probabilities of existence of complexes
-      of \ce{Ca2+} and $n$ POPC lipids. 
+  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
+  \caption{\label{fig:cacl-bind}
+    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .
     }
@@
   }
-  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca2+}}
 \end{figure}
@@
 
-It was found in the original work~\cite{altenbach84} that 
-a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
-provides the best fit to experimental measurements of all considered models in that study. 
-In such a model, there is a linear relationship between quantities 
-$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt{C_b/C_I}$, 
-where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
-The concentration $C_b$ was obtained from an extrapolation of linear relation 
-between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
-Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
-remains constant throughout the extrapolation range. 
-The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
-However, Boltzmann theory yields inaccurate results
-for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
-An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
-Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
-our simulations likely lays in the fact that 
-the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
-gradually do not hold for higher concentrations of \ce{Ca2+}.
-
-\begin{figure}[]
-  \centering
-  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
-  \caption{\label{fig:cacl-bind}
-    Ternary complex binding model of \ce{Ca2+} to a POPC membrane 
-    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
-    provides a good fit to experimental measurements~\cite{altenbach84}
-    and it also provides a good fit to our simulation data. 
-    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
-    and, hence, the observed slope of the linear relationship is different.
-    }
-\end{figure}
-
 % Create the reference section using BibTe

### 2017-09-08 Manuscript/manuscript.tex
Update to the Stoichiometry of binding section -- mainly corrections and comments on what to do.

--- 2017-09-08 Manuscript/manuscript.tex
@@
 
+This section is rough and will likely require editing. 
+Data -- up to noted exceptions -- shall be all there, however.
+
 Binding stoichiometry of \ce{Ca^{2+}} and POPC was 
-thoroughly studied in the experimental work~\cite{altenbach84}.
+thoroughly studied in the experimental work~\cite{altenbach84},
+in which the head group order parameter changes to cation binding are determined.
 Several binding models were proposed and tested of which only one,
 ternary complex binding model, provided a good fit of the experimental observations.
-In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
-with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
-In a later work~\cite{macdonald87}, 
-a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
-was found to provide as good fit as ternary complex model, 
-when only low concentrations of \ce{CaCl2} are considered.
 
-Ternary complex model also provides a good fit 
-to our simulations with ECC-lipids 
-(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
-Simulations, however, allow us to directly evaluate the stoichiometry 
+Simulations allow us to directly evaluate the stoichiometry 
 by calculating relative propensities of various \ce{Ca^{2+}}:$n\times$POPC clusters
+by evaluating contacts between cations and lipids with a cut off radius $0.3\,\mathrm{nm}$. 
 In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
@@
 $28\%$ for three lipids.
-The symmetry of such a distribution --  
+
+Several binding models were proposed and tested~\cite{altenbach84} of which only one,
+ternary complex binding model, provided a good fit of the experimental observations. 
+In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
+with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
+In a later work~\cite{macdonald87}, 
+a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
+was found to provide as good fit as ternary complex model, 
+when only low concentrations of \ce{CaCl2} are considered.
+Ternary complex model also provides a good fit 
+to our simulations with ECC-lipids 
+(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
+The symmetry of the distribution of complexes from simulation --  
 i.e. almost equal probabilities of complexes with 1 or 3 lipids 
@@
   }
+  \todo{Change this figure so that it contains the relative probabilities of finding the particular clusters.}
 \end{figure}
@@
 of several moieties in POPC towards \ce{Ca^{2+}}. 
+\todo{Add a simple analysis using number of contacts.}
 Based on the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
@@
 The residence time of \ce{Ca^{2+}} bound to a POPC membrane 
-is experimentally determined to be lower than $10\,\mu\mathrm{s}$~\cite{carrington67}. 
-Recent theoretical work with long enough simulations estimates this time
-in the order of $1\,\mu\mathrm{s}$~\cite{catte16}. 
+is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
+From recent theoretical work with long enough simulations this time can be roughly estimated
+in the order of $1$--$10\,\mu\mathrm{s}$~\cite{Javanainen2017}. 
 This is in contrast to our model, ECC-lipids, 
 which gives a mean residence time in the order of $10\,\mathrm{ns}$, 
+\todo{evaluate this number, mean residence time, accurately based on the contacts data.}
 i.e. at least two orders of magnitude lower than previous estimates.

### 2017-09-11 Manuscript/manuscript.tex

--- 2017-09-08 Manuscript/manuscript.tex
@@
 
-\begin{figure}[]
+%Binding probability:
+%\ce{Ca2+} complex with \\
+%1 POPC - 13\% -- 30\% \\
+%2 POPC - 18\% -- 42\% \\
+%3 POPC - 12\% -- 28\% \\
+
+This section is rough and will likely require editing. 
+Data -- up to noted exceptions -- shall be all there, however.
+
+Binding stoichiometry of \ce{Ca^{2+}} and POPC was 
+thoroughly studied in the experimental work~\cite{altenbach84},
+in which the head group order parameter changes to cation binding are determined.
+Several binding models were proposed and tested of which only one,
+ternary complex binding model, provided a good fit of the experimental observations.
+
+Simulations allow us to directly evaluate the stoichiometry 
+by calculating relative propensities of various \ce{Ca^{2+}}:$n\times$POPC clusters
+by evaluating contacts between cations and lipids with a cut off radius $0.3\,\mathrm{nm}$. 
+In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
+the most probable binding mode of calcium at 285~mM concentration. 
+Apart from this complex, we also find complexes with 1~and 3~lipids
+occuring with only a slightly lower but similar probability.
+The fractions of \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration 
+are then in order:
+$42\%$ for two lipids,
+$30\%$ for one lipid, and
+$28\%$ for three lipids.
+
+Several binding models were proposed and tested~\cite{altenbach84} of which only one,
+ternary complex binding model, provided a good fit of the experimental observations. 
+In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
+with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
+In a later work~\cite{macdonald87}, 
+a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
+was found to provide as good fit as ternary complex model, 
+when only low concentrations of \ce{CaCl2} are considered.
+Ternary complex model also provides a good fit 
+to our simulations with ECC-lipids 
+(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
+The symmetry of the distribution of complexes from simulation --  
+i.e. almost equal probabilities of complexes with 1 or 3 lipids 
+that behave in the total average picture as complexes with 2 lipids -- 
+provides clues why ternary complex binding model 
+fits both simulation and experimental results relatively well, 
+although it is apparently incorrect.
+
+\begin{figure}[tbp]
   \centering
-  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
-  \caption{\label{fig:cacl-bind}
-    Binding isotherm assuming stoichiometry of 2~POPC:1~\ce{Ca^{2+}} as used in \cite{altenbach84} fits our simulation data well .
+  \includegraphics[width=6.0cm]{../Fig/number_bound_lipids_p_cation_Ecc-lipids.eps} \\
+  \caption{\label{fig:cacl_complexes}
+      Histogram of relative probabilities of existence of complexes
+      of \ce{Ca^{2+}} and $n$ POPC lipids. 
   }
+  \todo{Change this figure so that it contains the relative probabilities of finding the particular clusters.}
 \end{figure}
 
-The good agreement of ECC-lipids model with experiments encourages us to analyse the
-molecular binding details from MD simulations. Direct analysis of contacts between ions and
-lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
-has stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
-As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
-with the ternary complex model suggested based on head group order parameter
-experiments~\cite{altenbach84}.
-In addition to the ternary complexes, there also is a non-negligable probability
-of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
-Technical details of the analysis are in the SI.
-
-In addition, we estimated the relative binding affinities of several moieties in POPC:
-phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
-Using the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
+In addition, we estimated relative binding affinities 
+of several moieties in POPC towards \ce{Ca^{2+}}. 
+\todo{Add a simple analysis using number of contacts.}
+Based on the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
 we estimate that the largest contribution
-to the binding of \ce{Ca2+} to POPC membranes likely comes from the phosphate group. 
+to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
 Although the isodensity plots are relatively easy to interpret,
 the contours shown in Fig.~\ref{fig:volmaps} cannot conclusively tell on the details
-of \ce{Ca2+} binding to any of the two carbonyl moieties.
-
-\todo{JOE: Put details of the cation-binding stoichiometry analysis to SI.} \\
-\todo{JOE: Update the binding isotherm figure with new simulations} \\
+of \ce{Ca^{2+}} binding to any of the two carbonyl moieties
+but their apparently lower affinity compared to phosphate oxygens.
+
+The residence time of \ce{Ca^{2+}} bound to a POPC membrane 
+is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
+From recent theoretical work with long enough simulations this time can be roughly estimated
+in the order of $1$--$10\,\mu\mathrm{s}$~\cite{Javanainen2017}. 
+This is in contrast to our model, ECC-lipids, 
+which gives a mean residence time in the order of $10\,\mathrm{ns}$, 
+\todo{evaluate this number, mean residence time, accurately based on the contacts data.}
+i.e. at least two orders of magnitude lower than previous estimates.
+Such a finding changes the point of view of calcium binding from
+very tight long-term stable binding with rare exchanges to 
+a relatively frequent exchange of cations 
+in equilibrium between membrane and solvent.
+
 \todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
@@
       Shown contours suggest that the dominant contribution 
-      to \ce{Ca2+} binding likely comes from the phosphate group, 
-      but remain inconclusive on the details of the interactions with any of the two carbonyl groups. 
+      to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
+      whereas the interactions with any of the two carbonyl groups are considerably milder.
   }
+  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca^{2+}}}
 \end{figure}
@@
 
+It was found in the original work~\cite{altenbach84} that 
+a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
+provides the best fit to experimental measurements of all considered models in that study. 
+In such a model, there is a linear relationship between quantities 
+$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
+where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
+The concentration $C_b$ was obtained from an extrapolation of linear relation 
+between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
+Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
+remains constant throughout the extrapolation range. 
+The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
+However, Boltzmann theory yields inaccurate results
+for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
+An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
+Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
+our simulations likely lays in the fact that 
+the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
+gradually do not hold for higher concentrations of \ce{Ca^{2+}}.
+
+\begin{figure}[]
+  \centering
+  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
+  \caption{\label{fig:cacl-bind}
+    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
+    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
+    provides a good fit to experimental measurements~\cite{altenbach84}
+    and it also provides a good fit to our simulation data. 
+    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
+    and, hence, the observed slope of the linear relationship is different.
+    }
+\end{figure}
+
 % Create the reference section using BibTe

### 2017-09-11 Manuscript/manuscript.tex
A small commit fix

@@
 has been previously shown to improve the behaviour of
-MD simulations of ions in bulk water~\cite{Jungwirth2017,Jungwirth2015,kohagen14,kohagen16}. 
+MD simulations of ions in bulk water~\cite{jungwirth17-new-paper-to-be-published,Pluharova2014,kohagen14,kohagen16}. 
 As a starting point we use the parameters from Lipid14 model \cite{dickson14},
@@
 From recent theoretical work with long enough simulations this time can be roughly estimated
-in the order of $1$--$10\,\mu\mathrm{s}$~\cite{Javanainen2017}. 
+in the order of $1$--$10\,\mu\mathrm{s}$~\cite{javanainen17}. 
 This is in contrast to our model, ECC-lipids, 

### 2017-09-11 Manuscript/manuscript.tex
Updated caption in the validation figure (x-ray form f and order params) with chem structure of POPC.

@@
   \centering
-  \includegraphics[width=7.0cm]{../Fig/POPCstructure.eps}
+  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
   \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
@@
   \caption{\label{simVSexpNOions}
-    X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
-    Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
-    compared with experimental order parameters from \cite{ferreira13}.}
-  \todo{There seems to a typo in experimental value for other order parameter for C2 carbon in sn-2 chain; it should be negative.
-    The scale of the y-axis should be then zoomed when this is fixed.} \\
-  \todo{I would put the acyl chain order parameters also for original Lipid14; JOE: Already done, check whether your figures are up to date. I usually push only *.eps.}\\
-  \todo{x-axis scale in form factor figure would be better from 0 to 0.6. The experimental data is available for this range.} \\
-  \todo{$S_{CD} -> S_{CH}$}
+    Top: chemical structure of POPC with definitions of headgroup order parameters.
+    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
+    compared with experimental order parameters from \cite{ferreira13}.
+    Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
+    }
 \end{figure}

### 2017-09-11 Manuscript/manuscript.tex
Removed TODOs
Improved figure with the probability histogram of lipid-Na/Ca complexes.

@@
   }
-  \todo{Labels are missing.} \\
-  \todo{Order should be the same as in other figures, i.e, $\beta$ segment on top.} \\
-  \todo{I would put x-axis from 0 to 51 and maybe zoom y-axis little bit as well.}
 \end{figure}
@@
   \centering
-  \includegraphics[width=6.0cm]{../Fig/number_bound_lipids_p_cation_Ecc-lipids.eps} \\
+  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
   \caption{\label{fig:cacl_complexes}
-      Histogram of relative probabilities of existence of complexes
-      of \ce{Ca^{2+}} and $n$ POPC lipids. 
+      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
+      with a certain number of POPC lipids. 
+      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
+      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
   }
-  \todo{Change this figure so that it contains the relative probabilities of finding the particular clusters.}
 \end{figure}

### 2017-09-19 Manuscript/manuscript.tex
Changes Hector

--- 2017-09-11 Manuscript/manuscript.tex
@@
 \usepackage{mhchem}
+\usepackage{chemfig}
 
@@
 effectively representing the newly introduced electronic continuum. 
-The value measured for water, $\epsilon _{el} = 1.78$, gives 
-a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}, which has been
-successfully used to improve the performance of force field for ions \cite{kohagen14,kohagen16,??}.
-
-While the scaling factor of $f_q = 0.75$ for ions in water is well justified and 
-improves the model performance against experimental scattering data \cite{kohagen14,kohagen16, Pluharova2014},
-it is not clear if the same factor should be used for partial charges in molecules,
-i.e, lipids in our case. 
-Unlike the total charge of molecules, atomic partial charges within each molecule are not experimental observables.
-\todo{Text should be clarified from here to the end of the paragraph}
-Although there have been proposed several schemes for the assignment of partial charges of molecules, 
-methods targetting at fitting of electrostatic potentials are 
-most popular among classical fixed charge force fields for biomolecules. \cite{Hu2007}
-Currently the most commonly employed scheme is the RESP scheme~\cite{RESP_paper,Singh1984}, 
-which fits the electrostatic potential from quantum mechanical calculations. 
-The resulting set of fixed partial atomic charges is hence 
-the best fit to electrostatic potentials of 
-an ensemble of conformations of a molecule. 
-Such a physical meaning is especially well suited for the application of ECC, 
-which targets at improving electrostatic interactions. 
-Depending to the details, partial charge calculations may also partially
-include some of the solvent electronic polarizability effects. 
-Thus, we expect that
-the correct ECC scaling factor, $f_q$, for molecular partial charges
-does not have to be striclty $f_q = \epsilon _{el} ^{-1/2}$, 
-but rather lies between this value (no electronic polarizability in the charge calculation)
+Assuming globally a  high frequency dielectric constant as measured in water (square of the refraction index),  $\epsilon _{el} = 1.78$, results 
+in a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
+successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??} which then agree with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+
+While the scaling factor of $f_q = 0.75$ for ions in water improves their performance and is physically justifiable within the MDECC theory \cite{Leontyev2011},
+it is not clear how the same factor should be used for partial charges in molecules,
+e.g., lipids in our case. Unlike the total charge of molecules, atomic partial charges within each molecule are not physical observables.
+There are several schemes for the assignment of partial charges for biomolecules. \cite{Hu2007}
+Currently, the most commonly employed scheme is RESP~\cite{RESP_paper,Singh1984}, 
+which aims to reproduce the electrostatic potential obtained from quantum mechanical calculations around an ensemble the target molecules. 
+The fixed partial atomic charges obtained by this method is hence 
+the best fit of these partial charges that reproduces the electrostatic potential around the provided target conformations. 
+Such averaging is well suited for the application of ECC, 
+which targets at improving electrostatic interactions using a mean field correction. 
+In practice, partial charges currently implemented in force fields may already 
+include  to some extend some of the solvent electronic polarizability effects, i.e.,
+the RESP charges are scaled to fit some experimental observable. 
+Thus, we expect that the application of the ECC scaling factor, $f_q$, on the molecular partial charges
+included in the available force fields do not necessarily require the integral, $f_q = \epsilon _{el} ^{-1/2}$, 
+but instead it lies between 0.75 (no electronic polarizability in the charge calculation)
 and $1$ (full electronic polarizability already included in the charge calculation).
@@
 
-In this work, we empirically sample this parameter space $f_q \in [0.75, 1.0]$
-to find an atomistic MD simulation force field parameters for lipids that
-accurately describe the lipid head group response to varying concentrations of cations
-when compared to NMR data \cite{akutsu81,altenbach84,scherer89}. Such data can be used to accurately asses
-ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
-We chose Lipid14~\cite{dickson14} force field parameters as a starting point,
-because their response to bound ions was apparently most realistic against NMR data
+In this work, we produce a phospholipid model for classical MD simulations that
+accurately describes the lipid head group response to varying concentrations
+of monovalent and divalent cations. This is critical membrane feature, which is
+poorly reproduced by currently available models, can affect not only membrane properties
+in the presence of ions but also modulates the interaction with charged moieties in the surface.
+Importantly, this response from simulations can be accurately
+compared against experimental NMR data \cite{akutsu81,altenbach84,scherer89},
+as discussed in Ref. \citenum{catte16} in section \ref{section:electrometer}.
+To this end, we empirically explore the scaling factor parameter space, $f_q \in [0.75, 1.0]$ on the Lipid14~\cite{dickson14} force field.
+We selected this force field as a starting point because their response to bound ions was apparently the most realistic against NMR data
 in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
@@
 glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
-are expected to have the largest contribution to the cation binding.
-The hydrocarbon chain parameters are not modified, because they are
+are expected to have the largest contribution to the cation binding \todo{Do we really need to scale charges in non charged methylens in the choline and glycerol backbones?
+why?
+Does it really make a difference?}.
+However, we do not modify the hydrocarbon chain parameters, because they are
 already highly optimized and give generally a good description for
-hydrophobic part of lipid bilayers in various conditions
-in most lipids models, including Lipid14 \cite{ollila16}.
-In contrast, improvements in glycerol backbone and head group
-parameters are required in all available lipid models~\cite{botan15}.
-
-Exploring different scaling factor values, applied to head group, glycerol backbone
-and carbonyl charges of Lipid14 model, we found out that ion binding and
+hydrophobic part of lipid bilayers in most lipids models, including Lipid14 \cite{ollila16}.
+This contrast with the behavior in glycerol backbone and head group
+parameters which call for improvements in all available lipid models~\cite{botan15}.
+
+Exploring different scaling factor values, we found out that ion binding and
 related head group order parameter responses become weaker. 
@@
 similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
+\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules}
 This decreases the solvation free energy by allowing water molecules to
@@
 
-\subsection{Ion concentrations and binding affinity}
-The salt concentrations are reported in two different ways
-in the used experimental data for the lipid head group order
-parameters. The data for DPPC is reported
-by using solution concentrations before solvating
-the lipids~\cite{akutsu81}, while the data
-for POPC~\cite{altenbach84} is reported by using
-supernatant concetration measured with atomic absorption
-spectroscopy after solvation of lipids.
-The former was used in comparison with experiments and
-calculated from simulations in previous work as 
-[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M~\cite{catte16}.
-Here we use the latter, because the focus is on POPC bilayers.
-This is determined from the ion concentrations, which reach
-a constant value at the edges of large enough simulation box.
-The two concentrations differ somewhat for CaCl$_2$ systems, because some
-amount of ions bind in lipids. However, the small difference do
-not affect the conclusions in this or in previous work~\cite{catte16}.
-
-The ion concentrations with constant value at the simulation box edges
-are also used to quantify the ion binding affinities is different simulation
-models by using the relative surface excess of ions with respect to water $\Gamma_i^w$.
-This is defined as \cite{chattorajBOOK}
-\begin{eqnarray}
-  \Gamma_i^w   &=& \Gamma_i-\Gamma_w \frac{C_i-C_i'}{C_w-C_w'} \\
-  \Gamma_{a} &=& \frac{n_{a} - C_{a} V - C_{a}' V'}{A} \, \, ; \, a=i,w \, ,
-\end{eqnarray}
-where index $i$ and $w$ denote ions and water respectively,
-prime accent denotes the phase inside bilayer, 
-whereas non-prime symbols account for bulk water phase,
-$n$ is the total amount of water or ions,
-$C$ denotes bulk concentrations, 
-$V$ is the volume of the respective phase separated by the Gibbs divinding plane, 
-$A$ is the area of the interface, 
-and $\Gamma_a$ is the surface excess of water or ions. 
-As $\Gamma_i^w$ does not depend on the definition of the interfacial region
-and the position of the Gibbs dividing plane~\cite{chattorajBOOK},
-we consider the whole simulation box as an interface. 
-Because bilayers have an interface
-in both leaflefts,
-the total area of the interface in the simulation
- is given by the simulation box area multiplied by two, $2A_b$.
-The concentrations of water and ions inside
-the bilayer can be expected to vanish, thus $c_{i,w}'=0$.
-This provides us a simplified relation for 
-the relative surface excess of lipid bilayers in simulations
+\subsection{Salt concentrations and binding affinity}
+Experimental works providing the lipid head group order
+parameters report the salt concentrations in two different ways.
+For DPPC, salt concentrations before solvating
+the lipids are used~\cite{akutsu81}. Instead, for POPC,
+the salt concentration of supernatant after solvation is
+measured using atomic absorption~\cite{altenbach84}.
+In this work, we will use the latter definition. For this we
+will measure the salt concentration in the water bulk region,
+\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
+Note that in our previous study, Ref.~\citenum{catte16},
+we calculated the concentration as in the DPPC experiments.
+Despite the two concentrations may differ significantly for CaCl$_2$ systems
+these differences do not affect qualitatively the conclusions neither in this
+nor in our previous work~\cite{catte16}.
+
+To quantify the ion binding affinity to a membrane, we
+calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
+This quantity does not depend on the definition of the interfacial region
+and the position of the Gibbs dividing plane. Therefore, we consider our whole simulation box as
+an interface flanked by a water phase and an ideal phase with the characteristics of to
+the hydrophobic region in the middle of the membrane where no ions or waters are present.
+This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
 \begin{equation}\label{surfexcess}
@@
 \end{equation}
-where the bulk concentrations, $C_w$ and $C_i$, are 
-concentrations with constant value at the simulation box edges.
+where $n_w$ and $n_i$ are the total number of waters and ions in the system;
+$C_w$ and $C_i$ are their respective bulk concentrations in the water phase;
+and $A_b$ is the size of the box in the membrane plane. Therefore, this quantity
+does not depend on any arbitrary definition of a bound ion to a membrane and
+therefore allow us to accurately compare between different models. Note that
+because bilayers have an interface in both leaflefts, the total area of the
+interface is twice the area of the membrane, $2A_b$.
 
@@
 
-Lipid bilayer structure without ions was validated against NMR experiments
-by order parameters for C-H bonds and \mbox{x-ray} scattering experiments by using
-form factors. The former validates the structures sampled by individual
-lipid molecules in simulations with segmental resolution, while the latter
+Lipid bilayer structure without ions was validated against NMR and 
+\mbox{x-ray} scattering experiments by calculating order parameters
+for C-H bonds and form factors from our simulations. The former validates the structures sampled by individual
+lipid molecules in simulations with atomic resolution, while the latter
 validates the dimensions of the lipid bilayer (thickness and area per
@@
 where $f_\alpha(q_z)$ is the density of atomic scattering length, 
-$\rho_s$ is the density of solvent scattering length,
+$\rho_s$ is the density of solvent scattering length in the bulk region,
 $n_\alpha (z)$ is the number density of atom $\alpha$ and
 $z$ is the distance from the membrane centre along its normal 
-spanning the membrane with thickness $D$.
+spanning until the water bulk region, $D$.
 
@@
 
-\subsubsection{Simulations with aqueous ions}
+\subsubsection{Simulations of POPC bilayers in aqueous ions}
 The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
-(POPC) bilayer and an aqueous salt solutions of varying concentrations.
-In particular, the periodic orthorhombic simulation box conatined 
+(POPC) bilayer and in aqueous solutions of varying salt concentrations.
+In particular, the periodic orthorhombic simulation box contained 
 128 POPC molecules and approximately 50~water molecules per each lipid.
-Water molecules were described by OPC3 model~\cite{Izadi16} 
-as it is currently the most accurate three site rigid water model. 
-In order to test transferability of the newly developed ECC-lipids model, 
+Our default, water molecules were described by OPC3 model~\cite{Izadi16} 
+which is currently the most accurate three site rigid water model. 
+In order to test transferability of our newly developed ECC-lipids model, 
 we also performed several additional simulations with water models 
@@
 
-\subsubsection{Simulations with cationic surfactants}
+\subsubsection{Simulations of POPC bilayers with cationic surfactants}
 Automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
+dihexadecyldimethylammonium bromide cationic surfactant,
+
+\vspace{0.5cm}
+\chemfig{
+\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{12}-CH_3)  (-[:0,3.5,,,draw=none]\chemabove{Br}{\scriptstyle\ominus}) 
+} .
+\vspace{0.5cm}
+
 AmberTools program \cite{amber} was then used to generate the Amber-type force field
@@
 6, 14, 21, 35 or 50 surfactants per leaflet.
-The systems were simulated for 200~ns using Lipid14 model for POPC.
+The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
 First 20~ns were omitted from the analysis.
@@
 
-\subsection{POPC membrane and its structure}
+\subsection{POPC membrane structure and dynamics}
 
@@
   \caption{\label{simVSexpNOions}
-    Top: chemical structure of POPC with definitions of headgroup order parameters.
+    Top: Chemical structure of POPC with definitions of different order parameters calculated.
     Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
-    compared with experimental order parameters from \cite{ferreira13}.
+    compared with experimental order parameters from \cite{ferreira13} {\color {red} X-label misses glycerol region}.
     Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
@@
 NMR order parameters measurements and x-ray scattering form factors 
-(Fig.~\ref{simVSexpNOions} and Table~\ref{tab:apls}). 
-The tail order parameters being already highly optimized in the original Lipid14 model~\cite{dickson14}
-are found to match the experimental values nicely.
+(Fig.~\ref{simVSexpNOions}). 
+Our tail order parameters, as in the original Lipid14 model~\cite{dickson14},
+match the experimental values. This is expected as the ECC model does not modify
+the already highly optimized parameters of this region in Lipid14. 
 \todo{The discussion about acyl chain and headgroup order parameters needs to be
@@
   size of the points are fixed in the figure.}
-The headgroup order parameters are slightly larger in the ECC-lipid model,
+The headgroup order parameters computed form our new are slightly larger in the ECC-lipid model in comparison ,
 which is in line with larger P-N vector angle in Fig.~\ref{OrderParameterCHANGESsurf}.
@@
 available in literature \cite{ollila16}. The agreement between the x-ray scattering
-form factors and the areas per molecule from simulations and experiments
+form factors (Fig.~\ref{simVSexpNOions}) and the areas per molecule from simulations and experiments (Table~\ref{tab:apls}) 
 confirm that the membrane structural properties are well captured. 
@@
 %by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
+
+\todo{Dynamics check is missing: MSD (Hector/Joe)}
 
@@
 The headgroup order parameter changes with increasing amount of 
-cationic surfactant dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) 
+cationic surfactant dihexadecyldimethylammonium bromide 
 is compared between experiments~\cite{scherer89} and simulations in
@@
 in Fig.~\ref{OrderParameterCHANGESsurf}.
-The headgroup orients more parallel to the membrane
-normal with increasing amount of bound cations, as expected~\cite{seelig87}. 
+The headgroup orients more towards the water phase with increasing
+amount of bound cations, as previously reported in Ref.~\citenum{seelig87}. 
 The effect is more pronounced in Lipid14 than 
@@
 bilayers can be measured by using the headgroup 
-order parameters, because they decrease linearly 
+order parameters, because they decrease monotonously  
 with the bound positive charge~\cite{seelig87,catte16}. 
@@
 that improvements also in lipid models are needed to 
-correctly describe cation binding in PC bilayers.
+correctly describe divalent cation binding in PC bilayers.
 
@@
 The binding affinities are quantified by using
-the ion density profiles along membrane normal shown
+the water density profiles along membrane normal shown
 in Fig.~\ref{fig:cacl-dens}. The density profiles show

### 2017-09-19 Manuscript/manuscript.tex

@@
 \textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
-Note that in our previous study, Ref.~\citenum{catte16},
-we calculated the concentration as in the DPPC experiments.
-Despite the two concentrations may differ significantly for CaCl$_2$ systems
-these differences do not affect qualitatively the conclusions neither in this
+Note that in the previous study, Ref.~\citenum{catte16},
+the concentration was calculated as in the DPPC experiments.
+Despite of the measurable differences between two concentrations for CaCl$_2$ systems,
+the qualitatively conclusions are not affected neither in this
 nor in our previous work~\cite{catte16}.
@@
 calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
-This quantity does not depend on the definition of the interfacial region
-and the position of the Gibbs dividing plane. Therefore, we consider our whole simulation box as
-an interface flanked by a water phase and an ideal phase with the characteristics of to
-the hydrophobic region in the middle of the membrane where no ions or waters are present.
+This quantity can be used to compare binding affinity between different models,
+because it does not depend on the definition of the interfacial region
+or the position of the Gibbs dividing plane between two bulk regions.
+Therefore, we consider the hydrophobic interior of a membrane having zero concentrations of ions and water
+as a bulk region, which is separated by a monolayer interface from 
+a another region with bulk concentration of ions in water.
+Since the boundary of the latter region is set to the edge of simulation box,
+all molecules in simulation are within the defined interfacial region. 
+%our whole simulation box as
+%an interface flanked by a water phase and an ideal phase with the characteristics of to
+%the hydrophobic region in the middle of the membrane where no ions or waters are present.
 This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
@@
 $C_w$ and $C_i$ are their respective bulk concentrations in the water phase;
-and $A_b$ is the size of the box in the membrane plane. Therefore, this quantity
-does not depend on any arbitrary definition of a bound ion to a membrane and
-therefore allow us to accurately compare between different models. Note that
-because bilayers have an interface in both leaflefts, the total area of the
-interface is twice the area of the membrane, $2A_b$.
+and $A_b$ is the size of the box in the membrane plane.
+The total area of the interface is twice the area of the membrane, $2A_b$,
+because bilayers have an interface in both leaflefts.
 
@@
 \chemfig{
-\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{12}-CH_3)  (-[:0,3.5,,,draw=none]\chemabove{Br}{\scriptstyle\ominus}) 
+(-[:0,1.0,,,draw=none]\chemabove{Br}{\scriptstyle\ominus}) -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{12}-CH_3)  
 } .
@@
 All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
-6, 14, 21, 35 or 50 surfactants per leaflet.
+6, 14, 21, 35 or 50 surfactants per leaflet. Parameters for Bromium ions were
+not available in the standard Gromacs files for Amber force field, thus the
+parameters for Chloride ions were used instead.
 The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
@@
 bilayers can be measured by using the headgroup 
-order parameters, because they decrease monotonously  
-with the bound positive charge~\cite{seelig87,catte16}. 
+order parameters, because they decrease proportionally  
+to the bound positive charge~\cite{seelig87,catte16}. 
 The headgroup order parameter responses 

### 2017-09-20 Manuscript/manuscript.tex
Joe - retouche petit up to results (not-included)

@@
 \todo{This could be maybe extended.}
-Since the direct measurements of ion-membrane interactions from biological systems
+\todo{JOE: following doesn't sound right to me. The reasoning is weird.}
+Since direct measurements of ion-membrane interactions from biological systems
 are difficult, lipid bilayers are often used as models systems for cellular
@@
 (ECC)~\cite{leontyev11}. 
+\todo{JOE: It would be a good idea to be consistent in using ECC resp. MDEC. I think that here shuld be MDEC as it cites Stuchebrukhov.}
 Technically, it is a similar approach to the phenomenological charge-scaling as applied in 
@@
 
-While the scaling factor of $f_q = 0.75$ for ions in water improves their performance and is physically justifiable within the MDECC theory \cite{Leontyev2011},
+While the scaling factor of $f_q = 0.75$ for ions in water improves their performance and is physically justifiable within the MDEC theory \cite{Leontyev2011},
 it is not clear how the same factor should be used for partial charges in molecules,
@@
 which aims to reproduce the electrostatic potential obtained from quantum mechanical calculations around an ensemble the target molecules. 
-The fixed partial atomic charges obtained by this method is hence 
-the best fit of these partial charges that reproduces the electrostatic potential around the provided target conformations. 
-Such averaging is well suited for the application of ECC, 
+The fixed partial atomic charges obtained by this method are hence 
+the best fit of pint charges reproducing the electrostatic potential around provided target conformations. 
+Such a scheme is well suited for the application of ECC, 
 which targets at improving electrostatic interactions using a mean field correction. 
@@
 include  to some extend some of the solvent electronic polarizability effects, i.e.,
-the RESP charges are scaled to fit some experimental observable. 
+the RESP charges are scaled to fit some experimental observables. 
 Thus, we expect that the application of the ECC scaling factor, $f_q$, on the molecular partial charges
-included in the available force fields do not necessarily require the integral, $f_q = \epsilon _{el} ^{-1/2}$, 
+included in the available force fields does not necessarily have to follow the relation $f_q = \epsilon _{el} ^{-1/2}$, 
 but instead it lies between 0.75 (no electronic polarizability in the charge calculation)
@@
 accurately describes the lipid head group response to varying concentrations
-of monovalent and divalent cations. This is critical membrane feature, which is
-poorly reproduced by currently available models, can affect not only membrane properties
+of monovalent and divalent cations. This is a critical membrane feature, which is
+poorly reproduced by currently available models; It affects not only membrane properties
 in the presence of ions but also modulates the interaction with charged moieties in the surface.
@@
 why?
-Does it really make a difference?}.
+Does it really make a difference?
+JOE: Cannot answer this without actually doing it.
+The selection was a matter of choice -- in the beginnig we argued with Samuli about this.
+I said that we shuld select the polar atoms in this selection and leave out the rest as ECC would mean little difference to them.
+Samuli then argued in a way that if it makes no difference, why not apply ECC onto such atoms too?
+Hence we applied ECC on all of the atoms.}.
 However, we do not modify the hydrocarbon chain parameters, because they are
@@
 hydrophobic part of lipid bilayers in most lipids models, including Lipid14 \cite{ollila16}.
-This contrast with the behavior in glycerol backbone and head group
+This contrasts with the behavior in glycerol backbone and head group order
 parameters which call for improvements in all available lipid models~\cite{botan15}.
@@
 the scaling factor $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
+\todo{JOE: following discussion shall be modified in the enlightenment of our recent ECC-discussions.}
 Interestingly, this scaling factor is in line with the estimate
@@
 While, the charge scaling improved the behaviour of lipid-ion interactions,
-it reduced the area per molecule of lipid bilayer without ions below experimental
+it reduced the area per molecule of the lipid bilayer without ions below experimental
 values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
@@
 similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
-\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules}
+\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules.
+JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer charges.
+So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever) and ``semi-scaled'' models.}
 This decreases the solvation free energy by allowing water molecules to
@@
 measured using atomic absorption~\cite{altenbach84}.
+\todo{JOE: I'm afraid that the above description might sound to the reader like for some reasons one method is used with DPPC, but a different one with POPC.
+I think simply writing one way is\dots another way is \dots (or similarly) will be less confusing.}
 In this work, we will use the latter definition. For this we
@@
 or the position of the Gibbs dividing plane between two bulk regions.
+\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
 Therefore, we consider the hydrophobic interior of a membrane having zero concentrations of ions and water
@@
 The total area of the interface is twice the area of the membrane, $2A_b$,
-because bilayers have an interface in both leaflefts.
+because bilayers have an interface in both leaflets.
 
@@
     Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
-    compared with experimental order parameters from \cite{ferreira13} {\color {red} X-label misses glycerol region}.
+    compared with experimental order parameters from \cite{ferreira13}.
     Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
     } 
+    \todo{X-label misses glycerol region. JOE: glycerol belongs to the headgroup part, but, yes, it wouldn't hurt to write it explicitly.}
 \end{figure}

### 2017-09-21 Manuscript/manuscript.tex
Retouche petit Results-valitaion section.

@@
   finished when the value for C2, the scale are fixed in the figure and
-  size of the points are fixed in the figure.}
-The headgroup order parameters computed form our new are slightly larger in the ECC-lipid model in comparison ,
-which is in line with larger P-N vector angle in Fig.~\ref{OrderParameterCHANGESsurf}.
+  size of the points are fixed in the figure.
+  JOE: I think this is fixed now.}
+The headgroup order parameters $\alpha$ and $\beta$ and P-N vector angle are slightly larger in the ECC-lipid model in comparison to Lipid14 (Figs.~\ref{simVSexpNOions} and \ref{OrderParameterCHANGESsurf}).
 With the current data we cannot, however, conclude which is the more realistic
-conformation since in both models other headgroup order parameter agrees
-with experiments within error bars and other does not.
+conformation since in both models one headgroup order parameter agrees
+with experiments within error bars whereas the other does not.
 Slight changes are also observed in glycerol backbone order parameters,
-bu the accuracy of both models is comparable to the state of art lipid models
+but the accuracy of both models is comparable to the state of art lipid models
 available in literature \cite{ollila16}. The agreement between the x-ray scattering
@@
 confirm that the membrane structural properties are well captured. 
+\todo{JOE: I think the following is actually in Fig.~\ref{simVSexpNOions}.}
 A structural comparison of ECC-lipids with Lipid14 can be found 

### 2017-09-27 Manuscript/manuscript.tex
Pavel's comments implemented until section E. 2.

@@
 \author{Hector Martinez-Seara Monne}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
 \author{Pavel Jungwirth}
@@
 Prague 6, Czech Republic}
+\affiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
+Tampere, Finland}
 
@@
 Cation interactions with cellular membranes play a key role in several biological processes,
-like in signal propagation in neurons and vesicle fusion.
-\todo{This could be maybe extended.}
+such as in signal propagation in neurons and vesicle fusion.
 \todo{JOE: following doesn't sound right to me. The reasoning is weird.}
@@
 cellular membranes. Interactions of biological cations, especially Na$^+$ and
-Ca$^{2+}$, with PC bilayers are widely studied with
+Ca$^{2+}$, with PC bilayers are widely studied in
 experiments~\cite{akutsu81,altenbach84,seelig90,cevc90,tocanne90,binder02,pabst07,uhrikova08}
 and classical MD simulations~\cite{bockmann03,bockmann04,Berkowitz12,melcrova16,javanainen17}.
-The details of ion binding are, however, not agreed in the literature.
+The details of ion binding are, however, not fully consistent in the literature.
 %not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
@@
 in PC lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
-Interpretation of these experiments suggests that Na$^+$ ions have
+Interpretation most of these experiments suggests that Na$^+$ ions exhibit
 negligible binding in PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
@@
 molecular dynamics (MD) simulation models, however, predict significantly
-stronger binding for the cations and interactions with 3-4 lipids,
+stronger binding for the cations than NMR experiments \cite{catte16}.
+On the other hand, some experiments have also been interpreted to support the predictions from
+MD simulations~\cite{bockmann03,vacha09a}. 
+Interactions of Calcium ions with 3-4 lipids,
 including also interactions with carbonyl
-oxygens~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
-Some experiments have also been interpreted to support the predictions from
-MD simulations~\cite{bockmann03,vacha09a}.
+oxygens, have been reported from simulations~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
 
 Recent work published by the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
-made an attempt to resolve the apparent controversies. The study presented 
+made an attempt to resolve the apparent controversies. This study presented 
 a direct comparison between ion binding affinity between simulations and
@@
 parameters and the electrometer concept~\cite{seelig87}.
-Using massive data collected by Open Collaboration method, it was concluded
+Using massive amounts of data collected by Open Collaboration method, it was concluded
 that the accuracy of the current state of the art lipid models for MD
@@
 MD simulations of ions in bulk water~\cite{jungwirth17-new-paper-to-be-published,Pluharova2014,kohagen14,kohagen16}. 
-As a starting point we use the parameters from Lipid14 model \cite{dickson14},
+As a starting point we use the parameters from the Lipid14 model \cite{dickson14},
 which gave the best cation binding behaviour in the previous study \cite{catte16}.
 The developed ECC-lipid parameters reproduce the experimentally
-measurable structural parameters of POPC lipid bilayer with the 
-accuracy comparable to the other state of the art lipid models.
-
-%We will validate such a model by comparing its lipid bilayer structure
-%to x-ray scattering data and its head group structure to NMR experiments. 
-%We will then quantify the accuracy of the head group order parameter response to the bound charge
-%and compare it to the experimental NMR data of sodium and calcium binding in PC lipid bilayers. 
-%In addition, we will quantify relative binding affinities of different moieties in POPC to such cations. 
-%This will be discussed in terms of binding modes, stoichiometry and binding isotherms and their models used in experiments.
-
-Since the cation binding affinity and the headgroup 
-order parameter changes are in good agreement with experiments
-in the proposed ECC-lipids, it can be used to interpret the
-related structural changes and ion binding stoichiometry.
-New lipid models with correct ion binding affinity in lipid
-bilayers are necessary in applications of 
-MD simulations with physiological salt conditions. 
-The overestimated cation binding in the current lipid models \cite{catte16}
-may lead to significant artefacts in MD simulations. For example,
-articifially positively charged membranes overestimate interactions
-with biomolecules having opposite sign.
-
-%MD simualtions (e.g.~divalent cation-induced charge inversion of bacterial membranes~\cite{Luan2016}
-
-%\todo{JOE: this may be a too daring statement. \\
-%  SAMULI: I think that the divalent cation induced charge inversion is not an artefact,
-%  but it might happen with too small concentrations. 
-%  On the other hand, it is not a new observation either, see e.g. literature cited in NMRlipids IV project.}).
+measurable structural parameters of an ion-free POPC lipid bilayer with the 
+accuracy comparable to the other state of the art lipid models, while surpassing
+them significantly for reproducing the membrane binding affinities and induced
+structural effects of Sodium and Calcium ions.
+
+%Since the cation binding affinity and the headgroup 
+%order parameter changes are in good agreement with experiments
+%in the proposed ECC-lipids, it can be used to interpret the
+%related structural changes and ion binding stoichiometry.
+%New lipid models with correct ion binding affinity in lipid
+%bilayers are necessary in applications of 
+%MD simulations with physiological salt conditions. 
+%The overestimated cation binding in the current lipid models \cite{catte16}
+%may lead to significant artefacts in MD simulations. For example,
+%articifially positively charged membranes overestimate interactions
+%with biomolecules having opposite sign.
 
@@
 The lack of electronic polarizability in the standard MD simulation
-force fields has been considered higly relevant issue since the early days of
+force fields has been considered a higly relevant issue since the early days of
 lipid bilayer simulations.
@@
 In this work we circumvent the rather demanding explicit inclusion
-of electronic polariization effects \cite{chowdhary13} by implicitly including
+of electronic polarization effects \cite{chowdhary13} by implicitly including
 electronic polarizability in lipid
@@
 the early studies where a scaling factor one half was used \cite{jonsson86,egberts94}.
-However, the concept of ECC is 
-physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11},
-but significantly more simple than the explicit inclusion of electronic
-polarizability \cite{chowdhary13} both for derivation and application.
+\todo{We should also cite papers where empirical scaling was used ionic liquids - but there the factor is not 0.5.}
+However, the present concept of ECC is 
+physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11}.
+%but significantly more simple than the explicit inclusion of electronic
+%polarizability \cite{chowdhary13} both for derivation and application.
 
@@
 effectively representing the newly introduced electronic continuum. 
-Assuming globally a  high frequency dielectric constant as measured in water (square of the refraction index),  $\epsilon _{el} = 1.78$, results 
+Assuming globally a  high frequency dielectric constant as measured in water (corresponding to the square of the refraction index),  $\epsilon _{el} = 1.78$, results 
 in a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
-successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??} which then agree with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
-
-While the scaling factor of $f_q = 0.75$ for ions in water improves their performance and is physically justifiable within the MDEC theory \cite{Leontyev2011},
-it is not clear how the same factor should be used for partial charges in molecules,
-e.g., lipids in our case. Unlike the total charge of molecules, atomic partial charges within each molecule are not physical observables.
+successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??} which then agree quantitatively with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+
+While the scaling factor of $f_q = 0.75$ for ions in water improves their performance and is physically justifiable within the ECC theory \cite{Leontyev2011},
+it is not clear whether the same factor should be used for partial charges in molecules,
+e.g., lipids in our case. Unlike the total charge of an atom or molecules, atomic partial charges within each molecule are not physical observables.
 There are several schemes for the assignment of partial charges for biomolecules. \cite{Hu2007}
-Currently, the most commonly employed scheme is RESP~\cite{RESP_paper,Singh1984}, 
-which aims to reproduce the electrostatic potential obtained from quantum mechanical calculations around an ensemble the target molecules. 
-The fixed partial atomic charges obtained by this method are hence 
-the best fit of pint charges reproducing the electrostatic potential around provided target conformations. 
-Such a scheme is well suited for the application of ECC, 
-which targets at improving electrostatic interactions using a mean field correction. 
+Currently, the most commonly employed scheme is the restrained electrostatic potential (RESP) method~\cite{RESP_paper,Singh1984}.
+%
+%PAVEL: i do not understand what you mean by "ensemble" - resp can be done for a single configuration of a single molecule.
+%In any case, tehre is no need to explain here how RESP works so i would drop this part (including the next two sentences).
+%
+%SAMULI: I commented out the suggested sentences
+%which aims to reproduce the electrostatic potential obtained from quantum mechanical calculations around an ensemble the target molecules. 
+%The fixed partial atomic charges obtained by this method are hence 
+%the best fit of pint charges reproducing the electrostatic potential around provided target conformations. 
+%Such a scheme is well suited for the application of ECC, 
+%which targets at improving electrostatic interactions using a mean field correction. 
 In practice, partial charges currently implemented in force fields may already 
 include  to some extend some of the solvent electronic polarizability effects, i.e.,
-the RESP charges are scaled to fit some experimental observables. 
-Thus, we expect that the application of the ECC scaling factor, $f_q$, on the molecular partial charges
+the RESP charges are often scaled to fit some experimental observables \todo{This needs a citation}. 
+Thus, we expect that the application of the ECC scaling factor, $f_q$, to the molecular partial charges
 included in the available force fields does not necessarily have to follow the relation $f_q = \epsilon _{el} ^{-1/2}$, 
-but instead it lies between 0.75 (no electronic polarizability in the charge calculation)
-and $1$ (full electronic polarizability already included in the charge calculation).
+but instead it lies between 0.75 (no electronic polarizability in the partial charge calculation)
+and $1$ (full electronic polarizability already included in the partial charge calculation).
 %
@@
 
-In this work, we produce a phospholipid model for classical MD simulations that
+In this work, we develop a phospholipid model for classical MD simulations that
 accurately describes the lipid head group response to varying concentrations
-of monovalent and divalent cations. This is a critical membrane feature, which is
-poorly reproduced by currently available models; It affects not only membrane properties
+of monovalent and divalent cations. This is a biologically highly relevant membrane feature, which is
+poorly reproduced by currently available models which can affect not only on membrane properties
 in the presence of ions but also modulates the interaction with charged moieties in the surface.
@@
 as discussed in Ref. \citenum{catte16} in section \ref{section:electrometer}.
-To this end, we empirically explore the scaling factor parameter space, $f_q \in [0.75, 1.0]$ on the Lipid14~\cite{dickson14} force field.
-We selected this force field as a starting point because their response to bound ions was apparently the most realistic against NMR data
+To this end, we empirically explore the scaling factor parameter space, $f_q \in [0.75, 1.0]$ for the Lipid14~\cite{dickson14} force field.
+We selected this force field as a starting point because its response to bound ions was apparently the most realistic against NMR data
 in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
@@
 glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
-are expected to have the largest contribution to the cation binding \todo{Do we really need to scale charges in non charged methylens in the choline and glycerol backbones?
-why?
-Does it really make a difference?
-JOE: Cannot answer this without actually doing it.
-The selection was a matter of choice -- in the beginnig we argued with Samuli about this.
-I said that we shuld select the polar atoms in this selection and leave out the rest as ECC would mean little difference to them.
-Samuli then argued in a way that if it makes no difference, why not apply ECC onto such atoms too?
-Hence we applied ECC on all of the atoms.}.
-However, we do not modify the hydrocarbon chain parameters, because they are
+are expected to have the largest contribution to the cation binding
+\todo{Do we really need to scale charges in non charged methylens in the choline and glycerol backbones? why? Does it really make a difference?
+  JOE: Cannot answer this without actually doing it. The selection was a matter of choice -- in the beginnig we argued with Samuli about this.
+  I said that we shuld select the polar atoms in this selection and leave out the rest as ECC would mean little difference to them. Samuli
+  then argued in a way that if it makes no difference, why not apply ECC onto such atoms too? Hence we applied ECC on all of the atoms.}.
+However, we do not modify the hydrocarbon chain parameters, because they
+do not come in contact with salt ions and are
 already highly optimized and give generally a good description for
 hydrophobic part of lipid bilayers in most lipids models, including Lipid14 \cite{ollila16}.
-This contrasts with the behavior in glycerol backbone and head group order
+This is in contrasts with the behavior in glycerol backbone and head group order
 parameters which call for improvements in all available lipid models~\cite{botan15}.
@@
 Exploring different scaling factor values, we found out that ion binding and
-related head group order parameter responses become weaker. 
+related head group order parameter responses become weakened in general. 
 The optimal behaviour of ion binding was observed with
-the scaling factor $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
+the scaling factor of $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
 \todo{JOE: following discussion shall be modified in the enlightenment of our recent ECC-discussions.}
@@
 backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
-which is significantly smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
+which is smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
 and the original Lipid14 value $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
-The decrease of area was found to arise from a lower hydration of the lipid head group region,
+The decrease of the area per lipid was found to arise from a lower hydration of the lipid head group region,
 which can be explained by the increased solvation free energy due to the lower polarity
-of molecules with scaled charges. The hydration can be increased
+of molecules with scaled charges. The hydration can be increased back
 by reducing the effective radius of atoms by changing the 
-the $\sigma$ parameters in Lennard-Jones potential for the selected atoms
+the $\sigma$ parameters in the Lennard-Jones potential for the selected atoms
 similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
 \todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules.
-JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer charges.
-So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever) and ``semi-scaled'' models.}
+  JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer
+  charges. So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever)
+  and ``semi-scaled'' models.}
 This decreases the solvation free energy by allowing water molecules to
-approach closer to lipid atoms and have stonger electrostatic interactions with them.
-After reducing the $\sigma$ parameters with a factor of $f_\sigma = 0.89$
+approach closer to lipid atoms and have stronger electrostatic interactions with them.
+After reducing the $\sigma$ parameters by a factor of $f_\sigma = 0.89$
 for the same atoms for which charges were scaled, the area per molecule value was
-again in agreement with experimental value (see Table~\ref{tab:apls}).
+back in agreement with the experimental value (see Table~\ref{tab:apls}).
 
 \subsection{Electrometer concept} \label{section:electrometer}
-Ion binding was compared between experiments and simulations by 
+Ion binding between experiments and simulations was compared by 
 using lipid head group order parameters and the
@@
 $\alpha$ and $\beta$ segments in head group, can be accurately measured
-by using $^2$H NMR or $^{13}$C NMR techniques. From MD simulations the
-order parameters can be calculated by using the definition
+using $^2$H NMR or $^{13}$C NMR techniques. From MD simulations the
+order parameters can be calculated using the definition
 \begin{equation}\label{OP}
@@
 The measured order parameter change depends on the response of 
-headgroup on bound charge, i.e. $m_i$ in Eq.~\ref{OPchangeEQ},
+the headgroup on the bound charge, i.e. $m_i$ in Eq.~\ref{OPchangeEQ},
 and the ion binding affinity. Thus, the former property has to be
-well quantified before using using the electrometer concept to
-analyze binding affinities. This is done experimentally for wide range
+well quantified before using the electrometer concept to
+analyze binding affinities. This is done experimentally for a wide range
 of systems~\cite{seelig87,beschiasvili91}. To calibrate 
@@
 we use experimental data for dihexadecyldimethylammonium (C$_{12}$C$_{16}$$^+$N2C$_1$Br$^-$)
-in POPC bilayer \cite{scherer89}. Dihexadecyldimethylammonium
+in a POPC bilayer \cite{scherer89}. Dihexadecyldimethylammonium
 is a cation surfactant having two acyl chains and bearing a unit
 charge in the hydrophilic end. Thus, it is expected to locate
-in bilayer similarly to the phospholipids and the molar ratio then 
+in the bilayer similarly to the phospholipids and the molar ratio then 
 gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems \cite{scherer87}.
@@
 \subsection{Salt concentrations and binding affinity}
-Experimental works providing the lipid head group order
-parameters report the salt concentrations in two different ways.
+Experimental studies providing the lipid head group order
+parameters report salt concentrations in two different ways.
 For DPPC, salt concentrations before solvating
-the lipids are used~\cite{akutsu81}. Instead, for POPC,
+the lipids are used~\cite{akutsu81}. Alternatively, for POPC,
 the salt concentration of supernatant after solvation is
 measured using atomic absorption~\cite{altenbach84}.
-\todo{JOE: I'm afraid that the above description might sound to the reader like for some reasons one method is used with DPPC, but a different one with POPC.
+\todo{JOE: I'm afraid that the above description might sound to the reader
+  like for some reasons one method is used with DPPC, but a different one with POPC.
   I think simply writing one way is\dots another way is \dots (or similarly) will be less confusing.}
 In this work, we will use the latter definition. For this we
-will measure the salt concentration in the water bulk region,
+will calculate the salt concentration in the aqueous bulk region,
 \textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
@@
 the concentration was calculated as in the DPPC experiments.
-Despite of the measurable differences between two concentrations for CaCl$_2$ systems,
-the qualitatively conclusions are not affected neither in this
+Despite of the measurable differences between these two concentrations for CaCl$_2$ systems,
+the qualitative conclusions are affected neither in this
 nor in our previous work~\cite{catte16}.
@@
 where $n_w$ and $n_i$ are the total number of waters and ions in the system;
-$C_w$ and $C_i$ are their respective bulk concentrations in the water phase;
+$C_w$ and $C_i$ are their respective bulk concentrations in the aqueous phase;
 and $A_b$ is the size of the box in the membrane plane.
 The total area of the interface is twice the area of the membrane, $2A_b$,
-because bilayers have an interface in both leaflets.
+because bilayers have an interface at both leaflets.
 
@@
 \subsubsection{Simulations of POPC bilayers in aqueous ions}
-The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
-(POPC) bilayer and in aqueous solutions of varying salt concentrations.
+The simulated systems consisted of POPC bilayer in pure water
+or in varying salt concentrations.
 In particular, the periodic orthorhombic simulation box contained 
 128 POPC molecules and approximately 50~water molecules per each lipid.
-Our default, water molecules were described by OPC3 model~\cite{Izadi16} 
+As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
 which is currently the most accurate three site rigid water model. 
 In order to test transferability of our newly developed ECC-lipids model, 
-we also performed several additional simulations with water models 
-OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}
+we also performed several additional simulations with 
+OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
 presented in Supporting Information (SI). 
-We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
 Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
-Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
+MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
 The simulation settings used in this work are 
@@
 \subsubsection{Simulations of POPC bilayers with cationic surfactants}
-Automated topology builder \cite{malde11} was first used to create the structure of
+An automated topology builder \cite{malde11} was first used to create the structure of
 dihexadecyldimethylammonium bromide cationic surfactant,

### 2017-09-28 Manuscript/manuscript.tex
Updates based on Pavel's comments until III. A.

@@
 the head group order parameter response also in simulations,
-we use experimental data for dihexadecyldimethylammonium (C$_{12}$C$_{16}$$^+$N2C$_1$Br$^-$)
+we use experimental data for dihexadecyldimethylammonium (2C$_{16}$$^+$N2C$_1$Br$^-$)
 in a POPC bilayer \cite{scherer89}. Dihexadecyldimethylammonium
@@
 An automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium bromide cationic surfactant,
-
-\vspace{0.5cm}
+dihexadecyldimethylammonium \\[0.5cm]
+%\vspace{0.5cm}
 \chemfig{
-(-[:0,1.0,,,draw=none]\chemabove{Br}{\scriptstyle\ominus}) -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{12}-CH_3)  
-} .
-\vspace{0.5cm}
-
-AmberTools program \cite{amber} was then used to generate the Amber-type force field
+ -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{16}-CH_3)
+} , 
+\vspace{0.5cm} \\
+which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
+The AmberTools program \cite{amber} was then used to generate the Amber-type force field
 parameters. The parameters were converted to the Gromacs format by using
-acpype tool \cite{acpype}. The partial charges were then manually modified
+the acpype tool \cite{acpype}. The partial charges were then manually modified
 to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
 The surfactants were randomly placed among the lipids to form bilayer structures with
-mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
+mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
 All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
-6, 14, 21, 35 or 50 surfactants per leaflet. Parameters for Bromium ions were
-not available in the standard Gromacs files for Amber force field, thus the
-parameters for Chloride ions were used instead.
+6, 14, 21, 35, or 50 surfactants per leaflet. Chloride counter ions were used
+in simulations, because parameters for bromide, which was used in the experiment \cite{scherer89},
+were not available in the standard Gromacs files for Amber force field.
 The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
@@
 
-The same systems were also simulated with ECC-lipid model for POPC using the same setup. 
+The same systems were also simulated with the ECC-lipid model for POPC using the same setup. 
 In these simulations the ECC correction was also applied to the cationic surfactant 
-by scaling all charges with the same factor as for ECC-lipids, $f_q=0.8$, 
+by scaling all charges with the same factor as for ECC-lipids, i.e., $f_q=0.8$, 
 and by using the atom types with reduced $\sigma$ parameters from ECC-lipids.
@@
     } 
-    \todo{X-label misses glycerol region. JOE: glycerol belongs to the headgroup part, but, yes, it wouldn't hurt to write it explicitly.}
+  \todo{X-label misses glycerol region.
+JOE: glycerol belongs to the headgroup part, but, yes, it wouldn't hurt to write it explicitly.
+SAMULI: I would maybe remove the whole text, tell the information in caption and increase the order parameter figure size in z-direction.}
+  \todo{Also acyl chain points should correspond the legends and maybe have the same size as HG and glycerol.}
+  \todo{Legends would be more clear without multiple points.}
 \end{figure}
@@
 
-In order to validate the newly developed model, ECC-lipids, 
-we compared our simulation results without any ions to
+In order to validate the structural properties of the newly developed ECC-lipids model
+we first compared our simulation results without any ions to
 NMR order parameters measurements and x-ray scattering form factors 
 (Fig.~\ref{simVSexpNOions}). 
-Our tail order parameters, as in the original Lipid14 model~\cite{dickson14},
-match the experimental values. This is expected as the ECC model does not modify
-the already highly optimized parameters of this region in Lipid14. 
-\todo{The discussion about acyl chain and headgroup order parameters needs to be
-  finished when the value for C2, the scale are fixed in the figure and
-  size of the points are fixed in the figure.
-  JOE: I think this is fixed now.}
-The headgroup order parameters $\alpha$ and $\beta$ and P-N vector angle are slightly larger in the ECC-lipid model in comparison to Lipid14 (Figs.~\ref{simVSexpNOions} and \ref{OrderParameterCHANGESsurf}).
+As in the original Lipid14 model~\cite{dickson14}, the acyl chain order parameters agree
+well with the experimental values. Notably,
+
+and in almost all the other available
+lipid models \cite{ollila16}
+
+%This is expected as the ECC model does not modify
+%the already highly optimized parameters of this region in Lipid14. 
+The headgroup order parameters $\alpha$ and $\beta$ and P-N vector angle are slightly larger
+in the ECC-lipid model in comparison to Lipid14 (Figs.~\ref{simVSexpNOions} and \ref{OrderParameterCHANGESsurf}). 
 With the current data we cannot, however, conclude which is the more realistic
@@
     Headgroup order parameter changes and P-N vector orientation as a function of
-    cationic surfactant (dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$)
+    cationic surfactant (dihexadecyldimethylammonium bromide, 2C$_{16}$$^+$N2C$_1$Br$^-$)
     in PC bilayer from simulations and experiments \cite{scherer89}.

### 2017-09-29 Manuscript/manuscript.tex
Modifications based on Pavel's comments until the end of section III A.

@@
 JOE: glycerol belongs to the headgroup part, but, yes, it wouldn't hurt to write it explicitly.
-SAMULI: I would maybe remove the whole text, tell the information in caption and increase the order parameter figure size in z-direction.}
-  \todo{Also acyl chain points should correspond the legends and maybe have the same size as HG and glycerol.}
+SAMULI: I would maybe remove the whole text, tell the information in caption and increase the order parameter figure size in z-direction.} \\
+  \todo{Also acyl chain points should correspond the legends and maybe have the same size as HG and glycerol.} \\
   \todo{Legends would be more clear without multiple points.}
@@
 
-In order to validate the structural properties of the newly developed ECC-lipids model
-we first compared our simulation results without any ions to
-NMR order parameters measurements and x-ray scattering form factors 
-(Fig.~\ref{simVSexpNOions}). 
-As in the original Lipid14 model~\cite{dickson14}, the acyl chain order parameters agree
-well with the experimental values. Notably,
-
-and in almost all the other available
-lipid models \cite{ollila16}
-
+%In order to validate the structural properties of the newly developed ECC-lipids model
+%we first compared our simulation results without any ions to
+%NMR order parameters measurements and x-ray scattering form factors 
+%(Fig.~\ref{simVSexpNOions}). 
+The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and
+the area per lipid (Table~\ref{tab:apls}) of POPC bilayer simulated
+with the ECC-lipid model are in good agreement with the experimental results,
+as well as with the Lipid14 model. Thus, we conclude that the lipid bilayer dimensions
+are well captured by the developed ECC-lipid model. 
+%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
+%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
+
+As in the original Lipid14 model~\cite{dickson14}, the acyl chain order parameters of the
+ECC-lipid model agree well with the experimental values in Fig.~\ref{simVSexpNOions}. Notably, the experimentally measured forking and
+small order parameter values of $C_2$ segment in {\it sn}-2 chain are relatively well
+reproduced by the both models. This has been suggested to indicate that the carbonyl
+region of {\it sn}-2 chain is directed towards the water phase, in contrast to the
+carbonyl in {\it sn}-1 chain, which would orient more along the bilayer
+plane~\cite{seelig75,schindler75,gawrisch92}. While this may be an important
+feature for the ion binding details, it is not necessarily reproduced by the
+available lipid models~\cite{ollila16}
 %This is expected as the ECC model does not modify
 %the already highly optimized parameters of this region in Lipid14.
-The headgroup order parameters $\alpha$ and $\beta$ and P-N vector angle are slightly larger
-in the ECC-lipid model in comparison to Lipid14 (Figs.~\ref{simVSexpNOions} and \ref{OrderParameterCHANGESsurf}). 
-With the current data we cannot, however, conclude which is the more realistic
-conformation since in both models one headgroup order parameter agrees
-with experiments within error bars whereas the other does not.
-Slight changes are also observed in glycerol backbone order parameters,
-but the accuracy of both models is comparable to the state of art lipid models
-available in literature \cite{ollila16}. The agreement between the x-ray scattering
-form factors (Fig.~\ref{simVSexpNOions}) and the areas per molecule from simulations and experiments (Table~\ref{tab:apls}) 
-confirm that the membrane structural properties are well captured. 
-\todo{JOE: I think the following is actually in Fig.~\ref{simVSexpNOions}.}
-A structural comparison of ECC-lipids with Lipid14 can be found 
-in SI along with results with other water models. 
-%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
-%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
+
+The headgroup order parameters of $\alpha$ and $\beta$ carbons are slightly larger
+in the ECC-lipid model than in Lipid14, which is apparently related to the
+smaller P-N vector angle with respect to the membrane normal in
+Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
+however, conclude which one of the models give the more realistic
+headgroup conformations. The ECC-lipid model gives
+the $\beta$ carbon order parameter value closer to experiments, while
+value for $\alpha$ carbon is better in Lipid14.
+Despite of some deviations from the experimental order parameter values
+in Fig. \ref{simVSexpNOions},
+the accuracy of the both models in the glycerol backbone region
+is comparable to the other state
+of art lipid models available in literature \cite{botan15}.
 
@@
 \todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
-  do this analysis by using the density profiles, but it is not necessary.}
+  do this analysis by using the density profiles, but it is not necessary.
+  The reason why I want to do something like this, is that it would be useful to have
+  some easier way to estimate the correct binding affinity than the electrometer concept,
+  which is quite tedious to apply in practise (simulations with different concenrations,
+  cationic surfactant check, etc.). I wuold like to be able to estimate much faster from a
+  simulation if the binding affinity is reasonable or not. This may not be the correct way for
+that anyway.}
 Rough estimates for the free energy difference between bound and unbound cation
@@
 Data -- up to noted exceptions -- shall be all there, however. 
+\todo{SAMULI: I think the we should make clear that the current binding
+  constants in, e.g., Marsh's handbook are based on the headgroup order parameter
+  changes intepreted with ternary complex binding model. I.e. the experimental raw data
+  is excatly the same as we have in this work. Our simulations enable much more versatile interpretation
+  of the binding phenomena. I think that this is one the main points of this work, so this comment
+  applies to the previous section, introduction, abstract and conclusions as well.
+}
 
@@
 
-\todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
-  be the only option. I will recheck and come back to this.
-  SAMULI: This is written in \cite{macdonald87}:
-  ''\ce{Ca^{2+}} binding to POPC bilayers over the whole concentration range can be best described in terms
-  of formation of a ternary complex involving complexation of two lipids to one calcium ion (Altenbach and Seelig, 1984).
-  The addition of a sodium competition term has not changed this conclusion. However, if \ce{Ca^{2+}} concentrations
-  up to 100 mM are considered, the data can be equally well explained by a 1:1 binding mechanism (cf. Figure 7).
-  In contrast, the Ca2+ binding to POPC-POPG mixtures can be best described by assuming a 1:l stoichiometry regardless of the range of Ca2+concentrations.''
-  We might or might not want to discuss about this.
-  JOE: I think this is adressed/acknowledged enough now.} \\
 \todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},

### 2017-09-29 Manuscript/manuscript.tex
One more comment

@@
 An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions.
+\todo{Did you really calculate the $C_I$ from simulations without severe assumptions?
+Note that this conctration at the plane of biding, which do not equal the concentration of free cations.}
 Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 

### 2017-10-03 Manuscript/manuscript.tex
Simple number of contacts analysis added to MS.

@@
 of several moieties in POPC towards \ce{Ca^{2+}}. 
-\todo{Add a simple analysis using number of contacts.}
-Based on the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
-we estimate that the largest contribution
+The dominant contribution
 to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
-Although the isodensity plots are relatively easy to interpret,
-the contours shown in Fig.~\ref{fig:volmaps} cannot conclusively tell on the details
-of \ce{Ca^{2+}} binding to any of the two carbonyl moieties
-but their apparently lower affinity compared to phosphate oxygens.
+Such a finding is supported by the calculated
+number of contacts between \ce{Ca^{2+}} and POPC oxygen atoms.
+The average number of contacts between \ce{Ca^{2+}} and any oxygen atom in POPC is 59.8, 
+whereas if only phosphate oxygen atoms are considered 
+the average number of contacts decreases only by a tiny amount to 57.2. 
+We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
+which are easier to interpret.
 

### 2017-10-04 Manuscript/manuscript.tex
Changed eps figures in manuscript to pdf

@@
   \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
   \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
@@
   \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.eps}
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
   \caption{\label{OrderParameterCHANGESsurf}

### 2017-10-04 Manuscript/manuscript.tex
Major change in order parameter changes plot -- split NaCl and CaCl2 plot to two separate figures as it follows the story better (and looks nicer, too)

@@
 %head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
-%observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
+%observed in  Ref. \citenum{catte16}, see also Fig. \ref{fig:delta_ordPar_NaCl}.
 %The present comparison of head group order parameter changes 
@@
 
-\begin{figure*}[tbp]
+\begin{figure}[tbp]
   \centering
-  \includegraphics[width=16.0cm]{../Fig/OrdParChanges_NaCl_CaCl2.eps}
-  \caption{\label{OrderParameterCHANGESnewMODELS}
-    Changes of head group order parameters of POPC bilayer as a function of NaCl and CaCl$_2$ concentrations
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
+  \caption{\label{fig:delta_ordPar_NaCl}
+    Changes of head group order parameters of POPC bilayer as a function of NaCl concentrations
+    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
+    Ion concentrations in bulk water are shown in x-axis. 
+    Values from simulations are calculated from the of cation number density $C_{np}$
+    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
+  }
+\end{figure}
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=7.5cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
+  \caption{\label{fig:delta_ordPar_CaCl}
+    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
     are shown from simulations with different force fields together with experimental data 
@@
   }
-\end{figure*}
+\end{figure}
 
@@
 to aqueous NaCl and CaCl$_2$ concentrations 
-are shown in Fig. \ref{OrderParameterCHANGESnewMODELS}
+are shown in Fig. \ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}
 from experiments (DPPC \cite{akutsu81} 
@@
 for the CaCl$_2$ interactions with Lipid14 model, as seen 
-in Figs. \ref{OrderParameterCHANGESnewMODELS} and \ref{??} (in SI), respectively.
+in Figs. \ref{fig:delta_ordPar_CaCl} and S\ref{fig:ordPars_waterModels} (in SI), respectively.
 \todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
@@
 as a function of CaCl$_2$ concentration from ECC-lipid model with ECC-ions
-show a good agreement with experiments in Fig. \ref{OrderParameterCHANGESnewMODELS}.
+show a good agreement with experiments in Fig. \ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}.
 As discussed in previous section, the model gives also
@@
 
+\begin{figure*}[tbp]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
+  \caption{\label{fig:ordPars_waterModels}
+    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
+    are shown from simulations with different force fields and water models together with experimental data 
+    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
+    Ion concentrations in bulk water are shown in x-axis. 
+    Values from simulations are calculated from the of cation number density $C_{np}$
+    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
+  }
+\end{figure*}
+
 % Create the reference section using BibTe

### 2017-10-05 Manuscript/manuscript.tex
Corrected DHDMA formula, changed APL tables, added extra APls in SI.

@@
 \chemfig{
- -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{16}-CH_3)
+ -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH2{)}_{15}-CH_3)(-[:30]{(}CH2{)}_{15}-CH_3)
 } , 
@@
 All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
-6, 14, 21, 35, or 50 surfactants per leaflet. Chloride counter ions were used
-in simulations, because parameters for bromide, which was used in the experiment \cite{scherer89},
-were not available in the standard Gromacs files for Amber force field.
+6, 14, 21, 35, or 50 surfactants per leaflet. 
+%Chloride counter ions were used
+%in simulations, because parameters for bromide, which was used in the experiment \cite{scherer89},
+%were not available in the standard Gromacs files for Amber force field.
 The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
@@
 \begin{table}
-  \caption{Area per lipid (APL) from different models of POPC without ions\label{tab:apls} }
+  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls} }
   \begin{tabular}{l|c c}
@@
     \hline
+    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
     Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
     \hline
-    ECC-lipids &        &  \\
-    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
-    ($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
+    ECC-lipids                & 62.2$\pm$ 0.6  &  300       \\
     \hline
-    (6.4 nm)$^2$, 128 lipids patch, OPC3           & 64.2   &  313       \\
-    (6.4 nm)$^2$, 128 lipids patch, SPC/E          & 65.1   &  313       \\
-    (6.4 nm)$^2$, 128 lipids patch, OPC            & 64.4   &  313       \\
-    (6.4 nm)$^2$, 128 lipids patch, TIP4p/2005     & 66.8   &  313       \\
-    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
-    %oMM small patch           & 63.65  &         \\
-    %oMM 4xbig patch           & 63.7   &         \\
-    \hline
-    %experiment   & 62.7  &  293    \\
     experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
-    experiment  & 67.3  &  323    \\
-    %experiment  & 68.1  &  333    \\
-    %experiment POPE  & 56.6 &  303    \\
     \hline
   \end{tabular}
-  \todo{SAMULI: I would put here Lipid14 in 303K, ECC-lipid in 303K and experiment in 303K. Rest in SI.
-  The best experimental value would be the one analyzed from the form factor shown in previous figure, 
-  if available.}
 \end{table}
@@
 
+\begin{table}
+  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
+  \begin{tabular}{l|c c}
+    model          & APL (\AA$^2$)   & Temperature [K] \\
+    \hline
+    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
+    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
+    \hline
+    ECC-lipids &        &  \\
+    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
+    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
+    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
+    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
+    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
+    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
+    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
+    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
+    %oMM small patch           & 63.65  &         \\
+    %oMM 4xbig patch           & 63.7   &         \\
+    \hline
+    experiment   & 62.7  &  293    \\
+    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
+    experiment  & 67.3  &  323    \\
+    experiment  & 68.1  &  333    \\
+    %experiment POPE  & 56.6 &  303    \\
+    \hline
+  \end{tabular}
+\end{table}
+
 % Create the reference section using BibTe

### 2017-10-05 Manuscript/manuscript.tex
Moved delta-OP plots into SI.

@@
   \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
-  \caption{\label{fig:delta_ordPar_NaCl}
-    Changes of head group order parameters of POPC bilayer as a function of NaCl concentrations
-    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
-    Ion concentrations in bulk water are shown in x-axis. 
-    Values from simulations are calculated from the of cation number density $C_{np}$
-    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
-  }
-\end{figure}
-\begin{figure}[tbp]
-  \centering
   \includegraphics[width=7.5cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
@@
 to aqueous NaCl and CaCl$_2$ concentrations 
-are shown in Fig. \ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}
+are shown in Fig.~\ref{fig:delta_ordPar_CaCl} and S\ref{fig:delta_ordPar_NaCl}
 from experiments (DPPC \cite{akutsu81} 
@@
 as a function of CaCl$_2$ concentration from ECC-lipid model with ECC-ions
-show a good agreement with experiments in Fig. \ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}.
+show a good agreement with experiments in Fig.~\ref{fig:delta_ordPar_CaCl} and S\ref{fig:delta_ordPar_NaCl}.
 As discussed in previous section, the model gives also
@@
 \end{figure*}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_NaCl.pdf}
+  \caption{\label{fig:delta_ordPar_NaCl}
+    Changes of head group order parameters of POPC bilayer as a function of NaCl concentrations
+    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
+    Ion concentrations in bulk water are shown in x-axis. 
+    Values from simulations are calculated from the of cation number density $C_{np}$
+    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
+  }
+\end{figure}
 

### 2017-10-05 Manuscript/manuscript.tex
Included Pavel's changes until the end of section III. C.

@@
 are well captured by the developed ECC-lipid model. 
+Areas per lipid of ECC-lipids with various other water models are documented in SI, table~\ref{tab:apls_si}. 
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 
-\subsection{Response of POPC head group to bound charge}
+\subsection{Response of POPC head groups to bound charge}
 
@@
 
-Before proceeding to the ion binding affinity studies, we quantify
+Before proceeding to the ion binding affinity, we quantify
 the response of headgroup order parameters to the amount of 
@@
 
-The headgroup order parameter changes with increasing amount of 
-cationic surfactant dihexadecyldimethylammonium bromide 
+The headgroup order parameter changes with an increasing amount of 
+the cationic surfactant dihexadecyldimethylammonium bromide 
 is compared between experiments~\cite{scherer89} and simulations in
 Fig. \ref{OrderParameterCHANGESsurf}. 
-The observed order parameter decrease is simulations and experiments
-can be approximated to be linear at least with the mole fractions
+The observed order parameter decrease in simulations and experiments
+can be approximated to be linear at least for mole fractions
 below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
-The slope is, however, too steep in Lipid14 model indicating that 
+The slope is, however, too steep in the Lipid14 model indicating that 
 the head group order parameters are too sensitive to a bound charge.
@@
 in Fig.~\ref{OrderParameterCHANGESsurf}.
-The headgroup orients more towards the water phase with increasing
+The headgroup orients more towards the water phase with an increasing
 amount of bound cations, as previously reported in Ref.~\citenum{seelig87}. 
@@
 in ECC-lipids model, which is in line with the order parameter 
-results and the reduced charge--dipole interactions in the ECC-lipid model.
+results and the reduced charge--dipole interactions in the latter model.
 The response of $\alpha$-order parameter to bound positive charge
@@
 
-\subsection{Cation binding affinity in POPC}
+\subsection{Cation binding affinities to POPC}
 
@@
 
-The binding affinity of aqueous cations in lipid 
-bilayers can be measured by using the headgroup 
+The binding affinities of aqueous cations to lipid 
+bilayers can be measured using the headgroup 
 order parameters, because they decrease proportionally  
@@
 The headgroup order parameter responses 
-to aqueous NaCl and CaCl$_2$ concentrations 
-are shown in Fig.~\ref{fig:delta_ordPar_CaCl} and S\ref{fig:delta_ordPar_NaCl}
+to increasing aqueous CaCl$_2$ concentrations 
 from experiments (DPPC \cite{akutsu81} 
-and POPC \cite{altenbach84}) and different simulation models for POPC.
+and POPC \cite{altenbach84}) and different simulation models for POPC 
+are shown in Fig.~\ref{fig:delta_ordPar_CaCl}.
+Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.
 
@@
 Ref.~\citenum{catte16}.
-Using ion model by Dang et al.~\cite{smith94,chang1999,dang2006} or
+Using the ion model by Dang et al.~\cite{smith94,chang1999,dang2006} or
 ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
 with more realistic bulk behaviour did not improve the results
-for the CaCl$_2$ interactions with Lipid14 model, as seen 
+for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model, as seen 
 in Figs. \ref{fig:delta_ordPar_CaCl} and S\ref{fig:ordPars_waterModels} (in SI), respectively.
@@
 that improvements also in lipid models are needed to 
-correctly describe divalent cation binding in PC bilayers.
-
-Significant improvement can be achieved by using the ECC
-approach also for lipids. The headgroup order parameter changes
-as a function of CaCl$_2$ concentration from ECC-lipid model with ECC-ions
-show a good agreement with experiments in Fig.~\ref{fig:delta_ordPar_CaCl} and S\ref{fig:delta_ordPar_NaCl}.
+correctly describe divalent cation binding to PC bilayers.
+
+Significant improvement can be achieved using the ECC
+approach also for lipids. The headgroup order parameters change
+as a function of CaCl$_2$ concentration for the ECC-lipid model with ECC-ions
+exhibit a good agreement with experiments in Fig.~\ref{fig:delta_ordPar_CaCl} and S\ref{fig:delta_ordPar_NaCl}.
 As discussed in previous section, the model gives also
@@
 the headgroup response to bound charge.
-Thus, the model can used for more detailed analysis of the
-binding affinity.
-
-The binding affinities are quantified by using
-the water density profiles along membrane normal shown
+%Thus, the model can used for more detailed analysis of the binding affinity.
+
+The binding affinities are quantified using
+the water density profiles along the membrane normal shown
 in Fig.~\ref{fig:cacl-dens}. The density profiles show
-larger Ca$^{2+}$ density peak in lipid headgroup region for
-Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
-The relative surface excess calculated from Eq.~\ref{surfexcess} gives 
+a larger Ca$^{2+}$ density peak in lipid headgroup region for
+the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
+The relative surface excess calculated from Eq.~\ref{surfexcess} is 
 $\Gamma_i^w = 0.07 \pm 0.01 nm^{-2}$ for the ECC-lipid model,
@@
 \begin{equation}
-  \Delta G={\rm k_b T}\log(\frac{p_o}{p_i}),
+  \Delta G = -{\rm k_b T}\log(\frac{p_i}{p_o}),
 \end{equation}
@@
 
-Since the lipid headgroup order parameter responses to the amount of bound charge
+Since the lipid headgroup order parameter responds to the amount of bound charge
 and to the aqueous ion concentrations are both in good agreement with experiments
 in the ECC-lipid model with ECC-ion parameters, we consider the Na$^+$ and Ca$^{2+}$
-binding affinities to be realistic in this model. On the other hand, the Ca$^{2+}$
+binding affinities to be realisticlly described in this model. In contrast, the Ca$^{2+}$
 binding affinity is overestimated by Lipid14 model when simulated with all the
@@
 the discrepancies with experiments in previous work could partly arise also from
-the inaccurate sensitivity of the headgroup to bound charge. Here we quantify this
-effect and conclude that the improvement due to ECC is partly, but not completely,
-caused by more realistic headgroup sensitivity to bound charge.
+the inaccurately described sensitivity of the headgroup to bound charge. Here we quantify this
+effect and conclude that the improvement due to ECC is to a large extent, but not entirely,
+caused by a more realistic headgroup sensitivity to bound charge.
 This indicates that the issue should be carefully considered also when
 the electrometer concept is used to compare ion binding between experiments
-and simulations wiht other models.
+and simulations with other models.
 

### 2017-10-06 Manuscript/manuscript.tex
Pavel's comments and corrections fully implemented into manuscript. This commit: Conclusions.

@@
 \section{Conclusions}
-We show that the \ce{Na^+} and \ce{Ca^2+} binding in phospholipid bilayers can
-be accurately described with classical MD simulation models, where electronic
-polarization is effectively included by using electronic continuum correction (ECC)~\cite{leontyev11}.
+We show that the \ce{Na^+} and \ce{Ca^2+} binding to phospholipid bilayers can
+be accurately described with classical force field MD simulation models, where electronic
+polarization is effectively included using the electronic continuum correction (ECC)~\cite{leontyev11}.
 This is a significant improvement over other available lipid models,
 which all overestimate specific cation binding affinities~\cite{catte16}.  
-The newly proposed model, which we denote as ''ECC-lipids~17'', 
+The newly proposed model, which we denote as ''ECC-lipids'', 
 exhibits accurate head group order parameter response to
-bound cations, monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
-and divalent \ce{Ca^2+}
-also quantifying their binding affinities.
-Moreover, ECC-lipids~17 reproduce the lipid bilayer structural details
+bound cations, namely monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
+and divalent \ce{Ca^2+},
+allowing also for quantifying their binding affinities.
+At the same time, ECC-lipids reproduce the lipid bilayer structural details
 with similar accuracy as other state of the art lipid models~\cite{catte16}.
@@
 were used to exemplify the transferability of 
-the parameters of the new ECC-lipids~17 force field.
-
-Direct analysis of calcium binding details from MD simulations is in agreement
+the parameters of the new ECC-lipids force field.
+
+Direct analysis of the details of calcium binding from MD simulations is in agreement
 with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
-In this model 1~calcium binds to 2~POPC molecules, which together form a ternary
+In this model one calcium binds to two POPC molecules, which together form a ternary
 complex.
@@
 
-The electronic continuum correction is applied here on Lipid14 POPC model \cite{dickson14},
-but we expect that the correction can be generalized also for other lipids
+The electronic continuum correction is built here upon the Lipid14 POPC model \cite{dickson14},
+but we expect that the correction can be used also for other lipids
 and force fields.
-The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. 
+\todo{JOE: Do we want to suggest that ECC-lipids can be used 
+with current force fields to study for example membrane proteins? 
+This was expressed with the following sentence: 
+The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. }
 We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
@@
 % Put your acknowledgments here.
+P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
+and 600 from the Academy of Finland via the FiDiPro award.
 \end{acknowledgments}
+
 \newpage
 \appendix
+
 \begin{center}
@@
 
-\begin{figure}[]
+\begin{figure}[tbp]
   \centering
@@
 
-\begin{figure*}[tbp]
+\begin{figure}[tbp]
   \centering

### 2017-10-08 Manuscript/manuscript.tex

@@
 such as in signal propagation in neurons and vesicle fusion.
-\todo{JOE: following doesn't sound right to me. The reasoning is weird.}
 Since direct measurements of ion-membrane interactions from biological systems
@@
 (ECC)~\cite{leontyev11}. 
-\todo{JOE: It would be a good idea to be consistent in using ECC resp. MDEC. I think that here shuld be MDEC as it cites Stuchebrukhov.}
 Technically, it is a similar approach to the phenomenological charge-scaling as applied in 
@@
 are expected to have the largest contribution to the cation binding
-\todo{Do we really need to scale charges in non charged methylens in the choline and glycerol backbones? why? Does it really make a difference?
-  JOE: Cannot answer this without actually doing it. The selection was a matter of choice -- in the beginnig we argued with Samuli about this.
-  I said that we shuld select the polar atoms in this selection and leave out the rest as ECC would mean little difference to them. Samuli
-  then argued in a way that if it makes no difference, why not apply ECC onto such atoms too? Hence we applied ECC on all of the atoms.}.
 However, we do not modify the hydrocarbon chain parameters, because they
@@
 The simulation settings used in this work are 
-summarized in Table \ref{tbl:mdpar}, 
-and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
-\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
-  this is done in Amber by default. If this is correct, the information in the
-  table do not match with this sentence. Based on semi-extensive testing I made few years
-ago this do not change anything. Anyway, this should be reported consistently.}
+summarized in Table \ref{tbl:mdpar}. 
+%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
+%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
+%this is done in Amber by default. If this is correct, the information in the
+%table do not match with this sentence. Based on semi-extensive testing I made few years
+%ago this do not change anything. Anyway, this should be reported consistently.}
 Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
-
-\begin{table}[btp]
-  \caption{Simulation parameters}
-  \label{tbl:mdpar}
-  \begin{tabular}{ll}
-    simulation property & parameter   \\
-    \hline
-    time-step           & 2~fs         \\
-    equilibration time  & 100~ns  \\
-    simulation time     & 200~ns  \\
-    temperature         & 313~K       \\
-    thermostat          & v-rescale  \cite{bussi07}   \\
-    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
-    long-range electrostatics & PME  \cite{darden93}  \\
-    cut-off scheme      & Verlet \cite{Pall13}      \\
-    Coulomb and VdW cut-off & 1.0~nm \\
-    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
-    constraints for water & SETTLE  \cite{miyamoto92} \\
-    \hline
-  \end{tabular}
-  \todo{This could be moved to SI. Only simulation lengths needs to be mentioned in the main paper.}
-\end{table}
 
@@
     Top: Chemical structure of POPC with definitions of different order parameters calculated.
-    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
+    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  
+    from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
     compared with experimental order parameters from \cite{ferreira13}.
+    Error bars are approximately the size of the points for head group order parameters.
+    Point size for the order parameters of tails are decreased by a factor of 3 to improve the clarity of the plot.
     Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
     } 
-  \todo{X-label misses glycerol region.
-JOE: glycerol belongs to the headgroup part, but, yes, it wouldn't hurt to write it explicitly.
-SAMULI: I would maybe remove the whole text, tell the information in caption and increase the order parameter figure size in z-direction.} \\
-  \todo{Also acyl chain points should correspond the legends and maybe have the same size as HG and glycerol.} \\
-  \todo{Legends would be more clear without multiple points.}
 \end{figure}
@@
   \todo{JOE: Change the figure so that it contains a membrane background} \\
-  \todo{The current data for Dang simulation seems to contain more ions than others.}
 \end{figure}
@@
 
+\begin{table}[btp]
+  \caption{Simulation parameters}
+  \label{tbl:mdpar}
+  \begin{tabular}{ll}
+    simulation property & parameter   \\
+    \hline
+    time-step           & 2~fs         \\
+    equilibration time  & 100~ns  \\
+    simulation time     & 200~ns  \\
+    temperature         & 313~K       \\
+    thermostat          & v-rescale  \cite{bussi07}   \\
+    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
+    long-range electrostatics & PME  \cite{darden93}  \\
+    cut-off scheme      & Verlet \cite{Pall13}      \\
+    Coulomb and VdW cut-off & 1.0~nm \\
+    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
+    constraints for water & SETTLE  \cite{miyamoto92} \\
+    \hline
+  \end{tabular}
+\end{table}
+
 % Create the reference section using BibTe

### 2017-10-09 Manuscript/manuscript.tex
average number of contacts normalized per lipid and reported in a slightly better way in stoichiometry section of MS.

@@
 to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
-Such a finding is supported by the calculated
-number of contacts between \ce{Ca^{2+}} and POPC oxygen atoms.
-The average number of contacts between \ce{Ca^{2+}} and any oxygen atom in POPC is 59.8, 
+Such a finding is based on the calculated
+number of contacts per lipid between \ce{Ca^{2+}} and POPC oxygen atoms, 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
+in which a contact between the two groups is counted 
+if their distance is less than or equal $0.3 \, \mathrm{nm}$ (robustness of this definition discussed in reference~\citenum{melcrova16}). 
+The average number of contacts between \ce{Ca^{2+}} and any oxygen atom in POPC is 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.47 $,    % = 59.8/128 
 whereas if only phosphate oxygen atoms are considered 
-the average number of contacts decreases only by a tiny amount to 57.2. 
+the average number of contacts decreases only by a tiny amount to 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.45 $,    % = 57.2/128 . 
+and if only carbony oxygen atoms are considered,
+this value is merely
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.02 $.    % = (59.8-57.2)/128 . 
 We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 

### 2017-10-09 Manuscript/manuscript.tex
Update conclusions, not ready

--- 2017-10-08 Manuscript/manuscript.tex
@@
 \section{Conclusions}
-We show that the \ce{Na^+} and \ce{Ca^2+} binding to phospholipid bilayers can
-be accurately described with classical force field MD simulation models, where electronic
-polarization is effectively included using the electronic continuum correction (ECC)~\cite{leontyev11}.
-This is a significant improvement over other available lipid models,
-which all overestimate specific cation binding affinities~\cite{catte16}.  
-The newly proposed model, which we denote as ''ECC-lipids'', 
-exhibits accurate head group order parameter response to
-bound cations, namely monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
-and divalent \ce{Ca^2+},
-allowing also for quantifying their binding affinities.
-At the same time, ECC-lipids reproduce the lipid bilayer structural details
-with similar accuracy as other state of the art lipid models~\cite{catte16}.
-Several water models 
-(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
-were used to exemplify the transferability of 
-the parameters of the new ECC-lipids force field.
-
+By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
+to a POPC lipid bilayer can be accurately described with classical MD simulation 
+force fields, where electronic polarization is implicitly included using 
+the electronic continuum correction (ECC)~\cite{leontyev11}.
+The proposed ECC-lipids model is a significant improvement over 
+other available lipid models, which all overestimate specific cation binding affinities~\cite{catte16}.  
+While the structural details of a POPC lipid bilayer simulated with the ECC-lipids
+model agree with experiments with the comparable accuracy to the other state of the art lipid models,
+it also reproduces the experimental lipid head group order parameter responses to
+the cationic surfactant, NaCl and CaCl$_2$ concentrations.
+
+The good agreement with experiments enables the atomic resolution 
+interpretation of NMR experiments by using MD simulations.
+In agreement with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
+the Ca$^{2+}$ ions mainly interact with phosphate oxygens.
+However, the stoichiometry of the binding is significantly more complicated 
+that the simple ternary complex model.
 Direct analysis of the details of calcium binding from MD simulations is in agreement
@@
 which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
+Several water models 
+(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
+were used to exemplify the transferability of 
+the parameters of the new ECC-lipids force field.
 

### 2017-10-10 Manuscript/manuscript.tex
New version of conclusions

@@
 By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
-to a POPC lipid bilayer can be accurately described with classical MD simulation 
-force fields, where electronic polarization is implicitly included using 
+to a POPC lipid bilayer can be accurately described with the classical MD simulation 
+force field, where the electronic polarization is implicitly included using 
 the electronic continuum correction (ECC)~\cite{leontyev11}.
@@
 However, the stoichiometry of the binding is significantly more complicated 
-that the simple ternary complex model.
-Direct analysis of the details of calcium binding from MD simulations is in agreement
-with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
-In this model one calcium binds to two POPC molecules, which together form a ternary
-complex.
-\todo{Continue summary using previous section once it is finished.}
-
-The electronic continuum correction is built here upon the Lipid14 POPC model \cite{dickson14},
-but we expect that the correction can be used also for other lipids
-and force fields.
-\todo{JOE: Do we want to suggest that ECC-lipids can be used 
-with current force fields to study for example membrane proteins? 
-This was expressed with the following sentence: 
-The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. }
-We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
-which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
-Several water models 
-(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
-were used to exemplify the transferability of 
-the parameters of the new ECC-lipids force field.
+than in the previous interpretation of the NMR data based on 
+the ternary complex model, where one calcium binds to two POPC molecules~\cite{altenbach84}.
+The complexes of one calcium bound to two lipids are the most probable also in the
+ECC-lipids model, but the complexes of one or three lipids per one calcium
+were observed to be almost equally likely. While the success of the ternary
+compex model is understandable based on the simulation results,
+a simple binding model cannot detec the complex binding
+oberved in the simulation.
+
+The improved cation binding behaviour to POPC bilayer pave the way for
+simulations of complex biochemical systems with correctly described
+electrostatic interactions in the vicinity of cellular membranes.
+The ECC-lipids model is build by scaling the partial charges and
+the LJ-radius of the headgroup, glycerol backbone and carbonyl atoms of
+the Lipid14 POPC model~\cite{dickson14}. While the Lipid14 model
+is compatible with the AMBER force field family, the compatibility of
+the ECC-lipids model may be compromised due to the changes in intermolecular
+interactions of the scaled atoms. On the other hand, a fully consistent
+ECC-force field should include the correction also in other than lipid
+molecules, including water. The work toward this direction and the extension
+to other lipid molecules and force fields is left for the future studies.
+
+%but we expect that the correction can be used also for other lipids
+%and force fields.
+
+%We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
+%which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
+%Several water models 
+%(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
+%were used to exemplify the transferability of 
+%the parameters of the new ECC-lipids force field.
 
@@
 
+\newpage
 \newpage

### 2017-10-10 Manuscript/manuscript.tex
Update introduction and electrometer concept section

@@
 such as in signal propagation in neurons and vesicle fusion.
-Since direct measurements of ion-membrane interactions from biological systems
+Since the direct measurements of ion-membrane interactions from biological systems
 are difficult, lipid bilayers are often used as models systems for cellular
-membranes. The detailed results from simple model systems can be then
-used to understand the role of ions in complex biological systems.
-
-Zwitterionic phosphocholine (PC) lipid bilayers are commonly used model systems for
-cellular membranes. Interactions of biological cations, especially Na$^+$ and
+membranes. Especially the zwitterionic phosphocholine (PC) lipid bilayers are
+often used as model systems to understand the role of ions in complex biological
+systems~\cite{scherer87,seelig90,cevc90}.
+
+Interactions of biological cations, especially Na$^+$ and
 Ca$^{2+}$, with PC bilayers are widely studied in
@@
 %ions, but do not affect the specific binding constant \cite{seelig90}.
-Non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
-and infrared scpectroscopy, give accurate information about ion binding
-in PC lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
-Interpretation most of these experiments suggests that Na$^+$ ions exhibit
-negligible binding in PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
-binds to phosphate groups of two lipid molecules. Atomistic resolution
-molecular dynamics (MD) simulation models, however, predict significantly
+Interpretations of non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
+and infrared scpectroscopy suggest that Na$^+$ ions exhibit
+negligible binding to PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
+binds to phosphate groups of two lipid molecules~\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
+Atomistic resolution molecular dynamics (MD) simulation models, however, predict significantly
 stronger binding for the cations than NMR experiments \cite{catte16}.
@@
 MD simulations~\cite{bockmann03,vacha09a}. 
-Interactions of Calcium ions with 3-4 lipids,
+Furthermore, interactions of Calcium ions with 3-4 lipids,
 including also interactions with carbonyl
@@
 Recent work published by the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
-made an attempt to resolve the apparent controversies. This study presented 
-a direct comparison between ion binding affinity between simulations and
-experiments by using NMR data for lipid headgroup order 
-parameters and the electrometer concept~\cite{seelig87}.
+made an attempt to resolve the apparent controversies. A direct comparison of
+ion binding affinities to PC bilayers was presented between simulations and
+experiments by using the electrometer concept, which is based on the experimental NMR data
+for the lipid headgroup order parameters~\cite{seelig87}.
 Using massive amounts of data collected by Open Collaboration method, it was concluded
@@
 in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
-improved by implicitly including electronic polarizability
+improved by implicitly including the electronic polarizability
 in the polar region of lipid molecules. The electronic
@@
 \subsection{Electrometer concept} \label{section:electrometer}
-Ion binding between experiments and simulations was compared by 
-using lipid head group order parameters and the
-''electrometer concept'' \cite{seelig87,catte16}.
-The concept is based on the experimental observation that the C-H bond
+Ion binding in lipid bilayers was compared between experiments and simulations by 
+using the lipid head group order parameters and the
+''electrometer concept'' \cite{seelig87,catte16}, which
+is based on the experimental observation that the C-H bond
 order parameters of $\alpha$ and $\beta$ carbons in PC lipid head group
 (see Fig.~\ref{simVSexpNOions}) are proportional to the amount 
-of unit charge bound per lipid,  $X^\pm$~\cite{seelig87}.
-Change in the order parameters measured with varying aqueous
+of unit charge bound per lipid~\cite{seelig87}.
+The change of order parameters measured with varying aqueous
 ion concentration can be then related to the amount of bound ions.
-
-The change of the head group order parameters is empirically quantified
-as~\cite{seelig87,ferreira16}
-\begin{equation}\label{OPchangeEQ}
-  \Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx \frac{4m_i }{3\chi}X^\pm,
-\end{equation}
-where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
-$i$ refers to either $\alpha$ or $\beta$ carbon,
-$m_i$ is an empirical constant depending on the valency and position of the bound charge,
-and the experimental value \cite{seelig77,davis83}, $\chi \approx$\,167\,kHz, is used for the quadrupole coupling constant.
-Atomic absorption spectra and $^2$H NMR data gave
-$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for Ca$^{2+}$ binding to POPC
-bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-The slopes are negative, because recent analysis concluded 
-that the order parameters decrease with bound positive charge 
-and increase with bound negative charge when the signs are taken 
-in account \cite{ollila16,catte16}. This is rationalized as a change
-of lipid head group dipole tilt toward water phase with bound positive
-charge and {\it vice versa} with negative charge \cite{seelig87}.
 
@@
 
-The measured order parameter change depends on the response of 
-the headgroup on the bound charge, i.e. $m_i$ in Eq.~\ref{OPchangeEQ},
-and the ion binding affinity. Thus, the former property has to be
-well quantified before using the electrometer concept to
-analyze binding affinities. This is done experimentally for a wide range
+The relation between bound charge per lipid,  $X^\pm$, and
+the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$,
+is empirically quantified as~\cite{seelig87,ferreira16}
+\begin{equation}\label{OPchangeEQ}
+  \Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx \frac{4m_i }{3\chi}X^\pm,
+\end{equation}
+where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
+$i$ refers to either $\alpha$ or $\beta$ carbon,
+$m_i$ is an empirical constant depending on the valency and position of the bound charge,
+and the experimental value \cite{seelig77,davis83}, $\chi \approx$\,167\,kHz, is used for the quadrupole coupling constant.
+Atomic absorption spectra and $^2$H NMR data gave
+$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for the binding of
+Ca$^{2+}$ to POPC bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
+The slopes are negative, because the head group order parameters
+decrease with bound positive charge and increase with bound negative
+charge~\cite{ollila16,catte16}. This can be rationalized as a change
+of the lipid head group dipole tilt toward water phase with the bound positive
+charge and {\it vice versa} with the negative charge~\cite{seelig87}.
+
+The measured order parameter change depends on both, the head group response 
+to the bound charge and the amount of bound charge, i.e. $m_i$ and $X^\pm$
+in Eq.~\ref{OPchangeEQ}, respectively. The former property has to be
+well quantified before the electrometer concept can used to 
+analyze the binding affinities, as done experimentally for a wide range
 of systems~\cite{seelig87,beschiasvili91}. To calibrate 
-the head group order parameter response also in simulations,
-we use experimental data for dihexadecyldimethylammonium (2C$_{16}$$^+$N2C$_1$Br$^-$)
-in a POPC bilayer \cite{scherer89}. Dihexadecyldimethylammonium
+the head group order parameter response to the bound charge in simulations,
+we use the experimental data for dihexadecyldimethylammoniumbromide %(2C$_{16}$$^+$N2C$_1$Br$^-$)
+mixed within a POPC bilayer~\cite{scherer89}. Dihexadecyldimethylammonium\\[0.5cm]
+%\vspace{0.5cm}
+\chemfig{
+ -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
+}  
+\vspace{0.5cm} \\
 is a cation surfactant having two acyl chains and bearing a unit
@@
 in the bilayer similarly to the phospholipids and the molar ratio then 
-gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems \cite{scherer87}.
+gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
 
@@
 An automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium \\[0.5cm]
-%\vspace{0.5cm}
-\chemfig{
- -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH2{)}_{15}-CH_3)(-[:30]{(}CH2{)}_{15}-CH_3)
-} , 
-\vspace{0.5cm} \\
-which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
+dihexadecyldimethylammonium, which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
 The AmberTools program \cite{amber} was then used to generate the Amber-type force field

### 2017-10-10 Manuscript/manuscript.tex
Update salt concentrations and binding affinity sections

@@
 similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
-\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules.
+\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules. \\
   JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer
   charges. So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever)
-  and ``semi-scaled'' models.}
+  and ``semi-scaled'' models. \\
+SAMULI: There is now a paragraph in the conclusions, which mentions this topic.}
 This decreases the solvation free energy by allowing water molecules to
@@
 \subsection{Salt concentrations and binding affinity}
-Experimental studies providing the lipid head group order
-parameters report salt concentrations in two different ways.
-For DPPC, salt concentrations before solvating
-the lipids are used~\cite{akutsu81}. Alternatively, for POPC,
-the salt concentration of supernatant after solvation is
-measured using atomic absorption~\cite{altenbach84}.
-\todo{JOE: I'm afraid that the above description might sound to the reader
-  like for some reasons one method is used with DPPC, but a different one with POPC.
-  I think simply writing one way is\dots another way is \dots (or similarly) will be less confusing.}
-In this work, we will use the latter definition. For this we
-will calculate the salt concentration in the aqueous bulk region,
-\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
+The early experimental electrometer concept data for the lipid head group
+order parameter changes was reported as a function 
+of the salt concentration in water before solvating
+the lipids~\cite{akutsu81}. The later study used atomic absorption
+spectroscopy and reported the order parameter changes as
+a function of the salt concentration in the supernatant after the
+solvation of lipids~\cite{altenbach84}.
+In this work we focus on POPC for which the latter definition was used~\cite{altenbach84}.
+The salt concentration in the aqueous bulk region was calculated from 
+the farthest point from both lipid leaflets in the water phase. 
 Note that in the previous study, Ref.~\citenum{catte16},
-the concentration was calculated as in the DPPC experiments.
+the ion concentrations were calculated in water before solvating the lipids 
+as in the earlier experiments~\cite{akutsu81}.
 Despite of the measurable differences between these two concentrations for CaCl$_2$ systems,
-the qualitative conclusions are affected neither in this
-nor in our previous work~\cite{catte16}.
-
-To quantify the ion binding affinity to a membrane, we
-calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
-This quantity can be used to compare binding affinity between different models,
-because it does not depend on the definition of the interfacial region
-or the position of the Gibbs dividing plane between two bulk regions.
-\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
-Therefore, we consider the hydrophobic interior of a membrane having zero concentrations of ions and water
-as a bulk region, which is separated by a monolayer interface from 
-a another region with bulk concentration of ions in water.
-Since the boundary of the latter region is set to the edge of simulation box,
-all molecules in simulation are within the defined interfacial region. 
+the qualitative conclusions in this or in the previous work~\cite{catte16} are
+not affected.
+
+To quantify the ion binding affinity to a lipid bilayer, we
+calculated the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
+This quantity was chosen because it does not depend on
+%the definition of the interfacial region or
+the position of the Gibbs dividing plane between two bulk regions.
+%\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
+%SAMULI: Here is a new trial.
+Here we assume that the interface locates between the hydrophobic interior
+of a lipid bilayer and the bulk water region far from the membrane.
+The bulk concentration of ions and water is zero inside the bilayer.
+The concentrations in bulk water region can be calculated from 
+the farthest point from both lipid leaflets in the water phase.
+The region between these boundaries contains all the ions and water
+molecules in the simulation box.
 %our whole simulation box as

### 2017-10-10 Manuscript/manuscript.tex
Updates and comments in methods section

@@
 
-Lipid bilayer structure without ions was validated against NMR and 
+The structure of lipid bilayers in simulations without ions were validated against NMR and 
 \mbox{x-ray} scattering experiments by calculating order parameters
-for C-H bonds and form factors from our simulations. The former validates the structures sampled by individual
+for C-H bonds and form factors. The former validates the structures sampled by individual
 lipid molecules in simulations with atomic resolution, while the latter
@@
 in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
-from equation
+from equation \todo{As Hector suggested, it might be better to write the simpler
+form for this equation.}
 \begin{equation}
@@
 \subsubsection{Simulations of POPC bilayers in aqueous ions}
-The simulated systems consisted of POPC bilayer in pure water
-or in varying salt concentrations.
-In particular, the periodic orthorhombic simulation box contained 
-128 POPC molecules and approximately 50~water molecules per each lipid.
+Simulations of POPC bilayer in pure water or in varying salt concentrations
+contained 128 POPC molecules and approximately 50~water molecules per each lipid
+in the periodic orthorhombic simulation box.
 As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
-which is currently the most accurate three site rigid water model. 
+which is currently the most accurate three site rigid water model
+\todo{More justification for the choosing the OPC3 water model are needed.
+It might be good to show the comparison with the scattering data in bulk water in SI.}. 
 In order to test transferability of our newly developed ECC-lipids model, 
@@
 OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
-presented in Supporting Information (SI). 
+\todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
 We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
@@
     \hline
-  \end{tabular}
+  \end{tabular} \\
+  \todo{Result with normal TIP3P missing?}
 \end{table}

### 2017-10-11 Manuscript/manuscript.tex
Update methods

@@
 \todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
-We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
+The ECC-ion models were used for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+Simulation with the Lipid14 model was also ran with the ion model by Dang et al.~\cite{smith94,chang1999,dang2006}.
+Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion model
+was taken directly from \cite{lipid14POPC0mMNaClfiles}. 
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
@@
 An automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium, which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
+dihexadecyldimethylammonium.%, which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
 The AmberTools program \cite{amber} was then used to generate the Amber-type force field
@@
 to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
+The parameters are available at \cite{??} \todo{To be added}.
 The surfactants were randomly placed among the lipids to form bilayer structures with
@@
 All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
-6, 14, 21, 35, or 50 surfactants per leaflet. 
-%Chloride counter ions were used
-%in simulations, because parameters for bromide, which was used in the experiment \cite{scherer89},
-%were not available in the standard Gromacs files for Amber force field.
-The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
-First 20~ns were omitted from the analysis.
+6, 14, 21, 35, or 50 surfactants per leaflet. Chloride counter ions were used
+in simulations, because parameters for bromide, the counterion in the experimental data \cite{scherer89},
+were not available in the standard Gromacs files for Amber force field.
+The Lipid14 model was used for POPC.
+The first 20~ns of the total simulation time of 200~ns
+was considered as an equlibration time and was omitted from the analysis.
+A reasonable lipid neighbor exchange occured during the simulation.
 
 The same systems were also simulated with the ECC-lipid model for POPC using the same setup.
+\todo{Which water model was used in these simulations?}
 In these simulations the ECC correction was also applied to the cationic surfactant 
@@
 
-%In order to validate the structural properties of the newly developed ECC-lipids model
-%we first compared our simulation results without any ions to
-%NMR order parameters measurements and x-ray scattering form factors 
-%(Fig.~\ref{simVSexpNOions}). 
 The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and

### 2017-10-11 Manuscript/manuscript.tex
Results and discussion updated

@@
   \caption{\label{simVSexpNOions}
-    Top: Chemical structure of POPC with definitions of different order parameters calculated.
-    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  
-    from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
-    compared with experimental order parameters from \cite{ferreira13}.
-    Error bars are approximately the size of the points for head group order parameters.
-    Point size for the order parameters of tails are decreased by a factor of 3 to improve the clarity of the plot.
-    Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
-    } 
+    Top: The chemical structure of POPC and the labeling of the carbon segments.
+    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
+    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipid models
+    compared with the experimental values from \cite{ferreira13}.
+    The size of the points for the head group order parameters correspond the
+    of the error bars ($\pm 0.02$ for experiments \cite{botan15,ollila16} and ?? for simulations).
+    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
+    Bottom: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
+    the ECC-lipid models compared with experiments~\cite{Kucerka2011}.
+  } 
+  \todo{Increase the height of the order parameter figure} \\
+  \todo{sn-1 and sn-2 order parameters should be somehow labelled in the order parameter figure. Maybe empty and filled points?} \\
+  \todo{Add size of the error bars in simulations to the caption.} \\
+  \todo{Would it be possible to increase the size of points for acyl chain order parameters only in z-direction such that
+    it would correspond the error bars?} \\
+  \todo{x-axis in form factor plot from 0 to 0.6, where experimental data ends. y-axis from 0 to $\sim$2.5 to remove the empty space.} \\
 \end{figure}
@@
 
-The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and
-the area per lipid (Table~\ref{tab:apls}) of POPC bilayer simulated
-with the ECC-lipid model are in good agreement with the experimental results,
-as well as with the Lipid14 model. Thus, we conclude that the lipid bilayer dimensions
-are well captured by the developed ECC-lipid model. 
-Areas per lipid of ECC-lipids with various other water models are documented in SI, table~\ref{tab:apls_si}. 
+The ECC-lipid and Lipid14 models reproduce the experimental x-ray scattering form factors
+of POPC bilayer with the similar accuracy in Fig.~\ref{simVSexpNOions}.
+The area per lipid values from the Lipid14 model is slightly larger than the
+experimental value in Table~\ref{tab:apls}, while the value from ECC-lipid model
+is slightly smaller. Area per lipid values of the ECC-lipid model show some variation
+when simulated with different water models in Table~\ref{tab:apls_si} (Supplementary Information),
+however, all the values are close to the experimentally reported values.
+In conclusion, the ECC-lipid model reproduces the experimental dimensions of POPC
+lipid bilayer with the accuracy comparable to the other state of the art lipid models \cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 
-As in the original Lipid14 model~\cite{dickson14}, the acyl chain order parameters of the
-ECC-lipid model agree well with the experimental values in Fig.~\ref{simVSexpNOions}. Notably, the experimentally measured forking and
+The acyl chain order parameters of the Lipid14 model~\cite{dickson14} and the
+ECC-lipid model agree with the experimental values with in the error bars
+in Fig.~\ref{simVSexpNOions}, althought the ECC-lipid model gives slightly larger
+values for {\it sn}-1 chain. Notably, the experimentally measured forking and
 small order parameter values of $C_2$ segment in {\it sn}-2 chain are relatively well
 reproduced by the both models. This has been suggested to indicate that the carbonyl
-region of {\it sn}-2 chain is directed towards the water phase, in contrast to the
+of {\it sn}-2 chain is directed towards the water phase, in contrast to the
 carbonyl in {\it sn}-1 chain, which would orient more along the bilayer
@@
 
-The headgroup order parameters of $\alpha$ and $\beta$ carbons are slightly larger
-in the ECC-lipid model than in Lipid14, which is apparently related to the
-smaller P-N vector angle with respect to the membrane normal in
-Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
+The order parameters of $\alpha$ and $\beta$ carbons in the headgroup are slightly larger
+in the ECC-lipid model than in the Lipid14 model, which is apparently related to the
+P-N vector orienting 7$^{\circ}$ more toward the water phase in the ECC-lipid model, 
+see Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
 however, conclude which one of the models give the more realistic
@@
 the $\beta$ carbon order parameter value closer to experiments, while
-value for $\alpha$ carbon is better in Lipid14.
+value for $\alpha$ carbon is better in the Lipid14 model.
 Despite of some deviations from the experimental order parameter values
@@
 
-\subsection{Response of POPC head groups to bound charge}
+\subsection{Response of POPC head groups to bound charge}\label{section:boundCHARGE}
 
@@
   \caption{\label{OrderParameterCHANGESsurf}
-    Headgroup order parameter changes and P-N vector orientation as a function of
-    cationic surfactant (dihexadecyldimethylammonium bromide, 2C$_{16}$$^+$N2C$_1$Br$^-$)
-    in PC bilayer from simulations and experiments \cite{scherer89}.
-  }
+    The changes of headgroup order parameters and P-N vector orientation as a function of
+    cationic surfactant (dihexadecyldimethylammonium) in POPC bilayer from simulations
+    and experiments \cite{scherer89}.
+  }
+  \todo{Labeling should be consistent with the previous figure, i.e. experimental data with diamonds.} \\
+  \todo{x-axis scale from -1 to 51 to make the point in zero fully visible and to remove empty space.} \\
+  \todo{Empty space between figures and from the right column could be reduced.}
 \end{figure}
@@
 Before proceeding to the ion binding affinity, we quantify
-the response of headgroup order parameters to the amount of 
-bound charge by using mixtures of monovalent cationic surfactants
+the response of the headgroup order parameters to the amount of 
+bound charge by using mixtures of monovalent cationic
+surfactants (dihexadecyldimethylammonium)
 and POPC~\cite{scherer89}. The amount of bound charge per PC 
 in these systems is given by the molar fraction of cationic 
-surfactants, because essentially all surfactants locate in the 
-lipid bilayers. Experimental data for these systems can be used to validate 
+surfactants, because essentially all surfactants with two hydrophopic
+acyl chains can be assumed to locate in the lipid bilayers.
+The experimental data for these systems can be used to validate 
 the sensitivity of lipid headgroup order parameters
@@
 
-%depends on both, the
-%binding affinity and the response of the head group on bound charge.
-%Thus, the head group response to bound charge has to be correctly
-%described in the model if the concept is used for quantitative studies of
-%binding affinity. Hence, we first quantify the response of head group order
-%parameters to the bound charge by using experimental data measured from
-%mixtures of cationinc surfactants and PC lipids~\cite{scherer89}.
-%
-%Thus, the head group order parameter response to bound charge can be first
-%quantified aginst experimental data before proceeding to binding affinity studies.
-%Headgroup order parameter response to the amount of bound charge can
-%be quantified by using ionic surfactants with a known valency~\cite{scherer89}.
-%All surfactants locate in lipid bilayers due to their amphiphilic nature,
-%thus the exact amount of bound charge in the membrane is known in these systems.
-%The head group order parameter changes as a function of bound charges can be then
-%explicitly determined from such systems.
-
-The headgroup order parameter changes with an increasing amount of 
-the cationic surfactant dihexadecyldimethylammonium bromide 
-is compared between experiments~\cite{scherer89} and simulations in
-Fig. \ref{OrderParameterCHANGESsurf}. 
-The observed order parameter decrease in simulations and experiments
-can be approximated to be linear at least for mole fractions
-below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
-The slope is, however, too steep in the Lipid14 model indicating that 
-the head group order parameters are too sensitive to a bound charge.
-The ECC-lipids model gives a slope in very good agreement with experiments
+The changes of the headgroup order parameters with an increasing amount of 
+the cationic surfactant is compared between experiments~\cite{scherer89} and
+simulations in Fig. \ref{OrderParameterCHANGESsurf}.
+Approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ},
+is observed in simulations and experiments at least for the mole fractions
+below $\sim$30\%. The slope is, however, too steep in the Lipid14 model indicating that 
+the head group order parameters respond too sensitively to the bound positive charge.
+The slope in the ECC-lipid model is in very good agreement with experiments
 for the $\alpha$ segment, while the slope is slightly
 underestimated for the $\beta$ segment.
-%This has to be taken into account when interpreting the
-%head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
-%observed in  Ref. \citenum{catte16}, see also Fig. \ref{fig:delta_ordPar_NaCl}.
-%The present comparison of head group order parameter changes 
-%suggests that the observed overestimated response of all models to 
-%increasing CaCl$_2$ concentration in \cite{catte16} 
-%can be explained at least in part by a high sensitivity of the head group order parameter response
-%to the bound charge.
-\todo{SAMULI: We could calculate the slopes from simulations, but I am not sure if
-we would actually learn anything useful from this.}
 
 The headgroup P-N vector angle with respect to the membrane normal
-is also shown as a function of cationic surfactant mole fraction
-in Fig.~\ref{OrderParameterCHANGESsurf}.
-The headgroup orients more towards the water phase with an increasing
-amount of bound cations, as previously reported in Ref.~\citenum{seelig87}. 
-The effect is more pronounced in Lipid14 than 
-in ECC-lipids model, which is in line with the order parameter 
-results and the reduced charge--dipole interactions in the latter model.
-The response of $\alpha$-order parameter to bound positive charge
-in ECC-lipid model is in good agreement with experiments.
-The model can be thus used to study changes of lipid P-N vector
-in varying conditions.
+is also shown in Fig.~\ref{OrderParameterCHANGESsurf}
+as a function of the mole fraction of the cationic surfactant.
+As suggested previously~\cite{seelig87}, the headgroup orients
+more towards the water phase with the increasing amount of positive
+charge in a PC lipid bilayer. The effect is more pronounced in the
+Lipid14 model, for which the addition of 50\% mole fraction of the cationic
+surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle,
+while the corresponding change in the ECC-lipid model is 11$^{\circ}$.
+The difference is in line with the smaller order parameter 
+changes and the reduced charge--dipole interactions in the latter model.
+The lesser sensitivity of the P-N vector angle response in the ECC-lipid
+model can be considered to be more realistic, because the changes of the headgroup
+order parameters as a function of the bound positive charge
+are in better agreement with experiments in this model.
+The results also imply that the validation and improvements of MD simulation
+models are generally required for the reliable simulation studies of the
+lipid headgroup responses to ions or other biomolecules, as also concluded
+previously \cite{botan15}.
 
@@
   }
+  \todo{I think that we must show the NaCl results in the main paper. I think that the best would
+    be to use similar two column format, which we had before. This is used also in NMRlipids II and makes the
+    comparison easier} \\
+  \todo{The DPPC and POPC should be shown and properly labelled as we had previously.} \\
+  \todo{The labeling should be consistent with previous figures, i.e., experiments with diamonds.} \\
+  \todo{Start x-axis scale from -1 to make the point in zero fully visible.} \\
+  \todo{Start y-axis scale of the two top figures from -0.13 to remove the empty space.} \\
+  \todo{y-axis scale of the bottom figure from 59 to 78 to remove the empty space.} \\
+  \todo{Empty space between figures and from the right column could be reduced.} \\
+  \todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
+    the value of $g_3$ order parameter of DPPC was -0.214 
+    in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
+    of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
+    at the C-2 segments of both fatty acyl chains.''
+  }
 \end{figure}
 
-The binding affinities of aqueous cations to lipid 
-bilayers can be measured using the headgroup 
-order parameters, because they decrease proportionally  
-to the bound positive charge~\cite{seelig87,catte16}. 
-The headgroup order parameter responses 
-to increasing aqueous CaCl$_2$ concentrations 
-from experiments (DPPC \cite{akutsu81} 
-and POPC \cite{altenbach84}) and different simulation models for POPC 
-are shown in Fig.~\ref{fig:delta_ordPar_CaCl}.
-Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.
-
-%The simulation results of Lipid14 with \AA{}qvist ions~\cite{aqvist90} (from Ref. \citenum{catte16})
-%are shown together with the results from Lipid14 simulation
-%with ions by Dang~\cite{smith94,chang1999,dang2006} and from ECC-lipids simulations with ECC-ions.
-
-Negligible changes of the headgroup order parameters are measured
-with submolar concentrations of NaCl due to the very low affinity of Na$^+$
-in PC bilayers \cite{akutsu81}. While Na$^+$ binding and the related 
-headgroup order parameter changes were overestimated in
-almost all the available simulation models, the low affinity
-and negligible order parameter changes were reproduced by Lipid14 
-model when simulated with \AA{}qvist ions~\cite{catte16}.
-However, the same combination of force field parameters
-overestimated the headgroup order parameter response to CaCl$_2$ 
-concentration, which was the case also in all other models tested in
-Ref.~\citenum{catte16}.
-Using the ion model by Dang et al.~\cite{smith94,chang1999,dang2006} or
-ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
-with more realistic bulk behaviour did not improve the results
-for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model, as seen 
-in Figs. \ref{fig:delta_ordPar_CaCl} and S\ref{fig:ordPars_waterModels} (in SI), respectively.
+The changes of the lipid bilayer headgroup order parameter from different simulations and
+experiments \cite{akutsu81,altenbach84} are shown in Fig.~\ref{fig:delta_ordPar_CaCl}
+as a function of NaCl and CaCl$_2$ concentrations.
+%Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.
+These results can be used to compare the ion binding affinities to lipid bilayers
+between simulations and experiments by using the electrometer concept, because 
+the order parameters decrease proportionally to the amount of bound positive charge~\cite{seelig87,catte16}. 
+The recent comparison of different simulation models to the experimental data
+revealed that most models significantly overestimate the Na$^+$ ion binding to PC
+lipid bilayers and that none of the available models correctly reproduce
+the details of the binding of Ca$^{2+}$ ions \cite{catte16}.
+A positive exception was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions,
+as seen from the data replotted from Ref.~\citenum{catte16} in Fig.~\ref{fig:delta_ordPar_CaCl}.
+The model reproduced the experimentally measured small order parameter changes with NaCl
+and imperceptible Na$^+$ binding to PC bilayers \cite{akutsu81,altenbach84}.
+However, the changes of the headgroup order parameters as a function of CaCl$_2$ 
+concentration were overestimated by the same combination of force field parameters.
+
+Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{??},
+we also simulated the Lipid14 model with the ion models by Dang et al.~\cite{smith94,chang1999,dang2006} and
+ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} having
+more realistic bulk behaviour. Instead of the improvement in the binding behaviour,
+we observed overbinding also of Na$^+$ with these ions models, while the results
+with Ca$^{2+}$ were barely affected,
+%for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model,
+as seen in Figs. \ref{fig:delta_ordPar_CaCl} and S??.%\ref{fig:ordPars_waterModels} (in SI), respectively.
 \todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
-The results support the conclusion of the previous work \cite{catte16}
-that improvements also in lipid models are needed to 
-correctly describe divalent cation binding to PC bilayers.
-
-Significant improvement can be achieved using the ECC
-approach also for lipids. The headgroup order parameters change
-as a function of CaCl$_2$ concentration for the ECC-lipid model with ECC-ions
-exhibit a good agreement with experiments in Fig.~\ref{fig:delta_ordPar_CaCl} and S\ref{fig:delta_ordPar_NaCl}.
-As discussed in previous section, the model gives also
-a good agreement with experiments for
-the headgroup response to bound charge.
-%Thus, the model can used for more detailed analysis of the binding affinity.
-
-The binding affinities are quantified using
-the water density profiles along the membrane normal shown
-in Fig.~\ref{fig:cacl-dens}. The density profiles show
-a larger Ca$^{2+}$ density peak in lipid headgroup region for
-the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
-The relative surface excess calculated from Eq.~\ref{surfexcess} is 
-$\Gamma_i^w = 0.07 \pm 0.01 nm^{-2}$ for the ECC-lipid model,
-which is significantly smaller than $\Gamma_i^w = 0.13 \pm 0.01 nm^{-2}$ for
-Lipid14 with \AA{}qvist and $\Gamma_i^w = 0.3 \pm 0.03 nm^{-2}$ with Dang ions.
-
-\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
-  do this analysis by using the density profiles, but it is not necessary.
-  The reason why I want to do something like this, is that it would be useful to have
-  some easier way to estimate the correct binding affinity than the electrometer concept,
-  which is quite tedious to apply in practise (simulations with different concenrations,
-  cationic surfactant check, etc.). I wuold like to be able to estimate much faster from a
-  simulation if the binding affinity is reasonable or not. This may not be the correct way for
-that anyway.}
-Rough estimates for the free energy difference between bound and unbound cation
-are given by
-\begin{equation}
-  \Delta G = -{\rm k_b T}\log(\frac{p_i}{p_o}),
-\end{equation}
-where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
-in the maximum density in bilayer, respectively. The density profiles in
-Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
-between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
-in Lipid14 with \AA{}qvist.
-
-\todo{SAMULI: Maybe we should discuss the repeat distances and
-  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
-  }
+The results are in line with the previous work \cite{catte16},
+suggesting that the improvements in the lipid parameters are
+required to correctly describe the divalent cation binding to PC lipid bilayers.
+
+To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
+polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
+The ECC-lipid model simulations with the ECC-ion models \cite{??}
+exhibit a good agreement with the experimental changes of 
+the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations
+in Fig.~\ref{fig:delta_ordPar_CaCl}. Since also the headgroup order parameter
+response the to the bound positive charge the ECC-lipid model
+was in good agreement with experiments in section \ref{section:boundCHARGE},
+we conclude that the model correctly reproduces the binding affinity of
+Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.
+
+The density distributions of water and ions along the membrane normal
+are shown in Fig.~\ref{fig:cacl-dens}.
+\todo{Densities and relative surface excess from NaCl systems to be added.
+  The discussion below to be finished after this.}
+%The density profiles show
+%a larger Ca$^{2+}$ density peak in lipid headgroup region for
+%the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
+To quantify the binding affinities in different models we calculated 
+the relative surface excess from Eq.~\ref{surfexcess}. The value for Calcium
+in the ECC-lipid model, $\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$, is significantly
+smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
+or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
+
+%\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
+%  do this analysis by using the density profiles, but it is not necessary.
+%  The reason why I want to do something like this, is that it would be useful to have
+%  some easier way to estimate the correct binding affinity than the electrometer concept,
+%  which is quite tedious to apply in practise (simulations with different concenrations,
+%  cationic surfactant check, etc.). I wuold like to be able to estimate much faster from a
+%  simulation if the binding affinity is reasonable or not. This may not be the correct way for
+%that anyway.}
+%Rough estimates for the free energy difference between bound and unbound cation
+%are given by
+%\begin{equation}
+%  \Delta G = -{\rm k_b T}\log(\frac{p_i}{p_o}),
+%\end{equation}
+%where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
+%in the maximum density in bilayer, respectively. The density profiles in
+%Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
+%between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
+%in Lipid14 with \AA{}qvist.
 
@@
 
-Since the lipid headgroup order parameter responds to the amount of bound charge
-and to the aqueous ion concentrations are both in good agreement with experiments
-in the ECC-lipid model with ECC-ion parameters, we consider the Na$^+$ and Ca$^{2+}$
-binding affinities to be realisticlly described in this model. In contrast, the Ca$^{2+}$
-binding affinity is overestimated by Lipid14 model when simulated with all the
-tested ion models. Similar conclusions were previously made based only on the headrogup
-order parameter data with aqueous cations~\cite{catte16}. However,
-the discrepancies with experiments in previous work could partly arise also from
-the inaccurately described sensitivity of the headgroup to bound charge. Here we quantify this
-effect and conclude that the improvement due to ECC is to a large extent, but not entirely,
-caused by a more realistic headgroup sensitivity to bound charge.
-This indicates that the issue should be carefully considered also when
-the electrometer concept is used to compare ion binding between experiments
-and simulations with other models.
-
-%The good agreement with experiments indicates
-%that the ECC corrected lipid model has a sufficient accuracy for realistic
-%studies of details of ions binding to lipid bilayers.
+In conclusion, the results suggest that 
+the ECC-lipid model together with the ECC-ion parameters gives a realistic
+description for the binding affinities of Na$^+$ and Ca$^{2+}$ ions to POPC lipid
+bilayer. Furthermore, the overestimated headgroup order parameter changes of POPC in the Lipid14
+model as a function of CaCl$_2$ concentration arise from both, the overestimated
+binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
+This probably applies also to the other lipid models considered in the previous study~\cite{catte16},
+emphasizing the importance of the comparison of the lipid headgroup order parameter response
+to the bound charge between simulations and experiments as done in section \ref{section:boundCHARGE}.
+
+\todo{SAMULI: Maybe we should discuss the repeat distances and
+  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
+  }
 

### 2017-10-11 Manuscript/manuscript.tex
Roughly skeched analysis with residence times of cations in stoichimetry section.

--- 2017-10-10 Manuscript/manuscript.tex
@@
 such as in signal propagation in neurons and vesicle fusion.
-Since the direct measurements of ion-membrane interactions from biological systems
+Since direct measurements of ion-membrane interactions from biological systems
 are difficult, lipid bilayers are often used as models systems for cellular
-membranes. Especially the zwitterionic phosphocholine (PC) lipid bilayers are
-often used as model systems to understand the role of ions in complex biological
-systems~\cite{scherer87,seelig90,cevc90}.
-
-Interactions of biological cations, especially Na$^+$ and
+membranes. The detailed results from simple model systems can be then
+used to understand the role of ions in complex biological systems.
+
+Zwitterionic phosphocholine (PC) lipid bilayers are commonly used model systems for
+cellular membranes. Interactions of biological cations, especially Na$^+$ and
 Ca$^{2+}$, with PC bilayers are widely studied in
@@
 %ions, but do not affect the specific binding constant \cite{seelig90}.
-Interpretations of non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
-and infrared scpectroscopy suggest that Na$^+$ ions exhibit
-negligible binding to PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
-binds to phosphate groups of two lipid molecules~\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
-Atomistic resolution molecular dynamics (MD) simulation models, however, predict significantly
+Non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
+and infrared scpectroscopy, give accurate information about ion binding
+in PC lipid bilayers\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
+Interpretation most of these experiments suggests that Na$^+$ ions exhibit
+negligible binding in PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
+binds to phosphate groups of two lipid molecules. Atomistic resolution
+molecular dynamics (MD) simulation models, however, predict significantly
 stronger binding for the cations than NMR experiments \cite{catte16}.
@@
 MD simulations~\cite{bockmann03,vacha09a}. 
-Furthermore, interactions of Calcium ions with 3-4 lipids,
+Interactions of Calcium ions with 3-4 lipids,
 including also interactions with carbonyl
@@
 Recent work published by the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
-made an attempt to resolve the apparent controversies. A direct comparison of
-ion binding affinities to PC bilayers was presented between simulations and
-experiments by using the electrometer concept, which is based on the experimental NMR data
-for the lipid headgroup order parameters~\cite{seelig87}.
+made an attempt to resolve the apparent controversies. This study presented 
+a direct comparison between ion binding affinity between simulations and
+experiments by using NMR data for lipid headgroup order 
+parameters and the electrometer concept~\cite{seelig87}.
 Using massive amounts of data collected by Open Collaboration method, it was concluded
@@
 in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
-improved by implicitly including the electronic polarizability
+improved by implicitly including electronic polarizability
 in the polar region of lipid molecules. The electronic
@@
 similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
-\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules. \\
+\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules.
   JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer
   charges. So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever)
-  and ``semi-scaled'' models. \\
-SAMULI: There is now a paragraph in the conclusions, which mentions this topic.}
+  and ``semi-scaled'' models.}
 This decreases the solvation free energy by allowing water molecules to
@@
 \subsection{Electrometer concept} \label{section:electrometer}
-Ion binding in lipid bilayers was compared between experiments and simulations by 
-using the lipid head group order parameters and the
-''electrometer concept'' \cite{seelig87,catte16}, which
-is based on the experimental observation that the C-H bond
+Ion binding between experiments and simulations was compared by 
+using lipid head group order parameters and the
+''electrometer concept'' \cite{seelig87,catte16}.
+The concept is based on the experimental observation that the C-H bond
 order parameters of $\alpha$ and $\beta$ carbons in PC lipid head group
 (see Fig.~\ref{simVSexpNOions}) are proportional to the amount 
-of unit charge bound per lipid~\cite{seelig87}.
-The change of order parameters measured with varying aqueous
+of unit charge bound per lipid,  $X^\pm$~\cite{seelig87}.
+Change in the order parameters measured with varying aqueous
 ion concentration can be then related to the amount of bound ions.
+
+The change of the head group order parameters is empirically quantified
+as~\cite{seelig87,ferreira16}
+\begin{equation}\label{OPchangeEQ}
+  \Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx \frac{4m_i }{3\chi}X^\pm,
+\end{equation}
+where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
+$i$ refers to either $\alpha$ or $\beta$ carbon,
+$m_i$ is an empirical constant depending on the valency and position of the bound charge,
+and the experimental value \cite{seelig77,davis83}, $\chi \approx$\,167\,kHz, is used for the quadrupole coupling constant.
+Atomic absorption spectra and $^2$H NMR data gave
+$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for Ca$^{2+}$ binding to POPC
+bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
+The slopes are negative, because recent analysis concluded 
+that the order parameters decrease with bound positive charge 
+and increase with bound negative charge when the signs are taken 
+in account \cite{ollila16,catte16}. This is rationalized as a change
+of lipid head group dipole tilt toward water phase with bound positive
+charge and {\it vice versa} with negative charge \cite{seelig87}.
 
@@
 
-The relation between bound charge per lipid,  $X^\pm$, and
-the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$,
-is empirically quantified as~\cite{seelig87,ferreira16}
-\begin{equation}\label{OPchangeEQ}
-  \Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx \frac{4m_i }{3\chi}X^\pm,
-\end{equation}
-where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
-$i$ refers to either $\alpha$ or $\beta$ carbon,
-$m_i$ is an empirical constant depending on the valency and position of the bound charge,
-and the experimental value \cite{seelig77,davis83}, $\chi \approx$\,167\,kHz, is used for the quadrupole coupling constant.
-Atomic absorption spectra and $^2$H NMR data gave
-$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for the binding of
-Ca$^{2+}$ to POPC bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-The slopes are negative, because the head group order parameters
-decrease with bound positive charge and increase with bound negative
-charge~\cite{ollila16,catte16}. This can be rationalized as a change
-of the lipid head group dipole tilt toward water phase with the bound positive
-charge and {\it vice versa} with the negative charge~\cite{seelig87}.
-
-The measured order parameter change depends on both, the head group response 
-to the bound charge and the amount of bound charge, i.e. $m_i$ and $X^\pm$
-in Eq.~\ref{OPchangeEQ}, respectively. The former property has to be
-well quantified before the electrometer concept can used to 
-analyze the binding affinities, as done experimentally for a wide range
+The measured order parameter change depends on the response of 
+the headgroup on the bound charge, i.e. $m_i$ in Eq.~\ref{OPchangeEQ},
+and the ion binding affinity. Thus, the former property has to be
+well quantified before using the electrometer concept to
+analyze binding affinities. This is done experimentally for a wide range
 of systems~\cite{seelig87,beschiasvili91}. To calibrate 
-the head group order parameter response to the bound charge in simulations,
-we use the experimental data for dihexadecyldimethylammoniumbromide %(2C$_{16}$$^+$N2C$_1$Br$^-$)
-mixed within a POPC bilayer~\cite{scherer89}. Dihexadecyldimethylammonium\\[0.5cm]
-%\vspace{0.5cm}
-\chemfig{
- -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
-}  
-\vspace{0.5cm} \\
+the head group order parameter response also in simulations,
+we use experimental data for dihexadecyldimethylammonium (2C$_{16}$$^+$N2C$_1$Br$^-$)
+in a POPC bilayer \cite{scherer89}. Dihexadecyldimethylammonium
 is a cation surfactant having two acyl chains and bearing a unit
@@
 in the bilayer similarly to the phospholipids and the molar ratio then 
-gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
+gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems \cite{scherer87}.
 
 \subsection{Salt concentrations and binding affinity}
-The early experimental electrometer concept data for the lipid head group
-order parameter changes was reported as a function 
-of the salt concentration in water before solvating
-the lipids~\cite{akutsu81}. The later study used atomic absorption
-spectroscopy and reported the order parameter changes as
-a function of the salt concentration in the supernatant after the
-solvation of lipids~\cite{altenbach84}.
-In this work we focus on POPC for which the latter definition was used~\cite{altenbach84}.
-The salt concentration in the aqueous bulk region was calculated from 
-the farthest point from both lipid leaflets in the water phase. 
+Experimental studies providing the lipid head group order
+parameters report salt concentrations in two different ways.
+For DPPC, salt concentrations before solvating
+the lipids are used~\cite{akutsu81}. Alternatively, for POPC,
+the salt concentration of supernatant after solvation is
+measured using atomic absorption~\cite{altenbach84}.
+\todo{JOE: I'm afraid that the above description might sound to the reader
+  like for some reasons one method is used with DPPC, but a different one with POPC.
+  I think simply writing one way is\dots another way is \dots (or similarly) will be less confusing.}
+In this work, we will use the latter definition. For this we
+will calculate the salt concentration in the aqueous bulk region,
+\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
 Note that in the previous study, Ref.~\citenum{catte16},
-the ion concentrations were calculated in water before solvating the lipids 
-as in the earlier experiments~\cite{akutsu81}.
+the concentration was calculated as in the DPPC experiments.
 Despite of the measurable differences between these two concentrations for CaCl$_2$ systems,
-the qualitative conclusions in this or in the previous work~\cite{catte16} are
-not affected.
-
-To quantify the ion binding affinity to a lipid bilayer, we
-calculated the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
-This quantity was chosen because it does not depend on
-%the definition of the interfacial region or
-the position of the Gibbs dividing plane between two bulk regions.
-%\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
-%SAMULI: Here is a new trial.
-Here we assume that the interface locates between the hydrophobic interior
-of a lipid bilayer and the bulk water region far from the membrane.
-The bulk concentration of ions and water is zero inside the bilayer.
-The concentrations in bulk water region can be calculated from 
-the farthest point from both lipid leaflets in the water phase.
-The region between these boundaries contains all the ions and water
-molecules in the simulation box.
+the qualitative conclusions are affected neither in this
+nor in our previous work~\cite{catte16}.
+
+To quantify the ion binding affinity to a membrane, we
+calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
+This quantity can be used to compare binding affinity between different models,
+because it does not depend on the definition of the interfacial region
+or the position of the Gibbs dividing plane between two bulk regions.
+\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
+Therefore, we consider the hydrophobic interior of a membrane having zero concentrations of ions and water
+as a bulk region, which is separated by a monolayer interface from 
+a another region with bulk concentration of ions in water.
+Since the boundary of the latter region is set to the edge of simulation box,
+all molecules in simulation are within the defined interfacial region. 
 %our whole simulation box as
@@
 
-The structure of lipid bilayers in simulations without ions were validated against NMR and 
+Lipid bilayer structure without ions was validated against NMR and 
 \mbox{x-ray} scattering experiments by calculating order parameters
-for C-H bonds and form factors. The former validates the structures sampled by individual
+for C-H bonds and form factors from our simulations. The former validates the structures sampled by individual
 lipid molecules in simulations with atomic resolution, while the latter
@@
 in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
-from equation \todo{As Hector suggested, it might be better to write the simpler
-form for this equation.}
+from equation
 \begin{equation}
@@
 \subsubsection{Simulations of POPC bilayers in aqueous ions}
-Simulations of POPC bilayer in pure water or in varying salt concentrations
-contained 128 POPC molecules and approximately 50~water molecules per each lipid
-in the periodic orthorhombic simulation box.
+The simulated systems consisted of POPC bilayer in pure water
+or in varying salt concentrations.
+In particular, the periodic orthorhombic simulation box contained 
+128 POPC molecules and approximately 50~water molecules per each lipid.
 As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
-which is currently the most accurate three site rigid water model
-\todo{More justification for the choosing the OPC3 water model are needed.
-It might be good to show the comparison with the scattering data in bulk water in SI.}. 
+which is currently the most accurate three site rigid water model. 
 In order to test transferability of our newly developed ECC-lipids model, 
@@
 OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
-\todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
+presented in Supporting Information (SI). 
 We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
@@
 An automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium, which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
+dihexadecyldimethylammonium \\[0.5cm]
+%\vspace{0.5cm}
+\chemfig{
+ -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH2{)}_{15}-CH_3)(-[:30]{(}CH2{)}_{15}-CH_3)
+} , 
+\vspace{0.5cm} \\
+which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
 The AmberTools program \cite{amber} was then used to generate the Amber-type force field
@@
 to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
-Such a finding is supported by the calculated
-number of contacts between \ce{Ca^{2+}} and POPC oxygen atoms.
-The average number of contacts between \ce{Ca^{2+}} and any oxygen atom in POPC is 59.8, 
+Such a finding is based on the calculated
+number of contacts per lipid between \ce{Ca^{2+}} and POPC oxygen atoms, 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
+in which a contact between the two groups is counted 
+if their distance is less than or equal $0.3 \, \mathrm{nm}$ (robustness of this definition discussed in reference~\citenum{javanainen17}). 
+the average number of contacts between \ce{ca^{2+}} and any oxygen atom in popc is 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.47 $,    % = 59.8/128 
 whereas if only phosphate oxygen atoms are considered 
-the average number of contacts decreases only by a tiny amount to 57.2. 
+the average number of contacts decreases only by a tiny amount to 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.45 $,    % = 57.2/128 . 
+and if only carbony oxygen atoms are considered,
+this value is merely
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.02 $.    % = (59.8-57.2)/128 . 
 We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
@@
 is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
-From recent theoretical work with long enough simulations this time can be roughly estimated
+From a recent theoretical work with long simulations 
+(Fig. S17 in \citenum{javanainen17} -- C36 with MK Ca2+, any other model, either different cation or Slipids/C36, just such calcium)
+this time can be roughly estimated -- get the simulation from \url{https://doi.org/10.5281/zenodo.259376} and analyze it -- 
 in the order of $1$--$10\,\mu\mathrm{s}$~\cite{javanainen17}. 
 This is in contrast to our model, ECC-lipids, 
-which gives a mean residence time in the order of $10\,\mathrm{ns}$, 
-\todo{evaluate this number, mean residence time, accurately based on the contacts data.}
+which estimates the residence time of calcium bound to a POPC membrane to be 
+smaller than $53\,\mathrm{ns}$ for 90\% of the time a cation is in a membrane-bound state. 
 i.e. at least two orders of magnitude lower than previous estimates.
+In addition, the longest observed residence time in our simulation is $120\,\mathrm{ns}$, 
+which is well below the length of our simulations.
+This is unfortunately not the case of many theoretical works,
+where the residence time cannot be estimated,
+or the simulation has not even decorrelated from it's 
+initial condition with a population of bound/unbound cations \cite{catte16}, 
+because it is limited by the simulation length, 
+even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
+Increasing time scales to the order of $\mu\mathrm{s}$ 
+doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used. 
 Such a finding changes the point of view of calcium binding from
@@
 in equilibrium between membrane and solvent. 
-
-\todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},
-  their interaction energies with the lipid membrane, etc,
-  and finalize the discussion after these results.}
-
-%It is also suggested that the addition of \ce{NaCl} to the solution of \ce{CaCl_2} enhances the hedgroup 
-%order parameter response compared to the solution with only \ce{CaCl_2}. \cite{altenbach84}
-%\todo{Simulate this effect and discuss it further}
-
-%\todo{The difference between DPPC and POPC -- simulate and compare with experiment. }
-
-%\todo{It might be worth acknowledging each experimental finding in \cite{Altenbach84} and observations in \cite{Javanainen2017-ChemComm}.}
+A histogram of residence times from our simulations is in Fig.~S\ref{fig:hist_residence_times} in SI. 
+\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
+Add data for Na+. }
 
@@
 \section{Conclusions}
-By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
-to a POPC lipid bilayer can be accurately described with the classical MD simulation 
-force field, where the electronic polarization is implicitly included using 
-the electronic continuum correction (ECC)~\cite{leontyev11}.
-The proposed ECC-lipids model is a significant improvement over 
-other available lipid models, which all overestimate specific cation binding affinities~\cite{catte16}.  
-While the structural details of a POPC lipid bilayer simulated with the ECC-lipids
-model agree with experiments with the comparable accuracy to the other state of the art lipid models,
-it also reproduces the experimental lipid head group order parameter responses to
-the cationic surfactant, NaCl and CaCl$_2$ concentrations.
-
-The good agreement with experiments enables the atomic resolution 
-interpretation of NMR experiments by using MD simulations.
-In agreement with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
-the Ca$^{2+}$ ions mainly interact with phosphate oxygens.
-However, the stoichiometry of the binding is significantly more complicated 
-than in the previous interpretation of the NMR data based on 
-the ternary complex model, where one calcium binds to two POPC molecules~\cite{altenbach84}.
-The complexes of one calcium bound to two lipids are the most probable also in the
-ECC-lipids model, but the complexes of one or three lipids per one calcium
-were observed to be almost equally likely. While the success of the ternary
-compex model is understandable based on the simulation results,
-a simple binding model cannot detec the complex binding
-oberved in the simulation.
-
-The improved cation binding behaviour to POPC bilayer pave the way for
-simulations of complex biochemical systems with correctly described
-electrostatic interactions in the vicinity of cellular membranes.
-The ECC-lipids model is build by scaling the partial charges and
-the LJ-radius of the headgroup, glycerol backbone and carbonyl atoms of
-the Lipid14 POPC model~\cite{dickson14}. While the Lipid14 model
-is compatible with the AMBER force field family, the compatibility of
-the ECC-lipids model may be compromised due to the changes in intermolecular
-interactions of the scaled atoms. On the other hand, a fully consistent
-ECC-force field should include the correction also in other than lipid
-molecules, including water. The work toward this direction and the extension
-to other lipid molecules and force fields is left for the future studies.
-
-%but we expect that the correction can be used also for other lipids
-%and force fields.
-
-%We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
-%which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
-%Several water models 
-%(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
-%were used to exemplify the transferability of 
-%the parameters of the new ECC-lipids force field.
+We show that the \ce{Na^+} and \ce{Ca^2+} binding to phospholipid bilayers can
+be accurately described with classical force field MD simulation models, where electronic
+polarization is effectively included using the electronic continuum correction (ECC)~\cite{leontyev11}.
+This is a significant improvement over other available lipid models,
+which all overestimate specific cation binding affinities~\cite{catte16}.  
+The newly proposed model, which we denote as ''ECC-lipids'', 
+exhibits accurate head group order parameter response to
+bound cations, namely monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
+and divalent \ce{Ca^2+},
+allowing also for quantifying their binding affinities.
+At the same time, ECC-lipids reproduce the lipid bilayer structural details
+with similar accuracy as other state of the art lipid models~\cite{catte16}.
+Several water models 
+(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
+were used to exemplify the transferability of 
+the parameters of the new ECC-lipids force field.
+
+Direct analysis of the details of calcium binding from MD simulations is in agreement
+with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
+In this model one calcium binds to two POPC molecules, which together form a ternary
+complex.
+\todo{Continue summary using previous section once it is finished.}
+
+The electronic continuum correction is built here upon the Lipid14 POPC model \cite{dickson14},
+but we expect that the correction can be used also for other lipids
+and force fields.
+\todo{JOE: Do we want to suggest that ECC-lipids can be used 
+with current force fields to study for example membrane proteins? 
+This was expressed with the following sentence: 
+The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. }
+We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
+which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
 
@@
 
-\newpage
 \newpage
@@
     \hline
-  \end{tabular} \\
-  \todo{Result with normal TIP3P missing?}
+  \end{tabular}
 \end{table}

### 2017-10-11 Manuscript/manuscript.tex
A dirty paragraph on the difference, when I consider carbonyl more as dipoles, added to stoichiometry section.

@@
 which are easier to interpret.
+
+Internal side-note: 
+However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
+when carbonyl dipoles are scaled only with 0.945 
+(roughly done through charge redistribution)
+ to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 . 
+This does not change the dominance of phosphates, which still interact more than 6x times more. That's good.
+In addition, the probability of states, containing sn-1:O increases from 6\% to 10\%.
+The residence times change to 75ns (90\% of bound time) and max is 185ns.
+That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
+The slight problem may be that with this tweak, there's approx. 20\%  more cations bound to the membrane,
+which means, I'd need to tune the sigma parameter a bit. 
+Not $f_q$ -- that's set by the simulations with the surfactant. 
+That could detune the membrane structure \dots
+I think that's within the accuracy.
 

### 2017-10-12 Manuscript/manuscript.tex
New abstract and small changes in the text

--- 2017-10-11 Manuscript/manuscript.tex
@@
   % insert abstract here
-  \todo{Abstract directly from Joe's conference abstracts. To be rewritten.}
-Classical molecular dynamics simulations give detailed information about membrane structure and dynamics. 
-However, there is still a room for improvements in current force fields – it is known from the literature,
-that the binding of ions, especially cations, to phopholipid membranes is overestimated in all classical models [1]. 
-We suggest that the membrane-ion interactions can be corrected by including implicit electronic polarizability into the
-lipid models through the electronic continuum correction (ECC) [2], which was already applied to monovalent and divalent
-ions yielding models that feature correct ion pairing [3]. 
-Using the electrometer concept [3, 4] and x-ray scattering form factors, our simulations point out that our hypothesis is
-correct and ECC is indeed a missing important contribution in current classical lipid models. 
-Moreover, the solid physical principles behind ECC are found not to hamper other relevant properties of a phospholipid bilayer. 
-The new lipid model, "ECC-lipids", shows accurate binding affinity to sodium and calcium cations and head group order parameter response to bound charge. 
-We also provide for the first time a realistic stochiometry of bound calcium cations to a POPC membrane, and their binding sites. 
-This work will continue as an open collaboration project NMRlipids VI (http://nmrlipids.blogspot.fi).
+  Despite of the significant biological relevance, the binding details of cations in cellular
+  lipid membranes are not fully understood. For example, consistent results for
+  the binding affinities and stoichimetries of Na$^+$ and Ca$^{2+}$ ions to phospholipid bilayers have not been
+  achieved with different experimental and theoretical methods. As recently shown in the 
+  NMRlipids project, a Open Collaboration running at \url{nmrlipids.blogspot.com},
+  the ion binding details could be resolved with unprecendent detail by intepreting
+  the experimental NMR data with classical molecular dynamics (MD) simulations.
+  However, the accuracy of the existing lipid models for MD simulations was not sufficient for this.
+  In this work we show that the binding details of Na$^+$ and Ca$^{2+}$ ions to
+  1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
+  can be accurately described with a MD simulation model having implicitly included electronic polarization.
+  This is demonstrated by applying the electronic continuum correction (ECC) to a
+  state of the art lipid model for MD simulations of POPC
+  lipid bilayer. The introduced ECC-lipid model reproduces the experimentally measured structural parameters for the ion-free membrane,
+  the response of lipid headgroup to the bound positive charge, and the binding affinities of
+  Na$^+$ and Ca$^{2+}$ ions. The imperceptible binding of Na$^+$ ions to POPC bilayer
+  and interactions of Ca$^{2+}$ mainly with phoshpate oxygens in the ECC-lipid model give
+  support to the early intepretations of the experimental spectroscopic data.
+  On the other hand, Ca$^{2+}$ ions form complexes with 1-3 lipid molecules with almost
+  equal probabilities, suggesting more complicated binding stoichimetry than 
+  the simple binding model used to interpret the NMR data.
+  The results in this work pave the way for MD simulations of complex biochemical systems
+  with realistic electrostatic interactions in the vicinity of cellular membranes.
 \end{abstract}
@@
 membranes. Especially the zwitterionic phosphocholine (PC) lipid bilayers are
-often used as model systems to understand the role of ions in complex biological
+used to understand the role of ions in complex biological
 systems~\cite{scherer87,seelig90,cevc90}.
@@
 bilayers between MD simulations and  NMR experiments, because the order
-parameters can be accurately determined from both techniques~\cite{ollila16}.
+parameters can be accurately determined from both techniques~\cite{catte16,ollila16}.
 The order parameters for all C-H bonds in lipid molecules, including
 $\alpha$ and $\beta$ segments in head group, can be accurately measured
-using $^2$H NMR or $^{13}$C NMR techniques. From MD simulations the
+using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}. From MD simulations the
 order parameters can be calculated using the definition
@@
 as in the earlier experiments~\cite{akutsu81}.
-Despite of the measurable differences between these two concentrations for CaCl$_2$ systems,
+Despite of the measurable differences between these two
+definitions of concentrations for CaCl$_2$ systems,
 the qualitative conclusions in this or in the previous work~\cite{catte16} are
-not affected.
+not affected by this.
 
@@
 The structure of lipid bilayers in simulations without ions were validated against NMR and 
-\mbox{x-ray} scattering experiments by calculating order parameters
-for C-H bonds and form factors. The former validates the structures sampled by individual
+\mbox{x-ray} scattering experiments by calculating the order parameters
+for C-H bonds and the scattering form factors. The former validates the structures sampled by individual
 lipid molecules in simulations with atomic resolution, while the latter
@@
 An automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium.%, which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
+dihexadecyldimethylammonium.
 The AmberTools program \cite{amber} was then used to generate the Amber-type force field
@@
 and by using the atom types with reduced $\sigma$ parameters from ECC-lipids. 
+\todo{Parameters are available at ??}
 
@@
 \begin{table}
-  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls} }
+  \caption{Area per lipid (APL) values of POPC with no ions from the Lipid14 simulation
+    ran in this work and from the literature, from ECC-lipid model and from experiments.\label{tab:apls} }
   \begin{tabular}{l|c c}

### 2017-10-12 Manuscript/manuscript.tex
New abstract and small changes in the text
Small update to the stochiometry section: added a few data (matti's C36 sim and mine sim with NaCl)

--- 2017-10-11 Manuscript/manuscript.tex
@@
 in the order of $1$--$10\,\mu\mathrm{s}$~\cite{javanainen17}.
+
+Quick result (Matti C36 with CaClMK, 800ns total):
+90\% of bound residence time is smaller than 787 ns, 
+The longest is 800 ns
+
 This is in contrast to our model, ECC-lipids, 
@@
 in equilibrium between membrane and solvent.
+
+Interestingly, \ce{Na+} cations exhibit much more rapid exchange between membrane and solvent.
+The residence time of sodium bound to a POPC membrane is
+smaller than $2\,\mathrm{ns}$ for 90\% of the time it is bound to the membrane 
+with the longest measured bound time being $13\,\mathrm{ns}$.
+
 A histogram of residence times from our simulations is in Fig.~S\ref{fig:hist_residence_times} in SI. 

### 2017-10-12 Manuscript/manuscript.tex
Added a few statements to stoichiometry section, now it shall be ready for re-writing it into a coherent text.

--- 2017-10-11 Manuscript/manuscript.tex
@@
   applies to the previous section, introduction, abstract and conclusions as well.
+  JOE: agree, this would be a good statement in general. I'll try to put it into the text.
 }
@@
 $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
-in which a contact between the two groups is counted 
-if their distance is less than or equal $0.3 \, \mathrm{nm}$ (robustness of this definition discussed in reference~\citenum{javanainen17}). 
-the average number of contacts between \ce{ca^{2+}} and any oxygen atom in popc is 
+in which a contact between the two atoms is counted 
+if their distance is less than or equal $0.3 \, \mathrm{nm}$ 
+(robustness of this definition discussed in reference~\citenum{javanainen17}). 
+The average number of contacts between \ce{Ca^{2+}} and any oxygen atom in POPC is 
 $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.47 $,    % = 59.8/128 
@@
 $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.45 $,    % = 57.2/128 . 
-and if only carbony oxygen atoms are considered,
+and if only carbonyl oxygen atoms are considered,
 this value is merely
@@
 
-Internal side-note: 
-However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
-when carbonyl dipoles are scaled only with 0.945 
-(roughly done through charge redistribution)
- to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 . 
-This does not change the dominance of phosphates, which still interact more than 6x times more. That's good.
-In addition, the probability of states, containing sn-1:O increases from 6\% to 10\%.
-The residence times change to 75ns (90\% of bound time) and max is 185ns.
-That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
-The slight problem may be that with this tweak, there's approx. 20\%  more cations bound to the membrane,
-which means, I'd need to tune the sigma parameter a bit. 
-Not $f_q$ -- that's set by the simulations with the surfactant. 
-That could detune the membrane structure \dots
-I think that's within the accuracy.
-
 The residence time of \ce{Ca^{2+}} bound to a POPC membrane 
@@
 (Fig. S17 in \citenum{javanainen17} -- C36 with MK Ca2+, any other model, either different cation or Slipids/C36, just such calcium)
-this time can be roughly estimated -- get the simulation from \url{https://doi.org/10.5281/zenodo.259376} and analyze it -- 
-in the order of $1$--$10\,\mu\mathrm{s}$~\cite{javanainen17}. 
+this time can be roughly estimated %-- get the simulation from \url{https://doi.org/10.5281/zenodo.259376} and analyze it -- 
+in the order of $1\,\mu\mathrm{s}$~\cite{javanainen17}.
+
+Improve the above statement so that it contains the reason why I estimate it to this value.
+Quick result (Matti C36 with CaClMK, 800ns total):
+90\% of bound residence time is smaller than 787 ns, 
+The longest is 800 ns (the simulation length)
+and it is only less than
+60\% of bound residence time is smaller than 400 ns -- half of the simulation length 
+and it is only less than
+50\% of bound residence time is smaller than 290 ns.
+Apparently the exchange of calcium between membrane and solvent 
+is apparently limited by the length of the simulation.
+
 This is in contrast to our model, ECC-lipids, 
@@
 in equilibrium between membrane and solvent.
+
+Interestingly, \ce{Na+} cations exhibit much more rapid exchange between membrane and solvent than calcium.
+The residence time of sodium bound to a POPC membrane is
+smaller than $2\,\mathrm{ns}$ for 90\% of the time it is bound to the membrane 
+with the longest measured bound time being $13\,\mathrm{ns}$. 
+This is an order of magnitude faster exchange than for calcium cations.
+
 A histogram of residence times from our simulations is in Fig.~S\ref{fig:hist_residence_times} in SI. 
 \todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
-Add data for Na+. }
+ }
+
+Internal side-note: 
+However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
+when carbonyl dipoles are scaled only with 0.945 
+(roughly done through charge redistribution)
+ to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 . 
+This does not change the dominance of phosphates, which still interact more than 6x times more. That's good.
+In addition, the states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
+so it can be thought of rather as a stabilizing moiety. 
+In addition, the probability of states, containing sn-1:O increases from 6\% to 10\%.
+The residence times change to 75ns (90\% of bound time) and max is 185ns.
+That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
+The slight problem may be that with this tweak, there's approx. 20\%  more cations bound to the membrane,
+which means, I'd need to tune the sigma parameter a bit (increase the scaling factor $f_\sigma$). 
+That could detune the membrane structure \dots
+Not $f_q$ -- that's set by the simulations with the surfactant. Good.
+I think that such a binding detail is within the accuracy of the proposed simple correction.
+It would be worth noting that the model may underestimate slightly the contribution from carbonyls due to this. 
+But we can't say to what extent\dots
 

### 2017-10-12 Manuscript/manuscript.tex
Corrected numbers of contacts in stoichiometry section and added probabilities of binding towards PO4 or carbonyls.

@@
 (robustness of this definition discussed in reference~\citenum{javanainen17}). 
-The average number of contacts between \ce{Ca^{2+}} and any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.47 $,    % = 59.8/128 
+The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
 whereas if only phosphate oxygen atoms are considered 
 the average number of contacts decreases only by a tiny amount to 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.45 $,    % = 57.2/128 . 
-and if only carbonyl oxygen atoms are considered,
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
+and if only carbonyl oxygen atoms are considered 
+(phosphate oxygens contribute, but are not counted as contacts),
 this value is merely
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.02 $.    % = (59.8-57.2)/128 . 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
+This means that calcium is in contact with a phosphate oxygen for 99\% of the time it is bound to the membrane, 
+whereas only 25\% of the time accounts for contact with a carbonyl oxygen,
+and purely carbonyl bound states account for less than 1\% of configurations. 
 We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
@@
  to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 .
-This does not change the dominance of phosphates, which still interact more than 6x times more. That's good.
-In addition, the states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
+
+The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 \rightarrow 0.309 $,
+whereas if only phosphate oxygen atoms are considered 
+the average number of contacts decreases only by a tiny amount to 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 \rightarrow 0.303 $,    % = 33.5/128 . 
+and if only carbonyl oxygen atoms are considered 
+(phosphate oxygens contribute, but are not counted as contacts),
+this value is merely
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 \rightarrow 0.139 $.    % = 8.4/128 .
+
+This means that calcium is in contact with a phosphate oxygen for 98\% of the time it is bound to the membrane, 
+whereas 46\% of the time accounts for contact with a carbonyl oxygen,
+and purely carbonyl bound states account for less than 2\% of configurations.
+
+The states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
 so it can be thought of rather as a stabilizing moiety. 
-In addition, the probability of states, containing sn-1:O increases from 6\% to 10\%.
+In addition, the probability of states (among bound states), containing sn-1:O increases from 20\% to 28\% (from MSM).
+Purely carbonyl-bound states make only 0.7\% and in carbonyl-dip corrected 1.4\%.
 The residence times change to 75ns (90\% of bound time) and max is 185ns.
 That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
-The slight problem may be that with this tweak, there's approx. 20\%  more cations bound to the membrane,
+The slight problem may be that with this tweak, there's approx. 17\%  more cations bound to the membrane,
 which means, I'd need to tune the sigma parameter a bit (increase the scaling factor $f_\sigma$). 

### 2017-10-13 Manuscript/manuscript.tex
Small updates to the text in Stoichiometry section.

@@
 (robustness of this definition discussed in reference~\citenum{javanainen17}). 
+% Numbers from simulation with 346mM conc. CaCl2
 The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
 whereas if only phosphate oxygen atoms are considered 
 the average number of contacts decreases only by a tiny amount to 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
 and if only carbonyl oxygen atoms are considered 
@@
 this value is merely
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
+% Numbers from simulation with 980/869mM conc. CaCl2
+%The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
+%whereas if only phosphate oxygen atoms are considered 
+%the average number of contacts decreases only by a tiny amount to 
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
+%and if only carbonyl oxygen atoms are considered 
+%(phosphate oxygens contribute, but are not counted as contacts),
+%this value is merely
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
 This means that calcium is in contact with a phosphate oxygen for 99\% of the time it is bound to the membrane, 
@@
 However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
+in simulation withy the highest \ce{CaCl2} concentration
 when carbonyl dipoles are scaled only with 0.945 

### 2017-10-13 Manuscript/manuscript.tex

@@
 such as in signal propagation in neurons and vesicle fusion.
-Since direct measurements of ion-membrane interactions from biological systems
+Since the direct measurements of ion-membrane interactions from biological systems
 are difficult, lipid bilayers are often used as models systems for cellular
-membranes. The detailed results from simple model systems can be then
-used to understand the role of ions in complex biological systems.
-
-Zwitterionic phosphocholine (PC) lipid bilayers are commonly used model systems for
-cellular membranes. Interactions of biological cations, especially Na$^+$ and
+membranes. Especially the zwitterionic phosphocholine (PC) lipid bilayers are
+used to understand the role of ions in complex biological
+systems~\cite{scherer87,seelig90,cevc90}.
+
+Interactions of biological cations, especially Na$^+$ and
 Ca$^{2+}$, with PC bilayers are widely studied in
@@
 \subsection{Electrometer concept} \label{section:electrometer}
-Ion binding between experiments and simulations was compared by 
-using lipid head group order parameters and the
-''electrometer concept'' \cite{seelig87,catte16}.
-The concept is based on the experimental observation that the C-H bond
+Ion binding in lipid bilayers was compared between experiments and simulations by 
+using the lipid head group order parameters and the
+''electrometer concept'' \cite{seelig87,catte16}, which
+is based on the experimental observation that the C-H bond
 order parameters of $\alpha$ and $\beta$ carbons in PC lipid head group
 (see Fig.~\ref{simVSexpNOions}) are proportional to the amount 
-of unit charge bound per lipid,  $X^\pm$~\cite{seelig87}.
-Change in the order parameters measured with varying aqueous
+of unit charge bound per lipid~\cite{seelig87}.
+The change of order parameters measured with varying aqueous
 ion concentration can be then related to the amount of bound ions.
 
-The change of the head group order parameters is empirically quantified
-as~\cite{seelig87,ferreira16}
+The concept can be used to compare the ion binding affinity in lipid 
+bilayers between MD simulations and  NMR experiments, because the order
+parameters can be accurately determined from both techniques~\cite{catte16,ollila16}.
+The order parameters for all C-H bonds in lipid molecules, including
+$\alpha$ and $\beta$ segments in head group, can be accurately measured
+using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}. From MD simulations the
+order parameters can be calculated using the definition
+\begin{equation}\label{OP}
+S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
+\end{equation}
+where $\theta$ is the angle between the bond and membrane
+normal and the average is taken over all sampled configurations \cite{ollila16}.
+
+The relation between bound charge per lipid,  $X^\pm$, and
+the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$,
+is empirically quantified as~\cite{seelig87,ferreira16}
 \begin{equation}\label{OPchangeEQ}
@@
 Atomic absorption spectra and $^2$H NMR data gave
-$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for Ca$^{2+}$ binding to POPC
-bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-The slopes are negative, because recent analysis concluded 
-that the order parameters decrease with bound positive charge 
-and increase with bound negative charge when the signs are taken 
-in account \cite{ollila16,catte16}. This is rationalized as a change
-of lipid head group dipole tilt toward water phase with bound positive
-charge and {\it vice versa} with negative charge \cite{seelig87}.
-
-The concept can be used to compare the ion binding affinity in lipid 
-bilayers between MD simulations and  NMR experiments, because the order
-parameters can be accurately determined from both techniques~\cite{ollila16}.
-The order parameters for all C-H bonds in lipid molecules, including
-$\alpha$ and $\beta$ segments in head group, can be accurately measured
-using $^2$H NMR or $^{13}$C NMR techniques. From MD simulations the
-order parameters can be calculated using the definition
-\begin{equation}\label{OP}
-S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
-\end{equation}
-where $\theta$ is the angle between the bond and membrane
-normal and the average is taken over all sampled configurations \cite{ollila16}.
-
-The measured order parameter change depends on the response of 
-the headgroup on the bound charge, i.e. $m_i$ in Eq.~\ref{OPchangeEQ},
-and the ion binding affinity. Thus, the former property has to be
-well quantified before using the electrometer concept to
-analyze binding affinities. This is done experimentally for a wide range
+$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for the binding of
+Ca$^{2+}$ to POPC bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
+The slopes are negative, because the head group order parameters
+decrease with bound positive charge and increase with bound negative
+charge~\cite{ollila16,catte16}. This can be rationalized as a change
+of the lipid head group dipole tilt toward water phase with the bound positive
+charge and {\it vice versa} with the negative charge~\cite{seelig87}.
+
+The measured order parameter change depends on both, the head group response 
+to the bound charge and the amount of bound charge, i.e. $m_i$ and $X^\pm$
+in Eq.~\ref{OPchangeEQ}, respectively. The former property has to be
+well quantified before the electrometer concept can used to 
+analyze the binding affinities, as done experimentally for a wide range
 of systems~\cite{seelig87,beschiasvili91}. To calibrate 
-the head group order parameter response also in simulations,
-we use experimental data for dihexadecyldimethylammonium (2C$_{16}$$^+$N2C$_1$Br$^-$)
-in a POPC bilayer \cite{scherer89}. Dihexadecyldimethylammonium
+the head group order parameter response to the bound charge in simulations,
+we use the experimental data for dihexadecyldimethylammoniumbromide %(2C$_{16}$$^+$N2C$_1$Br$^-$)
+mixed within a POPC bilayer~\cite{scherer89}. Dihexadecyldimethylammonium\\[0.5cm]
+%\vspace{0.5cm}
+\chemfig{
+ -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
+}  
+\vspace{0.5cm} \\
 is a cation surfactant having two acyl chains and bearing a unit
@@
 in the bilayer similarly to the phospholipids and the molar ratio then 
-gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems \cite{scherer87}.
+gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
 
 \subsection{Salt concentrations and binding affinity}
-Experimental studies providing the lipid head group order
-parameters report salt concentrations in two different ways.
-For DPPC, salt concentrations before solvating
-the lipids are used~\cite{akutsu81}. Alternatively, for POPC,
-the salt concentration of supernatant after solvation is
-measured using atomic absorption~\cite{altenbach84}.
-\todo{JOE: I'm afraid that the above description might sound to the reader
-  like for some reasons one method is used with DPPC, but a different one with POPC.
-  I think simply writing one way is\dots another way is \dots (or similarly) will be less confusing.}
-In this work, we will use the latter definition. For this we
-will calculate the salt concentration in the aqueous bulk region,
-\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
+The early experimental electrometer concept data for the lipid head group
+order parameter changes was reported as a function 
+of the salt concentration in water before solvating
+the lipids~\cite{akutsu81}. The later study used atomic absorption
+spectroscopy and reported the order parameter changes as
+a function of the salt concentration in the supernatant after the
+solvation of lipids~\cite{altenbach84}.
+In this work we focus on POPC for which the latter definition was used~\cite{altenbach84}.
+The salt concentration in the aqueous bulk region was calculated from 
+the farthest point from both lipid leaflets in the water phase. 
 Note that in the previous study, Ref.~\citenum{catte16},
-the concentration was calculated as in the DPPC experiments.
-Despite of the measurable differences between these two concentrations for CaCl$_2$ systems,
-the qualitative conclusions are affected neither in this
-nor in our previous work~\cite{catte16}.
-
-To quantify the ion binding affinity to a membrane, we
-calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
-This quantity can be used to compare binding affinity between different models,
-because it does not depend on the definition of the interfacial region
-or the position of the Gibbs dividing plane between two bulk regions.
-\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
-Therefore, we consider the hydrophobic interior of a membrane having zero concentrations of ions and water
-as a bulk region, which is separated by a monolayer interface from 
-a another region with bulk concentration of ions in water.
-Since the boundary of the latter region is set to the edge of simulation box,
-all molecules in simulation are within the defined interfacial region. 
+the ion concentrations were calculated in water before solvating the lipids 
+as in the earlier experiments~\cite{akutsu81}.
+Despite of the measurable differences between these two
+definitions of concentrations for CaCl$_2$ systems,
+the qualitative conclusions in this or in the previous work~\cite{catte16} are
+not affected by this.
+
+To quantify the ion binding affinity to a lipid bilayer, we
+calculated the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
+This quantity was chosen because it does not depend on
+%the definition of the interfacial region or
+the position of the Gibbs dividing plane between two bulk regions.
+%\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
+%SAMULI: Here is a new trial.
+Here we assume that the interface locates between the hydrophobic interior
+of a lipid bilayer and the bulk water region far from the membrane.
+The bulk concentration of ions and water is zero inside the bilayer.
+The concentrations in bulk water region can be calculated from 
+the farthest point from both lipid leaflets in the water phase.
+The region between these boundaries contains all the ions and water
+molecules in the simulation box.
 %our whole simulation box as
@@
 
-Lipid bilayer structure without ions was validated against NMR and 
-\mbox{x-ray} scattering experiments by calculating order parameters
-for C-H bonds and form factors from our simulations. The former validates the structures sampled by individual
+The structure of lipid bilayers in simulations without ions were validated against NMR and 
+\mbox{x-ray} scattering experiments by calculating the order parameters
+for C-H bonds and the scattering form factors. The former validates the structures sampled by individual
 lipid molecules in simulations with atomic resolution, while the latter
@@
 OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
-presented in Supporting Information (SI). 
-We used the ECC-ions model for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
+\todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
+The ECC-ion models were used for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+Simulation with the Lipid14 model was also ran with the ion model by Dang et al.~\cite{smith94,chang1999,dang2006}.
+Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion model
+was taken directly from \cite{lipid14POPC0mMNaClfiles}. 
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
@@
 An automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium \\[0.5cm]
-%\vspace{0.5cm}
-\chemfig{
- -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH2{)}_{15}-CH_3)(-[:30]{(}CH2{)}_{15}-CH_3)
-} , 
-\vspace{0.5cm} \\
-which is one of the cationic surfactants used to experimentally quantify the electrometer concept~\cite{scherer89}.
+dihexadecyldimethylammonium.
 The AmberTools program \cite{amber} was then used to generate the Amber-type force field
@@
 
-%In order to validate the structural properties of the newly developed ECC-lipids model
-%we first compared our simulation results without any ions to
-%NMR order parameters measurements and x-ray scattering form factors 
-%(Fig.~\ref{simVSexpNOions}). 
-The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and
-the area per lipid (Table~\ref{tab:apls}) of POPC bilayer simulated
-with the ECC-lipid model are in good agreement with the experimental results,
-as well as with the Lipid14 model. Thus, we conclude that the lipid bilayer dimensions
-are well captured by the developed ECC-lipid model. 
-Areas per lipid of ECC-lipids with various other water models are documented in SI, table~\ref{tab:apls_si}. 
+The ECC-lipid and Lipid14 models reproduce the experimental x-ray scattering form factors
+of POPC bilayer with the similar accuracy in Fig.~\ref{simVSexpNOions}.
+The area per lipid values from the Lipid14 model is slightly larger than the
+experimental value in Table~\ref{tab:apls}, while the value from ECC-lipid model
+is slightly smaller. Area per lipid values of the ECC-lipid model show some variation
+when simulated with different water models in Table~\ref{tab:apls_si} (Supplementary Information),
+however, all the values are close to the experimentally reported values.
+In conclusion, the ECC-lipid model reproduces the experimental dimensions of POPC
+lipid bilayer with the accuracy comparable to the other state of the art lipid models \cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 \section{Conclusions}
-We show that the \ce{Na^+} and \ce{Ca^2+} binding to phospholipid bilayers can
-be accurately described with classical force field MD simulation models, where electronic
-polarization is effectively included using the electronic continuum correction (ECC)~\cite{leontyev11}.
-This is a significant improvement over other available lipid models,
-which all overestimate specific cation binding affinities~\cite{catte16}.  
-The newly proposed model, which we denote as ''ECC-lipids'', 
-exhibits accurate head group order parameter response to
-bound cations, namely monovalent \ce{Na^+} and cationic surfactant dihexadecyldimethylammonium bromide, 
-and divalent \ce{Ca^2+},
-allowing also for quantifying their binding affinities.
-At the same time, ECC-lipids reproduce the lipid bilayer structural details
-with similar accuracy as other state of the art lipid models~\cite{catte16}.
-Several water models 
-(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
-were used to exemplify the transferability of 
-the parameters of the new ECC-lipids force field.
-
-Direct analysis of the details of calcium binding from MD simulations is in agreement
-with ternary complex model, which is suggested based on NMR data \cite{altenbach84}.
-In this model one calcium binds to two POPC molecules, which together form a ternary
-complex.
-\todo{Continue summary using previous section once it is finished.}
-
-The electronic continuum correction is built here upon the Lipid14 POPC model \cite{dickson14},
-but we expect that the correction can be used also for other lipids
-and force fields.
-\todo{JOE: Do we want to suggest that ECC-lipids can be used 
-with current force fields to study for example membrane proteins? 
-This was expressed with the following sentence: 
-The parameters can be used with existing standard nucleic acid and protein force fields, e.g. AMBER-FB15~\cite{Wang2017}. }
-We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
-which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
-
-This work can be reached as a repository containing all data at \url{zenodo.org:\dots\dots\dots} and as project \mbox{NMRlipids~VI} in \url{nmrlipids.blogspot.fi}.
+By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
+to a POPC lipid bilayer can be accurately described with the classical MD simulation 
+force field, where the electronic polarization is implicitly included using 
+the electronic continuum correction (ECC)~\cite{leontyev11}.
+The proposed ECC-lipids model is a significant improvement over 
+other available lipid models, which all overestimate specific cation binding affinities~\cite{catte16}.  
+While the structural details of a POPC lipid bilayer simulated with the ECC-lipids
+model agree with experiments with the comparable accuracy to the other state of the art lipid models,
+it also reproduces the experimental lipid head group order parameter responses to
+the cationic surfactant, NaCl and CaCl$_2$ concentrations.
+
+The good agreement with experiments enables the atomic resolution 
+interpretation of NMR experiments by using MD simulations.
+In agreement with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
+the Ca$^{2+}$ ions mainly interact with phosphate oxygens.
+However, the stoichiometry of the binding is significantly more complicated 
+than in the previous interpretation of the NMR data based on 
+the ternary complex model, where one calcium binds to two POPC molecules~\cite{altenbach84}.
+The complexes of one calcium bound to two lipids are the most probable also in the
+ECC-lipids model, but the complexes of one or three lipids per one calcium
+were observed to be almost equally likely. While the success of the ternary
+compex model is understandable based on the simulation results,
+a simple binding model cannot detec the complex binding
+oberved in the simulation.
+
+The improved cation binding behaviour to POPC bilayer pave the way for
+simulations of complex biochemical systems with correctly described
+electrostatic interactions in the vicinity of cellular membranes.
+The ECC-lipids model is build by scaling the partial charges and
+the LJ-radius of the headgroup, glycerol backbone and carbonyl atoms of
+the Lipid14 POPC model~\cite{dickson14}. While the Lipid14 model
+is compatible with the AMBER force field family, the compatibility of
+the ECC-lipids model may be compromised due to the changes in intermolecular
+interactions of the scaled atoms. On the other hand, a fully consistent
+ECC-force field should include the correction also in other than lipid
+molecules, including water. The work toward this direction and the extension
+to other lipid molecules and force fields is left for the future studies.
+
+%but we expect that the correction can be used also for other lipids
+%and force fields.
+
+%We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
+%which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
+%Several water models 
+%(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
+%were used to exemplify the transferability of 
+%the parameters of the new ECC-lipids force field.
+
+This work can be reached as a repository containing all data at \url{zenodo.org:\dots\dots\dots}.
+%and as project \mbox{NMRlipids~VI} in \url{nmrlipids.blogspot.fi}.
 

### 2017-10-14 Manuscript/manuscript.tex

@@
 All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
-6, 14, 21, 35, or 50 surfactants per leaflet. Chloride counter ions were used
+6, 14, 21, 35, or 50 surfactants per leaflet. 
+Chloride counter ions were used
 in simulations, because parameters for bromide, the counterion in the experimental data \cite{scherer89},
 were not available in the standard Gromacs files for Amber force field.
+\todo{JOE: We cannot present an intricate cool new model for a lipid and claim at the same time that we can't simulate bromide. }
 The Lipid14 model was used for POPC.
@@
 
-\subsection{Response of POPC head groups to bound charge}\label{section:boundCHARGE}
+\subsection{Calibration of lipid electrometer:
+            Response of POPC head groups to bound charge}\label{section:boundCHARGE}
 
@@
 
-\subsection{Cation binding affinities to POPC}
+\subsection{Cation binding affinities to POPC membrane validated through lipid electrometer}
 
@@
   \centering
-  \includegraphics[width=7.5cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
   \caption{\label{fig:delta_ordPar_CaCl}
@@
 
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
+  \caption{\label{fig:delta_ordPar_NaCl}
+    Changes of head group order parameters $\alpha$ and $\beta$ of POPC bilayer as a function of NaCl concentrations
+    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
+    Ion concentrations in bulk water are shown in x-axis. 
+    Values from simulations are calculated from the of cation number density $C_{np}$
+    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
+  }
+\end{figure}
+
 The changes of the lipid bilayer headgroup order parameter from different simulations and
-experiments \cite{akutsu81,altenbach84} are shown in Fig.~\ref{fig:delta_ordPar_CaCl}
+experiments \cite{akutsu81,altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}
 as a function of NaCl and CaCl$_2$ concentrations.
@@
 
-To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
-polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
-The ECC-lipid model simulations with the ECC-ion models \cite{??}
-exhibit a good agreement with the experimental changes of 
+%To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
+%polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
+Our simulations with the ECC-lipid and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
+exhibit an improved behaviour of cation binding to a POPC bilayer
+showing a good agreement with experiments in the changes of 
 the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations
@@
 Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.
-
-The density distributions of water and ions along the membrane normal
-are shown in Fig.~\ref{fig:cacl-dens}.
-\todo{Densities and relative surface excess from NaCl systems to be added.
-  The discussion below to be finished after this.}
+%In conclusion, the results suggest that 
+%the ECC-lipid model together with the ECC-ion parameters gives a realistic
+%description for the binding affinities of Na$^+$ and Ca$^{2+}$ ions to POPC lipid bilayer. 
+Furthermore, the overestimated headgroup order parameter changes of POPC in the Lipid14
+model as a function of CaCl$_2$ concentration arise from both, the overestimated
+binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
+This probably applies also to the other lipid models considered in the previous study~\cite{catte16},
+emphasizing the importance of the comparison of the lipid headgroup order parameter response
+to the bound charge between simulations and experiments as done in section \ref{section:boundCHARGE}.
+
+\todo{SAMULI: Maybe we should discuss the repeat distances and
+  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
+  }
+
+\subsection{Cation binding to POPC membrane in atomistic detail}
+
+To quantify the binding affinities in different models we calculated 
+the relative surface excess from Eq.~\ref{surfexcess}. 
+Calcium cations at a molar concentration of $350\,\mathrm{mM}$ provide
+$\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$ in the ECC-lipid model, which is significantly
+smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
+or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
+\todo{JOE: I'd put the $\Gamma$ values into a table}
+Interestingly, the relative surface excess of 
+\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) is qualitatively different from \ce{CaCl2}
+having a negative value $\Gamma_i^w = -0.1 \pm 0.01 \rm{nm}^{-2}$, 
+meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
+This is in contrast to previous works employing classical MD simulations,
+which generally agree together on the existence of an increased sodium density at the membrane-water interface.
+\todo{JOE: Add references to NaCl simulation papers that show peaks.
+Add a simple note that there is NBFIX trick in Charmm36 (?).} 
+This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
+which show the density profile of ions and water along the membrane normal. 
 %The density profiles show
@@
 %the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
-To quantify the binding affinities in different models we calculated 
-the relative surface excess from Eq.~\ref{surfexcess}. The value for Calcium
-in the ECC-lipid model, $\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$, is significantly
-smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
-or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
-
-%\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
-%  do this analysis by using the density profiles, but it is not necessary.
-%  The reason why I want to do something like this, is that it would be useful to have
-%  some easier way to estimate the correct binding affinity than the electrometer concept,
-%  which is quite tedious to apply in practise (simulations with different concenrations,
-%  cationic surfactant check, etc.). I wuold like to be able to estimate much faster from a
-%  simulation if the binding affinity is reasonable or not. This may not be the correct way for
-%that anyway.}
-%Rough estimates for the free energy difference between bound and unbound cation
-%are given by
-%\begin{equation}
-%  \Delta G = -{\rm k_b T}\log(\frac{p_i}{p_o}),
-%\end{equation}
-%where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
-%in the maximum density in bilayer, respectively. The density profiles in
-%Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
-%between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
-%in Lipid14 with \AA{}qvist.
+\todo{Densities and relative surface excess from NaCl systems to be added.
+  The discussion below to be finished after this.}
 
@@
 
-In conclusion, the results suggest that 
-the ECC-lipid model together with the ECC-ion parameters gives a realistic
-description for the binding affinities of Na$^+$ and Ca$^{2+}$ ions to POPC lipid
-bilayer. Furthermore, the overestimated headgroup order parameter changes of POPC in the Lipid14
-model as a function of CaCl$_2$ concentration arise from both, the overestimated
-binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
-This probably applies also to the other lipid models considered in the previous study~\cite{catte16},
-emphasizing the importance of the comparison of the lipid headgroup order parameter response
-to the bound charge between simulations and experiments as done in section \ref{section:boundCHARGE}.
-
-\todo{SAMULI: Maybe we should discuss the repeat distances and
-  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
-  }
-
-\subsection{Binding stoichiometry}
-
-%Binding probability:
-%\ce{Ca2+} complex with \\
-%1 POPC - 13\% -- 30\% \\
-%2 POPC - 18\% -- 42\% \\
-%3 POPC - 12\% -- 28\% \\
-
-This section is rough and will likely require editing. 
-Data -- up to noted exceptions -- shall be all there, however. 
-\todo{SAMULI: I think the we should make clear that the current binding
-  constants in, e.g., Marsh's handbook are based on the headgroup order parameter
-  changes intepreted with ternary complex binding model. I.e. the experimental raw data
-  is excatly the same as we have in this work. Our simulations enable much more versatile interpretation
-  of the binding phenomena. I think that this is one the main points of this work, so this comment
-  applies to the previous section, introduction, abstract and conclusions as well.
-  JOE: agree, this would be a good statement in general. I'll try to put it into the text.
-}
-
-Binding stoichiometry of \ce{Ca^{2+}} and POPC was 
-thoroughly studied in the experimental work~\cite{altenbach84},
-in which the head group order parameter changes to cation binding are determined.
-Several binding models were proposed and tested of which only one,
-ternary complex binding model, provided a good fit of the experimental observations.
-
-Simulations allow us to directly evaluate the stoichiometry 
-by calculating relative propensities of various \ce{Ca^{2+}}:$n\times$POPC clusters
-by evaluating contacts between cations and lipids with a cut off radius $0.3\,\mathrm{nm}$. 
-In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
-the most probable binding mode of calcium at 285~mM concentration. 
-Apart from this complex, we also find complexes with 1~and 3~lipids
-occuring with only a slightly lower but similar probability.
-The fractions of \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration 
-are then in order:
-$42\%$ for two lipids,
-$30\%$ for one lipid, and
-$28\%$ for three lipids.
-
-Several binding models were proposed and tested~\cite{altenbach84} of which only one,
-ternary complex binding model, provided a good fit of the experimental observations. 
-In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
-with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
-In a later work~\cite{macdonald87}, 
-a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
-was found to provide as good fit as ternary complex model, 
-when only low concentrations of \ce{CaCl2} are considered.
-Ternary complex model also provides a good fit 
-to our simulations with ECC-lipids 
-(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
-The symmetry of the distribution of complexes from simulation --  
-i.e. almost equal probabilities of complexes with 1 or 3 lipids 
-that behave in the total average picture as complexes with 2 lipids -- 
-provides clues why ternary complex binding model 
-fits both simulation and experimental results relatively well, 
-although it is apparently incorrect.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
-  \caption{\label{fig:cacl_complexes}
-      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
-      with a certain number of POPC lipids. 
-      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
-      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
-  }
-\end{figure}
-
-In addition, we estimated relative binding affinities 
-of several moieties in POPC towards \ce{Ca^{2+}}. 
-The dominant contribution
-to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
-Such a finding is based on the calculated
-number of contacts per lipid between \ce{Ca^{2+}} and POPC oxygen atoms, 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
+We can estimate relative binding affinities 
+of several moieties in POPC towards \ce{Ca^{2+}} and \ce{Na+} cations 
+using the number of contacts per lipid between cations and POPC oxygen atoms, 
+$\Sigma ^\mathrm{Na^{2+}} _\mathrm{O}$,
 in which a contact between the two atoms is counted 
@@
 (robustness of this definition discussed in reference~\citenum{javanainen17}).
+
+For sodium at $1000\,mM$ concentration 
+the average number of contacts per lipid with any oxygen atom in POPC is 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
+whereas if only phosphate oxygen atoms are considered 
+this number decreases to 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
+and if only carbonyl oxygen atoms are counted 
+it is even lower, 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
+This means that interaction of sodium and POPC 
+has balanced contributions of 55\% from only phosphate oxygens 
+(i.e. no contacts with carbonyl oxyens),
+and from only carbonyl oxygens with 20\% of configurations 
+leaving 25\% for mixed phosphate and carbonyl oxygen configurations.  
+This is in line with the calculated negative value of 
+relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
+their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
+suggesting that the populations of bound sodium cations 
+naturally decrease with the decreasing density of water.
+
+In contrast to sodium, the dominant contribution
+to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
 % Numbers from simulation with 346mM conc. CaCl2
-The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
+The average number of contacts per lipid 
+between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
 $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
@@
 and if only carbonyl oxygen atoms are considered 
-(phosphate oxygens contribute, but are not counted as contacts),
+%(phosphate oxygens contribute, but are not counted as contacts),
 this value is merely
 $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
-% Numbers from simulation with 980/869mM conc. CaCl2
+% Numbers from simulation with 980/845mM conc. CaCl2
 %The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
@@
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
-This means that calcium is in contact with a phosphate oxygen for 99\% of the time it is bound to the membrane, 
-whereas only 25\% of the time accounts for contact with a carbonyl oxygen,
-and purely carbonyl bound states account for less than 1\% of configurations. 
+Hence, calcium is in contact with phosphate oxygens for 99\% of the time it is bound to the membrane, 
+whereas only 25\% of the time accounts for contacts with carbonyl oxygens,
+and interactions purely with carbonyl oxygens account for less than 1\% of configurations. 
+Higher concentration of \ce{CaCl2} naturally increases the amount of contacts per lipid,
+for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
+the average number of contacts per lipid is increased to
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
+but the distribution of states between phosphate and carbonyl oxygens remains unchanged. 
 We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
-which are easier to interpret.
-
-The residence time of \ce{Ca^{2+}} bound to a POPC membrane 
-is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
-From a recent theoretical work with long simulations 
-(Fig. S17 in \citenum{javanainen17} -- C36 with MK Ca2+, any other model, either different cation or Slipids/C36, just such calcium)
-this time can be roughly estimated %-- get the simulation from \url{https://doi.org/10.5281/zenodo.259376} and analyze it -- 
-in the order of $1\,\mu\mathrm{s}$~\cite{javanainen17}.
-
-Improve the above statement so that it contains the reason why I estimate it to this value.
-Quick result (Matti C36 with CaClMK, 800ns total):
-90\% of bound residence time is smaller than 787 ns, 
-The longest is 800 ns (the simulation length)
-and it is only less than
-60\% of bound residence time is smaller than 400 ns -- half of the simulation length 
-and it is only less than
-50\% of bound residence time is smaller than 290 ns.
-Apparently the exchange of calcium between membrane and solvent 
-is apparently limited by the length of the simulation.
-
-This is in contrast to our model, ECC-lipids, 
-which estimates the residence time of calcium bound to a POPC membrane to be 
-smaller than $53\,\mathrm{ns}$ for 90\% of the time a cation is in a membrane-bound state. 
-i.e. at least two orders of magnitude lower than previous estimates.
-In addition, the longest observed residence time in our simulation is $120\,\mathrm{ns}$, 
-which is well below the length of our simulations.
-This is unfortunately not the case of many theoretical works,
-where the residence time cannot be estimated,
-or the simulation has not even decorrelated from it's 
-initial condition with a population of bound/unbound cations \cite{catte16}, 
-because it is limited by the simulation length, 
-even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
-Increasing time scales to the order of $\mu\mathrm{s}$ 
-doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used. 
-Such a finding changes the point of view of calcium binding from
-very tight long-term stable binding with rare exchanges to 
-a relatively frequent exchange of cations 
-in equilibrium between membrane and solvent.
-
-Interestingly, \ce{Na+} cations exhibit much more rapid exchange between membrane and solvent than calcium.
-The residence time of sodium bound to a POPC membrane is
-smaller than $2\,\mathrm{ns}$ for 90\% of the time it is bound to the membrane 
-with the longest measured bound time being $13\,\mathrm{ns}$. 
-This is an order of magnitude faster exchange than for calcium cations.
-
-A histogram of residence times from our simulations is in Fig.~S\ref{fig:hist_residence_times} in SI. 
-\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
- }
-
-Internal side-note: 
-However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
-in simulation withy the highest \ce{CaCl2} concentration
-when carbonyl dipoles are scaled only with 0.945 
-(roughly done through charge redistribution)
- to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 .
-
-The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 \rightarrow 0.309 $,
-whereas if only phosphate oxygen atoms are considered 
-the average number of contacts decreases only by a tiny amount to 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 \rightarrow 0.303 $,    % = 33.5/128 . 
-and if only carbonyl oxygen atoms are considered 
-(phosphate oxygens contribute, but are not counted as contacts),
-this value is merely
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 \rightarrow 0.139 $.    % = 8.4/128 .
-
-This means that calcium is in contact with a phosphate oxygen for 98\% of the time it is bound to the membrane, 
-whereas 46\% of the time accounts for contact with a carbonyl oxygen,
-and purely carbonyl bound states account for less than 2\% of configurations.
-
-The states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
-so it can be thought of rather as a stabilizing moiety. 
-In addition, the probability of states (among bound states), containing sn-1:O increases from 20\% to 28\% (from MSM).
-Purely carbonyl-bound states make only 0.7\% and in carbonyl-dip corrected 1.4\%.
-The residence times change to 75ns (90\% of bound time) and max is 185ns.
-That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
-The slight problem may be that with this tweak, there's approx. 17\%  more cations bound to the membrane,
-which means, I'd need to tune the sigma parameter a bit (increase the scaling factor $f_\sigma$). 
-That could detune the membrane structure \dots
-Not $f_q$ -- that's set by the simulations with the surfactant. Good.
-I think that such a binding detail is within the accuracy of the proposed simple correction.
-It would be worth noting that the model may underestimate slightly the contribution from carbonyls due to this. 
-But we can't say to what extent\dots
+and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.
 
@@
 \end{figure}
+
+\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in POPC membrane}
+
+The residence time of \ce{Ca^{2+}} in a POPC membrane 
+is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}. 
+From a recent theoretical work with long simulations \cite{javanainen17}
+this time can be roughly estimated 
+in the order of $1\,\mu\mathrm{s}$~\cite{javanainen17}. 
+Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
+we took the simulation with the apparent highest fluctuations of the number of bound calcium,
+which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
+and anaylsed the residence times.
+% simulation: (Matti C36 with CaClMK, 800ns total):
+We find out that
+approximately 20\% of bound residence times of \ce{Ca^{2+}} 
+are limited by the length of the simulation, $800\,\mathrm{ns}$, 
+and only less than
+60\% of bound residence time is smaller than half of the simulation length , $400\,\mathrm{ns}$. 
+Other simulations reported in that work show even longer residence times.
+
+This is in contrast to our model, ECC-lipids, 
+which estimates the residence time of calcium bound to a POPC membrane to be 
+smaller than $60\,\mathrm{ns}$ for 90\% of the time a cation is bound. % exactly $53\,\mathrm{ns}$
+i.e. at least two orders of magnitude lower than previous estimates.
+In addition, the longest observed residence time in our simulation is $120\,\mathrm{ns}$, 
+which is well below the length of our simulations.
+This is unfortunately not the case of many theoretical works,
+where the residence time cannot be estimated,
+or the simulation has not even decorrelated from it's 
+initial condition with a population of bound/unbound cations \cite{catte16}, 
+because it is limited by the simulation length, 
+even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
+Increasing time scales to the order of $\mu\mathrm{s}$ 
+doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used. 
+Such a finding changes the point of view on binding of calcium to PC membranes from
+a very strong long-term stable binding with rare exchanges to 
+a relatively frequent exchange of clacium cations 
+in equilibrium between membrane and solvent.
+
+Interestingly, \ce{Na+} cations exhibit much more rapid exchange between membrane and solvent than calcium.
+The residence time of sodium bound to a POPC membrane is
+smaller than $2\,\mathrm{ns}$ for 90\% of the time it is bound to the membrane 
+with the longest measured bound time being $13\,\mathrm{ns}$. 
+Such an exchange between membrane and solvent 
+is an order of magnitude faster than for calcium cations. 
+A histogram of residence times from our simulations is in Fig.~S\ref{fig:hist_residence_times} in SI. 
+\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
+ }
+
+\subsubsection{Internal side-note: }
+However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
+in simulation withy the highest \ce{CaCl2} concentration
+when carbonyl dipoles are scaled only with 0.945 
+(roughly done through charge redistribution)
+ to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 .
+
+The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 \rightarrow 0.309 $,
+whereas if only phosphate oxygen atoms are considered 
+the average number of contacts decreases only by a tiny amount to 
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 \rightarrow 0.303 $,    % = 33.5/128 . 
+and if only carbonyl oxygen atoms are considered 
+(phosphate oxygens contribute, but are not counted as contacts),
+this value is merely
+$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 \rightarrow 0.139 $.    % = 8.4/128 .
+
+This means that calcium is in contact with a phosphate oxygen for 98\% of the time it is bound to the membrane, 
+whereas 46\% of the time accounts for contact with a carbonyl oxygen,
+and purely carbonyl bound states account for less than 2\% of configurations. 
+The states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
+so it can be thought of rather as a stabilizing moiety.
+
+The residence times change to 75ns (90\% of bound time) and max is 185ns.
+That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
+The slight problem may be that with this tweak, there's approx. 17\%  more cations bound to the membrane,
+which means, I'd need to tune the sigma parameter a bit (increase the scaling factor $f_\sigma$). 
+That could detune the membrane structure \dots
+Not $f_q$ -- that's set by the simulations with the surfactant. Good.
+I think that such a binding detail is within the accuracy of the proposed simple correction.
+It would be worth noting that the model may underestimate slightly the contribution from carbonyls due to this. 
+But we can't say to what extent\dots
 
@@
 
+\subsection{Stoichiometry of binding of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
+
+%Binding probability:
+%\ce{Ca2+} complex with \\
+%1 POPC - 13\% -- 30\% \\
+%2 POPC - 18\% -- 42\% \\
+%3 POPC - 12\% -- 28\% \\
+
+Simulations allow us to directly evaluate the stoichiometry 
+by calculating relative propensities of various \ce{Ca^{2+}}:$n\times$POPC complexes
+by evaluating contacts between cations and lipids with a cut off radius $0.3\,\mathrm{nm}$. 
+The propensities of \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration 
+are in order:
+$42\%$ for two lipids,
+$30\%$ for one lipid, and
+$28\%$ for three lipids.
+Apart from a ternary complex, in which one calcium cation binds two lipids, 
+we also find complexes with 1~and 3~lipids
+occuring with only a slightly lower but similar probability
+as shown in Figure~\ref{fig:cacl_complexes}.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
+  \caption{\label{fig:cacl_complexes}
+      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
+      with a certain number of POPC lipids. 
+      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
+      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
+  }
+\end{figure}
+
+Several binding models were proposed and tested~\cite{altenbach84} of which only one,
+ternary complex binding model, provided a good fit of the experimental observations. 
+In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
+with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
+In a later work~\cite{macdonald87}, 
+a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
+was found to provide as good fit as ternary complex model, 
+when only low concentrations of \ce{CaCl2} are considered.
+Ternary complex model also provides a good fit 
+to our simulations with ECC-lipids 
+(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
+The symmetry of the distribution of complexes from simulation 
+as shown in Figure~\ref{fig:cacl_complexes} --   
+i.e. almost equal probabilities of complexes with 1 or 3 lipids 
+that behave in the total average picture as complexes with 2 lipids -- 
+provides clues why ternary complex binding model 
+fits both simulation and experimental results relatively well, 
+although it is apparently incorrect. 
+It is also worth noting that many binding constants in the literature
+(e.g., Marsh's handbook \cite{marsh13}) 
+are based on the headgroup order parameter changes 
+intepreted with the ternary complex binding model.
+
+\todo{SAMULI: I think the we should make clear that the current binding
+  constants in, e.g., Marsh's handbook are based on the headgroup order parameter
+  changes intepreted with ternary complex binding model. I.e. the experimental raw data
+  is excatly the same as we have in this work. Our simulations enable much more versatile interpretation
+  of the binding phenomena. I think that this is one the main points of this work, so this comment
+  applies to the previous section, introduction, abstract and conclusions as well.
+  JOE: agree, this would be a good statement in general. I'll try to put it into the text.
+}
+
+Binding of \ce{Na+} cations to a POPC membrane 
+mostly follows stoichiomtetry \ce{Na^{+}}:$1\times$POPC 
+as shown in Figure~\ref{fig:cacl_complexes}.
+
 \section{Conclusions}
+
 By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
@@
   \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_NaCl.pdf}
-  \caption{\label{fig:delta_ordPar_NaCl}
+  \caption{\label{fig:delta_ordPar_NaCl_si}
     Changes of head group order parameters of POPC bilayer as a function of NaCl concentrations

### 2017-10-16 Manuscript/manuscript.tex
Some updates in the results and discussion, not ready yet

@@
 
-The concept can be used to compare the ion binding affinity in lipid 
+The concept can be used to compare the ion binding affinity to lipid 
 bilayers between MD simulations and  NMR experiments, because the order
@@
 $C_w$ and $C_i$ are their respective bulk concentrations in the aqueous phase;
-and $A_b$ is the size of the box in the membrane plane.
+and $A_b$ is the area of the box in the membrane plane.
 The total area of the interface is twice the area of the membrane, $2A_b$,
@@
 \mbox{x-ray} scattering experiments by calculating the order parameters
-for C-H bonds and the scattering form factors. The former validates the structures sampled by individual
+for C-H bonds and the scattering form factors. The former validates the structures sampled by the individual
 lipid molecules in simulations with atomic resolution, while the latter
@@
 Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion model
-was taken directly from \cite{lipid14POPC0mMNaClfiles}. 
+was taken directly from \cite{lipid14POPC0mMNaClfiles}. \todo{Which water model was used in these simulations?} 
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
@@
 A positive exception was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions,
-as seen from the data replotted from Ref.~\citenum{catte16} in Fig.~\ref{fig:delta_ordPar_CaCl}.
+as seen from the data replotted from Ref.~\citenum{catte16} in Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}.
 The model reproduced the experimentally measured small order parameter changes with NaCl
@@
 %polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
-Our simulations with the ECC-lipid and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
-exhibit an improved behaviour of cation binding to a POPC bilayer
+The results from the simulations with the ECC-lipid and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
+exhibit an improved behaviour of cation binding to a POPC bilayer in Fig.~\ref{fig:delta_ordPar_CaCl},
 showing a good agreement with experiments in the changes of 
-the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations
-in Fig.~\ref{fig:delta_ordPar_CaCl}. Since also the headgroup order parameter
+the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations.
+Since also the headgroup order parameter
 response the to the bound positive charge the ECC-lipid model
@@
 %description for the binding affinities of Na$^+$ and Ca$^{2+}$ ions to POPC lipid bilayer. 
-Furthermore, the overestimated headgroup order parameter changes of POPC in the Lipid14
+Furthermore, the overestimated lipid headgroup order parameter changes of POPC in the Lipid14
 model as a function of CaCl$_2$ concentration arise from both, the overestimated
@@
 
+\begin{table}
+  \caption{The relative surface excess and the average number of contacts of Calcium with different
+    oxygens in POPC with different simulation models \label{tab:binding} }
+  \begin{tabular}{l|c c c}
+    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+    \hline
+    ECC-lipids            &  $0.07 \pm 0.01 $            &  0.133   &    0.033    \\
+    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
+    Lipid14/Dang          &  $ 0.3 \pm 0.03  $           &  -        &    -       \\
+  \end{tabular}
+\end{table}
 To quantify the binding affinities in different models we calculated 
@@
 or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
-\todo{JOE: I'd put the $\Gamma$ values into a table}
+\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
+with contacts of calcium with different oxygens. However, I am not sure if there is enough information for the
+table unless we analyze the contacts for Lipid14 models as well.}
 Interestingly, the relative surface excess of 
@@
 meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
-This is in contrast to previous works employing classical MD simulations,
-which generally agree together on the existence of an increased sodium density at the membrane-water interface.
-\todo{JOE: Add references to NaCl simulation papers that show peaks.
-Add a simple note that there is NBFIX trick in Charmm36 (?).} 
+This is in contrast to the most of the available lipid force fields,
+which which predict a specific binding of sodium on 
+PC lipid bilayers \cite{catte16}.
+%\todo{JOE: Add references to NaCl simulation papers that show peaks.
+%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
+%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
 This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
@@
 
-We can estimate relative binding affinities 
-of several moieties in POPC towards \ce{Ca^{2+}} and \ce{Na+} cations 
-using the number of contacts per lipid between cations and POPC oxygen atoms, 
-$\Sigma ^\mathrm{Na^{2+}} _\mathrm{O}$,
-in which a contact between the two atoms is counted 
-if their distance is less than or equal $0.3 \, \mathrm{nm}$ 
-(robustness of this definition discussed in reference~\citenum{javanainen17}).
-
-For sodium at $1000\,mM$ concentration 
-the average number of contacts per lipid with any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
-whereas if only phosphate oxygen atoms are considered 
-this number decreases to 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
-and if only carbonyl oxygen atoms are counted 
-it is even lower, 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
-This means that interaction of sodium and POPC 
-has balanced contributions of 55\% from only phosphate oxygens 
-(i.e. no contacts with carbonyl oxyens),
-and from only carbonyl oxygens with 20\% of configurations 
-leaving 25\% for mixed phosphate and carbonyl oxygen configurations.  
-This is in line with the calculated negative value of 
-relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
-their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
-suggesting that the populations of bound sodium cations 
-naturally decrease with the decreasing density of water.
+The contacts between 
+%\ce{Ca^{2+}} 
+ions and different oxygen moieties of POPC,
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
+were analyzed by calculating the probability of 
+having the smaller or equal distance than $0.3 \, \mathrm{nm}$, 
+as done previously in Ref.~\citenum{javanainen17}.
 
@@
 
+and \ce{Na+}
+For sodium at $1000\,mM$ concentration 
+the average number of contacts per lipid with any oxygen atom in POPC is 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
+whereas if only phosphate oxygen atoms are considered 
+this number decreases to 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
+and if only carbonyl oxygen atoms are counted 
+it is even lower, 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
+This means that interaction of sodium and POPC 
+has balanced contributions of 55\% from only phosphate oxygens 
+(i.e. no contacts with carbonyl oxyens),
+and from only carbonyl oxygens with 20\% of configurations 
+leaving 25\% for mixed phosphate and carbonyl oxygen configurations.  
+This is in line with the calculated negative value of 
+relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
+their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
+suggesting that the populations of bound sodium cations 
+naturally decrease with the decreasing density of water.
+
 \subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in POPC membrane}

### 2017-10-17 Manuscript/manuscript.tex
Updates to the results and discussion, not ready

@@
 or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
-\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
-with contacts of calcium with different oxygens. However, I am not sure if there is enough information for the
-table unless we analyze the contacts for Lipid14 models as well.}
+\todo{JOE: I'd put the $\Gamma$ values into a table}
 Interestingly, the relative surface excess of 
@@
 meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
-This is in contrast to the most of the available lipid force fields,
-which which predict a specific binding of sodium on 
-PC lipid bilayers \cite{catte16}.
-%\todo{JOE: Add references to NaCl simulation papers that show peaks.
-%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
-%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
+This is in contrast to previous works employing classical MD simulations,
+which generally agree together on the existence of an increased sodium density at the membrane-water interface.
+\todo{JOE: Add references to NaCl simulation papers that show peaks.
+Add a simple note that there is NBFIX trick in Charmm36 (?).} 
 This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
@@
 
-The contacts between 
-%\ce{Ca^{2+}} 
-ions and different oxygen moieties of POPC,
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
-were analyzed by calculating the probability of 
-having the smaller or equal distance than $0.3 \, \mathrm{nm}$, 
-as done previously in Ref.~\citenum{javanainen17}.
+We can estimate relative binding affinities 
+of several moieties in POPC towards \ce{Ca^{2+}} and \ce{Na+} cations 
+using the number of contacts per lipid between cations and POPC oxygen atoms, 
+$\Sigma ^\mathrm{Na^{2+}} _\mathrm{O}$,
+in which a contact between the two atoms is counted 
+if their distance is less than or equal $0.3 \, \mathrm{nm}$ 
+(robustness of this definition discussed in reference~\citenum{javanainen17}).
+
+For sodium at $1000\,mM$ concentration 
+the average number of contacts per lipid with any oxygen atom in POPC is 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
+whereas if only phosphate oxygen atoms are considered 
+this number decreases to 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
+and if only carbonyl oxygen atoms are counted 
+it is even lower, 
+$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
+This means that interaction of sodium and POPC 
+has balanced contributions of 55\% from only phosphate oxygens 
+(i.e. no contacts with carbonyl oxyens),
+and from only carbonyl oxygens with 20\% of configurations 
+leaving 25\% for mixed phosphate and carbonyl oxygen configurations.  
+This is in line with the calculated negative value of 
+relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
+their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
+suggesting that the populations of bound sodium cations 
+naturally decrease with the decreasing density of water.
 
@@
 
-and \ce{Na+}
-For sodium at $1000\,mM$ concentration 
-the average number of contacts per lipid with any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
-whereas if only phosphate oxygen atoms are considered 
-this number decreases to 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
-and if only carbonyl oxygen atoms are counted 
-it is even lower, 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
-This means that interaction of sodium and POPC 
-has balanced contributions of 55\% from only phosphate oxygens 
-(i.e. no contacts with carbonyl oxyens),
-and from only carbonyl oxygens with 20\% of configurations 
-leaving 25\% for mixed phosphate and carbonyl oxygen configurations.  
-This is in line with the calculated negative value of 
-relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
-their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
-suggesting that the populations of bound sodium cations 
-naturally decrease with the decreasing density of water.
-
 \subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in POPC membrane}

### 2017-10-17 Manuscript/manuscript.tex
Updates to the results and discussion, not ready yet

--- 2017-10-16 Manuscript/manuscript.tex
@@
 \begin{table}
-  \caption{The relative surface excess and the average number of contacts of Calcium with different
-    oxygens in POPC with different simulation models \label{tab:binding} }
+  \caption{The relative surface excess and the probabilities of contacts between bound Calcium and different
+    oxygen moieties in POPC from different simulation models. The CaCl$_2$ concentration of the ions in water
+    is 350mM in the analyzed systems\label{tab:binding} }
   \begin{tabular}{l|c c c}
-    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
     \hline
-    ECC-lipids            &  $0.07 \pm 0.01 $            &  0.133   &    0.033    \\
+    ECC-lipids            &  $0.07 \pm 0.01 $            &  99\%   &    25\%    \\
     Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
@@
 
-The contacts between 
+The interactions between 
 %\ce{Ca^{2+}} 
@@
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
-were analyzed by calculating the probability of 
-having the smaller or equal distance than $0.3 \, \mathrm{nm}$, 
+were analyzed by calculating the probability of the
+contacts with the smaller or equal distance than $0.3 \, \mathrm{nm}$, 
 as done previously in Ref.~\citenum{javanainen17}. 
-
-In contrast to sodium, the dominant contribution
-to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
-% Numbers from simulation with 346mM conc. CaCl2
-The average number of contacts per lipid 
-between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
-whereas if only phosphate oxygen atoms are considered 
-the average number of contacts decreases only by a tiny amount to 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
-and if only carbonyl oxygen atoms are considered 
-%(phosphate oxygens contribute, but are not counted as contacts),
-this value is merely
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
+The results are shown in Table \ref{tab:binding}.
+%
+%SAMULI: I commented out the average number of contacts. All the essential information
+%for this is in the probabilities.
+%
+%In contrast to sodium, the dominant contribution
+%to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
+%% Numbers from simulation with 346mM conc. CaCl2
+%The average number of contacts per lipid 
+%between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
+%whereas if only phosphate oxygen atoms are considered 
+%the average number of contacts decreases only by a tiny amount to 
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
+%and if only carbonyl oxygen atoms are considered 
+%%(phosphate oxygens contribute, but are not counted as contacts),
+%this value is merely
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
+%
 % Numbers from simulation with 980/845mM conc. CaCl2
@@
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
-Hence, calcium is in contact with phosphate oxygens for 99\% of the time it is bound to the membrane, 
-whereas only 25\% of the time accounts for contacts with carbonyl oxygens,
-and interactions purely with carbonyl oxygens account for less than 1\% of configurations. 
-Higher concentration of \ce{CaCl2} naturally increases the amount of contacts per lipid,
-for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
-the average number of contacts per lipid is increased to
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
-but the distribution of states between phosphate and carbonyl oxygens remains unchanged. 
-We corroborate this finding with the probability isodensity contours in Fig.~\ref{fig:volmaps}, 
-and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.
+The probability for the bound Ca$^{2+}$ ions to be in contact with the phosphate oxygens
+is 99\%, indicating the interactions of Calcium ions purely with carbonyl oxygens has the probability of
+less than 1\%. However, the probability for the bound Ca$^{2+}$ ion to be in contact with
+carbonyl oxygen is 25\%, indicating that roughly quarter of the bound Calcium ions are in close contact
+with both, phosphase and carbonyl oxygens. It should be noted that in this analysis carbonyl and
+phospate may or may not belong to the same lipid, and the carbonyl chain locate in either of the 
+acyl chains. While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
+%for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
+%the average number of contacts per lipid is increased to
+%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
+the distribution of contacts between phosphate and carbonyl oxygens was not affected.
+The most likely contacts between Ca$^{2+}$ ions and phosphate oxygens is visualized with
+the probability isodensity contours in Fig.~\ref{fig:volmaps}.
+%and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.
 
@@
   \centering
-  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} \\
-  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
-  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
+  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
+%  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
+%  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
   \caption{\label{fig:volmaps}
       Contours of probability isodensities of \ce{Ca^{2+}} with respect to 
-      various moieties fixed in space (highlighted with transparent spheres): 
-      phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
-      Shown contours suggest that the dominant contribution 
-      to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
-      whereas the interactions with any of the two carbonyl groups are considerably milder.
+      the phosphate oxygens of POPC from ECC-lipid simulation.
+      %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
+      %Shown contours suggest that the dominant contribution 
+      %to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
+      %whereas the interactions with any of the two carbonyl groups are considerably milder.
   }
-  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca^{2+}}}
+  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca^{2+}}
+  SAMULI: I am not sure if this would be needed anymore.}
 \end{figure}
 
-and \ce{Na+}
-For sodium at $1000\,mM$ concentration 
-the average number of contacts per lipid with any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
-whereas if only phosphate oxygen atoms are considered 
-this number decreases to 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
-and if only carbonyl oxygen atoms are counted 
-it is even lower, 
-$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
-This means that interaction of sodium and POPC 
-has balanced contributions of 55\% from only phosphate oxygens 
-(i.e. no contacts with carbonyl oxyens),
-and from only carbonyl oxygens with 20\% of configurations 
-leaving 25\% for mixed phosphate and carbonyl oxygen configurations.  
-This is in line with the calculated negative value of 
-relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
-their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
-suggesting that the populations of bound sodium cations 
-naturally decrease with the decreasing density of water.
+Even thought the \ce{Na+} ions bind to POPC bilayer with the very low probability,
+it is possible to analyze the distribution of contacts from simulations with a large
+sodium concetration of $1000\,mM$.
+%the average number of contacts per lipid with any oxygen atom in POPC is 
+%$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
+%whereas if only phosphate oxygen atoms are considered 
+%this number decreases to 
+%$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
+%and if only carbonyl oxygen atoms are counted 
+%it is even lower, 
+%$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
+The results show that the sodium ions bind purely to the phosphate oxygens 
+of POPC with the probability of 55\% and purely to the carbonyl oxygens
+with the probability of 20\%, leaving the probability of 25\% for the
+simultaneous binding to both, phosphate and carbonyl oxygens.  
+%This is in line with the calculated negative value of 
+%relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
+%their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
+%suggesting that the populations of bound sodium cations 
+%naturally decrease with the decreasing density of water.
+
+In conclusion, the results suggest that the Calcium ions specifically bind to phosphate
+oxygens, while occasionally interacting also with carbonyls. This is in good agreement
+with the previous conclusions based in large amount of experimental and theoretical
+studies \cite{hauser76,hauser78,herbette84,cevc90,binder02}, but slightly updates the
+picture given by the previous MD simulation, where the binding to carbonyls was found
+to be more pronounced \cite{bockmann03,bockmann04,melcrova16,javanainen17}.
+Also the Sodium ions bind most likely to the phosphate oxygens, while the probability
+of binding purely to carbonyls is also significant. The physiological relevance
+of the details of sodium binding is, however, uncertain due to the very weak binding.
+
+\subsection{Stoichiometry of binding of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
+
+%Binding probability:
+%\ce{Ca2+} complex with \\
+%1 POPC - 13\% -- 30\% \\
+%2 POPC - 18\% -- 42\% \\
+%3 POPC - 12\% -- 28\% \\
+
+The previous interpretation of the experimental data for the changes of lipid headgroup
+order parameters, shown in Fig. \ref{fig:delta_ordPar_CaCl}, concluded that
+the experimental data was best explained by the ternary complex binding model, 
+where \ce{Ca^{2+}} ion bind to a POPC membrane with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}~\cite{altenbach84}.
+However, also a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
+was successfully fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
+These binding models and the measured headgroup order parameter changes
+are also inluded in the tabulated experimental binding constants~\cite{marsh13}.
+
+Since the experimentally measured changes of the POPC lipid headgroup order parameters
+as a function of CaCl$_2$ concentration are reproduced by the ECC-lipid model,
+it can be used to give more versatile interepration of binding stoichiometry than
+the simple binding models. The stoichiometry was directly evaluated from simulations
+by calculating the relative propensities of various \ce{Ca^{2+}}:$n\times$POPC complexes,
+which were defined to occur when cations and lipids were found within a cut off distance of $0.3\,\mathrm{nm}$. 
+The observed propensities of
+$42\%$ for two lipids,
+$30\%$ for one lipid, and
+$28\%$ for three lipids 
+are shown in Figure~\ref{fig:cacl_complexes} for \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration.
+\todo{Also the larger complexes were analysed, right? So we can write ''Larger complexes were not existent in simulation.'', or something like that?}
+While the ternary complex had the largest propensity,
+the complexes with 1~and 3~lipids occur with only a slightly lower probability,
+suggesting more complex binding stoichiometry than the simple ternary complex model.
+%The symmetry of the distribution of complexes,
+%from simulation as shown in Figure~\ref{fig:cacl_complexes} --   
+%i.e.,
+It should be, however, noted that on average one Calcium binds roughly to two lipids,
+because the probabilities of forming the complexes with 1 or 3 lipids are almost equal.  
+This probably explains the successfull fit of the ternary complex model to the
+experimental data. 
+\todo{I think that we either have to do the analysis in SI properly,
+  i.e., using the real $C_I$ from simulations or remove the sentence below and
+the related concent from SI.}
+Ternary complex model also provides a good fit 
+to our simulations with ECC-lipids 
+(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
+  \caption{\label{fig:cacl_complexes}
+      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
+      with a certain number of POPC lipids. 
+      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
+      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
+  }
+\end{figure}
+
+Also the probabilities of different complexes formed by \ce{Na+} ions and POPC
+analyzed from the ECC-lipid model simulation with 1 M concentration of NaCl are
+shown in Fig.~\ref{fig:cacl_complexes}. In contrast to Calcium, the Sodium ions
+most likely binds to one POPC, while the probability to bind two lipids is less likely
+and very small for more than two.
+%follows stoichiomtetry \ce{Na^{+}}:$1\times$POPCas shown in
 
@@
 is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
-From a recent theoretical work with long simulations \cite{javanainen17}
-this time can be roughly estimated 
-in the order of $1\,\mu\mathrm{s}$~\cite{javanainen17}. 
-Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
-we took the simulation with the apparent highest fluctuations of the number of bound calcium,
-which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
-and anaylsed the residence times.
+The recent MD simulations indicate that the equilibration of \ce{Ca^{2+}}
+binding to POPC bilayer may take hundreds of nanoseconds~\cite{javanainen17},
+which could become an issue for the detailed binding studies with MD simulation
+trajectories with the having lenght less than several microseconds \cite{catte16}.
+
+%give a roughly estimate of the order
+%of $1\,\mu\mathrm{s}$
+%This is unfortunately not the case of many theoretical works,
+%where the residence time cannot be estimated,
+%or the simulation has not even decorrelated from it's 
+%initial condition with a population of bound/unbound cations \cite{catte16}, 
+%because it is limited by the simulation length, 
+%even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
+%Increasing time scales to the order of $\mu\mathrm{s}$ 
+%doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used. 
+To analyze the residence time in the ECC-lipid model we \todo{describe how the residence time analysis was done}.
+Histograms of the residence times analyzed from the ECC-lipid simulation and from the CHARMM36 simulation data
+\todo{specify the system, POPC?}
+taken from the previous study \cite{javanainen17} (the simulation data was downloaded from Ref.~\citenum{zenodo.259376})
+are shown in Fig.~S\ref{fig:hist_residence_times} in SI. 
+\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
+ }
+%Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
+%we took the simulation with the apparent highest fluctuations of the number of bound calcium,
+%which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
+%and anaylsed the residence times.
 % simulation: (Matti C36 with CaClMK, 800ns total):
-We find out that
+
+\todo{This paragraph should be clarified, but I cannot do it, because the residence time figure is not there yet.}
+In the CHARMM36 simulation
 approximately 20\% of bound residence times of \ce{Ca^{2+}} 
@@
 60\% of bound residence time is smaller than half of the simulation length , $400\,\mathrm{ns}$. 
-Other simulations reported in that work show even longer residence times.
-
-This is in contrast to our model, ECC-lipids, 
-which estimates the residence time of calcium bound to a POPC membrane to be 
-smaller than $60\,\mathrm{ns}$ for 90\% of the time a cation is bound. % exactly $53\,\mathrm{ns}$
-i.e. at least two orders of magnitude lower than previous estimates.
-In addition, the longest observed residence time in our simulation is $120\,\mathrm{ns}$, 
+Other simulations with CHARMM36 and Slipids models reported in the previous work show even longer residence times. 
+The ECC-lipids model estimates 90\% of the residence times of calcium bound to a POPC membrane to be 
+smaller than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
+which is at least two orders of magnitude lower than for the CHARMM36.
+In addition, the longest observed residence time in the ECC-lipid simulation is $120\,\mathrm{ns}$, 
 which is well below the length of our simulations.
-This is unfortunately not the case of many theoretical works,
-where the residence time cannot be estimated,
-or the simulation has not even decorrelated from it's 
-initial condition with a population of bound/unbound cations \cite{catte16}, 
-because it is limited by the simulation length, 
-even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
-Increasing time scales to the order of $\mu\mathrm{s}$ 
-doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used. 
-Such a finding changes the point of view on binding of calcium to PC membranes from
-a very strong long-term stable binding with rare exchanges to 
-a relatively frequent exchange of clacium cations 
-in equilibrium between membrane and solvent.
 
@@
 is an order of magnitude faster than for calcium cations.
-A histogram of residence times from our simulations is in Fig.~S\ref{fig:hist_residence_times} in SI. 
-\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
- }
+
+Such a finding changes the point of view on binding of calcium to PC membranes from
+a very strong long-term stable binding with rare exchanges to 
+a relatively frequent exchange of clacium cations 
+in equilibrium between membrane and solvent. 
+The reduced barrier for the Calcium binding is also visible in the density distributions
+in Fig.~\ref{fig:cacl-dens}, where the more shallow minimum for the Calcuim
+density is observed at the interface.
 
@@
 %\end{figure}
-
-\subsection{Stoichiometry of binding of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
-
-%Binding probability:
-%\ce{Ca2+} complex with \\
-%1 POPC - 13\% -- 30\% \\
-%2 POPC - 18\% -- 42\% \\
-%3 POPC - 12\% -- 28\% \\
-
-Simulations allow us to directly evaluate the stoichiometry 
-by calculating relative propensities of various \ce{Ca^{2+}}:$n\times$POPC complexes
-by evaluating contacts between cations and lipids with a cut off radius $0.3\,\mathrm{nm}$. 
-The propensities of \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration 
-are in order:
-$42\%$ for two lipids,
-$30\%$ for one lipid, and
-$28\%$ for three lipids.
-Apart from a ternary complex, in which one calcium cation binds two lipids, 
-we also find complexes with 1~and 3~lipids
-occuring with only a slightly lower but similar probability
-as shown in Figure~\ref{fig:cacl_complexes}.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
-  \caption{\label{fig:cacl_complexes}
-      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
-      with a certain number of POPC lipids. 
-      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
-      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
-  }
-\end{figure}
-
-Several binding models were proposed and tested~\cite{altenbach84} of which only one,
-ternary complex binding model, provided a good fit of the experimental observations. 
-In such a model, it is assumed that \ce{Ca^{2+}} cations bind to a POPC membrane 
-with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
-In a later work~\cite{macdonald87}, 
-a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
-was found to provide as good fit as ternary complex model, 
-when only low concentrations of \ce{CaCl2} are considered.
-Ternary complex model also provides a good fit 
-to our simulations with ECC-lipids 
-(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
-The symmetry of the distribution of complexes from simulation 
-as shown in Figure~\ref{fig:cacl_complexes} --   
-i.e. almost equal probabilities of complexes with 1 or 3 lipids 
-that behave in the total average picture as complexes with 2 lipids -- 
-provides clues why ternary complex binding model 
-fits both simulation and experimental results relatively well, 
-although it is apparently incorrect. 
-It is also worth noting that many binding constants in the literature
-(e.g., Marsh's handbook \cite{marsh13}) 
-are based on the headgroup order parameter changes 
-intepreted with the ternary complex binding model.
-
-\todo{SAMULI: I think the we should make clear that the current binding
-  constants in, e.g., Marsh's handbook are based on the headgroup order parameter
-  changes intepreted with ternary complex binding model. I.e. the experimental raw data
-  is excatly the same as we have in this work. Our simulations enable much more versatile interpretation
-  of the binding phenomena. I think that this is one the main points of this work, so this comment
-  applies to the previous section, introduction, abstract and conclusions as well.
-  JOE: agree, this would be a good statement in general. I'll try to put it into the text.
-}
-
-Binding of \ce{Na+} cations to a POPC membrane 
-mostly follows stoichiomtetry \ce{Na^{+}}:$1\times$POPC 
-as shown in Figure~\ref{fig:cacl_complexes}.
 

### 2017-10-17 Manuscript/manuscript.tex
Updates in results and discussion, not ready yet

--- 2017-10-16 Manuscript/manuscript.tex
@@
 \end{table}
-To quantify the binding affinities in different models we calculated 
-the relative surface excess from Eq.~\ref{surfexcess}. 
-Calcium cations at a molar concentration of $350\,\mathrm{mM}$ provide
-$\Gamma_i^w = 0.07 \pm 0.01 \rm{nm}^{-2}$ in the ECC-lipid model, which is significantly
-smaller than the values for the Lipid14 model with \AA{}qvist ions, $\Gamma_i^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
-or with Dang ions, $\Gamma_i^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
+The binding affinities of Ca$^{2+}$ ions to POPC bilayer in different simulation 
+models were quantified by calculating the relative surface excess using Eq.~\ref{surfexcess} 
+and the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
+All these simulation have the same, $350\,\mathrm{mM}$, concentration of CaCl$_2$ with 
+respect to water (concentration before solvating the lipids in experiments).
+As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
+the relative surface excess for the ECC-lipid model, $\Gamma_{\rm Ca}^w = 0.07 \pm 0.01 \rm{nm}^{-2}$,
+is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
+or with Dang ions, $\Gamma_{\rm Ca}^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
 \todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
-with contacts of calcium with different oxygens. However, I am not sure if there is enough information for the
-table unless we analyze the contacts for Lipid14 models as well.}
+together with probabilities of the calcium binding probabilities to different POPC oxygens. 
+However, I am not sure if there the table is useful in the end.}
 Interestingly, the relative surface excess of 
-\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) is qualitatively different from \ce{CaCl2}
+\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) 
+calculated using density profiles in Fig.~\ref{fig:nacl-dens}
+is qualitatively different from \ce{CaCl2}
 having a negative value $\Gamma_i^w = -0.1 \pm 0.01 \rm{nm}^{-2}$, 
@@
 This is in contrast to the most of the available lipid force fields,
-which which predict a specific binding of sodium on 
+which predict a specific binding of sodium on 
 PC lipid bilayers \cite{catte16}.
@@
 %SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
-This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
-which show the density profile of ions and water along the membrane normal. 
+%
+%This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
+%which show the density profile of ions and water along the membrane normal. 
+%
 %The density profiles show
@@
 \todo{Densities and relative surface excess from NaCl systems to be added.
-  The discussion below to be finished after this.}
+  The discussion above to be finished after this.}
 

### 2017-10-18 Manuscript/manuscript.tex

@@
 Technically, it is a similar approach to the phenomenological charge-scaling as applied in 
-the early studies where a scaling factor one half was used \cite{jonsson86,egberts94}.
-\todo{We should also cite papers where empirical scaling was used ionic liquids - but there the factor is not 0.5.}
-However, the present concept of ECC is 
+earlier studies \cite{jonsson86,egberts94}.
+\todo{We should also cite papers where empirical scaling was used ionic liquids DOI: 10.1002/anie.201308760}
+The present concept of ECC is, however, 
 physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11}.
@@
 
-According to ECC, electronic
-polarizability can be implicitly included in classical MD simulations by
+Following the ECC concept, electronic
+polarizability can be effectively included in classical MD simulations by
 placing all particles into a homogeneous dielectric continuum 
@@
 the media \cite{leontyev11}. Measurements of high frequency 
-dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
-for almost any biomaterial \cite{some_original_work, leontyev11}.
-Such a dielectric continuum can be easily included in standard MD simulation by
-a formal transformation of partial charges 
+dielectric constant gives values around 2 
+for almost any biologically relevant material \cite{some_original_work, leontyev11}.
+Following the Coulomb's law, such a dielectric continuum can be easily included in standard MD simulation by
+a scaling of charges 
 \begin{equation}
@@
 effectively representing the newly introduced electronic continuum. 
-Assuming globally a  high frequency dielectric constant as measured in water (corresponding to the square of the refraction index),  $\epsilon _{el} = 1.78$, results 
-in a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
+Given the  high frequency dielectric constant as measured in water (corresponding to the square of the refraction index) of  $\epsilon _{el} = 1.78$,  
+the scaling factor for aqueous ions is $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
 successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??} which then agree quantitatively with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
 
-While the scaling factor of $f_q = 0.75$ for ions in water improves their performance and is physically justifiable within the ECC theory \cite{Leontyev2011},
-it is not clear whether the same factor should be used for partial charges in molecules,
-e.g., lipids in our case. Unlike the total charge of an atom or molecules, atomic partial charges within each molecule are not physical observables.
-There are several schemes for the assignment of partial charges for biomolecules. \cite{Hu2007}
+While the scaling factor of $f_q = 0.75$ for ions in water improves their description and is physically well justified within the ECC theory \cite{Leontyev2011},
+it is nota priori  clear whether the same factor should be used for partial charges in molecules,
+e.g., lipids in our case. Unlike the total charge of an atom or molecules, atomic partial charges within molecules are not physical observables.
+There are thus several schemes for the assignment of partial charges for biomolecules. \cite{Hu2007}
 Currently, the most commonly employed scheme is the restrained electrostatic potential (RESP) method~\cite{RESP_paper,Singh1984}.
@@
 %which targets at improving electrostatic interactions using a mean field correction. 
-In practice, partial charges currently implemented in force fields may already 
-include  to some extend some of the solvent electronic polarizability effects, i.e.,
-the RESP charges are often scaled to fit some experimental observables \todo{This needs a citation}. 
-Thus, we expect that the application of the ECC scaling factor, $f_q$, to the molecular partial charges
-included in the available force fields does not necessarily have to follow the relation $f_q = \epsilon _{el} ^{-1/2}$, 
-but instead it lies between 0.75 (no electronic polarizability in the partial charge calculation)
-and $1$ (full electronic polarizability already included in the partial charge calculation).
+By construction, partial charges currently implemented in force fields may already 
+include  to some extend the solvent electronic polarizability effects, i.e.,
+the RESP charges are often scaled to fit certain experimental observables \todo{This needs a citation - Joe please pick 1-2 papers}. 
+Thus, a consistent application of the ECC scaling factor, $f_q$, to the molecular partial charges
+included in the available force fields does not necessarily have to follow the above relation $f_q = \epsilon _{el} ^{-1/2}$, 
+but instead it lies between 0.75 i.e., (no electronic polarizability included in the original partial charges)
+and $1$ (i.e., electronic polarizability fully included in the original partial charges).
 %
@@
 of monovalent and divalent cations. This is a biologically highly relevant membrane feature, which is
-poorly reproduced by currently available models which can affect not only on membrane properties
-in the presence of ions but also modulates the interaction with charged moieties in the surface.
-Importantly, this response from simulations can be accurately
+poorly reproduced by currently available models. This can affect not only on membrane properties
+in the presence of ions but also modulate interactions with charged moieties at the interface.
+
+Importantly, results from simulations can be directly
 compared against experimental NMR data \cite{akutsu81,altenbach84,scherer89},
-as discussed in Ref. \citenum{catte16} in section \ref{section:electrometer}.
-To this end, we empirically explore the scaling factor parameter space, $f_q \in [0.75, 1.0]$ for the Lipid14~\cite{dickson14} force field.
+as discussed in detail in Ref. \citenum{catte16}.
+To this end, we explore the above discussed scaling factor parameter space, $f_q \in (0.75, 1.0)$ for the Lipid14~\cite{dickson14} force field.
 We selected this force field as a starting point because its response to bound ions was apparently the most realistic against NMR data
-in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
-Also glycerol backbone and head group structures in Lipid14 model were
-relatively realistic when compared with other state of the art lipid models \cite{botan15}.
+in a recent test within NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
+Moreover, the glycerol backbone and head group structures in Lipid14 model were
+more realistically described compared with other state of the art lipid models \cite{botan15}.
 The ECC correction was applied to Lipid14 parameters by scaling partial charges of the head group,
-glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
-are expected to have the largest contribution to the cation binding
-However, we do not modify the hydrocarbon chain parameters, because they
-do not come in contact with salt ions and are
-already highly optimized and give generally a good description for
-hydrophobic part of lipid bilayers in most lipids models, including Lipid14 \cite{ollila16}.
+glycerol backbone, and carbonyl regions, which are the most polar parts in the  lipids and
+are thus expected to have the largest contribution to the cation binding
+In contrast, we do not modify the hydrocarbon chain parameters, as they
+do not come in direct contact with salt ions and are
+already highly optimized to provide a good description for the
+hydrophobic part of lipid bilayers\cite{ollila16}.
 This is in contrasts with the behavior in glycerol backbone and head group order
-parameters which call for improvements in all available lipid models~\cite{botan15}.
-
-Exploring different scaling factor values, we found out that ion binding and
+parameters in the presence of ions, which call for improvements in all available lipid models~\cite{botan15}.
+
+Exploring different values of the scaling factor, we found out that ion bindings and the
 related head group order parameter responses become weakened in general. 
-The optimal behaviour of ion binding was observed with
-the scaling factor of $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
-\todo{JOE: following discussion shall be modified in the enlightenment of our recent ECC-discussions.}
-Interestingly, this scaling factor is in line with the estimate
-given by ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013}
-combined with RESP calculations in vacuum and implicit solvent reported
-in \cite{maciejewski14}. IPolQ charges are obtained as the average of
-partial charges given by RESP calculation~\cite{RESP_paper}
-in vacuum and in a solvent. Applying the scaling factor of 0.75 to
-IPolQ charges calculated from the data in Ref.~\cite{maciejewski14},
-gives similar partial charges to ones obtained by scaling Lipid14 charges
-with a factor 0.8.
+The optimal behaviour  was observed for a
+scaling factor of $f_q = 0.8$.This value is only slighly higher than the original scaling factor of 0.75, which implies
+that electronic polarization effects are only to a small extent effectively accounted for in the Lipid14 force field. 
 %-----\footnote----------
@@
 
-While, the charge scaling improved the behaviour of lipid-ion interactions,
-it reduced the area per molecule of the lipid bilayer without ions below experimental
+While, the above charge scaling improved the description of lipid-ion interactions,
+it also led to a reduction of the area per molecule of the lipid bilayer in water below experimental
 values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
-backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
-which is smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
-and the original Lipid14 value $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
-The decrease of the area per lipid was found to arise from a lower hydration of the lipid head group region,
-which can be explained by the increased solvation free energy due to the lower polarity
-of molecules with scaled charges. The hydration can be increased back
-by reducing the effective radius of atoms by changing the 
-the $\sigma$ parameters in the Lennard-Jones potential for the selected atoms
-similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
+backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$, as well as
+which is somewhat smaller than the experimental value of 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
+the Lipid14 value of $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
+The decrease of the area per lipid is found to arise from a reduced hydration of the lipid head group region,
+due to the lower polarity of molecules with scaled charges. This can be compensated
+by reducing the effective radius of atoms with scaled charges by changing the 
+the $\sigma$ parameters in the Lennard-Jones potential, as done previously for ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
 \todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules. \\
@@
 SAMULI: There is now a paragraph in the conclusions, which mentions this topic.}
-This decreases the solvation free energy by allowing water molecules to
-approach closer to lipid atoms and have stronger electrostatic interactions with them.
-After reducing the $\sigma$ parameters by a factor of $f_\sigma = 0.89$
-for the same atoms for which charges were scaled, the area per molecule value was
-back in agreement with the experimental value (see Table~\ref{tab:apls}).
+After reducing these $\sigma$ parameters by a factor of $f_\sigma = 0.89$, the area per molecule is restored
+back close to the experimental value (see Table~\ref{tab:apls}).
 

### 2017-10-18 Manuscript/manuscript.tex
Updates to the results and discussion

@@
     oxygen moieties in POPC from different simulation models. The CaCl$_2$ concentration of the ions in water
-    is 350mM in the analyzed systems\label{tab:binding} }
+    is 350mM in the analyzed systems. \label{tab:binding}}
   \begin{tabular}{l|c c c}
@@
 \end{table}
+
 The binding affinities of Ca$^{2+}$ ions to POPC bilayer in different simulation 
@@
 All these simulation have the same, $350\,\mathrm{mM}$, concentration of CaCl$_2$ with 
-respect to water (concentration before solvating the lipids in experiments).
+respect to water, i.e., the concentration before solvating the lipids in experiments.
 As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
@@
 or with Dang ions, $\Gamma_{\rm Ca}^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
-\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \label{tab:binding} 
+\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \ref{tab:binding} 
 together with probabilities of the calcium binding probabilities to different POPC oxygens. 
-However, I am not sure if there the table is useful in the end.}
+However, I am not sure if the table is useful in the end.}
 Interestingly, the relative surface excess of 
@@
 
-The interactions between 
-%\ce{Ca^{2+}} 
-ions and different oxygen moieties of POPC,
+The probabilities for a bound \ce{Ca^{2+}} ion to interact with
+different oxygen moieties of POPC 
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
-were analyzed by calculating the probability of the
-contacts with the smaller or equal distance than $0.3 \, \mathrm{nm}$, 
+were analyzed by counting the contacts within the distance of $0.3 \, \mathrm{nm}$, 
 as done previously in Ref.~\citenum{javanainen17}. 
-The results are shown in Table \ref{tab:binding}.
 %
@@
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
-The probability for the bound Ca$^{2+}$ ions to be in contact with the phosphate oxygens
-is 99\%, indicating the interactions of Calcium ions purely with carbonyl oxygens has the probability of
-less than 1\%. However, the probability for the bound Ca$^{2+}$ ion to be in contact with
-carbonyl oxygen is 25\%, indicating that roughly quarter of the bound Calcium ions are in close contact
-with both, phosphase and carbonyl oxygens. It should be noted that in this analysis carbonyl and
-phospate may or may not belong to the same lipid, and the carbonyl chain locate in either of the 
-acyl chains. While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
+The results in Table \ref{tab:binding} show that almost all (99\%) of the 
+bound Ca$^{2+}$ ions are in contact with the phosphate oxygens and a quarter
+of them (25\%) interacts with the carbonyl oxygens. Thus, the interactions between
+Calcium ions with only carbonyl oxygens are very unlikely (1\%), but significant
+amount of the ions interact with both, phosphate and carbonyl oxygens.
+The individual lipids or acyl chains were not identified in the analysis,
+thus the ion contacts with both oxygen moieties may be inter or intra molecular
+and occur with the carbonyls in the either of the acyl chains.
+The most likely contacts between Ca$^{2+}$ ions and phosphate oxygens is visualized with
+the probability isodensity contours in Fig.~\ref{fig:volmaps}.
+While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
 %for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
@@
 the distribution of contacts between phosphate and carbonyl oxygens was not affected.
-The most likely contacts between Ca$^{2+}$ ions and phosphate oxygens is visualized with
-the probability isodensity contours in Fig.~\ref{fig:volmaps}.
 %and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.
@@
 
-Even thought the \ce{Na+} ions bind to POPC bilayer with the very low probability,
-it is possible to analyze the distribution of contacts from simulations with a large
-sodium concetration of $1000\,mM$.
+Even thought \ce{Na+} ions bind to POPC bilayer with a very low probability,
+it was possible to analyze the distribution of contacts from simulations with a large
+sodium concetration of~1\,M.
 %the average number of contacts per lipid with any oxygen atom in POPC is 
@@
 oxygens, while occasionally interacting also with carbonyls. This is in good agreement
-with the previous conclusions based in large amount of experimental and theoretical
-studies \cite{hauser76,hauser78,herbette84,cevc90,binder02}, but slightly updates the
-picture given by the previous MD simulation, where the binding to carbonyls was found
-to be more pronounced \cite{bockmann03,bockmann04,melcrova16,javanainen17}.
+with the previous conclusions from several experimental and theoretical
+studies~\cite{hauser76,hauser78,herbette84,cevc90,binder02}, but
+suggests lower relative binding affinity to carbonyls than the previous MD simulation
+studies~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
 Also the Sodium ions bind most likely to the phosphate oxygens, while the probability
-of binding purely to carbonyls is also significant. The physiological relevance
-of the details of sodium binding is, however, uncertain due to the very weak binding.
+of binding purely to carbonyls is also significant, in contrast to Calcium ions.
+The physiological relevance of the Sodium binding details is, however, uncertain
+due to the very weak binding affinity.
 
@@
 
-The previous interpretation of the experimental data for the changes of lipid headgroup
-order parameters, shown in Fig. \ref{fig:delta_ordPar_CaCl}, concluded that
-the experimental data was best explained by the ternary complex binding model, 
-where \ce{Ca^{2+}} ion bind to a POPC membrane with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}~\cite{altenbach84}.
-However, also a Langmuir adsorption model (i.e. stoichiometry 1~POPC:1~\ce{Ca^{2+}}) 
+The same experimental data for the lipid headgroup order parameters
+used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl},
+has been previous interpreted by using simple binding models to determine the
+binding constants and stoichimetry of \ce{Ca^{2+}} binding to PC lipid
+bilayers~\cite{altenbach84,macdonald87}. The experimental data was best
+explained by the ternary complex binding model with a binding
+stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
+However, also a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
 was successfully fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
-These binding models and the measured headgroup order parameter changes
+The results based on these binding models and the measured headgroup order parameter changes
 are also inluded in the tabulated experimental binding constants~\cite{marsh13}.
@@
 as a function of CaCl$_2$ concentration are reproduced by the ECC-lipid model,
-it can be used to give more versatile interepration of binding stoichiometry than
-the simple binding models. The stoichiometry was directly evaluated from simulations
-by calculating the relative propensities of various \ce{Ca^{2+}}:$n\times$POPC complexes,
-which were defined to occur when cations and lipids were found within a cut off distance of $0.3\,\mathrm{nm}$. 
-The observed propensities of
-$42\%$ for two lipids,
-$30\%$ for one lipid, and
-$28\%$ for three lipids 
-are shown in Figure~\ref{fig:cacl_complexes} for \ce{Ca^{2+}}:$n\times$POPC complexes at 285~mM concetration.
+it can be used to give a more versatile interepration of the binding stoichiometry than
+the previously used simple binding models. To directly evaluate the
+stoichiometry from simulations we calculated the relative propensities of \ce{Ca^{2+}}:POPC
+complexes with various stoichiometries by counting the individual lipid
+molecules within the distance of $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion.
+The results from POPC simulation with 285~mM concetration of CaCl$_2$ in
+Fig.~\ref{fig:cacl_complexes} show the largest propensity (42\%) for the
+ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
+the probabilities for the complexes with stoichimetries of 1:1~(30\%)~and
+1:3~(28\%)~are only slightly lower,
+suggesting more complex binding stoichiometry than the simple ternary complex model.
+%The observed propensities of
+%$42\%$ for two lipids,
+%$30\%$ for one lipid, and
+%$28\%$ for three lipids 
 \todo{Also the larger complexes were analysed, right? So we can write ''Larger complexes were not existent in simulation.'', or something like that?}
-While the ternary complex had the largest propensity,
-the complexes with 1~and 3~lipids occur with only a slightly lower probability,
-suggesting more complex binding stoichiometry than the simple ternary complex model.
 %The symmetry of the distribution of complexes,
@@
 %i.e.,
-It should be, however, noted that on average one Calcium binds roughly to two lipids,
+However, one Calcium binds roughly to two lipids on average also in simulations,
 because the probabilities of forming the complexes with 1 or 3 lipids are almost equal.  
@@
 analyzed from the ECC-lipid model simulation with 1 M concentration of NaCl are
-shown in Fig.~\ref{fig:cacl_complexes}. In contrast to Calcium, the Sodium ions
-most likely binds to one POPC, while the probability to bind two lipids is less likely
-and very small for more than two.
+shown in Fig.~\ref{fig:cacl_complexes}. In contrast to Calcium, the
+probability is largest (??\%) for 1:1 complex, significantly smaller (??\%)
+for 1:2 complexes and very small (??\%) for 1:3 of \ce{Na+}:POPC complexes.
 %follows stoichiomtetry \ce{Na^{+}}:$1\times$POPCas shown in 
+\todo{Add numbers for Sodium.}
 
@@
 is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
+\todo{This experimental estimate is actually for the residence time of Calcium
+  ion at the individual PC lipid. This section is to be finished when the analysis details
+and the results are here.}
 The recent MD simulations indicate that the equilibration of \ce{Ca^{2+}}

### 2017-10-19 scripts/surface_excess_calculate_oMM.sh
Adding rel.-surface-excess script adapted to openMM.

@@ -1,0 +1,88 @@
+#!/bin/bash
+
+# Gromacs 5.x version of ...
+# bash wrapper for calculating relative surface excess of a lipid bilayer 
+# using
+# python script/library get_conc_ion_bulk.py
+# meant for use with NMRlipids projects
+#------------------------------------------------------------
+# Made by J.Melcr,  Last edit 2017/08/21
+#------------------------------------------------------------
+
+scriptdir=`dirname $0`
+
+traj_file_name="traj_openMM.dcd" #"../traj.trr" 
+#traj_pbc_nonwat_file_name="traj_nonwat_pbc.xtc" #"../traj.trr" 
+#top_file_name="last_frame_nonwat.gro"
+tpr_file_name="topol.tpr"
+dens_file_name="density_ca_cl_water.xvg"
+#op_def_file="../../Headgroup_Glycerol_OPs.def"
+#op_def_file=${scriptdir}"/order_parameter_definitions_Lipid14_POPC_all.def"
+#op_out_file="OrdPars.dat"
+top="topol.top"
+f_conc=55430  # in mM/L
+
+if ! [ -s $tpr_file_name ] 
+then
+    echo "We really need " $tpr_file_name " , but we can't find it!"
+    exit 1
+fi
+
+# get densities of cation - anion - water centered around POPC
+# and start at 100ns
+if ! [ -s $dens_file_name ] 
+then
+    seq 2 5 | gmx density -sl 900 -dens number -ng 3 -f $traj_file_name -center -symm -relative -o $dens_file_name -xvg none #-b 100000 
+fi
+
+# calculate area and area-per-lipid
+area=`python $scriptdir/get_apl_fromDCD.py | grep -e "APL" | cut -d " " -f3`
+echo area = $area nm2
+
+# getting actual bulk concentration from number density profile
+python $scriptdir/get_conc_ion_bulk.py 
+conc=`cut -d " " -f1 conc_ion_bulk_mmolL.dat`
+conc_anion=`cut -d " " -f1 conc_anion_bulk_mmolL.dat`
+conc_wat=`cut -d " " -f1 conc_wat_bulk_mmolL.dat`
+
+
+#getting nominal concentration from topol.top file (if exists)
+if [ -f $top ]
+then
+    nwat=`grep -e "molecules" -A 10 $top | grep -e "^SOL" -e "^TIP" -e "^SPCE" -e "^OPC3" | cut -d " " -f1 --complement `
[…]

### 2017-10-19 Manuscript/manuscript.tex
Add references for the cationic surfactant simulations
Some updates in binding affinity details and stoichiometry sections. I will continue now with the residence times.

@@
 to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
-The parameters are available at \cite{??} \todo{To be added}.
+The parameters together with the simulation data are available
+at~\cite{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
 The surfactants were randomly placed among the lipids to form bilayer structures with
@@
 
-\subsection{Cation binding affinities to POPC membrane validated through lipid electrometer}
+\subsection{Binding affinities to POPC membrane validated through lipid electrometer}
 
@@
 
-\subsection{Cation binding to POPC membrane in atomistic detail}
+\subsection{Affinity and binding cites to POPC membrane}
 
@@
 The binding affinities of Ca$^{2+}$ ions to POPC bilayer in different simulation 
-models were quantified by calculating the relative surface excess using Eq.~\ref{surfexcess} 
-and the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
-All these simulation have the same, $350\,\mathrm{mM}$, concentration of CaCl$_2$ with 
+models were quantified by calculating the relative surface excess from Eq.~\ref{surfexcess} 
+using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
+The relative surface excess values, shown in Table~\ref{tab:binding},
+were calculated from simulations with the same, $350\,\mathrm{mM}$, concentration of CaCl$_2$ with 
 respect to water, i.e., the concentration before solvating the lipids in experiments.
@@
 The individual lipids or acyl chains were not identified in the analysis,
-thus the ion contacts with both oxygen moieties may be inter or intra molecular
+thus the simultaneous ion contacts with both oxygen moieties may be inter or intra molecular
 and occur with the carbonyls in the either of the acyl chains.
@@
 %$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
-The results show that the sodium ions bind purely to the phosphate oxygens 
-of POPC with the probability of 55\% and purely to the carbonyl oxygens
-with the probability of 20\%, leaving the probability of 25\% for the
-simultaneous binding to both, phosphate and carbonyl oxygens.  
+The results show that \ce{Na+} ions bind only to the phosphate oxygens 
+of POPC with the probablity of 55\% and only to the carbonyl oxygens with the
+probability of 20\%, leaving the probability of 25\% for the
+simultaneous binding to both.%, phosphate and carbonyl oxygens.  
 %This is in line with the calculated negative value of 
@@
 
-\subsection{Stoichiometry of binding of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
+\subsection{Binding stoichiometry to POPC membrane}
 
@@
 
-The same experimental data for the lipid headgroup order parameters
-used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl},
-has been previous interpreted by using simple binding models to determine the
-binding constants and stoichimetry of \ce{Ca^{2+}} binding to PC lipid
-bilayers~\cite{altenbach84,macdonald87}. The experimental data was best
+Simple binding models have been previous used to interpreted 
+the same experimental data, which is used in this work to validate
+the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}~\cite{altenbach84,macdonald87}.
+The experimental results from the lipid headgroup order parameters and from the
+atomic absorption spectroscopy were best
 explained by the ternary complex binding model with a binding
@@
 
-Since the experimentally measured changes of the POPC lipid headgroup order parameters
-as a function of CaCl$_2$ concentration are reproduced by the ECC-lipid model,
+Since the same experimental data is reproduced by the ECC-lipid model,
 it can be used to give a more versatile interepration of the binding stoichiometry than
-the previously used simple binding models. To directly evaluate the
+the previously used simple binding models~\cite{altenbach84,macdonald87}. To directly evaluate the
 stoichiometry from simulations we calculated the relative propensities of \ce{Ca^{2+}}:POPC
@@
 ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
-the probabilities for the complexes with stoichimetries of 1:1~(30\%)~and
+the probabilities for the complexes with the stoichimetries of 1:1~(30\%)~and
 1:3~(28\%)~are only slightly lower,
@@
 %$28\%$ for three lipids 
-\todo{Also the larger complexes were analysed, right? So we can write ''Larger complexes were not existent in simulation.'', or something like that?}
-%The symmetry of the distribution of complexes,
-%from simulation as shown in Figure~\ref{fig:cacl_complexes} --   
-%i.e.,
 However, one Calcium binds roughly to two lipids on average also in simulations,
@@
 experimental data. 
+Complexes with more than three lipids per one Calcium ion were not observed in 
+simulations with the ECC-lipids model.
 \todo{I think that we either have to do the analysis in SI properly,

### 2017-10-19 Manuscript/manuscript.tex
New residence time section

@@
 
-The residence time of \ce{Ca^{2+}} in a POPC membrane 
-is experimentally estimated to be lower than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
-\todo{This experimental estimate is actually for the residence time of Calcium
-  ion at the individual PC lipid. This section is to be finished when the analysis details
-and the results are here.}
-The recent MD simulations indicate that the equilibration of \ce{Ca^{2+}}
-binding to POPC bilayer may take hundreds of nanoseconds~\cite{javanainen17},
-which could become an issue for the detailed binding studies with MD simulation
-trajectories with the having lenght less than several microseconds \cite{catte16}.
+Equilibration of \ce{Ca^{2+}} binding to POPC bilayer takes hundreds of
+nanoseconds in MD simulations with CHARMM36 and Slipids force fields~\cite{javanainen17}, suggesting that
+simulations with the length of several microseconds would be required
+to study the binding phenomena in equilibrium. However, the density profiles of \ce{Ca^{2+}}
+from the ECC-lipid simulation in Fig.~\ref{fig:cacl-dens} show a relatively shallow
+minimum at the interface, suggesting a smaller barrier for the exchange of Calcium ions.
+To quantify the ion binding exchange in lipid bilayer simulations, we calculated 
+the residence times of ions within the distance of 0.3~nm from any oxygen atom
+belonging to a lipid molecule in a bilayer.
+
+The residence time histograms of \ce{Ca^{2+}}
+in POPC bilayer are shown in Fig.~S\ref{fig:hist_residence_times} in SI 
+from the ECC-lipid simulation with CaCl$_2$ concentration of 1~M
+and from the CHARMM36 simulation with 450~mM concentration of
+ECC-ions taken from the previous study \cite{javanainen17}
+(the simulation data was downloaded from Ref.~\citenum{zenodo.259376}).
+\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
+}
+\todo{It might be also better to do the analysis from ECC-lipid simulation with concentration closer
+to the data in literature, i.e. 450mM.}
+%Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
+%we took the simulation with the apparent highest fluctuations of the number of bound calcium,
+%which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
+%and anaylsed the residence times.
+% simulation: (Matti C36 with CaClMK, 800ns total):
+Approximately 20\% of the \ce{Ca^{2+}} residence times in the CHARMM36 simulation are longer than
+the lenght of the trajectory ($800\,\mathrm{ns}$),
+and only less than 60\% of the bound residence times are shorter than the half
+of the simulation length~(400~ns).
+Even longer residence times are observed
+in the other simulations with CHARMM36 and Slipids models reported in the previous work~\cite{javanainen17}.
+At least order of magnitude faster exchange is observed in the
+ECC-lipid model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are 
+shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
+and the longest observed residence time in the ECC-lipid simulation is $120\,\mathrm{ns}$, 
+which is well below the total length of the simulation (200~ns).
+The exchange of \ce{Na+} ions to a POPC membrane is
+an order of magnitude faster, giving 90\%
+of the residence times smaller than~$2\,\mathrm{ns}$,
+the longest residence time being~$13\,\mathrm{ns}$. 
+All the results are in line with the experimental estimation that 
+the residence time of \ce{Ca^{2+}} at the individual PC headgroup 
+is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
 
@@
 %doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used.
-To analyze the residence time in the ECC-lipid model we \todo{describe how the residence time analysis was done}.
-Histograms of the residence times analyzed from the ECC-lipid simulation and from the CHARMM36 simulation data
-\todo{specify the system, POPC?}
-taken from the previous study \cite{javanainen17} (the simulation data was downloaded from Ref.~\citenum{zenodo.259376})
-are shown in Fig.~S\ref{fig:hist_residence_times} in SI. 
-\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
- }
-%Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
-%we took the simulation with the apparent highest fluctuations of the number of bound calcium,
-%which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
-%and anaylsed the residence times.
-% simulation: (Matti C36 with CaClMK, 800ns total):
-
-\todo{This paragraph should be clarified, but I cannot do it, because the residence time figure is not there yet.}
-In the CHARMM36 simulation
-approximately 20\% of bound residence times of \ce{Ca^{2+}} 
-are limited by the length of the simulation, $800\,\mathrm{ns}$, 
-and only less than
-60\% of bound residence time is smaller than half of the simulation length , $400\,\mathrm{ns}$. 
-Other simulations with CHARMM36 and Slipids models reported in the previous work show even longer residence times. 
-The ECC-lipids model estimates 90\% of the residence times of calcium bound to a POPC membrane to be 
-smaller than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
-which is at least two orders of magnitude lower than for the CHARMM36.
-In addition, the longest observed residence time in the ECC-lipid simulation is $120\,\mathrm{ns}$, 
-which is well below the length of our simulations.
-
-Interestingly, \ce{Na+} cations exhibit much more rapid exchange between membrane and solvent than calcium.
-The residence time of sodium bound to a POPC membrane is
-smaller than $2\,\mathrm{ns}$ for 90\% of the time it is bound to the membrane 
-with the longest measured bound time being $13\,\mathrm{ns}$. 
-Such an exchange between membrane and solvent 
-is an order of magnitude faster than for calcium cations.
-
-Such a finding changes the point of view on binding of calcium to PC membranes from
-a very strong long-term stable binding with rare exchanges to 
-a relatively frequent exchange of clacium cations 
-in equilibrium between membrane and solvent. 
-The reduced barrier for the Calcium binding is also visible in the density distributions
-in Fig.~\ref{fig:cacl-dens}, where the more shallow minimum for the Calcuim
-density is observed at the interface.
+
+In conclusion, the results from ECC-lipid model suggest that the
+exchange of bound Calcium occurs within $\sim$100ns timeframe,
+which is much faster than previously reported~\cite{javanainen17}.
+The sodium exhibit even faster exchange.
+This suggests that simulations with the length of few hundred
+nanoseconds are sufficient to simulate equilibrium ion binding
+to PC lipid bilayers, when realistic force fields are used.
+This has not been the case with the available lipid force fields,
+which overestimate the binding streght of the cations \cite{javanainen17,catte16}.
+%Such a finding changes the point of view on binding of calcium to PC membranes from
+%a very strong long-term stable binding with rare exchanges to 
+%a relatively frequent exchange of clacium cations 
+%in equilibrium between membrane and solvent.
 

### 2017-10-19 Manuscript/manuscript.tex
Updates in ECC for lipids section, not ready

@@
 Prague 6, Czech Republic}
+\author{Ji{\v r}{\' i} Kolafa}
+\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
 \author{Pavel Jungwirth}
@@
 lipid bilayer simulations.
-%Some pioneering simulation studies scaled
-%the partial charges of atoms by one half to effectively include 
-%polarizability \cite{jonsson86,egberts94}. Also, approaches that
-%explicitly include polarizability has been introduced, but
-%this has turned out to be practically complicated \cite{chowdhary13}.
 In this work we circumvent the rather demanding explicit inclusion
-of electronic polarization effects \cite{chowdhary13} by implicitly including
+of electronic polarization effects \cite{lucas12,chowdhary13} by implicitly including
 electronic polarizability in lipid
@@
 physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11}.
-%but significantly more simple than the explicit inclusion of electronic
-%polarizability \cite{chowdhary13} both for derivation and application.
 
@@
 which is the electronic part of the dielectric constant of 
-the media \cite{leontyev11}. Measurements of high frequency 
-dielectric constant gives values around 2 
-for almost any biologically relevant material \cite{some_original_work, leontyev11}.
+the media \cite{leontyev11}. 
 Following the Coulomb's law, such a dielectric continuum can be easily included in standard MD simulation by
@@
 the scaling factor for aqueous ions is $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
-successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??} which then agree quantitatively with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??},
+which then agree quantitatively with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
 
 While the scaling factor of $f_q = 0.75$ for ions in water improves their description and is physically well justified within the ECC theory \cite{Leontyev2011},
-it is nota priori  clear whether the same factor should be used for partial charges in molecules,
+it is not a priori  clear whether the same factor should be used for partial charges in molecules,
 e.g., lipids in our case. Unlike the total charge of an atom or molecules, atomic partial charges within molecules are not physical observables.
-There are thus several schemes for the assignment of partial charges for biomolecules. \cite{Hu2007}
+There are thus several schemes for the assignment of partial charges for biomolecules~\cite{Hu2007}.
 Currently, the most commonly employed scheme is the restrained electrostatic potential (RESP) method~\cite{RESP_paper,Singh1984}.
-%
-%PAVEL: i do not understand what you mean by "ensemble" - resp can be done for a single configuration of a single molecule.
-%In any case, tehre is no need to explain here how RESP works so i would drop this part (including the next two sentences).
-%
-%SAMULI: I commented out the suggested sentences
-%which aims to reproduce the electrostatic potential obtained from quantum mechanical calculations around an ensemble the target molecules. 
-%The fixed partial atomic charges obtained by this method are hence 
-%the best fit of pint charges reproducing the electrostatic potential around provided target conformations. 
-%Such a scheme is well suited for the application of ECC, 
-%which targets at improving electrostatic interactions using a mean field correction. 
 By construction, partial charges currently implemented in force fields may already 
@@
 and $1$ (i.e., electronic polarizability fully included in the original partial charges).
-%
-% This is not really important for the paper:
-%
-%However, the RESP method has several remedies;  %and these are specific to RESP
-%RESP charges show a dependence on the orientation and position
-%and it does not provide numerically stable partial charges of burried atoms;~\cite{Hu2007}
-%the RESP-derived values depend, for example, on the conformational
-%sampling of the molecules and description of solvent 
-%in the calculations~\cite{RESP_paper, Singh1984, Hu2007, maciejewski14}. 
-%
-%Hirshfeld-E partitioning: \cite{Verstraelen2013}
-%In cases where the partial charges have some physical meaning 
-%(e.g. approximate electrostatic potential as in RESP~~\cite{RESP_paper,Singh1984}),
-%
-%
-
-In this work, we develop a phospholipid model for classical MD simulations that
-accurately describes the lipid head group response to varying concentrations
-of monovalent and divalent cations. This is a biologically highly relevant membrane feature, which is
-poorly reproduced by currently available models. This can affect not only on membrane properties
-in the presence of ions but also modulate interactions with charged moieties at the interface.
-
-Importantly, results from simulations can be directly
-compared against experimental NMR data \cite{akutsu81,altenbach84,scherer89},
-as discussed in detail in Ref. \citenum{catte16}.
-To this end, we explore the above discussed scaling factor parameter space, $f_q \in (0.75, 1.0)$ for the Lipid14~\cite{dickson14} force field.
-We selected this force field as a starting point because its response to bound ions was apparently the most realistic against NMR data
-in a recent test within NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
-Moreover, the glycerol backbone and head group structures in Lipid14 model were
-more realistically described compared with other state of the art lipid models \cite{botan15}.
-The ECC correction was applied to Lipid14 parameters by scaling partial charges of the head group,
+It is important to note that even thought the lipid bilayer interface has a large gradient in
+the {\it total} dielectric constant, the electronic polarizability part can be considered to be constant
+as the measurements of high frequency dielectric constant gives values around 2 
+for almost any biologically relevant material \cite{some_original_work, leontyev11}.
+
+In this work, we employ the ECC to develop a classical MD simulation model for
+a POPC lipid that accurately describes the binding of Sodium and Calcium ions to a lipid bilayer.
+%This is a biologically highly relevant membrane feature, which is
+%poorly reproduced by currently available models. This can affect not only on membrane properties
+%in the presence of ions but also modulate interactions with charged moieties at the interface.
+The quality of the model in terms of ion binding is measured by directly
+comparing simulations to the experimental NMR data used to measure the ion binding
+affinity~\cite{akutsu81,altenbach84,scherer89}, as discussed in detail in Ref. \citenum{catte16}.
+As a starting point we used the parameters of Lipid14~\cite{dickson14} force field,
+available in Gromacs format at \citenum{lipid14files}, because they apparently give the most
+realistic response to the bound ions and to the NaCl concentration 
+among the available lipid models (see Figs.~5~and~2 in Ref.~\citenum{catte16}).
+Moreover, the Lipid14 model gave relatively realistic glycerol backbone and
+head group structures when compared with the other state of the art lipid models~\cite{botan15}.
+
+The ECC correction was applied to the Lipid14 parameters by scaling partial charges of the head group,
 glycerol backbone, and carbonyl regions, which are the most polar parts in the  lipids and
-are thus expected to have the largest contribution to the cation binding
+are thus expected to have the largest contribution to the cation binding.
 In contrast, we do not modify the hydrocarbon chain parameters, as they
@@
 
+To find the best scaling factor to reproduce the ion binding affinity,
+we explored the above discussed scaling factor parameter space, $f_q \in (0.75, 1.0)$,
+and compared the results to the available NMR data~\cite{akutsu81,altenbach84,scherer89}.
 Exploring different values of the scaling factor, we found out that ion bindings and the
@@
 that electronic polarization effects are only to a small extent effectively accounted for in the Lipid14 force field.
-%-----\footnote----------
-  %{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
-  %types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
-  %partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
-  %I use 0.80 as an approximate round value, which accounts for the already lower charges in Lipid14. 
-  %The use of 0.78---.79 might be more appropriate, though.}
 

### 2017-10-19 Manuscript/manuscript.tex
New version of the ECC-lipid section

@@
 It is important to note that even thought the lipid bilayer interface has a large gradient in
-the {\it total} dielectric constant, the electronic polarizability part can be considered to be constant
+the total dielectric constant, the electronic polarizability part can be considered to be constant
 as the measurements of high frequency dielectric constant gives values around 2 
@@
 %in the presence of ions but also modulate interactions with charged moieties at the interface.
-The quality of the model in terms of ion binding is measured by directly
+The model quality in terms of the ion binding is measured by directly
 comparing simulations to the experimental NMR data used to measure the ion binding
 affinity~\cite{akutsu81,altenbach84,scherer89}, as discussed in detail in Ref. \citenum{catte16}.
-As a starting point we used the parameters of Lipid14~\cite{dickson14} force field,
-available in Gromacs format at \citenum{lipid14files}, because they apparently give the most
+The Lipid14~\cite{dickson14} force field parameters (available in Gromacs format from Ref. \citenum{lipid14files})
+were used as a starting point, because they give the most
 realistic response to the bound ions and to the NaCl concentration 
@@
 are thus expected to have the largest contribution to the cation binding.
-In contrast, we do not modify the hydrocarbon chain parameters, as they
+We do not modify the hydrocarbon chain parameters, as they
 do not come in direct contact with salt ions and are
 already highly optimized to provide a good description for the
-hydrophobic part of lipid bilayers\cite{ollila16}.
-This is in contrasts with the behavior in glycerol backbone and head group order
-parameters in the presence of ions, which call for improvements in all available lipid models~\cite{botan15}.
-
-To find the best scaling factor to reproduce the ion binding affinity,
-we explored the above discussed scaling factor parameter space, $f_q \in (0.75, 1.0)$,
-and compared the results to the available NMR data~\cite{akutsu81,altenbach84,scherer89}.
-Exploring different values of the scaling factor, we found out that ion bindings and the
-related head group order parameter responses become weakened in general. 
-The optimal behaviour  was observed for a
-scaling factor of $f_q = 0.8$.This value is only slighly higher than the original scaling factor of 0.75, which implies
-that electronic polarization effects are only to a small extent effectively accounted for in the Lipid14 force field.
-
-While, the above charge scaling improved the description of lipid-ion interactions,
-it also led to a reduction of the area per molecule of the lipid bilayer in water below experimental
-values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
-backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$, as well as
-which is somewhat smaller than the experimental value of 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
-the Lipid14 value of $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
+hydrophobic part of lipid bilayers \cite{ollila16}.
+This is in contrasts with the glycerol backbone and the head group regions, 
+which call for improvements in all available lipid models~\cite{botan15}.
+
+The values within the above discussed parameter space for the scaling factor, $f_q~\in~(0.75, 1.0)$,
+were explored to find the optimal value to reproduce the experimental ion binding affinity.
+We observed that the scaling
+of the partial charges reduced the ion binding affinity and the 
+related head group order parameter response in general.
+By comparing the results with different scaling factors to the experimental 
+NMR data~\cite{akutsu81,altenbach84,scherer89} we found the optimal 
+value of $f_q = 0.8$,
+which is only slighly higher than the scaling factor for the ions in water 0.75.
+%, which implies
+%that electronic polarization effects are only to a small extent effectively accounted for in the Lipid14 force field.
+
+While, the scaling of the partial charges improved the ion binding affinity,
+it reduced the bilayer area per lipid below the experimental value. 
+%Simulations with Lipid14 parameters having 
+After scaling the partial charges of the head group, the glycerol
+backbone and carbonyls with a factor of 0.8, the simulation gave 
+the approximately 60~\AA$^2$ for the area per lipid,
+which is somewhat smaller than the experimental value, 64.3~\AA$^2$ \cite{kucerka11}, 
+as well as the value given by the original Lipid14 parameters, $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
 The decrease of the area per lipid is found to arise from a reduced hydration of the lipid head group region,
 due to the lower polarity of molecules with scaled charges. This can be compensated
-by reducing the effective radius of atoms with scaled charges by changing the 
-the $\sigma$ parameters in the Lennard-Jones potential, as done previously for ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
+by reducing the effective radius of atoms with the scaled charges by changing 
+the $\sigma$ parameters in the Lennard-Jones potential, 
+as previously done for ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
+After reducing these $\sigma$ parameters by a factor of $f_\sigma = 0.89$, the area per molecule is restored
+back close to the experimental value (see Table~\ref{tab:apls}).
+
 \todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules. \\
@@
 SAMULI: There is now a paragraph in the conclusions, which mentions this topic.}
-After reducing these $\sigma$ parameters by a factor of $f_\sigma = 0.89$, the area per molecule is restored
-back close to the experimental value (see Table~\ref{tab:apls}).
 
@@
 \subsection{Salt concentrations and binding affinity}
-The early experimental electrometer concept data for the lipid head group
-order parameter changes was reported as a function 
+The early electrometer concept data for the lipid head group
+order parameter changes from the NMR experiments was reported as a function 
 of the salt concentration in water before solvating
-the lipids~\cite{akutsu81}. The later study used atomic absorption
+the lipids~\cite{akutsu81}. The later study used also the atomic absorption
 spectroscopy and reported the order parameter changes as
@@
     \hline
-    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
+    experiment \cite{kucerka11} & 64.3  &  303    \\
     \hline

### 2017-10-20 Manuscript/manuscript.tex
Updates

@@
 
-\title{Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization} %Title of paper
+\title{Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization} %Title of paper
 
@@
   the binding affinities and stoichimetries of Na$^+$ and Ca$^{2+}$ ions to phospholipid bilayers have not been
-  achieved with different experimental and theoretical methods. As recently shown in the 
-  NMRlipids project, a Open Collaboration running at \url{nmrlipids.blogspot.com},
-  the ion binding details could be resolved with unprecendent detail by intepreting
-  the experimental NMR data with classical molecular dynamics (MD) simulations.
-  However, the accuracy of the existing lipid models for MD simulations was not sufficient for this.
+  achieved with different experimental and theoretical methods.
+  The ion binding details could be resolved with classical molecular dynamics (MD) simulations,
+  however, the accuracy of the available lipid models has not been sufficient.
   In this work we show that the binding details of Na$^+$ and Ca$^{2+}$ ions to
@@
   equal probabilities, suggesting more complicated binding stoichimetry than 
-  the simple binding model used to interpret the NMR data.
+  the ternary complex binding model used to interpret the NMR data.
   The results in this work pave the way for MD simulations of complex biochemical systems
@@
 Cation interactions with cellular membranes play a key role in several biological processes,
-such as in signal propagation in neurons and vesicle fusion.
+such as signal propagation in neurons and vesicle fusion.
 Since the direct measurements of ion-membrane interactions from biological systems
@@
 Interpretations of non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
-and infrared scpectroscopy suggest that Na$^+$ ions exhibit
+and infrared spectroscopy suggest that Na$^+$ ions exhibit
 negligible binding to PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
@@
 which is somewhat smaller than the experimental value, 64.3~\AA$^2$ \cite{kucerka11}, 
-as well as the value given by the original Lipid14 parameters, $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
+as well as the value given by the original Lipid14 parameters, $65.6 \pm 0.5$~\AA$^2$~\cite{dickson14}.
 The decrease of the area per lipid is found to arise from a reduced hydration of the lipid head group region,
@@
 
-\subsubsection{Simulations of POPC bilayers in aqueous ions}
+\subsubsection{Simulations of POPC bilayers with aqueous ions}
 Simulations of POPC bilayer in pure water or in varying salt concentrations
 contained 128 POPC molecules and approximately 50~water molecules per each lipid
-in the periodic orthorhombic simulation box.
-As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
-which is currently the most accurate three site rigid water model
-\todo{More justification for the choosing the OPC3 water model are needed.
-It might be good to show the comparison with the scattering data in bulk water in SI.}. 
-In order to test transferability of our newly developed ECC-lipids model, 
-we also performed several additional simulations with 
-OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005} water models
-\todo{The normal TIP3P was tested as well, right?} presented in Supporting Information (SI). 
-The ECC-ion models were used for Sodium, Calcium and Chloride ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+in the periodic orthorhombic simulation box. The water was described using 
+SPC/E \cite{Berendsen1987} model. This model has the optimal dielectric constant
+to be used with ECC modified molecules \todo{Citation or discussion.} and has been 
+was also used in the parametrization of ECC
+ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.
+The results with other water models are available in SI.
+%As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
+%which is currently the most accurate three site rigid water model
+%In order to test transferability of our newly developed ECC-lipids model, 
+%we also performed several additional simulations with 
+%OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014},
+%and TIP4p/2005 \cite{Abascal2005} water models
+%\todo{The normal TIP3P was tested as well, right?}
+%presented in Supporting Information (SI). 
+The ECC-ion models \todo{Specify the exact version.} were used for Sodium, Calcium and Chloride
+ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
 Simulation with the Lipid14 model was also ran with the ion model by Dang et al.~\cite{smith94,chang1999,dang2006}.

### 2017-10-20 Manuscript/manuscript.tex
Updates

@@
 Simulations of POPC bilayer in pure water or in varying salt concentrations
-contained 128 POPC molecules and approximately 50~water molecules per each lipid
-in the periodic orthorhombic simulation box. The water was described using 
-SPC/E \cite{Berendsen1987} model. This model has the optimal dielectric constant
-to be used with ECC modified molecules \todo{Citation or discussion.} and has been 
-was also used in the parametrization of ECC
+contained 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per each lipid
+in the periodic orthorhombic simulation box. The chosen water model 
+has the optimal dielectric constant
+to be used with the ECC modified molecules \todo{Citation or discussion.} and it has been 
+was also used in the previous parametrization of ECC
 ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.
@@
 ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Simulation with the Lipid14 model was also ran with the ion model by Dang et al.~\cite{smith94,chang1999,dang2006}.
-Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion model
-was taken directly from \cite{lipid14POPC0mMNaClfiles}. \todo{Which water model was used in these simulations?} 
+Simulation with the Lipid14 model was also ran with the ion model by
+Dang et al.~\cite{smith94,chang1999,dang2006}\todo{Which water model was used in this simulation?}.
+Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion and TIP3P~\cite{jorgensen83}
+water models are reported in \cite{catte16} and
+were taken directly from \cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}. 
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
@@
 mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
-All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
+All systems contained 50 POPC molecules per leaflet, 6340 TIP3P~\cite{jorgensen83} water molecules and
 6, 14, 21, 35, or 50 surfactants per leaflet. 
@@
 were not available in the standard Gromacs files for Amber force field.
-\todo{JOE: We cannot present an intricate cool new model for a lipid and claim at the same time that we can't simulate bromide. }
+%\todo{JOE: We cannot present an intricate cool new model for a lipid and claim at the same time that we can't simulate bromide. }
+% Yes we can.
 The Lipid14 model was used for POPC.
@@
     compared with the experimental values from \cite{ferreira13}.
-    The size of the points for the head group order parameters correspond the
-    of the error bars ($\pm 0.02$ for experiments \cite{botan15,ollila16} and ?? for simulations).
+    The size of the points for the head group order parameters correspond 
+    the error bars ($\pm 0.02$ for experiments \cite{botan15,ollila16} and ?? for
+    simulations {\bf Add the size of the error bars in simulations.}).
     The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
@@
   \todo{sn-1 and sn-2 order parameters should be somehow labelled in the order parameter figure. Maybe empty and filled points?} \\
-  \todo{Add size of the error bars in simulations to the caption.} \\
   \todo{Would it be possible to increase the size of points for acyl chain order parameters only in z-direction such that
     it would correspond the error bars?} \\
-  \todo{x-axis in form factor plot from 0 to 0.6, where experimental data ends. y-axis from 0 to $\sim$2.5 to remove the empty space.} \\
+  \todo{x-axis in form factor plot from 0 to 0.6, where experimental data ends. y-axis from 0 to $\sim$2.2 to remove the empty space.} \\
 \end{figure}
@@
 \begin{table}
-  \caption{Area per lipid (APL) values of POPC with no ions from the Lipid14 simulation
-    ran in this work and from the literature, from ECC-lipid model and from experiments.\label{tab:apls} }
+  \caption{Area per lipid (APL) values of POPC bilayer without ions from the Lipid14 simulation
+    ran in this work and from the literature, from the ECC-lipid model and from experiments.\label{tab:apls} }
   \begin{tabular}{l|c c}
@@
   \caption{\label{fig:delta_ordPar_CaCl}
-    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
+    The changes of the head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
     are shown from simulations with different force fields together with experimental data 
@@
     from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
+    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
   }
-  \todo{I think that we must show the NaCl results in the main paper. I think that the best would
-    be to use similar two column format, which we had before. This is used also in NMRlipids II and makes the
-    comparison easier} \\
-  \todo{The DPPC and POPC should be shown and properly labelled as we had previously.} \\
-  \todo{The labeling should be consistent with previous figures, i.e., experiments with diamonds.} \\
-  \todo{Start x-axis scale from -1 to make the point in zero fully visible.} \\
-  \todo{Start y-axis scale of the two top figures from -0.13 to remove the empty space.} \\
+  \todo{Maybe experimental points and lines could be larger to make a more clear differentce to simulations} \\
+  \todo{Add temperatures for POPC and DPPC experiments} \\
   \todo{y-axis scale of the bottom figure from 59 to 78 to remove the empty space.} \\
-  \todo{Empty space between figures and from the right column could be reduced.} \\
+  \todo{Empty space from the right column could be reduced.} \\
   \todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
@@
   \caption{\label{fig:delta_ordPar_NaCl}
-    Changes of head group order parameters $\alpha$ and $\beta$ of POPC bilayer as a function of NaCl concentrations
-    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
-    Ion concentrations in bulk water are shown in x-axis. 
-    Values from simulations are calculated from the of cation number density $C_{np}$
-    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
+    The changes of the head group order parameters of POPC bilayer as a function of NaCl concentrations
+    from simulations with different force fields together with experimental data for DPPC~\cite{akutsu81} and POPC~\cite{altenbach84}.
+    Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from
+    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
   }
+  \todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for
+    POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
+  \todo{Add temperatures to the experimental data.} \\
+  \todo{Labeling should be consistent with the previous figure, i.e. experimental data with diamonds etc.} \\
+  \todo{x-axis scale from -1 to 1100 to make the point in zero fully visible and to remove empty space.} \\
+  \todo{Empty space between figures and from the right column could be reduced.}
 \end{figure}
@@
 
-Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{??},
+Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{??} \todo{Add citations},
 we also simulated the Lipid14 model with the ion models by Dang et al.~\cite{smith94,chang1999,dang2006} and
@@
 Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.
-%In conclusion, the results suggest that 
-%the ECC-lipid model together with the ECC-ion parameters gives a realistic
-%description for the binding affinities of Na$^+$ and Ca$^{2+}$ ions to POPC lipid bilayer. 
+Also the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was
+signifincantly overestimated in the original Lipid14 model, while the ECC-lipid
+gives a good agreement with exepriments, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
+Also the changes of the P-N vector angle are more prononced
+for the Lipid14 model, for which the maximum CaCl$_2$ induced tilting 
+toward water phase is approximately 17$^{\circ}$, while the corresponding
+value for ECC-lipid model is 6$^{\circ}$.
+
 Furthermore, the overestimated lipid headgroup order parameter changes of POPC in the Lipid14
@@
 
-\todo{SAMULI: Maybe we should discuss the repeat distances and
-  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
-  }
+The ion binding affinity of ECC-lipid model with different water models is compared
+in SI. The performance with all water models is better than in in the original Lipid14
+parameteres, but the order parameter changes are slightly overestimated
+with all four-site water models and with TIP3P.
+
+%\todo{SAMULI: Maybe we should discuss the repeat distances and
+%  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
+%  }
 
@@
 
+\todo{This should be either done properly with the correct $C_I$ or removed. Removing
+might be a better idea at this point.}
 It was found in the original work~\cite{altenbach84} that 
@@
 An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions.
-\todo{Did you really calculate the $C_I$ from simulations without severe assumptions?
-Note that this conctration at the plane of biding, which do not equal the concentration of free cations.}
 Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
@@
     experiment   & 62.7  &  293    \\
-    experiment \cite{jambeck12}\todoii{REF}{put original references, not Slipids param. paper.} & 64.3  &  303    \\
+    experiment \cite{kucerka11} & 64.3  &  303    \\
     experiment  & 67.3  &  323    \\
@@
     equilibration time  & 100~ns  \\
-    simulation time     & 200~ns  \\
+    total simulation time     & 300~ns  \\
     temperature         & 313~K       \\

### 2017-10-22 Manuscript/manuscript.tex

@@
 already highly optimized to provide a good description for the
-hydrophobic part of lipid bilayers \cite{ollila16}.
+hydrophobic part of lipid bilayers \cite{dickson14,ollila16}.
 This is in contrasts with the glycerol backbone and the head group regions, 
@@
 (see Fig.~\ref{simVSexpNOions}) are proportional to the amount 
-of unit charge bound per lipid~\cite{seelig87}.
+of charge bound per lipid~\cite{seelig87}.
 The change of order parameters measured with varying aqueous
@@
 
-The relation between bound charge per lipid,  $X^\pm$, and
+The relation between the amount of the bound charge per lipid,  $X^\pm$, and
 the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$,
@@
   \todo{x-axis scale from -1 to 51 to make the point in zero fully visible and to remove empty space.} \\
-  \todo{Empty space between figures and from the right column could be reduced.}
+  \todo{Empty space between figures and from the right column could be reduced.} \\
+  \todo{Put the experimental data with larger points and thicker line in this and in the next two figures to
+    make more clear separation between simulation and experimental data.}
 \end{figure}
@@
 electrostatic interactions in the vicinity of cellular membranes.
-The ECC-lipids model is build by scaling the partial charges and
-the LJ-radius of the headgroup, glycerol backbone and carbonyl atoms of
-the Lipid14 POPC model~\cite{dickson14}. While the Lipid14 model
+The ECC-lipids model is build by scaling the partial charges
+and the LJ-radius of the headgroup, glycerol backbone and carbonyl atoms of
+the Lipid14 POPC model~\cite{dickson14} with the constant factor
+of 0.8 and 0.89, respectively. While the Lipid14 model
 is compatible with the AMBER force field family, the compatibility of

### 2017-10-27 Manuscript/manuscript.tex
Improved density plot. Now in matplotlib style with membrane background and phosphate position. Gamma values still missing in the plot.

@@
   \centering
-  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities2.eps}
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
   \caption{\label{fig:cacl-dens}
-    Number density of \ce{Ca^{2+}} and \ce{Cl^-} as a function of membrane normal axis
+    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis
     for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
-    Densities of~\ce{Cl^-} and water are divided with 2 and 200, respectively, to visalize them
-    with the same scale as \ce{Ca^{2+}}. The molar concentration of the ions in water is 350~mM in all systems
+    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
+    densities of~\ce{Cl^-} and \ce{Na^{+}} are divided by 2, and
+    densities of phosphate group resp. water are divided by 5 resp. 200. 
+    The molar concentration of ions in water is 350~mM in all systems
     presented here. 
     }
-  \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
-  \todo{JOE: Change the figure so that it contains a membrane background} \\
+  \todo{PAVEL: include the number of$\Gamma$-surface access.} \\
 \end{figure}

### 2017-10-30 Manuscript/manuscript.tex
Big update of data: added several new figures, improved all main-text figures, finished transition to SPCE.

@@
 shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
-and the longest observed residence time in the ECC-lipid simulation is $120\,\mathrm{ns}$, 
+and the longest observed residence time in the ECC-lipid simulation is $141\,\mathrm{ns}$, 
 which is well below the total length of the simulation (200~ns).
@@
 an order of magnitude faster, giving 90\%
-of the residence times smaller than~$2\,\mathrm{ns}$,
-the longest residence time being~$13\,\mathrm{ns}$. 
+of the residence times smaller than~$1\,\mathrm{ns}$,
+the longest residence time being~$6\,\mathrm{ns}$. 
 All the results are in line with the experimental estimation that 

### 2017-11-01 Manuscript/manuscript.tex
Figure update2: done what was left to do in previou commit

@@
 %to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
-%% Numbers from simulation with 346mM conc. CaCl2
+%% Numbers from simulation with 346mM conc. CaCl2 (OPC3)
 %The average number of contacts per lipid 
@@
 %
-% Numbers from simulation with 980/845mM conc. CaCl2
+% Numbers from simulation with 980/845mM conc. CaCl2 (OPC3)
 %The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
@@
 The results in Table \ref{tab:binding} show that almost all (99\%) of the 
-bound Ca$^{2+}$ ions are in contact with the phosphate oxygens and a quarter
-of them (25\%) interacts with the carbonyl oxygens. Thus, the interactions between
+bound Ca$^{2+}$ ions are in contact with the phosphate oxygens,
+with two thirds (67\%) are interacting with only phosphate oxygens
+and the rest, one third (32\%), interacts also with the carbonyl oxygens. 
+Thus, the interactions between
 Calcium ions with only carbonyl oxygens are very unlikely (1\%), but significant

### 2017-11-01 Manuscript/manuscript.tex
Corrected/Added APL extra values.

@@
     \hline
-    ECC-lipids                & 62.2$\pm$ 0.6  &  300       \\
+    ECC-lipids                & 63.2$\pm$ 0.6  &  300       \\
     \hline
@@
     ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
+    ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
     ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
     ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
+    ~ TIP3p          & 66.2$\pm$ 0.6   &  313       \\
+    ~ TIP3p-FB       & 64.8$\pm$ 0.6   &  313       \\
+    ~ TIP4p-FB       & 65.6$\pm$ 0.6   &  313       \\
     %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89

### 2017-11-02 Manuscript/manuscript.tex
Modified and polished sections Intro and Methods. Introduced extra fixes and updated NaCl figure.

@@
   \centering
-  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
-  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89.pdf}
+  \includegraphics[width=7.4cm]{../Fig/POPCstructure.eps}
+  %\includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
+  %\includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
   \caption{\label{simVSexpNOions}
@@
   \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
+  \includegraphics[height=20.0cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
   \caption{\label{fig:cacl-dens}

### 2017-11-03 Manuscript/manuscript.tex
committing what should have been contained in previous commit.

@@
   % insert abstract here
-  Despite of the significant biological relevance, the binding details of cations in cellular
-  lipid membranes are not fully understood. For example, consistent results for
-  the binding affinities and stoichimetries of Na$^+$ and Ca$^{2+}$ ions to phospholipid bilayers have not been
-  achieved with different experimental and theoretical methods.
+  The binding affinities and stoichimetries of Na$^+$ and Ca$^{2+}$ ions to phospholipid bilayers have not been
+  consistently determined with different experimental and theoretical methods
+  despite their significance and biological relevance in cellular membranes. 
   The ion binding details could be resolved with classical molecular dynamics (MD) simulations,
@@
 and infrared spectroscopy suggest that Na$^+$ ions exhibit
-negligible binding to PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ specifically
-binds to phosphate groups of two lipid molecules~\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
+negligible binding to PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ is observed to specifically
+bind to phosphate groups of two lipid molecules~\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
 Atomistic resolution molecular dynamics (MD) simulation models, however, predict significantly
@@
 MD simulations~\cite{bockmann03,vacha09a}. 
-Furthermore, interactions of Calcium ions with 3-4 lipids,
+Furthermore, interactions of calcium ions with 3-4 lipids,
 including also interactions with carbonyl
@@
 in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
-improved by implicitly including the electronic polarizability
-in the polar region of lipid molecules. The electronic
-polarizability is included by using the electronic
-continuum correction (ECC)~\cite{leontyev11}, which
-has been previously shown to improve the behaviour of
+improved with an implicit inclusion of the electronic polarizability
+in the polar region of phospholipids,
+which we apply according to the 
+electronic continuum correction (ECC)~\cite{leontyev11}.
+Such an approach has been previously shown to improve the behaviour of
 MD simulations of ions in bulk water~\cite{jungwirth17-new-paper-to-be-published,Pluharova2014,kohagen14,kohagen16}. 
-As a starting point we use the parameters from the Lipid14 model \cite{dickson14},
-which gave the best cation binding behaviour in the previous study \cite{catte16}.
-The developed ECC-lipid parameters reproduce the experimentally
+We use the the Lipid14 model \cite{dickson14} as a starting point,
+because it provided one of the best available descriptions of cation binding \cite{catte16}.
+The developed ECC-lipid model reproduces the experimentally
 measurable structural parameters of an ion-free POPC lipid bilayer with the 
 accuracy comparable to the other state of the art lipid models, while surpassing
-them significantly for reproducing the membrane binding affinities and induced
-structural effects of Sodium and Calcium ions.
+them significantly in the description of the membrane binding affinities to cations and induced
+structural effects of sodium and calcium ions.
 
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-The lack of electronic polarizability in the standard MD simulation
-force fields has been considered a higly relevant issue since the early days of
+The lack of polarizability in standard MD simulation
+force fields has been considered an issue since the early days of
 lipid bilayer simulations.
-In this work we circumvent the rather demanding explicit inclusion
-of electronic polarization effects \cite{lucas12,chowdhary13} by implicitly including
-electronic polarizability in lipid
-bilayer simulations by using the electronic continuum correction
-(ECC)~\cite{leontyev11}. 
-Technically, it is a similar approach to the phenomenological charge-scaling as applied in 
+In this work, we circumvent the rather demanding explicit inclusion
+of electronic polarization effects \cite{lucas12,chowdhary13},
+including the electronic part of polarizability in lipid
+bilayer simulations implicitly through 
+the electronic continuum correction (ECC)~\cite{leontyev11}. 
+Technically, it is similar to the phenomenological charge-scaling applied in 
 earlier studies \cite{jonsson86,egberts94}.
 \todo{We should also cite papers where empirical scaling was used ionic liquids DOI: 10.1002/anie.201308760}
-The present concept of ECC is, however, 
-physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11}.
-
-Following the ECC concept, electronic
-polarizability can be effectively included in classical MD simulations by
+However, the present concept of ECC is
+physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
+
+According to ECC, electronic
+polarizability can be implicitly included in classical MD simulations by
 placing all particles into a homogeneous dielectric continuum 
@@
 the media \cite{leontyev11}. 
-Following the Coulomb's law, such a dielectric continuum can be easily included in standard MD simulation by
-a scaling of charges 
+Following the Coulomb's law, such a dielectric continuum can be 
+easily included in a standard MD simulation by
+the scaling of charges,
 \begin{equation}
-  Q^{ECC} = f_q \cdot Q
+  Q^{ECC} = f_q \cdot Q ,
 \end{equation}
@@
 include  to some extend the solvent electronic polarizability effects, i.e.,
-the RESP charges are often scaled to fit certain experimental observables \todo{This needs a citation - Joe please pick 1-2 papers}. 
+the RESP charges are subsequently modified to fit certain experimental observables \todo{This needs a citation - Joe please pick 1-2 papers}. 
 Thus, a consistent application of the ECC scaling factor, $f_q$, to the molecular partial charges
@@
 and $1$ (i.e., electronic polarizability fully included in the original partial charges).
-It is important to note that even thought the lipid bilayer interface has a large gradient in
-the total dielectric constant, the electronic polarizability part can be considered to be constant
+It is important to note that even though the lipid bilayer interface has a large gradient in
+the total dielectric constant, the electronic part of polarizability can be considered to be constant
 as the measurements of high frequency dielectric constant gives values around 2 
@@
 In this work, we employ the ECC to develop a classical MD simulation model for
-a POPC lipid that accurately describes the binding of Sodium and Calcium ions to a lipid bilayer.
+a POPC lipid that accurately describes the binding of sodium and calcium ions to a lipid bilayer.
 %This is a biologically highly relevant membrane feature, which is
@@
 %in the presence of ions but also modulate interactions with charged moieties at the interface.
-The model quality in terms of the ion binding is measured by directly
-comparing simulations to the experimental NMR data used to measure the ion binding
-affinity~\cite{akutsu81,altenbach84,scherer89}, as discussed in detail in Ref. \citenum{catte16}.
+The quality of the simulation model is compared 
+to the experimental NMR data used to measure the ion binding
+affinity~\cite{akutsu81,altenbach84,scherer89}, 
+as discussed in detail in Ref.~\citenum{catte16}.
 The Lipid14~\cite{dickson14} force field parameters (available in Gromacs format from Ref. \citenum{lipid14files})
-were used as a starting point, because they give the most
-realistic response to the bound ions and to the NaCl concentration 
-among the available lipid models (see Figs.~5~and~2 in Ref.~\citenum{catte16}).
-Moreover, the Lipid14 model gave relatively realistic glycerol backbone and
-head group structures when compared with the other state of the art lipid models~\cite{botan15}.
-
-The ECC correction was applied to the Lipid14 parameters by scaling partial charges of the head group,
-glycerol backbone, and carbonyl regions, which are the most polar parts in the  lipids and
-are thus expected to have the largest contribution to the cation binding.
-We do not modify the hydrocarbon chain parameters, as they
-do not come in direct contact with salt ions and are
-already highly optimized to provide a good description for the
+were used as a starting point, because they provide one of the most
+realistic descriptions of the head group order parameter response to ions 
+among the available lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}).
+Moreover, the Lipid14 model has a relatively realistic glycerol backbone and
+head group structure compareable to other state of the art lipid models~\cite{botan15}.
+
+We applied the ECC correction to the Lipid14 model of POPC as a scaling of partial charges of the head group,
+glycerol backbone, and carbonyl regions, which are the most polar parts in phospholipids
+being expected to have the largest contribution to the cation binding.
+We do not modify the hydrocarbon chain parameters, 
+as they do not come in direct contact with salt ions, 
+and as they are already highly optimized to provide a good description of the
 hydrophobic part of lipid bilayers \cite{dickson14,ollila16}.
@@
 related head group order parameter response in general.
-By comparing the results with different scaling factors to the experimental 
-NMR data~\cite{akutsu81,altenbach84,scherer89} we found the optimal 
-value of $f_q = 0.8$,
-which is only slighly higher than the scaling factor for the ions in water 0.75.
-%, which implies
-%that electronic polarization effects are only to a small extent effectively accounted for in the Lipid14 force field.
-
-While, the scaling of the partial charges improved the ion binding affinity,
-it reduced the bilayer area per lipid below the experimental value. 
-%Simulations with Lipid14 parameters having 
-After scaling the partial charges of the head group, the glycerol
-backbone and carbonyls with a factor of 0.8, the simulation gave 
-the approximately 60~\AA$^2$ for the area per lipid,
-which is somewhat smaller than the experimental value, 64.3~\AA$^2$ \cite{kucerka11}, 
-as well as the value given by the original Lipid14 parameters, $65.6 \pm 0.5$~\AA$^2$~\cite{dickson14}.
-The decrease of the area per lipid is found to arise from a reduced hydration of the lipid head group region,
-due to the lower polarity of molecules with scaled charges. This can be compensated
-by reducing the effective radius of atoms with the scaled charges by changing 
-the $\sigma$ parameters in the Lennard-Jones potential, 
-as previously done for ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
-After reducing these $\sigma$ parameters by a factor of $f_\sigma = 0.89$, the area per molecule is restored
-back close to the experimental value (see Table~\ref{tab:apls}).
-
-\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules. \\
-  JOE: I think that we rather increasingly see that there's nothing like ``fully non-scaled'' with the exception of ions with integer
-  charges. So the discussion shall be rather more about the interaction of our ``scaled'' (I'd still rather call it ECC-corrected or whatever)
-  and ``semi-scaled'' models. \\
-SAMULI: There is now a paragraph in the conclusions, which mentions this topic.}
+We compared the results with different scaling factors to the experimental 
+NMR data~\cite{akutsu81,altenbach84,scherer89} finding the optimal 
+value $f_q = 0.8$,
+which is only slighly higher than the scaling factor for the ions in water, 0.75.
+
+Although the scaling of the partial charges improved 
+the head group order parameter response and ion binding affinity,
+it also deteriorated some membrane properties; 
+namely the area per lipid decreased below the experimental value. 
+After only scaling the partial charges of the head group, the glycerol
+backbone and carbonyls with a factor of 0.8, 
+the area per lipid in the simulation drops to approximately 60~\AA$^2$,
+which is significantly smaller than the experimental value, 64.3~\AA$^2$ \cite{kucerka11}, 
+as well as the value from the original Lipid14 model, $65.6 \pm 0.5$~\AA$^2$~\cite{dickson14}.
+The decrease of the area per lipid is found to arise from a reduced hydration of the lipid head group region
+as a consequence of the lower polarity after scaling charges. 
+This effect can be compensated
+by reducing the effective radii of atoms with the scaled charges.
+This is in practice done by changing the parameters $\sigma$ in the Lennard-Jones potential 
+in a similar way as was previously done for the ECC-ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
+After reducing the $\sigma$ parameters of affected atoms with a scaling factor $f_\sigma = 0.89$, 
+the area per molecule is restored
+back close to the experimental value (Table~\ref{tab:apls}).
 
 \subsection{Electrometer concept} \label{section:electrometer}
-Ion binding in lipid bilayers was compared between experiments and simulations by 
+Ion binding in lipid bilayers was compared between experiments and simulations 
 using the lipid head group order parameters and the
@@
 is based on the experimental observation that the C-H bond
-order parameters of $\alpha$ and $\beta$ carbons in PC lipid head group
-(see Fig.~\ref{simVSexpNOions}) are proportional to the amount 
+order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group
+(Fig.~\ref{simVSexpNOions}) are proportional to the amount 
 of charge bound per lipid~\cite{seelig87}.
-The change of order parameters measured with varying aqueous
+The change of the order parameters measured with varying aqueous
 ion concentration can be then related to the amount of bound ions.
@@
 The slopes are negative, because the head group order parameters
-decrease with bound positive charge and increase with bound negative
+decrease with a bound positive charge and increase with a bound negative
 charge~\cite{ollila16,catte16}. This can be rationalized as a change
-of the lipid head group dipole tilt toward water phase with the bound positive
+of the lipid head group dipole tilt towards water phase with the bound positive
 charge and {\it vice versa} with the negative charge~\cite{seelig87}.
 
-The measured order parameter change depends on both, the head group response 
-to the bound charge and the amount of bound charge, i.e. $m_i$ and $X^\pm$
-in Eq.~\ref{OPchangeEQ}, respectively. The former property has to be
+The measured order parameter change depends on both the head group response 
+to the bound charge, and the amount of bound charge 
+(i.e. $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively). 
+The empirical factor $m_i$ has to be
 well quantified before the electrometer concept can used to 
-analyze the binding affinities, as done experimentally for a wide range
+analyze the binding affinities.
+This was done experimentally for a wide range
 of systems~\cite{seelig87,beschiasvili91}. To calibrate 
@@
 \subsection{Salt concentrations and binding affinity}
-The early electrometer concept data for the lipid head group
-order parameter changes from the NMR experiments was reported as a function 
-of the salt concentration in water before solvating
-the lipids~\cite{akutsu81}. The later study used also the atomic absorption
-spectroscopy and reported the order parameter changes as
-a function of the salt concentration in the supernatant after the
+The concentrations of salts are reported in two different ways
+in the NMR measurements of head group order parameters \cite{akutsu81, altenbach84}.
+The early work on the NMR electrometer concept 
+reports the salt concentrations in water before solvating
+the lipids~\cite{akutsu81}. 
+The later study uses the atomic absorption
+spectroscopy, and reports the salt concentration in the supernatant after the
 solvation of lipids~\cite{altenbach84}.
-In this work we focus on POPC for which the latter definition was used~\cite{altenbach84}.
+In this work, we focus on POPC for which the latter definition was used~\cite{altenbach84}.
 The salt concentration in the aqueous bulk region was calculated from 
 the farthest point from both lipid leaflets in the water phase. 
-Note that in the previous study, Ref.~\citenum{catte16},
+Note that in the previous study~\cite{catte16},
 the ion concentrations were calculated in water before solvating the lipids 
 as in the earlier experiments~\cite{akutsu81}.
-Despite of the measurable differences between these two
+Although there are measurable differences between these two
 definitions of concentrations for CaCl$_2$ systems,
-the qualitative conclusions in this or in the previous work~\cite{catte16} are
-not affected by this.
+they do not affect the qualitative conclusions in this or in the previous work~\cite{catte16}.
 
 To quantify the ion binding affinity to a lipid bilayer, we
-calculated the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
-This quantity was chosen because it does not depend on
-%the definition of the interfacial region or
-the position of the Gibbs dividing plane between two bulk regions.
-%\todo{JOE: following is very unclear to me. I think we can write it in a more concise way.}
-%SAMULI: Here is a new trial.
-Here we assume that the interface locates between the hydrophobic interior
+calculate the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
+Such a quantity compares the adsorption of ions to the adsorption of water molecules 
+at an interface without the necessity of defining the Gibbs dividing plane 
+between the membrane interior and the water bulk region. 
+In our simulations we assume that the interface locates between the hydrophobic interior
 of a lipid bilayer and the bulk water region far from the membrane.
@@
 %the hydrophobic region in the middle of the membrane where no ions or waters are present.
-This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
+Such a setup provides a simplified relation for $\Gamma_i^w$ for simulations of lipid bilayers, 
 \begin{equation}\label{surfexcess}
@@
 The order parameters were calculated from simulations for all C-H bonds
-in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
-from equation \todo{As Hector suggested, it might be better to write the simpler
-form for this equation.}
+in lipid molecules from Eq. \ref{OP}. Form factors were calculated 
+from the equation
 \begin{equation}
-  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
+  F(q) = \int _{-D/2} ^{D/2} \left ( \rho_{el}(z) - \rho_{el}^s \right ) \cos (zq_z) \mathrm{d}z,
+  %F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
 \end{equation}
-where $f_\alpha(q_z)$ is the density of atomic scattering length, 
-$\rho_s$ is the density of solvent scattering length in the bulk region,
-$n_\alpha (z)$ is the number density of atom $\alpha$ and
-$z$ is the distance from the membrane centre along its normal 
-spanning until the water bulk region, $D$.
+where $\rho_{el} (z)$ is the total electron density,
+$\rho_{el}^s$ is the electron density of the solvent far in the bulk,
+and $z$ is the distance from the membrane centre along its normal 
+spanning until the water bulk region, $D/2$.
 
@@
 \subsubsection{Simulations of POPC bilayers with aqueous ions}
-Simulations of POPC bilayer in pure water or in varying salt concentrations
+Simulations of a POPC bilayer in a pure water or in varying salt concentrations
 contained 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per each lipid
-in the periodic orthorhombic simulation box. The chosen water model 
-has the optimal dielectric constant
-to be used with the ECC modified molecules \todo{Citation or discussion.} and it has been 
-was also used in the previous parametrization of ECC
-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.
-The results with other water models are available in SI.
-%As a default, water molecules were described by the OPC3 force field~\cite{Izadi16} 
-%which is currently the most accurate three site rigid water model
-%In order to test transferability of our newly developed ECC-lipids model, 
-%we also performed several additional simulations with 
-%OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014},
-%and TIP4p/2005 \cite{Abascal2005} water models
-%\todo{The normal TIP3P was tested as well, right?}
-%presented in Supporting Information (SI). 
-The ECC-ion models \todo{Specify the exact version.} were used for Sodium, Calcium and Chloride
-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Simulation with the Lipid14 model was also ran with the ion model by
-Dang et al.~\cite{smith94,chang1999,dang2006}\todo{Which water model was used in this simulation?}.
-Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ion and TIP3P~\cite{jorgensen83}
-water models are reported in \cite{catte16} and
+in an orthorhombic simulation box with periodic boundary conditions. 
+The chosen water model was used in the previous parametrization of 
+ECC-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014},
+and it has dielectric properties consistent with ECC~\cite{leontyev11, leontyev14}.
+Results with other water models are available in SI
+(OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, 
+TIP3P~\cite{jorgensen83},
+TIP3p-FB and TIP4p-FB~\cite{Wang2014},
+and TIP4p/2005 \cite{Abascal2005}).
+Sodium, calcium and chloride ions 
+were modeled as ECC-ions with parameters from \citenum{jungwirth17-new-paper-to-be-published, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
+Simulation with the Lipid14 model used the TIP3p water model \cite{jorgensen83} and the ion model by
+Dang et al.~\cite{smith94,chang1999,dang2006}.
+Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ions and TIP3P~\cite{jorgensen83}
+water models reported in \cite{catte16} 
 were taken directly from \cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}. 
@@
 summarized in Table \ref{tbl:mdpar}. 
-%and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
-%\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
-%this is done in Amber by default. If this is correct, the information in the
-%table do not match with this sentence. Based on semi-extensive testing I made few years
-%ago this do not change anything. Anyway, this should be reported consistently.}
 Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
@@
 \subsubsection{Simulations of POPC bilayers with cationic surfactants}
-An automated topology builder \cite{malde11} was first used to create the structure of
-dihexadecyldimethylammonium.
+The structure of dihexadecyldimethylammonium
+was created with an automated topology builder \cite{malde11}. 
 The AmberTools program \cite{amber} was then used to generate the Amber-type force field
-parameters. The parameters were converted to the Gromacs format by using
+parameters. The parameters were converted to the Gromacs format with
 the acpype tool \cite{acpype}. The partial charges were then manually modified
@@
 6, 14, 21, 35, or 50 surfactants per leaflet. 
-Chloride counter ions were used
-in simulations, because parameters for bromide, the counterion in the experimental data \cite{scherer89},
-were not available in the standard Gromacs files for Amber force field.
-%\todo{JOE: We cannot present an intricate cool new model for a lipid and claim at the same time that we can't simulate bromide. }
-% Yes we can.
+Note that chloride counter ions were used in simulations, 
+but bromide was used as a counter ion in the experiment \cite{scherer89}.
 The Lipid14 model was used for POPC.
@@
 was considered as an equlibration time and was omitted from the analysis.
-A reasonable lipid neighbor exchange occured during the simulation.
-
-The same systems were also simulated with the ECC-lipid model for POPC using the same setup.
-\todo{Which water model was used in these simulations?}
-In these simulations the ECC correction was also applied to the cationic surfactant 
-by scaling all charges with the same factor as for ECC-lipids, i.e., $f_q=0.8$, 
-and by using the atom types with reduced $\sigma$ parameters from ECC-lipids. 
-\todo{Parameters are available at ??}
+A reasonable lipid neighbor exchange occured during the simulations.
+
+The same systems were also simulated with the ECC-lipid model for POPC 
+and SPC/E model for water using the same setup.
+In these simulations ECC was also applied to the cationic surfactant 
+in the same way as to POPC;
+all charges were scaled with the same factor as for ECC-lipids, i.e. $f_q=0.8$, 
+and the same atom types with reduced $\sigma$ parameters from ECC-lipids were used. 
+\todo{Provide the parameteres for ECC-surfactant somewere.}
 
@@
   } 
-  \todo{Increase the height of the order parameter figure} \\
-  \todo{sn-1 and sn-2 order parameters should be somehow labelled in the order parameter figure. Maybe empty and filled points?} \\
-  \todo{Would it be possible to increase the size of points for acyl chain order parameters only in z-direction such that
-    it would correspond the error bars?} \\
-  \todo{x-axis in form factor plot from 0 to 0.6, where experimental data ends. y-axis from 0 to $\sim$2.2 to remove the empty space.} \\
 \end{figure}
@@
 \begin{table}
-  \caption{Area per lipid (APL) values of POPC bilayer without ions from the Lipid14 simulation
-    ran in this work and from the literature, from the ECC-lipid model and from experiments.\label{tab:apls} }
+  \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
+    ran in this work and from the literature \cite{dickson14}, from the ECC-lipid model, and from experiments.\label{tab:apls} }
   \begin{tabular}{l|c c}
@@
   }
-  \todo{Labeling should be consistent with the previous figure, i.e. experimental data with diamonds.} \\
-  \todo{x-axis scale from -1 to 51 to make the point in zero fully visible and to remove empty space.} \\
-  \todo{Empty space between figures and from the right column could be reduced.} \\
-  \todo{Put the experimental data with larger points and thicker line in this and in the next two figures to
-    make more clear separation between simulation and experimental data.}
 \end{figure}
@@
   }
-  \todo{Maybe experimental points and lines could be larger to make a more clear differentce to simulations} \\
   \todo{Add temperatures for POPC and DPPC experiments} \\
-  \todo{y-axis scale of the bottom figure from 59 to 78 to remove the empty space.} \\
-  \todo{Empty space from the right column could be reduced.} \\
-  \todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
-    the value of $g_3$ order parameter of DPPC was -0.214 
-    in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
-    of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
-    at the C-2 segments of both fatty acyl chains.''
-  }
+  %\todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
+    %the value of $g_3$ order parameter of DPPC was -0.214 
+    %in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
+    %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
+    %at the C-2 segments of both fatty acyl chains.'' }
 \end{figure}
@@
   }
-  \todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for
-    POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
+  %\todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
   \todo{Add temperatures to the experimental data.} \\
-  \todo{Labeling should be consistent with the previous figure, i.e. experimental data with diamonds etc.} \\
-  \todo{x-axis scale from -1 to 1100 to make the point in zero fully visible and to remove empty space.} \\
-  \todo{Empty space between figures and from the right column could be reduced.}
 \end{figure}
@@
 
-Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{??} \todo{Add citations},
+Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{auffinger07},
 we also simulated the Lipid14 model with the ion models by Dang et al.~\cite{smith94,chang1999,dang2006} and
-ECC-ions~\cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} having
+ECC-ions~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014} having
 more realistic bulk behaviour. Instead of the improvement in the binding behaviour,
@@
 \begin{table}
-  \caption{The relative surface excess and the probabilities of contacts between bound Calcium and different
+  \caption{The relative surface excess and the probabilities of contacts between bound calcium and different
     oxygen moieties in POPC from different simulation models. The CaCl$_2$ concentration of the ions in water
@@
 or with Dang ions, $\Gamma_{\rm Ca}^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
-\todo{JOE: I'd put the $\Gamma$ values into a table SAMULI: They are now in the table \ref{tab:binding} 
-together with probabilities of the calcium binding probabilities to different POPC oxygens. 
-However, I am not sure if the table is useful in the end.}
 Interestingly, the relative surface excess of 
@@
 %the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
-\todo{Densities and relative surface excess from NaCl systems to be added.
-  The discussion above to be finished after this.}
 
@@
     }
-  \todo{PAVEL: include the number of$\Gamma$-surface access.} \\
 \end{figure}
@@
 Thus, the interactions between
-Calcium ions with only carbonyl oxygens are very unlikely (1\%), but significant
+calcium ions with only carbonyl oxygens are very unlikely (1\%), but significant
 amount of the ions interact with both, phosphate and carbonyl oxygens.
@@
 
-In conclusion, the results suggest that the Calcium ions specifically bind to phosphate
+In conclusion, the results suggest that the calcium ions specifically bind to phosphate
 oxygens, while occasionally interacting also with carbonyls. This is in good agreement
@@
 studies~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
-Also the Sodium ions bind most likely to the phosphate oxygens, while the probability
-of binding purely to carbonyls is also significant, in contrast to Calcium ions.
-The physiological relevance of the Sodium binding details is, however, uncertain
+Also the sodium ions bind most likely to the phosphate oxygens, while the probability
+of binding purely to carbonyls is also significant, in contrast to calcium ions.
+The physiological relevance of the sodium binding details is, however, uncertain
 due to the very weak binding affinity.
@@
 The results from POPC simulation with 285~mM concetration of CaCl$_2$ in
-Fig.~\ref{fig:cacl_complexes} show the largest propensity (42\%) for the
+Fig.~\ref{fig:cacl_complexes} show the largest propensity (41\%) for the
 ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
-the probabilities for the complexes with the stoichimetries of 1:1~(30\%)~and
-1:3~(28\%)~are only slightly lower,
+the probabilities for the complexes with the stoichimetries of 1:1~(25\%)~and
+1:3~(34\%)~are only slightly lower,
 suggesting more complex binding stoichiometry than the simple ternary complex model.
 %The observed propensities of
-%$42\%$ for two lipids,
-%$30\%$ for one lipid, and
-%$28\%$ for three lipids 
-However, one Calcium binds roughly to two lipids on average also in simulations,
+%$41\%$ for two lipids,
+%$25\%$ for one lipid, and
+%$34\%$ for three lipids 
+However, one calcium binds roughly to two lipids on average also in simulations,
 because the probabilities of forming the complexes with 1 or 3 lipids are almost equal.  
@@
 experimental data. 
-Complexes with more than three lipids per one Calcium ion were not observed in 
+Complexes with more than three lipids per one calcium ion were not observed in 
 simulations with the ECC-lipids model.
@@
 analyzed from the ECC-lipid model simulation with 1 M concentration of NaCl are
-shown in Fig.~\ref{fig:cacl_complexes}. In contrast to Calcium, the
-probability is largest (??\%) for 1:1 complex, significantly smaller (??\%)
-for 1:2 complexes and very small (??\%) for 1:3 of \ce{Na+}:POPC complexes.
-%follows stoichiomtetry \ce{Na^{+}}:$1\times$POPCas shown in 
-\todo{Add numbers for Sodium.}
+shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
+probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
+for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.
 
@@
 from the ECC-lipid simulation in Fig.~\ref{fig:cacl-dens} show a relatively shallow
-minimum at the interface, suggesting a smaller barrier for the exchange of Calcium ions.
+minimum at the interface, suggesting a smaller barrier for the exchange of calcium ions.
 To quantify the ion binding exchange in lipid bilayer simulations, we calculated 
@@
 In conclusion, the results from ECC-lipid model suggest that the
-exchange of bound Calcium occurs within $\sim$100ns timeframe,
+exchange of bound calcium occurs within $\sim$100ns timeframe,
 which is much faster than previously reported~\cite{javanainen17}.
@@
 %in equilibrium between membrane and solvent.
-
-\subsubsection{Internal side-note: }
-However, $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ is increased 
-in simulation withy the highest \ce{CaCl2} concentration
-when carbonyl dipoles are scaled only with 0.945 
-(roughly done through charge redistribution)
- to : $\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} \approx 0.08 $.    % = (70.-59.8)/128 .
-
-The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 \rightarrow 0.309 $,
-whereas if only phosphate oxygen atoms are considered 
-the average number of contacts decreases only by a tiny amount to 
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 \rightarrow 0.303 $,    % = 33.5/128 . 
-and if only carbonyl oxygen atoms are considered 
-(phosphate oxygens contribute, but are not counted as contacts),
-this value is merely
-$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 \rightarrow 0.139 $.    % = 8.4/128 .
-
-This means that calcium is in contact with a phosphate oxygen for 98\% of the time it is bound to the membrane, 
-whereas 46\% of the time accounts for contact with a carbonyl oxygen,
-and purely carbonyl bound states account for less than 2\% of configurations. 
-The states that have a contribution from carbonyl oxygens also contain a bound phosphate, 
-so it can be thought of rather as a stabilizing moiety.
-
-The residence times change to 75ns (90\% of bound time) and max is 185ns.
-That's not much, too. It certainly is not a game changer, I'd even call it to be within accuracy. 
-The slight problem may be that with this tweak, there's approx. 17\%  more cations bound to the membrane,
-which means, I'd need to tune the sigma parameter a bit (increase the scaling factor $f_\sigma$). 
-That could detune the membrane structure \dots
-Not $f_q$ -- that's set by the simulations with the surfactant. Good.
-I think that such a binding detail is within the accuracy of the proposed simple correction.
-It would be worth noting that the model may underestimate slightly the contribution from carbonyls due to this. 
-But we can't say to what extent\dots
-
-%\begin{figure}[tbp]
-%  \centering
-%  \includegraphics[height=6.0cm]{../Fig/POPC_binding_labels.eps} 
-%  \includegraphics[height=6.0cm]{../Fig/POPC_binding_ECC-lipids.eps}
-%  \caption{\label{fig:binding_conf}
-%      Configurations of \ce{Ca^{2+}} binding in POPC membranes in the ECC-lipids model.
-%      The binding sites are: 
-%      red circle for phosphate oxygens; 
-%      blue square for sn-1 carbony oxygen; and green triangle for sn-2 carbonyl oxygen.
-%      The configurations along with their probabilities are drawn in the following manner:
-%      each row corresponds to an individual lipid the cation is bound to;
-%      symbols denote the site the cations is bound to the respective lipid.
-%      An empty configuration denotes an unbound cation in a solvent. 
-%  }
-%\end{figure}
 
@@
 and 600 from the Academy of Finland via the FiDiPro award.
+\todo{Add acknowledgments of Jiri Kolafa.}
 \end{acknowledgments}
@@
   \end{tabular} \\
-  \todo{Result with normal TIP3P missing?}
 \end{table}

### 2017-11-03 Manuscript/manuscript.tex
Corrections to sections A-C in Results (i.e. until Affinity and binding headline)

@@
   state of the art lipid model for MD simulations of POPC
-  lipid bilayer. The introduced ECC-lipid model reproduces the experimentally measured structural parameters for the ion-free membrane,
+  lipid bilayer. The introduced ECC-lipids model reproduces the experimentally measured structural parameters for the ion-free membrane,
   the response of lipid headgroup to the bound positive charge, and the binding affinities of
   Na$^+$ and Ca$^{2+}$ ions. The imperceptible binding of Na$^+$ ions to POPC bilayer
-  and interactions of Ca$^{2+}$ mainly with phoshpate oxygens in the ECC-lipid model give
+  and interactions of Ca$^{2+}$ mainly with phoshpate oxygens in the ECC-lipids model give
   support to the early intepretations of the experimental spectroscopic data.
@@
 because it provided one of the best available descriptions of cation binding \cite{catte16}.
-The developed ECC-lipid model reproduces the experimentally
+The developed ECC-lipids model reproduces the experimentally
 measurable structural parameters of an ion-free POPC lipid bilayer with the 
@@
 
-The same systems were also simulated with the ECC-lipid model for POPC 
+The same systems were also simulated with the ECC-lipids model for POPC 
 and SPC/E model for water using the same setup.
@@
 
-\begin{figure}[tbp]
+\begin{figure}[tb]
   \centering
@@
   \caption{\label{simVSexpNOions}
-    Top: The chemical structure of POPC and the labeling of the carbon segments.
+    Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
+    the ECC-lipids models compared with experiments~\cite{Kucerka2011}.
     Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipid models
+    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
     compared with the experimental values from \cite{ferreira13}.
-    The size of the points for the head group order parameters correspond 
-    the error bars ($\pm 0.02$ for experiments \cite{botan15,ollila16} and ?? for
-    simulations {\bf Add the size of the error bars in simulations.}).
+    The size of the markers for the head group order parameters correspond to
+    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
+    while the error estimate for simulations is $\pm 0.005$.
     The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
-    Bottom: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-lipid models compared with experiments~\cite{Kucerka2011}.
+    Bottom: The chemical structure of POPC and the labeling of the carbon segments.
   } 
@@
   \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
-    ran in this work and from the literature \cite{dickson14}, from the ECC-lipid model, and from experiments.\label{tab:apls} }
+    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments.\label{tab:apls} }
   \begin{tabular}{l|c c}
@@
 
-The ECC-lipid and Lipid14 models reproduce the experimental x-ray scattering form factors
-of POPC bilayer with the similar accuracy in Fig.~\ref{simVSexpNOions}.
+The ECC-lipids and Lipid14 models reproduce the experimental x-ray scattering form factors
+of a POPC bilayer with a similar accuracy in Fig.~\ref{simVSexpNOions}.
 The area per lipid values from the Lipid14 model is slightly larger than the
-experimental value in Table~\ref{tab:apls}, while the value from ECC-lipid model
-is slightly smaller. Area per lipid values of the ECC-lipid model show some variation
+experimental value in Table~\ref{tab:apls}, while the value from ECC-lipids model
+is slightly smaller. The values of the area per lipid of the ECC-lipids model show some variation
 when simulated with different water models in Table~\ref{tab:apls_si} (Supplementary Information),
 however, all the values are close to the experimentally reported values.
-In conclusion, the ECC-lipid model reproduces the experimental dimensions of POPC
-lipid bilayer with the accuracy comparable to the other state of the art lipid models \cite{ollila16}.
+In conclusion, the ECC-lipids model reproduces the experimental dimensions of POPC
+lipid bilayer with the accuracy comparable to other state of the art lipid models \cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 The acyl chain order parameters of the Lipid14 model~\cite{dickson14} and the
-ECC-lipid model agree with the experimental values with in the error bars
-in Fig.~\ref{simVSexpNOions}, althought the ECC-lipid model gives slightly larger
+ECC-lipids model agree with the experimental values with in the error bars
+in Fig.~\ref{simVSexpNOions}, althought the ECC-lipids model gives slightly larger
 values for {\it sn}-1 chain. Notably, the experimentally measured forking and
 small order parameter values of $C_2$ segment in {\it sn}-2 chain are relatively well
-reproduced by the both models. This has been suggested to indicate that the carbonyl
+reproduced by both models. This has been suggested to indicate that the carbonyl
 of {\it sn}-2 chain is directed towards the water phase, in contrast to the
@@
 The order parameters of $\alpha$ and $\beta$ carbons in the headgroup are slightly larger
-in the ECC-lipid model than in the Lipid14 model, which is apparently related to the
-P-N vector orienting 7$^{\circ}$ more toward the water phase in the ECC-lipid model, 
+in the ECC-lipids model than in the Lipid14 model, which is apparently related to the
+P-N vector orienting 7$^{\circ}$ more toward the water phase in the ECC-lipids model, 
 see Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
 however, conclude which one of the models give the more realistic
-headgroup conformations. The ECC-lipid model gives
+headgroup conformations. The ECC-lipids model gives
 the $\beta$ carbon order parameter value closer to experiments, while
 value for $\alpha$ carbon is better in the Lipid14 model.
-Despite of some deviations from the experimental order parameter values
+Despite some deviations from the experimental order parameter values
 in Fig. \ref{simVSexpNOions},
@@
 the head group order parameters respond too sensitively to the bound positive charge.
-The slope in the ECC-lipid model is in very good agreement with experiments
+The slope in the ECC-lipids model is in a very good agreement with experiments
 for the $\alpha$ segment, while the slope is slightly
@@
 surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle,
-while the corresponding change in the ECC-lipid model is 11$^{\circ}$.
+while the corresponding change in the ECC-lipids model is 11$^{\circ}$.
 The difference is in line with the smaller order parameter 
 changes and the reduced charge--dipole interactions in the latter model.
-The lesser sensitivity of the P-N vector angle response in the ECC-lipid
+The lower sensitivity of the P-N vector angle response in the ECC-lipid
 model can be considered to be more realistic, because the changes of the headgroup
@@
 These results can be used to compare the ion binding affinities to lipid bilayers
-between simulations and experiments by using the electrometer concept, because 
-the order parameters decrease proportionally to the amount of bound positive charge~\cite{seelig87,catte16}. 
+between simulations and experiments using the electrometer concept, because 
+the order parameters decrease proportionally 
+to the amount of bound positive charge (Fig.~\ref{OrderParameterCHANGESsurf} and Ref.~\citenum{seelig87,catte16}). 
 The recent comparison of different simulation models to the experimental data
-revealed that most models significantly overestimate the Na$^+$ ion binding to PC
-lipid bilayers and that none of the available models correctly reproduce
+revealed that most models significantly overestimate Na$^+$ ion binding to PC
+lipid bilayers and that none of the available models correctly reproduces
 the details of the binding of Ca$^{2+}$ ions \cite{catte16}.
-A positive exception was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions,
-as seen from the data replotted from Ref.~\citenum{catte16} in Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}.
-The model reproduced the experimentally measured small order parameter changes with NaCl
+A positive exception in Ref.~\citenum{catte16} was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions.
+Such a combination of models reproduced the experimentally measured 
+small order parameter changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl})
 and imperceptible Na$^+$ binding to PC bilayers \cite{akutsu81,altenbach84}.
-However, the changes of the headgroup order parameters as a function of CaCl$_2$ 
-concentration were overestimated by the same combination of force field parameters.
+However, the changes of the headgroup order parameters as a function of CaCl$_2$ concentration 
+(Fig.~\ref{fig:delta_ordPar_CaCl}) were overestimated by the same combination of force field parameters.
 
 Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{auffinger07},
-we also simulated the Lipid14 model with the ion models by Dang et al.~\cite{smith94,chang1999,dang2006} and
-ECC-ions~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014} having
-more realistic bulk behaviour. Instead of the improvement in the binding behaviour,
-we observed overbinding also of Na$^+$ with these ions models, while the results
-with Ca$^{2+}$ were barely affected,
-%for the CaCl$_2$ interactions with a POPC bilayer described by the Lipid14 model,
-as seen in Figs. \ref{fig:delta_ordPar_CaCl} and S??.%\ref{fig:ordPars_waterModels} (in SI), respectively.
+we also simulated the Lipid14 model with more realistic ion models by Dang et al.~\cite{smith94,chang1999,dang2006}, and
+ECC-ions~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014}. 
+Instead of an improvement in the binding behaviour,
+we observed greatly overestimated binding for \ce{Ca^{2+}} and also for Na$^+$ with such ions models 
+(Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}). 
 \todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
@@
 suggesting that the improvements in the lipid parameters are
-required to correctly describe the divalent cation binding to PC lipid bilayers.
+required to correctly describe the binding of divalent cations to phospholipid bilayers.
 
@@
 %polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
-The results from the simulations with the ECC-lipid and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
+The results from the simulations with the ECC-lipids and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
 exhibit an improved behaviour of cation binding to a POPC bilayer in Fig.~\ref{fig:delta_ordPar_CaCl},
@@
 Since also the headgroup order parameter
-response the to the bound positive charge the ECC-lipid model
+response the to the bound positive charge the ECC-lipids model
 was in good agreement with experiments in section \ref{section:boundCHARGE},
@@
 Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.
-Also the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was
-signifincantly overestimated in the original Lipid14 model, while the ECC-lipid
-gives a good agreement with exepriments, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-Also the changes of the P-N vector angle are more prononced
-for the Lipid14 model, for which the maximum CaCl$_2$ induced tilting 
-toward water phase is approximately 17$^{\circ}$, while the corresponding
-value for ECC-lipid model is 6$^{\circ}$.
+
+In addition, the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was
+signifincantly overestimated in the original Lipid14 model, while the ECC-lipids model
+gives a greatly improved agreement with exepriments, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
+Also the changes of the P-N vector angle are more pronounced
+for the Lipid14 model, for which the largest tilting toward water phase 
+induced by a $780\,\mathrm{mM}$ CaCl$_2$ concentration
+is approximately 17$^{\circ}$, while the corresponding
+value for ECC-lipids model is 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
 
@@
 binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
-This probably applies also to the other lipid models considered in the previous study~\cite{catte16},
+This probably applies also to the other lipid models in a previous study~\cite{catte16},
 emphasizing the importance of the comparison of the lipid headgroup order parameter response
@@
 
-The ion binding affinity of ECC-lipid model with different water models is compared
-in SI. The performance with all water models is better than in in the original Lipid14
-parameteres, but the order parameter changes are slightly overestimated
-with all four-site water models and with TIP3P.
+The ion binding affinity of ECC-lipids model with different water models is compared in SI. 
+In general, the order parameter changes are slightly overestimated
+with the four-site water models, and with TIP3p.
+However, the performance of ECC-lipids with any of the water models 
+is better than the original Lipid14 model.
 
@@
 As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
-the relative surface excess for the ECC-lipid model, $\Gamma_{\rm Ca}^w = 0.07 \pm 0.01 \rm{nm}^{-2}$,
+the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^w = 0.07 \pm 0.01 \rm{nm}^{-2}$,
 is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
@@
 \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) 
-calculated using density profiles in Fig.~\ref{fig:nacl-dens}
+calculated using density profiles in Fig.~\ref{fig:cacl-dens}
 is qualitatively different from \ce{CaCl2}
@@
 %a larger Ca$^{2+}$ density peak in lipid headgroup region for
-%the Lipid14 model with Dang and ECC-ions than for the ECC-lipid model.
-
-\begin{figure}[tbp]
+%the Lipid14 model with Dang and ECC-ions than for the ECC-lipids model.
+
+\begin{figure}[tb]
   \centering
@@
       Contours of probability isodensities of \ce{Ca^{2+}} with respect to 
-      the phosphate oxygens of POPC from ECC-lipid simulation.
+      the phosphate oxygens of POPC from ECC-lipids simulation.
       %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
@@
 
-Since the same experimental data is reproduced by the ECC-lipid model,
+Since the same experimental data is reproduced by the ECC-lipids model,
 it can be used to give a more versatile interepration of the binding stoichiometry than
@@
 (see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
-\begin{figure}[tbp]
+\begin{figure}[tb]
   \centering
-  \includegraphics[width=6.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
   \caption{\label{fig:cacl_complexes}
@@
 Also the probabilities of different complexes formed by \ce{Na+} ions and POPC
-analyzed from the ECC-lipid model simulation with 1 M concentration of NaCl are
+analyzed from the ECC-lipids model simulation with 1 M concentration of NaCl are
 shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
@@
 to study the binding phenomena in equilibrium. However, the density profiles of \ce{Ca^{2+}}
-from the ECC-lipid simulation in Fig.~\ref{fig:cacl-dens} show a relatively shallow
+from the ECC-lipids simulation in Fig.~\ref{fig:cacl-dens} show a relatively shallow
 minimum at the interface, suggesting a smaller barrier for the exchange of calcium ions.
@@
 in POPC bilayer are shown in Fig.~S\ref{fig:hist_residence_times} in SI 
-from the ECC-lipid simulation with CaCl$_2$ concentration of 1~M
+from the ECC-lipids simulation with CaCl$_2$ concentration of 1~M
 and from the CHARMM36 simulation with 450~mM concentration of
@@
 }
-\todo{It might be also better to do the analysis from ECC-lipid simulation with concentration closer
+\todo{It might be also better to do the analysis from ECC-lipids simulation with concentration closer
 to the data in literature, i.e. 450mM.}
@@
 At least order of magnitude faster exchange is observed in the
-ECC-lipid model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are 
+ECC-lipids model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are 
 shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
-and the longest observed residence time in the ECC-lipid simulation is $141\,\mathrm{ns}$, 
+and the longest observed residence time in the ECC-lipids simulation is $141\,\mathrm{ns}$, 
 which is well below the total length of the simulation (200~ns).
@@
 
-In conclusion, the results from ECC-lipid model suggest that the
+In conclusion, the results from ECC-lipids model suggest that the
 exchange of bound calcium occurs within $\sim$100ns timeframe,

### 2017-11-03 Manuscript/manuscript.tex
Corrected section on binding affinites and oxygen moieties cations bind to.

@@
 
-\begin{figure}[tb]
+\begin{figure}[tb!]
   \centering
@@
 
-\begin{table}
+\begin{table}[tb!]
   \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
@@
 
-\begin{figure}[tb]
+\begin{figure}[htb!]
   \centering
@@
   }
-  \todo{Add temperatures for POPC and DPPC experiments} \\
+  \todo{Add temperatures for POPC and DPPC experiments} 
   %\todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
@@
   %\todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
-  \todo{Add temperatures to the experimental data.} \\
+  \todo{Add temperatures to the experimental data.} 
 \end{figure}
@@
 
-\subsection{Affinity and binding cites to POPC membrane}
-
-\begin{table}
-  \caption{The relative surface excess and the probabilities of contacts between bound calcium and different
-    oxygen moieties in POPC from different simulation models. The CaCl$_2$ concentration of the ions in water
+\subsection{Structure and affinity of cation binding to POPC membrane}
+
+\begin{table}[tb!]
+  \caption{Relative surface excess of calcium with respect to water, 
+  and the ratios of the amount of calcium cations 
+  bound to phosphate or carbonyl moieties 
+  to the amount of such cations bound to both moieties in a POPC bilayer 
+  from various simulation models. 
+  The CaCl$_2$ concentration of the ions in water before adsorption to the phospholipid bilayer
   is 350mM in the analyzed systems. \label{tab:binding}}
+  \todo{Add columns with concentrations of calcium -- bulk/molar fraction.}
+  \todo{Changing symbol $P$ to $R$ or $r$ might reflect better that it is rather a ratio, and not that much a probability.}
   \begin{tabular}{l|c c c}
@@
     \hline
-    ECC-lipids            &  $0.07 \pm 0.01 $            &  99\%   &    25\%    \\
+    ECC-lipids            &  $0.06 \pm 0.01 $            &  99\%   &    25\%    \\
     Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
     Lipid14/Dang          &  $ 0.3 \pm 0.03 $            &  -        &    -       \\
+    Lipid14/ECC-ions      &  $ \simeq 0.4   $            &  -        &    -       \\
   \end{tabular}
@@
 
-The binding affinities of Ca$^{2+}$ ions to POPC bilayer in different simulation 
-models were quantified by calculating the relative surface excess from Eq.~\ref{surfexcess} 
-using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
-The relative surface excess values, shown in Table~\ref{tab:binding},
-were calculated from simulations with the same, $350\,\mathrm{mM}$, concentration of CaCl$_2$ with 
+Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation 
+models were quantified by calculating the relative surface excess of calcium to water molecules, $\Gamma_{\rm Ca}^w$, from Eq.~\ref{surfexcess}.
+%using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
+The values of $\Gamma_{\rm Ca}^w$, shown in Table~\ref{tab:binding},
+were calculated from simulations with the same $350\,\mathrm{mM}$ concentration of CaCl$_2$ with 
 respect to water, i.e., the concentration before solvating the lipids in experiments.
 As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
-the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^w = 0.07 \pm 0.01 \rm{nm}^{-2}$,
-is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = 0.13 \pm 0.01 \rm{nm}^{-2}$,
-or with Dang ions, $\Gamma_{\rm Ca}^w = 0.3 \pm 0.03 \rm{nm}^{-2}$.
+the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^w = (0.06 \pm 0.01) \rm{nm}^{-2}$,
+is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = (0.13 \pm 0.01) \rm{nm}^{-2}$,
+or with Dang ions, $\Gamma_{\rm Ca}^w = (0.30 \pm 0.03) \rm{nm}^{-2}$,
+or with ECC-ions,  $\Gamma_{\rm Ca}^w \simeq 0.4 \rm{nm}^{-2}$,
+where the large uncertainity arises from a too small simulation box (Fig.~\ref{fig:cacl-dens}).
 Interestingly, the relative surface excess of 
 \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) 
-calculated using density profiles in Fig.~\ref{fig:cacl-dens}
+%calculated using density profiles in Fig.~\ref{fig:cacl-dens}
 is qualitatively different from \ce{CaCl2}
-having a negative value $\Gamma_i^w = -0.1 \pm 0.01 \rm{nm}^{-2}$, 
+having a negative value $\Gamma_i^w = (-0.11 \pm 0.01) \rm{nm}^{-2}$ 
 meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
 This is in contrast to the most of the available lipid force fields,
-which predict a specific binding of sodium on 
+which predict a specific binding of sodium to
 PC lipid bilayers \cite{catte16}.
@@
 
-\begin{figure}[tb]
+\begin{figure}[htbp!]
   \centering
-  \includegraphics[height=20.0cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
+  \includegraphics[height=19.5cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
   \caption{\label{fig:cacl-dens}
@@
 
-The probabilities for a bound \ce{Ca^{2+}} ion to interact with
-different oxygen moieties of POPC 
+The ratio of the amount of calcium cations 
+bound to phosphate or carbonyl moieties 
+to the amount of such cations bound to both moieties in a POPC bilayer 
+from various simulation models
+%The probabilities for a bound \ce{Ca^{2+}} ion to interact with
+%different oxygen moieties of POPC 
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
@@
 bound Ca$^{2+}$ ions are in contact with the phosphate oxygens,
-with two thirds (67\%) are interacting with only phosphate oxygens
-and the rest, one third (32\%), interacts also with the carbonyl oxygens. 
+with two thirds (67\%) interacting with only phosphate oxygens
+and the rest, one third (32\%), interacting also with the carbonyl oxygens. 
 Thus, the interactions between
-calcium ions with only carbonyl oxygens are very unlikely (1\%), but significant
-amount of the ions interact with both, phosphate and carbonyl oxygens.
-The individual lipids or acyl chains were not identified in the analysis,
-thus the simultaneous ion contacts with both oxygen moieties may be inter or intra molecular
-and occur with the carbonyls in the either of the acyl chains.
-The most likely contacts between Ca$^{2+}$ ions and phosphate oxygens is visualized with
-the probability isodensity contours in Fig.~\ref{fig:volmaps}.
+calcium ions with only carbonyl oxygens are very unlikely (1\%), however, a significant
+amount of the ions interacts with both, phosphate and carbonyl, oxygens moieties.
+The individual lipids or acyl chains were not distinguished in the analysis,
+hence, a simultaneous contacts of ions with both oxygen moieties may be inter or intra molecular,
+and may occur with carbonyls in either of the acyl chains.
+The most likely interactions between Ca$^{2+}$ ions and phosphate oxygens are visualized with
+the probability density isocontours in Fig.~\ref{fig:volmaps}.
 While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
@@
 
-\begin{figure}[tb]
+\begin{figure}[tb!]
   \centering
-  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
+  \includegraphics[width=8.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
 %  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
@@
   \caption{\label{fig:volmaps}
-      Contours of probability isodensities of \ce{Ca^{2+}} with respect to 
+      Isocontours of probability density of \ce{Ca^{2+}} with respect to 
       the phosphate oxygens of POPC from ECC-lipids simulation.
+      The probability density was evaluated around a single lipid, 
+      after its strucutral alignment using only phosphate group. 
       %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
@@
 
-Even thought \ce{Na+} ions bind to POPC bilayer with a very low probability,
-it was possible to analyze the distribution of contacts from simulations with a large
-sodium concetration of~1\,M.
+Even though \ce{Na+} ions do not specifically bind to a POPC bilayer,
+they still interact with its oxygen moieties. 
 %the average number of contacts per lipid with any oxygen atom in POPC is 
@@
 %$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
-The results show that \ce{Na+} ions bind only to the phosphate oxygens 
-of POPC with the probablity of 55\% and only to the carbonyl oxygens with the
-probability of 20\%, leaving the probability of 25\% for the
-simultaneous binding to both.%, phosphate and carbonyl oxygens.  
+The results  from simulation at $1\,$M \ce{NaCl} concentration show 
+that 55\% of \ce{Na+} ions interact with only phosphate oxygens of POPC 
+and 20\% with only carbonyl oxygens,
+and the rest, 25\%, is interacting with both.
+%, phosphate and carbonyl oxygens.  
 %This is in line with the calculated negative value of 
@@
 
-In conclusion, the results suggest that the calcium ions specifically bind to phosphate
-oxygens, while occasionally interacting also with carbonyls. This is in good agreement
+In conclusion, the results suggest that calcium ions specifically bind to phosphate
+oxygens occasionally interacting also with carbonyls. This is in good agreement
 with the previous conclusions from several experimental and theoretical
@@
 studies~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
-Also the sodium ions bind most likely to the phosphate oxygens, while the probability
-of binding purely to carbonyls is also significant, in contrast to calcium ions.
+Sodium ions also interact most likely with phosphate oxygens of POPC, 
+but in contrast to calcium, the interactions with only carbonyls are also significant.
 The physiological relevance of the sodium binding details is, however, uncertain
-due to the very weak binding affinity.
+due to its very weak binding affinity.
 
@@
 (see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
-\begin{figure}[tb]
+\begin{figure}[tb!]
   \centering
@@
 
-\begin{table}
+\begin{table}[tb!]
   \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
@@
 
-\begin{table}[btp]
+\begin{table}[btp!]
   \caption{Simulation parameters}

### 2017-11-04 Manuscript/manuscript.tex
Small correction to the Gamma values in MS.

@@
     Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
-    Lipid14/Dang          &  $ 0.3 \pm 0.03 $            &  -        &    -       \\
-    Lipid14/ECC-ions      &  $ \simeq 0.4   $            &  -        &    -       \\
+    Lipid14/Dang          &  $0.23 \pm 0.03 $            &  -        &    -       \\
+    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $            &  -        &    -       \\
   \end{tabular}
@@
 is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = (0.13 \pm 0.01) \rm{nm}^{-2}$,
-or with Dang ions, $\Gamma_{\rm Ca}^w = (0.30 \pm 0.03) \rm{nm}^{-2}$,
-or with ECC-ions,  $\Gamma_{\rm Ca}^w \simeq 0.4 \rm{nm}^{-2}$,
+or with Dang ions, $\Gamma_{\rm Ca}^w = (0.23 \pm 0.03) \rm{nm}^{-2}$,
+or with ECC-ions,  $\Gamma_{\rm Ca}^w (0.35 \pm 0.11) \rm{nm}^{-2}$,
 where the large uncertainity arises from a too small simulation box (Fig.~\ref{fig:cacl-dens}).

### 2017-11-04 Manuscript/manuscript.tex
Corrected section Stoichiometry.

@@
 \subsection{Structure and affinity of cation binding to POPC membrane}
+\label{sec:affinity}
 
@@
 
-\subsection{Binding stoichiometry to POPC membrane}
+\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
 
@@
 
-Simple binding models have been previous used to interpreted 
-the same experimental data, which is used in this work to validate
-the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}~\cite{altenbach84,macdonald87}.
+Simple binding models have been previously used to interpret
+the same experimental data \cite{altenbach84,macdonald87} 
+which are used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}.
 The experimental results from the lipid headgroup order parameters and from the
 atomic absorption spectroscopy were best
-explained by the ternary complex binding model with a binding
+explained by a ternary complex binding model with a binding
 stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
 However, also a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
-was successfully fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
+provided a good fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
 The results based on these binding models and the measured headgroup order parameter changes
@@
 complexes with various stoichiometries by counting the individual lipid
-molecules within the distance of $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion.
+molecules within the distance $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion
+as already used in previous section~\ref{sec:affinity}.
 The results from POPC simulation with 285~mM concetration of CaCl$_2$ in
@@
 ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
-the probabilities for the complexes with the stoichimetries of 1:1~(25\%)~and
-1:3~(34\%)~are only slightly lower,
-suggesting more complex binding stoichiometry than the simple ternary complex model.
+the probabilities of complexes with the stoichimetries of 1:1~(25\%)~and
+1:3~(34\%)~are only slightly lower
+suggesting a more complex binding model than the simple ternary complex model.
 %The observed propensities of
@@
 %$34\%$ for three lipids 
-However, one calcium binds roughly to two lipids on average also in simulations,
-because the probabilities of forming the complexes with 1 or 3 lipids are almost equal.  
-This probably explains the successfull fit of the ternary complex model to the
+However, the simulation data can also be viewed from such an angle that
+one calcium binds roughly to two lipids on average,
+because the probabilities of the complexes with 1 or 3 lipids are almost equal.  
+This probably explains the success of the ternary complex model fitting well the
 experimental data. 
@@
 simulations with the ECC-lipids model.
-\todo{I think that we either have to do the analysis in SI properly,
+\todo{Samuli: I think that we either have to do the analysis in SI properly,
   i.e., using the real $C_I$ from simulations or remove the sentence below and
-the related concent from SI.}
+the related concent from SI.\\
+Joe: I think it shows that ternary complex makes a good fit to our simulation data. 
+The fact that we use a different (and more accurate) estimate of $C_I$ and other quantities does not hamper this statement.}
 Ternary complex model also provides a good fit 
 to our simulations with ECC-lipids 
-(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
+(see Fig.~S\ref{fig:cacl-bind} in SI and its caption for details). 
 \begin{figure}[tb!]
@@
 
-Also the probabilities of different complexes formed by \ce{Na+} ions and POPC
-analyzed from the ECC-lipids model simulation with 1 M concentration of NaCl are
+The probabilities of different complexes formed by \ce{Na+} ions and POPC
+analyzed from the simulation with the ECC-lipids model at $1\,$M concentration of NaCl are also 
 shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the

### 2017-11-04 Manuscript/manuscript.tex
Corrected section Residence times and Conclusions.

@@
 Equilibration of \ce{Ca^{2+}} binding to POPC bilayer takes hundreds of
-nanoseconds in MD simulations with CHARMM36 and Slipids force fields~\cite{javanainen17}, suggesting that
-simulations with the length of several microseconds would be required
-to study the binding phenomena in equilibrium. However, the density profiles of \ce{Ca^{2+}}
-from the ECC-lipids simulation in Fig.~\ref{fig:cacl-dens} show a relatively shallow
-minimum at the interface, suggesting a smaller barrier for the exchange of calcium ions.
-To quantify the ion binding exchange in lipid bilayer simulations, we calculated 
-the residence times of ions within the distance of 0.3~nm from any oxygen atom
+nanoseconds in MD simulations with current state of the art force fields,
+e.g. CHARMM36 and Slipids force fields~\cite{javanainen17}, suggesting that
+simulations at a temporal scale of several microseconds are required
+for studies of cation binding in equilibrium. 
+This is in line with the density profiles of \ce{Ca^{2+}} 
+from simulations in Fig.~\ref{fig:cacl-dens}, 
+which all show a region of a very small calcium density 
+around the level of cholines in POPC (i.e. around $2.5\,\mathrm{nm}$)
+with the exception of ECC-lipids, 
+which exhibits only a relatively shallow local
+minimum at that level.
+This suggests that such a smaller barrier for the exchange of calcium ions 
+shall lead to their faster exchange with the solvent.
+To quantify the exchange of ions between the membrane and solvent in simulations, we calculated 
+residence times of ions bound to the membrane.
+We again use the definition for ions binding to the membrane as in the above section~\ref{sec:affinity},
+i.e. ion is bound if it is within 0.3~nm from any oxygen atom
 belonging to a lipid molecule in a bilayer.
 
-The residence time histograms of \ce{Ca^{2+}}
-in POPC bilayer are shown in Fig.~S\ref{fig:hist_residence_times} in SI 
-from the ECC-lipids simulation with CaCl$_2$ concentration of 1~M
-and from the CHARMM36 simulation with 450~mM concentration of
-ECC-ions taken from the previous study \cite{javanainen17}
-(the simulation data was downloaded from Ref.~\citenum{zenodo.259376}).
-\todo{JOE: Plot a histogram of residence times based on our simulations and simulation from \citenum{javanainen17} and put it into SI.
-}
-\todo{It might be also better to do the analysis from ECC-lipids simulation with concentration closer
-to the data in literature, i.e. 450mM.}
+The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer 
+at $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
+from simulations with ECC-lipids 
+and CHARMM36 (taken from previous study \cite{javanainen17}, 
+simulation data available at Ref.~\citenum{zenodo.259376})
+are shown in Fig.~S\ref{fig:hist_residence_times} in SI. 
 %Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
@@
 %and anaylsed the residence times.
-% simulation: (Matti C36 with CaClMK, 800ns total):
-Approximately 20\% of the \ce{Ca^{2+}} residence times in the CHARMM36 simulation are longer than
-the lenght of the trajectory ($800\,\mathrm{ns}$),
-and only less than 60\% of the bound residence times are shorter than the half
-of the simulation length~(400~ns).
+In the CHARMM36 simulation, approximately 20\% of the \ce{Ca^{2+}} residence times are longer than
+the length of the trajectory ($800\,\mathrm{ns}$).
+With such a simulation, which is hence relatively short for the employed model,
+we can merely estimate an upper bound that 
+only less than 60\% of the bound residence time 
+is formed by cations interacting with the phospholipid membrane 
+for less then half of the simulation length~(400~ns).
 Even longer residence times are observed
 in the other simulations with CHARMM36 and Slipids models reported in the previous work~\cite{javanainen17}.
-At least order of magnitude faster exchange is observed in the
+
+On the other hand, at least an order of magnitude faster exchange is observed in the
 ECC-lipids model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are 
@@
 and the longest observed residence time in the ECC-lipids simulation is $141\,\mathrm{ns}$, 
-which is well below the total length of the simulation (200~ns).
+which is well below the total length of the simulation used for analysis, 200~ns.
 The exchange of \ce{Na+} ions to a POPC membrane is
-an order of magnitude faster, giving 90\%
+another order of magnitude faster, giving 90\%
 of the residence times smaller than~$1\,\mathrm{ns}$,
 the longest residence time being~$6\,\mathrm{ns}$. 
-All the results are in line with the experimental estimation that 
+Note that the results from both analyzed models are in line with the experimental estimation that 
 the residence time of \ce{Ca^{2+}} at the individual PC headgroup 
@@
 
-In conclusion, the results from ECC-lipids model suggest that the
-exchange of bound calcium occurs within $\sim$100ns timeframe,
+In conclusion, the results from the ECC-lipids model suggest that the
+exchange of calcium between membrane and solvent occurs within $\sim 100\,$ns timeframe,
 which is much faster than previously reported~\cite{javanainen17}.
-The sodium exhibit even faster exchange.
-This suggests that simulations with the length of few hundred
-nanoseconds are sufficient to simulate equilibrium ion binding
-to PC lipid bilayers, when realistic force fields are used.
+Sodium cations exhibit even a faster exchange.
+This suggests that simulations with the length of a few hundreds of
+nanoseconds are sufficient to simulate ion binding
+to phospholipid bilayers in equilibrium, when realistic force fields are used.
 This has not been the case with the available lipid force fields,
-which overestimate the binding streght of the cations \cite{javanainen17,catte16}.
+which overestimate the binding strength of the cations \cite{javanainen17,catte16}.
 %Such a finding changes the point of view on binding of calcium to PC membranes from
@@
 
-By using the electrometer concept we show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
-to a POPC lipid bilayer can be accurately described with the classical MD simulation 
+We used the electrometer concept to show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
+to a POPC lipid bilayer can be accurately described with a classical MD simulation 
 force field, where the electronic polarization is implicitly included using 
@@
 While the structural details of a POPC lipid bilayer simulated with the ECC-lipids
-model agree with experiments with the comparable accuracy to the other state of the art lipid models,
+model agree with experiments with an accuracy comparable to the other state of the art lipid models,
 it also reproduces the experimental lipid head group order parameter responses to
-the cationic surfactant, NaCl and CaCl$_2$ concentrations.
-
-The good agreement with experiments enables the atomic resolution 
-interpretation of NMR experiments by using MD simulations.
-In agreement with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
-the Ca$^{2+}$ ions mainly interact with phosphate oxygens.
-However, the stoichiometry of the binding is significantly more complicated 
+a cationic surfactant, NaCl and CaCl$_2$ concentrations.
+
+The good agreement with experiments enables us to 
+interpret NMR experiments in atomistic details using MD simulations.
+In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
+Ca$^{2+}$ ions interact mainly with phosphate oxygens.
+However, the stoichiometry of calcium binding is significantly more complicated 
 than in the previous interpretation of the NMR data based on 
@@
 compex model is understandable based on the simulation results,
-a simple binding model cannot detec the complex binding
+a simple binding model cannot detect 
+the complex nature of calcium binding to phospholipid bilayers
 oberved in the simulation.
 
-The improved cation binding behaviour to POPC bilayer pave the way for
-simulations of complex biochemical systems with correctly described
+The accurate description of cation binding to POPC bilayer paves the way for
+simulations of complex biochemical systems with realistically described
 electrostatic interactions in the vicinity of cellular membranes.
 The ECC-lipids model is build by scaling the partial charges
-and the LJ-radius of the headgroup, glycerol backbone and carbonyl atoms of
+and the Van der Waals radii of the headgroup, glycerol backbone and carbonyl atoms of
 the Lipid14 POPC model~\cite{dickson14} with the constant factor
@@
 molecules, including water. The work toward this direction and the extension
-to other lipid molecules and force fields is left for the future studies.
+to other lipid molecules and force fields is left for future studies.
 
@@
 \todo{Add acknowledgments of Jiri Kolafa.}
+\todo{Joe: Thank Metacentrum.}
 \end{acknowledgments}
@@
     Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
+  }
+\end{figure}
+
+\begin{figure}[tbp!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_ECC-lipids_346mM_CaCl.pdf} \\
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_Charmm36_450mM_CaCl_Matti.pdf}
+  \caption{\label{fig:hist_residence_times}
+   Histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer 
+   at $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
+   from simulations with ECC-lipids/ECC-ions (top)
+   and CHARMM36/ECC-ions (bottom), which was taken from previous study \cite{javanainen17} 
+   (simulation data available at Ref.~\citenum{zenodo.259376}).
+   The maxima on the x-axis represent the lengths of the simulations used for analysis. 
+   For the ECC-lipids model, 90\% of the residence times of \ce{Ca^{2+}} in a POPC membrane are 
+   shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
+   with the longest observed residence time being $141\,\mathrm{ns}$, 
+   which is well below the total length of the simulation (200~ns).
+   This is, however not true for the simulation with CHARMM36 model,
+   where there are several calcium cations with their residence time 
+   apparently limited by the length of the simulation. 
+   With such a simulation, which is relatively short for the employed model,
+   we can merely estimate an upper bound that 
+   only less than 60\% of the bound residence time 
+   is formed by cations interacting with the phospholipid membrane 
+   for less then half of the simulation length~(400~ns).
   }

### 2017-11-06 Manuscript/manuscript.tex
Pavel's corrections to the MS.

@@
   The binding affinities and stoichimetries of Na$^+$ and Ca$^{2+}$ ions to phospholipid bilayers have not been
-  consistently determined with different experimental and theoretical methods
+  consistently determined with experimental and computational methods,
   despite their significance and biological relevance in cellular membranes. 
-  The ion binding details could be resolved with classical molecular dynamics (MD) simulations,
+  The details concerning ion binding could be resolved with classical molecular dynamics (MD) simulations,
   however, the accuracy of the available lipid models has not been sufficient.
-  In this work we show that the binding details of Na$^+$ and Ca$^{2+}$ ions to
+  In this work we show that the binding details of Na$^+$ and Ca$^{2+}$ ions to the
   1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
-  can be accurately described with a MD simulation model having implicitly included electronic polarization.
+  can be accurately described with a MD simulation model implicitly including electronic polarization.
   This is demonstrated by applying the electronic continuum correction (ECC) to a
-  state of the art lipid model for MD simulations of POPC
-  lipid bilayer. The introduced ECC-lipids model reproduces the experimentally measured structural parameters for the ion-free membrane,
-  the response of lipid headgroup to the bound positive charge, and the binding affinities of
-  Na$^+$ and Ca$^{2+}$ ions. The imperceptible binding of Na$^+$ ions to POPC bilayer
-  and interactions of Ca$^{2+}$ mainly with phoshpate oxygens in the ECC-lipids model give
+  state of the art lipid model for MD simulations of the POPC
+  lipid bilayer. The newly introduced ECC-lipids model reproduces the experimentally measured structural parameters for the ion-free membrane,
+  the response of lipid headgroup to a bound positive charge, and the binding affinities of
+  Na$^+$ and Ca$^{2+}$ ions. The observed negligible binding of Na$^+$ ions to POPC bilayer
+  and stronger interactions of Ca$^{2+}$ primarily with phoshpate oxygens give
   support to the early intepretations of the experimental spectroscopic data.
-  On the other hand, Ca$^{2+}$ ions form complexes with 1-3 lipid molecules with almost
-  equal probabilities, suggesting more complicated binding stoichimetry than 
+  At the same time, Ca$^{2+}$ ions form complexes with 1-3 lipid molecules with almost
+  equal probabilities, suggesting a more complicated binding stoichimetry than 
   the ternary complex binding model used to interpret the NMR data.
-  The results in this work pave the way for MD simulations of complex biochemical systems
+  The results of this work pave the way to accurate MD simulations of complex biochemical systems
   with realistic electrostatic interactions in the vicinity of cellular membranes.
@@
 \section{Introduction}
-Cation interactions with cellular membranes play a key role in several biological processes,
+Cation interactions with cellular membranes play a key role in important biological processes,
 such as signal propagation in neurons and vesicle fusion.
-Since the direct measurements of ion-membrane interactions from biological systems
+Since direct measurements of ion-membrane interactions in biological systems
 are difficult, lipid bilayers are often used as models systems for cellular
-membranes. Especially the zwitterionic phosphocholine (PC) lipid bilayers are
-used to understand the role of ions in complex biological
+membranes. In particular, the zwitterionic phosphocholine (PC) lipid bilayers are
+used as entry level models, helping to understand the role of ions in complex biological
 systems~\cite{scherer87,seelig90,cevc90}.
 
-Interactions of biological cations, especially Na$^+$ and
-Ca$^{2+}$, with PC bilayers are widely studied in
+Interactions of biologically relevant cations, especially Na$^+$ and
+Ca$^{2+}$, with PC bilayers have been widely studied in
 experiments~\cite{akutsu81,altenbach84,seelig90,cevc90,tocanne90,binder02,pabst07,uhrikova08}
@@
 %ions, but do not affect the specific binding constant \cite{seelig90}.
-Interpretations of non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering
-and infrared spectroscopy suggest that Na$^+$ ions exhibit
-negligible binding to PC lipid bilayers with submolar concentrations, while Ca$^{2+}$ is observed to specifically
+Interpretations of non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), as well as scattering
+and infrared spectroscopies, suggest that Na$^+$ ions exhibit only
+negligible binding to PC lipid bilayers wat submolar concentrations, while Ca$^{2+}$ is observed to specifically
 bind to phosphate groups of two lipid molecules~\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
-Atomistic resolution molecular dynamics (MD) simulation models, however, predict significantly
+Atomistic resolution molecular dynamics (MD) simulation models, however, typically predict a
 stronger binding for the cations than NMR experiments \cite{catte16}.
-On the other hand, some experiments have also been interpreted to support the predictions from
-MD simulations~\cite{bockmann03,vacha09a}. 
-Furthermore, interactions of calcium ions with 3-4 lipids,
-including also interactions with carbonyl
+On the other hand, several experiments have also been interpreted to support the predictions from
+earlier MD simulations~\cite{bockmann03,vacha09a}. 
+Furthermore, interactions of calcium ions with up to 3-4 lipids,
+including not only interactions with phosphates but also with carbonyl
 oxygens, have been reported from simulations~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
 
-Recent work published by the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
-made an attempt to resolve the apparent controversies. A direct comparison of
-ion binding affinities to PC bilayers was presented between simulations and
-experiments by using the electrometer concept, which is based on the experimental NMR data
+Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
+made an attempt to resolve these apparent controversies. A direct comparison of
+ion binding affinities to PC bilayers between simulations and
+experiments was done using the electrometer concept, which is based on experimental NMR data
 for the lipid headgroup order parameters~\cite{seelig87}.
-Using massive amounts of data collected by Open Collaboration method, it was concluded
+Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded
 that the accuracy of the current state of the art lipid models for MD
-simulations is not sufficient for the detailed interpretation of the cation
-interactions with PC lipid bilayers~\cite{catte16}.
+simulations is not sufficient for a detailed interpretation of the
+interactions of cations with PC lipid bilayers~\cite{catte16}.
 
@@
 
-In this work we show that the cation binding behavior
-in MD simulations of 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
-improved with an implicit inclusion of the electronic polarizability
+In this work, we show that the cation binding behavior
+in MD simulations of the 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
+improved via an implicit inclusion of the electronic polarizability
 in the polar region of phospholipids,
-which we apply according to the 
+which we apply using to the 
 electronic continuum correction (ECC)~\cite{leontyev11}.
 Such an approach has been previously shown to improve the behaviour of
-MD simulations of ions in bulk water~\cite{jungwirth17-new-paper-to-be-published,Pluharova2014,kohagen14,kohagen16}. 
-We use the the Lipid14 model \cite{dickson14} as a starting point,
+of ions in bulk water simulations ~\cite{jungwirth17-new-paper-to-be-published,Pluharova2014,kohagen14,kohagen16}. 
+We employ here the the Lipid14 model \cite{dickson14} as a starting point,
 because it provided one of the best available descriptions of cation binding \cite{catte16}.
-The developed ECC-lipids model reproduces the experimentally
+The newly developed ECC-lipids model reproduces the experimentally
 measurable structural parameters of an ion-free POPC lipid bilayer with the 
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-The lack of polarizability in standard MD simulation
-force fields has been considered an issue since the early days of
+The lack of electronic polarizability in standard MD simulation
+force fields has been considered a serious issue since the early days of
 lipid bilayer simulations.
 In this work, we circumvent the rather demanding explicit inclusion
-of electronic polarization effects \cite{lucas12,chowdhary13},
+of electronic polarization effects \cite{lucas12,chowdhary13} by
 including the electronic part of polarizability in lipid
-bilayer simulations implicitly through 
+bilayer simulations implicitly via 
 the electronic continuum correction (ECC)~\cite{leontyev11}. 
-Technically, it is similar to the phenomenological charge-scaling applied in 
-earlier studies \cite{jonsson86,egberts94}.
+Technically, this is similar to the phenomenological charge-scaling applied in 
+earlier studies of aqueous solutions or ionic liquids \cite{jonsson86,egberts94}.
 \todo{We should also cite papers where empirical scaling was used ionic liquids DOI: 10.1002/anie.201308760}
@@
 
-According to ECC, electronic
+According to ECC, the electronic
 polarizability can be implicitly included in classical MD simulations by
@@
 which is the electronic part of the dielectric constant of 
-the media \cite{leontyev11}. 
+the medium \cite{leontyev11}. 
 Following the Coulomb's law, such a dielectric continuum can be 
-easily included in a standard MD simulation by
-the scaling of charges,
+directly included in a standard MD simulation by
+scaling of charges,
 \begin{equation}
@@
 \end{equation}
-with a constant scaling factor $f_q = \epsilon _{el} ^{-1/2}$ 
-effectively representing the newly introduced electronic continuum. 
+with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$ 
+. 
 Given the  high frequency dielectric constant as measured in water (corresponding to the square of the refraction index) of  $\epsilon _{el} = 1.78$,  
 the scaling factor for aqueous ions is $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
-successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??},
+successfully used to improve the performance of force field simulations of ions in aqueous solution \cite{kohagen14,kohagen16,??},
 which then agree quantitatively with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
@@
 it is not a priori  clear whether the same factor should be used for partial charges in molecules,
-e.g., lipids in our case. Unlike the total charge of an atom or molecules, atomic partial charges within molecules are not physical observables.
-There are thus several schemes for the assignment of partial charges for biomolecules~\cite{Hu2007}.
+e.g., lipids in our case. Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
+Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}.
 Currently, the most commonly employed scheme is the restrained electrostatic potential (RESP) method~\cite{RESP_paper,Singh1984}.
 By construction, partial charges currently implemented in force fields may already 
-include  to some extend the solvent electronic polarizability effects, i.e.,
-the RESP charges are subsequently modified to fit certain experimental observables \todo{This needs a citation - Joe please pick 1-2 papers}. 
-Thus, a consistent application of the ECC scaling factor, $f_q$, to the molecular partial charges
-included in the available force fields does not necessarily have to follow the above relation $f_q = \epsilon _{el} ^{-1/2}$, 
-but instead it lies between 0.75 i.e., (no electronic polarizability included in the original partial charges)
+include  to some extend the electronic polarizability effects of the solvent, i.g.,
+the RESP charges may be subsequently modified to fit certain experimental observables \todo{This needs a citation - Joe please pick 1-2 papers}. 
+Thus, a consistent application of the ECC scaling factor, $f_q$, to the partial charges in molecules as
+included in the available force fields does not necessarily have to follow the above relation $f_q = \epsilon _{el} ^{-1/2}$. 
+Instead, a consistent scaling factor would lie between the value of 0.75 i.e., (no electronic polarizability included in the original partial charges)
 and $1$ (i.e., electronic polarizability fully included in the original partial charges).
 It is important to note that even though the lipid bilayer interface has a large gradient in
-the total dielectric constant, the electronic part of polarizability can be considered to be constant
-as the measurements of high frequency dielectric constant gives values around 2 
-for almost any biologically relevant material \cite{some_original_work, leontyev11}.
-
-In this work, we employ the ECC to develop a classical MD simulation model for
-a POPC lipid that accurately describes the binding of sodium and calcium ions to a lipid bilayer.
+the total dielectric constant, the electronic part of polarizability can be considered to be almost constant
+as measurements give values around 2 
+for almost any biologically relevant systems \cite{some_original_work, leontyev11}.
+
+In this work, we employ ECC to develop a force field model for
+the POPC lipid that accurately describes the binding of sodium and calcium ions to the corresponding lipid bilayer.
 %This is a biologically highly relevant membrane feature, which is
@@
 %in the presence of ions but also modulate interactions with charged moieties at the interface.
-The quality of the simulation model is compared 
-to the experimental NMR data used to measure the ion binding
+The quality of the simulation model is judged by comparison 
+to experimental NMR data reporting on the ion binding
 affinity~\cite{akutsu81,altenbach84,scherer89}, 
 as discussed in detail in Ref.~\citenum{catte16}.
-The Lipid14~\cite{dickson14} force field parameters (available in Gromacs format from Ref. \citenum{lipid14files})
-were used as a starting point, because they provide one of the most
+The Lipid14~\cite{dickson14} force field parameters (available in a Gromacs format from Ref. \citenum{lipid14files})
+were used as a starting point, as they provide one of the most
 realistic descriptions of the head group order parameter response to ions 
@@
 Moreover, the Lipid14 model has a relatively realistic glycerol backbone and
-head group structure compareable to other state of the art lipid models~\cite{botan15}.
-
-We applied the ECC correction to the Lipid14 model of POPC as a scaling of partial charges of the head group,
-glycerol backbone, and carbonyl regions, which are the most polar parts in phospholipids
-being expected to have the largest contribution to the cation binding.
-We do not modify the hydrocarbon chain parameters, 
-as they do not come in direct contact with salt ions, 
+head group structure comparable to other state of the art lipid models~\cite{botan15}.
+
+We applied the ECC correction to the Lipid14 model of POPC by scaling of partial charges of the head group,
+glycerol backbone, and carbonyl regions, which are the polar parts in phospholipids,
+expected to have a contribution to the cation binding.
+We did not modify the hydrocarbon chain parameters, 
+as they do not come in direct contact with ions from the solution 
 and as they are already highly optimized to provide a good description of the
 hydrophobic part of lipid bilayers \cite{dickson14,ollila16}.
-This is in contrasts with the glycerol backbone and the head group regions, 
-which call for improvements in all available lipid models~\cite{botan15}.
-
-The values within the above discussed parameter space for the scaling factor, $f_q~\in~(0.75, 1.0)$,
-were explored to find the optimal value to reproduce the experimental ion binding affinity.
-We observed that the scaling
+The latter is in contrasts with the situation for the glycerol backbone and the head group regions, 
+which calls for improvements in available lipid models~\cite{botan15}.
+
+The above discussed interval for the scaling factor, $f_q~\in~(0.75, 1.0)$,
+was explored to find the optimal value reproducing the experimental ion binding affinities.
+We first observed that the scaling
 of the partial charges reduced the ion binding affinity and the 
 related head group order parameter response in general.
-We compared the results with different scaling factors to the experimental 
+We then compared the results with different scaling factors to the experimental 
 NMR data~\cite{akutsu81,altenbach84,scherer89} finding the optimal 
 value $f_q = 0.8$,
-which is only slighly higher than the scaling factor for the ions in water, 0.75.
-
-Although the scaling of the partial charges improved 
+which is only slighly higher than the original scaling factor for the ions in water of 0.75.
+
+Although scaling of the partial charges improved 
 the head group order parameter response and ion binding affinity,
-it also deteriorated some membrane properties; 
+it at the same time deteriorated certain membrane properties; 
 namely the area per lipid decreased below the experimental value. 
-After only scaling the partial charges of the head group, the glycerol
-backbone and carbonyls with a factor of 0.8, 
-the area per lipid in the simulation drops to approximately 60~\AA$^2$,
-which is significantly smaller than the experimental value, 64.3~\AA$^2$ \cite{kucerka11}, 
-as well as the value from the original Lipid14 model, $65.6 \pm 0.5$~\AA$^2$~\cite{dickson14}.
+After directly scaling the partial charges of the head group, the glycerol
+backbone, and carbonyls with a factor of 0.8, 
+the area per lipid in the simulation drops approximately to 60~\AA$^2$,
+which is smaller than the experimental value of 64.3~\AA$^2$ \cite{kucerka11}, 
+as well as the value from the original Lipid14 model of $65.6 \pm 0.5$~\AA$^2$~\cite{dickson14}.
 The decrease of the area per lipid is found to arise from a reduced hydration of the lipid head group region
-as a consequence of the lower polarity after scaling charges. 
-This effect can be compensated
+due a reduced head group polarity after scaling charges. 
+This effect is compensated
 by reducing the effective radii of atoms with the scaled charges.
 This is in practice done by changing the parameters $\sigma$ in the Lennard-Jones potential 
-in a similar way as was previously done for the ECC-ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
-After reducing the $\sigma$ parameters of affected atoms with a scaling factor $f_\sigma = 0.89$, 
+in a similar way as was done previously for the ECC ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
+After reducing the $\sigma$ parameters of affected atoms by a factor of $f_\sigma = 0.89$, 
 the area per molecule is restored
-back close to the experimental value (Table~\ref{tab:apls}).
+back, being very close to the experimental value (Table~\ref{tab:apls}).
 
@@
 The change of the order parameters measured with varying aqueous
-ion concentration can be then related to the amount of bound ions.
-
-The concept can be used to compare the ion binding affinity to lipid 
+ion concentrations can be then related to the amount of bound ions.
+
+The electrometer concept can be used to compare the ion binding affinity to lipid 
 bilayers between MD simulations and  NMR experiments, because the order
@@
 decrease with a bound positive charge and increase with a bound negative
-charge~\cite{ollila16,catte16}. This can be rationalized as a change
+charge~\cite{ollila16,catte16}. This can be rationalized in terms of a change
 of the lipid head group dipole tilt towards water phase with the bound positive
@@
 
-The measured order parameter change depends on both the head group response 
-to the bound charge, and the amount of bound charge 
+The measured change of the order parameter depends both on the head group response 
+to the bound charge and on the amount of the bound charge 
 (i.e. $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively). 
 The empirical factor $m_i$ has to be
-well quantified before the electrometer concept can used to 
+properly quantified before the electrometer concept can used to 
 analyze the binding affinities.
-This was done experimentally for a wide range
+This has been done experimentally for a wide range
 of systems~\cite{seelig87,beschiasvili91}. To calibrate 
-the head group order parameter response to the bound charge in simulations,
-we use the experimental data for dihexadecyldimethylammoniumbromide %(2C$_{16}$$^+$N2C$_1$Br$^-$)
+the response of the head group order parameter to the bound charge in simulations,
+we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide %(2C$_{16}$$^+$N2C$_1$Br$^-$)
 mixed within a POPC bilayer~\cite{scherer89}. Dihexadecyldimethylammonium\\[0.5cm]
@@
 \vspace{0.5cm} \\
-is a cation surfactant having two acyl chains and bearing a unit
+is a cationic surfactant having two acyl chains and bearing a unit
 charge in the hydrophilic end. Thus, it is expected to locate
-in the bilayer similarly to the phospholipids and the molar ratio then 
+in the bilayer similarly to the phospholipids and its molar ratio then 
 gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
@@
 In this work, we focus on POPC for which the latter definition was used~\cite{altenbach84}.
-The salt concentration in the aqueous bulk region was calculated from 
-the farthest point from both lipid leaflets in the water phase. 
+The salt concentration in the aqueous bulk region was taken as that at 
+the farthest point from the lipid bilayer in the aqueous phase. 
 Note that in the previous study~\cite{catte16},
-the ion concentrations were calculated in water before solvating the lipids 
+the ion concentrations were calculated in water before solvating the lipids, 
 as in the earlier experiments~\cite{akutsu81}.
-Although there are measurable differences between these two
-definitions of concentrations for CaCl$_2$ systems,
-they do not affect the qualitative conclusions in this or in the previous work~\cite{catte16}.
-
-To quantify the ion binding affinity to a lipid bilayer, we
+Although there are noticable differences between these two
+definitions for concentrations of the CaCl$_2$ systems,
+they do not affect significantly conclusions drawn in this work or in the previous one~\cite{catte16}.
+
+To quantify the ion binding affinities to a lipid bilayer, we
 calculate the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
 Such a quantity compares the adsorption of ions to the adsorption of water molecules 
-at an interface without the necessity of defining the Gibbs dividing plane 
+at the interface without the necessity of defining a Gibbs dividing plane 
 between the membrane interior and the water bulk region. 
-In our simulations we assume that the interface locates between the hydrophobic interior
-of a lipid bilayer and the bulk water region far from the membrane.
-The bulk concentration of ions and water is zero inside the bilayer.
-The concentrations in bulk water region can be calculated from 
-the farthest point from both lipid leaflets in the water phase.
-The region between these boundaries contains all the ions and water
-molecules in the simulation box.
+In our simulations we only assume that the interface is located between the ion-free hydrophobic interior
+of the lipid bilayer and the aqueous region far from the membrane.
 %our whole simulation box as
@@
 %the hydrophobic region in the middle of the membrane where no ions or waters are present.
-Such a setup provides a simplified relation for $\Gamma_i^w$ for simulations of lipid bilayers, 
+Such a setup and the above definition of bulk ion concentration provide a simple relation for the normalized surface excess $\Gamma_i^w$ for simulations of lipid bilayers, 
 \begin{equation}\label{surfexcess}
-  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
+  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) .
 \end{equation}
-where $n_w$ and $n_i$ are the total number of waters and ions in the system;
-$C_w$ and $C_i$ are their respective bulk concentrations in the aqueous phase;
-and $A_b$ is the area of the box in the membrane plane.
-The total area of the interface is twice the area of the membrane, $2A_b$,
-because bilayers have an interface at both leaflets.
+Here, $n_w$ and $n_i$ are the total numbers of water molecules and ions in the system,
+$C_w$ and $C_i$ are their respective bulk concentrations in the aqueous phase,
+and $A_b$ is the area of the unit cell in the membrane plane.
+The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
+because the bilayer has an interface at each of the two leaflets.
 
@@
 
-The structure of lipid bilayers in simulations without ions were validated against NMR and 
+The structures of lipid bilayers in simulations without ions were validated against NMR and 
 \mbox{x-ray} scattering experiments by calculating the order parameters
-for C-H bonds and the scattering form factors. The former validates the structures sampled by the individual
+for the C-H bonds and the scattering form factors. The former validates the structures sampled by the individual
 lipid molecules in simulations with atomic resolution, while the latter
-validates the dimensions of the lipid bilayer (thickness and area per
+validates the dimensions of the lipid bilayer (i.e., its thickness and area per
 molecule)~\cite{ollila16}.
 
-The order parameters were calculated from simulations for all C-H bonds
+The simulated order parameters were calculated for all C-H bonds
 in lipid molecules from Eq. \ref{OP}. Form factors were calculated 
@@
 where $\rho_{el} (z)$ is the total electron density,
-$\rho_{el}^s$ is the electron density of the solvent far in the bulk,
+$\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk,
 and $z$ is the distance from the membrane centre along its normal 
-spanning until the water bulk region, $D/2$.
+with $D/2$ being half of the unit cell size.
 
@@
 \subsubsection{Simulations of POPC bilayers with aqueous ions}
-Simulations of a POPC bilayer in a pure water or in varying salt concentrations
-contained 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per each lipid
+Simulations of a POPC bilayer in pure water or at varying salt concentrations
+consisted of 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per each lipid
 in an orthorhombic simulation box with periodic boundary conditions. 
-The chosen water model was used in the previous parametrization of 
+The chosen SPC/E water model, used also in the previous parametrization of 
 ECC-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014},
-and it has dielectric properties consistent with ECC~\cite{leontyev11, leontyev14}.
-Results with other water models are available in SI
+has a lowered dielectric constant, consistent with the ECC concept~\cite{leontyev11, leontyev14}.
+Results from simulations with other water models 
 (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, 
@@
 TIP3p-FB and TIP4p-FB~\cite{Wang2014},
-and TIP4p/2005 \cite{Abascal2005}).
-Sodium, calcium and chloride ions 
+and TIP4p/2005 \cite{Abascal2005}), performed to check for robustness of our approach, are available in SI.
+Sodium, calcium, and chloride 
 were modeled as ECC-ions with parameters from \citenum{jungwirth17-new-paper-to-be-published, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Simulation with the Lipid14 model used the TIP3p water model \cite{jorgensen83} and the ion model by
+Scaled charges and van der Waals radii for atoms forming the POPC lipid were derived here, starting from the Lipid14 force field.
+For comparison, simulation with unscaled charges were perfomed using the Lipid14 model used the TIP3p water model \cite{jorgensen83} and the ion model by
 Dang et al.~\cite{smith94,chang1999,dang2006}.
@@
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
-The simulation settings used in this work are 
+The simulation settings employed in this work are 
 summarized in Table \ref{tbl:mdpar}. 
-Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+The simulatied trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
 
 \subsubsection{Simulations of POPC bilayers with cationic surfactants}
-The structure of dihexadecyldimethylammonium
+The topology of dihexadecyldimethylammonium
 was created with an automated topology builder \cite{malde11}. 
 The AmberTools program \cite{amber} was then used to generate the Amber-type force field
-parameters. The parameters were converted to the Gromacs format with
+parameters. These parameters were then converted to the Gromacs format with
 the acpype tool \cite{acpype}. The partial charges were then manually modified
@@
 at~\cite{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
-The surfactants were randomly placed among the lipids to form bilayer structures with
+The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with
 mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
-All systems contained 50 POPC molecules per leaflet, 6340 TIP3P~\cite{jorgensen83} water molecules and
+All these systems contained 50 POPC molecules per leaflet, 6340 TIP3P~\cite{jorgensen83} water molecules, and
 6, 14, 21, 35, or 50 surfactants per leaflet. 
-Note that chloride counter ions were used in simulations, 
-but bromide was used as a counter ion in the experiment \cite{scherer89}.
+Chloride counter ions were used in simulations, 
+while bromide was the counter ion in the experiment \cite{scherer89}.
 The Lipid14 model was used for POPC.
@@
 was considered as an equlibration time and was omitted from the analysis.
-A reasonable lipid neighbor exchange occured during the simulations.
-
-The same systems were also simulated with the ECC-lipids model for POPC 
+Sufficient lipid neighbor exchange occured during the simulations.
+
+The systems were simulated with the ECC-lipids model for POPC 
 and SPC/E model for water using the same setup.
 In these simulations ECC was also applied to the cationic surfactant 
-in the same way as to POPC;
-all charges were scaled with the same factor as for ECC-lipids, i.e. $f_q=0.8$, 
+in the same way as to POPC, i.e.,
+all partial charges were scaled with the same factor as for ECC-lipids, i.e. $f_q=0.8$, 
 and the same atom types with reduced $\sigma$ parameters from ECC-lipids were used. 
+For comparison, simulations with unscaled charges were also performed.
+
 \todo{Provide the parameteres for ECC-surfactant somewere.}
@@
 
-The ECC-lipids and Lipid14 models reproduce the experimental x-ray scattering form factors
-of a POPC bilayer with a similar accuracy in Fig.~\ref{simVSexpNOions}.
-The area per lipid values from the Lipid14 model is slightly larger than the
-experimental value in Table~\ref{tab:apls}, while the value from ECC-lipids model
-is slightly smaller. The values of the area per lipid of the ECC-lipids model show some variation
-when simulated with different water models in Table~\ref{tab:apls_si} (Supplementary Information),
-however, all the values are close to the experimentally reported values.
-In conclusion, the ECC-lipids model reproduces the experimental dimensions of POPC
-lipid bilayer with the accuracy comparable to other state of the art lipid models \cite{ollila16}.
+First, we present results for bilayers in pure water.
+The ECC-lipids and Lipid14 models both reproduce the experimental x-ray scattering form factors
+of a POPC bilayer with a comparable accuracy in Fig.~\ref{simVSexpNOions}.
+The area per lipid from the Lipid14 model is slightly larger than the
+experimental value in Table~\ref{tab:apls}, while the value from the ECC-lipids model
+is slightly smaller. The values of the area per lipid of the ECC-lipids model show small variations
+when simulated with different water models (see Table~\ref{tab:apls_si} in SI),
+with all being close to the experimentally reported values.
+We can thus conclude that the ECC-lipids model reproduces the experimental dimensions of the POPC
+lipid bilayer with accuracy comparable to that for other state of the art lipid models \cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 
-The acyl chain order parameters of the Lipid14 model~\cite{dickson14} and the
-ECC-lipids model agree with the experimental values with in the error bars
-in Fig.~\ref{simVSexpNOions}, althought the ECC-lipids model gives slightly larger
-values for {\it sn}-1 chain. Notably, the experimentally measured forking and
-small order parameter values of $C_2$ segment in {\it sn}-2 chain are relatively well
-reproduced by both models. This has been suggested to indicate that the carbonyl
+Similarly, the acyl chain order parameters of the
+ECC-lipids model, as well as the Lipid14 model~\cite{dickson14}, agree 
+with the experimental values with in the error bars, as presented
+in Fig.~\ref{simVSexpNOions}. Notably, the experimentally measured forking and
+small order parameter values of $C_2$ segment in {\it sn}-2 chain are well
+reproduced by both models. This feature has been suggested to indicate that the carbonyl
 of {\it sn}-2 chain is directed towards the water phase, in contrast to the
 carbonyl in {\it sn}-1 chain, which would orient more along the bilayer
-plane~\cite{seelig75,schindler75,gawrisch92}. While this may be an important
-feature for the ion binding details, it is not necessarily reproduced by the
+plane~\cite{seelig75,schindler75,gawrisch92}. This may be an important
+feature for the ion binding details, which it is not fully reproduced by other
 available lipid models~\cite{ollila16}
@@
 
-The order parameters of $\alpha$ and $\beta$ carbons in the headgroup are slightly larger
+The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger
 in the ECC-lipids model than in the Lipid14 model, which is apparently related to the
-P-N vector orienting 7$^{\circ}$ more toward the water phase in the ECC-lipids model, 
-see Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
-however, conclude which one of the models give the more realistic
+P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-lipids model, 
+see Fig.~\ref{OrderParameterCHANGESsurf}. This is a small change and with the current experimental data we cannot decide which of the two models provides more realistic
 headgroup conformations. The ECC-lipids model gives
-the $\beta$ carbon order parameter value closer to experiments, while
-value for $\alpha$ carbon is better in the Lipid14 model.
-Despite some deviations from the experimental order parameter values
+the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while
+the oposite is true for the $\alpha$ carbon.
+Despite small deviations from the experimental order parameter values
 in Fig. \ref{simVSexpNOions},
-the accuracy of the both models in the glycerol backbone region
-is comparable to the other state
+the accuracy of both models in the glycerol backbone region
+is comparable to other state
 of art lipid models available in literature \cite{botan15}.
@@
 
-Before proceeding to the ion binding affinity, we quantify
+Before proceeding to the sodium and calcium  ion binding affinities, we quantify
 the response of the headgroup order parameters to the amount of 
@@
 and POPC~\cite{scherer89}. The amount of bound charge per PC 
-in these systems is given by the molar fraction of cationic 
-surfactants, because essentially all surfactants with two hydrophopic
+in these systems is given simply by the molar fraction of cationic 
+surfactants, because essentially all the surfactants with two hydrophopic
 acyl chains can be assumed to locate in the lipid bilayers.
@@
 
-The changes of the headgroup order parameters with an increasing amount of 
-the cationic surfactant is compared between experiments~\cite{scherer89} and
-simulations in Fig. \ref{OrderParameterCHANGESsurf}.
-Approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ},
-is observed in simulations and experiments at least for the mole fractions
+The changes of the headgroup order parameters with increasing amount of 
+the cationic surfactant are compared for simulations and experiments~\cite{scherer89} in Fig. \ref{OrderParameterCHANGESsurf}.
+An approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ},
+is observed in both simulations and experiments at least for mole fractions
 below $\sim$30\%. The slope is, however, too steep in the Lipid14 model indicating that 
 the head group order parameters respond too sensitively to the bound positive charge.
-The slope in the ECC-lipids model is in a very good agreement with experiments
-for the $\alpha$ segment, while the slope is slightly
+In contrast, the slope for the ECC-lipids model is in a very good agreement with experiments
+for the $\alpha$ segment, while being slightly
 underestimated for the $\beta$ segment.
@@
 are in better agreement with experiments in this model.
-The results also imply that the validation and improvements of MD simulation
-models are generally required for the reliable simulation studies of the
-lipid headgroup responses to ions or other biomolecules, as also concluded
+The results also imply that validation and improvements of MD simulation
+models are generally required for reliable simulation studies of the
+lipid headgroup responses to ions or other biomolecules, as also suggested
 previously \cite{botan15}.
@@
   \caption{\label{fig:delta_ordPar_CaCl}
-    The changes of the head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
+    Changes of the head group order parameters of POPC bilayer as functions of the CaCl$_2$ concentration
     are shown from simulations with different force fields together with experimental data 
@@
 
-The changes of the lipid bilayer headgroup order parameter from different simulations and
+Changes of the lipid bilayer headgroup order parameters from different simulations and
 experiments \cite{akutsu81,altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}
-as a function of NaCl and CaCl$_2$ concentrations.
+as functions of NaCl or CaCl$_2$ concentrations.
 %Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.
@@
 the order parameters decrease proportionally 
-to the amount of bound positive charge (Fig.~\ref{OrderParameterCHANGESsurf} and Ref.~\citenum{seelig87,catte16}). 
-The recent comparison of different simulation models to the experimental data
-revealed that most models significantly overestimate Na$^+$ ion binding to PC
+to the amount of the bound positive charge (Fig.~\ref{OrderParameterCHANGESsurf} and Ref.~\citenum{seelig87,catte16}). 
+A recent comparison of different simulation models to the experimental data
+revealed that most force fields significantly overestimate Na$^+$ ion binding to PC
 lipid bilayers and that none of the available models correctly reproduces
 the details of the binding of Ca$^{2+}$ ions \cite{catte16}.
-A positive exception in Ref.~\citenum{catte16} was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions.
-Such a combination of models reproduced the experimentally measured 
+The only exception in Ref.~\citenum{catte16} was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions,
+which reproduced the experimentally measured 
 small order parameter changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl})
-and imperceptible Na$^+$ binding to PC bilayers \cite{akutsu81,altenbach84}.
-However, the changes of the headgroup order parameters as a function of CaCl$_2$ concentration 
-(Fig.~\ref{fig:delta_ordPar_CaCl}) were overestimated by the same combination of force field parameters.
-
-Since significant artefacts are reported in simulations with \AA{}qvist ions is water \cite{auffinger07},
+and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81,altenbach84}.
+However, the change of the headgroup order parameters as a function of CaCl$_2$ concentration 
+(Fig.~\ref{fig:delta_ordPar_CaCl}) was still grosly overestimated within this model.
+
+Since artifacts were reported in simulations with \AA{}qvist ions is water \cite{auffinger07},
 we also simulated the Lipid14 model with more realistic ion models by Dang et al.~\cite{smith94,chang1999,dang2006}, and
 ECC-ions~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014}. 
-Instead of an improvement in the binding behaviour,
-we observed greatly overestimated binding for \ce{Ca^{2+}} and also for Na$^+$ with such ions models 
+However, we still observed overestimated binding for both \ce{Ca^{2+}} and Na$^+$ with these ions models 
 (Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}). 
@@
 The results are in line with the previous work \cite{catte16},
-suggesting that the improvements in the lipid parameters are
-required to correctly describe the binding of divalent cations to phospholipid bilayers.
+suggesting that additional improvements in the lipid parameters are
+required to correctly describe the binding of cations to phospholipid bilayers.
 
@@
 %polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
-The results from the simulations with the ECC-lipids and the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
-exhibit an improved behaviour of cation binding to a POPC bilayer in Fig.~\ref{fig:delta_ordPar_CaCl},
+The results from the simulations combining the ECC-lipids with the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
+exhibit a significanlty improved behavior of cation binding to the POPC bilayer in Fig.~\ref{fig:delta_ordPar_CaCl},
 showing a good agreement with experiments in the changes of 
-the lipid headgroup order parameters as a function of NaCl and CaCl$_2$ concentrations.
+the lipid headgroup order parameters as functions of NaCl or CaCl$_2$ concentrations.
 Since also the headgroup order parameter
-response the to the bound positive charge the ECC-lipids model
-was in good agreement with experiments in section \ref{section:boundCHARGE},
-we conclude that the model correctly reproduces the binding affinity of
-Na$^{+}$ and Ca$^{2+}$ ions to POPC lipid bilayer.
-
-In addition, the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was
-signifincantly overestimated in the original Lipid14 model, while the ECC-lipids model
-gives a greatly improved agreement with exepriments, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-Also the changes of the P-N vector angle are more pronounced
+response of the ECC-lipids model the to the bound positive charge 
+is in good agreement with experiment (see above section \ref{section:boundCHARGE}),
+we can conclude that this model correctly reproduces the binding affinities of
+Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
+
+In addition, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was
+signifincantly overestimated in the original Lipid14 model, the ECC-lipids model
+provides a greatly improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
+Also the changes of the P-N vector angle are too pronounced
 for the Lipid14 model, for which the largest tilting toward water phase 
@@
 is approximately 17$^{\circ}$, while the corresponding
-value for ECC-lipids model is 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
-
-Furthermore, the overestimated lipid headgroup order parameter changes of POPC in the Lipid14
-model as a function of CaCl$_2$ concentration arise from both, the overestimated
-binding affinity and sensitivity of the headgroup tilt to the bound positive charge.
-This probably applies also to the other lipid models in a previous study~\cite{catte16},
-emphasizing the importance of the comparison of the lipid headgroup order parameter response
-to the bound charge between simulations and experiments as done in section \ref{section:boundCHARGE}.
-
-The ion binding affinity of ECC-lipids model with different water models is compared in SI. 
-In general, the order parameter changes are slightly overestimated
-with the four-site water models, and with TIP3p.
-However, the performance of ECC-lipids with any of the water models 
-is better than the original Lipid14 model.
+value for ECC-lipids model is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
+
+Within the Lipid14 model, 
+the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated
+binding affinity and the too strong sensitivity of the headgroup tilt to the bound positive charge.
+This likely applies also to the other lipid models tested in a previous study~\cite{catte16},
+which underlines the importance of the comparison of the lipid headgroup order parameter response
+to the bound charge between simulations and experiments as presented in section \ref{section:boundCHARGE}.
+
+Finally, the ion binding affinities for the ECC-lipids model with different water models are compared in SI. 
+In general, the performance of ECC-lipids with any of the water models 
+is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated
+with the four-site water models and with TIP3p model.
 
@@
 Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation 
-models were quantified by calculating the relative surface excess of calcium to water molecules, $\Gamma_{\rm Ca}^w$, from Eq.~\ref{surfexcess}.
+models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm Ca}^w$, from Eq.~\ref{surfexcess}.
 %using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
 The values of $\Gamma_{\rm Ca}^w$, shown in Table~\ref{tab:binding},
-were calculated from simulations with the same $350\,\mathrm{mM}$ concentration of CaCl$_2$ with 
-respect to water, i.e., the concentration before solvating the lipids in experiments.
+were calculated from simulations with the same concentration of CaCl$_2$ of $350\,\mathrm{mM}$  with 
+respect to water before solvating the lipids.
 As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
@@
 %calculated using density profiles in Fig.~\ref{fig:cacl-dens}
-is qualitatively different from \ce{CaCl2}
+is not only quantitatively but also qualitatively different from \ce{CaCl2}
 having a negative value $\Gamma_i^w = (-0.11 \pm 0.01) \rm{nm}^{-2}$ 
@@
 This is in contrast to the most of the available lipid force fields,
-which predict a specific binding of sodium to
+which predict a positive binding of sodium to
 PC lipid bilayers \cite{catte16}.
@@
 
-The ratio of the amount of calcium cations 
-bound to phosphate or carbonyl moieties 
-to the amount of such cations bound to both moieties in a POPC bilayer 
+In a POPC bilayer, the ratio of the amount of calcium cations 
+bound to either phosphate or carbonyl moieties 
+to the amount of such cations bound to both moieties
 from various simulation models
@@
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
-were analyzed by counting the contacts within the distance of $0.3 \, \mathrm{nm}$, 
+was analyzed by counting the contacts within the distance of $0.3 \, \mathrm{nm}$, 
 as done previously in Ref.~\citenum{javanainen17}. 
@@
 bound Ca$^{2+}$ ions are in contact with the phosphate oxygens,
-with two thirds (67\%) interacting with only phosphate oxygens
-and the rest, one third (32\%), interacting also with the carbonyl oxygens. 
-Thus, the interactions between
-calcium ions with only carbonyl oxygens are very unlikely (1\%), however, a significant
-amount of the ions interacts with both, phosphate and carbonyl, oxygens moieties.
+with two thirds (67\%) interacting only with these species,
+while the remaining one third (32\%), interacting also with the carbonyl oxygens. 
+The interactions between
+calcium ions with only carbonyl oxygens are very unlikely (1\%).
 The individual lipids or acyl chains were not distinguished in the analysis,
-hence, a simultaneous contacts of ions with both oxygen moieties may be inter or intra molecular,
+hence, simultaneous contacts of ions with both oxygen moieties may be inter or intra molecular,
 and may occur with carbonyls in either of the acyl chains.
@@
 the probability density isocontours in Fig.~\ref{fig:volmaps}.
-While the higher concentrations of \ce{CaCl2} naturally increased the amount of contacts per lipid,
+While the higher concentrations of \ce{CaCl2} naturally increases the amount of contacts per lipid,
 %for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
@@
 %$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
-the distribution of contacts between phosphate and carbonyl oxygens was not affected.
+the distribution of contacts between phosphate and carbonyl oxygens is not affected.
 %and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.
@@
 
-In conclusion, the results suggest that calcium ions specifically bind to phosphate
-oxygens occasionally interacting also with carbonyls. This is in good agreement
-with the previous conclusions from several experimental and theoretical
+In conclusion, the results suggest that calcium ions bind specifically to phosphate
+oxygens, occasionally interacting also with carbonyls. This is in good agreement
+with previous conclusions from several experimental and computational
 studies~\cite{hauser76,hauser78,herbette84,cevc90,binder02}, but
-suggests lower relative binding affinity to carbonyls than the previous MD simulation
+suggesting a lower relative binding affinity to the carbonyls than inferred from previous MD simulation
 studies~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
-Sodium ions also interact most likely with phosphate oxygens of POPC, 
-but in contrast to calcium, the interactions with only carbonyls are also significant.
-The physiological relevance of the sodium binding details is, however, uncertain
-due to its very weak binding affinity.
-
-\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
+Sodium ions also interact primarily with phosphate oxygens of the POPC, 
+but in contrast to calcium, the interactions purely with carbonyls are also significant.
+The physiological relevance of these details of sodium interactions is, however, unclear
+due to its overal very weak binding affinity.
+
+\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to the POPC membrane}
 
@@
 
-Simple binding models have been previously used to interpret
+Simple binding models have been used previously to interpret
 the same experimental data \cite{altenbach84,macdonald87} 
-which are used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}.
-The experimental results from the lipid headgroup order parameters and from the
+that are used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}.
+Results from measurements of the lipid headgroup order parameters and from
 atomic absorption spectroscopy were best
@@
 stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-However, also a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
-provided a good fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
-The results based on these binding models and the measured headgroup order parameter changes
+However, a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
+also provided a good fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
+Note that the results based on these binding models and the measured headgroup order parameter changes
 are also inluded in the tabulated experimental binding constants~\cite{marsh13}.
 
-Since the same experimental data is reproduced by the ECC-lipids model,
-it can be used to give a more versatile interepration of the binding stoichiometry than
+Since the same experimental data are reproduced by the present ECC-lipids model,
+we can use the model to provide a more detailed interepration of the binding stoichiometry than
 the previously used simple binding models~\cite{altenbach84,macdonald87}. To directly evaluate the
@@
 complexes with various stoichiometries by counting the individual lipid
-molecules within the distance $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion
-as already used in previous section~\ref{sec:affinity}.
-The results from POPC simulation with 285~mM concetration of CaCl$_2$ in
-Fig.~\ref{fig:cacl_complexes} show the largest propensity (41\%) for the
+molecules within the distance $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion,
+as already done in the previous section~\ref{sec:affinity}.
+The results from POPC simulation with a 285~mM concetration of CaCl$_2$ in
+Fig.~\ref{fig:cacl_complexes} demonstrate the largest propensity (41\%) for the
 ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
 the probabilities of complexes with the stoichimetries of 1:1~(25\%)~and
-1:3~(34\%)~are only slightly lower
-suggesting a more complex binding model than the simple ternary complex model.
+1:3~(34\%)~are only slightly lower. This
+suggests a more complex binding model than the previous simple ternary complex model.
 %The observed propensities of
@@
 %$34\%$ for three lipids 
-However, the simulation data can also be viewed from such an angle that
-one calcium binds roughly to two lipids on average,
-because the probabilities of the complexes with 1 or 3 lipids are almost equal.  
-This probably explains the success of the ternary complex model fitting well the
-experimental data. 
-Complexes with more than three lipids per one calcium ion were not observed in 
-simulations with the ECC-lipids model.
+With a broad brushstroke, the simulation data can, however, be viewed such that
+one calcium binds to two lipids on average,
+because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other 
+(and complexes with more than three lipids per one calcium ion were not observed).  
+This probably explains why the simple the ternary complex model fitts the
+experimental data, as well as the ECC-lipids simulation results, rather well
+(see Fig.~S\ref{fig:cacl-bind} in SI and its caption for details).
+
 \todo{Samuli: I think that we either have to do the analysis in SI properly,
@@
 The fact that we use a different (and more accurate) estimate of $C_I$ and other quantities does not hamper this statement.}
-Ternary complex model also provides a good fit 
-to our simulations with ECC-lipids 
-(see Fig.~S\ref{fig:cacl-bind} in SI and its caption for details). 
+
 \begin{figure}[tb!]
@@
 
-Equilibration of \ce{Ca^{2+}} binding to POPC bilayer takes hundreds of
+Equilibration of \ce{Ca^{2+}} ions at the POPC bilayer takes hundreds of
 nanoseconds in MD simulations with current state of the art force fields,
@@
 simulations at a temporal scale of several microseconds are required
-for studies of cation binding in equilibrium. 
+to reach the binding/unbinding equilibrium. 
 This is in line with the density profiles of \ce{Ca^{2+}} 
@@
 which all show a region of a very small calcium density 
-around the level of cholines in POPC (i.e. around $2.5\,\mathrm{nm}$)
-with the exception of ECC-lipids, 
-which exhibits only a relatively shallow local
+around the level of cholines in POPC (i.e. around $2.5\,\mathrm{nm}$).
+In contrast, simulations with  ECC-lipids 
+exhibits only a relatively shallow local
 minimum at that level.
-This suggests that such a smaller barrier for the exchange of calcium ions 
-shall lead to their faster exchange with the solvent.
-To quantify the exchange of ions between the membrane and solvent in simulations, we calculated 
+This suggests that the smaller barrier for the exchange of calcium ions 
+shall lead to their faster exchange with the solvent within ECC-lipids.
+To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated 
 residence times of ions bound to the membrane.
-We again use the definition for ions binding to the membrane as in the above section~\ref{sec:affinity},
-i.e. ion is bound if it is within 0.3~nm from any oxygen atom
+We use the same definition for ions bound to the membrane as in the above section~\ref{sec:affinity},
+i.e., an ion is considered as bound if it is within 0.3~nm from any oxygen atom
 belonging to a lipid molecule in a bilayer.
@@
 The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer 
-at $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
+at a $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
 from simulations with ECC-lipids 
-and CHARMM36 (taken from previous study \cite{javanainen17}, 
+compared to CHARMM36 (taken from previous study \cite{javanainen17}, 
 simulation data available at Ref.~\citenum{zenodo.259376})
@@
 In the CHARMM36 simulation, approximately 20\% of the \ce{Ca^{2+}} residence times are longer than
-the length of the trajectory ($800\,\mathrm{ns}$).
-With such a simulation, which is hence relatively short for the employed model,
-we can merely estimate an upper bound that 
-only less than 60\% of the bound residence time 
-is formed by cations interacting with the phospholipid membrane 
+the whole length of the trajectory ($800\,\mathrm{ns}$).
+Within such a relatively short simulation, 
+we can only estimate that 
+less than 60\% of the bound residence times 
+is due to cations interacting with the phospholipid membrane 
 for less then half of the simulation length~(400~ns).
 Even longer residence times are observed
-in the other simulations with CHARMM36 and Slipids models reported in the previous work~\cite{javanainen17}.
-
-On the other hand, at least an order of magnitude faster exchange is observed in the
+in  simulations with CHARMM36 and Slipids models reported in previous works~\cite{javanainen17}.
+
+In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the
 ECC-lipids model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are 
-shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
-and the longest observed residence time in the ECC-lipids simulation is $141\,\mathrm{ns}$, 
-which is well below the total length of the simulation used for analysis, 200~ns.
-The exchange of \ce{Na+} ions to a POPC membrane is
-another order of magnitude faster, giving 90\%
-of the residence times smaller than~$1\,\mathrm{ns}$,
+shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$.
+The longest observed residence time is now $141\,\mathrm{ns}$, 
+which is well below the total length of the simulation used for analysis, i.e., 200~ns.
+The exchange of \ce{Na+} ions at the POPC membrane is
+another order of magnitude faster, yielding 90\%
+of the residence times smaller than~$1\,\mathrm{ns}$, with
 the longest residence time being~$6\,\mathrm{ns}$. 
-Note that the results from both analyzed models are in line with the experimental estimation that 
-the residence time of \ce{Ca^{2+}} at the individual PC headgroup 
+Note that the ECC-lipids calcium results are in line with the experimental estimate that 
+the residence time of \ce{Ca^{2+}} at each individual PC headgroup 
 is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
@@
 In conclusion, the results from the ECC-lipids model suggest that the
-exchange of calcium between membrane and solvent occurs within $\sim 100\,$ns timeframe,
-which is much faster than previously reported~\cite{javanainen17}.
-Sodium cations exhibit even a faster exchange.
-This suggests that simulations with the length of a few hundreds of
+exchange of calcium between the POPC bilayer and the solvent occurs within $\sim 100\,$ns timeframe,
+which is significantly faster than previously reported~\cite{javanainen17}.
+Sodium cations exhibit an even a faster exchange.
+This all suggests that simulations with a length of a several hundreds of
 nanoseconds are sufficient to simulate ion binding
 to phospholipid bilayers in equilibrium, when realistic force fields are used.
-This has not been the case with the available lipid force fields,
-which overestimate the binding strength of the cations \cite{javanainen17,catte16}.
+This has not been the case with previous lipid force fields,
+which overestimate the binding strength of the sodium and calcium cations \cite{javanainen17,catte16}.
 %Such a finding changes the point of view on binding of calcium to PC membranes from
@@
 
-We used the electrometer concept to show that the binding of \ce{Na^+} and \ce{Ca^2+} ions
-to a POPC lipid bilayer can be accurately described with a classical MD simulation 
-force field, where the electronic polarization is implicitly included using 
-the electronic continuum correction (ECC)~\cite{leontyev11}.
-The proposed ECC-lipids model is a significant improvement over 
-other available lipid models, which all overestimate specific cation binding affinities~\cite{catte16}.  
-While the structural details of a POPC lipid bilayer simulated with the ECC-lipids
+In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions
+to a POPC lipid bilayer can be accurately described within a classical MD simulation 
+force field, provided that electronic polarization is implicitly included via 
+the electronic continuum correction~\cite{leontyev11}.
+While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-lipids
 model agree with experiments with an accuracy comparable to the other state of the art lipid models,
-it also reproduces the experimental lipid head group order parameter responses to
+the model also reproduces the experimental lipid head group order parameter responses to
 a cationic surfactant, NaCl and CaCl$_2$ concentrations. 
+It thus represents a significant improvement over 
+other available lipid models, which overestimate cation binding affinities~\cite{catte16}.  
+The ECC-lipids model is build upon
+the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8
+and reducing the Van der Waals radii by a factor of 0.89 for the headgroup, glycerol backbone and carbonyl atoms.
 
 The good agreement with experiments enables us to 
-interpret NMR experiments in atomistic details using MD simulations.
+interpret NMR experiments with atomistic details using MD simulations with the ECC-lipid model.
 In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
 Ca$^{2+}$ ions interact mainly with phosphate oxygens.
-However, the stoichiometry of calcium binding is significantly more complicated 
-than in the previous interpretation of the NMR data based on 
-the ternary complex model, where one calcium binds to two POPC molecules~\cite{altenbach84}.
-The complexes of one calcium bound to two lipids are the most probable also in the
-ECC-lipids model, but the complexes of one or three lipids per one calcium
-were observed to be almost equally likely. While the success of the ternary
-compex model is understandable based on the simulation results,
-a simple binding model cannot detect 
-the complex nature of calcium binding to phospholipid bilayers
-oberved in the simulation.
-
-The accurate description of cation binding to POPC bilayer paves the way for
-simulations of complex biochemical systems with realistically described
-electrostatic interactions in the vicinity of cellular membranes.
-The ECC-lipids model is build by scaling the partial charges
-and the Van der Waals radii of the headgroup, glycerol backbone and carbonyl atoms of
-the Lipid14 POPC model~\cite{dickson14} with the constant factor
-of 0.8 and 0.89, respectively. While the Lipid14 model
-is compatible with the AMBER force field family, the compatibility of
-the ECC-lipids model may be compromised due to the changes in intermolecular
-interactions of the scaled atoms. On the other hand, a fully consistent
-ECC-force field should include the correction also in other than lipid
-molecules, including water. The work toward this direction and the extension
-to other lipid molecules and force fields is left for future studies.
+However, the stoichiometry of calcium binding is significantly more complicated than
+the simple ternary complex model, used to interpret NMR data, where one calcium binds to two POPC molecules~\cite{altenbach84}.
+While complexes with one calcium ion bound to two lipids are the most probable also in the
+ECC-lipids model, complexes of one or three lipids per one calcium
+were observed to relatively abundant and also almost equally likely. While the success of the simple ternary
+compex model in fitting NMR data is understandable based on the simulation results,
+it cannot capture the complex nature of calcium binding to phospholipid bilayers
+oberved in the present simulation.
+
+Accurate description of cation binding to POPC bilayer paves the way for
+simulations of complex biochemical systems at cellular membranes with realistically described
+electrostatic interactions. To this end, the compatibility of
+the ECC-lipids model with existing models for proteins and other biological molecules should to be verified and, if necessary, further adjustments along the ECC concept in force fields of molecules interacting with the lipids need to be done.
 

### 2017-11-06 Manuscript/manuscript.tex

@@
 This is in practice done by changing the parameters $\sigma$ in the Lennard-Jones potential 
-in a similar way as was done previously for the ECC ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
+in a similar way as was done previously for the ECC-ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
 After reducing the $\sigma$ parameters of affected atoms by a factor of $f_\sigma = 0.89$, 
@@
 %the hydrophobic region in the middle of the membrane where no ions or waters are present.
-Such a setup and the above definition of bulk ion concentration provide a simple relation for the normalized surface excess $\Gamma_i^w$ for simulations of lipid bilayers, 
+Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_i^w$ for simulations of lipid bilayers, 
 \begin{equation}\label{surfexcess}
@@
 Scaled charges and van der Waals radii for atoms forming the POPC lipid were derived here, starting from the Lipid14 force field.
-For comparison, simulation with unscaled charges were perfomed using the Lipid14 model used the TIP3p water model \cite{jorgensen83} and the ion model by
+For comparison, simulations with unscaled charges were perfomed using the Lipid14 model, the TIP3p model for water \cite{jorgensen83} and the ion model by
 Dang et al.~\cite{smith94,chang1999,dang2006}.
@@
 and the same atom types with reduced $\sigma$ parameters from ECC-lipids were used. 
+\todo{Joe: I think that the followin is already said in the beginning of this section -- Lipid14 simulaitons.}
 For comparison, simulations with unscaled charges were also performed.
@@
 The physiological relevance of these details of sodium interactions is, however, unclear
-due to its overal very weak binding affinity.
+due to its overall very weak binding affinity.
 
@@
 (and complexes with more than three lipids per one calcium ion were not observed).  
-This probably explains why the simple the ternary complex model fitts the
+This probably explains why the simple the ternary complex model fits the
 experimental data, as well as the ECC-lipids simulation results, rather well
@@
 the longest residence time being~$6\,\mathrm{ns}$. 
+\todo{Joe: the following is a bit more subtle argument. 
+I deem it not necessary to include it in the paper, however, 
+the experimental estimate of less tan 10us holds likely also for Charmm36/ECC-ions, as stated before.}
 Note that the ECC-lipids calcium results are in line with the experimental estimate that 
@@
 electrostatic interactions. To this end, the compatibility of
-the ECC-lipids model with existing models for proteins and other biological molecules should to be verified and, if necessary, further adjustments along the ECC concept in force fields of molecules interacting with the lipids need to be done.
+the ECC-lipids model with existing models for proteins and other biological molecules should to be verified and, if necessary, 
+further adjustments along the ECC concept in force fields of molecules interacting with the lipids need to be done.
 

### 2017-11-06 Manuscript/manuscript.tex
Table with gammas and ratios of binding to phos/carbonyl filled with values now.

@@
 (Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}). 
-\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
 The results are in line with the previous work \cite{catte16},
@@
   from various simulation models. 
-  The CaCl$_2$ concentration of the ions in water before adsorption to the phospholipid bilayer
-  is 350mM in the analyzed systems. \label{tab:binding}}
-  \todo{Add columns with concentrations of calcium -- bulk/molar fraction.}
-  \todo{Changing symbol $P$ to $R$ or $r$ might reflect better that it is rather a ratio, and not that much a probability.}
-  \begin{tabular}{l|c c c}
-    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+  $C_w$ is the concentration of CaCl$_2$ in water before adsorbing to the phospholipid bilayer. 
+  \label{tab:binding}}
+  \begin{tabular}{l|c | c c | c c}
+    model                 & $\Gamma_{Ca}^w \,/\, \mathrm{nm}^{-2}$ & $C_{w}$ & $C_b\,/\,\mathrm{mM}$   & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
     \hline
-    ECC-lipids            &  $0.06 \pm 0.01 $            &  99\%   &    25\%    \\
-    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
-    Lipid14/Dang          &  $0.23 \pm 0.03 $            &  -        &    -       \\
-    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $            &  -        &    -       \\
+    ECC-lipids            &  $0.06 \pm 0.01 $ &  350  &  $280\pm 10 $  &  99\%  &    25\%    \\
+    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $ &  350  &  $210\pm 10 $  & 100\%  &    37\%     \\
+    Lipid14/Dang          &  $0.23 \pm 0.03 $ &  350  &  $160\pm 10 $  & 100\%  &    14\%    \\
+    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $ &  350  &  $120\pm 100$  & 100\%  &    23\%    \\
+%Average no. of bound cations:
+%L14+Aqvist: all:22.7, phos:22.7, carb:8.5
+%L14+Dang:   all:29.2, phos:29.2, carb:4.0
+%L14+ECC-ion:all:37.7, phos:37.7, carb:8.5
   \end{tabular}

### 2017-11-07 Manuscript/manuscript.tex
Removing TODOs: acknowledgments, SI figures, additional references, temperatures in captions etc.

@@
 Technically, this is similar to the phenomenological charge-scaling applied in 
-earlier studies of aqueous solutions or ionic liquids \cite{jonsson86,egberts94}.
-\todo{We should also cite papers where empirical scaling was used ionic liquids DOI: 10.1002/anie.201308760}
+earlier studies of aqueous solutions or ionic liquids \cite{jonsson86,egberts94,beichel14}.
 However, the present concept of ECC is
@@
 By construction, partial charges currently implemented in force fields may already 
-include  to some extend the electronic polarizability effects of the solvent, i.g.,
-the RESP charges may be subsequently modified to fit certain experimental observables \todo{This needs a citation - Joe please pick 1-2 papers}. 
+include  to some extent the electronic polarizability effects of the solvent, i.g.,
+the RESP charges may be modified during fitting or subsequently after it 
+to improve certain experimental observables 
+\cite{RESP_paper,Singh1984,jorgensen96,ipolq2013,benavides17}. 
 Thus, a consistent application of the ECC scaling factor, $f_q$, to the partial charges in molecules as
@@
     The changes of headgroup order parameters and P-N vector orientation as a function of
-    cationic surfactant (dihexadecyldimethylammonium) in POPC bilayer from simulations
-    and experiments \cite{scherer89}.
+    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
+    from simulations and experiments \cite{scherer89}.
   }
@@
   \caption{\label{fig:delta_ordPar_CaCl}
-    Changes of the head group order parameters of POPC bilayer as functions of the CaCl$_2$ concentration
+    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
+    as a function of the CaCl$_2$ concentration
     are shown from simulations with different force fields together with experimental data 
-    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
+    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
     Ion concentrations in bulk water are shown in x-axis. 
-    Values from simulations are calculated from the of cation number density $C_{np}$
-    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
+    Bulk concentrations from simulations are calculated 
+    from the farthest point from the lipid bilayer in the aqueous phase
+    with an error estimate of 10\,mM.
+    The error estimate for the simulation with Lipid14 and ECC-ions is an order of magnitude larger, 
+    because the simulation cell was too small (shown if Fig.~\ref{fig:cacl-dens}). 
     Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
@@
   }
-  \todo{Add temperatures for POPC and DPPC experiments} 
   %\todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
@@
   \caption{\label{fig:delta_ordPar_NaCl}
-    The changes of the head group order parameters of POPC bilayer as a function of NaCl concentrations
-    from simulations with different force fields together with experimental data for DPPC~\cite{akutsu81} and POPC~\cite{altenbach84}.
+    The changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
+    from simulations with different force fields together with 
+    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
     Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from
+    Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
     Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
@@
   %\todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
-  \todo{Add temperatures to the experimental data.} 
 \end{figure}
@@
 and 600 from the Academy of Finland via the FiDiPro award.
-\todo{Add acknowledgments of Jiri Kolafa.}
-\todo{Joe: Thank Metacentrum.}
+J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.
+Computational resources were supplied by the Ministry of Education, Youth and Sports of the Czech Republic under the Projects CESNET (Project No.
+LM2015042) and CERIT-Scientific Cloud (Project No.
+LM2015085) provided within the program Projects of Large Research, Development and Innovations Infrastructures.
 \end{acknowledgments}
@@
 
+\begin{figure}[tbp!]
+  \centering
+  \includegraphics[width=16.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89_GMX-oMM_compar.pdf}
+  \caption{\label{fig:ordPars_actual_GMX_oMM_compar}
+    Order parameters of POPC head group, glycerol backbone and acyl chains 
+    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
+    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
+    compared with the experimental values from \cite{ferreira13}.
+    The size of the markers for the head group order parameters correspond to
+    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
+    while the error estimate for simulations is $\pm 0.005$.
+    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
+  }
+\end{figure}
+
+\begin{figure}[tbp!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_surf_GMX-oMM_compar.pdf}
+  \caption{\label{fig:ordPars_surf_GMX_oMM_compar}
+    The changes of headgroup order parameters and P-N vector orientation as a function of
+    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
+    from simulations with the ECC-lipids models
+    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
+    compared with the experimental values from \cite{scherer89}.
+    The size of the markers for the head group order parameters correspond to
+    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
+    while the error estimate for simulations is $\pm 0.005$.
+  }
+\end{figure}
+
+\begin{figure}[tbp!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_GMX-oMM_compar.pdf}
+  \caption{\label{fig:ordPars_cacl_GMX_oMM_compar}
+    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
+    as a function of the CaCl$_2$ concentration
+    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
+    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
+    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
+    Ion concentrations in bulk water are shown in x-axis. 
+    Bulk concentrations from simulations are calculated 
+    from the farthest point from the lipid bilayer in the aqueous phase
+    with an error estimate of 10\,mM.
+  }
+\end{figure}
+
 % Create the reference section using BibTe

### 2017-11-08 Manuscript/manuscript.tex

@@
 
-\todo{Samuli: I think that we either have to do the analysis in SI properly,
-  i.e., using the real $C_I$ from simulations or remove the sentence below and
-the related concent from SI.\\
-Joe: I think it shows that ternary complex makes a good fit to our simulation data. 
-The fact that we use a different (and more accurate) estimate of $C_I$ and other quantities does not hamper this statement.}
-
 \begin{figure}[tb!]
@@
 
-\todo{This should be either done properly with the correct $C_I$ or removed. Removing
-might be a better idea at this point.}
 It was found in the original work~\cite{altenbach84} that 
@@
 remains constant throughout the extrapolation range. 
-The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
+The concentration $C_I$ is determined from the surface potential using the Boltzmann equation.
 However, Boltzmann theory yields inaccurate results
 for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
-An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions.
-Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
-our simulations likely lays in the fact that 
-the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
-gradually do not hold for higher concentrations of \ce{Ca^{2+}}.
+An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions;
+$C_b$ can be calculated from simulation using the robust definition in section~\ref{sec:affinity},
+and $C_I$ can be obtained from the density profiles in Fig.~\ref{fig:cacl-density} 
+as the minimum concentration at the shallow depression separating the bulk water phase 
+and the peak of adsorbed cations around 2.6\,nm from the membrene centre. 
+The small discrepancy between the results in the experiment~\cite{altenbach84} and 
+our simulations are likely caused by the difference in the evaluation of the concentrations.
 
@@
     and it also provides a good fit to our simulation data. 
+    This supports our statement that the success of the ternary complex model in the experiments
+    can be understood with the atomistic detail of our simulations as an average 
+    stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
     Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
-    and, hence, the observed slope of the linear relationship is different.
+    and, hence, the observed slope of the linear relationship is slightly different.
     }

### 2017-11-08 Manuscript/manuscript.tex
Rewritten Abstract

--- 2017-11-06 Manuscript/manuscript.tex
@@
   % insert abstract here
-  The binding affinities and stoichimetries of Na$^+$ and Ca$^{2+}$ ions to phospholipid bilayers have not been
-  consistently determined with experimental and computational methods,
-  despite their significance and biological relevance in cellular membranes. 
-  The details concerning ion binding could be resolved with classical molecular dynamics (MD) simulations,
-  however, the accuracy of the available lipid models has not been sufficient.
-  In this work we show that the binding details of Na$^+$ and Ca$^{2+}$ ions to the
-  1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
-  can be accurately described with a MD simulation model implicitly including electronic polarization.
-  This is demonstrated by applying the electronic continuum correction (ECC) to a
-  state of the art lipid model for MD simulations of the POPC
-  lipid bilayer. The newly introduced ECC-lipids model reproduces the experimentally measured structural parameters for the ion-free membrane,
-  the response of lipid headgroup to a bound positive charge, and the binding affinities of
-  Na$^+$ and Ca$^{2+}$ ions. The observed negligible binding of Na$^+$ ions to POPC bilayer
-  and stronger interactions of Ca$^{2+}$ primarily with phoshpate oxygens give
-  support to the early intepretations of the experimental spectroscopic data.
-  At the same time, Ca$^{2+}$ ions form complexes with 1-3 lipid molecules with almost
-  equal probabilities, suggesting a more complicated binding stoichimetry than 
-  the ternary complex binding model used to interpret the NMR data.
-  The results of this work pave the way to accurate MD simulations of complex biochemical systems
-  with realistic electrostatic interactions in the vicinity of cellular membranes.
+Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
+bilayers are of paramount significance in the properties and functionality of cellular
+membranes. For example, possible accumulation of ions at the membrane interface determines
+the interaction affinity with charged species. Current estimates of binding affinities and
+stoichiometries of cations are controversial due to limitations in the experimental and
+theoretical methods available. Experimentally one can assess these parameters by titrating
+several membrane properties as a function of the ions concentration. However, the
+interpretation of the obtained results requires theoretical models. Considering that the
+real behavior is unknown these models are at best idealizations. Classical molecular
+dynamics (MD) simulations provide details of the ions binding process with atomistic
+resolution. Therefore, MD a priory offers all necessary information to interpret
+experimental data without the need to resort to simplified models. However, the accuracy
+of the available lipid models when interacting with ions is currently limited. In this
+work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
+1-Palmitoyl-2-oleoyl-phosphatidylcholine POPC) bilayer by implicitly including electronic
+polarization. We do this by applying a mean field correction to the partial charges known
+as the electronic continuum correction (ECC) to a selected state of the art POPC lipid
+model for MD simulations. Our improved ECC-POPC model reproduces not only the
+experimentally measured structural parameters for the ion-free membrane, but also the
+response of lipid headgroup to a bound positive charge, and the binding affinities of 
+Na$^+$ and Ca$^{2+}$ ions. We also observe negligible binding of Na$^+$ ions to POPC
+bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens in our new
+model is in agreement with available experimental spectroscopic data. This new model
+results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
+equal probabilities which suggests that previous models used to interpret NMR data are too
+simple. The results of this work pave the way to quantitative MD simulations of complex
+biochemical systems with realistic electrostatic interactions in the vicinity of cellular
+membranes.
 \end{abstract}
@@
 (Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}). 
+\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
 The results are in line with the previous work \cite{catte16},
@@
   from various simulation models. 
-  $C_w$ is the concentration of CaCl$_2$ in water before adsorbing to the phospholipid bilayer. 
-  \label{tab:binding}}
-  \begin{tabular}{l|c | c c | c c}
-    model                 & $\Gamma_{Ca}^w \,/\, \mathrm{nm}^{-2}$ & $C_{w}$ & $C_b\,/\,\mathrm{mM}$   & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+  The CaCl$_2$ concentration of the ions in water before adsorption to the phospholipid bilayer
+  is 350mM in the analyzed systems. \label{tab:binding}}
+  \todo{Add columns with concentrations of calcium -- bulk/molar fraction.}
+  \todo{Changing symbol $P$ to $R$ or $r$ might reflect better that it is rather a ratio, and not that much a probability.}
+  \begin{tabular}{l|c c c}
+    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
     \hline
-    ECC-lipids            &  $0.06 \pm 0.01 $ &  350  &  $280\pm 10 $  &  99\%  &    25\%    \\
-    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $ &  350  &  $210\pm 10 $  & 100\%  &    37\%     \\
-    Lipid14/Dang          &  $0.23 \pm 0.03 $ &  350  &  $160\pm 10 $  & 100\%  &    14\%    \\
-    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $ &  350  &  $120\pm 100$  & 100\%  &    23\%    \\
-%Average no. of bound cations:
-%L14+Aqvist: all:22.7, phos:22.7, carb:8.5
-%L14+Dang:   all:29.2, phos:29.2, carb:4.0
-%L14+ECC-ion:all:37.7, phos:37.7, carb:8.5
+    ECC-lipids            &  $0.06 \pm 0.01 $            &  99\%   &    25\%    \\
+    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
+    Lipid14/Dang          &  $0.23 \pm 0.03 $            &  -        &    -       \\
+    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $            &  -        &    -       \\
   \end{tabular}

### 2017-11-08 Manuscript/manuscript.tex
Small edits to the abstract

@@
 the interaction affinity with charged species. Current estimates of binding affinities and
-stoichiometries of cations are controversial due to limitations in the experimental and
-theoretical methods available. Experimentally one can assess these parameters by titrating
-several membrane properties as a function of the ions concentration. However, the
-interpretation of the obtained results requires theoretical models. Considering that the
-real behavior is unknown these models are at best idealizations. Classical molecular
-dynamics (MD) simulations provide details of the ions binding process with atomistic
-resolution. Therefore, MD a priory offers all necessary information to interpret
+stoichiometries of cations are, however, controversial due to limitations in the available experimental and
+theoretical methods. Experimentally one can assess these parameters by titrating
+several membrane properties as a function of the ion concentration. However, the
+interpretation of the experiments relies on theoretical models, which give
+the best idealization of the ion binding as the direct experiments are not available. 
+Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
+resolution, therefore offering all the necessary information to interpret
 experimental data without the need to resort to simplified models. However, the accuracy
-of the available lipid models when interacting with ions is currently limited. In this
-work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
-1-Palmitoyl-2-oleoyl-phosphatidylcholine POPC) bilayer by implicitly including electronic
-polarization. We do this by applying a mean field correction to the partial charges known
-as the electronic continuum correction (ECC) to a selected state of the art POPC lipid
-model for MD simulations. Our improved ECC-POPC model reproduces not only the
+of the available lipid models when interacting with ions is not sufficient for
+such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
+1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
+polarization as a mean field correction, known as the electronic continuum correction (ECC).
+This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
+Our improved ECC-POPC model reproduces not only the
 experimentally measured structural parameters for the ion-free membrane, but also the
 response of lipid headgroup to a bound positive charge, and the binding affinities of 
-Na$^+$ and Ca$^{2+}$ ions. We also observe negligible binding of Na$^+$ ions to POPC
-bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens in our new
-model is in agreement with available experimental spectroscopic data. This new model
-results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
-equal probabilities which suggests that previous models used to interpret NMR data are too
-simple. The results of this work pave the way to quantitative MD simulations of complex
+Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
+bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens 
+, which are in agreement with the previous interpretations of the experimental spectroscopic data.
+On the other hand, the new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
+equal probabilities, suggesting more complex binding stoichiometry than simple models
+previously used to interpret the NMR data.
+The results of this work pave the way to quantitative MD simulations of complex
 biochemical systems with realistic electrostatic interactions in the vicinity of cellular

### 2017-11-09 Manuscript/manuscript.tex
Introduction changes

--- 2017-11-08 Manuscript/manuscript.tex
@@
 \section{Introduction}
-Cation interactions with cellular membranes play a key role in important biological processes,
-such as signal propagation in neurons and vesicle fusion.
-Since direct measurements of ion-membrane interactions in biological systems
-are difficult, lipid bilayers are often used as models systems for cellular
-membranes. In particular, the zwitterionic phosphocholine (PC) lipid bilayers are
-used as entry level models, helping to understand the role of ions in complex biological
-systems~\cite{scherer87,seelig90,cevc90}.
-
-Interactions of biologically relevant cations, especially Na$^+$ and
-Ca$^{2+}$, with PC bilayers have been widely studied in
-experiments~\cite{akutsu81,altenbach84,seelig90,cevc90,tocanne90,binder02,pabst07,uhrikova08}
-and classical MD simulations~\cite{bockmann03,bockmann04,Berkowitz12,melcrova16,javanainen17}.
+Ions interactions with cellular membranes play a key role in critical biological processes~\cite{REF_REVIEW}.
+Ions, especially cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{REF}.
+Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface.
+For example, Ca$^{2+}$  is crucial in neural signal propagation.
+It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}.
+Ca$^{2+}$ also participates in the T-cell receptor activation.
+It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013}.
+A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
+Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions.
+Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.
+
+Direct measurements of ion-membrane interactions in real biological systems are difficult.
+Simplified lipid bilayers are often used instead as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}.
+For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
 The details of ion binding are, however, not fully consistent in the literature.
-%not agreed in the literature \cite{cevc90,tocanne90,seelig90,binder02,catte16,??}.
-%General conclusion from experimental
-%studies has been that multivalent ions and lithium have weak specific binding in
-%phopholipid bilayers, while other monovalent ions do not essentially
-%bind 
-%The presence of anoinic lipids, like PS or PG,
-%increase the concentration close to the bilayer and thus the amount of bound
-%ions, but do not affect the specific binding constant \cite{seelig90}.
-Interpretations of non-invansive spectroscopic methods, like nuclear magnetic resonance (NMR), as well as scattering
-and infrared spectroscopies, suggest that Na$^+$ ions exhibit only
-negligible binding to PC lipid bilayers wat submolar concentrations, while Ca$^{2+}$ is observed to specifically
-bind to phosphate groups of two lipid molecules~\cite{hauser76,hauser78,herbette84,akutsu81,altenbach84,binder02,pabst07,uhrikova08}.
-Atomistic resolution molecular dynamics (MD) simulation models, however, typically predict a
-stronger binding for the cations than NMR experiments \cite{catte16}.
-On the other hand, several experiments have also been interpreted to support the predictions from
-earlier MD simulations~\cite{bockmann03,vacha09a}. 
-Furthermore, interactions of calcium ions with up to 3-4 lipids,
-including not only interactions with phosphates but also with carbonyl
-oxygens, have been reported from simulations~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
-
-Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16}
-made an attempt to resolve these apparent controversies. A direct comparison of
-ion binding affinities to PC bilayers between simulations and
-experiments was done using the electrometer concept, which is based on experimental NMR data
-for the lipid headgroup order parameters~\cite{seelig87}.
-Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded
-that the accuracy of the current state of the art lipid models for MD
-simulations is not sufficient for a detailed interpretation of the
-interactions of cations with PC lipid bilayers~\cite{catte16}.
+Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mainly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
+On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
+On the other hand, {\color{red} in general} \todo{Not CHARMM36 or amber14} atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
+For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
+{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}.
+\todo{remove?
+Why we need this?}}
+
+Recent studies within the NMRlipids project \url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
+A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept.
+Namely, the changes in NMR order parameters of the headgroup upon ions additions~\cite{seelig87} is directly compared to the MD simulations results.
+Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
 
@@
 
-In this work, we show that the cation binding behavior
-in MD simulations of the 1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer can be significantly 
-improved via an implicit inclusion of the electronic polarizability
-in the polar region of phospholipids,
-which we apply using to the 
-electronic continuum correction (ECC)~\cite{leontyev11}.
-Such an approach has been previously shown to improve the behaviour of
-of ions in bulk water simulations ~\cite{jungwirth17-new-paper-to-be-published,Pluharova2014,kohagen14,kohagen16}. 
-We employ here the the Lipid14 model \cite{dickson14} as a starting point,
-because it provided one of the best available descriptions of cation binding \cite{catte16}.
+In this work, we improve the cation binding of a POPC bilayer in MD simulations via the implicit inclusion of the electronic polarizability in the polar region of phospholipids.
+For this, we apply a simple mean-field electronic continuum correction (ECC)~\cite{leontyev11} which implies scaling the partial charges of the atoms to account for the missing electronic polarizability in MD simulations~\cite{leontyev11}.
+Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014, kohagen14, kohagen16}.
+As a starting template for our POPC model, we will use the Lipid14~\cite{dickson14}.
+This force field provides one of the best available descriptions of cation binding \cite{catte16}.
 The newly developed ECC-lipids model reproduces the experimentally
-measurable structural parameters of an ion-free POPC lipid bilayer with the 
-accuracy comparable to the other state of the art lipid models, while surpassing
-them significantly in the description of the membrane binding affinities to cations and induced
-structural effects of sodium and calcium ions.
+measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while improving the membrane binding affinities significantly to sodium and calcium cations.
 

### 2017-11-09 Manuscript/manuscript.tex
Small edits to Introduction

@@
 \section{Introduction}
-Ions interactions with cellular membranes play a key role in critical biological processes~\cite{REF_REVIEW}.
-Ions, especially cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{REF}.
+Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}.
+Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
 Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface.
@@
 On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-On the other hand, {\color{red} in general} \todo{Not CHARMM36 or amber14} atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
+On the other hand, in general atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
 For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
-{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}.
-\todo{remove?
-Why we need this?}}
-
-Recent studies within the NMRlipids project \url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
-A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept.
-Namely, the changes in NMR order parameters of the headgroup upon ions additions~\cite{seelig87} is directly compared to the MD simulations results.
+%{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}
+
+Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
+A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept~\cite{seelig87}.
+Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results.
 Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.

### 2017-11-09 Manuscript/manuscript.tex
Updated section: Electronic continuum correction for lipid bilayers

--- 2017-11-09 Manuscript/manuscript.tex
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-The lack of electronic polarizability in standard MD simulation
-force fields has been considered a serious issue since the early days of
-lipid bilayer simulations.
-In this work, we circumvent the rather demanding explicit inclusion
-of electronic polarization effects \cite{lucas12,chowdhary13} by
-including the electronic part of polarizability in lipid
-bilayer simulations implicitly via 
-the electronic continuum correction (ECC)~\cite{leontyev11}. 
-Technically, this is similar to the phenomenological charge-scaling applied in 
-earlier studies of aqueous solutions or ionic liquids \cite{jonsson86,egberts94,beichel14}.
-However, the present concept of ECC is
-physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
-
-According to ECC, the electronic
-polarizability can be implicitly included in classical MD simulations by
-placing all particles into a homogeneous dielectric continuum 
-with a dielectric constant $\epsilon _{el}$, 
-which is the electronic part of the dielectric constant of 
-the medium \cite{leontyev11}. 
-Following the Coulomb's law, such a dielectric continuum can be 
-directly included in a standard MD simulation by
-scaling of charges,
+The lack of electronic polarizability in standard MD simulation force fields has been considered a serious issue since the early days of lipid bilayer simulations.
+In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by including the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
+Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of aqueous solutions or ionic liquids \cite{jonsson86, egberts94, beichel14}.
+However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
+
+According to ECC mean-field correction, for atomic ions, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
+Following Coulomb's law,  we can directly incorporate the results by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$.
+
 \begin{equation}
@@
 \end{equation}
-with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$ 
-. 
-Given the  high frequency dielectric constant as measured in water (corresponding to the square of the refraction index) of  $\epsilon _{el} = 1.78$,  
-the scaling factor for aqueous ions is $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
-successfully used to improve the performance of force field simulations of ions in aqueous solution \cite{kohagen14,kohagen16,??},
-which then agree quantitatively with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
-
-While the scaling factor of $f_q = 0.75$ for ions in water improves their description and is physically well justified within the ECC theory \cite{Leontyev2011},
-it is not a priori  clear whether the same factor should be used for partial charges in molecules,
-e.g., lipids in our case. Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
-Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}.
-Currently, the most commonly employed scheme is the restrained electrostatic potential (RESP) method~\cite{RESP_paper,Singh1984}.
-By construction, partial charges currently implemented in force fields may already 
-include  to some extent the electronic polarizability effects of the solvent, i.g.,
-the RESP charges may be modified during fitting or subsequently after it 
-to improve certain experimental observables 
-\cite{RESP_paper,Singh1984,jorgensen96,ipolq2013,benavides17}. 
-Thus, a consistent application of the ECC scaling factor, $f_q$, to the partial charges in molecules as
-included in the available force fields does not necessarily have to follow the above relation $f_q = \epsilon _{el} ^{-1/2}$. 
-Instead, a consistent scaling factor would lie between the value of 0.75 i.e., (no electronic polarizability included in the original partial charges)
-and $1$ (i.e., electronic polarizability fully included in the original partial charges).
-It is important to note that even though the lipid bilayer interface has a large gradient in
-the total dielectric constant, the electronic part of polarizability can be considered to be almost constant
-as measurements give values around 2 
-for almost any biologically relevant systems \cite{some_original_work, leontyev11}.
-
-In this work, we employ ECC to develop a force field model for
-the POPC lipid that accurately describes the binding of sodium and calcium ions to the corresponding lipid bilayer.
-%This is a biologically highly relevant membrane feature, which is
-%poorly reproduced by currently available models. This can affect not only on membrane properties
-%in the presence of ions but also modulate interactions with charged moieties at the interface.
-The quality of the simulation model is judged by comparison 
-to experimental NMR data reporting on the ion binding
-affinity~\cite{akutsu81,altenbach84,scherer89}, 
-as discussed in detail in Ref.~\citenum{catte16}.
-The Lipid14~\cite{dickson14} force field parameters (available in a Gromacs format from Ref. \citenum{lipid14files})
-were used as a starting point, as they provide one of the most
-realistic descriptions of the head group order parameter response to ions 
-among the available lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}).
-Moreover, the Lipid14 model has a relatively realistic glycerol backbone and
-head group structure comparable to other state of the art lipid models~\cite{botan15}.
-
-We applied the ECC correction to the Lipid14 model of POPC by scaling of partial charges of the head group,
-glycerol backbone, and carbonyl regions, which are the polar parts in phospholipids,
-expected to have a contribution to the cation binding.
-We did not modify the hydrocarbon chain parameters, 
-as they do not come in direct contact with ions from the solution 
-and as they are already highly optimized to provide a good description of the
-hydrophobic part of lipid bilayers \cite{dickson14,ollila16}.
-The latter is in contrasts with the situation for the glycerol backbone and the head group regions, 
-which calls for improvements in available lipid models~\cite{botan15}.
-
-The above discussed interval for the scaling factor, $f_q~\in~(0.75, 1.0)$,
-was explored to find the optimal value reproducing the experimental ion binding affinities.
-We first observed that the scaling
-of the partial charges reduced the ion binding affinity and the 
-related head group order parameter response in general.
-We then compared the results with different scaling factors to the experimental 
-NMR data~\cite{akutsu81,altenbach84,scherer89} finding the optimal 
-value $f_q = 0.8$,
-which is only slighly higher than the original scaling factor for the ions in water of 0.75.
-
-Although scaling of the partial charges improved 
-the head group order parameter response and ion binding affinity,
-it at the same time deteriorated certain membrane properties; 
-namely the area per lipid decreased below the experimental value. 
-After directly scaling the partial charges of the head group, the glycerol
-backbone, and carbonyls with a factor of 0.8, 
-the area per lipid in the simulation drops approximately to 60~\AA$^2$,
-which is smaller than the experimental value of 64.3~\AA$^2$ \cite{kucerka11}, 
-as well as the value from the original Lipid14 model of $65.6 \pm 0.5$~\AA$^2$~\cite{dickson14}.
-The decrease of the area per lipid is found to arise from a reduced hydration of the lipid head group region
-due a reduced head group polarity after scaling charges. 
-This effect is compensated
-by reducing the effective radii of atoms with the scaled charges.
-This is in practice done by changing the parameters $\sigma$ in the Lennard-Jones potential 
-in a similar way as was done previously for the ECC-ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
-After reducing the $\sigma$ parameters of affected atoms by a factor of $f_\sigma = 0.89$, 
-the area per molecule is restored
-back, being very close to the experimental value (Table~\ref{tab:apls}).
+
+Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
+This scaling factor successfully improves the accuracy of the simulation of solvated ions \cite{kohagen14, kohagen16, REF}, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+
+While using a scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{Leontyev2011}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
+Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
+Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}.
+Considering that RESP charges are usually refined to improve certain experimental observables, to some extent,  the electronic polarizability effect of the solvent is included in modern force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
+Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$.
+Instead, a consistent scaling factor would lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
+It is important to note that the electronic part of polarizability is around 2 for almost any biologically relevant environment \cite{some_original_work, leontyev11}.
+
+Here we develop ECC-POPC lipid model that accurately describes the binding of sodium and calcium ions to the resulting lipid bilayer.
+We applied the ECC correction to the Lipid14 model of POPC by scaling of partial charges of the head group, glycerol backbone, and carbonyl regions.
+These are the polar parts of phospholipids which can contribute to the cation binding.
+The Lipid14~\cite{dickson14} force field parameters (available in a Gromacs format from Ref.
+\citenum{lipid14files}) were used as a starting point.
+This model provides the best response of the head group to ions among the available lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}).
+Additionally, the Lipid14 model has a relatively realistic head group, glycerol backbone, and tails structure~\cite{botan15}.
+
+To develop the ECC-POPC model, we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$, to reproduce the experimental ion binding affinities.
+In general, scaling down the partial charges reduces the ion binding affinity and the head group order parameter response which can be measured using NMR~\cite{akutsu81, altenbach84, scherer89}.
+We found an optimal value $f_q = 0.8$, which is only slightly higher the theoretical one ($f_q=0.75$).
+Directly applying the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and the carbonyls reduces the area per lipid to 60~\AA$^2$.
+This area is smaller than the original in Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and the experimental value (64.3~\AA$^2$) \cite{kucerka11}.
+The decrease of the area per lipid arises from a reduced hydration of the lipid head group region after scaling charges which effectively reduces head group polarity.
+We solve this difficulty by reducing the effective radii of the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a factor of $f_\sigma = 0.89$.
+The same solution was taken for the ECC-ions in solution \cite{kohagen14, kohagen16, Pluharova2014}.
+After reducing the $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
 

### 2017-11-10 Manuscript/manuscript.tex
Updates to Electronic continuum correction for lipid bilayers

@@
 In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by including the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
-Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of aqueous solutions or ionic liquids \cite{jonsson86, egberts94, beichel14}.
+Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
 However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
 
-According to ECC mean-field correction, for atomic ions, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
-Following Coulomb's law,  we can directly incorporate the results by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$.
-
+According to ECC, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
+Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
 \begin{equation}
-Q^{ECC} = f_q \cdot Q ,
+Q^{ECC} = f_q \cdot Q
 \end{equation}
-
+for the ECC corrected charges in simulation. 
 Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor successfully improves the accuracy of the simulation of solvated ions \cite{kohagen14, kohagen16, REF}, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+This scaling factor successfully improves the accuracy of simulations of solvated ions \cite{kohagen14, kohagen16, REF}, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+It is important to note that the 
+high frequency dielectric constant 
+is around 2 for almost any biologically relevant environment \cite{some_original_work, leontyev11}.
+The dielectric discontinuity in a lipid bilayer arises
+from the orientational polarization of the molecules. 
+Thus, the same correction for the electronic polarizability can be 
+applied throughout the lipid bilayer interface.
 
@@
 Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}.
-Considering that RESP charges are usually refined to improve certain experimental observables, to some extent,  the electronic polarizability effect of the solvent is included in modern force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
+Considering that water molecules are often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
 Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$.
 Instead, a consistent scaling factor would lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
-It is important to note that the electronic part of polarizability is around 2 for almost any biologically relevant environment \cite{some_original_work, leontyev11}.
-
-Here we develop ECC-POPC lipid model that accurately describes the binding of sodium and calcium ions to the resulting lipid bilayer.
-We applied the ECC correction to the Lipid14 model of POPC by scaling of partial charges of the head group, glycerol backbone, and carbonyl regions.
-These are the polar parts of phospholipids which can contribute to the cation binding.
-The Lipid14~\cite{dickson14} force field parameters (available in a Gromacs format from Ref.
-\citenum{lipid14files}) were used as a starting point.
-This model provides the best response of the head group to ions among the available lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}).
-Additionally, the Lipid14 model has a relatively realistic head group, glycerol backbone, and tails structure~\cite{botan15}.
-
-To develop the ECC-POPC model, we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$, to reproduce the experimental ion binding affinities.
-In general, scaling down the partial charges reduces the ion binding affinity and the head group order parameter response which can be measured using NMR~\cite{akutsu81, altenbach84, scherer89}.
-We found an optimal value $f_q = 0.8$, which is only slightly higher the theoretical one ($f_q=0.75$).
+
+Here we develop ECC-POPC lipid model that accurately describes the binding 
+of sodium and calcium ions to the lipid bilayer. 
+The Lipid14~\cite{dickson14} force field parameters 
+(available in a Gromacs format from Ref. \citenum{lipid14files}) were used as a starting 
+point. 
+This model provides the best response of the head group to ions among the available 
+lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
+has a relatively realistic head group, glycerol backbone, and acyl chain structure~\cite{dickson14,botan15}.
+We applied the ECC correction 
+to the Lipid14 model of POPC by scaling of 
+partial charges of the head group, glycerol 
+backbone, and carbonyl regions. 
+These are the polar parts of phospholipids which can 
+contribute to the cation binding.
+
+To reproduce the experimental ion binding affinities,
+we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
+In general, scaling down the partial charges reduces the ion binding affinity 
+and the experimentally measurable head group order parameter response to the bound charge~\cite{akutsu81, altenbach84, scherer89}.
+We found an optimal value $f_q = 0.8$, which is only slightly higher than the theoretical one ($f_q=0.75$).
 Directly applying the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and the carbonyls reduces the area per lipid to 60~\AA$^2$.
@@
 We solve this difficulty by reducing the effective radii of the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a factor of $f_\sigma = 0.89$.
-The same solution was taken for the ECC-ions in solution \cite{kohagen14, kohagen16, Pluharova2014}.
+The same solution was taken for the ECC-ions in aqueous solvent \cite{kohagen14, kohagen16, Pluharova2014}.
 After reducing the $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).

### 2017-11-10 Manuscript/manuscript.tex
Electrometer concept section update

--- 2017-11-09 Manuscript/manuscript.tex
@@
 \subsection{Electrometer concept} \label{section:electrometer}
-Ion binding in lipid bilayers was compared between experiments and simulations 
-using the lipid head group order parameters and the
-''electrometer concept'' \cite{seelig87,catte16}, which
-is based on the experimental observation that the C-H bond
-order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group
-(Fig.~\ref{simVSexpNOions}) are proportional to the amount 
-of charge bound per lipid~\cite{seelig87}.
-The change of the order parameters measured with varying aqueous
-ion concentrations can be then related to the amount of bound ions.
-
-The electrometer concept can be used to compare the ion binding affinity to lipid 
-bilayers between MD simulations and  NMR experiments, because the order
-parameters can be accurately determined from both techniques~\cite{catte16,ollila16}.
-The order parameters for all C-H bonds in lipid molecules, including
-$\alpha$ and $\beta$ segments in head group, can be accurately measured
-using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}. From MD simulations the
-order parameters can be calculated using the definition
+Comparing MD simulation to NMR experiments, we can validate the simulation ion binding affinity to lipid bilayers using the ''electrometer concept''~ \cite{seelig87, catte16}.
+This method relies on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
+Comparing this response is an excellent way to validate force fields as both NMR and MD can accurately determine order parameters ~\cite{catte16, ollila16}.
+The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
+From MD simulations the order parameters can be calculated using the definition
 \begin{equation}\label{OP}
@@
 \end{equation}
-where $\theta$ is the angle between the bond and membrane
-normal and the average is taken over all sampled configurations \cite{ollila16}.
-
-The relation between the amount of the bound charge per lipid,  $X^\pm$, and
-the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$,
-is empirically quantified as~\cite{seelig87,ferreira16}
+where $\theta$ is the angle between the C-H bond and membrane normal. Angular brackets point to the average over all sampled configurations.
+
+The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
 \begin{equation}\label{OPchangeEQ}
-  \Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx \frac{4m_i }{3\chi}X^\pm,
+\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
 \end{equation}
-where $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge,
-$i$ refers to either $\alpha$ or $\beta$ carbon,
-$m_i$ is an empirical constant depending on the valency and position of the bound charge,
-and the experimental value \cite{seelig77,davis83}, $\chi \approx$\,167\,kHz, is used for the quadrupole coupling constant.
-Atomic absorption spectra and $^2$H NMR data gave
-$m_\alpha=-20.5$~kHz  and $m_\beta=-10.0$~kHz for the binding of
-Ca$^{2+}$ to POPC bilayer (in the presence of 100~mM NaCl)~\cite{altenbach84,ollila16,catte16}.
-The slopes are negative, because the head group order parameters
-decrease with a bound positive charge and increase with a bound negative
-charge~\cite{ollila16,catte16}. This can be rationalized in terms of a change
-of the lipid head group dipole tilt towards water phase with the bound positive
-charge and {\it vice versa} with the negative charge~\cite{seelig87}.
-
-The measured change of the order parameter depends both on the head group response 
-to the bound charge and on the amount of the bound charge 
-(i.e. $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively). 
-The empirical factor $m_i$ has to be
-properly quantified before the electrometer concept can used to 
-analyze the binding affinities.
-This has been done experimentally for a wide range
-of systems~\cite{seelig87,beschiasvili91}. To calibrate 
-the response of the head group order parameter to the bound charge in simulations,
-we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide %(2C$_{16}$$^+$N2C$_1$Br$^-$)
-mixed within a POPC bilayer~\cite{scherer89}. Dihexadecyldimethylammonium\\[0.5cm]
+where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ empirical experimental constant depending on the valency and position of the bound charge.
+
+The measured change of the order parameter depends both on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).
+The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities.
+This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
+To calibrate the response of the head group order parameter to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide  (DHAB) bilayer~\cite{scherer89}.
+DHAB\\[0.5cm]
 %\vspace{0.5cm}
@@
 \vspace{0.5cm} \\
-is a cationic surfactant having two acyl chains and bearing a unit
-charge in the hydrophilic end. Thus, it is expected to locate
-in the bilayer similarly to the phospholipids and its molar ratio then 
-gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
+is a cationic surfactant having two acyl chains and bearing a unit charge in the hydrophilic end.
+Thus, it is expected to locate in the bilayer similarly to the phospholipids, and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
 

### 2017-11-13 Manuscript/manuscript.tex
Updated "Salt concentrations and binding affinity" and "Validation of lipid bilayer structure against experiments"

@@
 \subsection{Salt concentrations and binding affinity}
-The concentrations of salts are reported in two different ways
-in the NMR measurements of head group order parameters \cite{akutsu81, altenbach84}.
-The early work on the NMR electrometer concept 
-reports the salt concentrations in water before solvating
-the lipids~\cite{akutsu81}. 
-The later study uses the atomic absorption
-spectroscopy, and reports the salt concentration in the supernatant after the
-solvation of lipids~\cite{altenbach84}.
-In this work, we focus on POPC for which the latter definition was used~\cite{altenbach84}.
-The salt concentration in the aqueous bulk region was taken as that at 
-the farthest point from the lipid bilayer in the aqueous phase. 
-Note that in the previous study~\cite{catte16},
-the ion concentrations were calculated in water before solvating the lipids, 
-as in the earlier experiments~\cite{akutsu81}.
-Although there are noticable differences between these two
-definitions for concentrations of the CaCl$_2$ systems,
-they do not affect significantly conclusions drawn in this work or in the previous one~\cite{catte16}.
-
-To quantify the ion binding affinities to a lipid bilayer, we
-calculate the relative surface excess of ions with respect to water,~$\Gamma_i^w$~\cite{chattorajBOOK}.
-Such a quantity compares the adsorption of ions to the adsorption of water molecules 
-at the interface without the necessity of defining a Gibbs dividing plane 
-between the membrane interior and the water bulk region. 
-In our simulations we only assume that the interface is located between the ion-free hydrophobic interior
-of the lipid bilayer and the aqueous region far from the membrane.
-%our whole simulation box as
-%an interface flanked by a water phase and an ideal phase with the characteristics of to
-%the hydrophobic region in the middle of the membrane where no ions or waters are present.
-Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_i^w$ for simulations of lipid bilayers, 
+NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
+Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81} while others use atomic absorption spectroscopy and report the salt concentration in the supernatant after the solvation of lipids~\cite{altenbach84}.
+In this work, we use the latter definition.
+In particular, we measure the salt concentration in the aqueous bulk region as that at the farthest point from the lipid bilayer in the aqueous phase.
+Notice that this definition differs from the one used by Catte et al.~\cite{catte16} where the former definition is used.
+Although these two definitions vary when applied to Ca$^{2+}$, they do not affect significantly the conclusions of this work.
+
+To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}.
+Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing plane between the membrane interior and the water bulk region.
+In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.
+Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
+
 \begin{equation}\label{surfexcess}
-  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) .
+  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
 \end{equation}
-Here, $n_w$ and $n_i$ are the total numbers of water molecules and ions in the system,
-$C_w$ and $C_i$ are their respective bulk concentrations in the aqueous phase,
+Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
+$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase,
 and $A_b$ is the area of the unit cell in the membrane plane.
@@
 \subsection{Validation of lipid bilayer structure against experiments}
-
-The structures of lipid bilayers in simulations without ions were validated against NMR and 
-\mbox{x-ray} scattering experiments by calculating the order parameters
-for the C-H bonds and the scattering form factors. The former validates the structures sampled by the individual
-lipid molecules in simulations with atomic resolution, while the latter
-validates the dimensions of the lipid bilayer (i.e., its thickness and area per
-molecule)~\cite{ollila16}.
-
-The simulated order parameters were calculated for all C-H bonds
-in lipid molecules from Eq. \ref{OP}. Form factors were calculated 
-from the equation
+The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and \mbox{x-ray} scattering experiments by the scattering form factors.
+NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution.
+The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq.
+\ref{OP}.
+Scattering form factors validate the dimensions of the lipid bilayer (e.g., thickness and area per molecule).
+Form factors were calculated using
+
 \begin{equation}
@@
 \end{equation}
-where $\rho_{el} (z)$ is the total electron density,
-$\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk,
-and $z$ is the distance from the membrane centre along its normal 
-with $D/2$ being half of the unit cell size.
+
+\noindent where $\rho_{el} (z)$ is the total electron density, $\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk, and $z$ is the distance from the membrane center along its normal with $D/2$ being half of the unit cell size.
 

### 2017-11-14 Manuscript/manuscript.tex
Changed title methods subsection to be more accurate

@@
 
-\subsection{Validation of lipid bilayer structure against experiments}
+\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
 The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and \mbox{x-ray} scattering experiments by the scattering form factors.

### 2017-11-14 Manuscript/manuscript.tex
Updated sections: -Simulations of POPC bilayers with aqueous ions -Simulations of POPC bilayers with cationic surfactants

@@
 \subsubsection{Simulations of POPC bilayers with aqueous ions}
-Simulations of a POPC bilayer in pure water or at varying salt concentrations
-consisted of 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per each lipid
-in an orthorhombic simulation box with periodic boundary conditions. 
-The chosen SPC/E water model, used also in the previous parametrization of 
-ECC-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014},
-has a lowered dielectric constant, consistent with the ECC concept~\cite{leontyev11, leontyev14}.
-Results from simulations with other water models 
-(OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, 
-TIP3P~\cite{jorgensen83},
-TIP3p-FB and TIP4p-FB~\cite{Wang2014},
-and TIP4p/2005 \cite{Abascal2005}), performed to check for robustness of our approach, are available in SI.
-Sodium, calcium, and chloride 
-were modeled as ECC-ions with parameters from \citenum{jungwirth17-new-paper-to-be-published, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
-Scaled charges and van der Waals radii for atoms forming the POPC lipid were derived here, starting from the Lipid14 force field.
-For comparison, simulations with unscaled charges were perfomed using the Lipid14 model, the TIP3p model for water \cite{jorgensen83} and the ion model by
-Dang et al.~\cite{smith94,chang1999,dang2006}.
-Simulation data for the Lipid14 model with \AA{}qvist~\cite{aqvist90} ions and TIP3P~\cite{jorgensen83}
-water models reported in \cite{catte16} 
-were taken directly from \cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}. 
+Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
+The chosen SPC/E water model, also used in the previous parametrization of ECC-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}, has a lowered dielectric constant, consistent with the ECC concept~\cite{leontyev11, leontyev14}.
+We checked other water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}) to confirm the robustness of our approach (See SI).
+Sodium, calcium, and chloride ions were modeled as ECC-ions with parameters from Ref.
+\citenum{jungwirth17-new-paper-to-be-published, Pluharova2014}.
+% cite order: [Ca_2s, Na_s, Cl_2s] 
+Scaled charges and van der Waals radii for atoms forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{REF}.
+For comparison, simulations with unscaled charges were performed using the Lipid14 model~\cite{REF}, the TIP3p model for water \cite{jorgensen83}\todo{HECTOR: Do we compare this against scaled with SPC or TIP3P?}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
+Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from Ref.~\citenum{catte16}.
+
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).
-The simulation settings employed in this work are 
-summarized in Table \ref{tbl:mdpar}. 
-The simulatied trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.
+The simulation settings employed in this work are summarized in Table \ref{tbl:mdpar}\todo{HECTOR: Is this the SI?
+If so we have to indicate it.
+Now reference is weird.}.
+The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
 
 \subsubsection{Simulations of POPC bilayers with cationic surfactants}
-The topology of dihexadecyldimethylammonium
-was created with an automated topology builder \cite{malde11}. 
-The AmberTools program \cite{amber} was then used to generate the Amber-type force field
-parameters. These parameters were then converted to the Gromacs format with
-the acpype tool \cite{acpype}. The partial charges were then manually modified
-to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
-The parameters together with the simulation data are available
-at~\cite{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
-The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with
-mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
-All these systems contained 50 POPC molecules per leaflet, 6340 TIP3P~\cite{jorgensen83} water molecules, and
-6, 14, 21, 35, or 50 surfactants per leaflet. 
-Chloride counter ions were used in simulations, 
-while bromide was the counter ion in the experiment \cite{scherer89}.
-The Lipid14 model was used for POPC.
-The first 20~ns of the total simulation time of 200~ns
-was considered as an equlibration time and was omitted from the analysis.
-Sufficient lipid neighbor exchange occured during the simulations.
-
-The systems were simulated with the ECC-lipids model for POPC 
-and SPC/E model for water using the same setup.
-In these simulations ECC was also applied to the cationic surfactant 
-in the same way as to POPC, i.e.,
-all partial charges were scaled with the same factor as for ECC-lipids, i.e. $f_q=0.8$, 
-and the same atom types with reduced $\sigma$ parameters from ECC-lipids were used. 
-\todo{Joe: I think that the followin is already said in the beginning of this section -- Lipid14 simulaitons.}
-For comparison, simulations with unscaled charges were also performed.
+The topology of dihexadecyldimethylammonium was created with the automated topology builder \cite{malde11}.
+The AmberTools program~\cite{amber} was then used to generate the Amber-type force field parameters.
+These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
+The partial charges were then manually modified in the non-scaled version~\cite{catte16} to approximately match similar segments in Lipid14 \cite{dickson14}.
+For simulation with ECC-POPC, the partial charges of the headgroup were scaled, i.e., $f_q=0.8$, and the $\sigma$ of the scale atoms reduced as in POPC-ECC.
+
+The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
+All these systems contained 50 POPC molecules per leaflet, 6340 TIP3P~\cite{jorgensen83} water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
+Chloride counter ions were used in simulations, while bromide was the counter ion in the experiment \cite{scherer89}.
+For POPC either the Lipid14 or ECC-POPC model was used.
+The first 20~ns of the total simulation time of 200~ns was considered as an equilibration time and was omitted from the analysis.
+Sufficient lipid neighbor exchange occurred during the simulations.
+Further details for the unscaled simulations and the trajectories are in Ref.~\citenum{catte16}.
 

### 2017-11-14 Manuscript/manuscript.tex
More changes related to POPC membrane structure and dynamics section

@@
   \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
-    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments.\label{tab:apls} }
+    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments.{ \color{red} HECTOR:Are we using here the same water model?
+If not is it fair the comparison considering the large changes observed in the SI table?}\label{tab:apls} }
   \begin{tabular}{l|c c}
@@
     \hline
-    ECC-lipids                & 63.2$\pm$ 0.6  &  300       \\
+    ECC-lipids                & {\color{red}63.2}$\pm$ 0.6  &  300       \\
     \hline
@@
 of a POPC bilayer with a comparable accuracy in Fig.~\ref{simVSexpNOions}.
-The area per lipid from the Lipid14 model is slightly larger than the
+The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
 experimental value in Table~\ref{tab:apls}, while the value from the ECC-lipids model
-is slightly smaller. The values of the area per lipid of the ECC-lipids model show small variations
-when simulated with different water models (see Table~\ref{tab:apls_si} in SI),
-with all being close to the experimentally reported values.
+is $\approx$1\AA{} smaller\todo{The 63.2 value does not appear in the SI table.
+There it is 62.2 but I'm not sure corresponds to the same system}.
+The values of the area per lipid of the ECC-lipids model show larger variations
+when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~\ref{tab:apls_si} in SI\todo{HECTOR: change to real reference in SI}),
+while still being close to the experimentally reported values.
 We can thus conclude that the ECC-lipids model reproduces the experimental dimensions of the POPC
-lipid bilayer with accuracy comparable to that for other state of the art lipid models \cite{ollila16}.
+lipid bilayer with accuracy comparable to that for another state of the art lipid models \cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 
-Similarly, the acyl chain order parameters of the
-ECC-lipids model, as well as the Lipid14 model~\cite{dickson14}, agree 
-with the experimental values with in the error bars, as presented
-in Fig.~\ref{simVSexpNOions}. Notably, the experimentally measured forking and
-small order parameter values of $C_2$ segment in {\it sn}-2 chain are well
-reproduced by both models. This feature has been suggested to indicate that the carbonyl
-of {\it sn}-2 chain is directed towards the water phase, in contrast to the
-carbonyl in {\it sn}-1 chain, which would orient more along the bilayer
-plane~\cite{seelig75,schindler75,gawrisch92}. This may be an important
-feature for the ion binding details, which it is not fully reproduced by other
-available lipid models~\cite{ollila16}
-%This is expected as the ECC model does not modify
-%the already highly optimized parameters of this region in Lipid14.
-
-The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger
-in the ECC-lipids model than in the Lipid14 model, which is apparently related to the
-P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-lipids model, 
-see Fig.~\ref{OrderParameterCHANGESsurf}. This is a small change and with the current experimental data we cannot decide which of the two models provides more realistic
-headgroup conformations. The ECC-lipids model gives
-the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while
-the oposite is true for the $\alpha$ carbon.
-Despite small deviations from the experimental order parameter values
-in Fig. \ref{simVSexpNOions},
-the accuracy of both models in the glycerol backbone region
-is comparable to other state
-of art lipid models available in literature \cite{botan15}.
+Similarly, the acyl chain order parameters of the ECC-lipids model, as well as the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
+Notably, the experimentally measured forking and small order parameter values of $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
+This feature has been suggested to indicate that the carbonyl of {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in {\it sn}-1 chain, which would orient more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
+This arrangement may be a relevant feature for the ion binding details, which it is not fully reproduced by other available lipid models~\cite{ollila16}.
+%This is expected as the ECC model does not modify the already highly optimized parameters of this region in Lipid14
+
+The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger in the ECC-lipids model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-lipids model, see Fig.~\ref{OrderParameterCHANGESsurf}.
+Considering the available experimental evidence, it is not possible to suggest which of the two models provides more realistic headgroup orientation.
+The ECC-lipids model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
+The accuracy of both models in the glycerol backbone region is equally footed with any other state of art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
 

### 2017-11-14 Manuscript/manuscript.tex
Calibration of lipid electrometer: Response of POPC head groups to bound charge

@@
 
-Before proceeding to the sodium and calcium  ion binding affinities, we quantify
-the response of the headgroup order parameters to the amount of 
-bound charge by using mixtures of monovalent cationic
-surfactants (dihexadecyldimethylammonium)
-and POPC~\cite{scherer89}. The amount of bound charge per PC 
-in these systems is given simply by the molar fraction of cationic 
-surfactants, because essentially all the surfactants with two hydrophopic
-acyl chains can be assumed to locate in the lipid bilayers.
-The experimental data for these systems can be used to validate 
-the sensitivity of lipid headgroup order parameters
-to the amount of bound charge in simulations.
-
-The changes of the headgroup order parameters with increasing amount of 
-the cationic surfactant are compared for simulations and experiments~\cite{scherer89} in Fig. \ref{OrderParameterCHANGESsurf}.
-An approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ},
-is observed in both simulations and experiments at least for mole fractions
-below $\sim$30\%. The slope is, however, too steep in the Lipid14 model indicating that 
-the head group order parameters respond too sensitively to the bound positive charge.
-In contrast, the slope for the ECC-lipids model is in a very good agreement with experiments
-for the $\alpha$ segment, while being slightly
-underestimated for the $\beta$ segment.
-
-The headgroup P-N vector angle with respect to the membrane normal
-is also shown in Fig.~\ref{OrderParameterCHANGESsurf}
-as a function of the mole fraction of the cationic surfactant.
-As suggested previously~\cite{seelig87}, the headgroup orients
-more towards the water phase with the increasing amount of positive
-charge in a PC lipid bilayer. The effect is more pronounced in the
-Lipid14 model, for which the addition of 50\% mole fraction of the cationic
-surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle,
-while the corresponding change in the ECC-lipids model is 11$^{\circ}$.
-The difference is in line with the smaller order parameter 
-changes and the reduced charge--dipole interactions in the latter model.
-The lower sensitivity of the P-N vector angle response in the ECC-lipid
-model can be considered to be more realistic, because the changes of the headgroup
-order parameters as a function of the bound positive charge
-are in better agreement with experiments in this model.
-The results also imply that validation and improvements of MD simulation
-models are generally required for reliable simulation studies of the
-lipid headgroup responses to ions or other biomolecules, as also suggested
-previously \cite{botan15}.
+Before studying the sodium and calcium ion binding affinities, we have to quantify as reference the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
+These mixtures have a well-defined amount of bound charge per PC, i.e., the molar fraction of cationic surfactants.
+This direct relation results from the ability of dihexadecyldimethylammonium to be inserted in the lipid bilayers as any lipid because of its two hydrophobic acyl chains.
+Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid headgroup order parameters to the amount of bound charge in simulations \cite{REF}.
+
+The changes of the headgroup order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
+An approximately linear decrease of the order parameters, as expected from Eq.
+\ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
+The slope is, however, too steep in the Lipid14 model indicating that the head group order parameters response is too large to the bound positive charge.
+In contrast, the slope of the ECC-lipids model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.
+
+In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
+As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in a PC lipid bilayer.
+The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
+For example, the addition of 50\% mole fraction of the cationic surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only 11$^{\circ}$ in the ECC-POPC model.
+The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model.
+The lower sensitivity of the P-N vector angle response in the ECC-POPC model agrees better with experiments.
+The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules{\color{red}, as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
 

### 2017-11-14 Manuscript/manuscript.tex
Small edits from II B to III B

@@
 \end{equation}
-where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ empirical experimental constant depending on the valency and position of the bound charge.
+where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is a empirical constant depending on the valency and position of the bound charge.
 
@@
 This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
-To calibrate the response of the head group order parameter to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide  (DHAB) bilayer~\cite{scherer89}.
+To calibrate the response of the head group order parameter to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
 DHAB\\[0.5cm]
@@
 In this work, we use the latter definition.
-In particular, we measure the salt concentration in the aqueous bulk region as that at the farthest point from the lipid bilayer in the aqueous phase.
+In particular, we measure the salt concentration in the aqueous bulk region using the farthest point from the lipid bilayer in the aqueous phase.
 Notice that this definition differs from the one used by Catte et al.~\cite{catte16} where the former definition is used.
-Although these two definitions vary when applied to Ca$^{2+}$, they do not affect significantly the conclusions of this work.
+Although these two definitions vary when applied to Ca$^{2+}$, they do not significantly affect the conclusions of this work.
 
@@
 Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
-
  \begin{equation}\label{surfexcess}
@@
 The chosen SPC/E water model, also used in the previous parametrization of ECC-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}, has a lowered dielectric constant, consistent with the ECC concept~\cite{leontyev11, leontyev14}.
-We checked other water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}) to confirm the robustness of our approach (See SI).
+We tested also other water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}) to evaluate the robustness of our approach (See SI).
 Sodium, calcium, and chloride ions were modeled as ECC-ions with parameters from Ref.
@@
 % cite order: [Ca_2s, Na_s, Cl_2s] 
-Scaled charges and van der Waals radii for atoms forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{REF}.
-For comparison, simulations with unscaled charges were performed using the Lipid14 model~\cite{REF}, the TIP3p model for water \cite{jorgensen83}\todo{HECTOR: Do we compare this against scaled with SPC or TIP3P?}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
-Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from Ref.~\citenum{catte16}.
+Scaled charges and van der Waals radii for atoms forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
+For comparison, simulations with unscaled charges were performed using the Lipid14 model~\cite{dickson14}, the TIP3p model for water \cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
+\todo{HECTOR: Do we compare this against scaled with SPC or TIP3P? SAMULI: Joe, please fill the information about Lipid14 simulation with ECC-ions.}
+Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
+Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
 
@@
 These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
-The partial charges were then manually modified in the non-scaled version~\cite{catte16} to approximately match similar segments in Lipid14 \cite{dickson14}.
+The partial charges were then manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
 For simulation with ECC-POPC, the partial charges of the headgroup were scaled, i.e., $f_q=0.8$, and the $\sigma$ of the scale atoms reduced as in POPC-ECC.
@@
 The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
-All these systems contained 50 POPC molecules per leaflet, 6340 TIP3P~\cite{jorgensen83} water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
-Chloride counter ions were used in simulations, while bromide was the counter ion in the experiment \cite{scherer89}.
-For POPC either the Lipid14 or ECC-POPC model was used.
+All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
+Chloride counter ions were used in simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
+For POPC either the Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987}  were used.
 The first 20~ns of the total simulation time of 200~ns was considered as an equilibration time and was omitted from the analysis.
@@
   \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
-    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments.{ \color{red} HECTOR:Are we using here the same water model?
-If not is it fair the comparison considering the large changes observed in the SI table?}\label{tab:apls} }
+    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments.
+    { \color{red} HECTOR:Are we using here the same water model? If not is it fair the comparison considering the large changes observed in the SI table?
+      SAMULI: For Lipid14 simulations the water is TIP3P and SPC/E for ECC-lipids. I do understand why the comparison would
+    not be fair in this context?}\label{tab:apls} }
   \begin{tabular}{l|c c}
@@
 experimental value in Table~\ref{tab:apls}, while the value from the ECC-lipids model
-is $\approx$1\AA{} smaller\todo{The 63.2 value does not appear in the SI table.
-There it is 62.2 but I'm not sure corresponds to the same system}.
+is $\approx$1\AA{} smaller
+\todo{HECTOR: The 63.2 value does not appear in the SI table. There it is 62.2 but I'm not sure corresponds to the same system
+SAMULI: I can find the 63.2 from the SI table with SPE/E and the value 62.2 with OPC3, so everything seems to be ok for me.}.
 The values of the area per lipid of the ECC-lipids model show larger variations
@@
 We can thus conclude that the ECC-lipids model reproduces the experimental dimensions of the POPC
-lipid bilayer with accuracy comparable to that for another state of the art lipid models \cite{ollila16}.
+lipid bilayer with accuracy comparable to that for nother state of the art lipid models \cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 
-Before studying the sodium and calcium ion binding affinities, we have to quantify as reference the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
+Before studying the sodium and calcium ion binding affinities, we have to quantify the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
 These mixtures have a well-defined amount of bound charge per PC, i.e., the molar fraction of cationic surfactants.
 This direct relation results from the ability of dihexadecyldimethylammonium to be inserted in the lipid bilayers as any lipid because of its two hydrophobic acyl chains.
-Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid headgroup order parameters to the amount of bound charge in simulations \cite{REF}.
+Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid headgroup order parameters to the amount of bound charge in simulations \cite{scherer89}.
 
@@
 \ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
-The slope is, however, too steep in the Lipid14 model indicating that the head group order parameters response is too large to the bound positive charge.
+The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters is too large to the bound positive charge.
 In contrast, the slope of the ECC-lipids model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.
@@
 The lower sensitivity of the P-N vector angle response in the ECC-POPC model agrees better with experiments.
-The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules{\color{red}, as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
+The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules.
+%{\color{red}., as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
 

### 2017-11-14 Manuscript/manuscript.tex
Updated "Binding affinities to POPC membrane validated through lipid electrometer" section

--- 2017-11-14 Manuscript/manuscript.tex
@@
   \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
+  \caption{\label{fig:delta_ordPar_NaCl}
+    Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
+    from simulations with different force fields together with 
+    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
+    Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
+    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
+  }
+  %\todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
   \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
@@
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
-  \caption{\label{fig:delta_ordPar_NaCl}
-    The changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
-    from simulations with different force fields together with 
-    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
-    Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
-    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
-  }
-  %\todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
-\end{figure}
-
-Changes of the lipid bilayer headgroup order parameters from different simulations and
-experiments \cite{akutsu81,altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}
-as functions of NaCl or CaCl$_2$ concentrations.
+Changes of the lipid bilayer head group order parameters from different simulations and experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl} and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
 %Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.
-These results can be used to compare the ion binding affinities to lipid bilayers
-between simulations and experiments using the electrometer concept, because 
-the order parameters decrease proportionally 
-to the amount of the bound positive charge (Fig.~\ref{OrderParameterCHANGESsurf} and Ref.~\citenum{seelig87,catte16}). 
-A recent comparison of different simulation models to the experimental data
-revealed that most force fields significantly overestimate Na$^+$ ion binding to PC
-lipid bilayers and that none of the available models correctly reproduces
-the details of the binding of Ca$^{2+}$ ions \cite{catte16}.
-The only exception in Ref.~\citenum{catte16} was the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions,
-which reproduced the experimentally measured 
-small order parameter changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl})
-and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81,altenbach84}.
-However, the change of the headgroup order parameters as a function of CaCl$_2$ concentration 
-(Fig.~\ref{fig:delta_ordPar_CaCl}) was still grosly overestimated within this model.
-
-Since artifacts were reported in simulations with \AA{}qvist ions is water \cite{auffinger07},
-we also simulated the Lipid14 model with more realistic ion models by Dang et al.~\cite{smith94,chang1999,dang2006}, and
-ECC-ions~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014}. 
-However, we still observed overestimated binding for both \ce{Ca^{2+}} and Na$^+$ with these ions models 
-(Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}). 
-The results are in line with the previous work \cite{catte16},
-suggesting that additional improvements in the lipid parameters are
-required to correctly describe the binding of cations to phospholipid bilayers.
+These results can be used to compare the ion binding affinities to lipid bilayers between simulations and experiments using the electrometer concept because the order parameters decrease proportionally to the amount of the bound positive charge (Fig.~\ref{OrderParameterCHANGESsurf} and Ref.~\citenum{seelig87, catte16}).
+A recent comparison of different simulation models to the experimental data revealed that most force fields significantly overestimate Na$^+$ ion binding to PC lipid bilayers and that none of the available models correctly reproduces the details of the binding of Ca$^{2+}$ ions which is grossly overestimated \cite{catte16}.
+Only the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions reproduces the experimentally measured small order parameter changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81, altenbach84}.
+Since artifacts were reported in simulations involving \AA{}qvist ions is water \cite{auffinger07}, we tested the Lipid14 model with more realistic Na$^+$ models by Dang et al.~\cite{smith94, chang1999, dang2006}.
+In this case, overbinding reappeared (Fig.~\ref{fig:delta_ordPar_NaCl}).The results are in line with the previous work \cite{catte16}, suggesting that additional improvements in the lipid parameters are required to describe the binding of cations to phospholipid bilayers correctly.
 
@@
 %polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
-The results from the simulations combining the ECC-lipids with the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}
-exhibit a significanlty improved behavior of cation binding to the POPC bilayer in Fig.~\ref{fig:delta_ordPar_CaCl},
-showing a good agreement with experiments in the changes of 
-the lipid headgroup order parameters as functions of NaCl or CaCl$_2$ concentrations.
-Since also the headgroup order parameter
-response of the ECC-lipids model the to the bound positive charge 
-is in good agreement with experiment (see above section \ref{section:boundCHARGE}),
-we can conclude that this model correctly reproduces the binding affinities of
-Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
-
-In addition, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was
-signifincantly overestimated in the original Lipid14 model, the ECC-lipids model
-provides a greatly improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-Also the changes of the P-N vector angle are too pronounced
-for the Lipid14 model, for which the largest tilting toward water phase 
-induced by a $780\,\mathrm{mM}$ CaCl$_2$ concentration
-is approximately 17$^{\circ}$, while the corresponding
-value for ECC-lipids model is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
-
-Within the Lipid14 model, 
-the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated
-binding affinity and the too strong sensitivity of the headgroup tilt to the bound positive charge.
-This likely applies also to the other lipid models tested in a previous study~\cite{catte16},
-which underlines the importance of the comparison of the lipid headgroup order parameter response
-to the bound charge between simulations and experiments as presented in section \ref{section:boundCHARGE}.
+
+The results from the simulations combining the ECC-lipids with the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
+Considering that we can also reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC-model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
+Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-lipids model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
+
+\todo{HECTOR: I do not believe in the content of the following sentence as it is.
+How do we know that the changes of the P-N angle are too pronounced for Lipid14?.
+There is no direct correlation between angle and order parameter.}
+Also the changes of the P-N vector angle are too pronounced for the Lipid14 model, for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$ CaCl$_2$ concentration is approximately 17$^{\circ}$, while the corresponding value for ECC-lipids model is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
+\todo{HECTOR: I have a problem when comparing fig.
+2 and 3.
+If I read it right the behavior when adding calcium and surfactant is not the same.
+Is the latter changing the area per lipid and therefore the surface charge density?
+If this will be the case then besides the effect of the charge we will have a packing effect.
+This is more relevant for the more concentrated cases.
+Shouldn't we try to discuss the two figures and what the differences and similarities mean?
+Given our previous work, should we check the behavior respect the charge density?}
+
+Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge.
+It is plausible that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance to validate the lipid headgroup order parameter response to the bound charge.
 
 Finally, the ion binding affinities for the ECC-lipids model with different water models are compared in SI.
-In general, the performance of ECC-lipids with any of the water models 
-is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated
-with the four-site water models and with TIP3p model.
+In general, the performance of ECC-lipids with any of the water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3p model.
 

### 2017-11-14 Manuscript/manuscript.tex
Updated "Structure and affinity of cation binding to POPC membrane"

@@
 
+\todo{HECTOR: I'm not sure I understand the meaning of "structure" in the title.
+In general the titles in the result section are much alike between them, specially the first ones.
+This makes difficult to guess where to find a particular piece of information.
+I will make an effort to make them more clear.
+}
+
 \begin{table}[tb!]
-  \caption{Relative surface excess of calcium with respect to water, 
-  and the ratios of the amount of calcium cations 
-  bound to phosphate or carbonyl moieties 
-  to the amount of such cations bound to both moieties in a POPC bilayer 
-  from various simulation models. 
-  $C_w$ is the concentration of CaCl$_2$ in water before adsorbing to the phospholipid bilayer. 
+  \caption{Relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$) in different POPC bilayer models.
+All systems contain initially 350~mM of Ca$^{2+}$ (C$_w$).
+C$_b$ correspond to the end water bulk concentration of Ca$^{2+}$.
+$r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ show the location for bound calcium in percentages for phosphate and carbonyl oxygens.
+{\color{red}HECTOR:PLEASE CHECK}
   \label{tab:binding}}
   \begin{tabular}{l|c | c c | c c}
-    model                 & $\Gamma_{Ca}^w \,/\, \mathrm{nm}^{-2}$ & $C_{w}$ & $C_b\,/\,\mathrm{mM}$   & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+    model                 & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$ & $C_{w}$ & $C_b\,/\,\mathrm{mM}$   & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
     \hline
@@
 
-Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation 
-models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm Ca}^w$, from Eq.~\ref{surfexcess}.
+Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
+The values of $\Gamma_{\rm ion}^{\rm water}$, shown in Table~\ref{tab:binding}, were calculated from simulations with the same initial concentration of CaCl$_2$ of 350~mM before solvating the lipids.
+As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller than for the other models, 0.13--0.35~nm$^{-2}$.
+\todo{HECTOR: I have deleted the sentence relating the uncertainty and the error as in my opinion there are other factors which are more important.
+The most important the slow convergence.}.
+Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our POPC-ECC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having a negative value $\Gamma_{Na}^{water} = (-0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}).
+This value means that water molecules are preferred to sodium and chloride ions at the membrane-water interface.
+This is in contradiction with most of the available lipid force fields, which predict a positive binding of sodium to PC lipid bilayers \cite{catte16}.
+\todo{do we need to add a NaCl table or an extended table for CaCl2 using different concentrations in the SI?
+I think this is very important information}
+
 %using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
-The values of $\Gamma_{\rm Ca}^w$, shown in Table~\ref{tab:binding},
-were calculated from simulations with the same concentration of CaCl$_2$ of $350\,\mathrm{mM}$  with 
-respect to water before solvating the lipids.
-As expected from the changes of the lipid headgroup order parameters in Fig. \ref{fig:delta_ordPar_CaCl}, 
-the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^w = (0.06 \pm 0.01) \rm{nm}^{-2}$,
-is significantly smaller than for the Lipid14 model with \AA{}qvist ions, $\Gamma_{\rm Ca}^w = (0.13 \pm 0.01) \rm{nm}^{-2}$,
-or with Dang ions, $\Gamma_{\rm Ca}^w = (0.23 \pm 0.03) \rm{nm}^{-2}$,
-or with ECC-ions,  $\Gamma_{\rm Ca}^w (0.35 \pm 0.11) \rm{nm}^{-2}$,
-where the large uncertainity arises from a too small simulation box (Fig.~\ref{fig:cacl-dens}).
-Interestingly, the relative surface excess of 
-\ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) 
 %calculated using density profiles in Fig.~\ref{fig:cacl-dens}
-is not only quantitatively but also qualitatively different from \ce{CaCl2}
-having a negative value $\Gamma_i^w = (-0.11 \pm 0.01) \rm{nm}^{-2}$ 
-meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface. 
-This is in contrast to the most of the available lipid force fields,
-which predict a positive binding of sodium to
-PC lipid bilayers \cite{catte16}.
+
 %\todo{JOE: Add references to NaCl simulation papers that show peaks.
@@
 
-In a POPC bilayer, the ratio of the amount of calcium cations 
-bound to either phosphate or carbonyl moieties 
-to the amount of such cations bound to both moieties
-from various simulation models
-%The probabilities for a bound \ce{Ca^{2+}} ion to interact with
-%different oxygen moieties of POPC 
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O}$,
-was analyzed by counting the contacts within the distance of $0.3 \, \mathrm{nm}$, 
-as done previously in Ref.~\citenum{javanainen17}. 
-%
-%SAMULI: I commented out the average number of contacts. All the essential information
-%for this is in the probabilities.
-%
-%In contrast to sodium, the dominant contribution
-%to the binding of \ce{Ca^{2+}} to POPC membranes comes from the phosphate group. 
-%% Numbers from simulation with 346mM conc. CaCl2 (OPC3)
-%The average number of contacts per lipid 
-%between \ce{Ca^{2+}} at $290\,\mathrm{mM}$ concentration and any oxygen atom in POPC is 
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.134 $,    % = 17.1/128 
-%whereas if only phosphate oxygen atoms are considered 
-%the average number of contacts decreases only by a tiny amount to 
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.133 $,    % = 17.0/128 . 
-%and if only carbonyl oxygen atoms are considered 
-%%(phosphate oxygens contribute, but are not counted as contacts),
-%this value is merely
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.033 $.    % = 4.3/128 . 
-%
-% Numbers from simulation with 980/845mM conc. CaCl2 (OPC3)
-%The average number of contacts per lipid between \ce{Ca^{2+}} and any oxygen atom in POPC is 
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $,    % = 33.8/128 
-%whereas if only phosphate oxygen atoms are considered 
-%the average number of contacts decreases only by a tiny amount to 
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{PO_4} = 0.262 $,    % = 33.5/128 . 
-%and if only carbonyl oxygen atoms are considered 
-%(phosphate oxygens contribute, but are not counted as contacts),
-%this value is merely
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} = 0.066 $.    % = 8.4/128 . 
-The results in Table \ref{tab:binding} show that almost all (99\%) of the 
-bound Ca$^{2+}$ ions are in contact with the phosphate oxygens,
-with two thirds (67\%) interacting only with these species,
-while the remaining one third (32\%), interacting also with the carbonyl oxygens. 
-The interactions between
-calcium ions with only carbonyl oxygens are very unlikely (1\%).
-The individual lipids or acyl chains were not distinguished in the analysis,
-hence, simultaneous contacts of ions with both oxygen moieties may be inter or intra molecular,
-and may occur with carbonyls in either of the acyl chains.
-The most likely interactions between Ca$^{2+}$ ions and phosphate oxygens are visualized with
-the probability density isocontours in Fig.~\ref{fig:volmaps}.
-While the higher concentrations of \ce{CaCl2} naturally increases the amount of contacts per lipid,
-%for $850\,\mathrm{mM}$ concentration of \ce{CaCl2}
-%the average number of contacts per lipid is increased to
-%$\Sigma ^\mathrm{Ca^{2+}} _\mathrm{O} = 0.264 $, 
-the distribution of contacts between phosphate and carbonyl oxygens is not affected.
-%and density profiles of ions and water along the membrane normal in Fig.~\ref{fig:cacl-dens}.
+We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
+A maximum distance of 0.3~nm from any lipid oxygen is used to determine a bound calcium.
+The results in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in contact with the phosphate oxygens.
+From those only one third (32\%) also interact with the carbonyl oxygens.
+The interaction between calcium ions with only carbonyl oxygens is rare (1\%).
+Neither individual lipids nor source chain of carbonyl is distinguished in the analysis.
+The most likely interactions between Ca$^{2+}$ ions and phosphate oxygens are visualized with the probability density isocontours in Fig.~\ref{fig:volmaps}.
+While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.
 
@@
 
-Even though \ce{Na+} ions do not specifically bind to a POPC bilayer,
-they still interact with its oxygen moieties. 
-%the average number of contacts per lipid with any oxygen atom in POPC is 
-%$\Sigma ^\mathrm{Na^{+}} _\mathrm{O} = 0.106 $,    % = 17.1/128 
-%whereas if only phosphate oxygen atoms are considered 
-%this number decreases to 
-%$\Sigma ^\mathrm{Na^{+}} _\mathrm{PO_4} = 0.085 $,    % = 17.0/128 . 
-%and if only carbonyl oxygen atoms are counted 
-%it is even lower, 
-%$\Sigma ^\mathrm{Na^{+}} _\mathrm{O_{carb.}} = 0.048 $.    % = 4.3/128 . 
-The results  from simulation at $1\,$M \ce{NaCl} concentration show 
-that 55\% of \ce{Na+} ions interact with only phosphate oxygens of POPC 
-and 20\% with only carbonyl oxygens,
-and the rest, 25\%, is interacting with both.
-%, phosphate and carbonyl oxygens.  
-%This is in line with the calculated negative value of 
-%relative surface excess of sodium with respect to water, $\Gamma_i^w$, and
-%their density profiles along the membrane normal (Fig.~\ref{fig:nacl-dens})
-%suggesting that the populations of bound sodium cations 
-%naturally decrease with the decreasing density of water.
-
-In conclusion, the results suggest that calcium ions bind specifically to phosphate
-oxygens, occasionally interacting also with carbonyls. This is in good agreement
-with previous conclusions from several experimental and computational
-studies~\cite{hauser76,hauser78,herbette84,cevc90,binder02}, but
-suggesting a lower relative binding affinity to the carbonyls than inferred from previous MD simulation
-studies~\cite{bockmann03,bockmann04,melcrova16,javanainen17}.
-Sodium ions also interact primarily with phosphate oxygens of the POPC, 
-but in contrast to calcium, the interactions purely with carbonyls are also significant.
-The physiological relevance of these details of sodium interactions is, however, unclear
-due to its overall very weak binding affinity.
+Even though \ce{Na+} ions do not specifically bind to a POPC bilayer, they still interact mostly with its oxygen moieties.
+The results from the simulation at $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions interact with only phosphate oxygens of POPC and 20\% with only carbonyl oxygens, and the rest, 25\%, is interacting with both.
+
+In conclusion, the results suggest that calcium ions bind specifically to phosphate oxygens, occasionally interacting also with carbonyls.
+This is in good agreement with previous conclusions from several experimental and computational studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}, but suggesting a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
+Sodium ions also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
+\todo{This last sentence is weak and not relevant.
+If you want to keep it go ahead, but then I would ask you a simple question: Is it incompatible to have no overall binding (this relates to the whole surface volume} with having small cavities in the membrane with an increase Na+ concentration.
+If the answer is negative maybe the statement is incorrect and who knows it might be relevant for something.
+After all even Ca2+ affinity changes in the presence of sodium.} The physiological relevance of these details of sodium interactions is, however, unclear due to its overall very weak binding affinity.
 

### 2017-11-14 Manuscript/manuscript.tex
Corrected wrong bracket.

@@
 \todo{This last sentence is weak and not relevant.
-If you want to keep it go ahead, but then I would ask you a simple question: Is it incompatible to have no overall binding (this relates to the whole surface volume} with having small cavities in the membrane with an increase Na+ concentration.
+If you want to keep it go ahead, but then I would ask you a simple question: Is it incompatible to have no overall binding (this relates to the whole surface volume) with having small cavities in the membrane with an increase Na+ concentration.
 If the answer is negative maybe the statement is incorrect and who knows it might be relevant for something.

### 2017-11-14 Manuscript/manuscript.tex
Updated "Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to the POPC membrane"

--- 2017-11-14 Manuscript/manuscript.tex
@@
 
-Simple binding models have been used previously to interpret
-the same experimental data \cite{altenbach84,macdonald87} 
-that are used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}.
-Results from measurements of the lipid headgroup order parameters and from
-atomic absorption spectroscopy were best
-explained by a ternary complex binding model with a binding
-stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-However, a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
-also provided a good fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
-Note that the results based on these binding models and the measured headgroup order parameter changes
-are also inluded in the tabulated experimental binding constants~\cite{marsh13}.
-
-Since the same experimental data are reproduced by the present ECC-lipids model,
-we can use the model to provide a more detailed interepration of the binding stoichiometry than
-the previously used simple binding models~\cite{altenbach84,macdonald87}. To directly evaluate the
-stoichiometry from simulations we calculated the relative propensities of \ce{Ca^{2+}}:POPC
-complexes with various stoichiometries by counting the individual lipid
-molecules within the distance $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion,
-as already done in the previous section~\ref{sec:affinity}.
-The results from POPC simulation with a 285~mM concetration of CaCl$_2$ in
-Fig.~\ref{fig:cacl_complexes} demonstrate the largest propensity (41\%) for the
-ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
-the probabilities of complexes with the stoichimetries of 1:1~(25\%)~and
-1:3~(34\%)~are only slightly lower. This
-suggests a more complex binding model than the previous simple ternary complex model.
-%The observed propensities of
-%$41\%$ for two lipids,
-%$25\%$ for one lipid, and
-%$34\%$ for three lipids 
-With a broad brushstroke, the simulation data can, however, be viewed such that
-one calcium binds to two lipids on average,
-because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other 
-(and complexes with more than three lipids per one calcium ion were not observed).  
-This probably explains why the simple the ternary complex model fits the
-experimental data, as well as the ECC-lipids simulation results, rather well
-(see Fig.~S\ref{fig:cacl-bind} in SI and its caption for details).
+Simple binding models have been used previously to interpret the experimental data \cite{altenbach84,macdonald87} we use in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
+In particular, NMR data about PC headgroup order parameters response and atomic absorption spectroscopy were best explained using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
+However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2}low concentrations~\cite{macdonald87}.
+\todo{HECTOR: Why is the following sentence remotely significant.
+I will vote for removing it} Note that the results based on these binding models and the measured headgroup order parameter changes are also included in the tabulated experimental binding constants~\cite{marsh13}.
+
+In this work, we reproduce the same experimental data used to calculate binding stoichiometry with our ECC-POPC model.
+Using our simulations, we have direct access to exact details of the binding stoichiometry without requiring any middle model as in experiments~\cite{altenbach84, macdonald87}.
+To evaluate the relative propensities for each stoichiometric complex (e.i., 1 Ca$^{2+}$: n POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
+In Fig.~\ref{fig:cacl_complexes}, from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ we found that the largest propensity is for the 1:2 complex (41\%).
+\todo {I seriously doubt that this is what they call a ternary complex but correct me if you know better}.
+Still, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower.
+This variability suggests a more complex binding model than the previously considered in a simple ternary complex model.
+With a broad brushstroke, the simulation data can, however, be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
+This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S\ref{fig:cacl-bind} in SI).
 

### 2017-11-14 Manuscript/manuscript.tex
Updated results Section F -Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in POPC membrane

@@
 
-Equilibration of \ce{Ca^{2+}} ions at the POPC bilayer takes hundreds of
-nanoseconds in MD simulations with current state of the art force fields,
-e.g. CHARMM36 and Slipids force fields~\cite{javanainen17}, suggesting that
-simulations at a temporal scale of several microseconds are required
-to reach the binding/unbinding equilibrium. 
-This is in line with the density profiles of \ce{Ca^{2+}} 
-from simulations in Fig.~\ref{fig:cacl-dens}, 
-which all show a region of a very small calcium density 
-around the level of cholines in POPC (i.e. around $2.5\,\mathrm{nm}$).
-In contrast, simulations with  ECC-lipids 
-exhibits only a relatively shallow local
-minimum at that level.
-This suggests that the smaller barrier for the exchange of calcium ions 
-shall lead to their faster exchange with the solvent within ECC-lipids.
-To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated 
-residence times of ions bound to the membrane.
-We use the same definition for ions bound to the membrane as in the above section~\ref{sec:affinity},
-i.e., an ion is considered as bound if it is within 0.3~nm from any oxygen atom
-belonging to a lipid molecule in a bilayer.
-
-The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer 
-at a $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
-from simulations with ECC-lipids 
-compared to CHARMM36 (taken from previous study \cite{javanainen17}, 
-simulation data available at Ref.~\citenum{zenodo.259376})
-are shown in Fig.~S\ref{fig:hist_residence_times} in SI. 
-%Based on Fig. S17 in that work, Ref.~\citenum{javanainen17},
-%we took the simulation with the apparent highest fluctuations of the number of bound calcium,
-%which can be obtained at \url{https://doi.org/10.5281/zenodo.259376}
-%and anaylsed the residence times.
-In the CHARMM36 simulation, approximately 20\% of the \ce{Ca^{2+}} residence times are longer than
-the whole length of the trajectory ($800\,\mathrm{ns}$).
-Within such a relatively short simulation, 
-we can only estimate that 
-less than 60\% of the bound residence times 
-is due to cations interacting with the phospholipid membrane 
-for less then half of the simulation length~(400~ns).
-Even longer residence times are observed
-in  simulations with CHARMM36 and Slipids models reported in previous works~\cite{javanainen17}.
-
-In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the
-ECC-lipids model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are 
-shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$.
-The longest observed residence time is now $141\,\mathrm{ns}$, 
-which is well below the total length of the simulation used for analysis, i.e., 200~ns.
-The exchange of \ce{Na+} ions at the POPC membrane is
-another order of magnitude faster, yielding 90\%
-of the residence times smaller than~$1\,\mathrm{ns}$, with
-the longest residence time being~$6\,\mathrm{ns}$. 
-\todo{Joe: the following is a bit more subtle argument. 
-I deem it not necessary to include it in the paper, however, 
-the experimental estimate of less tan 10us holds likely also for Charmm36/ECC-ions, as stated before.}
-Note that the ECC-lipids calcium results are in line with the experimental estimate that 
-the residence time of \ce{Ca^{2+}} at each individual PC headgroup 
-is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
+Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current state of the art force fields, e.g., CHARMM36 and Slipids force fields~\cite{javanainen17}.
+This suggests that at least several microseconds are required to reach the binding/unbinding equilibrium.
+\todo{HECTOR: I disagree with the following lines.
+In the case of most of the unscale force fields the issue is that the binding of Calcium to the membrane is very strong.
+Therefore, the reorganization takes too long leading to a very slow equilibration process.
+In this case, the size of the hill in the choline region is nothing compared to the size between the bound and solvated state which is the real bottleneck.
+Even in the case of the scaled simulations, I believe the same applies but to a lesser extent.
+I might be wrong in this last part.
+If you do not agree we can discuss it tomorrow via skype.} {\color {red} This is in line with the density profiles of \ce{Ca^{2+}} from simulations in Fig.~\ref{fig:cacl-dens}, which all show a region of a very small calcium density around the level of choline groups in POPC (i.e., around $2.5\,\mathrm{nm}$).
+In contrast, simulations with  ECC-lipids exhibits only a relatively shallow local minimum at that level.
+This suggests that the smaller barrier for the exchange of calcium ions shall lead to their faster exchange with the solvent within ECC-lipids.} To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated residence times of ions bound to the membrane.
+In our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
+
+\todo{HECTOR: I do fully understand the content of this paragraph.
+I will discuss this with Joe tomorrow.
+Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_w$ = 450~nM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Ref.~\citenum{javanainen17}) are shown in Fig.~S\ref{fig:hist_residence_times} in SI.
+In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory ({\color{red} 800~ns}).
+In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-lipids model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$.
+The longest observed residence time is around 150~ns, which is well below the total length of the simulation used for analysis, i.e., 200~ns.
+Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC headgroup is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
+The exchange of \ce{Na+} ions at the POPC membrane is another order of magnitude faster, yielding 90\% of the residence times smaller than~1~ns, with the longest residence time being~6~ns.
 
@@
 
-In conclusion, the results from the ECC-lipids model suggest that the
-exchange of calcium between the POPC bilayer and the solvent occurs within $\sim 100\,$ns timeframe,
-which is significantly faster than previously reported~\cite{javanainen17}.
-Sodium cations exhibit an even a faster exchange.
-This all suggests that simulations with a length of a several hundreds of
-nanoseconds are sufficient to simulate ion binding
-to phospholipid bilayers in equilibrium, when realistic force fields are used.
-This has not been the case with previous lipid force fields,
-which overestimate the binding strength of the sodium and calcium cations \cite{javanainen17,catte16}.
+In conclusion, the results from the ECC-lipids model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within $\sim$100~ns timeframe, which is significantly faster than observed in most of the available lipid models~\cite{javanainen17}.
+Sodium cations exhibit an even a more rapid exchange.
+This all suggests that simulations with a length of several hundreds of nanoseconds are sufficient to simulate ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
+This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and calcium cations \cite{javanainen17, catte16}.
 %Such a finding changes the point of view on binding of calcium to PC membranes from

### 2017-11-14 Manuscript/manuscript.tex
Updated conclusions

@@
 
-In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions
-to a POPC lipid bilayer can be accurately described within a classical MD simulation 
-force field, provided that electronic polarization is implicitly included via 
-the electronic continuum correction~\cite{leontyev11}.
-While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-lipids
-model agree with experiments with an accuracy comparable to the other state of the art lipid models,
-the model also reproduces the experimental lipid head group order parameter responses to
-a cationic surfactant, NaCl and CaCl$_2$ concentrations. 
-It thus represents a significant improvement over 
-other available lipid models, which overestimate cation binding affinities~\cite{catte16}.  
-The ECC-lipids model is build upon
-the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8
-and reducing the Van der Waals radii by a factor of 0.89 for the headgroup, glycerol backbone and carbonyl atoms.
+In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a classical MD simulation force field, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
+While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-lipids model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations.
+It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}.
+The ECC-lipids model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Van der Waals radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
 
-The good agreement with experiments enables us to 
-interpret NMR experiments with atomistic details using MD simulations with the ECC-lipid model.
-In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02},
-Ca$^{2+}$ ions interact mainly with phosphate oxygens.
-However, the stoichiometry of calcium binding is significantly more complicated than
-the simple ternary complex model, used to interpret NMR data, where one calcium binds to two POPC molecules~\cite{altenbach84}.
-While complexes with one calcium ion bound to two lipids are the most probable also in the
-ECC-lipids model, complexes of one or three lipids per one calcium
-were observed to relatively abundant and also almost equally likely. While the success of the simple ternary
-compex model in fitting NMR data is understandable based on the simulation results,
-it cannot capture the complex nature of calcium binding to phospholipid bilayers
-oberved in the present simulation.
+The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-lipid model.
+In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens.
+However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data, where one calcium binds to two POPC molecules~\cite{altenbach84}.
+While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-lipids model, complexes of one or three lipids per one calcium were observed to relatively abundant and also almost equally likely.
+While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the complex nature of calcium binding to phospholipid bilayers observed in the present simulation.
 
-Accurate description of cation binding to POPC bilayer paves the way for
-simulations of complex biochemical systems at cellular membranes with realistically described
-electrostatic interactions. To this end, the compatibility of
-the ECC-lipids model with existing models for proteins and other biological molecules should to be verified and, if necessary, 
-further adjustments along the ECC concept in force fields of molecules interacting with the lipids need to be done.
+Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions.
+To this end, the compatibility of the ECC-lipids model with existing models for proteins and other biological molecules should be verified and, if necessary, further adjustments following ECC concept to the force fields of the interacting molecules with the lipids should be addressed.
 

### 2017-11-15 Manuscript/manuscript.tex

@@
 On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-On the other hand, {\color{red} in general} \todo{Not CHARMM36 or amber14} atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
+On the other hand, in general atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
 For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
-{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}.
-\todo{remove?
-Why we need this?}}
+%{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}
 
-Recent studies within the NMRlipids project \url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
-A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept.
-Namely, the changes in NMR order parameters of the headgroup upon ions additions~\cite{seelig87} is directly compared to the MD simulations results.
+Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
+A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept~\cite{seelig87}.
+Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results.
 Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
@@
 Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}.
-Considering that RESP charges are usually refined to improve certain experimental observables, to some extent,  the electronic polarizability effect of the solvent is included in modern force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
+Considering that water molecules are often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
 Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$.
 Instead, a consistent scaling factor would lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
-It is important to note that the electronic part of polarizability is around 2 for almost any biologically relevant environment \cite{some_original_work, leontyev11}.
 
-Here we develop ECC-POPC lipid model that accurately describes the binding of sodium and calcium ions to the resulting lipid bilayer.
-We applied the ECC correction to the Lipid14 model of POPC by scaling of partial charges of the head group, glycerol backbone, and carbonyl regions.
-These are the polar parts of phospholipids which can contribute to the cation binding.
-The Lipid14~\cite{dickson14} force field parameters (available in a Gromacs format from Ref.
-\citenum{lipid14files}) were used as a starting point.
-This model provides the best response of the head group to ions among the available lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}).
-Additionally, the Lipid14 model has a relatively realistic head group, glycerol backbone, and tails structure~\cite{botan15}.
+Here we develop ECC-POPC lipid model that accurately describes the binding 
+of sodium and calcium ions to the lipid bilayer. 
+The Lipid14~\cite{dickson14} force field parameters 
+(available in a Gromacs format from Ref. \citenum{lipid14files}) were used as a starting 
+point. 
+This model provides the best response of the head group to ions among the available 
+lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
+has a relatively realistic head group, glycerol backbone, and acyl chain structure~\cite{dickson14,botan15}.
+We applied the ECC correction 
+to the Lipid14 model of POPC by scaling of 
+partial charges of the head group, glycerol 
+backbone, and carbonyl regions. 
+These are the polar parts of phospholipids which can 
+contribute to the cation binding.
 
-To develop the ECC-POPC model, we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$, to reproduce the experimental ion binding affinities.
-In general, scaling down the partial charges reduces the ion binding affinity and the head group order parameter response which can be measured using NMR~\cite{akutsu81, altenbach84, scherer89}.
-We found an optimal value $f_q = 0.8$, which is only slightly higher the theoretical one ($f_q=0.75$).
+To reproduce the experimental ion binding affinities,
+we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
+In general, scaling down the partial charges reduces the ion binding affinity 
+and the experimentally measurable head group order parameter response to the bound charge~\cite{akutsu81, altenbach84, scherer89}.
+We found an optimal value $f_q = 0.8$, which is only slightly higher than the theoretical one ($f_q=0.75$).
 Directly applying the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and the carbonyls reduces the area per lipid to 60~\AA$^2$.

### 2017-11-15 Manuscript/manuscript.tex
Updates after discussing with Hector

@@
 Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.
+\todo{Hector, please deliver the missing citations for this paragraph.}
 
@@
 For this, we apply a simple mean-field electronic continuum correction (ECC)~\cite{leontyev11} which implies scaling the partial charges of the atoms to account for the missing electronic polarizability in MD simulations~\cite{leontyev11}.
-Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014, kohagen14, kohagen16}.
+Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
 As a starting template for our POPC model, we will use the Lipid14~\cite{dickson14}.
@@
 Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor successfully improves the accuracy of simulations of solvated ions \cite{kohagen14, kohagen16, REF}, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+This scaling factor successfully improves the accuracy of simulations of solvated ions \cite{kohagen14, kohagen16, martinek17}, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
 It is important to note that the 
 high frequency dielectric constant 
-is around 2 for almost any biologically relevant environment \cite{some_original_work, leontyev11}.
+is around 2 for almost any biologically relevant environment \cite{leontyev11}.
 The dielectric discontinuity in a lipid bilayer arises
@@
 
-While using a scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{Leontyev2011}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
+While using a scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
 Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
 Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}.
-Considering that water molecules are often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
+Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
 Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$.
@@
 \subsubsection{Simulations of POPC bilayers with aqueous ions}
-Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules and approximately 50~SPC/E~water~\cite{Berendsen1987} molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
-The chosen SPC/E water model, also used in the previous parametrization of ECC-ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}, has a lowered dielectric constant, consistent with the ECC concept~\cite{leontyev11, leontyev14}.
-We tested also other water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}) to evaluate the robustness of our approach (See SI).
-Sodium, calcium, and chloride ions were modeled as ECC-ions with parameters from Ref.
-\citenum{jungwirth17-new-paper-to-be-published, Pluharova2014}.
-% cite order: [Ca_2s, Na_s, Cl_2s] 
-Scaled charges and van der Waals radii for atoms forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
-For comparison, simulations with unscaled charges were performed using the Lipid14 model~\cite{dickson14}, the TIP3p model for water \cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
-\todo{HECTOR: Do we compare this against scaled with SPC or TIP3P? SAMULI: Joe, please fill the information about Lipid14 simulation with ECC-ions.}
+Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
+and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
+The SPC/E~\cite{Berendsen1987} water model was used in all ECC-lipid model simulations reported
+in the main text. This model was also used in the previous parametrization
+of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} and its lowered dielectric constant is consistent with
+the ECC concept~\cite{leontyev11, leontyev14}.
+The robustness of our approach with other water models (OPC \cite{Izadi14},
+OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
+is tested in SI.
+Sodium, calcium, and chloride ions were modeled as ECC-ions
+with parameters from Ref. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennart-Jones radii for atoms
+forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
+For comparison, simulations with the Lipid14 model~\cite{dickson14} and
+ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
+were also performed. The TIP3p water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
+\todo{Joe, please confirm that this is correct.}
 Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
@@
 
-MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).
-The simulation settings employed in this work are summarized in Table \ref{tbl:mdpar}\todo{HECTOR: Is this the SI?
-If so we have to indicate it.
-Now reference is weird.}.
+MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
+in this work are summarized in Table~S1.
 The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
@@
     Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-lipids models compared with experiments~\cite{Kucerka2011}.
+    the ECC-lipids models compared with experiments~\cite{kucerka11}.
     Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
@@
   \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
-    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments.
-    { \color{red} HECTOR:Are we using here the same water model? If not is it fair the comparison considering the large changes observed in the SI table?
-      SAMULI: For Lipid14 simulations the water is TIP3P and SPC/E for ECC-lipids. I do understand why the comparison would
-    not be fair in this context?}\label{tab:apls} }
+    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments. \label{tab:apls}
+  }
+    %    { \color{red} HECTOR:Are we using here the same water model? If not is it fair the comparison considering the large changes observed in the SI table?
+    %      SAMULI: For Lipid14 simulations the water is TIP3P and SPC/E for ECC-lipids. I do understand why the comparison would
+    %    not be fair in this context?}
   \begin{tabular}{l|c c}
@@
     \hline
-    ECC-lipids                & {\color{red}63.2}$\pm$ 0.6  &  300       \\
+    ECC-lipids                & 63.2$\pm$ 0.6  &  300       \\
     \hline
@@
 is $\approx$1\AA{} smaller
-\todo{HECTOR: The 63.2 value does not appear in the SI table. There it is 62.2 but I'm not sure corresponds to the same system
-SAMULI: I can find the 63.2 from the SI table with SPE/E and the value 62.2 with OPC3, so everything seems to be ok for me.}.
+%\todo{HECTOR: The 63.2 value does not appear in the SI table. There it is 62.2 but I'm not sure corresponds to the same system
+%SAMULI: I can find the 63.2 from the SI table with SPE/E and the value 62.2 with OPC3, so everything seems to be ok for me.}.
 The values of the area per lipid of the ECC-lipids model show larger variations
-when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~\ref{tab:apls_si} in SI\todo{HECTOR: change to real reference in SI}),
+when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI,
 while still being close to the experimentally reported values.
 We can thus conclude that the ECC-lipids model reproduces the experimental dimensions of the POPC
-lipid bilayer with accuracy comparable to that for nother state of the art lipid models \cite{ollila16}.
+lipid bilayer with accuracy comparable to that for nother state of the art lipid models~\cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 
-Changes of the lipid bilayer head group order parameters from different simulations and experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl} and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
-%Responses to NaCl concentrations are in SI, Fig.~S\ref{fig:delta_ordPar_NaCl}.
-These results can be used to compare the ion binding affinities to lipid bilayers between simulations and experiments using the electrometer concept because the order parameters decrease proportionally to the amount of the bound positive charge (Fig.~\ref{OrderParameterCHANGESsurf} and Ref.~\citenum{seelig87, catte16}).
-A recent comparison of different simulation models to the experimental data revealed that most force fields significantly overestimate Na$^+$ ion binding to PC lipid bilayers and that none of the available models correctly reproduces the details of the binding of Ca$^{2+}$ ions which is grossly overestimated \cite{catte16}.
-Only the Lipid14 model \cite{dickson14} simulated with \AA{}qvist ions reproduces the experimentally measured small order parameter changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81, altenbach84}.
-Since artifacts were reported in simulations involving \AA{}qvist ions is water \cite{auffinger07}, we tested the Lipid14 model with more realistic Na$^+$ models by Dang et al.~\cite{smith94, chang1999, dang2006}.
-In this case, overbinding reappeared (Fig.~\ref{fig:delta_ordPar_NaCl}).The results are in line with the previous work \cite{catte16}, suggesting that additional improvements in the lipid parameters are required to describe the binding of cations to phospholipid bilayers correctly.
+Changes of the lipid bilayer head group order parameters from different simulations and
+experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
+and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
+As seen in Fig.~\ref{OrderParameterCHANGESsurf}, the order parameters decrease
+proportionally to the amount of the bound positive charge.
+These results can be thus used to compare the ion binding affinities to lipid bilayers between
+simulations and experiments using the electrometer concept~\cite{seelig87, catte16}.
+
+The experimentally measured small order parameter
+changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) %and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81, altenbach84}.
+are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
+However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
+%
+%A recent comparison of different simulation
+%models to the experimental data revealed that most force fields significantly
+%overestimate Na$^+$ ion binding to PC lipid bilayers and that none of the available
+%models correctly reproduces the details of the binding of Ca$^{2+}$ ions which is
+%grossly overestimated \cite{catte16}.
+%
+%Since artifacts were reported
+%in simulations involving \AA{}qvist ions is water \cite{auffinger07}, we tested the
+%
+Replacing \AA{}qvist ions with models by Dang et al.~\cite{smith94, chang1999, dang2006}
+or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
+the results~(Figs.~\ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}).
+In line with the previous work \cite{catte16}, the results suggest that improvements
+in the lipid parameters are required to describe the binding of cations to phospholipid bilayers correctly.
 
@@
 
-The results from the simulations combining the ECC-lipids with the ECC-ion models \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
+The results from the simulations combining the ECC-lipids with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
 Considering that we can also reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC-model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
 Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-lipids model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-
-\todo{HECTOR: I do not believe in the content of the following sentence as it is.
+%\todo{HECTOR: I do not believe in the content of the following sentence as it is.
 How do we know that the changes of the P-N angle are too pronounced for Lipid14?.
 There is no direct correlation between angle and order parameter.}
-Also the changes of the P-N vector angle are too pronounced for the Lipid14 model, for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$ CaCl$_2$ concentration is approximately 17$^{\circ}$, while the corresponding value for ECC-lipids model is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
-\todo{HECTOR: I have a problem when comparing fig.
+% SAMULI: The sentence is now modified.
+Also the changes of the P-N vector angle are more pronounced for the Lipid14 model,
+for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
+CaCl$_2$ concentration is approximately 17$^{\circ}$. The corresponding value
+for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$). 
+%\todo{HECTOR: I have a problem when comparing fig.
 2 and 3.
@@
 Given our previous work, should we check the behavior respect the charge density?}
+%SAMULI: Not now, hopefully in the future.
 
@@
 In general, the performance of ECC-lipids with any of the water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3p model.
-
-%\todo{SAMULI: Maybe we should discuss the repeat distances and
-%  area per molecules measured at \cite{petrache06,pabst07,uhrikova08}
-%  }
 
@@
 I will make an effort to make them more clear.
-}
+SAMULI: I agree with the titles, especially the word ''structure'' here. Can you suggest new ones?}
 
@@
 \todo{HECTOR: I have deleted the sentence relating the uncertainty and the error as in my opinion there are other factors which are more important.
-The most important the slow convergence.}.
+The most important the slow convergence.
+SAMULI: I think that this was a good idea.}.
 Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our POPC-ECC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having a negative value $\Gamma_{Na}^{water} = (-0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}).
@@
 This is in contradiction with most of the available lipid force fields, which predict a positive binding of sodium to PC lipid bilayers \cite{catte16}.
-\todo{do we need to add a NaCl table or an extended table for CaCl2 using different concentrations in the SI?
-I think this is very important information}
-
-%using the bulk concentrations determined from the density profiles in Fig.~\ref{fig:cacl-dens}. 
-%calculated using density profiles in Fig.~\ref{fig:cacl-dens}
-
-%\todo{JOE: Add references to NaCl simulation papers that show peaks.
-%Add a simple note that there is NBFIX trick in Charmm36 (?).} 
-%SAMULI: I think that it is enough to cite NMRlipids II here, which shows all the results
-%
-%This finding is depicted in Figures~\ref{fig:nacl-dens} and~\ref{fig:cacl-dens},
-%which show the density profile of ions and water along the membrane normal. 
-%
-%The density profiles show
-%a larger Ca$^{2+}$ density peak in lipid headgroup region for
-%the Lipid14 model with Dang and ECC-ions than for the ECC-lipids model.
+%\todo{HECTOR: do we need to add a NaCl table or an extended table for CaCl2 using different concentrations in the SI? I think this is very important information
+%  SAMULI: I agree that this is important information, but we have quite a lot of stuff in the paper already. Maybe we could leave this for the future studies.}
 
@@
 Sodium ions also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
-\todo{This last sentence is weak and not relevant.
+%\todo{This last sentence is weak and not relevant.
 If you want to keep it go ahead, but then I would ask you a simple question: Is it incompatible to have no overall binding (this relates to the whole surface volume) with having small cavities in the membrane with an increase Na+ concentration.
@@
 However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2}low concentrations~\cite{macdonald87}.
-\todo{HECTOR: Why is the following sentence remotely significant.
+%\todo{HECTOR: Why is the following sentence remotely significant.
 I will vote for removing it} Note that the results based on these binding models and the measured headgroup order parameter changes are also included in the tabulated experimental binding constants~\cite{marsh13}.
@@
 Using our simulations, we have direct access to exact details of the binding stoichiometry without requiring any middle model as in experiments~\cite{altenbach84, macdonald87}.
-To evaluate the relative propensities for each stoichiometric complex (e.i., 1 Ca$^{2+}$: n POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
-In Fig.~\ref{fig:cacl_complexes}, from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ we found that the largest propensity is for the 1:2 complex (41\%).
-\todo {I seriously doubt that this is what they call a ternary complex but correct me if you know better}.
+To evaluate the relative propensities for each stoichiometric complex (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
+Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
+We found that the largest propensity is for the 1:2 complex (41\%).
 Still, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower.
@@
 This suggests that at least several microseconds are required to reach the binding/unbinding equilibrium.
-\todo{HECTOR: I disagree with the following lines.
+%\todo{HECTOR: I disagree with the following lines.
 In the case of most of the unscale force fields the issue is that the binding of Calcium to the membrane is very strong.
@@
 I might be wrong in this last part.
-If you do not agree we can discuss it tomorrow via skype.} {\color {red} This is in line with the density profiles of \ce{Ca^{2+}} from simulations in Fig.~\ref{fig:cacl-dens}, which all show a region of a very small calcium density around the level of choline groups in POPC (i.e., around $2.5\,\mathrm{nm}$).
+If you do not agree we can discuss it tomorrow via skype.} {\color{red} This is in line with the density profiles of \ce{Ca^{2+}} from simulations in Fig.~\ref{fig:cacl-dens}, which all show a region of a very small calcium density around the level of choline groups in POPC (i.e., around $2.5\,\mathrm{nm}$).
 In contrast, simulations with  ECC-lipids exhibits only a relatively shallow local minimum at that level.
-This suggests that the smaller barrier for the exchange of calcium ions shall lead to their faster exchange with the solvent within ECC-lipids.} To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated residence times of ions bound to the membrane.
+This suggests that the smaller barrier for the exchange of calcium ions shall lead to their faster exchange with the solvent within ECC-lipids.}
+To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated residence times of ions bound to the membrane.
 In our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
@@
 I will discuss this with Joe tomorrow.
-Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_w$ = 450~nM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Ref.~\citenum{javanainen17}) are shown in Fig.~S\ref{fig:hist_residence_times} in SI.
+Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_w$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S\ref{fig:hist_residence_times} in SI.
 In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory ({\color{red} 800~ns}).
@@
 
-\newpage
-\newpage
-\appendix
-
-\begin{center}
-{\bf SUPPLEMENTARY INFORMATION}
-\end{center}
-
-It was found in the original work~\cite{altenbach84} that 
-a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
-provides the best fit to experimental measurements of all considered models in that study. 
-In such a model, there is a linear relationship between quantities 
-$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
-where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
-The concentration $C_b$ was obtained from an extrapolation of linear relation 
-between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
-Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
-remains constant throughout the extrapolation range. 
-The concentration $C_I$ is determined from the surface potential using the Boltzmann equation.
-However, Boltzmann theory yields inaccurate results
-for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
-An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions;
-$C_b$ can be calculated from simulation using the robust definition in section~\ref{sec:affinity},
-and $C_I$ can be obtained from the density profiles in Fig.~\ref{fig:cacl-density} 
-as the minimum concentration at the shallow depression separating the bulk water phase 
-and the peak of adsorbed cations around 2.6\,nm from the membrene centre. 
-The small discrepancy between the results in the experiment~\cite{altenbach84} and 
-our simulations are likely caused by the difference in the evaluation of the concentrations.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
-  \caption{\label{fig:cacl-bind}
-    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
-    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
-    provides a good fit to experimental measurements~\cite{altenbach84}
-    and it also provides a good fit to our simulation data. 
-    This supports our statement that the success of the ternary complex model in the experiments
-    can be understood with the atomistic detail of our simulations as an average 
-    stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
-    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
-    and, hence, the observed slope of the linear relationship is slightly different.
-    }
-\end{figure}
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
-  \caption{\label{fig:ordPars_waterModels}
-    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
-    are shown from simulations with different force fields and water models together with experimental data 
-    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
-    Ion concentrations in bulk water are shown in x-axis. 
-    Values from simulations are calculated from the of cation number density $C_{np}$
-    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
-  }
-\end{figure}
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_NaCl.pdf}
-  \caption{\label{fig:delta_ordPar_NaCl_si}
-    Changes of head group order parameters of POPC bilayer as a function of NaCl concentrations
-    are shown from simulations with different force fields together with experimental data \cite{akutsu81}. 
-    Ion concentrations in bulk water are shown in x-axis. 
-    Values from simulations are calculated from the of cation number density $C_{np}$
-    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
-  }
-\end{figure}
-
-\begin{figure}[tbp!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_ECC-lipids_346mM_CaCl.pdf} \\
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_Charmm36_450mM_CaCl_Matti.pdf}
-  \caption{\label{fig:hist_residence_times}
-   Histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer 
-   at $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
-   from simulations with ECC-lipids/ECC-ions (top)
-   and CHARMM36/ECC-ions (bottom), which was taken from previous study \cite{javanainen17} 
-   (simulation data available at Ref.~\citenum{zenodo.259376}).
-   The maxima on the x-axis represent the lengths of the simulations used for analysis. 
-   For the ECC-lipids model, 90\% of the residence times of \ce{Ca^{2+}} in a POPC membrane are 
-   shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
-   with the longest observed residence time being $141\,\mathrm{ns}$, 
-   which is well below the total length of the simulation (200~ns).
-   This is, however not true for the simulation with CHARMM36 model,
-   where there are several calcium cations with their residence time 
-   apparently limited by the length of the simulation. 
-   With such a simulation, which is relatively short for the employed model,
-   we can merely estimate an upper bound that 
-   only less than 60\% of the bound residence time 
-   is formed by cations interacting with the phospholipid membrane 
-   for less then half of the simulation length~(400~ns).
-  }
-\end{figure}
-
-\begin{table}[tb!]
-  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
-  \begin{tabular}{l|c c}
-    model          & APL (\AA$^2$)   & Temperature [K] \\
-    \hline
-    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
-    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
-    \hline
-    ECC-lipids &        &  \\
-    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
-    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
-    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
-    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
-    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
-    ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
-    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
-    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
-    ~ TIP3p          & 66.2$\pm$ 0.6   &  313       \\
-    ~ TIP3p-FB       & 64.8$\pm$ 0.6   &  313       \\
-    ~ TIP4p-FB       & 65.6$\pm$ 0.6   &  313       \\
-    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
-    %oMM small patch           & 63.65  &         \\
-    %oMM 4xbig patch           & 63.7   &         \\
-    \hline
-    experiment   & 62.7  &  293    \\
-    experiment \cite{kucerka11} & 64.3  &  303    \\
-    experiment  & 67.3  &  323    \\
-    experiment  & 68.1  &  333    \\
-    %experiment POPE  & 56.6 &  303    \\
-    \hline
-  \end{tabular} \\
-\end{table}
-
-\begin{table}[btp!]
-  \caption{Simulation parameters}
-  \label{tbl:mdpar}
-  \begin{tabular}{ll}
-    simulation property & parameter   \\
-    \hline
-    time-step           & 2~fs         \\
-    equilibration time  & 100~ns  \\
-    total simulation time     & 300~ns  \\
-    temperature         & 313~K       \\
-    thermostat          & v-rescale  \cite{bussi07}   \\
-    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
-    long-range electrostatics & PME  \cite{darden93}  \\
-    cut-off scheme      & Verlet \cite{Pall13}      \\
-    Coulomb and VdW cut-off & 1.0~nm \\
-    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
-    constraints for water & SETTLE  \cite{miyamoto92} \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{figure}[tbp!]
-  \centering
-  \includegraphics[width=16.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89_GMX-oMM_compar.pdf}
-  \caption{\label{fig:ordPars_actual_GMX_oMM_compar}
-    Order parameters of POPC head group, glycerol backbone and acyl chains 
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
-    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
-    compared with the experimental values from \cite{ferreira13}.
-    The size of the markers for the head group order parameters correspond to
-    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
-    while the error estimate for simulations is $\pm 0.005$.
-    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
-  }
-\end{figure}
-
-\begin{figure}[tbp!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_surf_GMX-oMM_compar.pdf}
-  \caption{\label{fig:ordPars_surf_GMX_oMM_compar}
-    The changes of headgroup order parameters and P-N vector orientation as a function of
-    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
-    from simulations with the ECC-lipids models
-    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
-    compared with the experimental values from \cite{scherer89}.
-    The size of the markers for the head group order parameters correspond to
-    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
-    while the error estimate for simulations is $\pm 0.005$.
-  }
-\end{figure}
-
-\begin{figure}[tbp!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_GMX-oMM_compar.pdf}
-  \caption{\label{fig:ordPars_cacl_GMX_oMM_compar}
-    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
-    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
-    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    Ion concentrations in bulk water are shown in x-axis. 
-    Bulk concentrations from simulations are calculated 
-    from the farthest point from the lipid bilayer in the aqueous phase
-    with an error estimate of 10\,mM.
-  }
-\end{figure}
+%\newpage
+%\appendix
 
@@
 \bibliography{refs.bib}
-
-%\newpage
-%\section{APPENDIX: The NMR results reported by Tiago Ferreira}
 

### 2017-11-15 Manuscript/manuscript.tex
Updates

@@
 Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
-bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens 
-, which are in agreement with the previous interpretations of the experimental spectroscopic data.
+bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens,
+in agreement with the previous interpretations of the experimental spectroscopic data.
 On the other hand, the new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
@@
 On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-On the other hand, in general atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
+On the other hand, most atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
 For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
@@
 In this work, we improve the cation binding of a POPC bilayer in MD simulations via the implicit inclusion of the electronic polarizability in the polar region of phospholipids.
-For this, we apply a simple mean-field electronic continuum correction (ECC)~\cite{leontyev11} which implies scaling the partial charges of the atoms to account for the missing electronic polarizability in MD simulations~\cite{leontyev11}.
+For this, we apply a simple mean-field electronic continuum correction (ECC)~\cite{leontyev11} which implies scaling the partial charges of the atoms to account for the missing electronic polarizability in MD simulations.
 Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
@@
 Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor successfully improves the accuracy of simulations of solvated ions \cite{kohagen14, kohagen16, martinek17}, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.
+This scaling factor successfully improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
 It is important to note that the 
@@
 we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
-In general, scaling down the partial charges reduces the ion binding affinity 
-and the experimentally measurable head group order parameter response to the bound charge~\cite{akutsu81, altenbach84, scherer89}.
-We found an optimal value $f_q = 0.8$, which is only slightly higher than the theoretical one ($f_q=0.75$).
-Directly applying the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and the carbonyls reduces the area per lipid to 60~\AA$^2$.
-This area is smaller than the original in Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and the experimental value (64.3~\AA$^2$) \cite{kucerka11}.
-The decrease of the area per lipid arises from a reduced hydration of the lipid head group region after scaling charges which effectively reduces head group polarity.
-We solve this difficulty by reducing the effective radii of the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a factor of $f_\sigma = 0.89$.
-The same solution was taken for the ECC-ions in aqueous solvent \cite{kohagen14, kohagen16, Pluharova2014}.
-After reducing the $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
+The ion binding affinity was evaluated against experiments
+using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
+as discussed more detail in the next section.
+In general, scaling down the partial charges reduced the ion binding affinity.
+We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
+which is only slightly higher than the theoretical one ($f_q=0.75$).
+Directly applying
+the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
+the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
+original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
+(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
+reduced hydration of the lipid head group region after scaling charges which effectively
+reduces head group polarity. We solve this difficulty by reducing the effective radii of
+the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a
+factor of $f_\sigma = 0.89$. The same solution was taken for the ECC-ions in aqueous
+solvent \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the $\sigma$ parameters,
+the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
 
 \subsection{Electrometer concept} \label{section:electrometer}
-Comparing MD simulation to NMR experiments, we can validate the simulation ion binding affinity to lipid bilayers using the ''electrometer concept''~ \cite{seelig87, catte16}.
+Comparing MD simulation to NMR experiments, we can validate the ion
+binding affinity in lipid bilayer simulations using the ''electrometer concept''~ \cite{seelig87, catte16}.
 This method relies on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
-Comparing this response is an excellent way to validate force fields as both NMR and MD can accurately determine order parameters ~\cite{catte16, ollila16}.
+%Comparing this response is an excellent way to validate force fields as both NMR and MD can accurately determine order parameters ~\cite{catte16, ollila16}.
 The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
@@
 
-The measured change of the order parameter depends both on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).
+The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).
 The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities.
@@
 NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
-Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81} while others use atomic absorption spectroscopy and report the salt concentration in the supernatant after the solvation of lipids~\cite{altenbach84}.
-In this work, we use the latter definition.
-In particular, we measure the salt concentration in the aqueous bulk region using the farthest point from the lipid bilayer in the aqueous phase.
-Notice that this definition differs from the one used by Catte et al.~\cite{catte16} where the former definition is used.
-Although these two definitions vary when applied to Ca$^{2+}$, they do not significantly affect the conclusions of this work.
+Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_w$,
+while others use atomic absorption spectroscopy and report the salt concentration in
+the supernatant after the solvation of lipids~\cite{altenbach84}, $C_b$. In this work,
+%we use the latter definition. In particular,
+we measure the latter concentration in the aqueous
+bulk region using the farthest point from the lipid bilayer in the aqueous phase.
+Notice that former concentration was used by Catte et al.~\cite{catte16}.
+Although these two definitions vary when applied
+to Ca$^{2+}$, they do not significantly affect the conclusions of this work.
 
@@
 Sodium, calcium, and chloride ions were modeled as ECC-ions
-with parameters from Ref. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennart-Jones radii for atoms
+with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennart-Jones radii for atoms
 forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
@@
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
-in this work are summarized in Table~S1.
-The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
+in this work are summarized in Table~S1. Combability with openMM simulation package \cite{openmm7} was also
+tested in SI. The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
 
@@
 Sufficient lipid neighbor exchange occurred during the simulations.
-Further details for the unscaled simulations and the trajectories are in Ref.~\citenum{catte16}.
+The trajectories and simulation files for unscaled systems are available from
+Refs.~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
 
@@
     Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-lipids models compared with experiments~\cite{kucerka11}.
+    the ECC-lipids models compared with experiments~\cite{kucerka11} at {\color{red}??K}.
     Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
     from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
-    compared with the experimental values from \cite{ferreira13}.
+    compared with experiments \cite{ferreira13} at {\color{red}??K}.
     The size of the markers for the head group order parameters correspond to
@@
     Bottom: The chemical structure of POPC and the labeling of the carbon segments.
+    {\color{red}Joe, please add the temperatures}
   } 
@@
 We can thus conclude that the ECC-lipids model reproduces the experimental dimensions of the POPC
-lipid bilayer with accuracy comparable to that for nother state of the art lipid models~\cite{ollila16}.
+lipid bilayer with the comparable accuracy to other state of the art lipid models~\cite{ollila16}.
 %The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
@@
 This feature has been suggested to indicate that the carbonyl of {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in {\it sn}-1 chain, which would orient more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
-This arrangement may be a relevant feature for the ion binding details, which it is not fully reproduced by other available lipid models~\cite{ollila16}.
+While this arrangement may be a relevant feature for the ion binding details, it is not fully reproduced by other available lipid models~\cite{ollila16}.
 %This is expected as the ECC model does not modify the already highly optimized parameters of this region in Lipid14
@@
     a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
-    from simulations and experiments \cite{scherer89}.
+    from simulations and experiments \cite{scherer89} at 313 K.
   }
@@
 
-Before studying the sodium and calcium ion binding affinities, we have to quantify the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
+Before studying the sodium and calcium ion binding affinities, we quantify the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
 These mixtures have a well-defined amount of bound charge per PC, i.e., the molar fraction of cationic surfactants.
@@
 The lower sensitivity of the P-N vector angle response in the ECC-POPC model agrees better with experiments.
-The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules.
+%The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules.
 %{\color{red}., as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
@@
     Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
-    from simulations with different force fields together with 
+    in bulk ($C_b$) from simulations with different force fields at 313 K together with 
     experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
-    Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
+    %Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
     Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
@@
     Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration
-    are shown from simulations with different force fields together with experimental data 
+    as a function of the CaCl$_2$ concentration in bulk ($C_b$)
+    from simulations at 313 K together with experimental data 
     (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    Ion concentrations in bulk water are shown in x-axis. 
-    Bulk concentrations from simulations are calculated 
-    from the farthest point from the lipid bilayer in the aqueous phase
-    with an error estimate of 10\,mM.
-    The error estimate for the simulation with Lipid14 and ECC-ions is an order of magnitude larger, 
-    because the simulation cell was too small (shown if Fig.~\ref{fig:cacl-dens}). 
-    Simulation data with Lipid14 and \AA{}qvist ion parameters are taken directly from
+    %Ion concentrations in bulk water are shown in x-axis. 
+    %Bulk concentrations from simulations are calculated 
+    %from the farthest point from the lipid bilayer in the aqueous phase
+    The error estimate for bulk concentrations was approximately 10\,mM.
+    Exception was the order of magnitude larger error in
+    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) in too 
+    small simulation box. 
+    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
     Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
@@
 %
-Replacing \AA{}qvist ions with models by Dang et al.~\cite{smith94, chang1999, dang2006}
+Replacing \AA{}qvist ions with ions by Dang et al.~\cite{smith94, chang1999, dang2006}
 or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
@@
 In line with the previous work \cite{catte16}, the results suggest that improvements
-in the lipid parameters are required to describe the binding of cations to phospholipid bilayers correctly.
+in the lipid parameters are required to correctly describe the binding of cations to phospholipid bilayers.
 

### 2017-11-16 Manuscript/SI.tex

\documentclass[journal=jpcbfk]{achemso}

\usepackage[version=3]{mhchem}
\usepackage[T1]{fontenc}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}

\author{Josef Melcr}
\author{Hector Martinez-Seara}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Ji{\v r}{\' i} Kolafa}
\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
\author{Pavel Jungwirth}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
Tampere, Finland}

\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}

\SectionNumbersOn

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thepage}{S\arabic{page}}%

\title{Supporting Information:\\Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization}

\begin{document}

\newpage
%\tableofcontents

\section{Simulation details}

\begin{table}[!h]
  \caption{Simulation parameters}
  \label{tbl:mdpar}
  \begin{tabular}{ll}
    simulation property & parameter   \\
    \hline
    time-step           & 2~fs         \\
    equilibration time  & 100~ns  \\
    total simulation time     & 300~ns  \\
    temperature         & 313~K       \\
    thermostat          & v-rescale  \cite{bussi07}   \\
    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
    long-range electrostatics & PME  \cite{darden93}  \\
    cut-off scheme      & Verlet \cite{Pall13}      \\
    Coulomb and VdW cut-off & 1.0~nm \\
    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
    constraints for water & SETTLE  \cite{miyamoto92} \\
    \hline
  \end{tabular}
\end{table}

\section{Area per molecule and calcium binding with different water models}

\begin{table}[!h]
  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-lipids &        &  \\
    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
    ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
    ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
    ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
    ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
    ~ TIP4p/2005     & 66.8$\pm$ 0.6   &  313       \\
    ~ TIP3p          & 66.2$\pm$ 0.6   &  313       \\
    ~ TIP3p-FB       & 64.8$\pm$ 0.6   &  313       \\
    ~ TIP4p-FB       & 65.6$\pm$ 0.6   &  313       \\
    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
    %oMM small patch           & 63.65  &         \\
    %oMM 4xbig patch           & 63.7   &         \\
    \hline
    experiment   & 62.7  &  293    \\
    experiment \cite{kucerka11} & 64.3  &  303    \\
    experiment  & 67.3  &  323    \\
    experiment  & 68.1  &  333    \\
    %experiment POPE  & 56.6 &  303    \\
    \hline
  \end{tabular} \\
\end{table}

\begin{figure}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
  \caption{\label{fig:ordPars_waterModels}
    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
    are shown from simulations with different force fields and water models together with experimental data 
    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Values from simulations are calculated from the of cation number density $C_{np}$
    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
  }
\end{figure}

\section{Sodium binding to POPC with the glycerol carbon and changes in the head group angle}

\begin{figure}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_NaCl.pdf}
  \caption{\label{fig:delta_ordPar_NaCl_si}
    Changes of head group and glycerol $g_3$ order parameters and the P-N vector angle of POPC bilayer as a function of NaCl concentrations
    from simulations with different force fields together with experimental data \cite{akutsu81}. 
    Ion concentrations in bulk water are shown in x-axis. 
%    Values from simulations are calculated from the of cation number density $C_{np}$
%    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
  }
\end{figure}

\section{Histograms of residence times}

\begin{figure}[!h]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_ECC-lipids_346mM_CaCl.pdf} \\
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/histogram_bound_times_Charmm36_450mM_CaCl_Matti.pdf}
  \caption{\label{fig:hist_residence_times}
   Histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer 
   at $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
   from simulations with ECC-lipids/ECC-ions (top)
   and CHARMM36/ECC-ions (bottom), which was taken from previous study \cite{javanainen17} 
   (simulation data available at Ref.~\citenum{zenodo.259376}).
   The maxima on the x-axis represent the lengths of the simulations used for analysis. 
   For the ECC-lipids model, 90\% of the residence times of \ce{Ca^{2+}} in a POPC membrane are 
   shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$
   with the longest observed residence time being $141\,\mathrm{ns}$, 
   which is well below the total length of the simulation (200~ns).
   This is, however not true for the simulation with CHARMM36 model,
   where there are several calcium cations with their residence time 
   apparently limited by the length of the simulation. 
   With such a simulation, which is relatively short for the employed model,
   we can merely estimate an upper bound that 
   only less than 60\% of the bound residence time 
   is formed by cations interacting with the phospholipid membrane 
   for less then half of the simulation length~(400~ns).
  }
\end{figure}

\section{Ternary complex model in simulations}
It was found in the original work~\cite{altenbach84} that 
a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
provides the best fit to experimental measurements of all considered models in that study. 
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined from the surface potential using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions;
$C_b$ can be calculated from simulation using the robust definition in section~\ref{sec:affinity},
and $C_I$ can be obtained from the density profiles in Fig.~\ref{fig:cacl-density} 
as the minimum concentration at the shallow depression separating the bulk water phase 
and the peak of adsorbed cations around 2.6\,nm from the membrene centre. 
The small discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations are likely caused by the difference in the evaluation of the concentrations.

\begin{figure}[!h]
  \centering
  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
  \caption{\label{fig:cacl-bind}
    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
    provides a good fit to experimental measurements~\cite{altenbach84}
    and it also provides a good fit to our simulation data. 
    This supports our statement that the success of the ternary complex model in the experiments
    can be understood with the atomistic detail of our simulations as an average 
    stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
    and, hence, the observed slope of the linear relationship is slightly different.
    }
\end{figure}

\section{Comparison between Gromacs and openMM}

\begin{figure}[!h]
  \centering
  \includegraphics[width=16.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89_GMX-oMM_compar.pdf}
  \caption{\label{fig:ordPars_actual_GMX_oMM_compar}
    Order parameters of POPC head group, glycerol backbone and acyl chains 
    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
    compared with the experimental values from \cite{ferreira13}.
    The size of the markers for the head group order parameters correspond to
    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
    while the error estimate for simulations is $\pm 0.005$.
    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
  }
\end{figure}

\begin{figure}[!p]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_surf_GMX-oMM_compar.pdf}
  \caption{\label{fig:ordPars_surf_GMX_oMM_compar}
    The changes of headgroup order parameters and P-N vector orientation as a function of
    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
    from simulations with the ECC-lipids models
    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
    compared with the experimental values from \cite{scherer89}.
    The size of the markers for the head group order parameters correspond to
    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
    while the error estimate for simulations is $\pm 0.005$.
  }
\end{figure}

\begin{figure}[!p]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_GMX-oMM_compar.pdf}
  \caption{\label{fig:ordPars_cacl_GMX_oMM_compar}
    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
    as a function of the CaCl$_2$ concentration
    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
    Ion concentrations in bulk water are shown in x-axis. 
    Bulk concentrations from simulations are calculated 
    from the farthest point from the lipid bilayer in the aqueous phase
    with an error estimate of 10\,mM.
  }
\end{figure}

\newpage
\bibliography{refs.bib}

\end{document}

### 2017-11-16 Manuscript/manuscript.tex

@@
 This force field provides one of the best available descriptions of cation binding \cite{catte16}.
-The newly developed ECC-lipids model reproduces the experimentally
+The newly developed ECC-POPC model reproduces the experimentally
 measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while improving the membrane binding affinities significantly to sodium and calcium cations.
@@
 
-\subsection{POPC membrane structure and dynamics}
+\subsection{Structural parameters of pure ECC-POPC model membrane are in agreement with experiments}
 
@@
     Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-lipids models compared with experiments~\cite{kucerka11} at {\color{red}??K}.
+    the ECC-POPC models compared with experiments~\cite{kucerka11} at {\color{red}??K}.
     Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
+    from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
     compared with experiments \cite{ferreira13} at {\color{red}??K}.
@@
   \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
-    ran in this work and from the literature \cite{dickson14}, from the ECC-lipids model, and from experiments. \label{tab:apls}
+    ran in this work and from the literature \cite{dickson14}, from the ECC-POPC model, and from experiments. \label{tab:apls}
   }
@@
     \hline
-    ECC-lipids                & 63.2$\pm$ 0.6  &  300       \\
+    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
     \hline
@@
 First, we present results for bilayers in pure water.
-The ECC-lipids and Lipid14 models both reproduce the experimental x-ray scattering form factors
+The ECC-POPC and Lipid14 models both reproduce the experimental x-ray scattering form factors
 of a POPC bilayer with a comparable accuracy in Fig.~\ref{simVSexpNOions}.
 The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
-experimental value in Table~\ref{tab:apls}, while the value from the ECC-lipids model
+experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
 is $\approx$1\AA{} smaller
@@
 %SAMULI: I can find the 63.2 from the SI table with SPE/E and the value 62.2 with OPC3, so everything seems to be ok for me.}.
-The values of the area per lipid of the ECC-lipids model show larger variations
-when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI,
+The values of the area per lipid of the ECC-POPC model show larger variations
+when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI),
 while still being close to the experimentally reported values.
-We can thus conclude that the ECC-lipids model reproduces the experimental dimensions of the POPC
+We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
 lipid bilayer with the comparable accuracy to other state of the art lipid models~\cite{ollila16}.
@@
 
-Similarly, the acyl chain order parameters of the ECC-lipids model, as well as the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
+Similarly, the acyl chain order parameters of the ECC-POPC model, as well as the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
 Notably, the experimentally measured forking and small order parameter values of $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
@@
 
-The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger in the ECC-lipids model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-lipids model, see Fig.~\ref{OrderParameterCHANGESsurf}.
+The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-POPC model, see Fig.~\ref{OrderParameterCHANGESsurf}.
 Considering the available experimental evidence, it is not possible to suggest which of the two models provides more realistic headgroup orientation.
-The ECC-lipids model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
+The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
 The accuracy of both models in the glycerol backbone region is equally footed with any other state of art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
@@
 
-\subsection{Calibration of lipid electrometer:
-            Response of POPC head groups to bound charge}\label{section:boundCHARGE}
+\subsection{Calibration of head group response to membrane-bound charge using cationic surfactant}\label{section:boundCHARGE}
 
@@
 The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters is too large to the bound positive charge.
-In contrast, the slope of the ECC-lipids model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.
+In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.
 
@@
 
-\subsection{Binding affinities to POPC membrane validated through lipid electrometer}
+\subsection{Validation ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ Cations: the electrometer concept}
 
@@
 
-The results from the simulations combining the ECC-lipids with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
+The results from the simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
 Considering that we can also reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC-model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
-Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-lipids model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
+Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
 %\todo{HECTOR: I do not believe in the content of the following sentence as it is.
@@
 
-Finally, the ion binding affinities for the ECC-lipids model with different water models are compared in SI.
-In general, the performance of ECC-lipids with any of the water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3p model.
+Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI.
+In general, the performance of ECC-POPC with any of the water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3p model.
 
-\subsection{Structure and affinity of cation binding to POPC membrane}
+\subsection{Binding affinity of Na$^+$ and Ca$^{2+}$ cations to POPC membrane}
 \label{sec:affinity}
 
-\todo{HECTOR: I'm not sure I understand the meaning of "structure" in the title.
+%\todo{HECTOR: I'm not sure I understand the meaning of "structure" in the title.
 In general the titles in the result section are much alike between them, specially the first ones.
-This makes difficult to guess where to find a particular piece of information.
+This makes %difficult to guess where to find a particular piece of information.
 I will make an effort to make them more clear.
-SAMULI: I agree with the titles, especially the word ''structure'' here. Can you suggest new ones?}
+%SAMULI: I agree. There are now new titles.}
 
 \begin{table}[tb!]
-  \caption{Relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$) in different POPC bilayer models.
-All systems contain initially 350~mM of Ca$^{2+}$ (C$_w$).
-C$_b$ correspond to the end water bulk concentration of Ca$^{2+}$.
-$r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$ show the location for bound calcium in percentages for phosphate and carbonyl oxygens.
-{\color{red}HECTOR:PLEASE CHECK}
+  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
+    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
+    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ respect to water (C$_w$=350mM).
+    %{\color{red}HECTOR:PLEASE CHECK}
+    {\color{red}Joe, please confirm that $r^\mathrm{Ca^{2+}} _\mathrm{PO_4}$ is really 100\% for all the Lipid14 simulations
+      (i.e. that this is not a typo).}
   \label{tab:binding}}
-  \begin{tabular}{l|c | c c | c c}
-    model                 & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$ & $C_{w}$ & $C_b\,/\,\mathrm{mM}$   & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+  \begin{tabular}{l|c c | c | c c}
+    model                  & $C_{w}$ & $C_b\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
     \hline
-    ECC-lipids            &  $0.06 \pm 0.01 $ &  350  &  $280\pm 10 $  &  99\%  &    25\%    \\
-    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $ &  350  &  $210\pm 10 $  & 100\%  &    37\%     \\
-    Lipid14/Dang          &  $0.23 \pm 0.03 $ &  350  &  $160\pm 10 $  & 100\%  &    14\%    \\
-    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $ &  350  &  $120\pm 100$  & 100\%  &    23\%    \\
+    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    25\%    \\
+    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
+    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
+    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
 %Average no. of bound cations:
@@
 Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
-The values of $\Gamma_{\rm ion}^{\rm water}$, shown in Table~\ref{tab:binding}, were calculated from simulations with the same initial concentration of CaCl$_2$ of 350~mM before solvating the lipids.
-As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess for the ECC-lipids model, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller than for the other models, 0.13--0.35~nm$^{-2}$.
-\todo{HECTOR: I have deleted the sentence relating the uncertainty and the error as in my opinion there are other factors which are more important.
+The values of $\Gamma_{\rm ion}^{\rm water}$
+from different POPC simulations with then same molar concetration respect
+to water ($C_w$=350mM) are shown in Table~\ref{tab:binding}.
+As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess for the ECC-POPC model, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller than for the other models, 0.13--0.35~nm$^{-2}$.
+%\todo{HECTOR: I have deleted the sentence relating the uncertainty and the error as in my opinion there are other factors which are more important.
 The most important the slow convergence.
@@
     Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis
-    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
+    for different force fields.
+    %Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
+    {\color{red} Joe, please clarify the sentence below}
     In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
@@
     densities of phosphate group resp. water are divided by 5 resp. 200. 
-    The molar concentration of ions in water is 350~mM in all systems
-    presented here. 
+    All the simulations shown here have the same molar concentration of ions in water ($C_w$=350~mM).
     }
@@
 
+\subsection{Molecular interactions between Na$^+$ or Ca$^{2+}$ cations and POPC oxygens}
 We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
 A maximum distance of 0.3~nm from any lipid oxygen is used to determine a bound calcium.
-The results in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in contact with the phosphate oxygens.
+The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in contact with the phosphate oxygens.
 From those only one third (32\%) also interact with the carbonyl oxygens.
@@
       Isocontours of probability density of \ce{Ca^{2+}} with respect to 
-      the phosphate oxygens of POPC from ECC-lipids simulation.
+      the phosphate oxygens of POPC from ECC-POPC simulation.
       The probability density was evaluated around a single lipid, 
@@
   }
-  \todo{JOE: I'll update this figure with some ensemble of configuration to support binding preference of \ce{Ca^{2+}}
-  SAMULI: I am not sure if this would be needed anymore.}
+  \todo{The figure should be updated as discussed last week.}
 \end{figure}
@@
 
-Simple binding models have been used previously to interpret the experimental data \cite{altenbach84,macdonald87} we use in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
+Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} used in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
 In particular, NMR data about PC headgroup order parameters response and atomic absorption spectroscopy were best explained using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
@@
 Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
-We found that the largest propensity is for the 1:2 complex (41\%).
-Still, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower.
+We found the largest propensity for the 1:2 complex (41\%).
+However, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower.
 This variability suggests a more complex binding model than the previously considered in a simple ternary complex model.
 With a broad brushstroke, the simulation data can, however, be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
-This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S\ref{fig:cacl-bind} in SI).
+This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).
 
@@
 The probabilities of different complexes formed by \ce{Na+} ions and POPC
-analyzed from the simulation with the ECC-lipids model at $1\,$M concentration of NaCl are also 
+analyzed from the simulation with the ECC-POPC model at $1\,$M concentration of NaCl are also 
 shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
@@
 I will discuss this with Joe tomorrow.
-Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_w$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S\ref{fig:hist_residence_times} in SI.
+Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_w$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
 In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory ({\color{red} 800~ns}).
-In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-lipids model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$.
+In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$.
 The longest observed residence time is around 150~ns, which is well below the total length of the simulation used for analysis, i.e., 200~ns.
@@
 
-In conclusion, the results from the ECC-lipids model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within $\sim$100~ns timeframe, which is significantly faster than observed in most of the available lipid models~\cite{javanainen17}.
+In conclusion, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within $\sim$100~ns timeframe, which is significantly faster than observed in most of the available lipid models~\cite{javanainen17}.
 Sodium cations exhibit an even a more rapid exchange.
@@
 In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a classical MD simulation force field, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
-While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-lipids model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations.
+While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations.
 It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}.
-The ECC-lipids model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Van der Waals radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
+The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennart-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
 
@@
 However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data, where one calcium binds to two POPC molecules~\cite{altenbach84}.
-While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-lipids model, complexes of one or three lipids per one calcium were observed to relatively abundant and also almost equally likely.
+While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to relatively abundant and also almost equally likely.
 While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the complex nature of calcium binding to phospholipid bilayers observed in the present simulation.
@@
 Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions.
-To this end, the compatibility of the ECC-lipids model with existing models for proteins and other biological molecules should be verified and, if necessary, further adjustments following ECC concept to the force fields of the interacting molecules with the lipids should be addressed.
+To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules should be verified and, if necessary, further adjustments following ECC concept to the force fields of the interacting molecules with the lipids should be addressed.
 

### 2017-11-16 Manuscript/SI.tex

@@
 
+\newpage
 \section{Area per molecule and calcium binding with different water models}
@@
 \begin{table}[!h]
-  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
+  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si}
+  {\color{red}Joe, please deliver the citations for all experimental area per molecule values or remove them.}}
   \begin{tabular}{l|c c}
@@
     %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
+    ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
+    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
     ~ OPC3           & 62.2$\pm$ 0.6   &  300       \\
     ~ OPC3           & 64.2$\pm$ 0.6   &  313       \\
-    ~ SPC/E          & 65.1$\pm$ 0.6   &  313       \\
-    ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
     ~ OPC            & 64.4$\pm$ 0.6   &  313       \\
@@
 
-\begin{figure}[!h]
+\begin{figure}[!hp]
   \centering
@@
   \caption{\label{fig:ordPars_waterModels}
-    Changes of head group order parameters of POPC bilayer as a function of CaCl$_2$ concentrations
-    are shown from simulations with different force fields and water models together with experimental data 
-    (DPPC \cite{akutsu81} and POPC \cite{altenbach84}). 
-    Ion concentrations in bulk water are shown in x-axis. 
-    Values from simulations are calculated from the of cation number density $C_{np}$
-    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
+    Changes of head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of POPC bilayer
+    as a function of CaCl$_2$ concentrations in bulk ($C_b$) from ECC-POPC simulations at 313K with different
+    water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
+    together with experimental data (DPPC (323~K) \cite{akutsu81} and POPC (313~K) \cite{altenbach84}). 
+    Ion concentrations ($C_b$) are calculated from the cation number density $C_{np}$
+    at the farthest point from the lipid bilayer in the aqueous phase as [ion]=$C_{np}/0.602$.
   }
@@
 
-\section{Sodium binding to POPC with the glycerol carbon and changes in the head group angle}
+\newpage
+\section{Sodium binding to POPC with the glycerol carbon $g_3$ order parameters and the head group orientation}
 
@@
   \caption{\label{fig:delta_ordPar_NaCl_si}
-    Changes of head group and glycerol $g_3$ order parameters and the P-N vector angle of POPC bilayer as a function of NaCl concentrations
-    from simulations with different force fields together with experimental data \cite{akutsu81}. 
-    Ion concentrations in bulk water are shown in x-axis. 
+    Changes of head group and glycerol carbon $g_3$ order parameters, and the P-N vector orientation of POPC bilayer
+    as a function of NaCl concentrations in bulk ($C_b$)
+    from simulations with different force fields at 313~K together with experimental
+    data (DPPC (323~K)~\cite{akutsu81} and POPC (313~K)~\cite{altenbach84}). 
+%    Ion concentrations in bulk water are shown in x-axis. 
 %    Values from simulations are calculated from the of cation number density $C_{np}$
 %    from the region at the simulatin box edge with the constant ion concentration as [ion]=$C_{np}/0.602$.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters is taken directly from Ref. \cite{catte16}.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters at 293~K is taken directly
+    from Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
   }
@@
 
+\section{Ternary complex model in simulations}
+The NMR data about PC headgroup order parameters
+and atomic absorption spectroscopy data were previously
+best explained using a ternary complex binding model~\cite{altenbach84}.
+In this model, one calcium is assumed to form
+complexes with two lipids, i.e. with the binding stoichiometry of~1~\ce{Ca^{2+}}:2~POPC.
+The model predicts a linear relationship between quantities 
+$C_b$ and $\sqrt{C_b/C_I}$, where $C_b$ is the
+mole fraction of bound \ce{Ca^{2+}} per POPC
+and $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
+Experimentally detemined $C_b$ %was obtained from an extrapolation of linear relation 
+from NMR measurements and atomic absorption spectroscopy %for low concetrations of \ce{CaCl2}.
+%Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
+%remains constant throughout the extrapolation range. 
+together with $C_I$ calculated from the Poisson-Boltzmann equation
+gave a good agreement with the predictions of the
+ternary complex model~\cite{altenbach84}.
+
+To compare ECC-POPC simulations to the ternary complex model,
+%An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions;
+we calculated $C_b$ from simulations (as defined in the main text),
+and $C_I$ from the minimum CaCl$_2$ concentration at membrane-water interface,
+%the shallow depression separating the bulk water phase 
+locating around 2.6\,nm from the membrane center in
+the density profiles in Fig.~5 in the main text.
+
+The results from simulations are shown in Fig.~\ref{fig:cacl-bind}
+together with the line fitted to experimental data by Altenbach and Seelig~\cite{altenbach84}.
+Both results are in agreement with the prediction of the ternary complex model.
+The small discrepancy between the results from experiments and simulations probably
+arise from difference in the evaluation of the concentrations and
+inaccuracy of Poisson-Boltzmann theory for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
+In conclusion, the results suggest that the almost equal probabilities for
+\ce{Ca^{2+}} to form complexes with 1-3 lipids observed in the main text is,
+indeed, in line with the experimental observations supporting the ternary complex model.
+\newpage
+\begin{figure}[!h]
+  \centering
+  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
+  \caption{\label{fig:cacl-bind}
+    Fits of experimental~\cite{altenbach84} and ECC-POPC simulation data to
+    the prediction of the ternary complex model.
+    {\color{red} Joe, please change the axes labels to correspond the notation in the text, i.e. $C_b$ and $C_I$,
+    and write in the title of simulation fit (red dashed line) that it is the fit to simulation.}
+    %Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
+    %that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
+    %provides a good fit to experimental measurements~\cite{altenbach84}
+    %and it also provides a good fit to our simulation data. 
+    %This supports our statement that the success of the ternary complex model in the experiments
+    %can be understood with the atomistic detail of our simulations as an average 
+    %stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
+    %Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
+    %and, hence, the observed slope of the linear relationship is slightly different.
+    }
+\end{figure}
+
+\newpage
 \section{Histograms of residence times}
@@
    Histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer
-   at $450\,\mathrm{mM}$ concentration of \ce{CaCl_2} in water before adsorption to phospholipids
-   from simulations with ECC-lipids/ECC-ions (top)
-   and CHARMM36/ECC-ions (bottom), which was taken from previous study \cite{javanainen17} 
-   (simulation data available at Ref.~\citenum{zenodo.259376}).
-   The maxima on the x-axis represent the lengths of the simulations used for analysis. 
-   For the ECC-lipids model, 90\% of the residence times of \ce{Ca^{2+}} in a POPC membrane are 
+   from ECC-POPC (top) and CHARMM36 (bottom) simulations with ECC-ions.
+   Both simulations had the same concentration of \ce{CaCl_2} respect to water ($C_w$= 450~mM).
+   CHARMM36 simulation was directly taken from Refs.~\citenum{javanainen17,zenodo.259376}.
+   Scales of x-axes represent the lengths of the simulations used for analysis.
+   In ECC-POPC simulation, 90\% of the residence times are
    shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$                                                                          
@@
    which is well below the total length of the simulation (200~ns).
-   This is, however not true for the simulation with CHARMM36 model,
-   where there are several calcium cations with their residence time 
+   This is, however not the case in CHARMM36 simulation,
+   where residence times of several calcium cations are    
    apparently limited by the length of the simulation.
-   With such a simulation, which is relatively short for the employed model,
-   we can merely estimate an upper bound that 
-   only less than 60\% of the bound residence time 
-   is formed by cations interacting with the phospholipid membrane 
-   for less then half of the simulation length~(400~ns).
+   %With such a simulation, which is relatively short for the employed model,
+   %we can merely estimate an upper bound that
+   %only
+   Less than 60\% of the residence times are
+   %is formed by cations interacting with the phospholipid membrane
+   shorter than the half of the simulation length~(400~ns)
+   in CHARMM36 simulation.
   }
@@
 
-\section{Ternary complex model in simulations}
-It was found in the original work~\cite{altenbach84} that 
-a ternary complex binding model (i.e.~2~POPC:1~\ce{Ca^{2+}})
-provides the best fit to experimental measurements of all considered models in that study. 
-In such a model, there is a linear relationship between quantities 
-$C_b$, mole fraction of bound \ce{Ca^{2+}} per POPC, and $\sqrt{C_b/C_I}$, 
-where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
-The concentration $C_b$ was obtained from an extrapolation of linear relation 
-between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
-Such an extrapolation is valid as long as the mode of \ce{Ca^{2+}} binding 
-remains constant throughout the extrapolation range. 
-The concentration $C_I$ is determined from the surface potential using the Boltzmann equation.
-However, Boltzmann theory yields inaccurate results
-for divalent cations like \ce{Ca^{2+}}~\cite{Andelman1995}. 
-An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions;
-$C_b$ can be calculated from simulation using the robust definition in section~\ref{sec:affinity},
-and $C_I$ can be obtained from the density profiles in Fig.~\ref{fig:cacl-density} 
-as the minimum concentration at the shallow depression separating the bulk water phase 
-and the peak of adsorbed cations around 2.6\,nm from the membrene centre. 
-The small discrepancy between the results in the experiment~\cite{altenbach84} and 
-our simulations are likely caused by the difference in the evaluation of the concentrations.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[height=9.0cm,angle=-90]{../Fig/bound-CAs_conc-eccl17.eps}
-  \caption{\label{fig:cacl-bind}
-    Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
-    that assumes the stoichiometry of 2~POPC:1~\ce{Ca^{2+}} (details in reference~\citenum{altenbach84}) 
-    provides a good fit to experimental measurements~\cite{altenbach84}
-    and it also provides a good fit to our simulation data. 
-    This supports our statement that the success of the ternary complex model in the experiments
-    can be understood with the atomistic detail of our simulations as an average 
-    stoichiometry of otherwise almost evenly distributed clusters of one cation with 1,2 or 3 lipids (Fig.~\ref{fig:cacl_complexes}). 
-    Note that the units in the reference~\citenum{altenbach84} are different from the units presented here,
-    and, hence, the observed slope of the linear relationship is slightly different.
-    }
-\end{figure}
-
+\newpage
 \section{Comparison between Gromacs and openMM}
@@
     Order parameters of POPC head group, glycerol backbone and acyl chains 
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
-    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
-    compared with the experimental values from \cite{ferreira13}.
+    from Lipid14 \cite{dickson14} and ECC-POPC simulations ran
+    with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
+    together with experiments~\cite{ferreira13}.
     The size of the markers for the head group order parameters correspond to
@@
   \caption{\label{fig:ordPars_surf_GMX_oMM_compar}
-    The changes of headgroup order parameters and P-N vector orientation as a function of
+    Changes of headgroup and glycerol carbon $g_3$ order parameters, and P-N vector orientation as a function of
     a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
@@
   \caption{\label{fig:ordPars_cacl_GMX_oMM_compar}
-    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-lipids models
-    simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 
-    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    Ion concentrations in bulk water are shown in x-axis. 
+    Changes of the head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of a POPC bilayer 
+    as a function of the CaCl$_2$ concentration in bulk ($C_b$)
+    from Lipid14 \cite{dickson14} and ECC-POPC simulations ran
+    with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} at 313~K
+    together with experiments (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
+    %Ion concentrations in bulk water are shown in x-axis. 
     Bulk concentrations from simulations are calculated 

### 2017-11-16 Manuscript/manuscript.tex

@@
 bilayers are of paramount significance in the properties and functionality of cellular
-membranes. For example, possible accumulation of ions at the membrane interface determines
-the interaction affinity with charged species. Current estimates of binding affinities and
+membranes. Current estimates of binding affinities and
 stoichiometries of cations are, however, controversial due to limitations in the available experimental and
@@
 interpretation of the experiments relies on theoretical models, which give
-the best idealization of the ion binding as the direct experiments are not available. 
+the best idealization of the ion binding as direct experiments are not available. 
 Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
@@
 in agreement with the previous interpretations of the experimental spectroscopic data.
-On the other hand, the new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
+The new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
 equal probabilities, suggesting more complex binding stoichiometry than simple models
@@
 Ca$^{2+}$ also participates in the T-cell receptor activation.
-It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013}.
+It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013Ca2}.
 A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
@@
 Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.
-\todo{Hector, please deliver the missing citations for this paragraph.}
+\todo{Hector: I left one unknown citation for Pavel.}
 
@@
 
-According to ECC, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
+According to ECC mean-field correction\todo{HECTOR: I added this back because it is an important detail.}, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
 Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
@@
 as discussed more detail in the next section.
-In general, scaling down the partial charges reduced the ion binding affinity.
+Scaling down the partial charges reduced the ion binding affinity.
 We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
@@
 \end{equation}
-where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is a empirical constant depending on the valency and position of the bound charge.
+where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and position of the bound charge.
 
@@
 bulk region using the farthest point from the lipid bilayer in the aqueous phase.
-Notice that former concentration was used by Catte et al.~\cite{catte16}.
+Notice that the former concentration was used by Catte et al.~\cite{catte16}.
 Although these two definitions vary when applied
@@
 and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
-The SPC/E~\cite{Berendsen1987} water model was used in all ECC-lipid model simulations reported
+The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
 in the main text. This model was also used in the previous parametrization
-of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} and its lowered dielectric constant is consistent with
+of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
 the ECC concept~\cite{leontyev11, leontyev14}.
@@
 OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
-is tested in SI.
+is tested in the SI.
 Sodium, calcium, and chloride ions were modeled as ECC-ions
-with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennart-Jones radii for atoms
+with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
 forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
@@
 \begin{table}[tb!]
-  \caption{Values of the area per lipid (APL) of POPC bilayers without ions from the Lipid14 simulation
-    ran in this work and from the literature \cite{dickson14}, from the ECC-POPC model, and from experiments. \label{tab:apls}
+  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
   }
@@
 
-\subsection{Validation ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ Cations: the electrometer concept}
+\subsection{Validation ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
 
@@
     Exception was the order of magnitude larger error in
-    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) in too 
+    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) because the 
     small simulation box. 
@@
 It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}.
-The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennart-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
+The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
 

### 2017-11-16 Manuscript/SI.tex
Minor changes SI

@@
 In conclusion, the results suggest that the almost equal probabilities for
-\ce{Ca^{2+}} to form complexes with 1-3 lipids observed in the main text is,
-indeed, in line with the experimental observations supporting the ternary complex model.
+\ce{Ca^{2+}} to form complexes with one to three lipids detailed in the main text is in line with an averaged interpretation of the experimental observations which supported the ternary complex model.
 \newpage

### 2017-11-16 Manuscript/manuscript.tex
Hector and myself are now done with the modifications. Joe, plese check the todos and red text. Some of those require your attention.
One more thing: Consistent notation for concentrations throughout the manuscript and SI

@@
 \author{Josef Melcr}
-\author{Hector Martinez-Seara Monne}
+\author{Hector Martinez-Seara}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
-Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_w$,
+Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
 while others use atomic absorption spectroscopy and report the salt concentration in
-the supernatant after the solvation of lipids~\cite{altenbach84}, $C_b$. In this work,
+the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
 %we use the latter definition. In particular,
@@
     Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
-    in bulk ($C_b$) from simulations with different force fields at 313 K together with 
+    in bulk ($C_{ion}$) from simulations with different force fields at 313 K together with 
     experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
@@
     Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration in bulk ($C_b$)
+    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
     from simulations at 313 K together with experimental data 
@@
     and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
-    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ respect to water (C$_w$=350mM).
+    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ respect to water ($C_{ion}'$=350mM).
     %{\color{red}HECTOR:PLEASE CHECK}
@@
   \begin{tabular}{l|c c | c | c c}
-    model                  & $C_{w}$ & $C_b\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
     \hline
@@
 from different POPC simulations with then same molar concetration respect
-to water ($C_w$=350mM) are shown in Table~\ref{tab:binding}.
+to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
 As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess for the ECC-POPC model, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller than for the other models, 0.13--0.35~nm$^{-2}$.
@@
     densities of phosphate group resp. water are divided by 5 resp. 200. 
-    All the simulations shown here have the same molar concentration of ions in water ($C_w$=350~mM).
+    All the simulations shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
     }
@@
 I will discuss this with Joe tomorrow.
-Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_w$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
+Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_{ion}'$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
 In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory ({\color{red} 800~ns}).
@@
 J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.
-Computational resources were supplied by the Ministry of Education, Youth and Sports of the Czech Republic under the Projects CESNET (Project No.
-LM2015042) and CERIT-Scientific Cloud (Project No.
-LM2015085) provided within the program Projects of Large Research, Development and Innovations Infrastructures.
+Computational resources were supplied by the Ministry of Education, Youth and Sports
+of the Czech Republic under the Projects CESNET (Project No. LM2015042) and CERIT-Scientific
+Cloud (Project No. LM2015085) provided within the program Projects of Large Research,
+Development and Innovations Infrastructures.
+O.H.S.O. acknowledges financial support from
+Integrated Structural Biology Research Infrastructure of
+Helsinki Institute of Life Science (Instruct-HiLIFE).
 \end{acknowledgments}

### 2017-11-16 Manuscript/SI.tex

@@
     Changes of head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of POPC bilayer
-    as a function of CaCl$_2$ concentrations in bulk ($C_b$) from ECC-POPC simulations at 313K with different
+    as a function of CaCl$_2$ concentrations in bulk ($C_{ion}$) from ECC-POPC simulations at 313K with different
     water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
     together with experimental data (DPPC (323~K) \cite{akutsu81} and POPC (313~K) \cite{altenbach84}). 
-    Ion concentrations ($C_b$) are calculated from the cation number density $C_{np}$
+    Ion concentrations ($C_{ion}$) are calculated from the cation number density $C_{np}$
     at the farthest point from the lipid bilayer in the aqueous phase as [ion]=$C_{np}/0.602$.
@@
     Changes of head group and glycerol carbon $g_3$ order parameters, and the P-N vector orientation of POPC bilayer
-    as a function of NaCl concentrations in bulk ($C_b$)
+    as a function of NaCl concentrations in bulk ($C_{ion}$)
     from simulations with different force fields at 313~K together with experimental
@@
    from ECC-POPC (top) and CHARMM36 (bottom) simulations with ECC-ions.
-   Both simulations had the same concentration of \ce{CaCl_2} respect to water ($C_w$= 450~mM).
+   Both simulations had the same concentration of \ce{CaCl_2} respect to water ($C_{ion}'$= 450~mM).
    CHARMM36 simulation was directly taken from Refs.~\citenum{javanainen17,zenodo.259376}.
@@
     Changes of the head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration in bulk ($C_b$)
+    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
     from Lipid14 \cite{dickson14} and ECC-POPC simulations ran

### 2017-11-16 Manuscript/manuscript.tex
Removed a few red lines -- mostly checks

@@
 
-According to ECC mean-field correction\todo{HECTOR: I added this back because it is an important detail.}, the electronic polarizability can be modeled in non-polarizable MD by effectively embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
+According to ECC, the electronic polarizability can be modeled in non-polarizable MD in a mean-field way by 
+embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
 Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
@@
 were also performed. The TIP3p water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
-\todo{Joe, please confirm that this is correct.}
 Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
-Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
+Refs.~\citenum{catte16,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
 
@@
     Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-POPC models compared with experiments~\cite{kucerka11} at {\color{red}??K}.
+    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K.
     Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
     from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
-    compared with experiments \cite{ferreira13} at {\color{red}??K}.
+    compared with experiments \cite{ferreira13} at 300~K.
     The size of the markers for the head group order parameters correspond to
@@
     Bottom: The chemical structure of POPC and the labeling of the carbon segments.
-    {\color{red}Joe, please add the temperatures}
   } 
@@
     %{\color{red}HECTOR:PLEASE CHECK}
-    {\color{red}Joe, please confirm that $r^\mathrm{Ca^{2+}} _\mathrm{PO_4}$ is really 100\% for all the Lipid14 simulations
-      (i.e. that this is not a typo).}
+    %{\color{red}Joe, please confirm that $r^\mathrm{Ca^{2+}} _\mathrm{PO_4}$ is really 100\% for all the Lipid14 simulations
+    %  (i.e. that this is not a typo).}
   \label{tab:binding}}
@@
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
-    {\color{red} Joe, please clarify the sentence below}
-    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
-    densities of~\ce{Cl^-} and \ce{Na^{+}} are divided by 2, and
-    densities of phosphate group resp. water are divided by 5 resp. 200. 
-    All the simulations shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
+    %In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
+    The density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
+    the density profiles of phosphate groups resp. water are divided by 5 resp. 200. 
+    All simulations with \ce{CaCl2} shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
+    The simulation with \ce{NaCl} has $C_{ion}'$=1000~mM.
     }
@@
 
-\todo{HECTOR: I do fully understand the content of this paragraph.
-I will discuss this with Joe tomorrow.
-Please be carefull with my changes.} The histograms of residence times of \ce{Ca^{2+}} ($C_{ion}'$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
-In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory ({\color{red} 800~ns}).
+\todo{HECTOR: I do fully understand the content of this paragraph. I will discuss this with Joe tomorrow. Please be carefull with my changes.} 
+The histograms of residence times of \ce{Ca^{2+}} ($C_{ion}'$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
+In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory (800~ns).
 In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$.

### 2017-11-17 Manuscript/manuscript.tex
Edit to: abstract, intro

@@
 interpretation of the experiments relies on theoretical models, which give
-the best idealization of the ion binding as direct experiments are not available. 
+the \todo{Joe: I'm not sure about this: best idealization} of the ion binding as direct experiments are not available. 
 Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
@@
 Direct measurements of ion-membrane interactions in real biological systems are difficult. 
-Simplified lipid bilayers are often used instead as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}.
+Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
 For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
@@
 Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results.
-Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
+Analyzing massive amounts of data collected by an open collaboration method, 
+it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
 
@@
 
-In this work, we improve the cation binding of a POPC bilayer in MD simulations via the implicit inclusion of the electronic polarizability in the polar region of phospholipids.
-For this, we apply a simple mean-field electronic continuum correction (ECC)~\cite{leontyev11} which implies scaling the partial charges of the atoms to account for the missing electronic polarizability in MD simulations.
+In this work, we improve the cation binding in a POPC bilayer in MD simulations 
+via the implicit inclusion of the electronic polarizability in the polar region of phospholipids. 
+For this, we apply an electronic continuum correction (ECC)~\cite{leontyev11}, 
+in which scaling of the atomic partial charges is used to account for the missing electronic polarizability in MD simulations. 
 Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
@@
 The newly developed ECC-POPC model reproduces the experimentally
-measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while improving the membrane binding affinities significantly to sodium and calcium cations.
+measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while significantly improving the membrane binding affinities to sodium and calcium cations.
 

### 2017-11-17 Manuscript/manuscript.tex
Few things I noticed during scanning the paper.

@@
 
-\subsection{Validation ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
+\subsection{Validation of ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
 
@@
     and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
-    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ respect to water ($C_{ion}'$=350mM).
+    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
     %{\color{red}HECTOR:PLEASE CHECK}
@@
 
-\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to the POPC membrane}
+\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
 

### 2017-11-17 Manuscript/manuscript.tex
Some extra retouches in text.

@@
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
-in this work are summarized in Table~S1. Combability with openMM simulation package \cite{openmm7} was also
-tested in SI. The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
+in this work are summarized in Table~S1. 
+Compatibility with the openMM simulation package \cite{openmm7} was also
+tested in SI. 
+The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
 
@@
 
-\todo{Provide the parameteres for ECC-surfactant somewere.}
+\todo{Provide the parameteres for ECC-surfactant somewere.
+      Joe: They will appear along with the Zenodo upload of the trajectories.}
 
@@
   } 
+    \todo{Joe: I'm fine with the name change from ECC-lipids to ECC-POPC, but we have to be consistent throughout the text.
+I changed all accurences of ECC-lipids to ECC-POPC in the text, I yet need to change all figure legends then.}
 \end{figure}
@@
 %SAMULI: I can find the 63.2 from the SI table with SPE/E and the value 62.2 with OPC3, so everything seems to be ok for me.}.
-The values of the area per lipid of the ECC-POPC model show larger variations
+The values of the area per lipid of the ECC-POPC model vary
 when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI),
@@
 The values of $\Gamma_{\rm ion}^{\rm water}$
-from different POPC simulations with then same molar concetration respect
+from different simulations with the same molar concetration of cations with respect
 to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
@@
 SAMULI: I think that this was a good idea.}.
-Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our POPC-ECC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having a negative value $\Gamma_{Na}^{water} = (-0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}).
-This value means that water molecules are preferred to sodium and chloride ions at the membrane-water interface.
+Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having a negative value $\Gamma_{Na}^{water} = (-0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens})
+meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
 This is in contradiction with most of the available lipid force fields, which predict a positive binding of sodium to PC lipid bilayers \cite{catte16}.
@@
 The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in contact with the phosphate oxygens.
-From those only one third (32\%) also interact with the carbonyl oxygens.
-The interaction between calcium ions with only carbonyl oxygens is rare (1\%).
+From those only one third (32\%) also interacts with the carbonyl oxygens.
+The interaction of calcium ions with only carbonyl oxygens is rare (1\%).
 Neither individual lipids nor source chain of carbonyl is distinguished in the analysis.
@@
 In particular, NMR data about PC headgroup order parameters response and atomic absorption spectroscopy were best explained using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2}low concentrations~\cite{macdonald87}.
+However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2} at low concentrations~\cite{macdonald87}.
 %\todo{HECTOR: Why is the following sentence remotely significant.
@@
 In conclusion, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within $\sim$100~ns timeframe, which is significantly faster than observed in most of the available lipid models~\cite{javanainen17}.
-Sodium cations exhibit an even a more rapid exchange.
+Sodium cations exhibit an even more rapid exchange.
 This all suggests that simulations with a length of several hundreds of nanoseconds are sufficient to simulate ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
@@
 
-The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-lipid model.
+The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model.
 In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens.

### 2017-11-17 Manuscript/manuscript.tex
Add version to be given to Pavel

@@
 interpretation of the experiments relies on theoretical models, which give
-the \todo{Joe: I'm not sure about this: best idealization} of the ion binding as direct experiments are not available. 
+the best conception of the ion binding as direct experiments are not available. 
 Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
@@
 
-% Here I write pseudo-article statements that will make the main argument.
-% Beautiful polished sentences will be formed only afte we agree on these basic things
 \section{Introduction}
@@
 
-%\todo{PAVEL: introduce previous theoretical work that discusses cation binding to POPC w.r.t. its specific moiteties, e.g. Lukas' paper}.
-
-%While relative binding affinity of different ions in PC lipids is agreed to follow Hoffmeister
-%series, the molecular details of binding and binding energetics are
-
-%The binding details, like binding sites and stoichiometry are not yet fully
-%resolved but interpretation of NMR and scattering experiments suggest that one
-%Ca2+ interacts mainly with the choline groups \cite{hauser76,hauser78,herbette84} of two
-%phospholipid molecules \cite{altenbach84}.
-
 In this work, we improve the cation binding in a POPC bilayer in MD simulations 
@@
 measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while significantly improving the membrane binding affinities to sodium and calcium cations.
-
-%Since the cation binding affinity and the headgroup 
-%order parameter changes are in good agreement with experiments
-%in the proposed ECC-lipids, it can be used to interpret the
-%related structural changes and ion binding stoichiometry.
-%New lipid models with correct ion binding affinity in lipid
-%bilayers are necessary in applications of 
-%MD simulations with physiological salt conditions. 
-%The overestimated cation binding in the current lipid models \cite{catte16}
-%may lead to significant artefacts in MD simulations. For example,
-%articifially positively charged membranes overestimate interactions
-%with biomolecules having opposite sign.
 
@@
 Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
-Refs.~\citenum{catte16,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
+Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.  
+% SAMULI: I putted the trajectory without ions back, because it has been used as well.
 
@@
   }
-    %    { \color{red} HECTOR:Are we using here the same water model? If not is it fair the comparison considering the large changes observed in the SI table?
-    %      SAMULI: For Lipid14 simulations the water is TIP3P and SPC/E for ECC-lipids. I do understand why the comparison would
-    %    not be fair in this context?}
   \begin{tabular}{l|c c}
@@
 is $\approx$1\AA{} smaller
-%\todo{HECTOR: The 63.2 value does not appear in the SI table. There it is 62.2 but I'm not sure corresponds to the same system
-%SAMULI: I can find the 63.2 from the SI table with SPE/E and the value 62.2 with OPC3, so everything seems to be ok for me.}.
 The values of the area per lipid of the ECC-POPC model vary
@@
 lipid bilayer with the comparable accuracy to other state of the art lipid models~\cite{ollila16}.
-%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
-%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.
 
@@
 
-%
-%Therefore we use the data in Fig.~\ref{OrderParameterCHANGESsurf}
-%to suggest a linear relation between $\alpha$-carbon order parameter
-%and average P-N vector angle changes $\Delta \Theta_{\rm P-N}=?? \Delta S_{\rm CH}^{\alpha}$.
-%This can be used to estimate the change in P-N vector angle
-%from experiments measuring changes in $\alpha$-carbon order parameter. 
-%\todo{SAMULI: I think that we should establish the relation between 
-%order parameter and P-N vector angle changes by using ECC-lipids model.
-%This would be useful for people measuring aplha order parameter
-%changes from experiments.}
-%
-%For example, $\Delta S_{\rm CH}^{\alpha}=0.018$ measured due to the
-%addition of 66.8 $\mu$M concentration of Mellittin in bulk water \cite{kuchinka89}
-%gives $\Delta \Theta_{\rm P-N} = 2.5^{\circ}$.
-%\todo{I looked this roughly from the figure, should be calculated from 
-%the actual relation if we decide to put this.}
-
 \subsection{Validation of ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
@@
   }
-  %\todo{Add also the POPC result to the figure. Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.} \\
+  %Figure. in Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.
 \end{figure}
@@
     (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    %Ion concentrations in bulk water are shown in x-axis. 
-    %Bulk concentrations from simulations are calculated 
-    %from the farthest point from the lipid bilayer in the aqueous phase
     The error estimate for bulk concentrations was approximately 10\,mM.
@@
   }
-  %\todo{Experimental values from \cite{akutsu81} to be put in the $g_3$ figure:
+  %Experimental values from \cite{akutsu81} for $g_3$ figure:
   %the value of $g_3$ order parameter of DPPC was -0.214 
@@
   %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
-    %at the C-2 segments of both fatty acyl chains.'' }
+  %at the C-2 segments of both fatty acyl chains.'' 
 \end{figure}
@@
 
-%\todo{HECTOR: I'm not sure I understand the meaning of "structure" in the title.
-In general the titles in the result section are much alike between them, specially the first ones.
-This makes %difficult to guess where to find a particular piece of information.
-I will make an effort to make them more clear.
-%SAMULI: I agree. There are now new titles.}
-
 \begin{table}[tb!]
@@
     in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
-    %{\color{red}HECTOR:PLEASE CHECK}
-    %{\color{red}Joe, please confirm that $r^\mathrm{Ca^{2+}} _\mathrm{PO_4}$ is really 100\% for all the Lipid14 simulations
-    %  (i.e. that this is not a typo).}
   \label{tab:binding}}
@@
     %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
-    %In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
-    The density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
-    the density profiles of phosphate groups resp. water are divided by 5 resp. 200. 
+    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
+    the density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
+    the density profiles of phosphate groups and water are divided by 5 and 200, respectively. 
     All simulations with \ce{CaCl2} shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
@@
 
-\todo{HECTOR: I do fully understand the content of this paragraph. I will discuss this with Joe tomorrow. Please be carefull with my changes.} 
-The histograms of residence times of \ce{Ca^{2+}} ($C_{ion}'$ = 450~mM) in a POPC bilayer for simulations using either ECC-POPC or CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
+%\todo{HECTOR: I do fully understand the content of this paragraph. I will discuss this with Joe tomorrow. Please be carefull with my changes.} 
+The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
+ECC-POPC and CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
 In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory (800~ns).
-In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model, where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$.
-The longest observed residence time is around 150~ns, which is well below the total length of the simulation used for analysis, i.e., 200~ns.
-Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC headgroup is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}.
-The exchange of \ce{Na+} ions at the POPC membrane is another order of magnitude faster, yielding 90\% of the residence times smaller than~1~ns, with the longest residence time being~6~ns.
-
-%give a roughly estimate of the order
-%of $1\,\mu\mathrm{s}$
-%This is unfortunately not the case of many theoretical works,
-%where the residence time cannot be estimated,
-%or the simulation has not even decorrelated from it's 
-%initial condition with a population of bound/unbound cations \cite{catte16}, 
-%because it is limited by the simulation length, 
-%even though it can be very long (order of $\mu\mathrm{s}$) in some works, e.g. reference~\citenum{javanainen17}. 
-%Increasing time scales to the order of $\mu\mathrm{s}$ 
-%doesn't help to alleviate such a problem in current lipid models \cite{javanainen17}. %and also \cite{catte16}, but same simulation was used.
+In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
+where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
+residence time is around 150~ns, which is well below the total length of the simulation used for analysis, i.e., 200~ns.
+Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
+headgroup is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}. The exchange of \ce{Na+} ions at the POPC membrane
+is another order of magnitude faster, yielding 90\% of the residence times smaller than~1~ns, with the longest residence time being~6~ns.
 
@@
 
-%but we expect that the correction can be used also for other lipids
-%and force fields.
-
-%We suggest using state of the art water models like OPC3\cite{Izadi16} or OPC \cite{Izadi14},
-%which yield higher accuracy than the traditional TIP3p water model~\cite{jorgensen83}.
-%Several water models 
-%(OPC3\cite{Izadi16}, OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987} and TIP4p/2005 \cite{Abascal2005}) 
-%were used to exemplify the transferability of 
-%the parameters of the new ECC-lipids force field.
-
 This work can be reached as a repository containing all data at \url{zenodo.org:\dots\dots\dots}.
-%and as project \mbox{NMRlipids~VI} in \url{nmrlipids.blogspot.fi}.
 

### 2017-11-23 Manuscript/manuscript.tex
Fixing typos after from Pavel's latest version

@@
 membranes. Current estimates of binding affinities and
-stoichiometries of cations are, however, controversial due to limitations in the available experimental and
-theoretical methods. Experimentally one can assess these parameters by titrating
-several membrane properties as a function of the ion concentration. However, the
-interpretation of the experiments relies on theoretical models, which give
-the best conception of the ion binding as direct experiments are not available. 
-Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
-resolution, therefore offering all the necessary information to interpret
-experimental data without the need to resort to simplified models. However, the accuracy
-of the available lipid models when interacting with ions is not sufficient for
-such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
+stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
+computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
 1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
 polarization as a mean field correction, known as the electronic continuum correction (ECC).
-This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
+This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
 Our improved ECC-POPC model reproduces not only the
 experimentally measured structural parameters for the ion-free membrane, but also the
-response of lipid headgroup to a bound positive charge, and the binding affinities of 
-Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
-bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens,
+response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
+Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
+bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
 in agreement with the previous interpretations of the experimental spectroscopic data.
-The new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
-equal probabilities, suggesting more complex binding stoichiometry than simple models
-previously used to interpret the NMR data.
-The results of this work pave the way to quantitative MD simulations of complex
-biochemical systems with realistic electrostatic interactions in the vicinity of cellular
-membranes.
+The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
+equal probabilities, suggesting more complex binding stoichiometries than those from simple models
+used to interpret the NMR data previously.
+The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
+of complex biochemical systems at cellular membranes.
 \end{abstract}
@@
 \section{Introduction}
-Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}.
-Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
-Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface.
-For example, Ca$^{2+}$  is crucial in neural signal propagation.
+Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}.
+Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
+For example, Ca$^{2+}$  is crucial in propagation of neural signals.
 It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}.
 Ca$^{2+}$ also participates in the T-cell receptor activation.
-It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013Ca2}.
-A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
-Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions.
-Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.
-\todo{Hector: I left one unknown citation for Pavel.}
+It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}.
+Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
+Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions.
+Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.
 
-Direct measurements of ion-membrane interactions in real biological systems are difficult. 
+Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
 Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
-For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
+For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
 The details of ion binding are, however, not fully consistent in the literature.
-Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mainly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
-On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-On the other hand, most atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
-For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
-%{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}
+Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
+In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
+Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}.
+Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
 
-Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
-A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept~\cite{seelig87}.
-Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results.
-Analyzing massive amounts of data collected by an open collaboration method, 
-it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
+Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements.
+A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}.
+Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results.
+Analyzing massive amounts of data collected within an open collaboration approach, 
+it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
 
-In this work, we improve the cation binding in a POPC bilayer in MD simulations 
-via the implicit inclusion of the electronic polarizability in the polar region of phospholipids. 
-For this, we apply an electronic continuum correction (ECC)~\cite{leontyev11}, 
-in which scaling of the atomic partial charges is used to account for the missing electronic polarizability in MD simulations. 
-Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
-As a starting template for our POPC model, we will use the Lipid14~\cite{dickson14}.
-This force field provides one of the best available descriptions of cation binding \cite{catte16}.
+In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
+via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
+To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
+within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
+Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
+Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model.
+This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}.
 The newly developed ECC-POPC model reproduces the experimentally
-measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while significantly improving the membrane binding affinities to sodium and calcium cations.
+measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.
 
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-The lack of electronic polarizability in standard MD simulation force fields has been considered a serious issue since the early days of lipid bilayer simulations.
-In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by including the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
+The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations.
+In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
 Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
@@
 
-According to ECC, the electronic polarizability can be modeled in non-polarizable MD in a mean-field way by 
+According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
 embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
-Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
+Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
 \begin{equation}
@@
 \end{equation}
-for the ECC corrected charges in simulation. 
+for the ECC corrected charges. 
 Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor successfully improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
-It is important to note that the 
+This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
+It is important to note that the value of the
 high frequency dielectric constant 
 is around 2 for almost any biologically relevant environment \cite{leontyev11}.
-The dielectric discontinuity in a lipid bilayer arises
-from the orientational polarization of the molecules. 
-Thus, the same correction for the electronic polarizability can be 
-applied throughout the lipid bilayer interface.
+The dielectric discontinuity in a lipid bilayer thus arises only
+from the nuclear (orientational) polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
+Therefore, the same correction for the electronic polarizability can be 
+applied throughout the lipid bilayer/aqueous solution interface.
 
-While using a scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
+While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
 Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
-Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}.
-Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
-Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$.
-Instead, a consistent scaling factor would lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
+Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}.
+Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
+Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$.
+Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
 
-Here we develop ECC-POPC lipid model that accurately describes the binding 
-of sodium and calcium ions to the lipid bilayer. 
-The Lipid14~\cite{dickson14} force field parameters 
-(available in a Gromacs format from Ref. \citenum{lipid14files}) were used as a starting 
-point. 
-This model provides the best response of the head group to ions among the available 
+Here we develop a ECC-POPC lipid model that accurately describes binding 
+of sodium and calcium ions to the POPC  lipid bilayer. 
+The Lipid14~\cite{dickson14} force field 
+(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
+point since 
+it provides the most accurate response of the head groups to ions among the available 
 lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
-has a relatively realistic head group, glycerol backbone, and acyl chain structure~\cite{dickson14,botan15}.
+provides realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
 We applied the ECC correction 
-to the Lipid14 model of POPC by scaling of 
+to the Lipid14 model of POPC by scaling 
 partial charges of the head group, glycerol 
@@
 To reproduce the experimental ion binding affinities,
-we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
-The ion binding affinity was evaluated against experiments
+we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
+The ion binding affinity was benchmarked against experiments
 using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
@@
 Scaling down the partial charges reduced the ion binding affinity.
-We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
-which is only slightly higher than the theoretical one ($f_q=0.75$).
+We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
+which is only slightly higher than the ECC one ($f_q=0.75$).
 Directly applying
@@
 (64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
-reduced hydration of the lipid head group region after scaling charges which effectively
-reduces head group polarity. We solve this difficulty by reducing the effective radii of
-the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a
-factor of $f_\sigma = 0.89$. The same solution was taken for the ECC-ions in aqueous
-solvent \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the $\sigma$ parameters,
-the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
+reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
+reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
+the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
+factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
+solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}.
+After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
 
@@
 Comparing MD simulation to NMR experiments, we can validate the ion
-binding affinity in lipid bilayer simulations using the ''electrometer concept''~ \cite{seelig87, catte16}.
-This method relies on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
-%Comparing this response is an excellent way to validate force fields as both NMR and MD can accurately determine order parameters ~\cite{catte16, ollila16}.
+binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
+This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
 The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
-From MD simulations the order parameters can be calculated using the definition
+From MD simulations the order parameters can be calculated using the relation
 \begin{equation}\label{OP}
@@
 \end{equation}
-where $\theta$ is the angle between the C-H bond and membrane normal. Angular brackets point to the average over all sampled configurations.
+where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.
 
@@
 \end{equation}
-where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and position of the bound charge.
+where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.
 
@@
 This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
-To calibrate the response of the head group order parameter to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
+To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
 DHAB\\[0.5cm]
@@
 \vspace{0.5cm} \\
-is a cationic surfactant having two acyl chains and bearing a unit charge in the hydrophilic end.
-Thus, it is expected to locate in the bilayer similarly to the phospholipids, and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
+is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end.
+Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
 
-\subsection{Salt concentrations and binding affinity}
-NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
-Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
+\subsection{Salt concentrations and binding affinities}
+When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
+Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
 while others use atomic absorption spectroscopy and report the salt concentration in
 the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
-%we use the latter definition. In particular,
-we measure the latter concentration in the aqueous
+we stick to the latter definition by estimating the salt concentration in the aqueous
 bulk region using the farthest point from the lipid bilayer in the aqueous phase.
-Notice that the former concentration was used by Catte et al.~\cite{catte16}.
-Although these two definitions vary when applied
-to Ca$^{2+}$, they do not significantly affect the conclusions of this work.
+Note that the former definition was used by some of us previously.~\cite{catte16}.
+Although these two definitions provide somewhat different values when applied
+to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.
 
 To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}.
-Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing plane between the membrane interior and the water bulk region.
+Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region.
 In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.
-Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
+Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
  \begin{equation}\label{surfexcess}
@@
 \subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
-The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and \mbox{x-ray} scattering experiments by the scattering form factors.
+The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors.
 NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution.
@@
 \ref{OP}.
-Scattering form factors validate the dimensions of the lipid bilayer (e.g., thickness and area per molecule).
-Form factors were calculated using
+Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule).
+Form factors were calculated using a relation
 
@@
 The robustness of our approach with other water models (OPC \cite{Izadi14},
-OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
-is tested in the SI.
+OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3P-FB and TIP4P-FB~\cite{Wang2014}, and TIP4P/2005 \cite{Abascal2005})
+was verified and is presented in the Supporting Information (SI).
 Sodium, calcium, and chloride ions were modeled as ECC-ions
 with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
-forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
+forming the POPC lipid were derived in this work starting from the Lipid14 force field~\cite{dickson14}.
 For comparison, simulations with the Lipid14 model~\cite{dickson14} and
 ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
-were also performed. The TIP3p water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
-Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
+were also performed. The TIP3P water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
+Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and the TIP3P~\cite{jorgensen83} water model were taken directly from
 Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
-% SAMULI: I putted the trajectory without ions back, because it has been used as well.
 
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
-in this work are summarized in Table~S1. 
+in this work are summarized in Table~S1 in the SI. 
 Compatibility with the openMM simulation package \cite{openmm7} was also
-tested in SI. 
+tested and is presented in the SI. 
 The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
@@
 \subsubsection{Simulations of POPC bilayers with cationic surfactants}
-The topology of dihexadecyldimethylammonium was created with the automated topology builder \cite{malde11}.
-The AmberTools program~\cite{amber} was then used to generate the Amber-type force field parameters.
+The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
+The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters.
 These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
-The partial charges were then manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulation with ECC-POPC, the partial charges of the headgroup were scaled, i.e., $f_q=0.8$, and the $\sigma$ of the scale atoms reduced as in POPC-ECC.
+The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
+For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.
 
@@
 All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
-Chloride counter ions were used in simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
-For POPC either the Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987}  were used.
-The first 20~ns of the total simulation time of 200~ns was considered as an equilibration time and was omitted from the analysis.
-Sufficient lipid neighbor exchange occurred during the simulations.
+Chloride counter ions were used in all the simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
+For POPC, either Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987} parameters  were used.
+The first 20~ns of the total simulation time of 200~ns was considered as an equilibration period and was thus omitted from the analysis.
+We checked that a sufficient lipid  exchange occurred during the simulations.
 The trajectories and simulation files for unscaled systems are available from
@@
 
-\todo{Provide the parameteres for ECC-surfactant somewere.
+\todo{Provide the parameters for ECC-surfactant somewhere.
       Joe: They will appear along with the Zenodo upload of the trajectories.}
@@
 
-\subsection{Structural parameters of pure ECC-POPC model membrane are in agreement with experiments}
+\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}
 
@@
 First, we present results for bilayers in pure water.
-The ECC-POPC and Lipid14 models both reproduce the experimental x-ray scattering form factors
-of a POPC bilayer with a comparable accuracy in Fig.~\ref{simVSexpNOions}.
+The ECC-POPC and Lipid14 models both reproduce the experimental X-ray scattering form factors
+of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
 The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
 experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
-is $\approx$1\AA{} smaller
-The values of the area per lipid of the ECC-POPC model vary
-when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI),
+is $\approx$1\AA{} smaller than the experimental one.
+The values of the area per lipid of the ECC-POPC model vary slightly
+when simulated with different water models (i.e., within the interval of 62.2--66.8 \AA{}, see Table~S2 in SI),
 while still being close to the experimentally reported values.
 We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
-lipid bilayer with the comparable accuracy to other state of the art lipid models~\cite{ollila16}.
+lipid bilayer with a comparable accuracy to other state-of-the-art lipid models~\cite{ollila16}.
 
-Similarly, the acyl chain order parameters of the ECC-POPC model, as well as the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
-Notably, the experimentally measured forking and small order parameter values of $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
-This feature has been suggested to indicate that the carbonyl of {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in {\it sn}-1 chain, which would orient more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
-While this arrangement may be a relevant feature for the ion binding details, it is not fully reproduced by other available lipid models~\cite{ollila16}.
-%This is expected as the ECC model does not modify the already highly optimized parameters of this region in Lipid14
+Similarly, the acyl chain order parameters of the ECC-POPC model, as well as those of the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
+Notably, the experimentally measured forking and small order parameter values of the $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
+This feature has been suggested to indicate that the carbonyl of the {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in the {\it sn}-1 chain, which orients more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
+This arrangement, which is not fully reproduced by other available lipid models~\cite{ollila16}, may be a relevant feature for the ion binding details.
 
-The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-POPC model, see Fig.~\ref{OrderParameterCHANGESsurf}.
-Considering the available experimental evidence, it is not possible to suggest which of the two models provides more realistic headgroup orientation.
+The order parameters of the $\alpha$ and $\beta$ carbons in the head group are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the former model, see Fig.~\ref{OrderParameterCHANGESsurf}.
+While both models perform satisfactorily, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations.
 The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
-The accuracy of both models in the glycerol backbone region is equally footed with any other state of art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
+The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
 
@@
   \caption{\label{OrderParameterCHANGESsurf}
-    The changes of headgroup order parameters and P-N vector orientation as a function of
+    The changes of head group order parameters and P-N vector orientation as a function of
     a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
@@
 
-Before studying the sodium and calcium ion binding affinities, we quantify the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
-These mixtures have a well-defined amount of bound charge per PC, i.e., the molar fraction of cationic surfactants.
-This direct relation results from the ability of dihexadecyldimethylammonium to be inserted in the lipid bilayers as any lipid because of its two hydrophobic acyl chains.
-Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid headgroup order parameters to the amount of bound charge in simulations \cite{scherer89}.
+Before studying the sodium and calcium ion binding affinities, we quantify the response of the head group order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
+These mixtures have a well-defined amount of bound charge per PC, namely, the molar fraction of cationic surfactants.
+This is due to the ability of dihexadecyldimethylammonium to directly insert in the lipid bilayer due to its two hydrophobic acyl chains.
+Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid head group order parameters to the amount of bound charge in simulations \cite{scherer89}.
 
-The changes of the headgroup order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
+The changes of the head group order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
 An approximately linear decrease of the order parameters, as expected from Eq.
 \ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
-The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters is too large to the bound positive charge.
-In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.
+The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters to the bound positive charge is too strong.
+In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being only slightly underestimated for the $\beta$ segment.
 
 In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
-As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in a PC lipid bilayer.
+As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in the PC lipid bilayer.
 The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
-For example, the addition of 50\% mole fraction of the cationic surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only 11$^{\circ}$ in the ECC-POPC model.
+For example, the addition of 50\% mole fraction of the cationic surfactant leads to a decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only of 11$^{\circ}$ in the ECC-POPC model.
 The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model.
-The lower sensitivity of the P-N vector angle response in the ECC-POPC model agrees better with experiments.
-%The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules.
-%{\color{red}., as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
+The weaker sensitivity of the P-N vector angle response in the ECC-POPC model is in better agreement with experiments.
 
@@
 
-Changes of the lipid bilayer head group order parameters from different simulations and
+Changes of the lipid bilayer head group order parameters extracted from simulations and
 experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
@@
 The experimentally measured small order parameter
-changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) %and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81, altenbach84}.
+changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) 
 are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
 However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
-%
-%A recent comparison of different simulation
-%models to the experimental data revealed that most force fields significantly
-%overestimate Na$^+$ ion binding to PC lipid bilayers and that none of the available
-%models correctly reproduces the details of the binding of Ca$^{2+}$ ions which is
-%grossly overestimated \cite{catte16}.
-%
-%Since artifacts were reported
-%in simulations involving \AA{}qvist ions is water \cite{auffinger07}, we tested the
-%
-Replacing \AA{}qvist ions with ions by Dang et al.~\cite{smith94, chang1999, dang2006}
+Replacing \AA{}qvist ions with ion parameters by Dang et al.~\cite{smith94, chang1999, dang2006}
 or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
@@
 
-%To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
-%polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
-
-The results from the simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
-Considering that we can also reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC-model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
+The results from simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as a function of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
+Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
 Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-%\todo{HECTOR: I do not believe in the content of the following sentence as it is.
-How do we know that the changes of the P-N angle are too pronounced for Lipid14?.
-There is no direct correlation between angle and order parameter.}
-% SAMULI: The sentence is now modified.
-Also the changes of the P-N vector angle are more pronounced for the Lipid14 model,
+Also the changes of the P-N vector angle are too pronounced for the Lipid14 model,
 for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
@@
 for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
-%\todo{HECTOR: I have a problem when comparing fig.
-2 and 3.
-If I read it right the behavior when adding calcium and surfactant is not the same.
-Is the latter changing the area per lipid and therefore the surface charge density?
-If this will be the case then besides the effect of the charge we will have a packing effect.
-This is more relevant for the more concentrated cases.
-Shouldn't we try to discuss the two figures and what the differences and similarities mean?
-Given our previous work, should we check the behavior respect the charge density?}
-%SAMULI: Not now, hopefully in the future.
 
 Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge.
-It is plausible that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance to validate the lipid headgroup order parameter response to the bound charge.
+It is plausible to assume that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance of validation of the lipid headgroup order parameter response to the bound charge.
 
 Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI.
-In general, the performance of ECC-POPC with any of the water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3p model.
+In general, the performance of ECC-POPC with any of the tested water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3P model.
 
-\subsection{Binding affinity of Na$^+$ and Ca$^{2+}$ cations to POPC membrane}
+\subsection{Binding affinities of Na$^+$ and Ca$^{2+}$ cations to the POPC membrane}
 \label{sec:affinity}
@@
 
-Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
+Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
 The values of $\Gamma_{\rm ion}^{\rm water}$
@@
 to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
-As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess for the ECC-POPC model, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller than for the other models, 0.13--0.35~nm$^{-2}$.
-%\todo{HECTOR: I have deleted the sentence relating the uncertainty and the error as in my opinion there are other factors which are more important.
-The most important the slow convergence.
-SAMULI: I think that this was a good idea.}.
-Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having a negative value $\Gamma_{Na}^{water} = (-0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens})
-meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
-This is in contradiction with most of the available lipid force fields, which predict a positive binding of sodium to PC lipid bilayers \cite{catte16}.
-%\todo{HECTOR: do we need to add a NaCl table or an extended table for CaCl2 using different concentrations in the SI? I think this is very important information
-%  SAMULI: I agree that this is important information, but we have quite a lot of stuff in the paper already. Maybe we could leave this for the future studies.}
+As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess of calcium, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller for the ECC-POPC model than for the other models, 0.13--0.35~nm$^{-2}$.
+Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}.
+This
+means that on average water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
+This is in contradiction with most of the available lipid force fields, which predict a small but positive surface excess of sodium at PC lipid bilayers \cite{catte16}.
 
@@
     for different force fields.
-    %Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
-    %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
     In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
@@
 We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
-A maximum distance of 0.3~nm from any lipid oxygen is used to determine a bound calcium.
-The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in contact with the phosphate oxygens.
-From those only one third (32\%) also interacts with the carbonyl oxygens.
-The interaction of calcium ions with only carbonyl oxygens is rare (1\%).
-Neither individual lipids nor source chain of carbonyl is distinguished in the analysis.
-The most likely interactions between Ca$^{2+}$ ions and phosphate oxygens are visualized with the probability density isocontours in Fig.~\ref{fig:volmaps}.
+A maximum distance of 0.3~nm from any lipid oxygen is used to define a bound calcium.
+The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in direct contact with phosphate oxygens.
+From these ions, only one third (32\%) also interacts with the carbonyl oxygens, while the interaction of calcium ions with carbonyl oxygens only is rare (1\%).
+The most abundand interaction scenarios between Ca$^{2+}$ ions and phosphate oxygens are visualized using the probability density isocontours in Fig.~\ref{fig:volmaps}.
 While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.
@@
 
-Even though \ce{Na+} ions do not specifically bind to a POPC bilayer, they still interact mostly with its oxygen moieties.
-The results from the simulation at $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions interact with only phosphate oxygens of POPC and 20\% with only carbonyl oxygens, and the rest, 25\%, is interacting with both.
+Even though \ce{Na+} ions do not bind strongly to a POPC bilayer, they still interact mostly with its oxygen moieties.
+The results from a simulation at a $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions at the bilayer interact with phosphate oxygens of POPC only and 20\% with carbonyl oxygens only, with the remaining 25\%, is interacting with both negatively charged groups.
 
-In conclusion, the results suggest that calcium ions bind specifically to phosphate oxygens, occasionally interacting also with carbonyls.
-This is in good agreement with previous conclusions from several experimental and computational studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}, but suggesting a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
-Sodium ions also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
-%\todo{This last sentence is weak and not relevant.
-If you want to keep it go ahead, but then I would ask you a simple question: Is it incompatible to have no overall binding (this relates to the whole surface volume) with having small cavities in the membrane with an increase Na+ concentration.
-If the answer is negative maybe the statement is incorrect and who knows it might be relevant for something.
-After all even Ca2+ affinity changes in the presence of sodium.} The physiological relevance of these details of sodium interactions is, however, unclear due to its overall very weak binding affinity.
+In conclusion, the results suggest that calcium ions bind specifically to the phosphate oxygens, occasionally interacting also with the carbonyls of the PC lipids.
+This is in a qualitative agreement with previous conclusions from several experimental and computational studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}.
+However, the present results suggest, in
+agreement with experiments, an overally weaker binding to the bilayer, in particular with a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
+Sodium ions, which do not exhibit any appreciable affinity for the bilayer, also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
 
@@
 
-Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} used in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
-In particular, NMR data about PC headgroup order parameters response and atomic absorption spectroscopy were best explained using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2} at low concentrations~\cite{macdonald87}.
-%\todo{HECTOR: Why is the following sentence remotely significant.
-I will vote for removing it} Note that the results based on these binding models and the measured headgroup order parameter changes are also included in the tabulated experimental binding constants~\cite{marsh13}.
+Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} as employed in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
+In particular, NMR data concerning the PC headgroup order parameters response and atomic absorption spectra were explained best using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
+Nevertheless, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering \ce{CaCl2} at low concentrations only~\cite{macdonald87}.
 
-In this work, we reproduce the same experimental data used to calculate binding stoichiometry with our ECC-POPC model.
-Using our simulations, we have direct access to exact details of the binding stoichiometry without requiring any middle model as in experiments~\cite{altenbach84, macdonald87}.
-To evaluate the relative propensities for each stoichiometric complex (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
+In this work, we reproduce the same experimental data used to infer binding stoichiometries employing our ECC-POPC model.
+Thanks to our simulations, we have a direct access to atomistic details of the binding stoichiometry without a need for any binding model as employed for interpreting in experiments~\cite{altenbach84, macdonald87}.
+To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
 Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
-We found the largest propensity for the 1:2 complex (41\%).
-However, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower.
-This variability suggests a more complex binding model than the previously considered in a simple ternary complex model.
-With a broad brushstroke, the simulation data can, however, be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
+We found the largest propensity for the 1:2 complex (41\%), with probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~being only slightly lower.
+This suggests a more complex binding model than considered in a simple 1:2 ternary complex model previously.
+Nevertheless, with a broad brushstroke, the simulation data can be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
 This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).
@@
 
-\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in POPC membrane}
+\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in the POPC membrane}
 
-Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current state of the art force fields, e.g., CHARMM36 and Slipids force fields~\cite{javanainen17}.
-This suggests that at least several microseconds are required to reach the binding/unbinding equilibrium.
-%\todo{HECTOR: I disagree with the following lines.
-In the case of most of the unscale force fields the issue is that the binding of Calcium to the membrane is very strong.
-Therefore, the reorganization takes too long leading to a very slow equilibration process.
-In this case, the size of the hill in the choline region is nothing compared to the size between the bound and solvated state which is the real bottleneck.
-Even in the case of the scaled simulations, I believe the same applies but to a lesser extent.
-I might be wrong in this last part.
-If you do not agree we can discuss it tomorrow via skype.} {\color{red} This is in line with the density profiles of \ce{Ca^{2+}} from simulations in Fig.~\ref{fig:cacl-dens}, which all show a region of a very small calcium density around the level of choline groups in POPC (i.e., around $2.5\,\mathrm{nm}$).
-In contrast, simulations with  ECC-lipids exhibits only a relatively shallow local minimum at that level.
-This suggests that the smaller barrier for the exchange of calcium ions shall lead to their faster exchange with the solvent within ECC-lipids.}
-To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated residence times of ions bound to the membrane.
-In our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
+Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current force fields, such  as CHARMM36 and Slipids force fields~\cite{javanainen17}.
+This suggests that at least several microseconds are required to reach the ion binding/unbinding equilibrium.
+To quantify the exchange of ions between the membrane and aqueous solution in simulations, we evaluated residence times of ions bound to the membrane.
+Within our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
 
-%\todo{HECTOR: I do fully understand the content of this paragraph. I will discuss this with Joe tomorrow. Please be carefull with my changes.} 
 The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
-ECC-POPC and CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
-In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory (800~ns).
+ECC-POPC and CHARMM36 (simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
+In the CHARMM36 simulation, a significant number of the calcium ions is bound to the membrane for the whole length of the trajectory (800~ns).
 In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
 where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
-residence time is around 150~ns, which is well below the total length of the simulation used for analysis, i.e., 200~ns.
+residence time is around 150~ns, which is below the total length of the simulation used for analysis, i.e., 200~ns.
 Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
-headgroup is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}. The exchange of \ce{Na+} ions at the POPC membrane
-is another order of magnitude faster, yielding 90\% of the residence times smaller than~1~ns, with the longest residence time being~6~ns.
+headgroup is of the order of $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
+is yet another order of magnitude faster, with 90\% of the residence times smaller than~1~ns and the longest residence time being~6~ns.
 
-In conclusion, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within $\sim$100~ns timeframe, which is significantly faster than observed in most of the available lipid models~\cite{javanainen17}.
-Sodium cations exhibit an even more rapid exchange.
-This all suggests that simulations with a length of several hundreds of nanoseconds are sufficient to simulate ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
-This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and calcium cations \cite{javanainen17, catte16}.
-%Such a finding changes the point of view on binding of calcium to PC membranes from
-%a very strong long-term stable binding with rare exchanges to 
-%a relatively frequent exchange of clacium cations 
-%in equilibrium between membrane and solvent.
+In summary, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within the $\sim$100~ns timeframe, which is significantly faster than observed in simulations emloying most of the presently available lipid models~\cite{javanainen17}.
+Sodium cations exhibit an even more rapid exchange between the membrane and the aqueous solution.
+Our results suggest that simulations with a length of several hundreds of nanoseconds are sufficient to simulate alkali and alkali earth ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
+This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and, in particular,  calcium cations \cite{javanainen17, catte16}.
 
@@
 
-In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a classical MD simulation force field, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
-While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations.
-It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}.
-The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
+In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a force field MD simulation, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
+This ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
+While the structural details of a POPC lipid bilayer in pure water simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the new model in addition reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ at varying concentrations.
+It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
 
 The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model.
-In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens.
-However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data, where one calcium binds to two POPC molecules~\cite{altenbach84}.
-While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to relatively abundant and also almost equally likely.
-While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the complex nature of calcium binding to phospholipid bilayers observed in the present simulation.
+In line with previous interpretations of the experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens, with interactions to the carbonyls being of a secondary importance.
+However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data previously, within which one calcium binds to two POPC molecules~\cite{altenbach84}.
+While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to be relatively abundant and also almost equally likely to occur.
+While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the detailed nature of calcium binding to phospholipid bilayers observed in the present simulation.
 
-Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions.
-To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules should be verified and, if necessary, further adjustments following ECC concept to the force fields of the interacting molecules with the lipids should be addressed.
+Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions, including a mean-field account for electronic polarization effects.
+To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules needs to be verified and, if necessary, further adjustments following the ECC concept to the force fields of the biomolecules interacting with the lipids should be made.
 
@@
 P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
-and 600 from the Academy of Finland via the FiDiPro award.
+and from the Academy of Finland via the FiDiPro award.
 J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.

### 2017-11-23 Manuscript/manuscript.tex
Add version after Pavel's modifications

@@
 For example, Ca$^{2+}$  is crucial in propagation of neural signals.
-It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}.
+It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{papahadjopoulos90,brouwer15}.
 Ca$^{2+}$ also participates in the T-cell receptor activation.
@@
 The dielectric discontinuity in a lipid bilayer thus arises only
-from the nuclear (orientational) polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
+from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
 Therefore, the same correction for the electronic polarizability can be 
@@
 lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
-provides realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
+provides relatively realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
 We applied the ECC correction 
@@
 The order parameters of the $\alpha$ and $\beta$ carbons in the head group are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the former model, see Fig.~\ref{OrderParameterCHANGESsurf}.
-While both models perform satisfactorily, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations.
+While both models perform relatively well, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations.
 The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
@@
 means that on average water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
-This is in contradiction with most of the available lipid force fields, which predict a small but positive surface excess of sodium at PC lipid bilayers \cite{catte16}.
+This is in contradiction with most of the available lipid force fields, which predict a positive surface excess of sodium at PC lipid bilayers \cite{catte16}.
 
@@
 In conclusion, the results suggest that calcium ions bind specifically to the phosphate oxygens, occasionally interacting also with the carbonyls of the PC lipids.
-This is in a qualitative agreement with previous conclusions from several experimental and computational studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}.
+This is in a qualitative agreement with previous conclusions from several experimental studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}.
 However, the present results suggest, in

### 2017-11-29 Manuscript/manuscript.tex
Change manuscript to ACS format

@@
-%% ****** Start of file aiptemplate.tex ****** %
-%%
-%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
-%%   Version 4.1 of 9 October 2009.
-%%
-%
-% This is a template for producing documents for use with 
-% the REVTEX 4.1 document class and the AIP substyles.
-% 
-% Copy this file to another name and then work on that file.
-% That way, you always have this original template file to use.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% This is a (brief) model paper using the achemso class
+%% The document class accepts keyval options, which should include
+%% the target journal and optionally the manuscript type.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[journal=jpcbfk,manuscript=article]{achemso}
 
-%\documentclass[aip,graphicx]{revtex4-1}
-%\documentclass[aip,reprint]{revtex4-1}
-
-%\usepackage{graphicx}
-
-%\draft % marks overfull lines with a black rule on the right
-%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
-%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}
-
-\documentclass[aip,jcp,twocolumn]{revtex4}
-%\documentclass[aip,jcp]{revtex4}
-%\documentclass{article}
-
-%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional packages needed here.  Only include packages
+%% which are essential, to avoid problems later. Do NOT use any
+%% packages which require e-TeX (for example etoolbox): the e-TeX
+%% extensions are not currently available on the ACS conversion
+%% servers.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% If issues arise when submitting your manuscript, you may want to
+%% un-comment the next line.  This provides information on the
+%% version of every file you have used.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%\listfiles
 
-% Use the \preprint command to place your local institutional report number 
-% on the title page in preprint mode.
-% Multiple \preprint commands are allowed.
-%\preprint{}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional macros here.  Please use \newcommand* where
+%% possible, and avoid layout-changing macros (which are not used
+%% when typesetting).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
 
-\title{Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization} %Title of paper
-
-% repeat the \author .. \affiliation  etc. as needed
-% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
-% Explanatory text should go in the []'s, 
-% actual e-mail address or url should go in the {}'s for \email and \homepage.
-% Please use the appropriate macro for the type of information
-
-% \affiliation command applies to all authors since the last \affiliation command. 
-% The \affiliation command should follow the other information.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Meta-data block
+%% ---------------
+%% Each author should be given as a separate \author command.
+%%
+%% Corresponding authors should have an e-mail given after the author
+%% name as an \email command. Phone and fax numbers can be given
+%% using \phone and \fax, respectively; this information is optional.
+%%
+%% The affiliation of authors is given after the authors; each
+%% \affiliation command applies to all preceding authors not already
+%% assigned an affiliation.
+%%
+%% The affiliation takes an option argument for the short name.  This
+%% will typically be something like "University of Somewhere".
+%%
+%% The \altaffiliation macro should be used for new address, etc.
+%% On the other hand, \alsoaffiliation is used on a per author basis
+%% when authors are associated with multiple institutions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Prague 6, Czech Republic}
-\affiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
+\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
 Tampere, Finland}
@@
 \author{O. H. Samuli Ollila}
-\email[]{samuli.ollila@helsinki.fi}
+\email{samuli.ollila@helsinki.fi}
 %\homepage[]{Your web page}
@@
 Prague 6, Czech Republic}
-\affiliation{Institute of Biotechnology, University of Helsinki}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
 
-% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
-% \noaffiliation is required (may also be used with the \author command).
-%\collaboration{}
-%\noaffiliation
+%\author{Andrew N. Other}
+%\altaffiliation{A shared footnote}
+%\author{Fred T. Secondauthor}
+%\altaffiliation{Current address: Some other place, Othert\"own,
+%Germany}
+%\author{I. Ken Groupleader}
+%\altaffiliation{A shared footnote}
+%\email{i.k.groupleader@unknown.uu}
+%\phone{+123 (0)123 4445556}
+%\fax{+123 (0)123 4445557}
+%\affiliation[Unknown University]
+%{Department of Chemistry, Unknown University, Unknown Town}
+%\alsoaffiliation[Second University]
+%{Department of Chemistry, Second University, Nearby Town}
+%\author{Susanne K. Laborator}
+%\email{s.k.laborator@bigpharma.co}
+%\affiliation[BigPharma]
+%{Lead Discovery, BigPharma, Big Town, USA}
+%\author{Kay T. Finally}
+%\affiliation[Unknown University]
+%{Department of Chemistry, Unknown University, Unknown Town}
+%\alsoaffiliation[Second University]
+%{Department of Chemistry, Second University, Nearby Town}
 
-\date{\today}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The document title should be given as usual. Some journals require
+%% a running title from the author: this should be supplied as an
+%% optional argument to \title.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title[An \textsf{achemso} demo]
+  {Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Some journals require a list of abbreviations or keywords to be
+%% supplied. These should be set up here, and will be printed after
+%% the title and author information, if needed.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\abbreviations{IR,NMR,UV}
+\keywords{American Chemical Society, \LaTeX}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The manuscript does not need to include \maketitle, which is
+%% executed automatically.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The "tocentry" environment can be used to create an entry for the
+%% graphical table of contents. It is given here as some journals
+%% require that it is printed as part of the abstract page. It will
+%% be automatically moved as appropriate.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{tocentry}
+
+Some journals require a graphical entry for the Table of Contents.
+This should be laid out ``print ready'' so that the sizing of the
+text is correct.
+
+Inside the \texttt{tocentry} environment, the font used is Helvetica
+8\,pt, as required by \emph{Journal of the American Chemical
+Society}.
+
+The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
+permitted for  \emph{Journal of the American Chemical Society}
+graphical table of content entries. The box will not resize if the
+content is too big: instead it will overflow the edge of the box.
+
+This box and the associated title will always be printed on a
+separate page at the end of the document.
+
+\end{tocentry}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The abstract environment will automatically gobble the contents
+%% if an abstract is not used by the target journal.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{abstract}
-  % insert abstract here
 Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
@@
 
-%\pacs{}% insert suggested PACS numbers in braces on next line
+%\begin{abstract}
+%  This is an example document for the \textsf{achemso} document
+%  class, intended for submissions to the American Chemical Society
+%  for publication. The class is based on the standard \LaTeXe\
+%  \textsf{report} file, and does not seek to reproduce the appearance
+%  of a published paper.
 
-\maketitle %\maketitle must follow title, authors, abstract and \pacs
+%  This is an abstract for the \textsf{achemso} document class
+%  demonstration document.  An abstract is only allowed for certain
+%  manuscript types.  The selection of \texttt{journal} and
+%  \texttt{manuscript} will determine if an abstract is valid.  If
+%  not, the class will issue an appropriate error.
+%\end{abstract}
 
-% Body of paper goes here. Use proper sectioning commands. 
-% References should be done using the \cite, \ref, and \label commands
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Start the main part of the manuscript here.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 % If you have acknowledgments, this puts in the proper section head.
-\begin{acknowledgments}
+\begin{acknowledgement}
 % Put your acknowledgments here.
@@
 Helsinki Institute of Life Science (Instruct-HiLIFE).
-\end{acknowledgments}
+\end{acknowledgement}
 
-%\newpage
-%\appendix
+\begin{suppinfo}
 
-% Create the reference section using BibTe
+%A listing of the contents of each file supplied as Supporting Information
+%should be included. For instructions on what should be included in the
+%Supporting Information as well as how to prepare this material for
+%publications, refer to the journal's Instructions for Authors.
+
+The following files are available free of charge.
+\begin{itemize}
+  \item SI.pdf: Simulation details and further analysis.
+%  \item Filename: brief description
+\end{itemize}
+
+\end{suppinfo}
+
 \bibliography{refs.bib}
@@
 \end{document}
-%
-% ****** End of file aiptemplate.tex ******

### 2017-12-08 Manuscript/manuscript.tex
New isocontour figure with a slighlty improved caption. Not final, will update it with more smooth contours.

--- 2017-11-23 Manuscript/manuscript.tex
@@
 membranes. Current estimates of binding affinities and
-stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
-computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
+stoichiometries of cations are, however, controversial due to limitations in the available experimental and
+theoretical methods. Experimentally one can assess these parameters by titrating
+several membrane properties as a function of the ion concentration. However, the
+interpretation of the experiments relies on theoretical models, which give
+the best conception of the ion binding as direct experiments are not available. 
+Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
+resolution, therefore offering all the necessary information to interpret
+experimental data without the need to resort to simplified models. However, the accuracy
+of the available lipid models when interacting with ions is not sufficient for
+such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
 1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
 polarization as a mean field correction, known as the electronic continuum correction (ECC).
-This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
+This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
 Our improved ECC-POPC model reproduces not only the
 experimentally measured structural parameters for the ion-free membrane, but also the
-response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
-Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
-bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
+response of lipid headgroup to a bound positive charge, and the binding affinities of 
+Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
+bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens,
 in agreement with the previous interpretations of the experimental spectroscopic data.
-The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
-equal probabilities, suggesting more complex binding stoichiometries than those from simple models
-used to interpret the NMR data previously.
-The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
-of complex biochemical systems at cellular membranes.
+The new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
+equal probabilities, suggesting more complex binding stoichiometry than simple models
+previously used to interpret the NMR data.
+The results of this work pave the way to quantitative MD simulations of complex
+biochemical systems with realistic electrostatic interactions in the vicinity of cellular
+membranes.
 \end{abstract}
@@
 \section{Introduction}
-Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}.
-Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
-For example, Ca$^{2+}$  is crucial in propagation of neural signals.
+Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}.
+Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
+Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface.
+For example, Ca$^{2+}$  is crucial in neural signal propagation.
 It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}.
 Ca$^{2+}$ also participates in the T-cell receptor activation.
-It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}.
-Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
-Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions.
-Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.
+It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013Ca2}.
+A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
+Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions.
+Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.
+\todo{Hector: I left one unknown citation for Pavel.}
 
-Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
+Direct measurements of ion-membrane interactions in real biological systems are difficult. 
 Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
-For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
+For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
 The details of ion binding are, however, not fully consistent in the literature.
-Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
-In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}.
-Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
+Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mainly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
+On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
+On the other hand, most atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
+For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
+%{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}
 
-Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements.
-A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}.
-Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results.
-Analyzing massive amounts of data collected within an open collaboration approach, 
-it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
+Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
+A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept~\cite{seelig87}.
+Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results.
+Analyzing massive amounts of data collected by an open collaboration method, 
+it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
 
-In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
-via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
-To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
-within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
-Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
-Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model.
-This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}.
+In this work, we improve the cation binding in a POPC bilayer in MD simulations 
+via the implicit inclusion of the electronic polarizability in the polar region of phospholipids. 
+For this, we apply an electronic continuum correction (ECC)~\cite{leontyev11}, 
+in which scaling of the atomic partial charges is used to account for the missing electronic polarizability in MD simulations. 
+Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
+As a starting template for our POPC model, we will use the Lipid14~\cite{dickson14}.
+This force field provides one of the best available descriptions of cation binding \cite{catte16}.
 The newly developed ECC-POPC model reproduces the experimentally
-measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.
+measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while significantly improving the membrane binding affinities to sodium and calcium cations.
 
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations.
-In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
+The lack of electronic polarizability in standard MD simulation force fields has been considered a serious issue since the early days of lipid bilayer simulations.
+In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by including the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
 Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
@@
 
-According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
+According to ECC, the electronic polarizability can be modeled in non-polarizable MD in a mean-field way by 
 embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
-Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
+Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
 \begin{equation}
@@
 \end{equation}
-for the ECC corrected charges. 
+for the ECC corrected charges in simulation. 
 Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
-It is important to note that the value of the
+This scaling factor successfully improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
+It is important to note that the 
 high frequency dielectric constant 
 is around 2 for almost any biologically relevant environment \cite{leontyev11}.
-The dielectric discontinuity in a lipid bilayer thus arises only
-from the nuclear (orientational) polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
-Therefore, the same correction for the electronic polarizability can be 
-applied throughout the lipid bilayer/aqueous solution interface.
+The dielectric discontinuity in a lipid bilayer arises
+from the orientational polarization of the molecules. 
+Thus, the same correction for the electronic polarizability can be 
+applied throughout the lipid bilayer interface.
 
-While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
+While using a scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
 Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
-Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}.
-Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
-Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$.
-Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
+Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}.
+Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
+Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$.
+Instead, a consistent scaling factor would lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
 
-Here we develop a ECC-POPC lipid model that accurately describes binding 
-of sodium and calcium ions to the POPC  lipid bilayer. 
-The Lipid14~\cite{dickson14} force field 
-(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
-point since 
-it provides the most accurate response of the head groups to ions among the available 
+Here we develop ECC-POPC lipid model that accurately describes the binding 
+of sodium and calcium ions to the lipid bilayer. 
+The Lipid14~\cite{dickson14} force field parameters 
+(available in a Gromacs format from Ref. \citenum{lipid14files}) were used as a starting 
+point. 
+This model provides the best response of the head group to ions among the available 
 lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
-provides realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
+has a relatively realistic head group, glycerol backbone, and acyl chain structure~\cite{dickson14,botan15}.
 We applied the ECC correction 
-to the Lipid14 model of POPC by scaling 
+to the Lipid14 model of POPC by scaling of 
 partial charges of the head group, glycerol 
@@
 To reproduce the experimental ion binding affinities,
-we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
-The ion binding affinity was benchmarked against experiments
+we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
+The ion binding affinity was evaluated against experiments
 using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
@@
 Scaling down the partial charges reduced the ion binding affinity.
-We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
-which is only slightly higher than the ECC one ($f_q=0.75$).
+We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
+which is only slightly higher than the theoretical one ($f_q=0.75$).
 Directly applying
@@
 (64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
-reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
-reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
-the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
-factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
-solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}.
-After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
+reduced hydration of the lipid head group region after scaling charges which effectively
+reduces head group polarity. We solve this difficulty by reducing the effective radii of
+the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a
+factor of $f_\sigma = 0.89$. The same solution was taken for the ECC-ions in aqueous
+solvent \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the $\sigma$ parameters,
+the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
 
@@
 Comparing MD simulation to NMR experiments, we can validate the ion
-binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
-This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
+binding affinity in lipid bilayer simulations using the ''electrometer concept''~ \cite{seelig87, catte16}.
+This method relies on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
+%Comparing this response is an excellent way to validate force fields as both NMR and MD can accurately determine order parameters ~\cite{catte16, ollila16}.
 The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
-From MD simulations the order parameters can be calculated using the relation
+From MD simulations the order parameters can be calculated using the definition
 \begin{equation}\label{OP}
@@
 \end{equation}
-where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.
+where $\theta$ is the angle between the C-H bond and membrane normal. Angular brackets point to the average over all sampled configurations.
 
@@
 \end{equation}
-where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.
+where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and position of the bound charge.
 
@@
 This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
-To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
+To calibrate the response of the head group order parameter to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
 DHAB\\[0.5cm]
@@
 \vspace{0.5cm} \\
-is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end.
-Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
+is a cationic surfactant having two acyl chains and bearing a unit charge in the hydrophilic end.
+Thus, it is expected to locate in the bilayer similarly to the phospholipids, and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
 
-\subsection{Salt concentrations and binding affinities}
-When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
-Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
+\subsection{Salt concentrations and binding affinity}
+NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
+Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
 while others use atomic absorption spectroscopy and report the salt concentration in
 the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
-we stick to the latter definition by estimating the salt concentration in the aqueous
+%we use the latter definition. In particular,
+we measure the latter concentration in the aqueous
 bulk region using the farthest point from the lipid bilayer in the aqueous phase.
-Note that the former definition was used by some of us previously.~\cite{catte16}.
-Although these two definitions provide somewhat different values when applied
-to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.
+Notice that the former concentration was used by Catte et al.~\cite{catte16}.
+Although these two definitions vary when applied
+to Ca$^{2+}$, they do not significantly affect the conclusions of this work.
 
 To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}.
-Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region.
+Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing plane between the membrane interior and the water bulk region.
 In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.
-Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
+Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
  \begin{equation}\label{surfexcess}
@@
 \subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
-The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors.
+The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and \mbox{x-ray} scattering experiments by the scattering form factors.
 NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution.
@@
 \ref{OP}.
-Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule).
-Form factors were calculated using a relation
+Scattering form factors validate the dimensions of the lipid bilayer (e.g., thickness and area per molecule).
+Form factors were calculated using
 
@@
 The robustness of our approach with other water models (OPC \cite{Izadi14},
-OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3P-FB and TIP4P-FB~\cite{Wang2014}, and TIP4P/2005 \cite{Abascal2005})
-was verified and is presented in the Supporting Information (SI).
+OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
+is tested in the SI.
 Sodium, calcium, and chloride ions were modeled as ECC-ions
 with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
-forming the POPC lipid were derived in this work starting from the Lipid14 force field~\cite{dickson14}.
+forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
 For comparison, simulations with the Lipid14 model~\cite{dickson14} and
 ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
-were also performed. The TIP3P water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
-Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and the TIP3P~\cite{jorgensen83} water model were taken directly from
+were also performed. The TIP3p water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
+Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
 Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.  
+% SAMULI: I putted the trajectory without ions back, because it has been used as well.
 
 MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
-in this work are summarized in Table~S1 in the SI. 
+in this work are summarized in Table~S1. 
 Compatibility with the openMM simulation package \cite{openmm7} was also
-tested and is presented in the SI. 
+tested in SI. 
 The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
@@
 \subsubsection{Simulations of POPC bilayers with cationic surfactants}
-The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
-The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters.
+The topology of dihexadecyldimethylammonium was created with the automated topology builder \cite{malde11}.
+The AmberTools program~\cite{amber} was then used to generate the Amber-type force field parameters.
 These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
-The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.
+The partial charges were then manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
+For simulation with ECC-POPC, the partial charges of the headgroup were scaled, i.e., $f_q=0.8$, and the $\sigma$ of the scale atoms reduced as in POPC-ECC.
 
@@
 All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
-Chloride counter ions were used in all the simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
-For POPC, either Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987} parameters  were used.
-The first 20~ns of the total simulation time of 200~ns was considered as an equilibration period and was thus omitted from the analysis.
-We checked that a sufficient lipid  exchange occurred during the simulations.
+Chloride counter ions were used in simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
+For POPC either the Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987}  were used.
+The first 20~ns of the total simulation time of 200~ns was considered as an equilibration time and was omitted from the analysis.
+Sufficient lipid neighbor exchange occurred during the simulations.
 The trajectories and simulation files for unscaled systems are available from
@@
 
-\todo{Provide the parameters for ECC-surfactant somewhere.
+\todo{Provide the parameteres for ECC-surfactant somewere.
       Joe: They will appear along with the Zenodo upload of the trajectories.}
@@
 
-\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}
+\subsection{Structural parameters of pure ECC-POPC model membrane are in agreement with experiments}
 
@@
 First, we present results for bilayers in pure water.
-The ECC-POPC and Lipid14 models both reproduce the experimental X-ray scattering form factors
-of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
+The ECC-POPC and Lipid14 models both reproduce the experimental x-ray scattering form factors
+of a POPC bilayer with a comparable accuracy in Fig.~\ref{simVSexpNOions}.
 The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
 experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
-is $\approx$1\AA{} smaller than the experimental one.
-The values of the area per lipid of the ECC-POPC model vary slightly
-when simulated with different water models (i.e., within the interval of 62.2--66.8 \AA{}, see Table~S2 in SI),
+is $\approx$1\AA{} smaller
+The values of the area per lipid of the ECC-POPC model vary
+when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI),
 while still being close to the experimentally reported values.
 We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
-lipid bilayer with a comparable accuracy to other state-of-the-art lipid models~\cite{ollila16}.
+lipid bilayer with the comparable accuracy to other state of the art lipid models~\cite{ollila16}.
 
-Similarly, the acyl chain order parameters of the ECC-POPC model, as well as those of the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
-Notably, the experimentally measured forking and small order parameter values of the $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
-This feature has been suggested to indicate that the carbonyl of the {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in the {\it sn}-1 chain, which orients more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
-This arrangement, which is not fully reproduced by other available lipid models~\cite{ollila16}, may be a relevant feature for the ion binding details.
+Similarly, the acyl chain order parameters of the ECC-POPC model, as well as the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
+Notably, the experimentally measured forking and small order parameter values of $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
+This feature has been suggested to indicate that the carbonyl of {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in {\it sn}-1 chain, which would orient more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
+While this arrangement may be a relevant feature for the ion binding details, it is not fully reproduced by other available lipid models~\cite{ollila16}.
+%This is expected as the ECC model does not modify the already highly optimized parameters of this region in Lipid14
 
-The order parameters of the $\alpha$ and $\beta$ carbons in the head group are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the former model, see Fig.~\ref{OrderParameterCHANGESsurf}.
-While both models perform satisfactorily, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations.
+The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-POPC model, see Fig.~\ref{OrderParameterCHANGESsurf}.
+Considering the available experimental evidence, it is not possible to suggest which of the two models provides more realistic headgroup orientation.
 The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
-The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
+The accuracy of both models in the glycerol backbone region is equally footed with any other state of art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
 
@@
   \caption{\label{OrderParameterCHANGESsurf}
-    The changes of head group order parameters and P-N vector orientation as a function of
+    The changes of headgroup order parameters and P-N vector orientation as a function of
     a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
@@
 
-Before studying the sodium and calcium ion binding affinities, we quantify the response of the head group order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
-These mixtures have a well-defined amount of bound charge per PC, namely, the molar fraction of cationic surfactants.
-This is due to the ability of dihexadecyldimethylammonium to directly insert in the lipid bilayer due to its two hydrophobic acyl chains.
-Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid head group order parameters to the amount of bound charge in simulations \cite{scherer89}.
+Before studying the sodium and calcium ion binding affinities, we quantify the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
+These mixtures have a well-defined amount of bound charge per PC, i.e., the molar fraction of cationic surfactants.
+This direct relation results from the ability of dihexadecyldimethylammonium to be inserted in the lipid bilayers as any lipid because of its two hydrophobic acyl chains.
+Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid headgroup order parameters to the amount of bound charge in simulations \cite{scherer89}.
 
-The changes of the head group order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
+The changes of the headgroup order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
 An approximately linear decrease of the order parameters, as expected from Eq.
 \ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
-The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters to the bound positive charge is too strong.
-In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being only slightly underestimated for the $\beta$ segment.
+The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters is too large to the bound positive charge.
+In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.
 
 In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
-As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in the PC lipid bilayer.
+As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in a PC lipid bilayer.
 The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
-For example, the addition of 50\% mole fraction of the cationic surfactant leads to a decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only of 11$^{\circ}$ in the ECC-POPC model.
+For example, the addition of 50\% mole fraction of the cationic surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only 11$^{\circ}$ in the ECC-POPC model.
 The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model.
-The weaker sensitivity of the P-N vector angle response in the ECC-POPC model is in better agreement with experiments.
+The lower sensitivity of the P-N vector angle response in the ECC-POPC model agrees better with experiments.
+%The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules.
+%{\color{red}., as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
 
@@
 
-Changes of the lipid bilayer head group order parameters extracted from simulations and
+Changes of the lipid bilayer head group order parameters from different simulations and
 experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
@@
 The experimentally measured small order parameter
-changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) 
+changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) %and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81, altenbach84}.
 are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
 However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
-Replacing \AA{}qvist ions with ion parameters by Dang et al.~\cite{smith94, chang1999, dang2006}
+%
+%A recent comparison of different simulation
+%models to the experimental data revealed that most force fields significantly
+%overestimate Na$^+$ ion binding to PC lipid bilayers and that none of the available
+%models correctly reproduces the details of the binding of Ca$^{2+}$ ions which is
+%grossly overestimated \cite{catte16}.
+%
+%Since artifacts were reported
+%in simulations involving \AA{}qvist ions is water \cite{auffinger07}, we tested the
+%
+Replacing \AA{}qvist ions with ions by Dang et al.~\cite{smith94, chang1999, dang2006}
 or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
@@
 
-The results from simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as a function of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
-Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
+%To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
+%polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
+
+The results from the simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
+Considering that we can also reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC-model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
 Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-Also the changes of the P-N vector angle are too pronounced for the Lipid14 model,
+%\todo{HECTOR: I do not believe in the content of the following sentence as it is.
+How do we know that the changes of the P-N angle are too pronounced for Lipid14?.
+There is no direct correlation between angle and order parameter.}
+% SAMULI: The sentence is now modified.
+Also the changes of the P-N vector angle are more pronounced for the Lipid14 model,
 for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
@@
 for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$). 
+%\todo{HECTOR: I have a problem when comparing fig.
+2 and 3.
+If I read it right the behavior when adding calcium and surfactant is not the same.
+Is the latter changing the area per lipid and therefore the surface charge density?
+If this will be the case then besides the effect of the charge we will have a packing effect.
+This is more relevant for the more concentrated cases.
+Shouldn't we try to discuss the two figures and what the differences and similarities mean?
+Given our previous work, should we check the behavior respect the charge density?}
+%SAMULI: Not now, hopefully in the future.
 
 Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge.
-It is plausible to assume that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance of validation of the lipid headgroup order parameter response to the bound charge.
+It is plausible that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance to validate the lipid headgroup order parameter response to the bound charge.
 
 Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI.
-In general, the performance of ECC-POPC with any of the tested water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3P model.
+In general, the performance of ECC-POPC with any of the water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3p model.
 
-\subsection{Binding affinities of Na$^+$ and Ca$^{2+}$ cations to the POPC membrane}
+\subsection{Binding affinity of Na$^+$ and Ca$^{2+}$ cations to POPC membrane}
 \label{sec:affinity}
@@
 
-Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
+Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
 The values of $\Gamma_{\rm ion}^{\rm water}$
@@
 to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
-As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess of calcium, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller for the ECC-POPC model than for the other models, 0.13--0.35~nm$^{-2}$.
-Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}.
-This
-means that on average water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
-This is in contradiction with most of the available lipid force fields, which predict a small but positive surface excess of sodium at PC lipid bilayers \cite{catte16}.
+As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess for the ECC-POPC model, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller than for the other models, 0.13--0.35~nm$^{-2}$.
+%\todo{HECTOR: I have deleted the sentence relating the uncertainty and the error as in my opinion there are other factors which are more important.
+The most important the slow convergence.
+SAMULI: I think that this was a good idea.}.
+Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having a negative value $\Gamma_{Na}^{water} = (-0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens})
+meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
+This is in contradiction with most of the available lipid force fields, which predict a positive binding of sodium to PC lipid bilayers \cite{catte16}.
+%\todo{HECTOR: do we need to add a NaCl table or an extended table for CaCl2 using different concentrations in the SI? I think this is very important information
+%  SAMULI: I agree that this is important information, but we have quite a lot of stuff in the paper already. Maybe we could leave this for the future studies.}
 
@@
     for different force fields.
+    %Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
+    %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
     In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
@@
 We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
-A maximum distance of 0.3~nm from any lipid oxygen is used to define a bound calcium.
-The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in direct contact with phosphate oxygens.
-From these ions, only one third (32\%) also interacts with the carbonyl oxygens, while the interaction of calcium ions with carbonyl oxygens only is rare (1\%).
-The most abundand interaction scenarios between Ca$^{2+}$ ions and phosphate oxygens are visualized using the probability density isocontours in Fig.~\ref{fig:volmaps}.
+A maximum distance of 0.3~nm from any lipid oxygen is used to determine a bound calcium.
+The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in contact with the phosphate oxygens.
+From those only one third (32\%) also interacts with the carbonyl oxygens.
+The interaction of calcium ions with only carbonyl oxygens is rare (1\%).
+Neither individual lipids nor source chain of carbonyl is distinguished in the analysis.
+The most likely interactions between Ca$^{2+}$ ions and phosphate oxygens are visualized with the probability density isocontours in Fig.~\ref{fig:volmaps}.
 While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.
@@
   \centering
-  \includegraphics[width=8.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
+  \includegraphics[width=8.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
+%  \includegraphics[width=8.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
 %  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
@@
   \caption{\label{fig:volmaps}
-      Isocontours of probability density of \ce{Ca^{2+}} with respect to 
-      the phosphate oxygens of POPC from ECC-POPC simulation.
-      The probability density was evaluated around a single lipid, 
+      Isocontours of spatial number density of \ce{Ca^{2+}} (cyan, 0.0015~\AA$^{-3}$) 
+      and POPC carbonyl oxygen atoms (red, 0.0075~\AA$^{-3}$). 
+      Calcium cations localize mostly around phosphate oxygens.
+      Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
+      and it is contributed more by other neighbouring phospholipids than by the same lipid. 
+      Transparent structures are shown to depict the variability of POPC configurations. 
+      The number density was evaluated for each lipid, 
       after its strucutral alignment using only phosphate group. 
@@
   }
-  \todo{The figure should be updated as discussed last week.}
+  \todo{The figure should be updated as discussed last week. Joe: Will only smoothen the isosurfaces now. }
 \end{figure}
 
-Even though \ce{Na+} ions do not bind strongly to a POPC bilayer, they still interact mostly with its oxygen moieties.
-The results from a simulation at a $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions at the bilayer interact with phosphate oxygens of POPC only and 20\% with carbonyl oxygens only, with the remaining 25\%, is interacting with both negatively charged groups.
+Even though \ce{Na+} ions do not specifically bind to a POPC bilayer, they still interact mostly with its oxygen moieties.
+The results from the simulation at $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions interact with only phosphate oxygens of POPC and 20\% with only carbonyl oxygens, and the rest, 25\%, is interacting with both.
 
-In conclusion, the results suggest that calcium ions bind specifically to the phosphate oxygens, occasionally interacting also with the carbonyls of the PC lipids.
-This is in a qualitative agreement with previous conclusions from several experimental and computational studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}.
-However, the present results suggest, in
-agreement with experiments, an overally weaker binding to the bilayer, in particular with a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
-Sodium ions, which do not exhibit any appreciable affinity for the bilayer, also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
+In conclusion, the results suggest that calcium ions bind specifically to phosphate oxygens, occasionally interacting also with carbonyls.
+This is in good agreement with previous conclusions from several experimental and computational studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}, but suggesting a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
+Sodium ions also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
+%\todo{This last sentence is weak and not relevant.
+If you want to keep it go ahead, but then I would ask you a simple question: Is it incompatible to have no overall binding (this relates to the whole surface volume) with having small cavities in the membrane with an increase Na+ concentration.
+If the answer is negative maybe the statement is incorrect and who knows it might be relevant for something.
+After all even Ca2+ affinity changes in the presence of sodium.} The physiological relevance of these details of sodium interactions is, however, unclear due to its overall very weak binding affinity.
 
@@
 
-Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} as employed in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
-In particular, NMR data concerning the PC headgroup order parameters response and atomic absorption spectra were explained best using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-Nevertheless, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering \ce{CaCl2} at low concentrations only~\cite{macdonald87}.
+Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} used in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
+In particular, NMR data about PC headgroup order parameters response and atomic absorption spectroscopy were best explained using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
+However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2} at low concentrations~\cite{macdonald87}.
+%\todo{HECTOR: Why is the following sentence remotely significant.
+I will vote for removing it} Note that the results based on these binding models and the measured headgroup order parameter changes are also included in the tabulated experimental binding constants~\cite{marsh13}.
 
-In this work, we reproduce the same experimental data used to infer binding stoichiometries employing our ECC-POPC model.
-Thanks to our simulations, we have a direct access to atomistic details of the binding stoichiometry without a need for any binding model as employed for interpreting in experiments~\cite{altenbach84, macdonald87}.
-To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
+In this work, we reproduce the same experimental data used to calculate binding stoichiometry with our ECC-POPC model.
+Using our simulations, we have direct access to exact details of the binding stoichiometry without requiring any middle model as in experiments~\cite{altenbach84, macdonald87}.
+To evaluate the relative propensities for each stoichiometric complex (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
 Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
-We found the largest propensity for the 1:2 complex (41\%), with probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~being only slightly lower.
-This suggests a more complex binding model than considered in a simple 1:2 ternary complex model previously.
-Nevertheless, with a broad brushstroke, the simulation data can be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
+We found the largest propensity for the 1:2 complex (41\%).
+However, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower.
+This variability suggests a more complex binding model than the previously considered in a simple ternary complex model.
+With a broad brushstroke, the simulation data can, however, be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
 This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).
@@
 
-\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in the POPC membrane}
+\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in POPC membrane}
 
-Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current force fields, such  as CHARMM36 and Slipids force fields~\cite{javanainen17}.
-This suggests that at least several microseconds are required to reach the ion binding/unbinding equilibrium.
-To quantify the exchange of ions between the membrane and aqueous solution in simulations, we evaluated residence times of ions bound to the membrane.
-Within our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
+Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current state of the art force fields, e.g., CHARMM36 and Slipids force fields~\cite{javanainen17}.
+This suggests that at least several microseconds are required to reach the binding/unbinding equilibrium.
+%\todo{HECTOR: I disagree with the following lines.
+In the case of most of the unscale force fields the issue is that the binding of Calcium to the membrane is very strong.
+Therefore, the reorganization takes too long leading to a very slow equilibration process.
+In this case, the size of the hill in the choline region is nothing compared to the size between the bound and solvated state which is the real bottleneck.
+Even in the case of the scaled simulations, I believe the same applies but to a lesser extent.
+I might be wrong in this last part.
+If you do not agree we can discuss it tomorrow via skype.} {\color{red} This is in line with the density profiles of \ce{Ca^{2+}} from simulations in Fig.~\ref{fig:cacl-dens}, which all show a region of a very small calcium density around the level of choline groups in POPC (i.e., around $2.5\,\mathrm{nm}$).
+In contrast, simulations with  ECC-lipids exhibits only a relatively shallow local minimum at that level.
+This suggests that the smaller barrier for the exchange of calcium ions shall lead to their faster exchange with the solvent within ECC-lipids.}
+To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated residence times of ions bound to the membrane.
+In our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
 
+%\todo{HECTOR: I do fully understand the content of this paragraph. I will discuss this with Joe tomorrow. Please be carefull with my changes.} 
 The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
-ECC-POPC and CHARMM36 (simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
-In the CHARMM36 simulation, a significant number of the calcium ions is bound to the membrane for the whole length of the trajectory (800~ns).
+ECC-POPC and CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
+In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory (800~ns).
 In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
 where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
-residence time is around 150~ns, which is below the total length of the simulation used for analysis, i.e., 200~ns.
+residence time is around 150~ns, which is well below the total length of the simulation used for analysis, i.e., 200~ns.
 Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
-headgroup is of the order of $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
-is yet another order of magnitude faster, with 90\% of the residence times smaller than~1~ns and the longest residence time being~6~ns.
+headgroup is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}. The exchange of \ce{Na+} ions at the POPC membrane
+is another order of magnitude faster, yielding 90\% of the residence times smaller than~1~ns, with the longest residence time being~6~ns.
 
-In summary, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within the $\sim$100~ns timeframe, which is significantly faster than observed in simulations emloying most of the presently available lipid models~\cite{javanainen17}.
-Sodium cations exhibit an even more rapid exchange between the membrane and the aqueous solution.
-Our results suggest that simulations with a length of several hundreds of nanoseconds are sufficient to simulate alkali and alkali earth ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
-This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and, in particular,  calcium cations \cite{javanainen17, catte16}.
+In conclusion, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within $\sim$100~ns timeframe, which is significantly faster than observed in most of the available lipid models~\cite{javanainen17}.
+Sodium cations exhibit an even more rapid exchange.
+This all suggests that simulations with a length of several hundreds of nanoseconds are sufficient to simulate ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
+This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and calcium cations \cite{javanainen17, catte16}.
+%Such a finding changes the point of view on binding of calcium to PC membranes from
+%a very strong long-term stable binding with rare exchanges to 
+%a relatively frequent exchange of clacium cations 
+%in equilibrium between membrane and solvent.
 
@@
 
-In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a force field MD simulation, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
-This ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
-While the structural details of a POPC lipid bilayer in pure water simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the new model in addition reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ at varying concentrations.
-It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
+In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a classical MD simulation force field, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
+While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations.
+It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}.
+The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
 
 The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model.
-In line with previous interpretations of the experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens, with interactions to the carbonyls being of a secondary importance.
-However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data previously, within which one calcium binds to two POPC molecules~\cite{altenbach84}.
-While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to be relatively abundant and also almost equally likely to occur.
-While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the detailed nature of calcium binding to phospholipid bilayers observed in the present simulation.
+In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens.
+However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data, where one calcium binds to two POPC molecules~\cite{altenbach84}.
+While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to relatively abundant and also almost equally likely.
+While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the complex nature of calcium binding to phospholipid bilayers observed in the present simulation.
 
-Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions, including a mean-field account for electronic polarization effects.
-To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules needs to be verified and, if necessary, further adjustments following the ECC concept to the force fields of the biomolecules interacting with the lipids should be made.
+Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions.
+To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules should be verified and, if necessary, further adjustments following ECC concept to the force fields of the interacting molecules with the lipids should be addressed.
 
@@
 P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
-and from the Academy of Finland via the FiDiPro award.
+and 600 from the Academy of Finland via the FiDiPro award.
 J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.

### 2017-12-11 Manuscript/SI.tex
SI cleanup

@@
 \begin{table}[!h]
-  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si}
-  {\color{red}Joe, please deliver the citations for all experimental area per molecule values or remove them.}}
+  \caption{Area per lipid (APL) from different models of POPC with no ions\label{tab:apls_si} }
   \begin{tabular}{l|c c}
@@
     \hline
-    ECC-lipids &        &  \\
-    %Lipid14ecc0.80+sigma0.875 &        &  313    \\
-    %($4.6\cdot 5.1 \, \mathrm{nm}^2$), 72 lipids patch, OPC3           & 63.2   &   313      \\
+    ECC-POPC &        &  \\
     ~ SPC/E          & 63.2$\pm$ 0.6   &  300       \\
@@
     ~ TIP4p-FB       & 65.6$\pm$ 0.6   &  313       \\
-    %($9.2\cdot 10.2 \, \mathrm{nm}^2$), 288 lipids patch           & 65.5   &  313       \\ %% not done for this model with f_sigma=0.89
-    %oMM small patch           & 63.65  &         \\
-    %oMM 4xbig patch           & 63.7   &         \\
     \hline
-    experiment   & 62.7  &  293    \\
-    experiment \cite{kucerka11} & 64.3  &  303    \\
-    experiment  & 67.3  &  323    \\
-    experiment  & 68.1  &  333    \\
+    experiment \cite{kucerka11}  & 62.7$\pm$ 1.3  &  293    \\
+                                 & 64.3$\pm$ 1.3  &  303    \\
+                                 & 67.3$\pm$ 1.3  &  323    \\
+                                 & 68.1$\pm$ 1.4  &  333    \\
     %experiment POPE  & 56.6 &  303    \\
@@
     the prediction of the ternary complex model.
-    {\color{red} Joe, please change the axes labels to correspond the notation in the text, i.e. $C_b$ and $C_I$,
-    and write in the title of simulation fit (red dashed line) that it is the fit to simulation.}
+    %{\color{red} Joe, please change the axes labels to correspond the notation in the text, i.e. $C_b$ and $C_I$,
+    %and write in the title of simulation fit (red dashed line) that it is the fit to simulation.}
     %Ternary complex binding model of \ce{Ca^{2+}} to a POPC membrane 
@@
     a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
-    from simulations with the ECC-lipids models
+    from simulations with the ECC-POPC model
     simulated with GROMACS~5.1.4 \cite{Abraham15} and openMM~7 \cite{openmm7} 

### 2017-12-11 Manuscript/manuscript.tex
Manuscript TODO cleanup

@@
   } 
-    \todo{Joe: I'm fine with the name change from ECC-lipids to ECC-POPC, but we have to be consistent throughout the text.
-I changed all accurences of ECC-lipids to ECC-POPC in the text, I yet need to change all figure legends then.}
 \end{figure}
@@
 
-\todo{Dynamics check is missing: MSD (Hector/Joe)}
+%\todo{Dynamics check is missing: MSD (Hector/Joe)}
 
@@
   }
-  \todo{The figure should be updated as discussed last week. Joe: Will only smoothen the isosurfaces now. }
+  %\todo{The figure should be updated as discussed last week. Joe: Will only smoothen the isosurfaces now. }
 \end{figure}

### 2017-12-11 Manuscript/manuscript.tex
Manuscript updates : Figure caption polishing and clarifying.

@@
     Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K.
+    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K. \\
     Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
@@
     The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
-    Bottom: The chemical structure of POPC and the labeling of the carbon segments.
+    Open/closed symbols are used for palmitoyl/oleoyl chains of POPC. \\
+    Bottom: Chemical structure of POPC and the labeling of hydrocarbon segments.
   } 
@@
     Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
-    from simulations at 313 K together with experimental data 
+    as a function of the CaCl$_2$ concentration in a bulk water ($C_{ion}$)
+    from simulations at 313~K together with experimental data 
     (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    The error estimate for bulk concentrations was approximately 10\,mM.
-    Exception was the order of magnitude larger error in
-    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) because the 
-    small simulation box. 
-    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
-    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
+    The error estimate for bulk concentrations is approximately 10\,mM.
+    The order of magnitude larger error in the
+    simulation with Lipid14 and ECC-ions is due to unconverged bulk densities  
+    (shown if Fig.~\ref{fig:cacl-dens}) limited by the simulation box. 
+    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298~K are taken directly from
+    Refs.~\citenum{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
   }
@@
   \centering
-  \includegraphics[height=19.5cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
+  \includegraphics[height=19.9cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
   \caption{\label{fig:cacl-dens}
@@
       Isocontours of spatial number density of \ce{Ca^{2+}} (cyan, 0.0015~\AA$^{-3}$) 
-      and POPC carbonyl oxygen atoms (red, 0.0075~\AA$^{-3}$). 
-      Calcium cations localize mostly around phosphate oxygens.
+      and POPC carbonyl oxygen atoms (red, 0.0075~\AA$^{-3}$, all POPC lipids contribute). 
+      Calcium cations localize mostly around phosphate oxygens (red and bronze).
       Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
@@
       after its strucutral alignment using only phosphate group. 
+      Carbon atoms are cyan, hydrogen atoms white, oxygen atoms red, nitrogen blue.
       %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.

### 2017-12-12 Manuscript/manuscript.tex
Added reference for MDAnalysis, cited VMD.

@@
       after its strucutral alignment using only phosphate group. 
+      MDAnalysis \cite{mdanalysis2011} library was used for 
+      the calculations of the structural alignment and the spatial number density. 
+      VMD \cite{hump96} was used for visualisation. 
       Carbon atoms are cyan, hydrogen atoms white, oxygen atoms red, nitrogen blue.
@@
 
-This work can be reached as a repository containing all data at \url{zenodo.org:\dots\dots\dots}.
+This work can be reached as a repository containing all data and scripts for analysis at \url{zenodo.org:\dots\dots\dots}.
 

### 2017-12-12 Manuscript/manuscript.tex
Turned fig1 into a single image (PDF)
Improved isocontour figure and its caption.

@@
   \centering
-  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89.pdf}
-  \includegraphics[width=7.4cm]{../Fig/POPCstructure.eps}
+  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89_POPC-struct.pdf}
+  %\includegraphics[width=7.4cm]{../Fig/POPCstructure.eps}
   %\includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
@@
   \centering
-  \includegraphics[width=8.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
+  \includegraphics[width=9.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
 %  \includegraphics[width=8.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
@@
   \caption{\label{fig:volmaps}
-      Isocontours of spatial number density of \ce{Ca^{2+}} (cyan, 0.0015~\AA$^{-3}$) 
-      and POPC carbonyl oxygen atoms (red, 0.0075~\AA$^{-3}$, all POPC lipids contribute). 
+      Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.0016~\AA$^{-3}$) 
+      and POPC carbonyl oxygen atoms (light semi-transparent red, 0.007~\AA$^{-3}$, all POPC lipids contribute). 
       Calcium cations localize mostly around phosphate oxygens (red and bronze).
@@
       and it is contributed more by other neighbouring phospholipids than by the same lipid. 
-      Transparent structures are shown to depict the variability of POPC configurations. 
+      Transparent structures are shown to depict the variability of choline configurations 
+      (colour warps from red to blue along the simulation time). 
       The number density was evaluated for each lipid, 

### 2017-12-12 Manuscript/manuscript.tex
Added references for our simulations -- only placeholders!! Real links TBA.

@@
 tested in SI. 
-The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
+The simulated trajectories and parameters files are available at 
+Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files} 
+\todo{To be uploaded to Zenodo, references are only placeholders. }.
 
@@
 Sufficient lipid neighbor exchange occurred during the simulations.
-The trajectories and simulation files for unscaled systems are available from
+The simulated trajectories and parameters files are available at Ref.~\citenum{ECC-POPC_cationic-surf_files} 
+\todo{To be uploaded to Zenodo, references are only placeholders. }.
+The trajectories and simulation files for the unscaled systems are available from
 Refs.~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
-
-\todo{Provide the parameteres for ECC-surfactant somewere.
-      Joe: They will appear along with the Zenodo upload of the trajectories.}
 

### 2017-12-14 Manuscript/manuscript.tex
reference updates, warning to the cationic surfactant figure.

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% This is a (brief) model paper using the achemso class
+%% The document class accepts keyval options, which should include
+%% the target journal and optionally the manuscript type.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[journal=jpcbfk,manuscript=article]{achemso}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional packages needed here.  Only include packages
+%% which are essential, to avoid problems later. Do NOT use any
+%% packages which require e-TeX (for example etoolbox): the e-TeX
+%% extensions are not currently available on the ACS conversion
+%% servers.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\usepackage{rotating} 
+\usepackage{times}
+\usepackage{graphicx}
+\usepackage{setspace}
+\usepackage{amsmath}
+\usepackage{epstopdf}
+\usepackage[obeyFinal]{easy-todo}
+\usepackage{csquotes}
+\usepackage{mhchem}
+\usepackage{chemfig}
+
+%\usepackage{markdown}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% If issues arise when submitting your manuscript, you may want to
+%% un-comment the next line.  This provides information on the
+%% version of every file you have used.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%\listfiles
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional macros here.  Please use \newcommand* where
+%% possible, and avoid layout-changing macros (which are not used
+%% when typesetting).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Meta-data block
+%% ---------------
+%% Each author should be given as a separate \author command.
+%%
+%% Corresponding authors should have an e-mail given after the author
+%% name as an \email command. Phone and fax numbers can be given
+%% using \phone and \fax, respectively; this information is optional.
+%%
+%% The affiliation of authors is given after the authors; each
+%% \affiliation command applies to all preceding authors not already
+%% assigned an affiliation.
+%%
+%% The affiliation takes an option argument for the short name.  This
+%% will typically be something like "University of Somewhere".
+%%
+%% The \altaffiliation macro should be used for new address, etc.
+%% On the other hand, \alsoaffiliation is used on a per author basis
+%% when authors are associated with multiple institutions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\author{Josef Melcr}
+\author{Hector Martinez-Seara}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
+\author{Ji{\v r}{\' i} Kolafa}
+\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
+\author{Pavel Jungwirth}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
+\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
+Tampere, Finland}
+
+\author{O. H. Samuli Ollila}
+\email{samuli.ollila@helsinki.fi}
+%\homepage[]{Your web page}
+\affiliation{Institute of Organic Chemistry and Biochemistry,
+Academy of Sciences of the Czech Republic, 
+Prague 6, Czech Republic}
+\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
+
+%\author{Andrew N. Other}
+%\altaffiliation{A shared footnote}
+%\author{Fred T. Secondauthor}
+%\altaffiliation{Current address: Some other place, Othert\"own,
+%Germany}
+%\author{I. Ken Groupleader}
+%\altaffiliation{A shared footnote}
+%\email{i.k.groupleader@unknown.uu}
+%\phone{+123 (0)123 4445556}
+%\fax{+123 (0)123 4445557}
+%\affiliation[Unknown University]
+%{Department of Chemistry, Unknown University, Unknown Town}
+%\alsoaffiliation[Second University]
+%{Department of Chemistry, Second University, Nearby Town}
+%\author{Susanne K. Laborator}
+%\email{s.k.laborator@bigpharma.co}
+%\affiliation[BigPharma]
+%{Lead Discovery, BigPharma, Big Town, USA}
+%\author{Kay T. Finally}
+%\affiliation[Unknown University]
+%{Department of Chemistry, Unknown University, Unknown Town}
+%\alsoaffiliation[Second University]
+%{Department of Chemistry, Second University, Nearby Town}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The document title should be given as usual. Some journals require
+%% a running title from the author: this should be supplied as an
+%% optional argument to \title.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title[An \textsf{achemso} demo]
+  {Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Some journals require a list of abbreviations or keywords to be
+%% supplied. These should be set up here, and will be printed after
+%% the title and author information, if needed.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\abbreviations{IR,NMR,UV}
+\keywords{American Chemical Society, \LaTeX}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The manuscript does not need to include \maketitle, which is
+%% executed automatically.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The "tocentry" environment can be used to create an entry for the
+%% graphical table of contents. It is given here as some journals
+%% require that it is printed as part of the abstract page. It will
+%% be automatically moved as appropriate.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{tocentry}
+
+Some journals require a graphical entry for the Table of Contents.
+This should be laid out ``print ready'' so that the sizing of the
+text is correct.
+
+Inside the \texttt{tocentry} environment, the font used is Helvetica
+8\,pt, as required by \emph{Journal of the American Chemical
+Society}.
+
+The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
+permitted for  \emph{Journal of the American Chemical Society}
+graphical table of content entries. The box will not resize if the
+content is too big: instead it will overflow the edge of the box.
+
+This box and the associated title will always be printed on a
+separate page at the end of the document.
+
+\end{tocentry}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The abstract environment will automatically gobble the contents
+%% if an abstract is not used by the target journal.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{abstract}
+Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
+bilayers are of paramount significance in the properties and functionality of cellular
+membranes. Current estimates of binding affinities and
+stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
+computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
+1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
+polarization as a mean field correction, known as the electronic continuum correction (ECC).
+This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
+Our improved ECC-POPC model reproduces not only the
+experimentally measured structural parameters for the ion-free membrane, but also the
+response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
+Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
+bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
+in agreement with the previous interpretations of the experimental spectroscopic data.
+The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
+equal probabilities, suggesting more complex binding stoichiometries than those from simple models
+used to interpret the NMR data previously.
+The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
+of complex biochemical systems at cellular membranes.
+\end{abstract}
+
+%\begin{abstract}
+%  This is an example document for the \textsf{achemso} document
+%  class, intended for submissions to the American Chemical Society
+%  for publication. The class is based on the standard \LaTeXe\
+%  \textsf{report} file, and does not seek to reproduce the appearance
+%  of a published paper.
+
+%  This is an abstract for the \textsf{achemso} document class
+%  demonstration document.  An abstract is only allowed for certain
+%  manuscript types.  The selection of \texttt{journal} and
+%  \texttt{manuscript} will determine if an abstract is valid.  If
+%  not, the class will issue an appropriate error.
+%\end{abstract}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Start the main part of the manuscript here.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Introduction}
+Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}.
+Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
+For example, Ca$^{2+}$  is crucial in propagation of neural signals.
+It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{papahadjopoulos90,brouwer15}.
+Ca$^{2+}$ also participates in the T-cell receptor activation.
+It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}.
+Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
+Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions.
+Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.
+
+Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
+Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
+For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
+The details of ion binding are, however, not fully consistent in the literature.
+Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
+In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
+Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}.
+Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
+
+Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements.
+A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}.
+Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results.
+Analyzing massive amounts of data collected within an open collaboration approach, 
+it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
+
+In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
+via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
+To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
+within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
+Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
+Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model.
+This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}.
+The newly developed ECC-POPC model reproduces the experimentally
+measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.
+
+\section{Methods}
+
+\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
+The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations.
+In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
+Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
+However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
+
+According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
+embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
+Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
+\begin{equation}
+Q^{ECC} = f_q \cdot Q
+\end{equation}
+for the ECC corrected charges. 
+Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
+This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
+It is important to note that the value of the
+high frequency dielectric constant 
+is around 2 for almost any biologically relevant environment \cite{leontyev11}.
+The dielectric discontinuity in a lipid bilayer thus arises only
+from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
+Therefore, the same correction for the electronic polarizability can be 
+applied throughout the lipid bilayer/aqueous solution interface.
+
+While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
+Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
+Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}.
+Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
+Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$.
+Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
+
+Here we develop a ECC-POPC lipid model that accurately describes binding 
+of sodium and calcium ions to the POPC  lipid bilayer. 
+The Lipid14~\cite{dickson14} force field 
+(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
+point since 
+it provides the most accurate response of the head groups to ions among the available 
+lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
+provides relatively realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
+We applied the ECC correction 
+to the Lipid14 model of POPC by scaling 
+partial charges of the head group, glycerol 
+backbone, and carbonyl regions. 
+These are the polar parts of phospholipids which can 
+contribute to the cation binding.
+
+To reproduce the experimental ion binding affinities,
+we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
+The ion binding affinity was benchmarked against experiments
+using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
+as discussed more detail in the next section.
+Scaling down the partial charges reduced the ion binding affinity.
+We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
+which is only slightly higher than the ECC one ($f_q=0.75$).
+Directly applying
+the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
+the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
+original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
+(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
+reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
+reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
+the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
+factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
+solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}.
+After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
+
+\subsection{Electrometer concept} \label{section:electrometer}
+Comparing MD simulation to NMR experiments, we can validate the ion
+binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
+This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
+The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
+From MD simulations the order parameters can be calculated using the relation
+\begin{equation}\label{OP}
+S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
+\end{equation}
+where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.
+
+The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
+\begin{equation}\label{OPchangeEQ}
+\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
+\end{equation}
+where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.
+
+The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).
+The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities.
+This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
+To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
+DHAB\\[0.5cm]
+%\vspace{0.5cm}
+\chemfig{
+  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
+}  
+\vspace{0.5cm} \\
+is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end.
+Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
+
+\subsection{Salt concentrations and binding affinities}
+When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
+Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
+while others use atomic absorption spectroscopy and report the salt concentration in
+the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
+we stick to the latter definition by estimating the salt concentration in the aqueous
+bulk region using the farthest point from the lipid bilayer in the aqueous phase.
+Note that the former definition was used by some of us previously.~\cite{catte16}.
+Although these two definitions provide somewhat different values when applied
+to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.
+
+To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}.
+Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region.
+In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.
+Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
+ \begin{equation}\label{surfexcess}
+  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
+\end{equation}
+Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
+$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase,
+and $A_b$ is the area of the unit cell in the membrane plane.
+The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
+because the bilayer has an interface at each of the two leaflets.
+
+\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
+The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors.
+NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution.
+The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq.
+\ref{OP}.
+Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule).
+Form factors were calculated using a relation
+
+\begin{equation}
+  F(q) = \int _{-D/2} ^{D/2} \left ( \rho_{el}(z) - \rho_{el}^s \right ) \cos (zq_z) \mathrm{d}z,
+  %F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
+\end{equation}
+
+\noindent where $\rho_{el} (z)$ is the total electron density, $\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk, and $z$ is the distance from the membrane center along its normal with $D/2$ being half of the unit cell size.
+
+\subsection{Simulation details}
+
+\subsubsection{Simulations of POPC bilayers with aqueous ions}
+Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
+and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
+The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
+in the main text. This model was also used in the previous parametrization
+of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
+the ECC concept~\cite{leontyev11, leontyev14}.
+The robustness of our approach with other water models (OPC \cite{Izadi14},
+OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3P-FB and TIP4P-FB~\cite{Wang2014}, and TIP4P/2005 \cite{Abascal2005})
+was verified and is presented in the Supporting Information (SI).
+Sodium, calcium, and chloride ions were modeled as ECC-ions
+with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
+forming the POPC lipid were derived in this work starting from the Lipid14 force field~\cite{dickson14}.
+For comparison, simulations with the Lipid14 model~\cite{dickson14} and
+ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
+were also performed. The TIP3P water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
+Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and the TIP3P~\cite{jorgensen83} water model were taken directly from
+Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
+
+MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
+in this work are summarized in Table~S1 in the SI. 
+Compatibility with the openMM simulation package \cite{openmm7} was also
+tested and is presented in the SI. 
+The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
+
+\subsubsection{Simulations of POPC bilayers with cationic surfactants}
+The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
+The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters.
+These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
+The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
+For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.
+
+The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
+All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
+Chloride counter ions were used in all the simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
+For POPC, either Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987} parameters  were used.
+The first 20~ns of the total simulation time of 200~ns was considered as an equilibration period and was thus omitted from the analysis.
+We checked that a sufficient lipid  exchange occurred during the simulations.
+The trajectories and simulation files for unscaled systems are available from
+Refs.~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
+
+\todo{Provide the parameters for ECC-surfactant somewhere.
+      Joe: They will appear along with the Zenodo upload of the trajectories.}
+
+\section{Results and Discussion}
+
+\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}
+
+\begin{figure}[tb!]
+  \centering
+  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89.pdf}
+  \includegraphics[width=7.4cm]{../Fig/POPCstructure.eps}
+  %\includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
+  %\includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
+  \caption{\label{simVSexpNOions}
+    Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
+    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K.
+    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
+    from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
+    compared with experiments \cite{ferreira13} at 300~K.
+    The size of the markers for the head group order parameters correspond to
+    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
+    while the error estimate for simulations is $\pm 0.005$.
+    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
+    Bottom: The chemical structure of POPC and the labeling of the carbon segments.
+  } 
+    \todo{Joe: I'm fine with the name change from ECC-lipids to ECC-POPC, but we have to be consistent throughout the text.
+I changed all accurences of ECC-lipids to ECC-POPC in the text, I yet need to change all figure legends then.}
+\end{figure}
+
+\begin{table}[tb!]
+  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
+  }
+  \begin{tabular}{l|c c}
+    model          & APL (\AA$^2$)   & Temperature [K] \\
+    \hline
+    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
+    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
+    \hline
+    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
+    \hline
+    experiment \cite{kucerka11} & 64.3  &  303    \\
+    \hline
+  \end{tabular}
+\end{table}
+
+First, we present results for bilayers in pure water.
+The ECC-POPC and Lipid14 models both reproduce the experimental X-ray scattering form factors
+of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
+The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
+experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
+is $\approx$1\AA{} smaller than the experimental one.
+The values of the area per lipid of the ECC-POPC model vary slightly
+when simulated with different water models (i.e., within the interval of 62.2--66.8 \AA{}, see Table~S2 in SI),
+while still being close to the experimentally reported values.
+We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
+lipid bilayer with a comparable accuracy to other state-of-the-art lipid models~\cite{ollila16}.
+
+Similarly, the acyl chain order parameters of the ECC-POPC model, as well as those of the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
+Notably, the experimentally measured forking and small order parameter values of the $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
+This feature has been suggested to indicate that the carbonyl of the {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in the {\it sn}-1 chain, which orients more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
+This arrangement, which is not fully reproduced by other available lipid models~\cite{ollila16}, may be a relevant feature for the ion binding details.
+
+The order parameters of the $\alpha$ and $\beta$ carbons in the head group are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the former model, see Fig.~\ref{OrderParameterCHANGESsurf}.
+While both models perform relatively well, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations.
+The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
+The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
+
+\todo{Dynamics check is missing: MSD (Hector/Joe)}
+
+\subsection{Calibration of head group response to membrane-bound charge using cationic surfactant}\label{section:boundCHARGE}
+
+\begin{figure}[tb!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
+  \caption{\label{OrderParameterCHANGESsurf}
+    The changes of head group order parameters and P-N vector orientation as a function of
+    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
+    from simulations and experiments \cite{scherer89} at 313 K.
+  }
+\end{figure}
+
+Before studying the sodium and calcium ion binding affinities, we quantify the response of the head group order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
+These mixtures have a well-defined amount of bound charge per PC, namely, the molar fraction of cationic surfactants.
+This is due to the ability of dihexadecyldimethylammonium to directly insert in the lipid bilayer due to its two hydrophobic acyl chains.
+Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid head group order parameters to the amount of bound charge in simulations \cite{scherer89}.
+
+The changes of the head group order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
+An approximately linear decrease of the order parameters, as expected from Eq.
+\ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
+The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters to the bound positive charge is too strong.
+In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being only slightly underestimated for the $\beta$ segment.
+
+In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
+As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in the PC lipid bilayer.
+The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
+For example, the addition of 50\% mole fraction of the cationic surfactant leads to a decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only of 11$^{\circ}$ in the ECC-POPC model.
+The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model.
+The weaker sensitivity of the P-N vector angle response in the ECC-POPC model is in better agreement with experiments.
+
+\subsection{Validation of ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
+  \caption{\label{fig:delta_ordPar_NaCl}
+    Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
+    in bulk ($C_{ion}$) from simulations with different force fields at 313 K together with 
+    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
+    %Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
+    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
+    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
+  }
+  %Figure. in Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.
+\end{figure}
+
+\begin{figure}[htb!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
+  \caption{\label{fig:delta_ordPar_CaCl}
+    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
+    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
+    from simulations at 313 K together with experimental data 
+    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
+    The error estimate for bulk concentrations was approximately 10\,mM.
+    Exception was the order of magnitude larger error in
+    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) because the 
+    small simulation box. 
+    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
+    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
+  }
+  %Experimental values from \cite{akutsu81} for $g_3$ figure:
+  %the value of $g_3$ order parameter of DPPC was -0.214 
+  %in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
+  %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
+  %at the C-2 segments of both fatty acyl chains.'' 
+\end{figure}
+
+Changes of the lipid bilayer head group order parameters extracted from simulations and
+experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
+and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
+As seen in Fig.~\ref{OrderParameterCHANGESsurf}, the order parameters decrease
+proportionally to the amount of the bound positive charge.
+These results can be thus used to compare the ion binding affinities to lipid bilayers between
+simulations and experiments using the electrometer concept~\cite{seelig87, catte16}.
+
+The experimentally measured small order parameter
+changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) 
+are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
+However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
+Replacing \AA{}qvist ions with ion parameters by Dang et al.~\cite{smith94, chang1999, dang2006}
+or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
+the results~(Figs.~\ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}).
+In line with the previous work \cite{catte16}, the results suggest that improvements
+in the lipid parameters are required to correctly describe the binding of cations to phospholipid bilayers.
+
+The results from simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as a function of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
+Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
+Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
+Also the changes of the P-N vector angle are too pronounced for the Lipid14 model,
+for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
+CaCl$_2$ concentration is approximately 17$^{\circ}$. The corresponding value
+for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
+
+Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge.
+It is plausible to assume that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance of validation of the lipid headgroup order parameter response to the bound charge.
+
+Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI.
+In general, the performance of ECC-POPC with any of the tested water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3P model.
+
+\subsection{Binding affinities of Na$^+$ and Ca$^{2+}$ cations to the POPC membrane}
+\label{sec:affinity}
+
+\begin{table}[tb!]
+  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
+    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
+    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
+  \label{tab:binding}}
+  \begin{tabular}{l|c c | c | c c}
+    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+    \hline
+    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    25\%    \\
+    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
+    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
+    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
+%Average no. of bound cations:
+%L14+Aqvist: all:22.7, phos:22.7, carb:8.5
+%L14+Dang:   all:29.2, phos:29.2, carb:4.0
+%L14+ECC-ion:all:37.7, phos:37.7, carb:8.5
+  \end{tabular}
+\end{table}
+
+Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
+The values of $\Gamma_{\rm ion}^{\rm water}$
+from different simulations with the same molar concetration of cations with respect
+to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
+As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess of calcium, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller for the ECC-POPC model than for the other models, 0.13--0.35~nm$^{-2}$.
+Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}.
+This
+means that on average water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
+This is in contradiction with most of the available lipid force fields, which predict a positive surface excess of sodium at PC lipid bilayers \cite{catte16}.
+
+\begin{figure}[htbp!]
+  \centering
+  \includegraphics[height=19.5cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
+  \caption{\label{fig:cacl-dens}
+    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis
+    for different force fields.
+    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
+    the density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
+    the density profiles of phosphate groups and water are divided by 5 and 200, respectively. 
+    All simulations with \ce{CaCl2} shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
+    The simulation with \ce{NaCl} has $C_{ion}'$=1000~mM.
+    }
+\end{figure}
+
+\subsection{Molecular interactions between Na$^+$ or Ca$^{2+}$ cations and POPC oxygens}
+We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
+A maximum distance of 0.3~nm from any lipid oxygen is used to define a bound calcium.
+The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in direct contact with phosphate oxygens.
+From these ions, only one third (32\%) also interacts with the carbonyl oxygens, while the interaction of calcium ions with carbonyl oxygens only is rare (1\%).
+The most abundand interaction scenarios between Ca$^{2+}$ ions and phosphate oxygens are visualized using the probability density isocontours in Fig.~\ref{fig:volmaps}.
+While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.
+
+\begin{figure}[tb!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
+%  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
+%  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
+  \caption{\label{fig:volmaps}
+      Isocontours of probability density of \ce{Ca^{2+}} with respect to 
+      the phosphate oxygens of POPC from ECC-POPC simulation.
+      The probability density was evaluated around a single lipid, 
+      after its strucutral alignment using only phosphate group. 
+      %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
+      %Shown contours suggest that the dominant contribution 
+      %to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
+      %whereas the interactions with any of the two carbonyl groups are considerably milder.
+  }
+  \todo{The figure should be updated as discussed last week.}
+\end{figure}
+
+Even though \ce{Na+} ions do not bind strongly to a POPC bilayer, they still interact mostly with its oxygen moieties.
+The results from a simulation at a $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions at the bilayer interact with phosphate oxygens of POPC only and 20\% with carbonyl oxygens only, with the remaining 25\%, is interacting with both negatively charged groups.
+
+In conclusion, the results suggest that calcium ions bind specifically to the phosphate oxygens, occasionally interacting also with the carbonyls of the PC lipids.
+This is in a qualitative agreement with previous conclusions from several experimental studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}.
+However, the present results suggest, in
+agreement with experiments, an overally weaker binding to the bilayer, in particular with a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
+Sodium ions, which do not exhibit any appreciable affinity for the bilayer, also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
+
+\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
+
+%Binding probability:
+%\ce{Ca2+} complex with \\
+%1 POPC - 13\% -- 30\% \\
+%2 POPC - 18\% -- 42\% \\
+%3 POPC - 12\% -- 28\% \\
+
+Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} as employed in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
+In particular, NMR data concerning the PC headgroup order parameters response and atomic absorption spectra were explained best using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
+Nevertheless, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering \ce{CaCl2} at low concentrations only~\cite{macdonald87}.
+
+In this work, we reproduce the same experimental data used to infer binding stoichiometries employing our ECC-POPC model.
+Thanks to our simulations, we have a direct access to atomistic details of the binding stoichiometry without a need for any binding model as employed for interpreting in experiments~\cite{altenbach84, macdonald87}.
+To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
+Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
+We found the largest propensity for the 1:2 complex (41\%), with probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~being only slightly lower.
+This suggests a more complex binding model than considered in a simple 1:2 ternary complex model previously.
+Nevertheless, with a broad brushstroke, the simulation data can be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
+This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).
+
+\begin{figure}[tb!]
+  \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
+  \caption{\label{fig:cacl_complexes}
+      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
+      with a certain number of POPC lipids. 
+      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
+      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
+  }
+\end{figure}
+
+The probabilities of different complexes formed by \ce{Na+} ions and POPC
+analyzed from the simulation with the ECC-POPC model at $1\,$M concentration of NaCl are also 
+shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
+probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
+for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.
+
+\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in the POPC membrane}
+
+Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current force fields, such  as CHARMM36 and Slipids force fields~\cite{javanainen17}.
+This suggests that at least several microseconds are required to reach the ion binding/unbinding equilibrium.
+To quantify the exchange of ions between the membrane and aqueous solution in simulations, we evaluated residence times of ions bound to the membrane.
+Within our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
+
+The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
+ECC-POPC and CHARMM36 (simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
+In the CHARMM36 simulation, a significant number of the calcium ions is bound to the membrane for the whole length of the trajectory (800~ns).
+In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
+where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
+residence time is around 150~ns, which is below the total length of the simulation used for analysis, i.e., 200~ns.
+Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
+headgroup is of the order of $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
+is yet another order of magnitude faster, with 90\% of the residence times smaller than~1~ns and the longest residence time being~6~ns.
+
+In summary, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within the $\sim$100~ns timeframe, which is significantly faster than observed in simulations emloying most of the presently available lipid models~\cite{javanainen17}.
+Sodium cations exhibit an even more rapid exchange between the membrane and the aqueous solution.
+Our results suggest that simulations with a length of several hundreds of nanoseconds are sufficient to simulate alkali and alkali earth ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
+This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and, in particular,  calcium cations \cite{javanainen17, catte16}.
+
+\section{Conclusions}
+
+In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a force field MD simulation, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
+This ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
+While the structural details of a POPC lipid bilayer in pure water simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the new model in addition reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ at varying concentrations.
+It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
+
+The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model.
+In line with previous interpretations of the experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens, with interactions to the carbonyls being of a secondary importance.
+However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data previously, within which one calcium binds to two POPC molecules~\cite{altenbach84}.
+While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to be relatively abundant and also almost equally likely to occur.
+While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the detailed nature of calcium binding to phospholipid bilayers observed in the present simulation.
+
+Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions, including a mean-field account for electronic polarization effects.
+To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules needs to be verified and, if necessary, further adjustments following the ECC concept to the force fields of the biomolecules interacting with the lipids should be made.
+
+This work can be reached as a repository containing all data at \url{zenodo.org:\dots\dots\dots}.
+
+% If you have acknowledgments, this puts in the proper section head.
+\begin{acknowledgement}
+% Put your acknowledgments here.
+P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
+and from the Academy of Finland via the FiDiPro award.
+J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.
+Computational resources were supplied by the Ministry of Education, Youth and Sports
+of the Czech Republic under the Projects CESNET (Project No. LM2015042) and CERIT-Scientific
+Cloud (Project No. LM2015085) provided within the program Projects of Large Research,
+Development and Innovations Infrastructures.
+O.H.S.O. acknowledges financial support from
+Integrated Structural Biology Research Infrastructure of
+Helsinki Institute of Life Science (Instruct-HiLIFE).
+\end{acknowledgement}
+
+\begin{suppinfo}
+
+%A listing of the contents of each file supplied as Supporting Information
+%should be included. For instructions on what should be included in the
+%Supporting Information as well as how to prepare this material for
+%publications, refer to the journal's Instructions for Authors.
+
+The following files are available free of charge.
+\begin{itemize}
+  \item SI.pdf: Simulation details and further analysis.
+%  \item Filename: brief description
+\end{itemize}
+
+\end{suppinfo}
+
+\bibliography{refs.bib}
+
+\listoftodos
+
+\end{document}
+
+=======
 %% ****** Start of file aiptemplate.tex ****** %
@@
   }
+  \todo{Joe: currenct figure depicts simulations with OPC3 water. Simulations with SPCE are running now.}
 \end{figure}

### 2017-12-15 Manuscript/manuscript.tex

@@
 tested and is presented in the SI. 
-The simulated trajectories and parameters files are available at \cite{??} \todo{To be uploaded to Zenodo}.
+The simulated trajectories and parameters files are available at
+Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files}.
+\todo{To be uploaded to Zenodo, references are only placeholders. }.
 
@@
 We checked that a sufficient lipid  exchange occurred during the simulations.
-The trajectories and simulation files for unscaled systems are available from
+The simulated trajectories and parameters files are available at Ref.~\citenum{ECC-POPC_cationic-surf_files} 
+\todo{To be uploaded to Zenodo, references are only placeholders. }.
+The trajectories and simulation files for unscaled systems are available at
 Refs.~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
-
-\todo{Provide the parameters for ECC-surfactant somewhere.
-      Joe: They will appear along with the Zenodo upload of the trajectories.}
 
@@
     The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
+    Open/closed symbols are used for palmitoyl/oleoyl chains of POPC. \\
     Bottom: The chemical structure of POPC and the labeling of the carbon segments.
   } 
-    \todo{Joe: I'm fine with the name change from ECC-lipids to ECC-POPC, but we have to be consistent throughout the text.
-I changed all accurences of ECC-lipids to ECC-POPC in the text, I yet need to change all figure legends then.}
 \end{figure}
@@
     (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    The error estimate for bulk concentrations was approximately 10\,mM.
-    Exception was the order of magnitude larger error in
-    simulation with Lipid14 and ECC-ions due to the unconverged densities  (shown if Fig.~\ref{fig:cacl-dens}) because the 
-    small simulation box. 
+    The error estimate for bulk concentrations is approximately 10\,mM.
+    The order of magnitude larger error in the
+    simulation with Lipid14 and ECC-ions is due to unconverged bulk densities  (shown if Fig.~\ref{fig:cacl-dens}) limited by
+    the simulation box. 
     Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
-    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
+    Refs.~\citenum{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
   }
@@
   \centering
-  \includegraphics[width=8.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
+  \includegraphics[width=8.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
 %  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
@@
   \caption{\label{fig:volmaps}
-      Isocontours of probability density of \ce{Ca^{2+}} with respect to 
-      the phosphate oxygens of POPC from ECC-POPC simulation.
-      The probability density was evaluated around a single lipid, 
+    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.0016~\AA$^{-3}$) 
+    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.007~\AA$^{-3}$, all POPC lipids contribute). 
+    Calcium cations localize mostly around phosphate oxygens (red and bronze).
+    Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
+    and it is contributed more by other neighbouring phospholipids than by the same lipid. 
+    Transparent structures are shown to depict the variability of choline configurations 
+    (colour warps from red to blue along the simulation time). 
+    The number density was evaluated for each lipid, 
     after its strucutral alignment using only phosphate group.
-      %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
-      %Shown contours suggest that the dominant contribution 
-      %to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
-      %whereas the interactions with any of the two carbonyl groups are considerably milder.
-  }
-  \todo{The figure should be updated as discussed last week.}
+    MDAnalysis \cite{mdanalysis2011} library was used for 
+    the calculations of the structural alignment and the spatial number density. 
+    VMD \cite{hump96} was used for visualisation. 
+    Carbon atoms are cyan, hydrogen atoms white, oxygen atoms red, nitrogen blue.
+  }
+ %\todo{The figure should be updated as discussed last week. Joe: Will only smoothen the isosurfaces now. }
 \end{figure}

### 2017-12-18 Manuscript/manuscript.tex
Update after comments by Kolafa

@@
 Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
-However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
+However, the present concept of ECC is physically well justified and has obtained theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
 
@@
 
-While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
+While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case \cite{leontyev14}.
 Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
-Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}.
+Several computational schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}.
 Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
@@
 which is only slightly higher than the ECC one ($f_q=0.75$).
+Note that common scaling factors for monovalent ions in water are 0.8 or
+larger \cite{benavides17,skinner14,nacleps}.  In contrast, modern force fields
+for ionic liquids suggest values 0.6--0.65 which are lower than $\epsilon^{-1/2}_{el}$\cite{holm14}.
 Directly applying
@@
 \begin{equation}\label{OP}
-S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
+S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
 \end{equation}
@@
 Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
-$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase,
+$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase
+(measured from the density of farthest point between periodic bilayers),
 and $A_b$ is the area of the unit cell in the membrane plane.
@@
 of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
-The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
+The area per lipid from the Lipid14 model is by $\approx$1\AA{} larger than the
 experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
-is $\approx$1\AA{} smaller than the experimental one.
+is by $\approx$1\AA{} smaller than the experimental one.
 The values of the area per lipid of the ECC-POPC model vary slightly
@@
     and POPC carbonyl oxygen atoms (light semi-transparent red, 0.007~\AA$^{-3}$, all POPC lipids contribute). 
-    Calcium cations localize mostly around phosphate oxygens (red and bronze).
+    Calcium cations localize mostly around phosphate oxygens (oxygens red, phosphate bronze).
     Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
@@
 Thanks to our simulations, we have a direct access to atomistic details of the binding stoichiometry without a need for any binding model as employed for interpreting in experiments~\cite{altenbach84, macdonald87}.
-To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
+To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC),
+we calculated for each bound Ca$^{2+}$ the number of POPC molecules having oxygen atoms within a distance of 0.3~nm.
 Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
@@
 Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
-However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
+However, the present concept of ECC is physically well justified and has
+obtained theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
 

### 2017-12-18 Manuscript/SI.tex
Updated plots after OPC3->SPCE correction in simulations with cat.surf. According changes in MS and SI.

@@
 
+OpenMM simulations were run with 4~fs time step using 4 times heavier hydrogen atoms (mass subtracted from neighbouring atoms).
+
 \newpage
@@
   \centering
+  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_surf_waterModels_compar.pdf}
+  \caption{\label{fig:ordPars_waterModels_surf}
+    Changes of head group and glycerol carbon $g_3$ order parameters, and P-N vector orientation of POPC bilayer
+    as a function of a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
+    from simulations with different water models (OPC3~\cite{Izadi16}, SPC/E~\cite{Berendsen1987})
+    and experiments \cite{scherer89} at 313 K.
+  }
+\end{figure}
+
+\begin{figure}[!hp]
+  \centering
   \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl_waterModels.pdf}
@@
     as a function of CaCl$_2$ concentrations in bulk ($C_{ion}$) from ECC-POPC simulations at 313K with different
-    water models (OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
+    water models (SPC/E~\cite{Berendsen1987}, OPC \cite{Izadi14}, OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
     together with experimental data (DPPC (323~K) \cite{akutsu81} and POPC (313~K) \cite{altenbach84}). 

### 2017-12-18 Manuscript/manuscript.tex
Phospate to phosphorus in figure caption
Updated plots after OPC3->SPCE correction in simulations with cat.surf. According changes in MS and SI.

--- 2017-12-15 Manuscript/manuscript.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% This is a (brief) model paper using the achemso class
-%% The document class accepts keyval options, which should include
-%% the target journal and optionally the manuscript type.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jpcbfk,manuscript=article]{achemso}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Place any additional packages needed here.  Only include packages
-%% which are essential, to avoid problems later. Do NOT use any
-%% packages which require e-TeX (for example etoolbox): the e-TeX
-%% extensions are not currently available on the ACS conversion
-%% servers.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\usepackage{rotating} 
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{setspace}
-\usepackage{amsmath}
-\usepackage{epstopdf}
-\usepackage[obeyFinal]{easy-todo}
-\usepackage{csquotes}
-\usepackage{mhchem}
-\usepackage{chemfig}
-
-%\usepackage{markdown}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% If issues arise when submitting your manuscript, you may want to
-%% un-comment the next line.  This provides information on the
-%% version of every file you have used.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\listfiles
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Place any additional macros here.  Please use \newcommand* where
-%% possible, and avoid layout-changing macros (which are not used
-%% when typesetting).
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Meta-data block
-%% ---------------
-%% Each author should be given as a separate \author command.
-%%
-%% Corresponding authors should have an e-mail given after the author
-%% name as an \email command. Phone and fax numbers can be given
-%% using \phone and \fax, respectively; this information is optional.
-%%
-%% The affiliation of authors is given after the authors; each
-%% \affiliation command applies to all preceding authors not already
-%% assigned an affiliation.
-%%
-%% The affiliation takes an option argument for the short name.  This
-%% will typically be something like "University of Somewhere".
-%%
-%% The \altaffiliation macro should be used for new address, etc.
-%% On the other hand, \alsoaffiliation is used on a per author basis
-%% when authors are associated with multiple institutions.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\author{Josef Melcr}
-\author{Hector Martinez-Seara}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\author{Ji{\v r}{\' i} Kolafa}
-\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
-\author{Pavel Jungwirth}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
-Tampere, Finland}
-
-\author{O. H. Samuli Ollila}
-\email{samuli.ollila@helsinki.fi}
-%\homepage[]{Your web page}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
-
-%\author{Andrew N. Other}
-%\altaffiliation{A shared footnote}
-%\author{Fred T. Secondauthor}
-%\altaffiliation{Current address: Some other place, Othert\"own,
-%Germany}
-%\author{I. Ken Groupleader}
-%\altaffiliation{A shared footnote}
-%\email{i.k.groupleader@unknown.uu}
-%\phone{+123 (0)123 4445556}
-%\fax{+123 (0)123 4445557}
-%\affiliation[Unknown University]
-%{Department of Chemistry, Unknown University, Unknown Town}
-%\alsoaffiliation[Second University]
-%{Department of Chemistry, Second University, Nearby Town}
-%\author{Susanne K. Laborator}
-%\email{s.k.laborator@bigpharma.co}
-%\affiliation[BigPharma]
-%{Lead Discovery, BigPharma, Big Town, USA}
-%\author{Kay T. Finally}
-%\affiliation[Unknown University]
-%{Department of Chemistry, Unknown University, Unknown Town}
-%\alsoaffiliation[Second University]
-%{Department of Chemistry, Second University, Nearby Town}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The document title should be given as usual. Some journals require
-%% a running title from the author: this should be supplied as an
-%% optional argument to \title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[An \textsf{achemso} demo]
-  {Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Some journals require a list of abbreviations or keywords to be
-%% supplied. These should be set up here, and will be printed after
-%% the title and author information, if needed.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\abbreviations{IR,NMR,UV}
-\keywords{American Chemical Society, \LaTeX}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The manuscript does not need to include \maketitle, which is
-%% executed automatically.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "tocentry" environment can be used to create an entry for the
-%% graphical table of contents. It is given here as some journals
-%% require that it is printed as part of the abstract page. It will
-%% be automatically moved as appropriate.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{tocentry}
-
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
-
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
-
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
-
-This box and the associated title will always be printed on a
-separate page at the end of the document.
-
-\end{tocentry}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The abstract environment will automatically gobble the contents
-%% if an abstract is not used by the target journal.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstract}
-Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
-bilayers are of paramount significance in the properties and functionality of cellular
-membranes. Current estimates of binding affinities and
-stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
-computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
-1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
-polarization as a mean field correction, known as the electronic continuum correction (ECC).
-This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
-Our improved ECC-POPC model reproduces not only the
-experimentally measured structural parameters for the ion-free membrane, but also the
-response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
-Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
-bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
-in agreement with the previous interpretations of the experimental spectroscopic data.
-The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
-equal probabilities, suggesting more complex binding stoichiometries than those from simple models
-used to interpret the NMR data previously.
-The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
-of complex biochemical systems at cellular membranes.
-\end{abstract}
-
-%\begin{abstract}
-%  This is an example document for the \textsf{achemso} document
-%  class, intended for submissions to the American Chemical Society
-%  for publication. The class is based on the standard \LaTeXe\
-%  \textsf{report} file, and does not seek to reproduce the appearance
-%  of a published paper.
-
-%  This is an abstract for the \textsf{achemso} document class
-%  demonstration document.  An abstract is only allowed for certain
-%  manuscript types.  The selection of \texttt{journal} and
-%  \texttt{manuscript} will determine if an abstract is valid.  If
-%  not, the class will issue an appropriate error.
-%\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Start the main part of the manuscript here.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction}
-Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}.
-Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
-For example, Ca$^{2+}$  is crucial in propagation of neural signals.
-It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{papahadjopoulos90,brouwer15}.
-Ca$^{2+}$ also participates in the T-cell receptor activation.
-It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}.
-Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
-Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions.
-Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.
-
-Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
-Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
-For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
-The details of ion binding are, however, not fully consistent in the literature.
-Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
-In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}.
-Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
-
-Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements.
-A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}.
-Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results.
-Analyzing massive amounts of data collected within an open collaboration approach, 
-it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
-
-In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
-via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
-To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
-within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
-Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
-Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model.
-This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}.
-The newly developed ECC-POPC model reproduces the experimentally
-measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.
-
-\section{Methods}
-
-\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations.
-In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
-Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
-However, the present concept of ECC is physically well justified and rigorously derived~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
-
-According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
-embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
-Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
-\begin{equation}
-Q^{ECC} = f_q \cdot Q
-\end{equation}
-for the ECC corrected charges. 
-Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
-It is important to note that the value of the
-high frequency dielectric constant 
-is around 2 for almost any biologically relevant environment \cite{leontyev11}.
-The dielectric discontinuity in a lipid bilayer thus arises only
-from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
-Therefore, the same correction for the electronic polarizability can be 
-applied throughout the lipid bilayer/aqueous solution interface.
-
-While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
-Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
-Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}.
-Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
-Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$.
-Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
-
-Here we develop a ECC-POPC lipid model that accurately describes binding 
-of sodium and calcium ions to the POPC  lipid bilayer. 
-The Lipid14~\cite{dickson14} force field 
-(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
-point since 
-it provides the most accurate response of the head groups to ions among the available 
-lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
-provides relatively realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
-We applied the ECC correction 
-to the Lipid14 model of POPC by scaling 
-partial charges of the head group, glycerol 
-backbone, and carbonyl regions. 
-These are the polar parts of phospholipids which can 
-contribute to the cation binding.
-
-To reproduce the experimental ion binding affinities,
-we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
-The ion binding affinity was benchmarked against experiments
-using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
-as discussed more detail in the next section.
-Scaling down the partial charges reduced the ion binding affinity.
-We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
-which is only slightly higher than the ECC one ($f_q=0.75$).
-Directly applying
-the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
-the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
-original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
-(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
-reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
-reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
-the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
-factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
-solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}.
-After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
-
-\subsection{Electrometer concept} \label{section:electrometer}
-Comparing MD simulation to NMR experiments, we can validate the ion
-binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
-This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
-The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
-From MD simulations the order parameters can be calculated using the relation
-\begin{equation}\label{OP}
-S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
-\end{equation}
-where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.
-
-The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
-\begin{equation}\label{OPchangeEQ}
-\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
-\end{equation}
-where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.
-
-The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).
-The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities.
-This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
-To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
-DHAB\\[0.5cm]
-%\vspace{0.5cm}
-\chemfig{
-  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
-}  
-\vspace{0.5cm} \\
-is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end.
-Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
-
-\subsection{Salt concentrations and binding affinities}
-When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
-Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
-while others use atomic absorption spectroscopy and report the salt concentration in
-the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
-we stick to the latter definition by estimating the salt concentration in the aqueous
-bulk region using the farthest point from the lipid bilayer in the aqueous phase.
-Note that the former definition was used by some of us previously.~\cite{catte16}.
-Although these two definitions provide somewhat different values when applied
-to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.
-
-To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}.
-Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region.
-In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.
-Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
- \begin{equation}\label{surfexcess}
-  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
-\end{equation}
-Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
-$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase,
-and $A_b$ is the area of the unit cell in the membrane plane.
-The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
-because the bilayer has an interface at each of the two leaflets.
-
-\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
-The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors.
-NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution.
-The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq.
-\ref{OP}.
-Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule).
-Form factors were calculated using a relation
-
-\begin{equation}
-  F(q) = \int _{-D/2} ^{D/2} \left ( \rho_{el}(z) - \rho_{el}^s \right ) \cos (zq_z) \mathrm{d}z,
-  %F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
-\end{equation}
-
-\noindent where $\rho_{el} (z)$ is the total electron density, $\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk, and $z$ is the distance from the membrane center along its normal with $D/2$ being half of the unit cell size.
-
-\subsection{Simulation details}
-
-\subsubsection{Simulations of POPC bilayers with aqueous ions}
-Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
-and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
-The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
-in the main text. This model was also used in the previous parametrization
-of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
-the ECC concept~\cite{leontyev11, leontyev14}.
-The robustness of our approach with other water models (OPC \cite{Izadi14},
-OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3P-FB and TIP4P-FB~\cite{Wang2014}, and TIP4P/2005 \cite{Abascal2005})
-was verified and is presented in the Supporting Information (SI).
-Sodium, calcium, and chloride ions were modeled as ECC-ions
-with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
-forming the POPC lipid were derived in this work starting from the Lipid14 force field~\cite{dickson14}.
-For comparison, simulations with the Lipid14 model~\cite{dickson14} and
-ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
-were also performed. The TIP3P water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
-Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and the TIP3P~\cite{jorgensen83} water model were taken directly from
-Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
-
-MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
-in this work are summarized in Table~S1 in the SI. 
-Compatibility with the openMM simulation package \cite{openmm7} was also
-tested and is presented in the SI. 
-The simulated trajectories and parameters files are available at
-Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files}.
-\todo{To be uploaded to Zenodo, references are only placeholders. }.
-
-\subsubsection{Simulations of POPC bilayers with cationic surfactants}
-The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
-The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters.
-These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
-The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.
-
-The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
-All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
-Chloride counter ions were used in all the simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
-For POPC, either Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987} parameters  were used.
-The first 20~ns of the total simulation time of 200~ns was considered as an equilibration period and was thus omitted from the analysis.
-We checked that a sufficient lipid  exchange occurred during the simulations.
-The simulated trajectories and parameters files are available at Ref.~\citenum{ECC-POPC_cationic-surf_files} 
-\todo{To be uploaded to Zenodo, references are only placeholders. }.
-The trajectories and simulation files for unscaled systems are available at
-Refs.~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
-
-\section{Results and Discussion}
-
-\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89_POPC-struct.pdf}
-  %\includegraphics[width=7.4cm]{../Fig/POPCstructure.eps}
-  %\includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
-  %\includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-  \caption{\label{simVSexpNOions}
-    Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K. \\
-    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
-    compared with experiments \cite{ferreira13} at 300~K.
-    The size of the markers for the head group order parameters correspond to
-    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
-    while the error estimate for simulations is $\pm 0.005$.
-    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
-    Open/closed symbols are used for palmitoyl/oleoyl chains of POPC. \\
-    Bottom: The chemical structure of POPC and the labeling of the carbon segments.
-  } 
-\end{figure}
-
-\begin{table}[tb!]
-  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
-  }
-  \begin{tabular}{l|c c}
-    model          & APL (\AA$^2$)   & Temperature [K] \\
-    \hline
-    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
-    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
-    \hline
-    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
-    \hline
-    experiment \cite{kucerka11} & 64.3  &  303    \\
-    \hline
-  \end{tabular}
-\end{table}
-
-First, we present results for bilayers in pure water.
-The ECC-POPC and Lipid14 models both reproduce the experimental X-ray scattering form factors
-of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
-The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
-experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
-is $\approx$1\AA{} smaller than the experimental one.
-The values of the area per lipid of the ECC-POPC model vary slightly
-when simulated with different water models (i.e., within the interval of 62.2--66.8 \AA{}, see Table~S2 in SI),
-while still being close to the experimentally reported values.
-We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
-lipid bilayer with a comparable accuracy to other state-of-the-art lipid models~\cite{ollila16}.
-
-Similarly, the acyl chain order parameters of the ECC-POPC model, as well as those of the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
-Notably, the experimentally measured forking and small order parameter values of the $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
-This feature has been suggested to indicate that the carbonyl of the {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in the {\it sn}-1 chain, which orients more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
-This arrangement, which is not fully reproduced by other available lipid models~\cite{ollila16}, may be a relevant feature for the ion binding details.
-
-The order parameters of the $\alpha$ and $\beta$ carbons in the head group are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the former model, see Fig.~\ref{OrderParameterCHANGESsurf}.
-While both models perform relatively well, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations.
-The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
-The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
-
-%\todo{Dynamics check is missing: MSD (Hector/Joe)}
-
-\subsection{Calibration of head group response to membrane-bound charge using cationic surfactant}\label{section:boundCHARGE}
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
-  \caption{\label{OrderParameterCHANGESsurf}
-    The changes of head group order parameters and P-N vector orientation as a function of
-    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
-    from simulations and experiments \cite{scherer89} at 313 K.
-  }
-  \todo{Joe: currenct figure depicts simulations with OPC3 water. Simulations with SPCE are running now.}
-\end{figure}
-
-Before studying the sodium and calcium ion binding affinities, we quantify the response of the head group order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
-These mixtures have a well-defined amount of bound charge per PC, namely, the molar fraction of cationic surfactants.
-This is due to the ability of dihexadecyldimethylammonium to directly insert in the lipid bilayer due to its two hydrophobic acyl chains.
-Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid head group order parameters to the amount of bound charge in simulations \cite{scherer89}.
-
-The changes of the head group order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
-An approximately linear decrease of the order parameters, as expected from Eq.
-\ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
-The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters to the bound positive charge is too strong.
-In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being only slightly underestimated for the $\beta$ segment.
-
-In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
-As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in the PC lipid bilayer.
-The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
-For example, the addition of 50\% mole fraction of the cationic surfactant leads to a decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only of 11$^{\circ}$ in the ECC-POPC model.
-The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model.
-The weaker sensitivity of the P-N vector angle response in the ECC-POPC model is in better agreement with experiments.
-
-\subsection{Validation of ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
-  \caption{\label{fig:delta_ordPar_NaCl}
-    Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
-    in bulk ($C_{ion}$) from simulations with different force fields at 313 K together with 
-    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
-    %Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
-    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
-  }
-  %Figure. in Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
-  \caption{\label{fig:delta_ordPar_CaCl}
-    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
-    from simulations at 313 K together with experimental data 
-    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    The error estimate for bulk concentrations is approximately 10\,mM.
-    The order of magnitude larger error in the
-    simulation with Lipid14 and ECC-ions is due to unconverged bulk densities  (shown if Fig.~\ref{fig:cacl-dens}) limited by
-    the simulation box. 
-    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
-    Refs.~\citenum{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
-  }
-  %Experimental values from \cite{akutsu81} for $g_3$ figure:
-  %the value of $g_3$ order parameter of DPPC was -0.214 
-  %in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
-  %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
-  %at the C-2 segments of both fatty acyl chains.'' 
-\end{figure}
-
-Changes of the lipid bilayer head group order parameters extracted from simulations and
-experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
-and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
-As seen in Fig.~\ref{OrderParameterCHANGESsurf}, the order parameters decrease
-proportionally to the amount of the bound positive charge.
-These results can be thus used to compare the ion binding affinities to lipid bilayers between
-simulations and experiments using the electrometer concept~\cite{seelig87, catte16}.
-
-The experimentally measured small order parameter
-changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) 
-are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
-However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
-Replacing \AA{}qvist ions with ion parameters by Dang et al.~\cite{smith94, chang1999, dang2006}
-or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
-the results~(Figs.~\ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}).
-In line with the previous work \cite{catte16}, the results suggest that improvements
-in the lipid parameters are required to correctly describe the binding of cations to phospholipid bilayers.
-
-The results from simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as a function of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
-Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
-Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-Also the changes of the P-N vector angle are too pronounced for the Lipid14 model,
-for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
-CaCl$_2$ concentration is approximately 17$^{\circ}$. The corresponding value
-for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
-
-Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge.
-It is plausible to assume that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance of validation of the lipid headgroup order parameter response to the bound charge.
-
-Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI.
-In general, the performance of ECC-POPC with any of the tested water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3P model.
-
-\subsection{Binding affinities of Na$^+$ and Ca$^{2+}$ cations to the POPC membrane}
-\label{sec:affinity}
-
-\begin{table}[tb!]
-  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
-    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
-    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
-  \label{tab:binding}}
-  \begin{tabular}{l|c c | c | c c}
-    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
-    \hline
-    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    25\%    \\
-    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
-    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
-    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
-%Average no. of bound cations:
-%L14+Aqvist: all:22.7, phos:22.7, carb:8.5
-%L14+Dang:   all:29.2, phos:29.2, carb:4.0
-%L14+ECC-ion:all:37.7, phos:37.7, carb:8.5
-  \end{tabular}
-\end{table}
-
-Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
-The values of $\Gamma_{\rm ion}^{\rm water}$
-from different simulations with the same molar concetration of cations with respect
-to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
-As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess of calcium, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller for the ECC-POPC model than for the other models, 0.13--0.35~nm$^{-2}$.
-Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}.
-This
-means that on average water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
-This is in contradiction with most of the available lipid force fields, which predict a positive surface excess of sodium at PC lipid bilayers \cite{catte16}.
-
-\begin{figure}[htbp!]
-  \centering
-  \includegraphics[height=19.9cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
-  \caption{\label{fig:cacl-dens}
-    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis
-    for different force fields.
-    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
-    the density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
-    the density profiles of phosphate groups and water are divided by 5 and 200, respectively. 
-    All simulations with \ce{CaCl2} shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
-    The simulation with \ce{NaCl} has $C_{ion}'$=1000~mM.
-    }
-\end{figure}
-
-\subsection{Molecular interactions between Na$^+$ or Ca$^{2+}$ cations and POPC oxygens}
-We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
-A maximum distance of 0.3~nm from any lipid oxygen is used to define a bound calcium.
-The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in direct contact with phosphate oxygens.
-From these ions, only one third (32\%) also interacts with the carbonyl oxygens, while the interaction of calcium ions with carbonyl oxygens only is rare (1\%).
-The most abundand interaction scenarios between Ca$^{2+}$ ions and phosphate oxygens are visualized using the probability density isocontours in Fig.~\ref{fig:volmaps}.
-While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
-%  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
-%  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
-  \caption{\label{fig:volmaps}
-    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.0016~\AA$^{-3}$) 
-    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.007~\AA$^{-3}$, all POPC lipids contribute). 
-    Calcium cations localize mostly around phosphate oxygens (red and bronze).
-    Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
-    and it is contributed more by other neighbouring phospholipids than by the same lipid. 
-    Transparent structures are shown to depict the variability of choline configurations 
-    (colour warps from red to blue along the simulation time). 
-    The number density was evaluated for each lipid, 
-    after its strucutral alignment using only phosphate group.
-    MDAnalysis \cite{mdanalysis2011} library was used for 
-    the calculations of the structural alignment and the spatial number density. 
-    VMD \cite{hump96} was used for visualisation. 
-    Carbon atoms are cyan, hydrogen atoms white, oxygen atoms red, nitrogen blue.
-  }
- %\todo{The figure should be updated as discussed last week. Joe: Will only smoothen the isosurfaces now. }
-\end{figure}
-
-Even though \ce{Na+} ions do not bind strongly to a POPC bilayer, they still interact mostly with its oxygen moieties.
-The results from a simulation at a $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions at the bilayer interact with phosphate oxygens of POPC only and 20\% with carbonyl oxygens only, with the remaining 25\%, is interacting with both negatively charged groups.
-
-In conclusion, the results suggest that calcium ions bind specifically to the phosphate oxygens, occasionally interacting also with the carbonyls of the PC lipids.
-This is in a qualitative agreement with previous conclusions from several experimental studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}.
-However, the present results suggest, in
-agreement with experiments, an overally weaker binding to the bilayer, in particular with a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
-Sodium ions, which do not exhibit any appreciable affinity for the bilayer, also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
-
-\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
-
-%Binding probability:
-%\ce{Ca2+} complex with \\
-%1 POPC - 13\% -- 30\% \\
-%2 POPC - 18\% -- 42\% \\
-%3 POPC - 12\% -- 28\% \\
-
-Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} as employed in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
-In particular, NMR data concerning the PC headgroup order parameters response and atomic absorption spectra were explained best using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-Nevertheless, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering \ce{CaCl2} at low concentrations only~\cite{macdonald87}.
-
-In this work, we reproduce the same experimental data used to infer binding stoichiometries employing our ECC-POPC model.
-Thanks to our simulations, we have a direct access to atomistic details of the binding stoichiometry without a need for any binding model as employed for interpreting in experiments~\cite{altenbach84, macdonald87}.
-To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
-Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
-We found the largest propensity for the 1:2 complex (41\%), with probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~being only slightly lower.
-This suggests a more complex binding model than considered in a simple 1:2 ternary complex model previously.
-Nevertheless, with a broad brushstroke, the simulation data can be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
-This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
-  \caption{\label{fig:cacl_complexes}
-      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
-      with a certain number of POPC lipids. 
-      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
-      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
-  }
-\end{figure}
-
-The probabilities of different complexes formed by \ce{Na+} ions and POPC
-analyzed from the simulation with the ECC-POPC model at $1\,$M concentration of NaCl are also 
-shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
-probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
-for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.
-
-\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in the POPC membrane}
-
-Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current force fields, such  as CHARMM36 and Slipids force fields~\cite{javanainen17}.
-This suggests that at least several microseconds are required to reach the ion binding/unbinding equilibrium.
-To quantify the exchange of ions between the membrane and aqueous solution in simulations, we evaluated residence times of ions bound to the membrane.
-Within our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
-
-The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
-ECC-POPC and CHARMM36 (simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
-In the CHARMM36 simulation, a significant number of the calcium ions is bound to the membrane for the whole length of the trajectory (800~ns).
-In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
-where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
-residence time is around 150~ns, which is below the total length of the simulation used for analysis, i.e., 200~ns.
-Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
-headgroup is of the order of $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
-is yet another order of magnitude faster, with 90\% of the residence times smaller than~1~ns and the longest residence time being~6~ns.
-
-In summary, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within the $\sim$100~ns timeframe, which is significantly faster than observed in simulations emloying most of the presently available lipid models~\cite{javanainen17}.
-Sodium cations exhibit an even more rapid exchange between the membrane and the aqueous solution.
-Our results suggest that simulations with a length of several hundreds of nanoseconds are sufficient to simulate alkali and alkali earth ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
-This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and, in particular,  calcium cations \cite{javanainen17, catte16}.
-
-\section{Conclusions}
-
-In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a force field MD simulation, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
-This ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
-While the structural details of a POPC lipid bilayer in pure water simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the new model in addition reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ at varying concentrations.
-It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
-
-The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model.
-In line with previous interpretations of the experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens, with interactions to the carbonyls being of a secondary importance.
-However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data previously, within which one calcium binds to two POPC molecules~\cite{altenbach84}.
-While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to be relatively abundant and also almost equally likely to occur.
-While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the detailed nature of calcium binding to phospholipid bilayers observed in the present simulation.
-
-Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions, including a mean-field account for electronic polarization effects.
-To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules needs to be verified and, if necessary, further adjustments following the ECC concept to the force fields of the biomolecules interacting with the lipids should be made.
-
-This work can be reached as a repository containing all data and scripts for analysis at \url{zenodo.org:\dots\dots\dots}.
-
-% If you have acknowledgments, this puts in the proper section head.
-\begin{acknowledgement}
-% Put your acknowledgments here.
-P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
-and from the Academy of Finland via the FiDiPro award.
-J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.
-Computational resources were supplied by the Ministry of Education, Youth and Sports
-of the Czech Republic under the Projects CESNET (Project No. LM2015042) and CERIT-Scientific
-Cloud (Project No. LM2015085) provided within the program Projects of Large Research,
-Development and Innovations Infrastructures.
-O.H.S.O. acknowledges financial support from
-Integrated Structural Biology Research Infrastructure of
-Helsinki Institute of Life Science (Instruct-HiLIFE).
-\end{acknowledgement}
-
-\begin{suppinfo}
-
-%A listing of the contents of each file supplied as Supporting Information
-%should be included. For instructions on what should be included in the
-%Supporting Information as well as how to prepare this material for
-%publications, refer to the journal's Instructions for Authors.
-
-The following files are available free of charge.
-\begin{itemize}
-  \item SI.pdf: Simulation details and further analysis.
-%  \item Filename: brief description
-\end{itemize}
-
-\end{suppinfo}
-
-\bibliography{refs.bib}
-
-\listoftodos
-
-\end{document}
-
-=======
 %% ****** Start of file aiptemplate.tex ****** %
@@
     the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
-    while the error estimate for simulations is $\pm 0.005$.
+    while the error estimate for simulations is $\pm 0.005$ 
+    (Bayesian estimate of 95\% confidence interval \cite{scipy}).
     The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
@@
   }
-  \todo{Joe: currenct figure depicts simulations with OPC3 water. Simulations with SPCE are running now.}
 \end{figure}

### 2017-12-18 Manuscript/manuscript.tex
New smooth isocontours figure with according changes in MS.

--- 2017-12-15 Manuscript/manuscript.tex
@@
   \caption{\label{fig:volmaps}
-    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.0016~\AA$^{-3}$) 
-    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.007~\AA$^{-3}$, all POPC lipids contribute). 
+    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.001~\AA$^{-3}$) 
+    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.008~\AA$^{-3}$, all POPC lipids contribute). 
     Calcium cations localize mostly around phosphate oxygens (red and bronze).

### 2017-12-18 Manuscript/manuscript.tex

--- 2017-12-18 Manuscript/manuscript.tex
@@
   \caption{\label{fig:volmaps}
-    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.0016~\AA$^{-3}$) 
-    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.007~\AA$^{-3}$, all POPC lipids contribute). 
-    Calcium cations localize mostly around phosphate oxygens (oxygens red, phosphate bronze).
+    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.001~\AA$^{-3}$) 
+    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.008~\AA$^{-3}$, all POPC lipids contribute). 
+    Calcium cations localize mostly around phosphate oxygens (oxygens red, phosphorus bronze).
     Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 

### 2017-12-19 Manuscript/manuscript.tex

@@
     the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
-    while the error estimate for simulations is $\pm 0.005$.
+    while the error estimate for simulations is $\pm 0.005$
+    (Bayesian estimate of 95\% confidence interval \cite{scipy}).
     The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
@@
 \end{document}
-
-=======
-%% ****** Start of file aiptemplate.tex ****** %
-%%
-%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
-%%   Version 4.1 of 9 October 2009.
-%%
-%
-% This is a template for producing documents for use with 
-% the REVTEX 4.1 document class and the AIP substyles.
-% 
-% Copy this file to another name and then work on that file.
-% That way, you always have this original template file to use.
-
-%\documentclass[aip,graphicx]{revtex4-1}
-%\documentclass[aip,reprint]{revtex4-1}
-
-%\usepackage{graphicx}
-
-%\draft % marks overfull lines with a black rule on the right
-%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
-%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}
-
-\documentclass[aip,jcp,twocolumn]{revtex4}
-%\documentclass[aip,jcp]{revtex4}
-%\documentclass{article}
-
-%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}
-
-\usepackage{rotating} 
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{setspace}
-\usepackage{amsmath}
-\usepackage{epstopdf}
-\usepackage[obeyFinal]{easy-todo}
-\usepackage{csquotes}
-\usepackage{mhchem}
-\usepackage{chemfig}
-
-%\usepackage{markdown}
-
-\begin{document}
-
-% Use the \preprint command to place your local institutional report number 
-% on the title page in preprint mode.
-% Multiple \preprint commands are allowed.
-%\preprint{}
-
-\title{Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization} %Title of paper
-
-% repeat the \author .. \affiliation  etc. as needed
-% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
-% Explanatory text should go in the []'s, 
-% actual e-mail address or url should go in the {}'s for \email and \homepage.
-% Please use the appropriate macro for the type of information
-
-% \affiliation command applies to all authors since the last \affiliation command. 
-% The \affiliation command should follow the other information.
-
-\author{Josef Melcr}
-\author{Hector Martinez-Seara}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\author{Ji{\v r}{\' i} Kolafa}
-\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
-\author{Pavel Jungwirth}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\affiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
-Tampere, Finland}
-
-\author{O. H. Samuli Ollila}
-\email[]{samuli.ollila@helsinki.fi}
-%\homepage[]{Your web page}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
-\affiliation{Institute of Biotechnology, University of Helsinki}
-
-% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
-% \noaffiliation is required (may also be used with the \author command).
-%\collaboration{}
-%\noaffiliation
-
-\date{\today}
-
-\begin{abstract}
-  % insert abstract here
-Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
-bilayers are of paramount significance in the properties and functionality of cellular
-membranes. Current estimates of binding affinities and
-stoichiometries of cations are, however, controversial due to limitations in the available experimental and
-theoretical methods. Experimentally one can assess these parameters by titrating
-several membrane properties as a function of the ion concentration. However, the
-interpretation of the experiments relies on theoretical models, which give
-the best conception of the ion binding as direct experiments are not available. 
-Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic
-resolution, therefore offering all the necessary information to interpret
-experimental data without the need to resort to simplified models. However, the accuracy
-of the available lipid models when interacting with ions is not sufficient for
-such interpretation. In this work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
-1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
-polarization as a mean field correction, known as the electronic continuum correction (ECC).
-This is applied to the partial charges of a selected state of the art POPC lipid model for MD simulations.
-Our improved ECC-POPC model reproduces not only the
-experimentally measured structural parameters for the ion-free membrane, but also the
-response of lipid headgroup to a bound positive charge, and the binding affinities of 
-Na$^+$ and Ca$^{2+}$ ions. With our new model we also observe negligible binding of Na$^+$ ions to POPC
-bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens,
-in agreement with the previous interpretations of the experimental spectroscopic data.
-The new model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
-equal probabilities, suggesting more complex binding stoichiometry than simple models
-previously used to interpret the NMR data.
-The results of this work pave the way to quantitative MD simulations of complex
-biochemical systems with realistic electrostatic interactions in the vicinity of cellular
-membranes.
-\end{abstract}
-
-%\pacs{}% insert suggested PACS numbers in braces on next line
-
-\maketitle %\maketitle must follow title, authors, abstract and \pacs
-
-% Body of paper goes here. Use proper sectioning commands. 
-% References should be done using the \cite, \ref, and \label commands
-
-\section{Introduction}
-Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}.
-Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
-Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface.
-For example, Ca$^{2+}$  is crucial in neural signal propagation.
-It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}.
-Ca$^{2+}$ also participates in the T-cell receptor activation.
-It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013Ca2}.
-A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
-Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions.
-Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.
-\todo{Hector: I left one unknown citation for Pavel.}
-
-Direct measurements of ion-membrane interactions in real biological systems are difficult. 
-Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
-For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
-The details of ion binding are, however, not fully consistent in the literature.
-Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mainly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
-On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-On the other hand, most atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}.
-For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
-%{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}
-
-Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies.
-A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept~\cite{seelig87}.
-Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results.
-Analyzing massive amounts of data collected by an open collaboration method, 
-it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
-
-In this work, we improve the cation binding in a POPC bilayer in MD simulations 
-via the implicit inclusion of the electronic polarizability in the polar region of phospholipids. 
-For this, we apply an electronic continuum correction (ECC)~\cite{leontyev11}, 
-in which scaling of the atomic partial charges is used to account for the missing electronic polarizability in MD simulations. 
-Such an approach has been previously shown to improve the behavior of ions in bulk water simulations ~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
-As a starting template for our POPC model, we will use the Lipid14~\cite{dickson14}.
-This force field provides one of the best available descriptions of cation binding \cite{catte16}.
-The newly developed ECC-POPC model reproduces the experimentally
-measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state of the art lipid models, while significantly improving the membrane binding affinities to sodium and calcium cations.
-
-\section{Methods}
-
-\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-The lack of electronic polarizability in standard MD simulation force fields has been considered a serious issue since the early days of lipid bilayer simulations.
-In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by including the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
-Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
-However, the present concept of ECC is physically well justified and has
-obtained theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
-
-According to ECC, the electronic polarizability can be modeled in non-polarizable MD in a mean-field way by 
-embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
-Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, giving
-\begin{equation}
-Q^{ECC} = f_q \cdot Q
-\end{equation}
-for the ECC corrected charges in simulation. 
-Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor successfully improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
-It is important to note that the 
-high frequency dielectric constant 
-is around 2 for almost any biologically relevant environment \cite{leontyev11}.
-The dielectric discontinuity in a lipid bilayer arises
-from the orientational polarization of the molecules. 
-Thus, the same correction for the electronic polarizability can be 
-applied throughout the lipid bilayer interface.
-
-While using a scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case.
-Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
-Several schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, where the restrained electrostatic potential method (RESP) is the most commonly used~\cite{RESP_paper, Singh1984}.
-Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effect of the solvent is included in modern force fields to some extent~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
-Thus, the application of the ECC scaling factor, $f_q$, to the existing partial charges in molecules does not universally follow $f_q = \epsilon _{el} ^{-1/2}$.
-Instead, a consistent scaling factor would lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
-
-Here we develop ECC-POPC lipid model that accurately describes the binding 
-of sodium and calcium ions to the lipid bilayer. 
-The Lipid14~\cite{dickson14} force field parameters 
-(available in a Gromacs format from Ref. \citenum{lipid14files}) were used as a starting 
-point. 
-This model provides the best response of the head group to ions among the available 
-lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
-has a relatively realistic head group, glycerol backbone, and acyl chain structure~\cite{dickson14,botan15}.
-We applied the ECC correction 
-to the Lipid14 model of POPC by scaling of 
-partial charges of the head group, glycerol 
-backbone, and carbonyl regions. 
-These are the polar parts of phospholipids which can 
-contribute to the cation binding.
-
-To reproduce the experimental ion binding affinities,
-we scanned the optimal interval for the scaling factor, $f_q~\in~(0.75, 1.0)$.
-The ion binding affinity was evaluated against experiments
-using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
-as discussed more detail in the next section.
-Scaling down the partial charges reduced the ion binding affinity.
-We found an optimal binding affinity with the scaling factor of $f_q = 0.8$,
-which is only slightly higher than the theoretical one ($f_q=0.75$).
-Directly applying
-the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
-the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
-original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
-(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
-reduced hydration of the lipid head group region after scaling charges which effectively
-reduces head group polarity. We solve this difficulty by reducing the effective radii of
-the modified atoms by lowering the $\sigma$ parameters in the Lennard-Jones potential by a
-factor of $f_\sigma = 0.89$. The same solution was taken for the ECC-ions in aqueous
-solvent \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the $\sigma$ parameters,
-the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
-
-\subsection{Electrometer concept} \label{section:electrometer}
-Comparing MD simulation to NMR experiments, we can validate the ion
-binding affinity in lipid bilayer simulations using the ''electrometer concept''~ \cite{seelig87, catte16}.
-This method relies on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
-%Comparing this response is an excellent way to validate force fields as both NMR and MD can accurately determine order parameters ~\cite{catte16, ollila16}.
-The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
-From MD simulations the order parameters can be calculated using the definition
-\begin{equation}\label{OP}
-S_{\rm CH}=\frac{3}{2}\langle \cos^2\theta -1 \rangle,
-\end{equation}
-where $\theta$ is the angle between the C-H bond and membrane normal. Angular brackets point to the average over all sampled configurations.
-
-The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
-\begin{equation}\label{OPchangeEQ}
-\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
-\end{equation}
-where $i$ refers to either $\alpha$ or $\beta$ carbon, $S_{\rm{CH}}^{i}(0)$ denote the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and position of the bound charge.
-
-The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).
-The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities.
-This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
-To calibrate the response of the head group order parameter to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammoniumbromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
-DHAB\\[0.5cm]
-%\vspace{0.5cm}
-\chemfig{
-  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
-}  
-\vspace{0.5cm} \\
-is a cationic surfactant having two acyl chains and bearing a unit charge in the hydrophilic end.
-Thus, it is expected to locate in the bilayer similarly to the phospholipids, and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
-
-\subsection{Salt concentrations and binding affinity}
-NMR experiments report the used salt concentrations in two different ways when measuring head group order parameters.
-Some use the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
-while others use atomic absorption spectroscopy and report the salt concentration in
-the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
-%we use the latter definition. In particular,
-we measure the latter concentration in the aqueous
-bulk region using the farthest point from the lipid bilayer in the aqueous phase.
-Notice that the former concentration was used by Catte et al.~\cite{catte16}.
-Although these two definitions vary when applied
-to Ca$^{2+}$, they do not significantly affect the conclusions of this work.
-
-To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}.
-Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing plane between the membrane interior and the water bulk region.
-In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.
-Such a setup and the above definition of bulk ion concentration provide a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
- \begin{equation}\label{surfexcess}
-  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
-\end{equation}
-Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
-$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase,
-and $A_b$ is the area of the unit cell in the membrane plane.
-The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
-because the bilayer has an interface at each of the two leaflets.
-
-\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
-The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and \mbox{x-ray} scattering experiments by the scattering form factors.
-NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution.
-The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq.
-\ref{OP}.
-Scattering form factors validate the dimensions of the lipid bilayer (e.g., thickness and area per molecule).
-Form factors were calculated using
-
-\begin{equation}
-  F(q) = \int _{-D/2} ^{D/2} \left ( \rho_{el}(z) - \rho_{el}^s \right ) \cos (zq_z) \mathrm{d}z,
-  %F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
-\end{equation}
-
-\noindent where $\rho_{el} (z)$ is the total electron density, $\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk, and $z$ is the distance from the membrane center along its normal with $D/2$ being half of the unit cell size.
-
-\subsection{Simulation details}
-
-\subsubsection{Simulations of POPC bilayers with aqueous ions}
-Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
-and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
-The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
-in the main text. This model was also used in the previous parametrization
-of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
-the ECC concept~\cite{leontyev11, leontyev14}.
-The robustness of our approach with other water models (OPC \cite{Izadi14},
-OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005})
-is tested in the SI.
-Sodium, calcium, and chloride ions were modeled as ECC-ions
-with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
-forming the POPC lipid are derived in this work starting from the Lipid14 force field~\cite{dickson14}.
-For comparison, simulations with the Lipid14 model~\cite{dickson14} and
-ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
-were also performed. The TIP3p water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
-Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and TIP3P~\cite{jorgensen83} water models were taken directly from
-Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.  
-% SAMULI: I putted the trajectory without ions back, because it has been used as well.
-
-MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
-in this work are summarized in Table~S1. 
-Compatibility with the openMM simulation package \cite{openmm7} was also
-tested in SI. 
-The simulated trajectories and parameters files are available at 
-Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files} 
-\todo{To be uploaded to Zenodo, references are only placeholders. }.
-
-\subsubsection{Simulations of POPC bilayers with cationic surfactants}
-The topology of dihexadecyldimethylammonium was created with the automated topology builder \cite{malde11}.
-The AmberTools program~\cite{amber} was then used to generate the Amber-type force field parameters.
-These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
-The partial charges were then manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulation with ECC-POPC, the partial charges of the headgroup were scaled, i.e., $f_q=0.8$, and the $\sigma$ of the scale atoms reduced as in POPC-ECC.
-
-The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
-All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
-Chloride counter ions were used in simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
-For POPC either the Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987}  were used.
-The first 20~ns of the total simulation time of 200~ns was considered as an equilibration time and was omitted from the analysis.
-Sufficient lipid neighbor exchange occurred during the simulations.
-The simulated trajectories and parameters files are available at Ref.~\citenum{ECC-POPC_cationic-surf_files} 
-\todo{To be uploaded to Zenodo, references are only placeholders. }.
-The trajectories and simulation files for the unscaled systems are available from
-Refs.~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
-
-\section{Results and Discussion}
-
-\subsection{Structural parameters of pure ECC-POPC model membrane are in agreement with experiments}
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89_POPC-struct.pdf}
-  %\includegraphics[width=7.4cm]{../Fig/POPCstructure.eps}
-  %\includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
-  %\includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
-  \caption{\label{simVSexpNOions}
-    Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K. \\
-    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
-    compared with experiments \cite{ferreira13} at 300~K.
-    The size of the markers for the head group order parameters correspond to
-    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
-    while the error estimate for simulations is $\pm 0.005$ 
-    (Bayesian estimate of 95\% confidence interval \cite{scipy}).
-    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
-    Open/closed symbols are used for palmitoyl/oleoyl chains of POPC. \\
-    Bottom: Chemical structure of POPC and the labeling of hydrocarbon segments.
-  } 
-\end{figure}
-
-\begin{table}[tb!]
-  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
-  }
-  \begin{tabular}{l|c c}
-    model          & APL (\AA$^2$)   & Temperature [K] \\
-    \hline
-    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
-    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
-    \hline
-    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
-    \hline
-    experiment \cite{kucerka11} & 64.3  &  303    \\
-    \hline
-  \end{tabular}
-\end{table}
-
-First, we present results for bilayers in pure water.
-The ECC-POPC and Lipid14 models both reproduce the experimental x-ray scattering form factors
-of a POPC bilayer with a comparable accuracy in Fig.~\ref{simVSexpNOions}.
-The area per lipid from the Lipid14 model is $\approx$1\AA{} larger than the
-experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
-is $\approx$1\AA{} smaller
-The values of the area per lipid of the ECC-POPC model vary
-when simulated with different water models (i.e., 62.2--66.8 \AA{}, see Table~S2 in SI),
-while still being close to the experimentally reported values.
-We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
-lipid bilayer with the comparable accuracy to other state of the art lipid models~\cite{ollila16}.
-
-Similarly, the acyl chain order parameters of the ECC-POPC model, as well as the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
-Notably, the experimentally measured forking and small order parameter values of $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
-This feature has been suggested to indicate that the carbonyl of {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in {\it sn}-1 chain, which would orient more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
-While this arrangement may be a relevant feature for the ion binding details, it is not fully reproduced by other available lipid models~\cite{ollila16}.
-%This is expected as the ECC model does not modify the already highly optimized parameters of this region in Lipid14
-
-The order parameters of the $\alpha$ and $\beta$ carbons in the headgroup are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the ECC-POPC model, see Fig.~\ref{OrderParameterCHANGESsurf}.
-Considering the available experimental evidence, it is not possible to suggest which of the two models provides more realistic headgroup orientation.
-The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
-The accuracy of both models in the glycerol backbone region is equally footed with any other state of art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
-
-%\todo{Dynamics check is missing: MSD (Hector/Joe)}
-
-\subsection{Calibration of head group response to membrane-bound charge using cationic surfactant}\label{section:boundCHARGE}
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
-  \caption{\label{OrderParameterCHANGESsurf}
-    The changes of headgroup order parameters and P-N vector orientation as a function of
-    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
-    from simulations and experiments \cite{scherer89} at 313 K.
-  }
-\end{figure}
-
-Before studying the sodium and calcium ion binding affinities, we quantify the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
-These mixtures have a well-defined amount of bound charge per PC, i.e., the molar fraction of cationic surfactants.
-This direct relation results from the ability of dihexadecyldimethylammonium to be inserted in the lipid bilayers as any lipid because of its two hydrophobic acyl chains.
-Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid headgroup order parameters to the amount of bound charge in simulations \cite{scherer89}.
-
-The changes of the headgroup order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
-An approximately linear decrease of the order parameters, as expected from Eq.
-\ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
-The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters is too large to the bound positive charge.
-In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.
-
-In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
-As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in a PC lipid bilayer.
-The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
-For example, the addition of 50\% mole fraction of the cationic surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only 11$^{\circ}$ in the ECC-POPC model.
-The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model.
-The lower sensitivity of the P-N vector angle response in the ECC-POPC model agrees better with experiments.
-%The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules.
-%{\color{red}., as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
-
-\subsection{Validation of ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
-  \caption{\label{fig:delta_ordPar_NaCl}
-    Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
-    in bulk ($C_{ion}$) from simulations with different force fields at 313 K together with 
-    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
-    %Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
-    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
-  }
-  %Figure. in Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.
-\end{figure}
-
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
-  \caption{\label{fig:delta_ordPar_CaCl}
-    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration in a bulk water ($C_{ion}$)
-    from simulations at 313~K together with experimental data 
-    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    The error estimate for bulk concentrations is approximately 10\,mM.
-    The order of magnitude larger error in the
-    simulation with Lipid14 and ECC-ions is due to unconverged bulk densities  
-    (shown if Fig.~\ref{fig:cacl-dens}) limited by the simulation box. 
-    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298~K are taken directly from
-    Refs.~\citenum{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
-  }
-  %Experimental values from \cite{akutsu81} for $g_3$ figure:
-  %the value of $g_3$ order parameter of DPPC was -0.214 
-  %in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
-  %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
-  %at the C-2 segments of both fatty acyl chains.'' 
-\end{figure}
-
-Changes of the lipid bilayer head group order parameters from different simulations and
-experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
-and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
-As seen in Fig.~\ref{OrderParameterCHANGESsurf}, the order parameters decrease
-proportionally to the amount of the bound positive charge.
-These results can be thus used to compare the ion binding affinities to lipid bilayers between
-simulations and experiments using the electrometer concept~\cite{seelig87, catte16}.
-
-The experimentally measured small order parameter
-changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) %and the negligible Na$^+$ binding to PC bilayers \cite{akutsu81, altenbach84}.
-are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
-However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
-%
-%A recent comparison of different simulation
-%models to the experimental data revealed that most force fields significantly
-%overestimate Na$^+$ ion binding to PC lipid bilayers and that none of the available
-%models correctly reproduces the details of the binding of Ca$^{2+}$ ions which is
-%grossly overestimated \cite{catte16}.
-%
-%Since artifacts were reported
-%in simulations involving \AA{}qvist ions is water \cite{auffinger07}, we tested the
-%
-Replacing \AA{}qvist ions with ions by Dang et al.~\cite{smith94, chang1999, dang2006}
-or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
-the results~(Figs.~\ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}).
-In line with the previous work \cite{catte16}, the results suggest that improvements
-in the lipid parameters are required to correctly describe the binding of cations to phospholipid bilayers.
-
-%To improve the cation binding behaviour to POPC lipid bilayer we implicitly included the electronic
-%polarization to the Lipid14 model by using the ECC approach as described in section \ref{section:ecc}.
-
-The results from the simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as functions of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
-Considering that we can also reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC-model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
-Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-%\todo{HECTOR: I do not believe in the content of the following sentence as it is.
-How do we know that the changes of the P-N angle are too pronounced for Lipid14?.
-There is no direct correlation between angle and order parameter.}
-% SAMULI: The sentence is now modified.
-Also the changes of the P-N vector angle are more pronounced for the Lipid14 model,
-for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
-CaCl$_2$ concentration is approximately 17$^{\circ}$. The corresponding value
-for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$). 
-%\todo{HECTOR: I have a problem when comparing fig.
-2 and 3.
-If I read it right the behavior when adding calcium and surfactant is not the same.
-Is the latter changing the area per lipid and therefore the surface charge density?
-If this will be the case then besides the effect of the charge we will have a packing effect.
-This is more relevant for the more concentrated cases.
-Shouldn't we try to discuss the two figures and what the differences and similarities mean?
-Given our previous work, should we check the behavior respect the charge density?}
-%SAMULI: Not now, hopefully in the future.
-
-Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge.
-It is plausible that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance to validate the lipid headgroup order parameter response to the bound charge.
-
-Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI.
-In general, the performance of ECC-POPC with any of the water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3p model.
-
-\subsection{Binding affinity of Na$^+$ and Ca$^{2+}$ cations to POPC membrane}
-\label{sec:affinity}
-
-\begin{table}[tb!]
-  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
-    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
-    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
-  \label{tab:binding}}
-  \begin{tabular}{l|c c | c | c c}
-    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
-    \hline
-    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    25\%    \\
-    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
-    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
-    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
-%Average no. of bound cations:
-%L14+Aqvist: all:22.7, phos:22.7, carb:8.5
-%L14+Dang:   all:29.2, phos:29.2, carb:4.0
-%L14+ECC-ion:all:37.7, phos:37.7, carb:8.5
-  \end{tabular}
-\end{table}
-
-Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
-The values of $\Gamma_{\rm ion}^{\rm water}$
-from different simulations with the same molar concetration of cations with respect
-to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
-As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess for the ECC-POPC model, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller than for the other models, 0.13--0.35~nm$^{-2}$.
-%\todo{HECTOR: I have deleted the sentence relating the uncertainty and the error as in my opinion there are other factors which are more important.
-The most important the slow convergence.
-SAMULI: I think that this was a good idea.}.
-Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having a negative value $\Gamma_{Na}^{water} = (-0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens})
-meaning that water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
-This is in contradiction with most of the available lipid force fields, which predict a positive binding of sodium to PC lipid bilayers \cite{catte16}.
-%\todo{HECTOR: do we need to add a NaCl table or an extended table for CaCl2 using different concentrations in the SI? I think this is very important information
-%  SAMULI: I agree that this is important information, but we have quite a lot of stuff in the paper already. Maybe we could leave this for the future studies.}
-
-\begin{figure}[htbp!]
-  \centering
-  \includegraphics[height=19.9cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
-  \caption{\label{fig:cacl-dens}
-    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis
-    for different force fields.
-    %Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
-    %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
-    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
-    the density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
-    the density profiles of phosphate groups and water are divided by 5 and 200, respectively. 
-    All simulations with \ce{CaCl2} shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
-    The simulation with \ce{NaCl} has $C_{ion}'$=1000~mM.
-    }
-\end{figure}
-
-\subsection{Molecular interactions between Na$^+$ or Ca$^{2+}$ cations and POPC oxygens}
-We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
-A maximum distance of 0.3~nm from any lipid oxygen is used to determine a bound calcium.
-The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in contact with the phosphate oxygens.
-From those only one third (32\%) also interacts with the carbonyl oxygens.
-The interaction of calcium ions with only carbonyl oxygens is rare (1\%).
-Neither individual lipids nor source chain of carbonyl is distinguished in the analysis.
-The most likely interactions between Ca$^{2+}$ ions and phosphate oxygens are visualized with the probability density isocontours in Fig.~\ref{fig:volmaps}.
-While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=9.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
-%  \includegraphics[width=8.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} %\\
-%  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
-%  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
-  \caption{\label{fig:volmaps}
-      Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.0016~\AA$^{-3}$) 
-      and POPC carbonyl oxygen atoms (light semi-transparent red, 0.007~\AA$^{-3}$, all POPC lipids contribute). 
-      Calcium cations localize mostly around phosphate oxygens (red and bronze).
-      Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
-      and it is contributed more by other neighbouring phospholipids than by the same lipid. 
-      Transparent structures are shown to depict the variability of choline configurations 
-      (colour warps from red to blue along the simulation time). 
-      The number density was evaluated for each lipid, 
-      after its strucutral alignment using only phosphate group. 
-      MDAnalysis \cite{mdanalysis2011} library was used for 
-      the calculations of the structural alignment and the spatial number density. 
-      VMD \cite{hump96} was used for visualisation. 
-      Carbon atoms are cyan, hydrogen atoms white, oxygen atoms red, nitrogen blue.
-      %moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
-      %Shown contours suggest that the dominant contribution 
-      %to \ce{Ca^{2+}} binding comes from the phosphate oxygens,
-      %whereas the interactions with any of the two carbonyl groups are considerably milder.
-  }
-  %\todo{The figure should be updated as discussed last week. Joe: Will only smoothen the isosurfaces now. }
-\end{figure}
-
-Even though \ce{Na+} ions do not specifically bind to a POPC bilayer, they still interact mostly with its oxygen moieties.
-The results from the simulation at $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions interact with only phosphate oxygens of POPC and 20\% with only carbonyl oxygens, and the rest, 25\%, is interacting with both.
-
-In conclusion, the results suggest that calcium ions bind specifically to phosphate oxygens, occasionally interacting also with carbonyls.
-This is in good agreement with previous conclusions from several experimental and computational studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}, but suggesting a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
-Sodium ions also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
-%\todo{This last sentence is weak and not relevant.
-If you want to keep it go ahead, but then I would ask you a simple question: Is it incompatible to have no overall binding (this relates to the whole surface volume) with having small cavities in the membrane with an increase Na+ concentration.
-If the answer is negative maybe the statement is incorrect and who knows it might be relevant for something.
-After all even Ca2+ affinity changes in the presence of sodium.} The physiological relevance of these details of sodium interactions is, however, unclear due to its overall very weak binding affinity.
-
-\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
-
-%Binding probability:
-%\ce{Ca2+} complex with \\
-%1 POPC - 13\% -- 30\% \\
-%2 POPC - 18\% -- 42\% \\
-%3 POPC - 12\% -- 28\% \\
-
-Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} used in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
-In particular, NMR data about PC headgroup order parameters response and atomic absorption spectroscopy were best explained using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-However, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering only \ce{CaCl2} at low concentrations~\cite{macdonald87}.
-%\todo{HECTOR: Why is the following sentence remotely significant.
-I will vote for removing it} Note that the results based on these binding models and the measured headgroup order parameter changes are also included in the tabulated experimental binding constants~\cite{marsh13}.
-
-In this work, we reproduce the same experimental data used to calculate binding stoichiometry with our ECC-POPC model.
-Using our simulations, we have direct access to exact details of the binding stoichiometry without requiring any middle model as in experiments~\cite{altenbach84, macdonald87}.
-To evaluate the relative propensities for each stoichiometric complex (i.e.,~1~Ca$^{2+}$:~n~POPC), we calculated for each bound Ca$^{2+}$ the number of POPC molecules within a distance of 0.3~nm.
-Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
-We found the largest propensity for the 1:2 complex (41\%).
-However, the probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~are only slightly lower.
-This variability suggests a more complex binding model than the previously considered in a simple ternary complex model.
-With a broad brushstroke, the simulation data can, however, be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
-This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
-  \caption{\label{fig:cacl_complexes}
-      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
-      with a certain number of POPC lipids. 
-      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
-      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
-  }
-\end{figure}
-
-The probabilities of different complexes formed by \ce{Na+} ions and POPC
-analyzed from the simulation with the ECC-POPC model at $1\,$M concentration of NaCl are also 
-shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
-probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
-for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.
-
-\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in POPC membrane}
-
-Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current state of the art force fields, e.g., CHARMM36 and Slipids force fields~\cite{javanainen17}.
-This suggests that at least several microseconds are required to reach the binding/unbinding equilibrium.
-%\todo{HECTOR: I disagree with the following lines.
-In the case of most of the unscale force fields the issue is that the binding of Calcium to the membrane is very strong.
-Therefore, the reorganization takes too long leading to a very slow equilibration process.
-In this case, the size of the hill in the choline region is nothing compared to the size between the bound and solvated state which is the real bottleneck.
-Even in the case of the scaled simulations, I believe the same applies but to a lesser extent.
-I might be wrong in this last part.
-If you do not agree we can discuss it tomorrow via skype.} {\color{red} This is in line with the density profiles of \ce{Ca^{2+}} from simulations in Fig.~\ref{fig:cacl-dens}, which all show a region of a very small calcium density around the level of choline groups in POPC (i.e., around $2.5\,\mathrm{nm}$).
-In contrast, simulations with  ECC-lipids exhibits only a relatively shallow local minimum at that level.
-This suggests that the smaller barrier for the exchange of calcium ions shall lead to their faster exchange with the solvent within ECC-lipids.}
-To quantify the exchange of ions between the membrane and solvent in simulations, we thus calculated residence times of ions bound to the membrane.
-In our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
-
-%\todo{HECTOR: I do fully understand the content of this paragraph. I will discuss this with Joe tomorrow. Please be carefull with my changes.} 
-The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
-ECC-POPC and CHARMM36 (Simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
-In the CHARMM36 simulation, a significant number of the calciums are bound for the whole length of the trajectory (800~ns).
-In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
-where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
-residence time is around 150~ns, which is well below the total length of the simulation used for analysis, i.e., 200~ns.
-Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
-headgroup is shorter than $10\,\mu\mathrm{s}$~\cite{altenbach84}. The exchange of \ce{Na+} ions at the POPC membrane
-is another order of magnitude faster, yielding 90\% of the residence times smaller than~1~ns, with the longest residence time being~6~ns.
-
-In conclusion, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within $\sim$100~ns timeframe, which is significantly faster than observed in most of the available lipid models~\cite{javanainen17}.
-Sodium cations exhibit an even more rapid exchange.
-This all suggests that simulations with a length of several hundreds of nanoseconds are sufficient to simulate ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
-This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and calcium cations \cite{javanainen17, catte16}.
-%Such a finding changes the point of view on binding of calcium to PC membranes from
-%a very strong long-term stable binding with rare exchanges to 
-%a relatively frequent exchange of clacium cations 
-%in equilibrium between membrane and solvent.
-
-\section{Conclusions}
-
-In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a classical MD simulation force field, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
-While the structural details of a POPC lipid bilayer simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the model also reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ concentrations.
-It thus represents a significant improvement over other available lipid models, which overestimate cation binding affinities~\cite{catte16}.
-The ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
-
-The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model.
-In line with previous interpretations of experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens.
-However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data, where one calcium binds to two POPC molecules~\cite{altenbach84}.
-While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to relatively abundant and also almost equally likely.
-While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the complex nature of calcium binding to phospholipid bilayers observed in the present simulation.
-
-Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions.
-To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules should be verified and, if necessary, further adjustments following ECC concept to the force fields of the interacting molecules with the lipids should be addressed.
-
-This work can be reached as a repository containing all data and scripts for analysis at \url{zenodo.org:\dots\dots\dots}.
-
-% If you have acknowledgments, this puts in the proper section head.
-\begin{acknowledgments}
-% Put your acknowledgments here.
-P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
-and 600 from the Academy of Finland via the FiDiPro award.
-J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.
-Computational resources were supplied by the Ministry of Education, Youth and Sports
-of the Czech Republic under the Projects CESNET (Project No. LM2015042) and CERIT-Scientific
-Cloud (Project No. LM2015085) provided within the program Projects of Large Research,
-Development and Innovations Infrastructures.
-O.H.S.O. acknowledges financial support from
-Integrated Structural Biology Research Infrastructure of
-Helsinki Institute of Life Science (Instruct-HiLIFE).
-\end{acknowledgments}
-
-%\newpage
-%\appendix
-
-% Create the reference section using BibTe
-\bibliography{refs.bib}
-
-\listoftodos
-
-\end{document}
-%
-% ****** End of file aiptemplate.tex ******

### 2017-12-19 Manuscript/manuscript.tex
Cosmetic updates by Pavel

@@
 \title[An \textsf{achemso} demo]
-  {Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization}
+  {Accurate Binding of Sodium and Calcium to Phospholipid Bilayers by Effective Inclusion of Electronic Polarization}
 
@@
 Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
-However, the present concept of ECC is physically well justified and has obtained theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
+However, the present concept of ECC is physically well justified and has theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
 
@@
 which is only slightly higher than the ECC one ($f_q=0.75$).
-Note that common scaling factors for monovalent ions in water are 0.8 or
-larger \cite{benavides17,skinner14,nacleps}.  In contrast, modern force fields
-for ionic liquids suggest values 0.6--0.65 which are lower than $\epsilon^{-1/2}_{el}$\cite{holm14}.
+Note that common empirical scaling factors for monovalent ions in water are 0.8 or
+even closer to unity \cite{benavides17,skinner14,nacleps}.  In contrast, modern force fields
+for ionic liquids suggest values of 0.6--0.65, which are lower than $\epsilon^{-1/2}_{el}$\cite{holm14}.
 Directly applying
@@
 $C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase
-(measured from the density of farthest point between periodic bilayers),
+(estimated from the solution density of farthest point between periodic bilayers),
 and $A_b$ is the area of the unit cell in the membrane plane.
@@
 tested and is presented in the SI. 
-The simulated trajectories and parameters files are available at
+For deposited trajectories and parameter files see 
 Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files}.
@@
 We checked that a sufficient lipid  exchange occurred during the simulations.
-The simulated trajectories and parameters files are available at Ref.~\citenum{ECC-POPC_cationic-surf_files} 
+For deposited trajectories and parameter files of scaled and unscaled simulations see
+Refs.~\citenum{ECC-POPC_cationic-surf_files} and~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}, respectively.
 \todo{To be uploaded to Zenodo, references are only placeholders. }.
-The trajectories and simulation files for unscaled systems are available at
-Refs.~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
 
@@
     The number density was evaluated for each lipid, 
-    after its strucutral alignment using only phosphate group.
+    after its structural alignment using only phosphate group.
     MDAnalysis \cite{mdanalysis2011} library was used for 
@@
     VMD \cite{hump96} was used for visualisation. 
-    Carbon atoms are cyan, hydrogen atoms white, oxygen atoms red, nitrogen blue.
+    Carbon atoms are depicted in cyan, hydrogen atoms in white, oxygen atoms in red, nitrogen in blue.
   }

### 2017-12-19 Manuscript/manuscript.tex
Updated Zenodo references with simulation files ECC-POPC

--- 2017-12-19 Manuscript/manuscript.tex
@@
 Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files}.
-\todo{To be uploaded to Zenodo, references are only placeholders. }.
 
@@
 The simulated trajectories and parameters files are available at Ref.~\citenum{ECC-POPC_cationic-surf_files} 
-\todo{To be uploaded to Zenodo, references are only placeholders. }.
 The trajectories and simulation files for unscaled systems are available at

### 2018-02-23 Manuscript/manuscriptFIRSTsubmission.tex
Updates after the reviewers' comments

--- 2017-12-19 Manuscript/manuscript.tex
@@
 \usepackage{epstopdf}
-\usepackage[obeyFinal]{easy-todo}
 \usepackage{csquotes}
@@
 Refs.~\citenum{ECC-POPC_cationic-surf_files} and~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}, respectively.
-\todo{To be uploaded to Zenodo, references are only placeholders. }.
 
@@
   \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89_POPC-struct.pdf}
-  %\includegraphics[width=7.4cm]{../Fig/POPCstructure.eps}
-  %\includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.pdf}
-  %\includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
   \caption{\label{simVSexpNOions}
@@
 
-%\todo{Dynamics check is missing: MSD (Hector/Joe)}
-
 \subsection{Calibration of head group response to membrane-bound charge using cationic surfactant}\label{section:boundCHARGE}
@@
     experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
-    %Ion concentrations in x-axis are calculated as in Fig.~\ref{fig:delta_ordPar_CaCl}.
     Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
@@
   }
-  %Figure. in Ref.\cite{altenbach84} gives for POPC at 1M of NaCl $\Delta S^\alpha=(6.1-5.5)*0.00784=0.004704$.
 \end{figure}
@@
   }
-  %Experimental values from \cite{akutsu81} for $g_3$ figure:
-  %the value of $g_3$ order parameter of DPPC was -0.214 
-  %in the absence of ions [average of two closely spaced splittings] and -0.211 in the presence
-  %of 0.35 M CaCl$_2$ (at 59 Celcius). No effect of ions could be detected on DPPC bilayers labeled
-  %at the C-2 segments of both fatty acyl chains.'' 
 \end{figure}
@@
     Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
-%Average no. of bound cations:
-%L14+Aqvist: all:22.7, phos:22.7, carb:8.5
-%L14+Dang:   all:29.2, phos:29.2, carb:4.0
-%L14+ECC-ion:all:37.7, phos:37.7, carb:8.5
   \end{tabular}
@@
   \includegraphics[width=8.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
-%  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
-%  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
   \caption{\label{fig:volmaps}
@@
   }
- %\todo{The figure should be updated as discussed last week. Joe: Will only smoothen the isosurfaces now. }
 \end{figure}
@@
 \subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
-
-%Binding probability:
-%\ce{Ca2+} complex with \\
-%1 POPC - 13\% -- 30\% \\
-%2 POPC - 18\% -- 42\% \\
-%3 POPC - 12\% -- 28\% \\
-
 Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} as employed in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
@@
 
-This work can be reached as a repository containing all data and scripts for analysis at \url{zenodo.org:\dots\dots\dots}.
+This work can be reached as a repository containing all data and scripts for analysis at Ref. \citenum{github}.
 
@@
 
-\listoftodos
-
 \end{document}

### 2018-03-12 Manuscript/SI.tex
Updates based on reviewers' comments

@@
 \author{Hector Martinez-Seara}
+\author{Ricky Nencini}
 \affiliation{Institute of Organic Chemistry and Biochemistry,
@@
 
-\title{Supporting Information:\\Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization}
+\title{Supporting Information:\\ Accurate Binding of Sodium and Calcium to a POPC Bilayer by Effective Inclusion of Electronic Polarization}
 
@@
 \newpage
+\section{Electrostatic potential of lipid bilayer}
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=16.0cm]{../Fig/potential.eps}
+  \caption{\label{potential}
+    Electrostatic potential calculated by evaluating the double integral of the charge density
+    using {\it gmx potential} tool from Gromacs package \cite{gromacsMANUAL}.
+    Comparing the Lipid14 and ECC-POPC simulations with TIP3P water model
+    reveals that applying ECC to the headgroup and glycerol backbone charges slightly decreases
+    the dipole potential inside the bilayer. On the other hand, ECC-POPC simulation
+    with SPC/E water model has larger dipole pontential than with TIP3P.
+    The stronger dependence of water model is expected as the dipole potential
+    mainly originates from orientation of water molecules \cite{gawrisch92}.
+  }
+\end{figure}
+
+\newpage
+\section{Binding of potassium to ECC-POPC lipid bilayer}
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=13.0cm]{../Fig/density_profiles_KCl-water.eps}
+  \caption{\label{Kdens}
+    Number density profiles of K$^+$, Cl$^-$ and water along membrane normal
+    axis from ECC-POPC simulation with KCl ($C_{ion}'$=1000mM, simulation parameters as in Table S1).
+    Parameters for ions are taken from Ref. \citenum{mason12}.
+    In consistent with figure 5 in the main text, the density profiles of ions and water
+    are divided with 2 and 200, respectively. Similarity of potassium
+    and chloride curves together with the bulk concentrations indicate lower binding 
+    affinity for potassium than for sodium (Fig. 5 in the main text).
+    Since this is in line with wide range of experimental data \cite{eisenberg79,binder02,klasczyk10}, we conclude
+    that the ECC-POPC model reproduces the correct binding affinity also for potassium.
+    In line with the low binding affinity, the headgroup order parameters were not
+    affected by the presense of KCl in simulations. The order parameter 
+    data is not shown because we are not aware of experimental data for PC headgroup
+    order parameters as a function of KCl concentration.
+  }
+\end{figure}
+
+\newpage
 \bibliography{refs.bib}

### 2018-03-12 Manuscript/manuscript.tex
Remove todo list and update pdf
Updates based on reviewers' comments

--- 2017-12-19 Manuscript/manuscript.tex
@@
 \title[An \textsf{achemso} demo]
-  {Accurate Binding of Sodium and Calcium to Phospholipid Bilayers by Effective Inclusion of Electronic Polarization}
+  {Accurate Binding of Sodium and Calcium to a POPC Bilayer by Effective Inclusion of Electronic Polarization}
 
@@
 
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
+%Some journals require a graphical entry for the Table of Contents.
+%This should be laid out ``print ready'' so that the sizing of the
+%text is correct.
 
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
+%Inside the \texttt{tocentry} environment, the font used is Helvetica
+%8\,pt, as required by \emph{Journal of the American Chemical
+%Society}.
 
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
+%The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
+%permitted for  \emph{Journal of the American Chemical Society}
+%graphical table of content entries. The box will not resize if the
+%content is too big: instead it will overflow the edge of the box.
 
-This box and the associated title will always be printed on a
-separate page at the end of the document.
+%This box and the associated title will always be printed on a
+%separate page at the end of the document.
 
+%\begin{figure}[tb!]
+%  \centering
+  \includegraphics[width=9.0cm]{../Fig/toc_figure.pdf}
+%  \caption{\label{tocFIG}
+%    The changes of head group order parameters and P-N vector orientation as a function of
+%    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
+%    from simulations and experiments \cite{scherer89} at 313 K.
+%  }
+%\end{figure}
 \end{tocentry}
@@
 Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
-For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
+For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD)
+simulations~\cite{bockmann03, bockmann04, sachs04, berkowitz06, cordomi08,cordomi09,vacha09a,valley11,Berkowitz12,knecht13, melcrova16, javanainen17}.
 The details of ion binding are, however, not fully consistent in the literature.
@@
 contribute to the cation binding.
+The similar approach is applied  also to the CHARMM36 model of POPC~\cite{klauda10} in SI.
 
@@
 The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
-The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters.
-These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
-The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.
+The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters. These parameters were
+then converted to the Gromacs format with the acpype tool \cite{acpype}. The partial charges were manually modified in
+the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
+For simulations of cations mixed with ECC-POPC, the ECC version of cation force field was made
+by scaling the head group parameters with the same scaling factors that were
+used for ECC-POPC, i.e., $f_q=0.8$ for charges and 0.89 for $\sigma$.
 
@@
 The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
-
-%\todo{Dynamics check is missing: MSD (Hector/Joe)}
+The electrostatic potential of lipid bilayer is compared between ECC-POPC and Lipid14 simulations in SI.
 
@@
 In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
-As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in the PC lipid bilayer.
+As suggested previously~\cite{seelig87} and observed in MD simulations \cite{cordomi08,cordomi09,gurtovenko05,zhao12}, the headgroup orients
+more towards the water phase with the increasing amount of positive charge in
+the PC lipid bilayer.
 The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
@@
 Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
+Also the binding affinity of K$^+$ ions to POPC bilayer is in line with the experimental data \cite{eisenberg79,binder02,klasczyk10}, as shown in SI.
 Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
@@
 \begin{table}[tb!]
-  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
+  \caption{Bulk concentrations of Ca$^{2+}$ ($C_{ion}$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
     and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
-    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
+    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350 mM).
   \label{tab:binding}}
@@
     \hline
-    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    25\%    \\
+    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    32\%    \\
     Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
@@
 As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess of calcium, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller for the ECC-POPC model than for the other models, 0.13--0.35~nm$^{-2}$.
-Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}.
+Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}).
 This
@@
 Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
-headgroup is of the order of $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
+headgroup is less than $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
 is yet another order of magnitude faster, with 90\% of the residence times smaller than~1~ns and the longest residence time being~6~ns.
@@
 It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
+Similar improvement is demonstrated for the POPC model of CHARMM36 \cite{klauda10} in SI.
 
@@
 To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules needs to be verified and, if necessary, further adjustments following the ECC concept to the force fields of the biomolecules interacting with the lipids should be made.
+In future studies, the ECC concept will be applied also to bilayers composed of other lipids than POPC.
 
-This work can be reached as a repository containing all data and scripts for analysis at \url{zenodo.org:\dots\dots\dots}.
+This work can be reached as a repository containing all data and scripts for analysis at Ref. \citenum{github}.
 

### 2018-03-13 Manuscript/SI.tex
Updates to repy reviwers' comments

@@
 \newpage
+\section{Electronic continuum correction applied to CHARMM36 model}
+In order to demonstrate that ECC can be applied also to other than Lipid14 model,
+we scaled the charges and Lennard-Jones radii of headgroup, glycerol backbone and
+carbonyl atoms of the CHARMM36 model of POPC~\cite{klauda10}. We used the
+scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
+radius. Changes of the headgroup order parameters as a function of added ions
+and number density profiles along membrane normal from the ECC-CHARMM model are shown 
+in Figs. \ref{scaledCHARMM} and \ref{scaledCHARMMdensity}, respectively,
+together with the results from CHARMM36 simulations \cite{catte16,charmmPOPC450mMCaClfiles}
+and experiments \cite{akutsu81,altenbach84}. As shown previously \cite{catte16},
+the binding of Ca$^{2+}$ is significantly overestimated in CHARMM36 simulations.
+In ECC-CHARMM simulations, the Ca$^{2+}$ binding affinity 
+is significantly reduced (Fig. \ref{scaledCHARMMdensity}) and the headgroup order parameter response to the added CaCl$_2$
+ is in better agreement with experiments (Fig. \ref{scaledCHARMM}).
+In line with the results in main text,
+Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration time
+for Ca$^{2+}$ binding in ECC-CHARMM model than in the original CHARMM36.
+
+In conclusion, applying ECC to CHARMM36 simulation
+improved the ion binding affinity to a lipid bilayer, even thought the
+scaling factors optimized for Lipid14 were used. Also the area per molecule in
+ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
+lower than the experimental value (64.3 $\AA^2$). 
+The optimal scaling factors for CHARMM36 are expected to be different
+than for Lipid14 because the initial partial charges are not the same
+in different models. Our results indicate that the performance of ion
+binding in CHARMM36, and probably in other lipid models as well, could
+be further improved by finding the optimal scaling factors for the partial
+charges and Lennard-Jones radii.
+
+\newpage
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=16.0cm]{../Fig/OP.pdf}
+  \caption{\label{scaledCHARMM}
+    Changes of the headgroup order parameters as a function of NaCl (left) and CaCl$_2$ (right)
+    from simulations with CHARMM36 \cite{klauda10}, simulations with ECC-CHARMM parameters and experiments \cite{akutsu81,altenbach84}. 
+    CHARMM36 force field parameters for simulations with and without NaCl were taken from CHARMM-GUI \cite{lee15}.
+    These were ran 320 ns and 280 ns, respectively, and the first 100 ns was omitted from the analysis.
+    Results for CHARMM36 with CaCl$_2$ were calculated from the last 100 ns from the data available at Ref.~\citenum{charmmPOPC450mMCaClfiles}.
+    ECC-CHARMM simulations with NaCl and CaCl$_2$ were ran 260 ns and 600 ns, respectively, and the first 100 ns were omitted from the analysis.
+    Sodium, calcium, and chloride ions were modeled as ECC-ions in these simulations with parameters from Refs. \citenum{martinek17, Pluharova2014}.
+    All the other simulation parameters in simulations in this figure were equal to the available dataset at Ref.~\citenum{charmmPOPC450mMCaClfiles}.
+    The temperature in all simulations was set to 310 K.
+  }
+\end{figure}
+
+\newpage
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=10.0cm]{../Fig/density2.pdf}
+  \caption{\label{scaledCHARMMdensity}
+    Number density profiles along membrane normal from CHARMM36 \cite{klauda10} and ECC-CHARMM models with NaCl and CaCl$_2$.
+    The values for the relative surface excess are also shown.
+  }
+\end{figure}
+
+\newpage
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=16.0cm]{../Fig/contacts_ca.pdf}
+  \caption{\label{scaledCHARMMcontacts}
+    Number of bound Ca$^{2+}$ during simulations.
+  }
+\end{figure}
+
+\newpage
 \section{Sodium binding to POPC with the glycerol carbon $g_3$ order parameters and the head group orientation}
@@
 
+\newpage
 \section{Ternary complex model in simulations}

### 2018-03-13 Manuscript/manuscript.tex
Updates to repy reviwers' comments

@@
 \begin{table}[tb!]
-  \caption{Bulk concentrations of Ca$^{2+}$ ($C_{ion}$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
+  \caption{Concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$), bulk concentrations of Ca$^{2+}$ ($C_{ion}$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
     and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
-    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350 mM).
+    in different POPC bilayer models. 
   \label{tab:binding}}

### 2018-03-13 Manuscript/SI.tex
Updates based on reviewers comments

@@
 \section{Electronic continuum correction applied to CHARMM36 model}
-In order to demonstrate that ECC can be applied also to other than Lipid14 model,
+In order to demonstrate that the ECC method can be applied also to other than Lipid14 model,
 we scaled the charges and Lennard-Jones radii of headgroup, glycerol backbone and
@@
 scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
-radius. Changes of the headgroup order parameters as a function of added ions
+radii. Changes of the headgroup order parameters as a function of added ions
 and number density profiles along membrane normal from the ECC-CHARMM model are shown 
@@
 ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
-lower than the experimental value (64.3 $\AA^2$). 
+lower than the experimental value (64.3 $\AA^2$) \cite{kucerka11}. 
 The optimal scaling factors for CHARMM36 are expected to be different
@@
 
-\newpage
+It should be noted that the calcium binding to POPC bilayer in ECC-CHARMM model
+is compared here to the CHARMM36 simulation, which was originally ran for
+the previous study \cite{catte16,charmmPOPC450mMCaClfiles}.
+After the latter simulation has been performed, spesific interaction parameters
+between calcium and oxygen \cite{kim16} have been introduced in CHARMM-gui
+to reduce the binding of calcium to lipid bilayers. These parameters
+reduce the calcium binding to a POPC bilayer to the same level as sodium (data not shown).
+Further studies of ion binding with different parameters are in progress
+in NMRlipids IV project (\url{nmrlipids.blogspot.fi}), but are beyond the scope of this work.
+
+%\newpage
 
@@
     from simulations with CHARMM36 \cite{klauda10}, simulations with ECC-CHARMM parameters and experiments \cite{akutsu81,altenbach84}. 
-    CHARMM36 force field parameters for simulations with and without NaCl were taken from CHARMM-GUI \cite{lee15}.
+    CHARMM36 force field parameters for simulations with and without NaCl were taken from CHARMM-gui \cite{lee15}.
     These were ran 320 ns and 280 ns, respectively, and the first 100 ns was omitted from the analysis.

### 2018-03-15 Manuscript/SI.tex
Updates for reviwers' comments

--- 2018-03-13 Manuscript/SI.tex
@@
 \section{Electronic continuum correction applied to CHARMM36 model}
-In order to demonstrate that ECC can be applied also to other than Lipid14 model,
+In order to demonstrate that the ECC method can be applied also to other models than Lipid14,
 we scaled the charges and Lennard-Jones radii of headgroup, glycerol backbone and
@@
 scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
-radius. Changes of the headgroup order parameters as a function of added ions
+radii. Changes of the headgroup order parameters as functions of added ions
 and number density profiles along membrane normal from the ECC-CHARMM model are shown 
@@
 is significantly reduced (Fig. \ref{scaledCHARMMdensity}) and the headgroup order parameter response to the added CaCl$_2$
- is in better agreement with experiments (Fig. \ref{scaledCHARMM}).
+is in a better agreement with experiments than for the original CHARMM model (Fig. \ref{scaledCHARMM}).
 In line with the results in main text,
-Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration time
-for Ca$^{2+}$ binding in ECC-CHARMM model than in the original CHARMM36.
+Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration times
+for Ca$^{2+}$ binding in the ECC-CHARMM model than in the original CHARMM36.
 
@@
 scaling factors optimized for Lipid14 were used. Also the area per molecule in
-ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
-lower than the experimental value (64.3 $\AA^2$). 
-The optimal scaling factors for CHARMM36 are expected to be different
+the ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
+lower than the experimental value (64.3 $\AA^2$) \cite{kucerka11}. 
+The optimal scaling factors for CHARMM36 are expected to be somewhat different
 than for Lipid14 because the initial partial charges are not the same
 in different models. Our results indicate that the performance of ion
-binding in CHARMM36, and probably in other lipid models as well, could
+binding in ECC-CHARMM36, and probably in other lipid models as well, could
 be further improved by finding the optimal scaling factors for the partial
@@
 
-\newpage
+It should be noted that the calcium binding to POPC bilayer within the ECC-CHARMM model
+is compared here to the CHARMM36 simulation, which was originally done in a 
+previous study \cite{catte16,charmmPOPC450mMCaClfiles}.
+After the latter simulation has been performed, specific interaction parameters
+between calcium and oxygen \cite{kim16} have been introduced in CHARMM-gui
+to reduce the binding of calcium to lipid bilayers. These parameters
+reduce the calcium binding to a POPC bilayer to the same level as that of sodium (data not shown).
+Further analyses of ion binding with different parameters (including the newly modified CHARMM) are in progress
+within the NMRlipids IV project (\url{nmrlipids.blogspot.fi}), being beyond the scope of this work.
+
+%\newpage
 
@@
     The temperature in all simulations was set to 310 K.
+    The simulation data is available at Refs. \citenum{charmm36popcT310K,charmm36popc940mMNaClT310K,ECCcharmm36popcT310K,ECCcharmm36popc940mMNaClT310K,ECCcharmm36popc450mMCaClT310K}.
   }

### 2018-03-15 Manuscript/manuscript.tex
Updates based on reviewers comments
Updates for reviwers' comments

@@
 contribute to the cation binding.
-The similar approach is applied  also to the CHARMM36 model of POPC~\cite{klauda10} in SI.
+A similar approach is also applied to the CHARMM36 model of POPC~\cite{klauda10} in SI.
 
@@
 The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
-The electrostatic potential of lipid bilayer is compared between ECC-POPC and Lipid14 simulations in SI.
+The electrostatic potentials as obtained in ECC-POPC and Lipid14 simulations are compared in the SI.
 
@@
 It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
-Similar improvement is demonstrated for the POPC model of CHARMM36 \cite{klauda10} in SI.
+A similar improvement using the ECC procedure is demonstrated for the POPC model of CHARMM36 \cite{klauda10} in SI.
 

### 2018-03-15 Manuscript/SI.tex
Final changes Hector SI and Manuscript

@@
 \section{Electronic continuum correction applied to CHARMM36 model}
-In order to demonstrate that the ECC method can be applied also to other models than Lipid14,
+To demonstrate that the ECC method can also be applied to other models than Lipid14,
 we scaled the charges and Lennard-Jones radii of headgroup, glycerol backbone and
 carbonyl atoms of the CHARMM36 model of POPC~\cite{klauda10}. We used the
-scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
-radii. Changes of the headgroup order parameters as functions of added ions
-and number density profiles along membrane normal from the ECC-CHARMM model are shown 
-in Figs. \ref{scaledCHARMM} and \ref{scaledCHARMMdensity}, respectively,
-together with the results from CHARMM36 simulations \cite{catte16,charmmPOPC450mMCaClfiles}
+same scaling factors optimized for Lipid14 in the main text, i.e., 0.8 for charges and 0.89 for Lennard-Jones
+radii. In Figs. \ref{scaledCHARMM} and \ref{scaledCHARMMdensity} we show
+changes of the headgroup order parameters as functions of added ions
+and number density profiles along membrane normal for our ECC-CHARMM model,
+together with the CHARMM36 model \cite{catte16,charmmPOPC450mMCaClfiles}
 and experiments \cite{akutsu81,altenbach84}. As shown previously \cite{catte16},
-the binding of Ca$^{2+}$ is significantly overestimated in CHARMM36 simulations.
+the binding of Ca$^{2+}$ is significantly overestimated in CHARMM36 simulations
+where no explicit repulsive Ca$^{2+}$--oxygen Lennard-Jones parameters are included.
 In ECC-CHARMM simulations, the Ca$^{2+}$ binding affinity 
-is significantly reduced (Fig. \ref{scaledCHARMMdensity}) and the headgroup order parameter response to the added CaCl$_2$
-is in a better agreement with experiments than for the original CHARMM model (Fig. \ref{scaledCHARMM}).
-In line with the results in main text,
+is reduced substantially (Fig. \ref{scaledCHARMMdensity}), and the headgroup order parameter response to the added CaCl$_2$
+is in a better agreement with experiments than for the original CHARMM36 model (Fig. \ref{scaledCHARMM}).
 Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration times
 for Ca$^{2+}$ binding in the ECC-CHARMM model than in the original CHARMM36.
+Both results are in line with the findings described in the main text.
 
 In conclusion, applying ECC to CHARMM36 simulation
-improved the ion binding affinity to a lipid bilayer, even thought the
-scaling factors optimized for Lipid14 were used. Also the area per molecule in
+improved the ion binding affinity to a lipid bilayer, despite we
+directly applied the scaling factors optimized for Lipid14 without further optimizations. Also, the area per molecule in
 the ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
@@
 binding in ECC-CHARMM36, and probably in other lipid models as well, could
-be further improved by finding the optimal scaling factors for the partial
+be further improved by finding the optimal values for the partial
 charges and Lennard-Jones radii.
 
-It should be noted that the calcium binding to POPC bilayer within the ECC-CHARMM model
-is compared here to the CHARMM36 simulation, which was originally done in a 
-previous study \cite{catte16,charmmPOPC450mMCaClfiles}.
-After the latter simulation has been performed, specific interaction parameters
-between calcium and oxygen \cite{kim16} have been introduced in CHARMM-gui
-to reduce the binding of calcium to lipid bilayers. These parameters
-reduce the calcium binding to a POPC bilayer to the same level as that of sodium (data not shown).
-Further analyses of ion binding with different parameters (including the newly modified CHARMM) are in progress
-within the NMRlipids IV project (\url{nmrlipids.blogspot.fi}), being beyond the scope of this work.
+It should be noted that the CHARMM36 simulation used here uses the original paramter
+as in our previous study \cite{catte16,charmmPOPC450mMCaClfiles}.
+Currently (February 2018), specific interaction parameters
+between calcium and oxygen \cite{kim16} have been introduced in CHARMM-GUI
+to reduce the binding of calcium to lipid bilayers. The impact of this changes that
+dramatically reduce the calcium binding to levels similar to sodium is beyond the
+scope of this work.
 

### 2018-03-15 Manuscript/manuscript.tex
Final changes Hector SI and Manuscript

@@
 the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulations of cations mixed with ECC-POPC, the ECC version of cation force field was made
-by scaling the head group parameters with the same scaling factors that were
-used for ECC-POPC, i.e., $f_q=0.8$ for charges and 0.89 for $\sigma$.
+For simulations of cationic surfactant mixed with ECC-POPC, the ECC version of the cationic surgactnat force field was made
+by scaling the partial charges and Lennard-Jones parameters of the head group parameters with the same scaling factors as in
+ECC-POPC, i.e., $f_q=0.8$ for charges and 0.89 for $\sigma$.
 

### 2018-03-16 Manuscript/SI.tex
Almost ready for second submission

@@
 changes of the headgroup order parameters as functions of added ions
-and number density profiles along membrane normal for our ECC-CHARMM model,
+and number density profiles along membrane normal for our ECC-CHARMM36 model,
 together with the CHARMM36 model \cite{catte16,charmmPOPC450mMCaClfiles}
@@
 where no explicit repulsive Ca$^{2+}$--oxygen Lennard-Jones parameters are included.
-In ECC-CHARMM simulations, the Ca$^{2+}$ binding affinity 
+In ECC-CHARMM36 simulations, the Ca$^{2+}$ binding affinity 
 is reduced substantially (Fig. \ref{scaledCHARMMdensity}), and the headgroup order parameter response to the added CaCl$_2$
@@
 Fig. \ref{scaledCHARMMcontacts} also shows faster equilibration times
-for Ca$^{2+}$ binding in the ECC-CHARMM model than in the original CHARMM36.
+for Ca$^{2+}$ binding in the ECC-CHARMM36 model than in the original CHARMM36.
 Both results are in line with the findings described in the main text.
@@
 directly applied the scaling factors optimized for Lipid14 without further optimizations. Also, the area per molecule in
-the ECC-CHARMM simulation without ions (61 $\AA^2$) is only slightly
+the ECC-CHARMM36 simulation without ions (61 $\AA^2$) is only slightly
 lower than the experimental value (64.3 $\AA^2$) \cite{kucerka11}. 
@@
 than for Lipid14 because the initial partial charges are not the same
-in different models. Our results indicate that the performance of ion
+in the two models. Our results indicate that the performance of ion
 binding in ECC-CHARMM36, and probably in other lipid models as well, could
@@
 Currently (February 2018), specific interaction parameters
-between calcium and oxygen \cite{kim16} have been introduced in CHARMM-GUI
-to reduce the binding of calcium to lipid bilayers. The impact of this changes that
-dramatically reduce the calcium binding to levels similar to sodium is beyond the
+between calcium and oxygen \cite{kim16} have been introduced in CHARMM-GUI \cite{lee15}
+to reduce the binding of calcium to lipid bilayers. The further discussion of these
+changes that dramatically reduce the calcium binding to levels similar to sodium is beyond the
 scope of this work.
@@
     Changes of the headgroup order parameters as a function of NaCl (left) and CaCl$_2$ (right)
-    from simulations with CHARMM36 \cite{klauda10}, simulations with ECC-CHARMM parameters and experiments \cite{akutsu81,altenbach84}. 
-    CHARMM36 force field parameters for simulations with and without NaCl were taken from CHARMM-gui \cite{lee15}.
+    from simulations with CHARMM36 \cite{klauda10}, simulations with ECC-CHARMM36 parameters and experiments \cite{akutsu81,altenbach84}. 
+    CHARMM36 force field parameters for simulations with and without NaCl were taken from CHARMM-GUI \cite{lee15}.
     These were ran 320 ns and 280 ns, respectively, and the first 100 ns was omitted from the analysis.
     Results for CHARMM36 with CaCl$_2$ were calculated from the last 100 ns from the data available at Ref.~\citenum{charmmPOPC450mMCaClfiles}.
-    ECC-CHARMM simulations with NaCl and CaCl$_2$ were ran 260 ns and 600 ns, respectively, and the first 100 ns were omitted from the analysis.
+    ECC-CHARMM36 simulations with NaCl and CaCl$_2$ were ran 260 ns and 600 ns, respectively, and the first 100 ns were omitted from the analysis.
     Sodium, calcium, and chloride ions were modeled as ECC-ions in these simulations with parameters from Refs. \citenum{martinek17, Pluharova2014}.
@@
   \caption{\label{scaledCHARMMdensity}
-    Number density profiles along membrane normal from CHARMM36 \cite{klauda10} and ECC-CHARMM models with NaCl and CaCl$_2$.
+    Number density profiles along membrane normal from CHARMM36 \cite{klauda10} and ECC-CHARMM36 models with NaCl and CaCl$_2$.
     The values for the relative surface excess are also shown.

### 2018-03-16 Manuscript/manuscript.tex
Almost ready for second submission

--- 2018-03-15 Manuscript/manuscript.tex
@@
 contribute to the cation binding.
-A similar approach is also applied to the CHARMM36 model of POPC~\cite{klauda10} in SI.
+A similar approach is also applied to the CHARMM36 model of POPC~\cite{klauda10} in the the SI.
 
@@
 the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulations of cations mixed with ECC-POPC, the ECC version of cation force field was made
-by scaling the head group parameters with the same scaling factors that were
-used for ECC-POPC, i.e., $f_q=0.8$ for charges and 0.89 for $\sigma$.
+For simulations of cationic surfactants mixed with ECC-POPC,
+all charges of cationic surfactants were scaled by a factor $f_q$=0.75
+consistently with the counterions. The Lennard-Jones radii were scaled with
+the same factor as for ECC-POPC, i.e. $f_{\sigma}$=0.89.
+
+%the ECC version of the cationic surgactnat force field was made
+%by scaling the partial charges and Lennard-Jones parameters of the head group parameters with the same scaling factors as in
+%ECC-POPC, i.e., $f_q=0.8$ for charges and 0.89 for $\sigma$.
+
+%For simulations of THE cations mixed with ECC-POPC,
 
@@
 Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
-Also the binding affinity of K$^+$ ions to POPC bilayer is in line with the experimental data \cite{eisenberg79,binder02,klasczyk10}, as shown in SI.
+Also the binding affinity of K$^+$ ions to POPC bilayer is in line with the experimental data \cite{eisenberg79,binder02,klasczyk10} (shown in SI).
 Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
@@
 It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
-A similar improvement using the ECC procedure is demonstrated for the POPC model of CHARMM36 \cite{klauda10} in SI.
+A similar improvement using the ECC procedure is demonstrated for the POPC model of CHARMM36 \cite{klauda10} in the SI.
 

### 2018-03-20 Manuscript/SI.tex
Second submission to J. Phys. Chem. B

@@
     The temperature in all simulations was set to 310 K.
-    The simulation data is available at Refs. \citenum{charmm36popcT310K,charmm36popc940mMNaClT310K,ECCcharmm36popcT310K,ECCcharmm36popc940mMNaClT310K,ECCcharmm36popc450mMCaClT310K}.
+    The simulation data is available at Refs.~\citenum{charmm36popcT310K,charmm36popc940mMNaClT310K,ECCcharmm36popcT310K,ECCcharmm36popc940mMNaClT310K,ECCcharmm36popc450mMCaClT310K}.
   }
@@
     Number density profiles along membrane normal from CHARMM36 \cite{klauda10} and ECC-CHARMM36 models with NaCl and CaCl$_2$.
-    The values for the relative surface excess are also shown.
+    The values for the relative surface excess are also shown. Simulation details are as in Fig. \ref{scaledCHARMM}.
   }
@@
   \caption{\label{scaledCHARMMcontacts}
-    Number of bound Ca$^{2+}$ during simulations.
+    Number of bound Ca$^{2+}$ during simulations. The data for CHARMM36 is taken directly from Refs. \citenum{catte16,charmmPOPC450mMCaClfiles,githubIONpaper} 
   }
